### 2018-09-07 draft.tex
Initial commit

\documentclass[10pt]{book}

\author{Polonio Davide}
\title{A thesis draft}
% The outline'd be something like this:
%  Thesis outline
% 
% - Introduction about the VIBES project (context, 5G, the need for a new 
% network 
% architecture, proposed solution, requisites)
%   + Related works
% - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
% docker, docker swarm)
%   + Choseen tecnhologies and description of the 'reason why'
% - Project description
%   + Initial solution with Openstack and Tacker
%   + From Tacker to Openbaton
%   + From Openbaton with Openstack to Openbaton with Docker
%   + Exploration of docker orchestrators (Swam vs Kubernetes)
%   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
%   + Openbaton and Kubernetes integration (lack of a implemented Driver, 
% integration with existinf VNFM)
% - Case study
%   + Analysis of TCP/UDP performance
% - Results
% - Future works
% - Conclusions

\makeindex
\begin{document}
 \maketitle

 % This is just a draft, nothing suitable for the final product

 % --- INTRODUCTION ---
 \chapter{Introduction}

 In the modern mobile networking, these is the continuous necessity to 
improve the overall speed connection: important goals are low latency and a 
good throughtput. Mobile data is continuously sourging every year, with the 
backbones having more traffic to process. In this context, in-hardware backbone 
solutions are becoming less efficent. Their time-to-deploy, in fact, is very 
high compared to the requirement of the network. With this trend, in-hardware 
solutions will not be able to keep the

 \section{The VIBES project}
\end{document}

### 2018-09-07 draft.tex
Add abstract

@@
 
- In the modern mobile networking, these is the continuous necessity to 
+ In the modern mobile networking, these is a continuous necessity to 
 improve the overall speed connection: important goals are low latency and a 
-good throughtput. Mobile data is continuously sourging every year, with the 
-backbones having more traffic to process. In this context, in-hardware backbone 
-solutions are becoming less efficent. Their time-to-deploy, in fact, is very 
-high compared to the requirement of the network. With this trend, in-hardware 
-solutions will not be able to keep the
+high throughtput. Mobile traffic is continuously surging every year, with 
+the backbones having more data to process. In this context, in-hardware 
+backbone solutions are becoming less efficent. Their time-to-deploy, in fact, is 
+very high compared to the requirements of the network. With this trend, 
+in-hardware solutions will not be able to keep up. Is with this considerations 
+that software solutions are being proposed to become the new way to deploy 
+network components. While less efficent, they are faster to bring them up. On 
+top of that, new virtualization technologies such as Docker are becoming more 
+and more mature to be employable for large-scale production environments, where 
+new frameworks, called orchestrators, allow to easily manage the application 
+lifecyle, i.e. deployment, scale up/down and finally, removal.
 

### 2018-09-10 draft.tex
Fix sentence

@@
 top of that, new virtualization technologies such as Docker are becoming more 
-and more mature to be employable for large-scale production environments, where 
-new frameworks, called orchestrators, allow to easily manage the application 
-lifecyle, i.e. deployment, scale up/down and finally, removal.
+and more mature for being employed in large-scale production environments,
+where new frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm),
+allow to easily manage the application lifecyle, i.e. deployment, scale up/down
+and finally, removal.
 

### 2018-09-11 draft.tex
Add introduction about VIBES project
Add 'todonotes' package and move content in the right section

@@
 \documentclass[10pt]{book}
+
+\usepackage{todonotes}
 
@@
 
+ Through the history of internet, many factor changed and new innovations came,
+but at the end of the day a foundamental key remained: TCP/IP performance are, 
+indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP 
+performes quite well in wired connections, while in wireless one, such as 
+wireless/satellite, sucks ass}
+
  \section{The VIBES project}

### 2018-09-11 draft.tex
Start talking about the VIBES project

--- 2018-09-10 draft.tex
@@
 
- In the modern mobile networking, these is a continuous necessity to 
-improve the overall speed connection: important goals are low latency and a 
-high throughtput. Mobile traffic is continuously surging every year, with 
-the backbones having more data to process. In this context, in-hardware 
-backbone solutions are becoming less efficent. Their time-to-deploy, in fact, is 
-very high compared to the requirements of the network. With this trend, 
+  Through the history of internet, many factor changed and new innovations came,
+but at the end of the day a foundamental key remained: TCP/IP performance are, 
+indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP 
+performes quite well in wired connections, while in wireless one, such as 
+wireless/satellite, sucks ass} In this context network optimization plays an 
+important role, and it is for this reason that network packets are processed 
+while traversing the network in order to make their transmission as fast as 
+possible. On top of that, mobile traffic is continuously surging every year, 
+with the backbones having more and more data to process. In this context, 
+in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in 
+fact, is very high compared to the requirements of the network. With this trend, 
 in-hardware solutions will not be able to keep up. Is with this considerations 
 that software solutions are being proposed to become the new way to deploy 
-network components. While less efficent, they are faster to bring them up. On 
-top of that, new virtualization technologies such as Docker are becoming more 
-and more mature for being employed in large-scale production environments,
-where new frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm),
-allow to easily manage the application lifecyle, i.e. deployment, scale up/down
+network components. While less efficent, they are faster to bring them up, 
+offering an unprecedented flexibility. Furthermore, new virtualization 
+technologies such as Docker are becoming more and more mature for being employed 
+in large-scale production environments, where new frameworks, called 
+orchestrators (e.g. Kubernetes, Docker Swarm), allow to easily manage the 
+application lifecyle, i.e. deployment, scale up/down
 and finally, removal.
@@
  \section{The VIBES project}
+
+ This thesis initially begun with the goals descripted in the VIBES\todo{Talk 
+about VIBES project, add some reference, explain what it is.} project, where 
+the necessity for better TCP/IP transmission through satellite connections was 
+the main requirement. Starting from that, we encountered many that, many 
+networking tools already present in the market, required some tweaking and some 
+integrations, shifting our goal to create a complete testbed for software VNF 
+communications, following the specifications suggested in the RFC 7665.
 \end{document}

### 2018-09-12 draft.tex
Minor improvement

@@
 
- % This is just a draft, nothing suitable for the final product
+ % This is just a draft, nothing suitable for the final product: sources are in 
+ % form of a footnote
 
@@
 network components. While less efficent, they are faster to bring them up, 
-offering an unprecedented flexibility. Furthermore, new virtualization 
-technologies such as Docker are becoming more and more mature for being employed 
-in large-scale production environments, where new frameworks, called 
-orchestrators (e.g. Kubernetes, Docker Swarm), allow to easily manage the 
-application lifecyle, i.e. deployment, scale up/down
+offering an unprecedented flexibility\footnote{SDN/NFV-Based Mobile Packet 
+Core Network Architectures: A Survey}. Furthermore, new virtualization 
+technologies such as Cloud computing and Docker are becoming more and more 
+mature for being employed in large-scale production environments, where new 
+frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to 
+easily manage the application lifecyle, i.e. deployment, scale up/down
 and finally, removal.
@@
 the necessity for better TCP/IP transmission through satellite connections was 
-the main requirement. Starting from that, we encountered many that, many 
+the main requirement. Starting with this objective, we encountered that, many 
 networking tools already present in the market, required some tweaking and some 
-integrations, shifting our goal to create a complete testbed for software VNF 
-communications, following the specifications suggested in the RFC 7665.
+integrations, shifting our goal to create a complete  ETSI (European
+Telecommunications Standards Institutes) MANO testbed, following the 
+specifications suggested in the RFC 7665.
 \end{document}

### 2018-09-13 draft.tex
Add content about VIBES project

@@
 % - Introduction about the VIBES project (context, 5G, the need for a new 
-% network 
-% architecture, proposed solution, requisites)
+% network architecture, proposed solution, requisites)
 %   + Related works
@@
 
- This thesis initially begun with the goals descripted in the VIBES\todo{Talk 
-about VIBES project, add some reference, explain what it is.} project, where 
-the necessity for better TCP/IP transmission through satellite connections was 
-the main requirement. Starting with this objective, we encountered that, many 
-networking tools already present in the market, required some tweaking and some 
-integrations, shifting our goal to create a complete  ETSI (European
-Telecommunications Standards Institutes) MANO testbed, following the 
-specifications suggested in the RFC 7665.
+ This thesis is part of the VIBES project\todo{Talk about VIBES project, add 
+some reference, explain what it is.}, where the necessity for better TCP/IP 
+transmission through satellite connections is the main requirement. To 
+reach this goal, the project specifications suggest to exploit the 5G incoming 
+technology and use the NFV-MANO architecture to perform first packets 
+elaboration and performance improvment and finally TCP/IP satellite chunk
+optimization with the Performance Enhancing Proxy (PEP). The VIBES project 
+proposed five technical requirements\todo{Finish list!}:
+\begin{itemize}
+ \item Analysis of the applicability of current and new Internet protocols in
+the proposed VNF-PEP architecture
+\end{itemize}
+
+Starting with the first step of creating a MANO able to process incoming data 
+packets through VNF functions, we encountered that many networking tools already 
+present in the market required some tweaking and some integrations, shifting our 
+goal to create a complete  European Telecommunications Standards Institutes 
+(ETSI) Management and orchestrator (MANO) testbed instead, following the 
+specifications suggested in the RFC 7665, thus implementing only the first three 
+requisites, without digging in the satellite data flow optimization.
 \end{document}

### 2018-09-14 draft.tex
Talk about requirements and Kubernetes

@@
 optimization with the Performance Enhancing Proxy (PEP). The VIBES project 
-proposed five technical requirements\todo{Finish list!}:
-\begin{itemize}
+proposed five technical requirements:
+\begin{enumerate}
  \item Analysis of the applicability of current and new Internet protocols in
 the proposed VNF-PEP architecture
-\end{itemize}
+ \item Implementation of a VNF-PEP prototype
+ \item Building of a PoC test platform
+ \item VNF-PEP validation and performance tests on 5G use-cases
+ \item Demonstration testbed management
+\end{enumerate}
+
+\subsection{VNF-PEP architecture and internet protocols}
+
+The analysis of this topic revealed to be trivial: since internet has many 
+different protocols that would become infeasible to support all of them at the 
+same time, packet incapsulation revealed to be the solution: every packet 
+incoming in the VNFs have been previously incapsulated by the MANO, making the 
+whole architecture indipendent from the protocol a particolar data-flux uses. 
+To achieve this, TUN/TAP\todo{Find out acronym} interfaces were used to deploy 
+a tunnel between VNFs, allowing them to pass packets encapsulated in UDP ones.
+
+\subsection{VNF-PEP prototype}
+
+Since our goal shifted into creating a MANO testbed, our prototype doesn't 
+include PEP. The VNF architecture was shaped following the container 
+orchestrator we decided to use: Kubernetes.
+
+\subsubsection{About kubernetes}
+
+As already introduced before, Kubernetes is an open source software framework 
+specialized in container management and orchestration. Developed by Google in 
+??\todo{Find out the actual date} and written in Go, it allows different 
+machines (called nodes from now on) to create an abstraction layer and to form a 
+cluster where is possible to run Docker containers without brothering about 
+hardware constraints. Nodes can have specific roles and specialization, all of 
+them managed by one or many master nodes, that actually manage the 
+orchestration.
+On top of that, Kubernetes astrae from the virtualization in-use, giving the 
+possibility to change the virtualization software in use (i.e. from 
+Container-based technology to Virtual machine software).
 

### 2018-09-17 draft.tex
Add content about docker

@@
 orchestration.
-On top of that, Kubernetes astrae from the virtualization in-use, giving the 
-possibility to change the virtualization software in use (i.e. from 
+On top of that, Kubernetes is virtualization-agnostic, meaning that it offers 
+the possibility to change the virtualization software in use (i.e. from 
 Container-based technology to Virtual machine software).
+
+\subsubsection{About Docker}
+
+In recent years, with the new hardware capabilities and the recent development 
+of in-kernel virtualization systems (such as Hyper-V) this technology 
+begin to be adopted. Virtualization allows to run different systems on the 
+same machine, making them completely isolated and then more resilient to 
+failure. The back of the medal, though, is that virtual machines requires large 
+amount of resources, especially memory, because solutions like copy-and-write 
+of in-memory pages are not viable anymore, leading to duplication of loaded 
+libraries and assets in the memory of the host. For large deployment containing 
+only simple services (e.g. a backend application serving a website or a 
+database) this result in a waste of resources. \todo{Talk about lxc containers 
+and how Docker solved magically the problem.}
+
+, finally leading to , especially 
+regarding memory consumption.
 

### 2018-09-17 draft.tex
Little fixes and additions

@@
 amount of resources, especially memory, because solutions like copy-and-write 
-of in-memory pages are not viable anymore, leading to duplication of loaded 
-libraries and assets in the memory of the host. For large deployment containing 
-only simple services (e.g. a backend application serving a website or a 
-database) this result in a waste of resources. \todo{Talk about lxc containers 
-and how Docker solved magically the problem.}
+of in-memory pages are not viable anymore (since the Kernel gets duplicated 
+too), leading to duplication of loaded libraries and assets in the memory of the 
+host. For large deployment containing only simple services (e.g. a backend 
+application serving a website or a database) this ends up in a waste of 
+resources. \todo{Talk about lxc containers and how Docker solved magically the 
+problem.} It's here that, in ??\todo{Find Docker date of birth}, Docker was 
+created, basing its solution on an already existing product: Linux Container 
+(LXC). From this framework, docker built an entire ecosystem, consisting in a 
+client/server model giving the possibility for users to simply launch, scale and 
+delete containers \todo{Talk about Docker-Machine}, a repository system where 
+images, the snapshot of a defined image, can be stored and tagged, in a sort of 
+system version control and of another additional solutions, called 
+\textit{Docker Compose}, and \textit{Docker Swarm} to, respectively, orchestrate 
+containers in a single node or a clustered system. \vspace{0.5cm}
 
@@
 requisites, without digging in the satellite data flow optimization.
+
+\subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
 \end{document}

### 2018-09-18 draft.tex
Finish Docker introduction. Start talking about Swarm

--- 2018-09-14 draft.tex
@@
 % - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
-% docker, docker swarm)
+% Docker, Docker swarm)
 %   + Choseen tecnhologies and description of the 'reason why'
@@
 %   + From Openbaton with Openstack to Openbaton with Docker
-%   + Exploration of docker orchestrators (Swam vs Kubernetes)
+%   + Exploration of Docker orchestrators (Swam vs Kubernetes)
 %   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
@@
 in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in 
-fact, is very high compared to the requirements of the network. With this trend, 
-in-hardware solutions will not be able to keep up. Is with this considerations 
-that software solutions are being proposed to become the new way to deploy 
-network components. While less efficent, they are faster to bring them up, 
-offering an unprecedented flexibility\footnote{SDN/NFV-Based Mobile Packet 
+fact, is very high compared to the requirements of the network. With this 
+trend, in-hardware solutions will not be able to keep up. Is with this 
+considerations that software solutions are being proposed to become the new way 
+to deploy network components. While less efficent, they are faster to bring them 
+up, offering an unprecedented flexibility\footnote{SDN/NFV-Based Mobile Packet 
 Core Network Architectures: A Survey}. Furthermore, new virtualization 
@@
 ??\todo{Find out the actual date} and written in Go, it allows different 
-machines (called nodes from now on) to create an abstraction layer and to form a 
-cluster where is possible to run Docker containers without brothering about 
+machines (called nodes from now on) to create an abstraction layer and to form 
+a cluster where is possible to run Docker containers without brothering about 
 hardware constraints. Nodes can have specific roles and specialization, all of 
@@
 orchestration.
-On top of that, Kubernetes astrae from the virtualization in-use, giving the 
-possibility to change the virtualization software in use (i.e. from 
+On top of that, Kubernetes is virtualization-agnostic, meaning that it offers 
+the possibility to change the virtualization software in use (i.e. from 
 Container-based technology to Virtual machine software).
+
+\subsubsection{About Docker}
+
+In recent years, with the new hardware capabilities and the recent development 
+of in-kernel virtualization systems (such as Hyper-V) this technology 
+begin to be adopted. Virtualization allows to run different systems on the 
+same machine, making them completely isolated and then more resilient to 
+failures. The back of the medal, though, is that virtual machines require large 
+amount of resources, especially memory, because solutions like copy-and-write 
+of in-memory pages are not viable anymore (since the kernel gets duplicated 
+too), leading to duplication of loaded libraries and assets in the memory of 
+the host. 
+For large deployment containing only simple services (e.g. a backend 
+application serving a website or a database) this ends up in a waste of 
+resources. \todo{Talk about lxc containers and how Docker solved magically the 
+problem.} It's here that, in ??\todo{Find Docker date of birth}, Docker was 
+created, basing its solution on an already existing product: Linux Container 
+(LXC). From this framework, Docker built an entire ecosystem, consisting in a 
+client/server model giving the possibility for users to simply launch, scale and 
+delete containers (locally or remotely with Docker-machine), a repository 
+system where images, a layer-based ``core'' where containers start running from 
+when they are launched, can be stored and tagged in a sort of version control 
+system. Finally, another couple of solutions, called Docker-compose, and Docker 
+Swarm have the purpose to, respectively, orchestrate containers in a single or 
+a clustered system.
+
+\paragraph{Docker Swarm} \todo{Talk about its characteristic and introduce it}
+
+\vspace{0.5cm}
 

### 2018-09-19 draft.tex
Add Docker Swarm software description

@@
 
-\paragraph{Docker Swarm} \todo{Talk about its characteristic and introduce it}
-
+\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
+birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
+seen as one logical unit. This layer completely makes the underlying
+infrastructure transparent: data management, as network one, are totaly managed
+by Docker. Recently, Docker-compose support has been introduced, making
+this one-node tool available for clustered systems. At the end of the day, on
+one hand Docker Swarm provides an easy way to set up a cluster system with all
+the tools configured out-of-the-box, without the necessity to set networking
+configurations or installing ad-hoc storage solutions. On the other hand,
+though, it doesn't have all the personalization options that Kuberentes offers,
+thus making the product less flexible. This foundamental characteristic,
+although, makes the two solutions have different use cases leading to different
+market needs.
 \vspace{0.5cm}

### 2018-09-19 draft.tex
Fix lowercase word
Add WIP docker-compose part

@@
 
-\subsubsection{About kubernetes}
+\subsubsection{About Kubernetes}
 
@@
 market needs.
+
+\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool 
+that allows services orchestration. It automates most of the tasks that should 
+be performed by hand when launching one or multiple docker containers. It's 
+particulary useful when more continers have to \textbf{INTERAGIRE}\todo{Find 
+translation} between them, because\todo{Write down docker-commpose 
+functionalities and describe the product a little bit}
+
 \vspace{0.5cm}
@@
 Starting with the first step of creating a MANO able to process incoming data 
-packets through VNF functions, we encountered that many networking tools already 
-present in the market required some tweaking and some integrations, shifting our 
+packets through VNF functions, we encountered that many networking tools 
+already 
+present in the market required some tweaking and some integrations, shifting 
+our 
 goal to create a complete  European Telecommunications Standards Institutes 
 (ETSI) Management and orchestrator (MANO) testbed instead, following the 
-specifications suggested in the RFC 7665, thus implementing only the first three 
+specifications suggested in the RFC 7665, thus implementing only the first 
+three 
 requisites, without digging in the satellite data flow optimization.

### 2018-09-20 draft.tex
Complete docker-compose paragraph

@@
 different protocols that would become infeasible to support all of them at the
-same time, packet incapsulation revealed to be the solution: every packet 
-incoming in the VNFs have been previously incapsulated by the MANO, making the 
-whole architecture indipendent from the protocol a particolar data-flux uses. 
-To achieve this, TUN/TAP\todo{Find out acronym} interfaces were used to deploy 
-a tunnel between VNFs, allowing them to pass packets encapsulated in UDP ones.
+same time, packet incapsulation present itself as the only faseable solution:
+every packet incoming in the VNFs has already been incapsulated by the MANO,
+\todo{Not sure about this, please give it a check} making the whole
+architecture indipendent from the protocol a particolar flux of data uses. To
+achieve this, TUN/TAP\todo{Find out acronym} interfaces were used to deploy a
+tunnel between VNFs, allowing them to pass packets encapsulated in UDP ones.
 
@@
 the possibility to change the virtualization software in use (i.e. from
-Container-based technology to Virtual machine software).
+Container-based technology to Virtual machine software).\todo{Add Kubernetes
+logo}
 
@@
 Swarm have the purpose to, respectively, orchestrate containers in a single or
-a clustered system.
+a clustered system. \todo{Add Docker, docker-machine logos}
 
@@
 although, makes the two solutions have different use cases leading to different
-market needs.
+market needs. \todo{Add Docker Swarm}
 
@@
 be performed by hand when launching one or multiple docker containers. It's
-particulary useful when more continers have to \textbf{INTERAGIRE}\todo{Find 
-translation} between them, because\todo{Write down docker-commpose 
-functionalities and describe the product a little bit}
+particulary useful when more continers have to interact together,
+because\todo{Write down docker-commpose functionalities and describe the product
+a little bit} all the services rules are described in a simple YAML file.
+With the birth of Docker Swarm, work has started to port this tool to the Swarm
+framework. At the time of this writing, the procedure to port a Docker-compose
+configuration in Docker Swarm is still not completely transparent, and it
+requires a sort of compilation, where Docker-compose bundles all the
+necessaries resources into a binary file that Docker Swarm will use to allocate
+the necessary resources.
 
@@
 packets through VNF functions, we encountered that many networking tools
-already 
-present in the market required some tweaking and some integrations, shifting 
-our 
-goal to create a complete  European Telecommunications Standards Institutes 
-(ETSI) Management and orchestrator (MANO) testbed instead, following the 
-specifications suggested in the RFC 7665, thus implementing only the first 
-three 
-requisites, without digging in the satellite data flow optimization.
+already present in the market required some tweaking and some integrations,
+shifting our goal to create a complete  European Telecommunications Standards
+Institutes (ETSI) Management and orchestrator (MANO) testbed instead, following
+the specifications suggested in the RFC 7665, thus implementing only the first
+three requisites, without digging in the satellite data flow optimization.
+\todo{This section needs to be expanded a lot!!!}
 

### 2018-09-21 draft.tex
Add new content, new section about MANO

@@
 the specifications suggested in the RFC 7665, thus implementing only the first
-three requisites, without digging in the satellite data flow optimization.
-\todo{This section needs to be expanded a lot!!!}
+three requisites, without digging in the satellite data flow optimization. In 
+particular, we discovered how, these tools, were suitable to create ETSI MANO 
+and VNFs using virtual machine or exploiting cloud tecnhologies, while they 
+were not designed with enough flexibility to be able to accept new 
+virtualization technologies such as Docker. In the next session we are going to 
+take a deep analysis of the cited technologies, in order to provide to the 
+reader enough context to be able to understand our design choices we are going 
+to describe later.
+
+\section{MANO}
+
+\todo{This section needs to be expanded a lot!!! In particular, we should put 
+some focus about VNF and MANO, introducing them}
+
+\noindent With this in mind, we performed a requirements analysis described in 
+the next chapter.\todo{This should be put at the end of the introduction.}
 

### 2018-09-24 draft.tex
Fix phrase

--- 2018-09-20 draft.tex
@@
 the specifications suggested in the RFC 7665, thus implementing only the first
-three requisites, without digging in the satellite data flow optimization.
-\todo{This section needs to be expanded a lot!!!}
+three requisites, without digging in the satellite data flow optimization. In 
+particular, we discovered how, these tools, were suitable to create ETSI MANO 
+and VNFs using virtual machine or exploiting cloud tecnhologies, while they 
+were not designed with enough flexibility to integrate with Docker. In the next 
+section we are going to take a deep analysis of the cited technologies, in 
+order to provide to the reader enough context to be able to understand our 
+design choices we are going to describe later.
+
+\section{MANO}
+
+\todo{This section needs to be expanded a lot!!! In particular, we should put 
+some focus about VNF and MANO, introducing them}
+
+\noindent With this in mind, we performed a requirements analysis described in 
+the next chapter.\todo{This should be put at the end of the introduction.}
 

### 2018-09-24 draft.tex
Add chap 2

@@
 \chapter{Introduction}
+\label{chap:intro}
 
@@
 were not designed with enough flexibility to integrate with Docker. In the next 
-section we are going to take a deep analysis of the cited technologies, in 
+chapter we are going to take a deep analysis of the cited technologies, in 
 order to provide to the reader enough context to be able to understand our 
@@
 
-\section{MANO}
+\noindent With this in mind, we performed a requirements analysis described in 
+the next chapter.\todo{This should be put at the end of the introduction.}
+
+\chapter{Project Analysis}
+\label{chap:prjan}
+
+As in~\ref{chap:intro}, the creation of a full ETSI MANO compliant
+architecture, following the suggestions described in RFC 7665, require the 
+integration of multiple tools (some of them have been already briefly 
+introduced in~\ref{chap:intro}). The tools were choosen after their 
+analysis, where pro and cons have been studied and compared to the necessity of 
+this thesis.
 
@@
 
-\noindent With this in mind, we performed a requirements analysis described in 
-the next chapter.\todo{This should be put at the end of the introduction.}
-
 \subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
+\paragraph{Openstack}
+\paragraph{Openvswitch}
 \end{document}

### 2018-09-27 draft.tex
Add part about Openstack

@@
 analysis, where pro and cons have been studied and compared to the necessity of 
-this thesis.
+this thesis. Since the project has been developed with the cloud provided by 
+the University of Padua, Openstack has been used as base for our deployments.
+
+\paragraph{Openstack} Created in ??\todo{Find out Openstack date of creation}, 
+Openstack allows on-permise cloud installation, using bare-metal resources to 
+provide common cloud services as object storage, virtual machine deploy, 
+virtual networking. Furthermore its modularity offers the possibility to add 
+additional components, even proprietary, to achieve a complete solution. 
+\todo{Finish him! :D}
 

### 2018-09-28 draft.tex
Add technical analysis

--- 2018-09-24 draft.tex
@@
 analysis, where pro and cons have been studied and compared to the necessity of 
-this thesis.
+this thesis. Since the project has been developed with the cloud provided by 
+the University of Padua, Openstack (which a product analysis can be found later 
+in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
+
+\section{Technical analysis}
+
+As today, when a packets gets send from a seder $S$, to a receiver $R$, the 
+packet gets travelling to te backbone receive appropriate in-hardware packet 
+processing before being received by $R$. During this elaboration, it is 
+important to maintain the transparency of the connection, so that the network 
+users are not aware of the packets elaboration at all. To perform that, many 
+solutions are available: TUN/TAP packet incapsulation, IP over IP or TCP 
+session recreation\todo{Check if it's the same of Split TCP}.
+
+TUN/TAP and IP over IP maintain the connection transparency incapsulating the 
+packet in another one. In this way, the original packet remains preserved and it 
+becomes the new packet payload\todo{Make an image showing a packet inside 
+another packet, to make easy to understand what we're meaning here}, that can 
+be elaborated and modified accordingly, until it gets decapsulated and sent to 
+$R$ eventually. When $R$ receives the packet, it is not be aware of the packet 
+elaboration, since all the original headers were preserved (or slightly 
+modified).
+
+\todo{Explain TCP session recreation}
+
+\section{Openstack}
+\label{chap:prjan:sec:openstack}
+Created in ??\todo{Find out Openstack date of creation}, 
+Openstack allows on-permise cloud installation, using bare-metal resources to 
+provide common cloud services as object storage, virtual machine deploy, 
+virtual networking. Furthermore its modularity offers the possibility to add 
+additional components, even proprietary, to achieve a complete cloud solution.
 

### 2018-10-01 draft.tex
Add TCP session recreation

@@
 every packet incoming in the VNFs has already been incapsulated by the MANO,
-\todo{Not sure about this, please give it a check} making the whole
-architecture indipendent from the protocol a particolar flux of data uses. To
-achieve this, TUN/TAP\todo{Find out acronym} interfaces were used to deploy a
-tunnel between VNFs, allowing them to pass packets encapsulated in UDP ones.
+\todo{Not sure about this, please give it a check - Not sure but I think that
+it is not part of the MANO the encapsulation of the packets. Also I think that
+the main purpose of UDP encapsulation is not to hide the protocol used on the
+edges but 1) usa a ``quick protocol to exchange data among VNFs and 2) we are
+hiding the path not the protocol itself. As we were discussing, we are creating
+some sort of proxy, so the aim will be the same even if we support a plethora
+of protocols.''} making the whole architecture indipendent from the protocol a
+particolar flux of data uses. To achieve this, TUN/TAP
+\todo{Find out acronym - EPKT? (Encapsulated PacKeT)}
+interfaces were used to deploy a tunnel between VNFs, allowing them to pass
+packets encapsulated in UDP ones.
 
@@
 elaboration, since all the original headers were preserved (or slightly 
-modified).
+modified). TCP session recreation, instead, ``split'' the TCP session in two 
+endpoints: an Ingress, that manage the connection with the client and an 
+Egress, that communicate with the packet receiver. In order to achieve that, a 
+session table needs to be established and continuously updated, while TCP 
+headers need to be accordingly modified to allow data transmission through the 
+Ingress and Egress points. TCP session recreation seems burdersome at the first 
+sight, but it offers the flexibility to perform extensive packet manipulation: 
+since the packet its recreated every time at the Ingress and Egress points, 
+while in the Network Function Virtualization Infrastructure (NFVI), the 
+original packet can be stripped of its headers (that could be saved in the 
+session table) to slightly improve internal packet transmission. \todo{Insert 
+TCP session recreation schema}
+
+During the testbed development we first tried the TUN/TAP solution, that 
+revealed to be not completely transparent and with major performance drawbacks 
+(i.e. packet transmission in a LAN connection reached latencies of 50ms) 
+\todo{Explain the constraint of point-to-point connections, while we need 
+dynamic SFC)}
+
+In TCP Splitting, instead, we split the path from sender and receiver dividing
+the end-to-end TCP connection into a multi-overlay-hop path where to each hop
+we have an independant TCP cconnection. Each segment that composes the whole
+path can use different flavours of TCP, e.g. using different enhancement based
+on the tipology of the connection (in case of satellite connection we TCP Hybla
+can be used, HighSpeed TCP in case of optical fiber). That could cause
+performance issues depending on the intermediate proxies that create the route:
+it is the case, for example, when an intermediate node sends back a spoofed ACK
+before the original packet is delivered to the real consegnee. In that senario,
+the sender unaware of the proxy, will miscalculate the sending window based on
+the too low RTT. Other problems of TCP splitting are reliability andd security:
+breacking the end to end TCP logic: a server failure is may cause to an unaware
+client to believe that all packets reach the destination.
 
@@
 \paragraph{Openvswitch}
+\paragraph{Openbaton}
+\paragraph{SDN?}
 \end{document}

### 2018-10-02 draft.tex
Add content

@@
 
-\section{Technical analysis}
+\section{Architecture overview}
+
+During the testbed development the software architecture diverged from the 
+original one, experiencing different evolutions and modifications.
+
+\section{SFC packet transmission strategies}
 
@@
 \todo{Explain the constraint of point-to-point connections, while we need 
-dynamic SFC)}
+dynamic SFC)}, so we choose to opt for a TCP session recreation solution, 
+implementing a little back-end with the aim of saving in a persistent storage 
+the SFC the packet has to perform and additional metadata.
 

### 2018-10-03 draft.tex
Add section on tcp split

--- 2018-10-01 draft.tex
@@
 elaboration, since all the original headers were preserved (or slightly 
-modified). TCP session recreation, instead, ``split'' the TCP session in two 
-endpoints: an Ingress, that manage the connection with the client and an 
-Egress, that communicate with the packet receiver. In order to achieve that, a 
-session table needs to be established and continuously updated, while TCP 
-headers need to be accordingly modified to allow data transmission through the 
-Ingress and Egress points. TCP session recreation seems burdersome at the first 
-sight, but it offers the flexibility to perform extensive packet manipulation: 
-since the packet its recreated every time at the Ingress and Egress points, 
-while in the Network Function Virtualization Infrastructure (NFVI), the 
-original packet can be stripped of its headers (that could be saved in the 
-session table) to slightly improve internal packet transmission. \todo{Insert 
-TCP session recreation schema}
-
-During the testbed development we first tried the TUN/TAP solution, that 
-revealed to be not completely transparent and with major performance drawbacks 
-(i.e. packet transmission in a LAN connection reached latencies of 50ms) 
-\todo{Explain the constraint of point-to-point connections, while we need 
-dynamic SFC)}
+modified).
 

### 2018-10-03 draft.tex
Start talking about VIBES architecture

--- 2018-10-02 draft.tex
@@
 
+\subsection{VIBES architectural description}
+
+In the original VIBES architectural design, based on leveraging virtualization 
+technologies, is possible to identify different macro-components, each of these 
+specialized in fullfilling a specific goal. For the ETSI standards, the Network 
+Function Virtualization (NFV) design should be composed as follows:\todo{Insert 
+image page 11 ETSI vibes specification}
+\begin{itemize}
+ \item Network Function Virtualization Infrastructure (NFVI): it's the heart of 
+the whole computation component, and here three sub-domains can be identified:
+\todo{Expand this section}
+\begin{itemize} 
+ \item Hypervisor domain that is composed of the Virtual Compute, the Virtual 
+Storage, and the Virtualization Layer. These components
+ \item Compute domain composed by the Physical Compute and the Physical Storage
+ \item Infrastructure domain which has the Physical Network, and its 
+virtualized counterpart as made available by the Virtualization Layer
+\end{itemize}
+
+\end{itemize}
+
 \section{SFC packet transmission strategies}

### 2018-10-04 draft.tex
Add description about ETSI software architecture

@@
 \begin{itemize} 
- \item Hypervisor domain that is composed of the Virtual Compute, the Virtual 
-Storage, and the Virtualization Layer. These components
- \item Compute domain composed by the Physical Compute and the Physical Storage
- \item Infrastructure domain which has the Physical Network, and its 
-virtualized counterpart as made available by the Virtualization Layer
+ \item \textbf{Hypervisor domain} that is composed of the Virtual Compute, the 
+Virtual Storage, and the Virtualization Layer. These components are used to 
+virtualize the computational and storage resources, providing a layer for 
+accessing it.
+ \item \textbf{Compute domain} composed by the Physical Compute and the 
+Physical Storage. These are the real resources which are virtualized via 
+software.
+ \item \textbf{Infrastructure domain} which has the Physical Network, and its 
+virtualized counterpart as made available by the Virtualization Layer.
 \end{itemize}
+
+In conventional virtualization systems these resources are usually separated 
+from the host operating system, providing better isolation but worste 
+performance. The ETSI infrastructure, instead, tries with container 
+virtualization technology to push for a better NFV responsiveness\todo{Check 
+term - I don't think it actually exists}, removing the additional operating 
+system (called ``guest'' OS) required by the usual hypervisor-oriented approach.
 

### 2018-10-05 draft.tex
Add additional content about architectural description

@@
 In conventional virtualization systems these resources are usually separated 
-from the host operating system, providing better isolation but worste 
+from the host operating system, providing better isolation but wrost
 performance. The ETSI infrastructure, instead, tries with container 
@@
 system (called ``guest'' OS) required by the usual hypervisor-oriented approach.
+
+To realize the required components, specific tools were suggested in the VNF 
+technical proposal: regarding the NFV (network infrastructure) domain 
+Docker-Compose was proposed as a viable solution, while to store PEP instances 
+docker swarm was described as a possible candidate.
+
+We performed an accured analysis of the tools choosen based on their maturity, 
+on the community and on the technical support offered (e.g. user manuals, 
+developer documentation), which, at the end, commited us to choose different 
+tools from the suggested ones. A detailed description of the architectural 
+implementation can be found in chapter~\ref{chap:archimpl}.
 
@@
 \paragraph{SDN?}
+
+\chapter{Architectural implementation}
+\label{chap:archimpl}
 \end{document}

### 2018-10-09 draft.tex
Add docker logo and vnf architecture schema
Fix some TODO and add images to the doc

@@
 \usepackage{todonotes}
+
+\graphicspath{{res/img/}}
 
@@
 the possibility to change the virtualization software in use (i.e. from
-Container-based technology to Virtual machine software).\todo{Add Kubernetes
-logo}
+Container-based technology to Virtual machine software).
+\begin{figure}[h]
+ \centering
+ \includegraphics[scale=0.35]{kubernetes_logo}
+ \caption{Kubernetes Logo}
+ \label{chap:intro:img:k8s_logo}
+\end{figure}
 
@@
 Swarm have the purpose to, respectively, orchestrate containers in a single or
-a clustered system. \todo{Add Docker, docker-machine logos}
+a clustered system.
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.7]{docker_logo}
+ \caption{Official Docker logo}
+ \label{chap:intro:img:docker_logo}
+\end{figure}
 
@@
 specialized in fullfilling a specific goal. For the ETSI standards, the Network 
-Function Virtualization (NFV) design should be composed as follows:\todo{Insert 
-image page 11 ETSI vibes specification}
+Function Virtualization (NFV) design should be composed as follows:
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=2]{etsi_arch}
+ \caption{A schematic image representing the high level NFV architecture 
+proposed by ESA.}
+ \label{chap:prjan:img:etsi_arch}
+\end{figure}
+
 \begin{itemize}
@@
 Created in ??\todo{Find out Openstack date of creation}, 
-Openstack allows on-permise cloud installation, using bare-metal resources to 
+Openstack allows on-permise cloud installations, using bare-metal resources to 
 provide common cloud services as object storage, virtual machine deploy, 
@@
 additional components, even proprietary, to achieve a complete cloud solution.
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.58]{openstack_logo}
+ \caption{Openstack logo}
+ \label{chap:prjan:img:openstack_logo}
+\end{figure}
 
@@
 \paragraph{Openbaton}
+\begin{figure}[h]
+ \centering
+ \includegraphics[scale=0.45]{openbaton_logo}
+ \caption{Openbaton logo}
+ \label{chap:prjan:img:openbaton_logo}
+\end{figure}
+
 \paragraph{SDN?}

### 2018-10-09 draft.tex
Merge Marco's TCP split part with mine

@@
 modified). TCP session recreation, instead, ``split'' the TCP session in two
-endpoints: an Ingress, that manage the connection with the client and an 
-Egress, that communicate with the packet receiver. In order to achieve that, a 
-session table needs to be established and continuously updated, while TCP 
-headers need to be accordingly modified to allow data transmission through the 
-Ingress and Egress points. TCP session recreation seems burdersome at the first 
+endpoints: an Ingress, that manage the connection with the client and an Egress,
+that communicate with the packet receiver creating a multi-overlay-hop path
+where for each hop there is an possible indipendent TCP connection. In order to
+achieve that, a session table needs to be established and continuously updated,
+while TCP headers need to be accordingly modified to allow data transmission
+through the Ingress and Egress points. Nonetheless, care should be given to
+packets path: in order to avoid ill-calculated congestion windows, the incoming
+packets need to follow the same (or an equivalent one) path for the same
+instatiated connection. This phenomenom is based on the intermediate proxies
+that create the route: it is the case, for example, when an intermediate node
+sends back a spoofed ACK before the original packet is delivered to the real
+consegnee. In this senario, the sender, unaware of the proxy, will miscalculate
+the sending window based on the too low RTT. Other problems of TCP splitting are
+reliability and security, especially because the end to end TCP logic is no
+longer maintained: a server failure may cause to an unaware client to believe
+that all the packets have reached the destination. TCP session recreation seems
+burdersome and it doesn't seem to add any significant benefit at the first
 sight, but it offers the flexibility to perform extensive packet manipulation:
 since the packet its recreated every time at the Ingress and Egress points,
-while in the Network Function Virtualization Infrastructure (NFVI), the 
-original packet can be stripped of its headers (that could be saved in the 
-session table) to slightly improve internal packet transmission. \todo{Insert 
-TCP session recreation schema}
+while in the Network Function Virtualization Infrastructure (NFVI), the original
+packet can be stripped of its headers (that could be saved in the session table)
+to slightly improve internal packet transmission. \todo{Insert TCP session
+  recreation schema} On top of that, it offers the possibility to use different
+TCP flavours, hence increasing the communication speed in heterogeneous networks
+(e.g. in case of satellite connections TCP Hybla can be employed, while it is
+known that HighSpeed TCP gives best results when applied in optical fiber-based
+connections).
 
@@
 the SFC the packet has to perform and additional metadata.
-
-In TCP Splitting, instead, we split the path from sender and receiver dividing
-the end-to-end TCP connection into a multi-overlay-hop path where to each hop
-we have an independant TCP cconnection. Each segment that composes the whole
-path can use different flavours of TCP, e.g. using different enhancement based
-on the tipology of the connection (in case of satellite connection we TCP Hybla
-can be used, HighSpeed TCP in case of optical fiber). That could cause
-performance issues depending on the intermediate proxies that create the route:
-it is the case, for example, when an intermediate node sends back a spoofed ACK
-before the original packet is delivered to the real consegnee. In that senario,
-the sender unaware of the proxy, will miscalculate the sending window based on
-the too low RTT. Other problems of TCP splitting are reliability andd security:
-breacking the end to end TCP logic: a server failure is may cause to an unaware
-client to believe that all packets reach the destination.
 

### 2018-10-09 draft.tex
Remove TODO
Add IP over IP image, create misc folder for draw.io source files

@@
 
-As in~\ref{chap:intro}, the creation of a full ETSI MANO compliant
+As in chapter~\ref{chap:intro}, the creation of a full ETSI MANO compliant
 architecture, following the suggestions described in RFC 7665, require the
-integration of multiple tools (some of them have been already briefly 
-introduced in~\ref{chap:intro}). The tools were choosen after their 
-analysis, where pro and cons have been studied and compared to the necessity of 
-this thesis. Since the project has been developed with the cloud provided by 
-the University of Padua, Openstack (which a product analysis can be found later 
+integration of multiple tools (some of them have been already briefly introduced
+in~\ref{chap:intro}). The tools were choosen after their analysis, where pro and
+cons have been studied and compared to the necessity of this thesis. Since the
+project has been developed with the cloud provided by the University of Padua,
+Openstack (which a product analysis can be found later
 in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
@@
 solutions are available: TUN/TAP packet incapsulation, IP over IP or TCP 
-session recreation\todo{Check if it's the same of Split TCP}.
+session recreation (know also as TCP split).
 
 TUN/TAP and IP over IP maintain the connection transparency incapsulating the
-packet in another one. In this way, the original packet remains preserved and it
-becomes the new packet payload\todo{Make an image showing a packet inside
-  another packet, to make easy to understand what we're meaning here}, that can
-be elaborated and modified accordingly, until it gets decapsulated and sent to
-$R$ eventually. When $R$ receives the packet, it is not be aware of the packet
-elaboration, since all the original headers were preserved (or slightly
-modified). TCP session recreation, instead, ``split'' the TCP session in two
-endpoints: an Ingress, that manage the connection with the client and an Egress,
-that communicate with the packet receiver creating a multi-overlay-hop path
-where for each hop there is an possible indipendent TCP connection. In order to
-achieve that, a session table needs to be established and continuously updated,
-while TCP headers need to be accordingly modified to allow data transmission
-through the Ingress and Egress points. Nonetheless, care should be given to
-packets path: in order to avoid ill-calculated congestion windows, the incoming
-packets need to follow the same (or an equivalent one) path for the same
-instatiated connection. This phenomenom is based on the intermediate proxies
-that create the route: it is the case, for example, when an intermediate node
-sends back a spoofed ACK before the original packet is delivered to the real
-consegnee. In this senario, the sender, unaware of the proxy, will miscalculate
-the sending window based on the too low RTT. Other problems of TCP splitting are
-reliability and security, especially because the end to end TCP logic is no
-longer maintained: a server failure may cause to an unaware client to believe
-that all the packets have reached the destination. TCP session recreation seems
-burdersome and it doesn't seem to add any significant benefit at the first
-sight, but it offers the flexibility to perform extensive packet manipulation:
-since the packet its recreated every time at the Ingress and Egress points,
-while in the Network Function Virtualization Infrastructure (NFVI), the original
-packet can be stripped of its headers (that could be saved in the session table)
-to slightly improve internal packet transmission. \todo{Insert TCP session
-  recreation schema} On top of that, it offers the possibility to use different
-TCP flavours, hence increasing the communication speed in heterogeneous networks
-(e.g. in case of satellite connections TCP Hybla can be employed, while it is
-known that HighSpeed TCP gives best results when applied in optical fiber-based
-connections).
+packet in another one.
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{IPoverIP}
+  \caption[IP over IP schema]{An example of IP over IP. Is noticeable how the
+    original packet becomes the payload of a new one, thus being encapsulated.
+    All the informations regarding the original packet remains untouched: when
+    this kind of packet gets send from a machine to another, the OS will remove
+    the new packet header and will present to the user-space program receiving
+    the data its payload: a IP packet. Packets encapsulation can be applied with
+    higher layers too, thus allowing for TCP and UDP packet incapsulation.}
+  \label{chap:prjan:img:ip_over_ip}
+\end{figure}
+
+In this way, the original packet remains preserved and it becomes the new packet
+payload\todo{Make an image showing a packet inside another packet, to make easy
+  to understand what we're meaning here}, that can be elaborated and modified
+accordingly, until it gets decapsulated and sent to $R$ eventually. When $R$
+receives the packet, it is not be aware of the packet elaboration, since all the
+original headers were preserved (or slightly modified). TCP session recreation,
+instead, ``split'' the TCP session in two endpoints: an Ingress, that manage the
+connection with the client and an Egress, that communicate with the packet
+receiver creating a multi-overlay-hop path where for each hop there is an
+possible indipendent TCP connection. In order to achieve that, a session table
+needs to be established and continuously updated, while TCP headers need to be
+accordingly modified to allow data transmission through the Ingress and Egress
+points. Nonetheless, care should be given to packets path: in order to avoid
+ill-calculated congestion windows, the incoming packets need to follow the same
+(or an equivalent one) path for the same instatiated connection. This phenomenom
+is based on the intermediate proxies that create the route: it is the case, for
+example, when an intermediate node sends back a spoofed ACK before the original
+packet is delivered to the real consegnee. In this senario, the sender, unaware
+of the proxy, will miscalculate the sending window based on the too low RTT.
+Other problems of TCP splitting are reliability and security, especially because
+the end to end TCP logic is no longer maintained: a server failure may cause to
+an unaware client to believe that all the packets have reached the destination.
+TCP session recreation seems burdersome and it doesn't seem to add any
+significant benefit at the first sight, but it offers the flexibility to perform
+extensive packet manipulation: since the packet its recreated every time at the
+Ingress and Egress points, while in the Network Function Virtualization
+Infrastructure (NFVI), the original packet can be stripped of its headers (that
+could be saved in the session table) to slightly improve internal packet
+transmission. \todo{Insert TCP session recreation schema} On top of that, it
+offers the possibility to use different TCP flavours, hence increasing the
+communication speed in heterogeneous networks (e.g. in case of satellite
+connections TCP Hybla can be employed, while it is known that HighSpeed TCP
+gives best results when applied in optical fiber-based connections).
 

### 2018-10-10 draft.tex
Minor fix about SFC packet transmission

@@
 users are not aware of the packets elaboration at all. To perform that, many 
-solutions are available: TUN/TAP packet incapsulation, IP over IP or TCP 
-session recreation (know also as TCP split).
+solutions are available: TUN/TAP packet incapsulation, IP Encapsulation within 
+IP or TCP session recreation (know also as TCP split).
 
-TUN/TAP and IP over IP maintain the connection transparency incapsulating the
-packet in another one.
+TUN/TAP and IP Encapsulation within IP (initially defined in RFC 2003 with the 
+aim to deliver IP packets over Mobile IP for mobile networks) maintain the 
+connection transparency incapsulating the packet in another one.
 \begin{figure}[t]
@@
   \includegraphics[scale=0.5]{IPoverIP}
-  \caption[IP over IP schema]{An example of IP over IP. Is noticeable how the
-    original packet becomes the payload of a new one, thus being encapsulated.
-    All the informations regarding the original packet remains untouched: when
-    this kind of packet gets send from a machine to another, the OS will remove
-    the new packet header and will present to the user-space program receiving
-    the data its payload: a IP packet. Packets encapsulation can be applied with
-    higher layers too, thus allowing for TCP and UDP packet incapsulation.}
+  \caption[IP Encapsulation within IP packet schema]{An schema of IP 
+Encapsulation within IP. Is noticeable how the original packet becomes the 
+payload of a new one, thus being encapsulated. All the informations regarding 
+the original packet remains untouched: when this kind of packet gets send 
+from a machine to another, the OS will remove the new packet header and will 
+present to the user-space program receiving the data its payload: a IP 
+packet. Packets encapsulation can be applied with higher layers too, thus 
+allowing for TCP and UDP packet incapsulation.}
   \label{chap:prjan:img:ip_over_ip}
@@
 In this way, the original packet remains preserved and it becomes the new packet
-payload\todo{Make an image showing a packet inside another packet, to make easy
-  to understand what we're meaning here}, that can be elaborated and modified
-accordingly, until it gets decapsulated and sent to $R$ eventually. When $R$
-receives the packet, it is not be aware of the packet elaboration, since all the
-original headers were preserved (or slightly modified). TCP session recreation,
-instead, ``split'' the TCP session in two endpoints: an Ingress, that manage the
-connection with the client and an Egress, that communicate with the packet
-receiver creating a multi-overlay-hop path where for each hop there is an
-possible indipendent TCP connection. In order to achieve that, a session table
-needs to be established and continuously updated, while TCP headers need to be
-accordingly modified to allow data transmission through the Ingress and Egress
-points. Nonetheless, care should be given to packets path: in order to avoid
-ill-calculated congestion windows, the incoming packets need to follow the same
-(or an equivalent one) path for the same instatiated connection. This phenomenom
-is based on the intermediate proxies that create the route: it is the case, for
-example, when an intermediate node sends back a spoofed ACK before the original
-packet is delivered to the real consegnee. In this senario, the sender, unaware
-of the proxy, will miscalculate the sending window based on the too low RTT.
+payload, that can be elaborated and modified accordingly, until it gets 
+decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is 
+not be aware of the packet elaboration, since all the original headers were 
+preserved (or slightly modified). In particular, IP Encapsulation within IP 
+allow the packets to make intermediate destinations that otherwise would not be 
+selected.
+TCP session recreation, instead, ``split'' the TCP session in two endpoints: an 
+Ingress, that manage the connection with the client and an Egress, that 
+communicate with the packet receiver creating a multi-overlay-hop path where 
+for each hop there is an possible indipendent TCP connection. In order to 
+achieve that, a session table needs to be established and continuously updated, 
+while TCP headers need to be accordingly modified to allow data transmission 
+through the Ingress and Egress points. Nonetheless, care should be given to 
+packets path: in order to avoid ill-calculated congestion windows, the incoming 
+packets need to follow the same (or an equivalent one) path for the same 
+instatiated connection. This phenomenom is based on the intermediate proxies 
+that create the route: it is the case, for example, when an intermediate node 
+sends back a spoofed ACK before the original packet is delivered to the real 
+consegnee. In this senario, the sender, unaware of the proxy, will miscalculate 
+the sending window based on the too low RTT.
 Other problems of TCP splitting are reliability and security, especially because

### 2018-10-10 draft.tex
Add something about Openbaton

@@
 \subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
-\paragraph{Openstack}
 \paragraph{Openvswitch}
-\paragraph{Openbaton}
+\paragraph{Openbaton} Created by the Institute Fraunhofer, it's a open-source, 
+customizable NFV MANO-compliant framework, that offers many configurations 
+options, and since it's written in Java, there is the possibility to add 
+components (plug-ins) dynamically. Openbaton can be installed on-permise 
+installations but it can also run on Docker containers. It's designed to work 
+with cloud providers like Amazon or Google Compute engine, but it offers 
+compability with on-permise cloud solutions like Openstack too.
 \begin{figure}[h]

### 2018-10-10 draft.tex
Fix grammar error
Add openbaton paragraph

@@
   \includegraphics[scale=0.5]{IPoverIP}
-  \caption[IP Encapsulation within IP packet schema]{An schema of IP 
+  \caption[IP Encapsulation within IP packet schema]{A schema of IP 
 Encapsulation within IP. Is noticeable how the original packet becomes the 
@@
 options, and since it's written in Java, there is the possibility to add 
-components (plug-ins) dynamically. Openbaton can be installed on-permise 
-installations but it can also run on Docker containers. It's designed to work 
-with cloud providers like Amazon or Google Compute engine, but it offers 
-compability with on-permise cloud solutions like Openstack too.
+components (plug-ins) dynamically. Openbaton can be installed as a normal 
+computer program (they offer packages for the most common distros) but it can 
+be deployed on Docker containers too. It's designed to work with cloud 
+providers like Amazon or Google Compute engine, but it offers compability with 
+on-permise cloud solutions like Openstack, where it deploys virtual machines 
+using the API the platform offers. It can store VNF configurations saved as 
+TOSCA YAML\todo{What does TOSCA mean? Find out} and it can handle multiple 
+PoPs\todo{Write down PoP meaning}. It offers a VNF lifecyle management 
+out-of-the-box, with the possibility to customize it based on the user needs. 
+Its modular design allows developers to change parts of the codebase with 
+custom ones, making the product really flexible to different platforms. 
+Performing this operation requires a deep knowlege of the framework though.
 \begin{figure}[h]

### 2018-10-10 draft.tex
Fix 'SFC pts' introduction

@@
 
-As today, when a packets gets send from a seder $S$, to a receiver $R$, the 
-packet gets travelling to te backbone receive appropriate in-hardware packet 
+When a packets are sent from a seder $S$, to a receiver $R$, they start
+travelling to the backbone receiving appropriate in-hardware packet 
 processing before being received by $R$. During this elaboration, it is 

### 2018-10-10 draft.tex
Fix repetition

@@
 \paragraph{Openvswitch}
-\paragraph{Openbaton} Created by the Institute Fraunhofer, it's a open-source, 
-customizable NFV MANO-compliant framework, that offers many configurations 
-options, and since it's written in Java, there is the possibility to add 
-components (plug-ins) dynamically. Openbaton can be installed as a normal 
-computer program (they offer packages for the most common distros) but it can 
-be deployed on Docker containers too. It's designed to work with cloud 
+\paragraph{Openbaton} Patrocined by the Institute Fraunhofer, it's a 
+open-source, customizable NFV MANO-compliant framework, that offers many 
+configurations options, and since it's written in Java, there is the possibility 
+to add components (plug-ins) dynamically. Openbaton can be installed as a 
+normal computer program (they offer packages for the most common distros) but it 
+can be deployed on Docker containers too. It's designed to work with cloud 
 providers like Amazon or Google Compute engine, but it offers compability with 

### 2018-10-10 draft.tex
Add bullet points

@@
 tools from the suggested ones. A detailed description of the architectural 
-implementation can be found in chapter~\ref{chap:archimpl}.
+implementation can be found in chapter~\ref{chap:archimpl}.\todo{Include prof. 
+survey about community components here}
+
+\item VNFs this are the core components that perform packets manipulation 
+\todo{Expand}
+
+\item NFV Management and Orchestration (MANO) \todo{Expand this part}
 

### 2018-10-10 draft.tex
Fix typo

@@
 particulary useful when more continers have to interact together,
-because\todo{Write down docker-commpose functionalities and describe the product
+because\todo{Write down docker-compose functionalities and describe the product
 a little bit} all the services rules are described in a simple YAML file.

### 2018-10-10 draft.tex
Add content about MANO and VNFs

@@
 \begin{itemize}
- \item Network Function Virtualization Infrastructure (NFVI): it's the heart of 
-the whole computation component, and here three sub-domains can be identified:
-\todo{Expand this section}
+ \item \textbf{Network Function Virtualization Infrastructure (NFVI)}: it's the
+   heart of the whole computation component, and here three sub-domains can be
+   identified: \todo{Expand this section}
 \begin{itemize} 
@@
 
-\item VNFs this are the core components that perform packets manipulation 
-\todo{Expand}
+\item \textbf{VNFs} this are the core components that perform packets
+  manipulation. This components are design to be lightweight services able to
+  elaborate a great amount of packets per seconds. Manupulation can be of two
+  typologies: stateful or stateless. In the first case, a state is maintained,
+  and the VNF can perform heuristics about data coming through it (especially
+  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
+  case, no state is maintained, and every packet is treated without any
+  particular assumption. \todo{Add ref to chap about SFC?}
 
-\item NFV Management and Orchestration (MANO) \todo{Expand this part}
+\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
+  Orchestrator (NFVO) is responsible for the end-to-end management and
+  orchestration of network services (NS), also called SFC. The MANO has
+  different responsabilities, such as managing the
+  istantiation/destruction/modification of the SFCs, that achieves interacting
+  with the VIM. \todo{Expand this part}
 

### 2018-10-11 draft.tex
Use Emacs formatting
Add footnote
Add part about MANO, add papers to read before adding new content

@@
 
-\author{Polonio Davide}
-\title{A thesis draft}
+\author{Polonio Davide} \title{A thesis draft}
 % The outline'd be something like this:
@@
  considerations that software solutions are being proposed to become the new way
-to deploy network components. While less efficent, they are faster to bring them 
-up, offering an unprecedented flexibility\footnote{SDN/NFV-Based Mobile Packet 
-Core Network Architectures: A Survey}. Furthermore, new virtualization 
+ to deploy network components. While less efficent, they are faster to bring
+ them up, offering an unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile
+   Packet Core Network Architectures: A Survey}. Furthermore, new virtualization
  technologies such as Cloud computing and Docker are becoming more and more
@@
  frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to
-easily manage the application lifecyle, i.e. deployment, scale up/down
-and finally, removal.
+ easily manage the application lifecyle, i.e. deployment, scale up/down and
+ finally, removal.
 
@@
    some reference, explain what it is.}, where the necessity for better TCP/IP
-transmission through satellite connections is the main requirement. To
-reach this goal, the project specifications suggest to exploit the 5G incoming
+ transmission through satellite connections is the main requirement. To reach
+ this goal, the project specifications suggest to exploit the 5G incoming
  technology and use the NFV-MANO architecture to perform first packets
@@
 same time, packet incapsulation present itself as the only faseable solution:
-every packet incoming in the VNFs has already been incapsulated by the MANO,
-\todo{Not sure about this, please give it a check - Not sure but I think that
-it is not part of the MANO the encapsulation of the packets. Also I think that
+every packet incoming in the VNFs has already been encapsulated by the MANO,
+\todo{Not sure about this, please give it a check - Not sure but I think that it
+  is not part of the MANO the encapsulation of the packets. Also I think that
   the main purpose of UDP encapsulation is not to hide the protocol used on the
   edges but 1) usa a ``quick protocol to exchange data among VNFs and 2) we are
-hiding the path not the protocol itself. As we were discussing, we are creating
-some sort of proxy, so the aim will be the same even if we support a plethora
-of protocols.''} making the whole architecture indipendent from the protocol a
-particolar flux of data uses. To achieve this, TUN/TAP
-\todo{Find out acronym - EPKT? (Encapsulated PacKeT)}
-interfaces were used to deploy a tunnel between VNFs, allowing them to pass
-packets encapsulated in UDP ones.
+  hiding the path not the protocol itself. As we were discussing, we are
+  creating some sort of proxy, so the aim will be the same even if we support a
+  plethora of protocols.''} making the whole architecture indipendent from the
+protocol a particolar flux of data uses. To achieve this, TUN/TAP interfaces 
+were used to deploy a tunnel between VNFs, allowing them to pass packets 
+encapsulated in UDP ones.
 
@@
 ??\todo{Find out the actual date} and written in Go, it allows different
-machines (called nodes from now on) to create an abstraction layer and to form
-a cluster where is possible to run Docker containers without brothering about
+machines (called nodes from now on) to create an abstraction layer and to form a
+cluster where is possible to run Docker containers without brothering about
 hardware constraints. Nodes can have specific roles and specialization, all of
 them managed by one or many master nodes, that actually manage the
-orchestration.
-On top of that, Kubernetes is virtualization-agnostic, meaning that it offers
-the possibility to change the virtualization software in use (i.e. from
-Container-based technology to Virtual machine software).
+orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
+that it offers the possibility to change the virtualization software in use
+(i.e. from Container-based technology to Virtual machine software).
 \begin{figure}[h]
- \centering
- \includegraphics[scale=0.35]{kubernetes_logo}
+ \centering \includegraphics[scale=0.35]{kubernetes_logo}
  \caption{Kubernetes Logo}
@@
 In recent years, with the new hardware capabilities and the recent development
-of in-kernel virtualization systems (such as Hyper-V) this technology
-begin to be adopted. Virtualization allows to run different systems on the
-same machine, making them completely isolated and then more resilient to
-failures. The back of the medal, though, is that virtual machines require large
-amount of resources, especially memory, because solutions like copy-and-write
-of in-memory pages are not viable anymore (since the kernel gets duplicated
-too), leading to duplication of loaded libraries and assets in the memory of
-the host.
-For large deployment containing only simple services (e.g. a backend
-application serving a website or a database) this ends up in a waste of
-resources. \todo{Talk about lxc containers and how Docker solved magically the
-problem.} It's here that, in ??\todo{Find Docker date of birth}, Docker was
-created, basing its solution on an already existing product: Linux Container
-(LXC). From this framework, Docker built an entire ecosystem, consisting in a
-client/server model giving the possibility for users to simply launch, scale and
-delete containers (locally or remotely with Docker-machine), a repository
-system where images, a layer-based ``core'' where containers start running from
-when they are launched, can be stored and tagged in a sort of version control
-system. Finally, another couple of solutions, called Docker-compose, and Docker
-Swarm have the purpose to, respectively, orchestrate containers in a single or
-a clustered system.
+of in-kernel virtualization systems (such as Hyper-V) this technology begin to
+be adopted. Virtualization allows to run different systems on the same machine,
+making them completely isolated and then more resilient to failures. The back of
+the medal, though, is that virtual machines require large amount of resources,
+especially memory, because solutions like copy-and-write of in-memory pages are
+not viable anymore (since the kernel gets duplicated too), leading to
+duplication of loaded libraries and assets in the memory of the host. For large
+deployment containing only simple services (e.g. a backend application serving a
+website or a database) this ends up in a waste of resources. \todo{Talk about
+  lxc containers and how Docker solved magically the problem.} It's here that,
+in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
+on an already existing product: Linux Container (LXC). From this framework,
+Docker built an entire ecosystem, consisting in a client/server model giving the
+possibility for users to simply launch, scale and delete containers (locally or
+remotely with Docker-machine), a repository system where images, a layer-based
+``core'' where containers start running from when they are launched, can be
+stored and tagged in a sort of version control system. Finally, another couple
+of solutions, called Docker-compose, and Docker Swarm have the purpose to,
+respectively, orchestrate containers in a single or a clustered system.
 \begin{figure}[t]
- \centering
- \includegraphics[scale=0.7]{docker_logo}
+ \centering \includegraphics[scale=0.7]{docker_logo}
  \caption{Official Docker logo}
@@
 infrastructure transparent: data management, as network one, are totaly managed
-by Docker. Recently, Docker-compose support has been introduced, making
-this one-node tool available for clustered systems. At the end of the day, on
-one hand Docker Swarm provides an easy way to set up a cluster system with all
-the tools configured out-of-the-box, without the necessity to set networking
+by Docker. Recently, Docker-compose support has been introduced, making this
+one-node tool available for clustered systems. At the end of the day, on one
+hand Docker Swarm provides an easy way to set up a cluster system with all the
+tools configured out-of-the-box, without the necessity to set networking
 configurations or installing ad-hoc storage solutions. On the other hand,
@@
 because\todo{Write down docker-compose functionalities and describe the product
-a little bit} all the services rules are described in a simple YAML file.
-With the birth of Docker Swarm, work has started to port this tool to the Swarm
+  a little bit} all the services rules are described in a simple YAML file. With
+the birth of Docker Swarm, work has started to port this tool to the Swarm
 framework. At the time of this writing, the procedure to port a Docker-compose
 configuration in Docker Swarm is still not completely transparent, and it
-requires a sort of compilation, where Docker-compose bundles all the
-necessaries resources into a binary file that Docker Swarm will use to allocate
-the necessary resources.
+requires a sort of compilation, where Docker-compose bundles all the necessaries
+resources into a binary file that Docker Swarm will use to allocate the
+necessary resources.
 
@@
 Starting with the first step of creating a MANO able to process incoming data
-packets through VNF functions, we encountered that many networking tools
-already present in the market required some tweaking and some integrations,
-shifting our goal to create a complete  European Telecommunications Standards
-Institutes (ETSI) Management and orchestrator (MANO) testbed instead, following
-the specifications suggested in the RFC 7665, thus implementing only the first
-three requisites, without digging in the satellite data flow optimization. In 
+packets through VNF functions, we encountered that many networking tools already
+present in the market required some tweaking and some integrations, shifting our
+goal to create a complete European Telecommunications Standards Institutes
+(ETSI) Management and orchestrator (MANO) testbed instead, following the
+specifications suggested in the RFC 7665, thus implementing only the first three
+requisites, without digging in the satellite data flow optimization. In
 particular, we discovered how, these tools, were suitable to create ETSI MANO
-and VNFs using virtual machine or exploiting cloud tecnhologies, while they 
-were not designed with enough flexibility to integrate with Docker. In the next 
-chapter we are going to take a deep analysis of the cited technologies, in 
-order to provide to the reader enough context to be able to understand our 
-design choices we are going to describe later.
+and VNFs using virtual machine or exploiting cloud tecnhologies, while they were
+not designed with enough flexibility to integrate with Docker. In the next
+chapter we are going to take a deep analysis of the cited technologies, in order
+to provide to the reader enough context to be able to understand our design
+choices we are going to describe later.
 
@@
 \begin{figure}[t]
- \centering
- \includegraphics[scale=2]{etsi_arch}
+ \centering \includegraphics[scale=2]{etsi_arch}
  \caption{A schematic image representing the high level NFV architecture
@@
    accessing it.
- \item \textbf{Compute domain} composed by the Physical Compute and the 
-Physical Storage. These are the real resources which are virtualized via 
-software.
+ \item \textbf{Compute domain} composed by the Physical Compute and the Physical
+   Storage. These are the real resources which are virtualized via software.
  \item \textbf{Infrastructure domain} which has the Physical Network, and its
@@
   case, no state is maintained, and every packet is treated without any
-  particular assumption. \todo{Add ref to chap about SFC?}
+  particular assumption. Following the RFC 7665, the VNF are also called Service
+  Function (SF)\footnote{From now on the two terminologies will be use in a
+    interchangeable way.}. \todo{Add ref to chap about SFC?}
 
@@
   istantiation/destruction/modification of the SFCs, that achieves interacting
-  with the VIM. \todo{Expand this part}
+  with the VIM.\todo{CITE: SDN/NFV-Based Mobile Packet Core Network
+Architectures: A Survey}. Finally, the mano can integrate with 
+Operational and Business Support Systems (OSS/BSS), \todo{Expand this part with 
+the following papers: https://www.tandfonline.com/doi/abs/10.1057/jors.1995.176 
+https://ieeexplore.ieee.org/abstract/document/6963800 
+https://www.computer.org/csdl/proceedings/ifcsta/2009/3930/01/3930a466-abs.html}
 
@@
 When a packets are sent from a seder $S$, to a receiver $R$, they start
-travelling to the backbone receiving appropriate in-hardware packet 
-processing before being received by $R$. During this elaboration, it is 
-important to maintain the transparency of the connection, so that the network 
-users are not aware of the packets elaboration at all. To perform that, many 
-solutions are available: TUN/TAP packet incapsulation, IP Encapsulation within 
-IP or TCP session recreation (know also as TCP split).
+travelling to the backbone receiving appropriate in-hardware packet processing
+before being received by $R$. During this elaboration, it is important to
+maintain the transparency of the connection, so that the network users are not
+aware of the packets elaboration at all. To perform that, many solutions are
+available: TUN/TAP packet incapsulation, IP Encapsulation within IP or TCP
+session recreation (know also as TCP split).
 
@@
 \begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.5]{IPoverIP}
+  \centering \includegraphics[scale=0.5]{IPoverIP}
   \caption[IP Encapsulation within IP packet schema]{A schema of IP
     Encapsulation within IP. Is noticeable how the original packet becomes the
-payload of a new one, thus being encapsulated. All the informations regarding 
-the original packet remains untouched: when this kind of packet gets send 
-from a machine to another, the OS will remove the new packet header and will 
-present to the user-space program receiving the data its payload: a IP 
-packet. Packets encapsulation can be applied with higher layers too, thus 
-allowing for TCP and UDP packet incapsulation.}
+    payload of a new one, thus being encapsulated. All the informations
+    regarding the original packet remains untouched: when this kind of packet
+    gets send from a machine to another, the OS will remove the new packet
+    header and will present to the user-space program receiving the data its
+    payload: a IP packet. Packets encapsulation can be applied with higher
+    layers too, thus allowing for TCP and UDP packet incapsulation.}
   \label{chap:prjan:img:ip_over_ip}
@@
 payload, that can be elaborated and modified accordingly, until it gets
-decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is 
-not be aware of the packet elaboration, since all the original headers were 
+decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is not
+be aware of the packet elaboration, since all the original headers were
 preserved (or slightly modified). In particular, IP Encapsulation within IP
 allow the packets to make intermediate destinations that otherwise would not be
-selected.
-TCP session recreation, instead, ``split'' the TCP session in two endpoints: an 
-Ingress, that manage the connection with the client and an Egress, that 
-communicate with the packet receiver creating a multi-overlay-hop path where 
-for each hop there is an possible indipendent TCP connection. In order to 
+selected. TCP session recreation, instead, ``split'' the TCP session in two
+endpoints: an Ingress, that manage the connection with the client and an Egress,
+that communicate with the packet receiver creating a multi-overlay-hop path
+where for each hop there is an possible indipendent TCP connection. In order to
 achieve that, a session table needs to be established and continuously updated,
@@
 consegnee. In this senario, the sender, unaware of the proxy, will miscalculate
-the sending window based on the too low RTT.
-Other problems of TCP splitting are reliability and security, especially because
-the end to end TCP logic is no longer maintained: a server failure may cause to
-an unaware client to believe that all the packets have reached the destination.
-TCP session recreation seems burdersome and it doesn't seem to add any
-significant benefit at the first sight, but it offers the flexibility to perform
-extensive packet manipulation: since the packet its recreated every time at the
-Ingress and Egress points, while in the Network Function Virtualization
-Infrastructure (NFVI), the original packet can be stripped of its headers (that
-could be saved in the session table) to slightly improve internal packet
-transmission. \todo{Insert TCP session recreation schema} On top of that, it
-offers the possibility to use different TCP flavours, hence increasing the
-communication speed in heterogeneous networks (e.g. in case of satellite
-connections TCP Hybla can be employed, while it is known that HighSpeed TCP
-gives best results when applied in optical fiber-based connections).
+the sending window based on the too low RTT. Other problems of TCP splitting are
+reliability and security, especially because the end to end TCP logic is no
+longer maintained: a server failure may cause to an unaware client to believe
+that all the packets have reached the destination. TCP session recreation seems
+burdersome and it doesn't seem to add any significant benefit at the first
+sight, but it offers the flexibility to perform extensive packet manipulation:
+since the packet its recreated every time at the Ingress and Egress points,
+while in the Network Function Virtualization Infrastructure (NFVI), the original
+packet can be stripped of its headers (that could be saved in the session table)
+to slightly improve internal packet transmission. \todo{Insert TCP session
+  recreation schema} On top of that, it offers the possibility to use different
+TCP flavours, hence increasing the communication speed in heterogeneous networks
+(e.g. in case of satellite connections TCP Hybla can be employed, while it is
+known that HighSpeed TCP gives best results when applied in optical fiber-based
+connections).
 
@@
 \label{chap:prjan:sec:openstack}
-Created in ??\todo{Find out Openstack date of creation}, 
-Openstack allows on-permise cloud installations, using bare-metal resources to 
-provide common cloud services as object storage, virtual machine deploy, 
-virtual networking. Furthermore its modularity offers the possibility to add 
-additional components, even proprietary, to achieve a complete cloud solution.
+Created in ??\todo{Find out Openstack date of creation}, Openstack allows
+on-permise cloud installations, using bare-metal resources to provide common
+cloud services as object storage, virtual machine deploy, virtual networking.
+Furthermore its modularity offers the possibility to add additional components,
+even proprietary, to achieve a complete cloud solution.
 \begin{figure}[t]
- \centering
- \includegraphics[scale=0.58]{openstack_logo}
+ \centering \includegraphics[scale=0.58]{openstack_logo}
  \caption{Openstack logo}
@@
 configurations options, and since it's written in Java, there is the possibility
-to add components (plug-ins) dynamically. Openbaton can be installed as a 
-normal computer program (they offer packages for the most common distros) but it 
-can be deployed on Docker containers too. It's designed to work with cloud 
-providers like Amazon or Google Compute engine, but it offers compability with 
-on-permise cloud solutions like Openstack, where it deploys virtual machines 
-using the API the platform offers. It can store VNF configurations saved as 
-TOSCA YAML\todo{What does TOSCA mean? Find out} and it can handle multiple 
+to add components (plug-ins) dynamically. Openbaton can be installed as a normal
+computer program (they offer packages for the most common distros) but it can be
+deployed on Docker containers too. It's designed to work with cloud providers
+like Amazon or Google Compute engine, but it offers compability with on-permise
+cloud solutions like Openstack, where it deploys virtual machines using the API
+the platform offers. It can store VNF configurations saved as TOSCA
+YAML\todo{What does TOSCA mean? Find out} and it can handle multiple
 PoPs\todo{Write down PoP meaning}. It offers a VNF lifecyle management
 out-of-the-box, with the possibility to customize it based on the user needs.
-Its modular design allows developers to change parts of the codebase with 
-custom ones, making the product really flexible to different platforms. 
-Performing this operation requires a deep knowlege of the framework though.
+Its modular design allows developers to change parts of the codebase with custom
+ones, making the product really flexible to different platforms. Performing this
+operation requires a deep knowlege of the framework though.
 \begin{figure}[h]
- \centering
- \includegraphics[scale=0.45]{openbaton_logo}
+ \centering \includegraphics[scale=0.45]{openbaton_logo}
  \caption{Openbaton logo}

### 2018-10-11 draft.tex
Start studies about available technologies in the market

@@
 % network architecture, proposed solution, requisites)
-%   + Related works
+%   + Related works (how are we gonna do that??)
 % - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
@@
 %   + Choseen tecnhologies and description of the 'reason why'
-% - Project description
+% - Project description (tell how we've done stuff)
 %   + Initial solution with Openstack and Tacker
@@
 
-\section{SFC packet transmission strategies}
+\section{SFC}
+
+\todo{Here it's missing an introduction about SFC and its terminology, need to 
+include parts of RFC 7665}
+
+\subsection{Packet transmission strategies}
 
@@
 
-\todo{Explain TCP session recreation}
+\section{Available technologies in the market}
 
-\section{Openstack}
+Before starting any effective work on the project it has been decided to 
+perform an analysis of the possible technologies avvailable in the market, 
+performing a choice between them and exploiting their possibile ``weak 
+points''. Proceding with a top-down approach, the frameworks were choosen from 
+the one that had the biggest role in the testbed to conclude with the 
+less-important components.
+
+\subsection{NFVI}
+
+A critical component, as the MANO one, is the NFVI. Without it, network 
+resources can't be allocated, and since our project was focused on dynamically 
+allocating containerized applications on cloud platforms, we decided to start 
+analizing the right framework for this component first.
+
+\subsection{Openstack}
 \label{chap:prjan:sec:openstack}

### 2018-10-14 draft.tex
Add content about 5G

@@
 \usepackage{todonotes}
+\usepackage{graphicx}
 
@@
  % This is just a draft, nothing suitable for the final product: sources are in
- % form of a footnote
+ % form of a todo
 
@@
 
- Through the history of internet, many factor changed and new innovations came,
- but at the end of the day a foundamental key remained: TCP/IP performance are,
- indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP
+ Through the history of internet, many factors changed and new innovations came,
+ but at the end of the day a foundamental key remained: connection performances
+ are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
    performes quite well in wired connections, while in wireless one, such as
    wireless/satellite, sucks ass} In this context network optimization plays an
- important role, and it is for this reason that network packets are processed
- while traversing the network in order to make their transmission as fast as
- possible. On top of that, mobile traffic is continuously surging every year,
- with the backbones having more and more data to process. In this context,
- in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in
- fact, is very high compared to the requirements of the network. With this
- trend, in-hardware solutions will not be able to keep up. Is with this
- considerations that software solutions are being proposed to become the new way
- to deploy network components. While less efficent, they are faster to bring
- them up, offering an unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile
-   Packet Core Network Architectures: A Survey}. Furthermore, new virtualization
- technologies such as Cloud computing and Docker are becoming more and more
- mature for being employed in large-scale production environments, where new
- frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to
- easily manage the application lifecyle, i.e. deployment, scale up/down and
- finally, removal.
+ important role, and it is for this reason that network packets are elaborated,
+ compressed and processed while traversing the network in order to make their
+ transmission as fast as possible. On top of that, mobile traffic is
+ continuously surging every year, with the backbones having more and more data
+ to deliver, making in-hardware backbone solutions less efficent. Their
+ time-to-deploy, in fact, is very high compared to the requirements of the
+ network. With this trend, in-hardware solutions will not be able to keep up. Is
+ with this considerations that software a software approach is being proposed to
+ become the new way to deploy network components. While less efficent, they are
+ faster to bring them up, offering an unprecedented flexibility\todo{CITE:
+   SDN/NFV-Based Mobile Packet Core Network Architectures: A Survey}.
+ Furthermore, new virtualization technologies such as Cloud computing and Docker
+ are becoming more and more mature for being employed in large-scale production
+ environments, where new frameworks, called orchestrators (e.g. Kubernetes,
+ Docker Swarm), allow to easily manage the application lifecyle, i.e.
+ deployment, scale up/down and finally, removal.
+
+\section{The upcoming connectivity standard: 5G}
+The continuous innovation in the mobile network connectivity is leading to the
+creation of a new standard, the 5G, which is estimated to arrive to the market
+constumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
+alliace, a group composed by the major players in the field of mobile
+connectivity, the 5G aims to offer at the end-user a new way to browse the web,
+download and watch interactive content. Relatively to 4G, 5G points to have data
+rates $10$ times better, with $10$ times smaller end-to-end latency and
+increased connection densisty by $100$ times\todo{CITE}.
+
+\begin{table}[t]
+\centering
+\resizebox{\textwidth}{!}{%
+  \begin{tabular}{p{4cm}|p{5,5cm}|p{5,5cm}}
+\textbf{Attribute}                                                    & \textbf{LTE capability}                                                                                                                                     & \textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
+\textbf{Data rate (per user)}                                         & Up to 100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).
+& 10X expected on average and peak rates and 100X expected on cell edge \\
+\textbf{End-to-end latency}                                           & 10 ms for two-way RAN (pre-scheduled).
+Typically, up to 50 ms end-to-end if other factors are considered (e.g., transmission, CN, internet, proxy servers).
+& 10X (smaller)                                                         \\
+\textbf{Mobility}                                                     & Functional up to 350 km/h (for certain bands up to 500 km/h).
+No support for civil aviation.
+& 1.5X                                                                  \\
+\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & Typically $\sim$2,000 active users/km2.
+& 100X
+\end{tabular}%
+}
+\caption[5G improvement over LTE]{An extract from the official 5G white paper illustating the improvements of 5G relatively to LTE connections.}
+\label{chap:intro:table:ltevs5g}
+\end{table}
+
+Part of the new requirements can be satisfied using a large radio spectrum with
+higher frequencies. But the utilization of higher frequencies mean that the
+radio signals can be easly interrupted by physical objects, like buildings and
+many geographical elements (such as hills and mountains), clashing with the
+expectation of an ever-reachable connectivity. It is here where virtualization
+plays an important role. In fact, the re-design of some network components today
+existing via hardware can transform a monolitic networking approach to a modular
+one, exploting the flexibility that Virtual Network Functions (VNFs)\todo{CITE
+  VNF-P: A Model for Efficient Placement of Virtualized Network Functions} can
+offer thanks to virtualization, closing the gap to the use-case fullfilment
+defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture}.
+
+\subsection{5G architecture}
+
+With the constraints placed on the requirements formulated by the NGMN alliance,
+5G envisage a multi-layered architecture, based on three main layer:\todo{CITE:
+  NGMN View on 5G architecture}
+\begin{itemize}
+\item \textbf{infrastructure resouce layer}: physical resources that are exposed
+  via a virtualized interface, and that can be monitored using specific APIs
+\item \textbf{business enablement layer}: where a library of functions and
+  deployment is contained, and its configuration is accessible via APIs
+\item \textbf{business application layer}: layer that contains specific
+  applications and services of the operator
+\end{itemize}
+
+\subsubsection{Network slicing}
+The role of a ``network slice'' in a 5G architecture is to specificatly handle a
+Control-place of a praticular service (e.g. smartphones traffic, autonomous
+driving, massive IoT), deploying resources in a manner that assure the required
+latency, security and reliability. While some legacy of very peculiar services
+could require specific hardware, the common resources between services could be
+shared in a virtualized way, providing auto-scaling capabilities in services
+that are under heavy network pressure.
 

### 2018-10-14 draft.tex
Fix and add TODOs

@@
 download and watch interactive content. Relatively to 4G, 5G points to have data
-rates $10$ times better, with $10$ times smaller end-to-end latency and
+rates $10$ times better, with $10$ times smaller end-to-end latency and an
 increased connection densisty by $100$ times\todo{CITE}.
@@
 Part of the new requirements can be satisfied using a large radio spectrum with
-higher frequencies. But the utilization of higher frequencies mean that the
-radio signals can be easly interrupted by physical objects, like buildings and
+higher frequencies. The utilization of higher frequencies, though, mean that the
+radio signals can be easly disrupted by physical objects, like buildings and
 many geographical elements (such as hills and mountains), clashing with the
@@
 With the constraints placed on the requirements formulated by the NGMN alliance,
-5G envisage a multi-layered architecture, based on three main layer:\todo{CITE:
+5G envisage a multi-layered architecture, based on three main layers:\todo{CITE:
   NGMN View on 5G architecture}
@@
 \end{itemize}
+
+\todo{Add three-layer image here?}
 

### 2018-10-14 res/sections/03-Inscription.tex
Applied thesis template

%**************************************************************
% Dedica
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Dedica}{Dedica}

\vspace*{3cm}

\begin{flushright}{
	\slshape
	dedica} \\
	\medskip
\end{flushright}

\medskip

### 2018-10-14 res/sections/04-Credits.tex
Applied thesis template

%**************************************************************
% Ringraziamenti
%**************************************************************
\cleardoublepage
\phantomsection
\thispagestyle{empty}
\pdfbookmark{Credits}{credits}

%\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\section*{Credits}

\noindent \textit{Ringraziamenti}\\
\bigskip

\noindent\textit{\myLocation, \myTime}
\hfill \myName

\endgroup

### 2018-10-15 res/sections/07-Introduction.tex
Move chapters in separate files, fix gitignore

--- 2018-10-14 draft.tex
@@
-\documentclass[10pt]{book}
-
-\usepackage{todonotes}
-\usepackage{graphicx}
-
-\graphicspath{{res/img/}}
-
-\author{Polonio Davide} \title{A thesis draft}
-% The outline'd be something like this:
-%  Thesis outline
-% 
-% - Introduction about the VIBES project (context, 5G, the need for a new 
-% network architecture, proposed solution, requisites)
-%   + Related works (how are we gonna do that??)
-% - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
-% Docker, Docker swarm)
-%   + Choseen tecnhologies and description of the 'reason why'
-% - Project description (tell how we've done stuff)
-%   + Initial solution with Openstack and Tacker
-%   + From Tacker to Openbaton
-%   + From Openbaton with Openstack to Openbaton with Docker
-%   + Exploration of Docker orchestrators (Swam vs Kubernetes)
-%   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
-%   + Openbaton and Kubernetes integration (lack of a implemented Driver, 
-% integration with existinf VNFM)
-% - Case study
-%   + Analysis of TCP/UDP performance
-% - Results
-% - Future works
-% - Conclusions
-
-\makeindex
-\begin{document}
- \maketitle
-
  % This is just a draft, nothing suitable for the final product: sources are in
@@
 the next chapter.\todo{This should be put at the end of the introduction.}
-
-\chapter{Project Analysis}
-\label{chap:prjan}
-
-As in chapter~\ref{chap:intro}, the creation of a full ETSI MANO compliant
-architecture, following the suggestions described in RFC 7665, require the
-integration of multiple tools (some of them have been already briefly introduced
-in~\ref{chap:intro}). The tools were choosen after their analysis, where pro and
-cons have been studied and compared to the necessity of this thesis. Since the
-project has been developed with the cloud provided by the University of Padua,
-Openstack (which a product analysis can be found later
-in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
-
-\section{Architecture overview}
-
-During the testbed development the software architecture diverged from the
-original one, experiencing different evolutions and modifications.
-
-\subsection{VIBES architectural description}
-
-In the original VIBES architectural design, based on leveraging virtualization
-technologies, is possible to identify different macro-components, each of these
-specialized in fullfilling a specific goal. For the ETSI standards, the Network
-Function Virtualization (NFV) design should be composed as follows:
-\begin{figure}[t]
- \centering \includegraphics[scale=2]{etsi_arch}
- \caption{A schematic image representing the high level NFV architecture
-   proposed by ESA.}
- \label{chap:prjan:img:etsi_arch}
-\end{figure}
-
-\begin{itemize}
- \item \textbf{Network Function Virtualization Infrastructure (NFVI)}: it's the
-   heart of the whole computation component, and here three sub-domains can be
-   identified: \todo{Expand this section}
-\begin{itemize} 
- \item \textbf{Hypervisor domain} that is composed of the Virtual Compute, the
-   Virtual Storage, and the Virtualization Layer. These components are used to
-   virtualize the computational and storage resources, providing a layer for
-   accessing it.
- \item \textbf{Compute domain} composed by the Physical Compute and the Physical
-   Storage. These are the real resources which are virtualized via software.
- \item \textbf{Infrastructure domain} which has the Physical Network, and its
-   virtualized counterpart as made available by the Virtualization Layer.
-\end{itemize}
-
-In conventional virtualization systems these resources are usually separated
-from the host operating system, providing better isolation but wrost
-performance. The ETSI infrastructure, instead, tries with container
-virtualization technology to push for a better NFV responsiveness\todo{Check
-  term - I don't think it actually exists}, removing the additional operating
-system (called ``guest'' OS) required by the usual hypervisor-oriented approach.
-
-To realize the required components, specific tools were suggested in the VNF
-technical proposal: regarding the NFV (network infrastructure) domain
-Docker-Compose was proposed as a viable solution, while to store PEP instances
-docker swarm was described as a possible candidate.
-
-We performed an accured analysis of the tools choosen based on their maturity,
-on the community and on the technical support offered (e.g. user manuals,
-developer documentation), which, at the end, commited us to choose different
-tools from the suggested ones. A detailed description of the architectural
-implementation can be found in chapter~\ref{chap:archimpl}.\todo{Include prof.
-  survey about community components here}
-
-\item \textbf{VNFs} this are the core components that perform packets
-  manipulation. This components are design to be lightweight services able to
-  elaborate a great amount of packets per seconds. Manupulation can be of two
-  typologies: stateful or stateless. In the first case, a state is maintained,
-  and the VNF can perform heuristics about data coming through it (especially
-  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
-  case, no state is maintained, and every packet is treated without any
-  particular assumption. Following the RFC 7665, the VNF are also called Service
-  Function (SF)\footnote{From now on the two terminologies will be use in a
-    interchangeable way.}. \todo{Add ref to chap about SFC?}
-
-\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
-  Orchestrator (NFVO) is responsible for the end-to-end management and
-  orchestration of network services (NS), also called SFC. The MANO has
-  different responsabilities, such as managing the
-  istantiation/destruction/modification of the SFCs, that achieves interacting
-  with the VIM.\todo{CITE: SDN/NFV-Based Mobile Packet Core Network
-Architectures: A Survey}. Finally, the mano can integrate with 
-Operational and Business Support Systems (OSS/BSS), \todo{Expand this part with 
-the following papers: https://www.tandfonline.com/doi/abs/10.1057/jors.1995.176 
-https://ieeexplore.ieee.org/abstract/document/6963800 
-https://www.computer.org/csdl/proceedings/ifcsta/2009/3930/01/3930a466-abs.html}
-
-\end{itemize}
-
-\section{SFC}
-
-\todo{Here it's missing an introduction about SFC and its terminology, need to 
-include parts of RFC 7665}
-
-\subsection{Packet transmission strategies}
-
-When a packets are sent from a seder $S$, to a receiver $R$, they start
-travelling to the backbone receiving appropriate in-hardware packet processing
-before being received by $R$. During this elaboration, it is important to
-maintain the transparency of the connection, so that the network users are not
-aware of the packets elaboration at all. To perform that, many solutions are
-available: TUN/TAP packet incapsulation, IP Encapsulation within IP or TCP
-session recreation (know also as TCP split).
-
-TUN/TAP and IP Encapsulation within IP (initially defined in RFC 2003 with the
-aim to deliver IP packets over Mobile IP for mobile networks) maintain the
-connection transparency incapsulating the packet in another one.
-\begin{figure}[t]
-  \centering \includegraphics[scale=0.5]{IPoverIP}
-  \caption[IP Encapsulation within IP packet schema]{A schema of IP
-    Encapsulation within IP. Is noticeable how the original packet becomes the
-    payload of a new one, thus being encapsulated. All the informations
-    regarding the original packet remains untouched: when this kind of packet
-    gets send from a machine to another, the OS will remove the new packet
-    header and will present to the user-space program receiving the data its
-    payload: a IP packet. Packets encapsulation can be applied with higher
-    layers too, thus allowing for TCP and UDP packet incapsulation.}
-  \label{chap:prjan:img:ip_over_ip}
-\end{figure}
-
-In this way, the original packet remains preserved and it becomes the new packet
-payload, that can be elaborated and modified accordingly, until it gets
-decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is not
-be aware of the packet elaboration, since all the original headers were
-preserved (or slightly modified). In particular, IP Encapsulation within IP
-allow the packets to make intermediate destinations that otherwise would not be
-selected. TCP session recreation, instead, ``split'' the TCP session in two
-endpoints: an Ingress, that manage the connection with the client and an Egress,
-that communicate with the packet receiver creating a multi-overlay-hop path
-where for each hop there is an possible indipendent TCP connection. In order to
-achieve that, a session table needs to be established and continuously updated,
-while TCP headers need to be accordingly modified to allow data transmission
-through the Ingress and Egress points. Nonetheless, care should be given to
-packets path: in order to avoid ill-calculated congestion windows, the incoming
-packets need to follow the same (or an equivalent one) path for the same
-instatiated connection. This phenomenom is based on the intermediate proxies
-that create the route: it is the case, for example, when an intermediate node
-sends back a spoofed ACK before the original packet is delivered to the real
-consegnee. In this senario, the sender, unaware of the proxy, will miscalculate
-the sending window based on the too low RTT. Other problems of TCP splitting are
-reliability and security, especially because the end to end TCP logic is no
-longer maintained: a server failure may cause to an unaware client to believe
-that all the packets have reached the destination. TCP session recreation seems
-burdersome and it doesn't seem to add any significant benefit at the first
-sight, but it offers the flexibility to perform extensive packet manipulation:
-since the packet its recreated every time at the Ingress and Egress points,
-while in the Network Function Virtualization Infrastructure (NFVI), the original
-packet can be stripped of its headers (that could be saved in the session table)
-to slightly improve internal packet transmission. \todo{Insert TCP session
-  recreation schema} On top of that, it offers the possibility to use different
-TCP flavours, hence increasing the communication speed in heterogeneous networks
-(e.g. in case of satellite connections TCP Hybla can be employed, while it is
-known that HighSpeed TCP gives best results when applied in optical fiber-based
-connections).
-
-During the testbed development we first tried the TUN/TAP solution, that
-revealed to be not completely transparent and with major performance drawbacks
-(i.e. packet transmission in a LAN connection reached latencies of 50ms)
-\todo{Explain the constraint of point-to-point connections, while we need
-  dynamic SFC)}, so we choose to opt for a TCP session recreation solution,
-implementing a little back-end with the aim of saving in a persistent storage
-the SFC the packet has to perform and additional metadata.
-
-\section{Available technologies in the market}
-
-Before starting any effective work on the project it has been decided to 
-perform an analysis of the possible technologies avvailable in the market, 
-performing a choice between them and exploiting their possibile ``weak 
-points''. Proceding with a top-down approach, the frameworks were choosen from 
-the one that had the biggest role in the testbed to conclude with the 
-less-important components.
-
-\subsection{NFVI}
-
-A critical component, as the MANO one, is the NFVI. Without it, network 
-resources can't be allocated, and since our project was focused on dynamically 
-allocating containerized applications on cloud platforms, we decided to start 
-analizing the right framework for this component first.
-
-\subsection{Openstack}
-\label{chap:prjan:sec:openstack}
-Created in ??\todo{Find out Openstack date of creation}, Openstack allows
-on-permise cloud installations, using bare-metal resources to provide common
-cloud services as object storage, virtual machine deploy, virtual networking.
-Furthermore its modularity offers the possibility to add additional components,
-even proprietary, to achieve a complete cloud solution.
-\begin{figure}[t]
- \centering \includegraphics[scale=0.58]{openstack_logo}
- \caption{Openstack logo}
- \label{chap:prjan:img:openstack_logo}
-\end{figure}
-
-\todo{This section needs to be expanded a lot!!! In particular, we should put
-  some focus about VNF and MANO, introducing them}
-
-\subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
-\paragraph{Openvswitch}
-\paragraph{Openbaton} Patrocined by the Institute Fraunhofer, it's a 
-open-source, customizable NFV MANO-compliant framework, that offers many
-configurations options, and since it's written in Java, there is the possibility
-to add components (plug-ins) dynamically. Openbaton can be installed as a normal
-computer program (they offer packages for the most common distros) but it can be
-deployed on Docker containers too. It's designed to work with cloud providers
-like Amazon or Google Compute engine, but it offers compability with on-permise
-cloud solutions like Openstack, where it deploys virtual machines using the API
-the platform offers. It can store VNF configurations saved as TOSCA
-YAML\todo{What does TOSCA mean? Find out} and it can handle multiple
-PoPs\todo{Write down PoP meaning}. It offers a VNF lifecyle management
-out-of-the-box, with the possibility to customize it based on the user needs.
-Its modular design allows developers to change parts of the codebase with custom
-ones, making the product really flexible to different platforms. Performing this
-operation requires a deep knowlege of the framework though.
-\begin{figure}[h]
- \centering \includegraphics[scale=0.45]{openbaton_logo}
- \caption{Openbaton logo}
- \label{chap:prjan:img:openbaton_logo}
-\end{figure}
-
-\paragraph{SDN?}
-
-\chapter{Architectural implementation}
-\label{chap:archimpl}
-\end{document}

### 2018-10-15 res/sections/10-ProjectAnalysis.tex
Move chapters in separate files, fix gitignore

--- 2018-10-11 draft.tex
@@
-\documentclass[10pt]{book}
-
-\usepackage{todonotes}
-
-\graphicspath{{res/img/}}
-
-\author{Polonio Davide} \title{A thesis draft}
-% The outline'd be something like this:
-%  Thesis outline
-% 
-% - Introduction about the VIBES project (context, 5G, the need for a new 
-% network architecture, proposed solution, requisites)
-%   + Related works (how are we gonna do that??)
-% - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
-% Docker, Docker swarm)
-%   + Choseen tecnhologies and description of the 'reason why'
-% - Project description (tell how we've done stuff)
-%   + Initial solution with Openstack and Tacker
-%   + From Tacker to Openbaton
-%   + From Openbaton with Openstack to Openbaton with Docker
-%   + Exploration of Docker orchestrators (Swam vs Kubernetes)
-%   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
-%   + Openbaton and Kubernetes integration (lack of a implemented Driver, 
-% integration with existinf VNFM)
-% - Case study
-%   + Analysis of TCP/UDP performance
-% - Results
-% - Future works
-% - Conclusions
-
-\makeindex
-\begin{document}
- \maketitle
-
- % This is just a draft, nothing suitable for the final product: sources are in
- % form of a footnote
-
- % --- INTRODUCTION ---
-\chapter{Introduction}
-\label{chap:intro}
-
- Through the history of internet, many factor changed and new innovations came,
- but at the end of the day a foundamental key remained: TCP/IP performance are,
- indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP
-   performes quite well in wired connections, while in wireless one, such as
-   wireless/satellite, sucks ass} In this context network optimization plays an
- important role, and it is for this reason that network packets are processed
- while traversing the network in order to make their transmission as fast as
- possible. On top of that, mobile traffic is continuously surging every year,
- with the backbones having more and more data to process. In this context,
- in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in
- fact, is very high compared to the requirements of the network. With this
- trend, in-hardware solutions will not be able to keep up. Is with this
- considerations that software solutions are being proposed to become the new way
- to deploy network components. While less efficent, they are faster to bring
- them up, offering an unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile
-   Packet Core Network Architectures: A Survey}. Furthermore, new virtualization
- technologies such as Cloud computing and Docker are becoming more and more
- mature for being employed in large-scale production environments, where new
- frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to
- easily manage the application lifecyle, i.e. deployment, scale up/down and
- finally, removal.
-
-\section{The VIBES project}
-
- This thesis is part of the VIBES project\todo{Talk about VIBES project, add
-   some reference, explain what it is.}, where the necessity for better TCP/IP
- transmission through satellite connections is the main requirement. To reach
- this goal, the project specifications suggest to exploit the 5G incoming
- technology and use the NFV-MANO architecture to perform first packets
- elaboration and performance improvment and finally TCP/IP satellite chunk
- optimization with the Performance Enhancing Proxy (PEP). The VIBES project
- proposed five technical requirements:
-\begin{enumerate}
- \item Analysis of the applicability of current and new Internet protocols in
-   the proposed VNF-PEP architecture
- \item Implementation of a VNF-PEP prototype
- \item Building of a PoC test platform
- \item VNF-PEP validation and performance tests on 5G use-cases
- \item Demonstration testbed management
-\end{enumerate}
-
-\subsection{VNF-PEP architecture and internet protocols}
-
-The analysis of this topic revealed to be trivial: since internet has many
-different protocols that would become infeasible to support all of them at the
-same time, packet incapsulation present itself as the only faseable solution:
-every packet incoming in the VNFs has already been encapsulated by the MANO,
-\todo{Not sure about this, please give it a check - Not sure but I think that it
-  is not part of the MANO the encapsulation of the packets. Also I think that
-  the main purpose of UDP encapsulation is not to hide the protocol used on the
-  edges but 1) usa a ``quick protocol to exchange data among VNFs and 2) we are
-  hiding the path not the protocol itself. As we were discussing, we are
-  creating some sort of proxy, so the aim will be the same even if we support a
-  plethora of protocols.''} making the whole architecture indipendent from the
-protocol a particolar flux of data uses. To achieve this, TUN/TAP interfaces 
-were used to deploy a tunnel between VNFs, allowing them to pass packets 
-encapsulated in UDP ones.
-
-\subsection{VNF-PEP prototype}
-
-Since our goal shifted into creating a MANO testbed, our prototype doesn't
-include PEP. The VNF architecture was shaped following the container
-orchestrator we decided to use: Kubernetes.
-
-\subsubsection{About Kubernetes}
-
-As already introduced before, Kubernetes is an open source software framework
-specialized in container management and orchestration. Developed by Google in
-??\todo{Find out the actual date} and written in Go, it allows different
-machines (called nodes from now on) to create an abstraction layer and to form a
-cluster where is possible to run Docker containers without brothering about
-hardware constraints. Nodes can have specific roles and specialization, all of
-them managed by one or many master nodes, that actually manage the
-orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
-that it offers the possibility to change the virtualization software in use
-(i.e. from Container-based technology to Virtual machine software).
-\begin{figure}[h]
- \centering \includegraphics[scale=0.35]{kubernetes_logo}
- \caption{Kubernetes Logo}
- \label{chap:intro:img:k8s_logo}
-\end{figure}
-
-\subsubsection{About Docker}
-
-In recent years, with the new hardware capabilities and the recent development
-of in-kernel virtualization systems (such as Hyper-V) this technology begin to
-be adopted. Virtualization allows to run different systems on the same machine,
-making them completely isolated and then more resilient to failures. The back of
-the medal, though, is that virtual machines require large amount of resources,
-especially memory, because solutions like copy-and-write of in-memory pages are
-not viable anymore (since the kernel gets duplicated too), leading to
-duplication of loaded libraries and assets in the memory of the host. For large
-deployment containing only simple services (e.g. a backend application serving a
-website or a database) this ends up in a waste of resources. \todo{Talk about
-  lxc containers and how Docker solved magically the problem.} It's here that,
-in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
-on an already existing product: Linux Container (LXC). From this framework,
-Docker built an entire ecosystem, consisting in a client/server model giving the
-possibility for users to simply launch, scale and delete containers (locally or
-remotely with Docker-machine), a repository system where images, a layer-based
-``core'' where containers start running from when they are launched, can be
-stored and tagged in a sort of version control system. Finally, another couple
-of solutions, called Docker-compose, and Docker Swarm have the purpose to,
-respectively, orchestrate containers in a single or a clustered system.
-\begin{figure}[t]
- \centering \includegraphics[scale=0.7]{docker_logo}
- \caption{Official Docker logo}
- \label{chap:intro:img:docker_logo}
-\end{figure}
-
-\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
-birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
-seen as one logical unit. This layer completely makes the underlying
-infrastructure transparent: data management, as network one, are totaly managed
-by Docker. Recently, Docker-compose support has been introduced, making this
-one-node tool available for clustered systems. At the end of the day, on one
-hand Docker Swarm provides an easy way to set up a cluster system with all the
-tools configured out-of-the-box, without the necessity to set networking
-configurations or installing ad-hoc storage solutions. On the other hand,
-though, it doesn't have all the personalization options that Kuberentes offers,
-thus making the product less flexible. This foundamental characteristic,
-although, makes the two solutions have different use cases leading to different
-market needs. \todo{Add Docker Swarm}
-
-\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool
-that allows services orchestration. It automates most of the tasks that should
-be performed by hand when launching one or multiple docker containers. It's
-particulary useful when more continers have to interact together,
-because\todo{Write down docker-compose functionalities and describe the product
-  a little bit} all the services rules are described in a simple YAML file. With
-the birth of Docker Swarm, work has started to port this tool to the Swarm
-framework. At the time of this writing, the procedure to port a Docker-compose
-configuration in Docker Swarm is still not completely transparent, and it
-requires a sort of compilation, where Docker-compose bundles all the necessaries
-resources into a binary file that Docker Swarm will use to allocate the
-necessary resources.
-
-\vspace{0.5cm}
-
-Starting with the first step of creating a MANO able to process incoming data
-packets through VNF functions, we encountered that many networking tools already
-present in the market required some tweaking and some integrations, shifting our
-goal to create a complete European Telecommunications Standards Institutes
-(ETSI) Management and orchestrator (MANO) testbed instead, following the
-specifications suggested in the RFC 7665, thus implementing only the first three
-requisites, without digging in the satellite data flow optimization. In
-particular, we discovered how, these tools, were suitable to create ETSI MANO
-and VNFs using virtual machine or exploiting cloud tecnhologies, while they were
-not designed with enough flexibility to integrate with Docker. In the next
-chapter we are going to take a deep analysis of the cited technologies, in order
-to provide to the reader enough context to be able to understand our design
-choices we are going to describe later.
-
-\noindent With this in mind, we performed a requirements analysis described in
-the next chapter.\todo{This should be put at the end of the introduction.}
-
 \chapter{Project Analysis}
@@
 \paragraph{SDN?}
-
-\chapter{Architectural implementation}
-\label{chap:archimpl}
-\end{document}

### 2018-10-15 res/sections/15-ArchitecturalImplementation.tex
Move chapters in separate files, fix gitignore

--- 2018-10-11 draft.tex
@@
-\documentclass[10pt]{book}
-
-\usepackage{todonotes}
-
-\graphicspath{{res/img/}}
-
-\author{Polonio Davide} \title{A thesis draft}
-% The outline'd be something like this:
-%  Thesis outline
-% 
-% - Introduction about the VIBES project (context, 5G, the need for a new 
-% network architecture, proposed solution, requisites)
-%   + Related works (how are we gonna do that??)
-% - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
-% Docker, Docker swarm)
-%   + Choseen tecnhologies and description of the 'reason why'
-% - Project description (tell how we've done stuff)
-%   + Initial solution with Openstack and Tacker
-%   + From Tacker to Openbaton
-%   + From Openbaton with Openstack to Openbaton with Docker
-%   + Exploration of Docker orchestrators (Swam vs Kubernetes)
-%   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
-%   + Openbaton and Kubernetes integration (lack of a implemented Driver, 
-% integration with existinf VNFM)
-% - Case study
-%   + Analysis of TCP/UDP performance
-% - Results
-% - Future works
-% - Conclusions
-
-\makeindex
-\begin{document}
- \maketitle
-
- % This is just a draft, nothing suitable for the final product: sources are in
- % form of a footnote
-
- % --- INTRODUCTION ---
-\chapter{Introduction}
-\label{chap:intro}
-
- Through the history of internet, many factor changed and new innovations came,
- but at the end of the day a foundamental key remained: TCP/IP performance are,
- indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP
-   performes quite well in wired connections, while in wireless one, such as
-   wireless/satellite, sucks ass} In this context network optimization plays an
- important role, and it is for this reason that network packets are processed
- while traversing the network in order to make their transmission as fast as
- possible. On top of that, mobile traffic is continuously surging every year,
- with the backbones having more and more data to process. In this context,
- in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in
- fact, is very high compared to the requirements of the network. With this
- trend, in-hardware solutions will not be able to keep up. Is with this
- considerations that software solutions are being proposed to become the new way
- to deploy network components. While less efficent, they are faster to bring
- them up, offering an unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile
-   Packet Core Network Architectures: A Survey}. Furthermore, new virtualization
- technologies such as Cloud computing and Docker are becoming more and more
- mature for being employed in large-scale production environments, where new
- frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to
- easily manage the application lifecyle, i.e. deployment, scale up/down and
- finally, removal.
-
-\section{The VIBES project}
-
- This thesis is part of the VIBES project\todo{Talk about VIBES project, add
-   some reference, explain what it is.}, where the necessity for better TCP/IP
- transmission through satellite connections is the main requirement. To reach
- this goal, the project specifications suggest to exploit the 5G incoming
- technology and use the NFV-MANO architecture to perform first packets
- elaboration and performance improvment and finally TCP/IP satellite chunk
- optimization with the Performance Enhancing Proxy (PEP). The VIBES project
- proposed five technical requirements:
-\begin{enumerate}
- \item Analysis of the applicability of current and new Internet protocols in
-   the proposed VNF-PEP architecture
- \item Implementation of a VNF-PEP prototype
- \item Building of a PoC test platform
- \item VNF-PEP validation and performance tests on 5G use-cases
- \item Demonstration testbed management
-\end{enumerate}
-
-\subsection{VNF-PEP architecture and internet protocols}
-
-The analysis of this topic revealed to be trivial: since internet has many
-different protocols that would become infeasible to support all of them at the
-same time, packet incapsulation present itself as the only faseable solution:
-every packet incoming in the VNFs has already been encapsulated by the MANO,
-\todo{Not sure about this, please give it a check - Not sure but I think that it
-  is not part of the MANO the encapsulation of the packets. Also I think that
-  the main purpose of UDP encapsulation is not to hide the protocol used on the
-  edges but 1) usa a ``quick protocol to exchange data among VNFs and 2) we are
-  hiding the path not the protocol itself. As we were discussing, we are
-  creating some sort of proxy, so the aim will be the same even if we support a
-  plethora of protocols.''} making the whole architecture indipendent from the
-protocol a particolar flux of data uses. To achieve this, TUN/TAP interfaces 
-were used to deploy a tunnel between VNFs, allowing them to pass packets 
-encapsulated in UDP ones.
-
-\subsection{VNF-PEP prototype}
-
-Since our goal shifted into creating a MANO testbed, our prototype doesn't
-include PEP. The VNF architecture was shaped following the container
-orchestrator we decided to use: Kubernetes.
-
-\subsubsection{About Kubernetes}
-
-As already introduced before, Kubernetes is an open source software framework
-specialized in container management and orchestration. Developed by Google in
-??\todo{Find out the actual date} and written in Go, it allows different
-machines (called nodes from now on) to create an abstraction layer and to form a
-cluster where is possible to run Docker containers without brothering about
-hardware constraints. Nodes can have specific roles and specialization, all of
-them managed by one or many master nodes, that actually manage the
-orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
-that it offers the possibility to change the virtualization software in use
-(i.e. from Container-based technology to Virtual machine software).
-\begin{figure}[h]
- \centering \includegraphics[scale=0.35]{kubernetes_logo}
- \caption{Kubernetes Logo}
- \label{chap:intro:img:k8s_logo}
-\end{figure}
-
-\subsubsection{About Docker}
-
-In recent years, with the new hardware capabilities and the recent development
-of in-kernel virtualization systems (such as Hyper-V) this technology begin to
-be adopted. Virtualization allows to run different systems on the same machine,
-making them completely isolated and then more resilient to failures. The back of
-the medal, though, is that virtual machines require large amount of resources,
-especially memory, because solutions like copy-and-write of in-memory pages are
-not viable anymore (since the kernel gets duplicated too), leading to
-duplication of loaded libraries and assets in the memory of the host. For large
-deployment containing only simple services (e.g. a backend application serving a
-website or a database) this ends up in a waste of resources. \todo{Talk about
-  lxc containers and how Docker solved magically the problem.} It's here that,
-in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
-on an already existing product: Linux Container (LXC). From this framework,
-Docker built an entire ecosystem, consisting in a client/server model giving the
-possibility for users to simply launch, scale and delete containers (locally or
-remotely with Docker-machine), a repository system where images, a layer-based
-``core'' where containers start running from when they are launched, can be
-stored and tagged in a sort of version control system. Finally, another couple
-of solutions, called Docker-compose, and Docker Swarm have the purpose to,
-respectively, orchestrate containers in a single or a clustered system.
-\begin{figure}[t]
- \centering \includegraphics[scale=0.7]{docker_logo}
- \caption{Official Docker logo}
- \label{chap:intro:img:docker_logo}
-\end{figure}
-
-\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
-birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
-seen as one logical unit. This layer completely makes the underlying
-infrastructure transparent: data management, as network one, are totaly managed
-by Docker. Recently, Docker-compose support has been introduced, making this
-one-node tool available for clustered systems. At the end of the day, on one
-hand Docker Swarm provides an easy way to set up a cluster system with all the
-tools configured out-of-the-box, without the necessity to set networking
-configurations or installing ad-hoc storage solutions. On the other hand,
-though, it doesn't have all the personalization options that Kuberentes offers,
-thus making the product less flexible. This foundamental characteristic,
-although, makes the two solutions have different use cases leading to different
-market needs. \todo{Add Docker Swarm}
-
-\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool
-that allows services orchestration. It automates most of the tasks that should
-be performed by hand when launching one or multiple docker containers. It's
-particulary useful when more continers have to interact together,
-because\todo{Write down docker-compose functionalities and describe the product
-  a little bit} all the services rules are described in a simple YAML file. With
-the birth of Docker Swarm, work has started to port this tool to the Swarm
-framework. At the time of this writing, the procedure to port a Docker-compose
-configuration in Docker Swarm is still not completely transparent, and it
-requires a sort of compilation, where Docker-compose bundles all the necessaries
-resources into a binary file that Docker Swarm will use to allocate the
-necessary resources.
-
-\vspace{0.5cm}
-
-Starting with the first step of creating a MANO able to process incoming data
-packets through VNF functions, we encountered that many networking tools already
-present in the market required some tweaking and some integrations, shifting our
-goal to create a complete European Telecommunications Standards Institutes
-(ETSI) Management and orchestrator (MANO) testbed instead, following the
-specifications suggested in the RFC 7665, thus implementing only the first three
-requisites, without digging in the satellite data flow optimization. In
-particular, we discovered how, these tools, were suitable to create ETSI MANO
-and VNFs using virtual machine or exploiting cloud tecnhologies, while they were
-not designed with enough flexibility to integrate with Docker. In the next
-chapter we are going to take a deep analysis of the cited technologies, in order
-to provide to the reader enough context to be able to understand our design
-choices we are going to describe later.
-
-\noindent With this in mind, we performed a requirements analysis described in
-the next chapter.\todo{This should be put at the end of the introduction.}
-
-\chapter{Project Analysis}
-\label{chap:prjan}
-
-As in chapter~\ref{chap:intro}, the creation of a full ETSI MANO compliant
-architecture, following the suggestions described in RFC 7665, require the
-integration of multiple tools (some of them have been already briefly introduced
-in~\ref{chap:intro}). The tools were choosen after their analysis, where pro and
-cons have been studied and compared to the necessity of this thesis. Since the
-project has been developed with the cloud provided by the University of Padua,
-Openstack (which a product analysis can be found later
-in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
-
-\section{Architecture overview}
-
-During the testbed development the software architecture diverged from the
-original one, experiencing different evolutions and modifications.
-
-\subsection{VIBES architectural description}
-
-In the original VIBES architectural design, based on leveraging virtualization
-technologies, is possible to identify different macro-components, each of these
-specialized in fullfilling a specific goal. For the ETSI standards, the Network
-Function Virtualization (NFV) design should be composed as follows:
-\begin{figure}[t]
- \centering \includegraphics[scale=2]{etsi_arch}
- \caption{A schematic image representing the high level NFV architecture
-   proposed by ESA.}
- \label{chap:prjan:img:etsi_arch}
-\end{figure}
-
-\begin{itemize}
- \item \textbf{Network Function Virtualization Infrastructure (NFVI)}: it's the
-   heart of the whole computation component, and here three sub-domains can be
-   identified: \todo{Expand this section}
-\begin{itemize} 
- \item \textbf{Hypervisor domain} that is composed of the Virtual Compute, the
-   Virtual Storage, and the Virtualization Layer. These components are used to
-   virtualize the computational and storage resources, providing a layer for
-   accessing it.
- \item \textbf{Compute domain} composed by the Physical Compute and the Physical
-   Storage. These are the real resources which are virtualized via software.
- \item \textbf{Infrastructure domain} which has the Physical Network, and its
-   virtualized counterpart as made available by the Virtualization Layer.
-\end{itemize}
-
-In conventional virtualization systems these resources are usually separated
-from the host operating system, providing better isolation but wrost
-performance. The ETSI infrastructure, instead, tries with container
-virtualization technology to push for a better NFV responsiveness\todo{Check
-  term - I don't think it actually exists}, removing the additional operating
-system (called ``guest'' OS) required by the usual hypervisor-oriented approach.
-
-To realize the required components, specific tools were suggested in the VNF
-technical proposal: regarding the NFV (network infrastructure) domain
-Docker-Compose was proposed as a viable solution, while to store PEP instances
-docker swarm was described as a possible candidate.
-
-We performed an accured analysis of the tools choosen based on their maturity,
-on the community and on the technical support offered (e.g. user manuals,
-developer documentation), which, at the end, commited us to choose different
-tools from the suggested ones. A detailed description of the architectural
-implementation can be found in chapter~\ref{chap:archimpl}.\todo{Include prof.
-  survey about community components here}
-
-\item \textbf{VNFs} this are the core components that perform packets
-  manipulation. This components are design to be lightweight services able to
-  elaborate a great amount of packets per seconds. Manupulation can be of two
-  typologies: stateful or stateless. In the first case, a state is maintained,
-  and the VNF can perform heuristics about data coming through it (especially
-  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
-  case, no state is maintained, and every packet is treated without any
-  particular assumption. Following the RFC 7665, the VNF are also called Service
-  Function (SF)\footnote{From now on the two terminologies will be use in a
-    interchangeable way.}. \todo{Add ref to chap about SFC?}
-
-\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
-  Orchestrator (NFVO) is responsible for the end-to-end management and
-  orchestration of network services (NS), also called SFC. The MANO has
-  different responsabilities, such as managing the
-  istantiation/destruction/modification of the SFCs, that achieves interacting
-  with the VIM.\todo{CITE: SDN/NFV-Based Mobile Packet Core Network
-Architectures: A Survey}. Finally, the mano can integrate with 
-Operational and Business Support Systems (OSS/BSS), \todo{Expand this part with 
-the following papers: https://www.tandfonline.com/doi/abs/10.1057/jors.1995.176 
-https://ieeexplore.ieee.org/abstract/document/6963800 
-https://www.computer.org/csdl/proceedings/ifcsta/2009/3930/01/3930a466-abs.html}
-
-\end{itemize}
-
-\section{SFC}
-
-\todo{Here it's missing an introduction about SFC and its terminology, need to 
-include parts of RFC 7665}
-
-\subsection{Packet transmission strategies}
-
-When a packets are sent from a seder $S$, to a receiver $R$, they start
-travelling to the backbone receiving appropriate in-hardware packet processing
-before being received by $R$. During this elaboration, it is important to
-maintain the transparency of the connection, so that the network users are not
-aware of the packets elaboration at all. To perform that, many solutions are
-available: TUN/TAP packet incapsulation, IP Encapsulation within IP or TCP
-session recreation (know also as TCP split).
-
-TUN/TAP and IP Encapsulation within IP (initially defined in RFC 2003 with the
-aim to deliver IP packets over Mobile IP for mobile networks) maintain the
-connection transparency incapsulating the packet in another one.
-\begin{figure}[t]
-  \centering \includegraphics[scale=0.5]{IPoverIP}
-  \caption[IP Encapsulation within IP packet schema]{A schema of IP
-    Encapsulation within IP. Is noticeable how the original packet becomes the
-    payload of a new one, thus being encapsulated. All the informations
-    regarding the original packet remains untouched: when this kind of packet
-    gets send from a machine to another, the OS will remove the new packet
-    header and will present to the user-space program receiving the data its
-    payload: a IP packet. Packets encapsulation can be applied with higher
-    layers too, thus allowing for TCP and UDP packet incapsulation.}
-  \label{chap:prjan:img:ip_over_ip}
-\end{figure}
-
-In this way, the original packet remains preserved and it becomes the new packet
-payload, that can be elaborated and modified accordingly, until it gets
-decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is not
-be aware of the packet elaboration, since all the original headers were
-preserved (or slightly modified). In particular, IP Encapsulation within IP
-allow the packets to make intermediate destinations that otherwise would not be
-selected. TCP session recreation, instead, ``split'' the TCP session in two
-endpoints: an Ingress, that manage the connection with the client and an Egress,
-that communicate with the packet receiver creating a multi-overlay-hop path
-where for each hop there is an possible indipendent TCP connection. In order to
-achieve that, a session table needs to be established and continuously updated,
-while TCP headers need to be accordingly modified to allow data transmission
-through the Ingress and Egress points. Nonetheless, care should be given to
-packets path: in order to avoid ill-calculated congestion windows, the incoming
-packets need to follow the same (or an equivalent one) path for the same
-instatiated connection. This phenomenom is based on the intermediate proxies
-that create the route: it is the case, for example, when an intermediate node
-sends back a spoofed ACK before the original packet is delivered to the real
-consegnee. In this senario, the sender, unaware of the proxy, will miscalculate
-the sending window based on the too low RTT. Other problems of TCP splitting are
-reliability and security, especially because the end to end TCP logic is no
-longer maintained: a server failure may cause to an unaware client to believe
-that all the packets have reached the destination. TCP session recreation seems
-burdersome and it doesn't seem to add any significant benefit at the first
-sight, but it offers the flexibility to perform extensive packet manipulation:
-since the packet its recreated every time at the Ingress and Egress points,
-while in the Network Function Virtualization Infrastructure (NFVI), the original
-packet can be stripped of its headers (that could be saved in the session table)
-to slightly improve internal packet transmission. \todo{Insert TCP session
-  recreation schema} On top of that, it offers the possibility to use different
-TCP flavours, hence increasing the communication speed in heterogeneous networks
-(e.g. in case of satellite connections TCP Hybla can be employed, while it is
-known that HighSpeed TCP gives best results when applied in optical fiber-based
-connections).
-
-During the testbed development we first tried the TUN/TAP solution, that
-revealed to be not completely transparent and with major performance drawbacks
-(i.e. packet transmission in a LAN connection reached latencies of 50ms)
-\todo{Explain the constraint of point-to-point connections, while we need
-  dynamic SFC)}, so we choose to opt for a TCP session recreation solution,
-implementing a little back-end with the aim of saving in a persistent storage
-the SFC the packet has to perform and additional metadata.
-
-\section{Available technologies in the market}
-
-Before starting any effective work on the project it has been decided to 
-perform an analysis of the possible technologies avvailable in the market, 
-performing a choice between them and exploiting their possibile ``weak 
-points''. Proceding with a top-down approach, the frameworks were choosen from 
-the one that had the biggest role in the testbed to conclude with the 
-less-important components.
-
-\subsection{NFVI}
-
-A critical component, as the MANO one, is the NFVI. Without it, network 
-resources can't be allocated, and since our project was focused on dynamically 
-allocating containerized applications on cloud platforms, we decided to start 
-analizing the right framework for this component first.
-
-\subsection{Openstack}
-\label{chap:prjan:sec:openstack}
-Created in ??\todo{Find out Openstack date of creation}, Openstack allows
-on-permise cloud installations, using bare-metal resources to provide common
-cloud services as object storage, virtual machine deploy, virtual networking.
-Furthermore its modularity offers the possibility to add additional components,
-even proprietary, to achieve a complete cloud solution.
-\begin{figure}[t]
- \centering \includegraphics[scale=0.58]{openstack_logo}
- \caption{Openstack logo}
- \label{chap:prjan:img:openstack_logo}
-\end{figure}
-
-\todo{This section needs to be expanded a lot!!! In particular, we should put
-  some focus about VNF and MANO, introducing them}
-
-\subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
-\paragraph{Openvswitch}
-\paragraph{Openbaton} Patrocined by the Institute Fraunhofer, it's a 
-open-source, customizable NFV MANO-compliant framework, that offers many
-configurations options, and since it's written in Java, there is the possibility
-to add components (plug-ins) dynamically. Openbaton can be installed as a normal
-computer program (they offer packages for the most common distros) but it can be
-deployed on Docker containers too. It's designed to work with cloud providers
-like Amazon or Google Compute engine, but it offers compability with on-permise
-cloud solutions like Openstack, where it deploys virtual machines using the API
-the platform offers. It can store VNF configurations saved as TOSCA
-YAML\todo{What does TOSCA mean? Find out} and it can handle multiple
-PoPs\todo{Write down PoP meaning}. It offers a VNF lifecyle management
-out-of-the-box, with the possibility to customize it based on the user needs.
-Its modular design allows developers to change parts of the codebase with custom
-ones, making the product really flexible to different platforms. Performing this
-operation requires a deep knowlege of the framework though.
-\begin{figure}[h]
- \centering \includegraphics[scale=0.45]{openbaton_logo}
- \caption{Openbaton logo}
- \label{chap:prjan:img:openbaton_logo}
-\end{figure}
-
-\paragraph{SDN?}
-
 \chapter{Architectural implementation}
 \label{chap:archimpl}
-\end{document}

### 2018-10-16 res/sections/07-Introduction.tex
Add content to intro and created background empty file

@@
    performes quite well in wired connections, while in wireless one, such as
-   wireless/satellite, sucks ass} In this context network optimization plays an
- important role, and it is for this reason that network packets are elaborated,
- compressed and processed while traversing the network in order to make their
- transmission as fast as possible. On top of that, mobile traffic is
- continuously surging every year, with the backbones having more and more data
- to deliver, making in-hardware backbone solutions less efficent. Their
- time-to-deploy, in fact, is very high compared to the requirements of the
- network. With this trend, in-hardware solutions will not be able to keep up. Is
- with this considerations that software a software approach is being proposed to
- become the new way to deploy network components. While less efficent, they are
- faster to bring them up, offering an unprecedented flexibility\todo{CITE:
-   SDN/NFV-Based Mobile Packet Core Network Architectures: A Survey}.
+   wireless/satellite, sucks ass}
+With the arrival of mobile connectivity, wireless networks started to have an 
+important role in everyone's life. In this context network efficiency plays 
+an important role with the development of new connectivity standards (e.g 2G, 
+3G, LTE) and the creation of common communication protocols ad-hoc versions 
+(i.e. TCP versions specialized for exploit the characteristic of wireless 
+networks). It is for this reason that, beside improved radio communications, 
+backend infrastructure began to be refined too. During their travel to the 
+destination packets are elaborated, compressed and processed in order to make 
+their transmission as fast as possible.
+On top of that, mobile traffic is continuously surging every year, with the 
+backbones having more and more data to deliver, making in-hardware backbone 
+solutions less efficent. Their time-to-deploy, in fact, is very high compared 
+to the requirements of the network. With this trend, in-hardware solutions 
+will not be able to keep up. Is with this considerations that software a 
+software approach is being proposed to become the new way to deploy network 
+components. While less efficent, they are faster to bring them up, offering an 
+unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network 
+Architectures: A Survey}.
  Furthermore, new virtualization technologies such as Cloud computing and Docker

### 2018-10-17 res/sections/07-Introduction.tex
Move 5G introduction in a 'Background' chapter

@@
 
-\section{The upcoming connectivity standard: 5G}
-The continuous innovation in the mobile network connectivity is leading to the
-creation of a new standard, the 5G, which is estimated to arrive to the market
-constumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
-alliace, a group composed by the major players in the field of mobile
-connectivity, the 5G aims to offer at the end-user a new way to browse the web,
-download and watch interactive content. Relatively to 4G, 5G points to have data
-rates $10$ times better, with $10$ times smaller end-to-end latency and an
-increased connection densisty by $100$ times\todo{CITE}.
+ \section{Document organization}
 
-\begin{table}[t]
-\centering
-\resizebox{\textwidth}{!}{%
-  \begin{tabular}{p{4cm}|p{5,5cm}|p{5,5cm}}
-\textbf{Attribute}                                                    & \textbf{LTE capability}                                                                                                                                     & \textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
-\textbf{Data rate (per user)}                                         & Up to 100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).
-& 10X expected on average and peak rates and 100X expected on cell edge \\
-\textbf{End-to-end latency}                                           & 10 ms for two-way RAN (pre-scheduled).
-Typically, up to 50 ms end-to-end if other factors are considered (e.g., transmission, CN, internet, proxy servers).
-& 10X (smaller)                                                         \\
-\textbf{Mobility}                                                     & Functional up to 350 km/h (for certain bands up to 500 km/h).
-No support for civil aviation.
-& 1.5X                                                                  \\
-\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & Typically $\sim$2,000 active users/km2.
-& 100X
-\end{tabular}%
-}
-\caption[5G improvement over LTE]{An extract from the official 5G white paper illustating the improvements of 5G relatively to LTE connections.}
-\label{chap:intro:table:ltevs5g}
-\end{table}
-
-Part of the new requirements can be satisfied using a large radio spectrum with
-higher frequencies. The utilization of higher frequencies, though, mean that the
-radio signals can be easly disrupted by physical objects, like buildings and
-many geographical elements (such as hills and mountains), clashing with the
-expectation of an ever-reachable connectivity. It is here where virtualization
-plays an important role. In fact, the re-design of some network components today
-existing via hardware can transform a monolitic networking approach to a modular
-one, exploting the flexibility that Virtual Network Functions (VNFs)\todo{CITE
-  VNF-P: A Model for Efficient Placement of Virtualized Network Functions} can
-offer thanks to virtualization, closing the gap to the use-case fullfilment
-defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture}.
-
-\subsection{5G architecture}
-
-With the constraints placed on the requirements formulated by the NGMN alliance,
-5G envisage a multi-layered architecture, based on three main layers:\todo{CITE:
-  NGMN View on 5G architecture}
+ This thesis is organized as follows:
  \begin{itemize}
-\item \textbf{infrastructure resouce layer}: physical resources that are exposed
-  via a virtualized interface, and that can be monitored using specific APIs
-\item \textbf{business enablement layer}: where a library of functions and
-  deployment is contained, and its configuration is accessible via APIs
-\item \textbf{business application layer}: layer that contains specific
-  applications and services of the operator
+  \item Chapter~\ref{chap:background} provides background informations about 5G 
+and the technologies used to build an ETSI compliant architecture
  \end{itemize}
-
-\todo{Add three-layer image here?}
-
-\subsubsection{Network slicing}
-The role of a ``network slice'' in a 5G architecture is to specificatly handle a
-Control-place of a praticular service (e.g. smartphones traffic, autonomous
-driving, massive IoT), deploying resources in a manner that assure the required
-latency, security and reliability. While some legacy of very peculiar services
-could require specific hardware, the common resources between services could be
-shared in a virtualized way, providing auto-scaling capabilities in services
-that are under heavy network pressure.
-
-\section{The VIBES project}
-
- This thesis is part of the VIBES project\todo{Talk about VIBES project, add
-   some reference, explain what it is.}, where the necessity for better TCP/IP
- transmission through satellite connections is the main requirement. To reach
- this goal, the project specifications suggest to exploit the 5G incoming
- technology and use the NFV-MANO architecture to perform first packets
- elaboration and performance improvment and finally TCP/IP satellite chunk
- optimization with the Performance Enhancing Proxy (PEP). The VIBES project
- proposed five technical requirements:
-\begin{enumerate}
- \item Analysis of the applicability of current and new Internet protocols in
-   the proposed VNF-PEP architecture
- \item Implementation of a VNF-PEP prototype
- \item Building of a PoC test platform
- \item VNF-PEP validation and performance tests on 5G use-cases
- \item Demonstration testbed management
-\end{enumerate}
-
-\subsection{VNF-PEP architecture and internet protocols}
-
-The analysis of this topic revealed to be trivial: since internet has many
-different protocols that would become infeasible to support all of them at the
-same time, packet incapsulation present itself as the only faseable solution:
-every packet incoming in the VNFs has already been encapsulated by the MANO,
-\todo{Not sure about this, please give it a check - Not sure but I think that it
-  is not part of the MANO the encapsulation of the packets. Also I think that
-  the main purpose of UDP encapsulation is not to hide the protocol used on the
-  edges but 1) usa a ``quick protocol to exchange data among VNFs and 2) we are
-  hiding the path not the protocol itself. As we were discussing, we are
-  creating some sort of proxy, so the aim will be the same even if we support a
-  plethora of protocols.''} making the whole architecture indipendent from the
-protocol a particolar flux of data uses. To achieve this, TUN/TAP interfaces 
-were used to deploy a tunnel between VNFs, allowing them to pass packets 
-encapsulated in UDP ones.
-
-\subsection{VNF-PEP prototype}
-
-Since our goal shifted into creating a MANO testbed, our prototype doesn't
-include PEP. The VNF architecture was shaped following the container
-orchestrator we decided to use: Kubernetes.
-
-\subsubsection{About Kubernetes}
-
-As already introduced before, Kubernetes is an open source software framework
-specialized in container management and orchestration. Developed by Google in
-??\todo{Find out the actual date} and written in Go, it allows different
-machines (called nodes from now on) to create an abstraction layer and to form a
-cluster where is possible to run Docker containers without brothering about
-hardware constraints. Nodes can have specific roles and specialization, all of
-them managed by one or many master nodes, that actually manage the
-orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
-that it offers the possibility to change the virtualization software in use
-(i.e. from Container-based technology to Virtual machine software).
-\begin{figure}[h]
- \centering \includegraphics[scale=0.35]{kubernetes_logo}
- \caption{Kubernetes Logo}
- \label{chap:intro:img:k8s_logo}
-\end{figure}
-
-\subsubsection{About Docker}
-
-In recent years, with the new hardware capabilities and the recent development
-of in-kernel virtualization systems (such as Hyper-V) this technology begin to
-be adopted. Virtualization allows to run different systems on the same machine,
-making them completely isolated and then more resilient to failures. The back of
-the medal, though, is that virtual machines require large amount of resources,
-especially memory, because solutions like copy-and-write of in-memory pages are
-not viable anymore (since the kernel gets duplicated too), leading to
-duplication of loaded libraries and assets in the memory of the host. For large
-deployment containing only simple services (e.g. a backend application serving a
-website or a database) this ends up in a waste of resources. \todo{Talk about
-  lxc containers and how Docker solved magically the problem.} It's here that,
-in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
-on an already existing product: Linux Container (LXC). From this framework,
-Docker built an entire ecosystem, consisting in a client/server model giving the
-possibility for users to simply launch, scale and delete containers (locally or
-remotely with Docker-machine), a repository system where images, a layer-based
-``core'' where containers start running from when they are launched, can be
-stored and tagged in a sort of version control system. Finally, another couple
-of solutions, called Docker-compose, and Docker Swarm have the purpose to,
-respectively, orchestrate containers in a single or a clustered system.
-\begin{figure}[t]
- \centering \includegraphics[scale=0.7]{docker_logo}
- \caption{Official Docker logo}
- \label{chap:intro:img:docker_logo}
-\end{figure}
-
-\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
-birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
-seen as one logical unit. This layer completely makes the underlying
-infrastructure transparent: data management, as network one, are totaly managed
-by Docker. Recently, Docker-compose support has been introduced, making this
-one-node tool available for clustered systems. At the end of the day, on one
-hand Docker Swarm provides an easy way to set up a cluster system with all the
-tools configured out-of-the-box, without the necessity to set networking
-configurations or installing ad-hoc storage solutions. On the other hand,
-though, it doesn't have all the personalization options that Kuberentes offers,
-thus making the product less flexible. This foundamental characteristic,
-although, makes the two solutions have different use cases leading to different
-market needs. \todo{Add Docker Swarm}
-
-\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool
-that allows services orchestration. It automates most of the tasks that should
-be performed by hand when launching one or multiple docker containers. It's
-particulary useful when more continers have to interact together,
-because\todo{Write down docker-compose functionalities and describe the product
-  a little bit} all the services rules are described in a simple YAML file. With
-the birth of Docker Swarm, work has started to port this tool to the Swarm
-framework. At the time of this writing, the procedure to port a Docker-compose
-configuration in Docker Swarm is still not completely transparent, and it
-requires a sort of compilation, where Docker-compose bundles all the necessaries
-resources into a binary file that Docker Swarm will use to allocate the
-necessary resources.
-
-\vspace{0.5cm}
-
-Starting with the first step of creating a MANO able to process incoming data
-packets through VNF functions, we encountered that many networking tools already
-present in the market required some tweaking and some integrations, shifting our
-goal to create a complete European Telecommunications Standards Institutes
-(ETSI) Management and orchestrator (MANO) testbed instead, following the
-specifications suggested in the RFC 7665, thus implementing only the first three
-requisites, without digging in the satellite data flow optimization. In
-particular, we discovered how, these tools, were suitable to create ETSI MANO
-and VNFs using virtual machine or exploiting cloud tecnhologies, while they were
-not designed with enough flexibility to integrate with Docker. In the next
-chapter we are going to take a deep analysis of the cited technologies, in order
-to provide to the reader enough context to be able to understand our design
-choices we are going to describe later.
-
-\noindent With this in mind, we performed a requirements analysis described in
-the next chapter.\todo{This should be put at the end of the introduction.}

### 2018-10-17 res/sections/08-Background.tex
Move 5G introduction in a 'Background' chapter

--- 2018-10-14 draft.tex
@@
-\documentclass[10pt]{book}
+\chapter{Background}
+\label{chap:background}
 
-\usepackage{todonotes}
-\usepackage{graphicx}
-
-\graphicspath{{res/img/}}
-
-\author{Polonio Davide} \title{A thesis draft}
-% The outline'd be something like this:
-%  Thesis outline
-% 
-% - Introduction about the VIBES project (context, 5G, the need for a new 
-% network architecture, proposed solution, requisites)
-%   + Related works (how are we gonna do that??)
-% - Analysis of the available technologies (kubernetes, openstack, openvswitch, 
-% Docker, Docker swarm)
-%   + Choseen tecnhologies and description of the 'reason why'
-% - Project description (tell how we've done stuff)
-%   + Initial solution with Openstack and Tacker
-%   + From Tacker to Openbaton
-%   + From Openbaton with Openstack to Openbaton with Docker
-%   + Exploration of Docker orchestrators (Swam vs Kubernetes)
-%   + Kubernetes internal composition (GlusterFS, Ingress, Harbor, metrics)
-%   + Openbaton and Kubernetes integration (lack of a implemented Driver, 
-% integration with existinf VNFM)
-% - Case study
-%   + Analysis of TCP/UDP performance
-% - Results
-% - Future works
-% - Conclusions
-
-\makeindex
-\begin{document}
- \maketitle
-
- % This is just a draft, nothing suitable for the final product: sources are in
- % form of a todo
-
- % --- INTRODUCTION ---
-\chapter{Introduction}
-\label{chap:intro}
-
- Through the history of internet, many factors changed and new innovations came,
- but at the end of the day a foundamental key remained: connection performances
- are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
-   performes quite well in wired connections, while in wireless one, such as
-   wireless/satellite, sucks ass} In this context network optimization plays an
- important role, and it is for this reason that network packets are elaborated,
- compressed and processed while traversing the network in order to make their
- transmission as fast as possible. On top of that, mobile traffic is
- continuously surging every year, with the backbones having more and more data
- to deliver, making in-hardware backbone solutions less efficent. Their
- time-to-deploy, in fact, is very high compared to the requirements of the
- network. With this trend, in-hardware solutions will not be able to keep up. Is
- with this considerations that software a software approach is being proposed to
- become the new way to deploy network components. While less efficent, they are
- faster to bring them up, offering an unprecedented flexibility\todo{CITE:
-   SDN/NFV-Based Mobile Packet Core Network Architectures: A Survey}.
- Furthermore, new virtualization technologies such as Cloud computing and Docker
- are becoming more and more mature for being employed in large-scale production
- environments, where new frameworks, called orchestrators (e.g. Kubernetes,
- Docker Swarm), allow to easily manage the application lifecyle, i.e.
- deployment, scale up/down and finally, removal.
+In this chapter we are going to explain the most important concepts of the 
+upcoming 5G, talking about the Virtual Network Functions (VNFs) and how first 
+experiments with this new technology already begun. Finally we are going to 
+briefly introduce and explain the role of the technologies used in this 
+Thesis.
 
@@
   \begin{tabular}{p{4cm}|p{5,5cm}|p{5,5cm}}
-\textbf{Attribute}                                                    & \textbf{LTE capability}                                                                                                                                     & \textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
-\textbf{Data rate (per user)}                                         & Up to 100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).
-& 10X expected on average and peak rates and 100X expected on cell edge \\
-\textbf{End-to-end latency}                                           & 10 ms for two-way RAN (pre-scheduled).
-Typically, up to 50 ms end-to-end if other factors are considered (e.g., transmission, CN, internet, proxy servers).
-& 10X (smaller)                                                         \\
-\textbf{Mobility}                                                     & Functional up to 350 km/h (for certain bands up to 500 km/h).
-No support for civil aviation.
-& 1.5X                                                                  \\
-\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & Typically $\sim$2,000 active users/km2.
-& 100X
+\textbf{Attribute}                                                    & 
+\textbf{LTE capability}                                                          
+                                                                           & 
+\textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
+\textbf{Data rate (per user)}                                         & Up to 
+100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).                               
+                                                                     & 10X 
+expected on average and peak rates and 100X expected on cell edge \\
+\textbf{End-to-end latency}                                           & 10 ms 
+for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other 
+factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X 
+(smaller)                                                         \\
+\textbf{Mobility}                                                     & 
+Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
+civil aviation.                                                                & 
+1.5X                                                                  \\
+\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & 
+Typically $\sim$2,000 active users/km2.                                          
+                                                                           & 
+100X                                                                 
 \end{tabular}%
 }
-\caption[5G improvement over LTE]{An extract from the official 5G white paper illustating the improvements of 5G relatively to LTE connections.}
+\caption[5G improvement over LTE]{An extract from the official 5G white paper 
+illustating the improvements of 5G relatively to LTE connections.}
 \label{chap:intro:table:ltevs5g}

### 2018-10-17 res/sections/07-Introduction.tex
Add document organization section

@@
  \begin{itemize}
-  \item Chapter~\ref{chap:background} provides background informations about 5G 
+  \item Chapter~\ref{chap:background} describes background information about 5G 
 and the technologies used to build an ETSI compliant architecture
+  \item Chapter~\ref{chap:prjan} explains the analysis performed on the 
+requirements provided and the design choices performed.
+  \item Chapter~\ref{chap:archimpl} provides an in-depth view of the 
+high-level architectural implementation.
+  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.
  \end{itemize}

### 2018-10-17 res/sections/07-Introduction.tex
Perform some proof-reading

@@
 an important role with the development of new connectivity standards (e.g 2G, 
-3G, LTE) and the creation of common communication protocols ad-hoc versions 
-(i.e. TCP versions specialized for exploit the characteristic of wireless 
-networks). It is for this reason that, beside improved radio communications, 
-backend infrastructure began to be refined too. During their travel to the 
-destination packets are elaborated, compressed and processed in order to make 
-their transmission as fast as possible.
+3G, LTE) and the creation of common, ad-hoc versions, transmission protocols 
+ad-hoc versions (i.e. TCP versions specialized for exploit the characteristic of 
+wireless networks). It is for this reason that, beside improved radio 
+communications, backend infrastructure began to be refined too. During their 
+communication to the destination packets are elaborated, compressed and 
+processed in order to make their transmission as fast as possible.
 On top of that, mobile traffic is continuously surging every year, with the 

### 2018-10-18 res/sections/07-Introduction.tex
Add vibes logo, proof reading, add content

@@
 On top of that, mobile traffic is continuously surging every year, with the 
-backbones having more and more data to deliver, making in-hardware backbone 
-solutions less efficent. Their time-to-deploy, in fact, is very high compared 
+backbones having more and more data to deliver, making in-hardware solutions 
+less efficent. Their time-to-deploy, in fact, is very high compared 
 to the requirements of the network. With this trend, in-hardware solutions 
-will not be able to keep up. Is with this considerations that software a 
-software approach is being proposed to become the new way to deploy network 
-components. While less efficent, they are faster to bring them up, offering an 
-unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network 
-Architectures: A Survey}.
+will not be able to keep up the Internet needs. Is with this considerations 
+that software-based approach, based on virtualization technologies, is being 
+proposed to become the new way to deploy network components. While less 
+efficent, they are faster to bring them up, offering an unprecedented 
+flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network Architectures: A 
+Survey}.
  Furthermore, new virtualization technologies such as Cloud computing and Docker
@@
  environments, where new frameworks, called orchestrators (e.g. Kubernetes,
- Docker Swarm), allow to easily manage the application lifecyle, i.e.
+ Docker Swarm), allow to easily manage the application lifecyle, i.e. its
  deployment, scale up/down and finally, removal.
@@
 high-level architectural implementation.
-  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.
+  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience 
+and takes into account final considerations.
  \end{itemize}

### 2018-10-18 res/sections/08-Background.tex
Change column width in table
Add vibes logo, proof reading, add content

@@
 In this chapter we are going to explain the most important concepts of the 
-upcoming 5G, talking about the Virtual Network Functions (VNFs) and how first 
-experiments with this new technology already begun. Finally we are going to 
-briefly introduce and explain the role of the technologies used in this 
-Thesis.
+upcoming 5G standard, talking about the Virtual Network Functions (VNFs) and 
+how first experiments with this new technology already begun. Finally we are 
+going to briefly introduce and explain the technologies involved and their role.
 
@@
 alliace, a group composed by the major players in the field of mobile
-connectivity, the 5G aims to offer at the end-user a new way to browse the web,
-download and watch interactive content. Relatively to 4G, 5G points to have data
-rates $10$ times better, with $10$ times smaller end-to-end latency and an
-increased connection densisty by $100$ times\todo{CITE}.
+connectivity, the 5G aims to offer not only at the end-user a new way to browse 
+the web,
+download and watch interactive content but also to create an ad-hoc solution 
+for Machine-to-Machine (M2M) data traffic, which is increasing more than ever 
+thanks to the spreading of IoT devices, which sensors need to continuosly send 
+data to servers/datacenters. Relatively to LTE\todo{Explain acronym}, 5G points 
+to have data rates $10$ times better, with $10$ times smaller end-to-end 
+latency and an increased connection densisty by $100$ times\todo{CITE}.
 
@@
 \resizebox{\textwidth}{!}{%
-  \begin{tabular}{p{4cm}|p{5,5cm}|p{5,5cm}}
+  \begin{tabular}{p{4,5cm}|p{5,5cm}|p{5cm}}
 \textbf{Attribute}                                                    & 
@@
 offer thanks to virtualization, closing the gap to the use-case fullfilment
-defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture}.
+defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture} which 
+require a descreased time to set up and deploy network services (specificatly, 
+from 90 hours to 90 minutes\todo{Search this requirement, check 90h 'cause 
+I'm not sure about it}).
 
@@
 driving, massive IoT), deploying resources in a manner that assure the required
-latency, security and reliability. While some legacy of very peculiar services
+latency, security and reliability. While some very peculiar legacy services
 could require specific hardware, the common resources between services could be
@@
 \end{enumerate}
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=1]{vibes_logo}
+ \caption{ViBeS project logo}
+ \label{chap:background:img:vibes_logo}
+\end{figure}
 

### 2018-10-19 res/sections/08-Background.tex
Add new file, add content in background section

@@
 
+This separation in layers allows to easly manage different identities 
+differently: an Internet Service Provider (ISP) could manage different physical
+infrastructure and have multiple business emblements dislocated 
+along an entire continet for example, but it could decide to have only a single 
+centralized business application deployment that manage all the other layers 
+resources.
+
 \subsubsection{Network slicing}
-The role of a ``network slice'' in a 5G architecture is to specificatly handle a
-Control-place of a praticular service (e.g. smartphones traffic, autonomous
-driving, massive IoT), deploying resources in a manner that assure the required
-latency, security and reliability. While some very peculiar legacy services
-could require specific hardware, the common resources between services could be
-shared in a virtualized way, providing auto-scaling capabilities in services
-that are under heavy network pressure.
+The role of a ``network slice'' in a 5G architecture is to specificatly handle 
+the Control-plane\todo{Explain what a control plane is?} of a praticular 
+service (e.g. smartphones traffic, autonomous driving, massive IoT), deploying 
+resources in a manner that assure the required latency, security and 
+reliability. While some very peculiar legacy services could require specific 
+hardware, the common resources between services could be shared in a 
+virtualized way, providing auto-scaling capabilities in services that are under 
+heavy network pressure.
 
@@
 same time, packet incapsulation present itself as the only faseable solution:
-every packet incoming in the VNFs has already been encapsulated by the MANO,
-\todo{Not sure about this, please give it a check - Not sure but I think that it
-  is not part of the MANO the encapsulation of the packets. Also I think that
-  the main purpose of UDP encapsulation is not to hide the protocol used on the
-  edges but 1) usa a ``quick protocol to exchange data among VNFs and 2) we are
-  hiding the path not the protocol itself. As we were discussing, we are
-  creating some sort of proxy, so the aim will be the same even if we support a
-  plethora of protocols.''} making the whole architecture indipendent from the
-protocol a particolar flux of data uses. To achieve this, TUN/TAP interfaces 
-were used to deploy a tunnel between VNFs, allowing them to pass packets 
-encapsulated in UDP ones.
+every packet incoming in the VNFs has already been encapsulated by a generic
+packet incapsulator/decapsulator,
+\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
+the protocol used on the edges but 1) usa a ``quick protocol to exchange data 
+among VNFs and 2) we are hiding the path not the protocol itself. As we were 
+discussing, we are creating some sort of proxy, so the aim will be the same 
+even if we support a plethora of protocols.''} making the whole architecture 
+indipendent from the protocol a particolar flux of data. To achieve this, 
+several solutions have been studied, and at the end packet incapsulation with 
+TCP split (for TCP sessions) have been choosen. The rationale that guided us on 
+this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
+reference with the section of the explanation}
 
@@
 
-Since our goal shifted into creating a MANO testbed, our prototype doesn't
-include PEP. The VNF architecture was shaped following the container
-orchestrator we decided to use: Kubernetes.
+Since the requirement for the whole system (MANO+PEP) were too challenging, the 
+goal shifted into creating a MANO testbed and a working NFVI, excluding PEP. 
+The VNF architecture was shaped following the container orchestrator we decided 
+to use. An more detailed architectural implementation can be found at 
+Chapter~\ref{chap:archimpl}.
+
+\section{Technologies used}
+
+The development of the ETSI MANO testbed involved the use of several 
+tecnhologies already in the market, and in some cases the creation of ad-hoc 
+solutions to solve particular issues. Here, a brief listing of the 
+well-estabilished technologies used is performed, describing even solutions 
+that at the end weren't choosen because they didn't fit with the defined goals.
 

### 2018-10-19 res/sections/20-VNFImplementaton.tex
Add new file, add content in background section

\chapter{VNF and Network Service implementation} 
\label{chap:vnf_ns_impl}

### 2018-10-24 res/sections/08-Background.tex
Move technologies, add harbor description, minor fixes

@@
 
-\section{Technologies used}
-
-The development of the ETSI MANO testbed involved the use of several 
-tecnhologies already in the market, and in some cases the creation of ad-hoc 
-solutions to solve particular issues. Here, a brief listing of the 
-well-estabilished technologies used is performed, describing even solutions 
-that at the end weren't choosen because they didn't fit with the defined goals.
-
-\subsubsection{About Kubernetes}
-
-As already introduced before, Kubernetes is an open source software framework
-specialized in container management and orchestration. Developed by Google in
-??\todo{Find out the actual date} and written in Go, it allows different
-machines (called nodes from now on) to create an abstraction layer and to form a
-cluster where is possible to run Docker containers without brothering about
-hardware constraints. Nodes can have specific roles and specialization, all of
-them managed by one or many master nodes, that actually manage the
-orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
-that it offers the possibility to change the virtualization software in use
-(i.e. from Container-based technology to Virtual machine software).
-\begin{figure}[h]
- \centering \includegraphics[scale=0.35]{kubernetes_logo}
- \caption{Kubernetes Logo}
- \label{chap:intro:img:k8s_logo}
-\end{figure}
-
-\subsubsection{About Docker}
-
-In recent years, with the new hardware capabilities and the recent development
-of in-kernel virtualization systems (such as Hyper-V) this technology begin to
-be adopted. Virtualization allows to run different systems on the same machine,
-making them completely isolated and then more resilient to failures. The back of
-the medal, though, is that virtual machines require large amount of resources,
-especially memory, because solutions like copy-and-write of in-memory pages are
-not viable anymore (since the kernel gets duplicated too), leading to
-duplication of loaded libraries and assets in the memory of the host. For large
-deployment containing only simple services (e.g. a backend application serving a
-website or a database) this ends up in a waste of resources. \todo{Talk about
-  lxc containers and how Docker solved magically the problem.} It's here that,
-in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
-on an already existing product: Linux Container (LXC). From this framework,
-Docker built an entire ecosystem, consisting in a client/server model giving the
-possibility for users to simply launch, scale and delete containers (locally or
-remotely with Docker-machine), a repository system where images, a layer-based
-``core'' where containers start running from when they are launched, can be
-stored and tagged in a sort of version control system. Finally, another couple
-of solutions, called Docker-compose, and Docker Swarm have the purpose to,
-respectively, orchestrate containers in a single or a clustered system.
-\begin{figure}[t]
- \centering \includegraphics[scale=0.7]{docker_logo}
- \caption{Official Docker logo}
- \label{chap:intro:img:docker_logo}
-\end{figure}
-
-\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
-birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
-seen as one logical unit. This layer completely makes the underlying
-infrastructure transparent: data management, as network one, are totaly managed
-by Docker. Recently, Docker-compose support has been introduced, making this
-one-node tool available for clustered systems. At the end of the day, on one
-hand Docker Swarm provides an easy way to set up a cluster system with all the
-tools configured out-of-the-box, without the necessity to set networking
-configurations or installing ad-hoc storage solutions. On the other hand,
-though, it doesn't have all the personalization options that Kuberentes offers,
-thus making the product less flexible. This foundamental characteristic,
-although, makes the two solutions have different use cases leading to different
-market needs. \todo{Add Docker Swarm}
-
-\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool
-that allows services orchestration. It automates most of the tasks that should
-be performed by hand when launching one or multiple docker containers. It's
-particulary useful when more continers have to interact together,
-because\todo{Write down docker-compose functionalities and describe the product
-  a little bit} all the services rules are described in a simple YAML file. With
-the birth of Docker Swarm, work has started to port this tool to the Swarm
-framework. At the time of this writing, the procedure to port a Docker-compose
-configuration in Docker Swarm is still not completely transparent, and it
-requires a sort of compilation, where Docker-compose bundles all the necessaries
-resources into a binary file that Docker Swarm will use to allocate the
-necessary resources.
+% \section{Technologies used}
 
@@
 and VNFs using virtual machine or exploiting cloud tecnhologies, while they were
-not designed with enough flexibility to integrate with Docker. In the next
-chapter we are going to take a deep analysis of the cited technologies, in order
-to provide to the reader enough context to be able to understand our design
-choices we are going to describe later.
+not designed with enough flexibility an integration with Docker. In the next
+chapter we are going to take a deep analysis of the cited technologies 
+(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
+context to be able to understand our design choices we are going to describe 
+later.
 

### 2018-10-24 res/sections/10-ProjectAnalysis.tex
Move technologies, add harbor description, minor fixes

@@
 \section{Available technologies in the market}
+\label{chap:prjan:sec:tech}
 
@@
 
-\subsection{Openstack}
+\subsubsection{Openstack}
 \label{chap:prjan:sec:openstack}
@@
 
-\todo{This section needs to be expanded a lot!!! In particular, we should put
-  some focus about VNF and MANO, introducing them}
+\subsubsection{Kubernetes}
 
-\subsubsection{Other technologies}\todo{Talk somehow of the other tecnhologies}
-\paragraph{Openvswitch}
-\paragraph{Openbaton} Patrocined by the Institute Fraunhofer, it's a 
-open-source, customizable NFV MANO-compliant framework, that offers many
-configurations options, and since it's written in Java, there is the possibility
-to add components (plug-ins) dynamically. Openbaton can be installed as a normal
-computer program (they offer packages for the most common distros) but it can be
-deployed on Docker containers too. It's designed to work with cloud providers
-like Amazon or Google Compute engine, but it offers compability with on-permise
-cloud solutions like Openstack, where it deploys virtual machines using the API
+As already introduced before, Kubernetes is an open source software framework
+specialized in container management and orchestration. Developed by Google in
+??\todo{Find out the actual date} and written in Go, it allows different
+machines (called nodes from now on) to create an abstraction layer and to form a
+cluster where is possible to run Docker containers without brothering about
+hardware constraints. Nodes can have specific roles and specialization, all of
+them managed by one or many master nodes, that actually manage the
+orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
+that it offers the possibility to change the virtualization software in use
+(i.e. from Container-based technology to Virtual machine software).
+\begin{figure}[h]
+ \centering \includegraphics[scale=0.35]{kubernetes_logo}
+ \caption{Kubernetes Logo}
+ \label{chap:intro:img:k8s_logo}
+\end{figure}
+
+\subsubsection{Docker}
+
+In recent years, with the new hardware capabilities and the recent development
+of in-kernel virtualization systems (such as Hyper-V) this technology begin to
+be adopted. Virtualization allows to run different systems on the same machine,
+making them completely isolated and then more resilient to failures. The back of
+the medal, though, is that virtual machines require large amount of resources,
+especially memory, because solutions like copy-and-write of in-memory pages are
+not viable anymore (since the kernel gets duplicated too), leading to
+duplication of loaded libraries and assets in the memory of the host. For large
+deployment containing only simple services (e.g. a backend application serving a
+website or a database) this ends up in a waste of resources. \todo{Talk about
+  lxc containers and how Docker solved magically the problem.} It's here that,
+in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
+on an already existing product: Linux Container (LXC). From this framework,
+Docker built an entire ecosystem, consisting in a client/server model giving the
+possibility for users to simply launch, scale and delete containers (locally or
+remotely with Docker-machine), a repository system where images, a layer-based
+``core'' where containers start running from when they are launched, can be
+stored and tagged in a sort of version control system. Finally, another couple
+of solutions, called Docker-compose, and Docker Swarm have the purpose to,
+respectively, orchestrate containers in a single or a clustered system.
+\begin{figure}[t]
+ \centering \includegraphics[scale=0.7]{docker_logo}
+ \caption{Official Docker logo}
+ \label{chap:intro:img:docker_logo}
+\end{figure}
+
+\vspace{0.5cm}
+
+% The development of the ETSI MANO testbed involved the use of several 
+% tecnhologies already in the market, and in some cases the creation of ad-hoc 
+% solutions to solve particular issues. Here, a brief listing of the 
+% well-estabilished technologies used is performed, describing even solutions 
+% that at the end weren't choosen because they didn't fit with the defined goals.
+
+\subsection{MANO}
+
+Another foundamental component in the architecture is the MANO, that has to 
+manage the Network Services and the deployment of new VNFs. In coordination 
+with Kubernetes, it composes the core of the testbed architecture. Initially 
+identified with Openbaton (the component is described later 
+in~\ref{chap:prjan:sec:tech:sub:other:sub:openbaton}), afterword it has been 
+discovered to be not fexible enough to support other kinds of virtualization 
+tecniques than pure or in-cloud virtual machines\todo{AAAAAA rewrite this}. 
+Since this was a strong project requisite that could not be easily changed, it 
+has been decided to opt for the creation of a smaller and simpler clone that 
+offered a full integration with Docker and Kubernetes.
+
+\subsubsection{Harbor} \todo{Talk about harbor!}
+Started as a Virtual VNFM (Network Function Virtualization Manager) for 
+Openbaton, during the project it has been expanded to become a complete 
+replacement of the original product, adding the possibility to manage Network 
+Services and to update the route tables, Roulette Controller (the tool 
+description can be found in \todo{Insert roulette ref!}).
+
+\subsection{Technologies used during development}
+
+Since the testbed is composed by different components and thanks to its 
+modularity, the implementation of the single components has been done with the 
+help of additional tools, illustrated below
+
+\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
+birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
+seen as one logical unit. This layer completely makes the underlying
+infrastructure transparent: data management, as network one, are totaly managed
+by Docker. Recently, Docker-compose support has been introduced, making this
+one-node tool available for clustered systems. At the end of the day, on one
+hand Docker Swarm provides an easy way to set up a cluster system with all the
+tools configured out-of-the-box, without the necessity to set networking
+configurations or installing ad-hoc storage solutions. On the other hand,
+though, it doesn't have all the personalization options that Kuberentes offers,
+thus making the product less flexible. This foundamental characteristic,
+although, makes the two solutions have different use cases leading to different
+market needs. \todo{Add Docker Swarm}
+
+\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool
+that allows services orchestration. It automates most of the tasks that should
+be performed by hand when launching one or multiple docker containers. It's
+particulary useful when more continers have to interact together,
+because\todo{Write down docker-compose functionalities and describe the product
+  a little bit} all the services rules are described in a simple YAML file. With
+the birth of Docker Swarm, work has started to port this tool to the Swarm
+framework. At the time of this writing, the procedure to port a Docker-compose
+configuration in Docker Swarm is still not completely transparent, and it
+requires a sort of compilation, where Docker-compose bundles all the necessaries
+resources into a binary file that Docker Swarm will use to allocate the
+necessary resources.
+
+\subsection{Other technologies evaluated}
+
+Additional components have been evaluated during the development of the 
+testbed, and here we propose a list of the most important ones. Some of them 
+where discarded as no more necessary, or during the evolution of the project 
+itself they have become obsolete.
+
+\subsubsection{Openvswitch} \todo{Talk about Openvswitch!!}
+\subsubsection{Openbaton}
+\label{chap:prjan:sec:tech:sub:other:sub:openbaton}
+Patrocined by the Institute Fraunhofer, it's a open-source, customizable NFV 
+MANO-compliant framework, that offers manyconfigurations options, and since it's 
+written in Java, there is the possibility to add components (plug-ins) 
+dynamically. Openbaton can be installed as a normal computer program (they 
+offer packages for the most common distros) but it can be deployed on Docker 
+containers too. It's designed to work with cloud providers like Amazon or 
+Google Compute engine, but it offers compability with on-permise cloud 
+solutions like Openstack, where it deploys virtual machines using the API
 the platform offers. It can store VNF configurations saved as TOSCA
@@
 \end{figure}
-
-\paragraph{SDN?}
+% \paragraph{SDN?}

### 2018-10-25 res/sections/10-ProjectAnalysis.tex
Add about Harbor

@@
 
-\subsubsection{Harbor} \todo{Talk about harbor!}
-Started as a Virtual VNFM (Network Function Virtualization Manager) for 
-Openbaton, during the project it has been expanded to become a complete 
-replacement of the original product, adding the possibility to manage Network 
-Services and to update the route tables, Roulette Controller (the tool 
-description can be found in \todo{Insert roulette ref!}).
+\subsubsection{Harbor}
+Started as a Virtual VNFM (Network Function Virtualization Manager) to offer 
+a bridge for Openbaton to interact with Kubernetes, during the project it has 
+been expanded to become a complete replacement of the original product, adding 
+the possibility to manage Network Services and to update the route tables, 
+Roulette Controller (the tool description can be found below
+in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web 
+interface, but it has a RESTful API that can allow users to interact with him 
+using CRUD operations.
+
+\subsection{SFC}
+
+\subsubsection{Roulette Controller}
+\label{chap:prjan:sec:Stech:sub:SFC:sub:roulette}
+
+\subsubsection{Astaire}
+
+\subsubsection{Ironhide}
 

### 2018-10-25 res/sections/10-ProjectAnalysis.tex
Add additional Harbor content

--- 2018-10-24 res/sections/10-ProjectAnalysis.tex
@@
 
-\subsubsection{Harbor} \todo{Talk about harbor!}
-Started as a Virtual VNFM (Network Function Virtualization Manager) for 
-Openbaton, during the project it has been expanded to become a complete 
-replacement of the original product, adding the possibility to manage Network 
-Services and to update the route tables, Roulette Controller (the tool 
-description can be found in \todo{Insert roulette ref!}).
+\subsubsection{Harbor}
+Started as a VNFM (Virtual Network Function Virtualization Manager) to offer 
+a bridge for Openbaton to interact with Kubernetes, during the project it has 
+been expanded to become a complete replacement of the original product, adding 
+the possibility to manage Network Services and to update the routing tables 
+contained in the Roulette Controller (the tool description can be found below
+in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web 
+interface, but it has a RESTful API that allow users to interact with him 
+using CRUD operations. It offers functionalities similar to Openbaton: 
+possibility to upload VNF definitions (basically, Kuberentes YAML deployments), 
+and possibility to upload NS definitions. When an NS is saved, a consistency 
+check is performed to assure that alle the VNF are already present in Harbor. 
+Similarly to Openbaton an NS is deployed, all the associated VNFs are deployed 
+with it, making the whole operation atomic from the user point-of-view. 
+What Openbaton doesn't do is a smart resource utilization: Harbor, in fact, 
+during NS is deletion, it checks if any related VNF remains orphan (which means 
+without any NS using it). In that case the VNF will enter in a state of 
+pruning: after five minutes, if the VNF has not be reutilized by another NS it 
+will be deleted. This mechanism assure faster deployments in environments where 
+NSs are continuously launched and deleted.
+
+\subsection{SFC}
+
+\subsubsection{Roulette Controller}
+\label{chap:prjan:sec:Stech:sub:SFC:sub:roulette}
+
+\subsubsection{Astaire}
+
+\subsubsection{Ironhide}
 

### 2018-10-26 res/sections/10-ProjectAnalysis.tex
Fix English

@@
 \subsubsection{Harbor}
-Started as a VNFM (Virtual Network Function Virtualization Manager) to offer 
-a bridge for Openbaton to interact with Kubernetes, during the project it has 
-been expanded to become a complete replacement of the original product, adding 
-the possibility to manage Network Services and to update the routing tables 
+Started as a VNFM (Virtual Network Function Virtualization Manager) to offer a
+bridge for Openbaton to interact with Kubernetes, during the project it has been
+expanded to become a complete replacement of the original product, adding the
+possibility to manage Network Services and to update the routing tables
 contained in the Roulette Controller (the tool description can be found below
 in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web
-interface, but it has a RESTful API that allow users to interact with him 
-using CRUD operations. It offers functionalities similar to Openbaton: 
-possibility to upload VNF definitions (basically, Kuberentes YAML deployments), 
-and possibility to upload NS definitions. When an NS is saved, a consistency 
-check is performed to assure that alle the VNF are already present in Harbor. 
-Similarly to Openbaton an NS is deployed, all the associated VNFs are deployed 
-with it, making the whole operation atomic from the user point-of-view. 
-What Openbaton doesn't do is a smart resource utilization: Harbor, in fact, 
-during NS is deletion, it checks if any related VNF remains orphan (which means 
-without any NS using it). In that case the VNF will enter in a state of 
-pruning: after five minutes, if the VNF has not be reutilized by another NS it 
-will be deleted. This mechanism assure faster deployments in environments where 
-NSs are continuously launched and deleted.
+interface, but it has a RESTful API that allow users to interact with it using
+CRUD operations. It offers functionalities similar to Openbaton: possibility to
+save and retrieve VNF and NS definitions (basically, Kuberentes YAML
+deployments). When an NS is saved, a consistency check is performed to assure
+that all the VNFs have already been defined in Harbor. On an NS deployment, with
+a mechanism similar to the Openbaton one, all the associated VNFs are set up
+too, making the whole operation of launching a NS atomic from the user
+point-of-view. What Openbaton doesn't do is a smart resource utilization:
+Harbor, in fact, during NS is deletion, it checks if any related VNF remains
+orphan (which means without any NS using it). In that case the VNF will enter in
+a state of pruning: after five minutes, if the VNF has not be reutilized by any
+other NS it will be deleted. This mechanism assure faster deployments in
+environments where NSs are continuously taken up and down.
 
@@
 
-\paragraph{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
+\subsubsection{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
 birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
@@
 
-\paragraph{Docker-compose} As briefly discussed above, Docker-compose is a tool
+\subsubsection{Docker-compose} As briefly discussed above, Docker-compose is a tool
 that allows services orchestration. It automates most of the tasks that should

### 2018-10-26 res/sections/10-ProjectAnalysis.tex
Add paragraph about roulette

@@
 \subsubsection{Harbor}
+\label{chap:prjan:sec:tech:sub:MANO:sub:harbor}
 Started as a VNFM (Virtual Network Function Virtualization Manager) to offer a
@@
 in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web
-interface, but it has a RESTful API that allow users to interact with it using
+interface, but it has a RESTful API that allows users to interact with it using
 CRUD operations. It offers functionalities similar to Openbaton: possibility to
@@
 point-of-view. What Openbaton doesn't do is a smart resource utilization:
-Harbor, in fact, during NS is deletion, it checks if any related VNF remains
+Harbor, in fact, during a NS deletion, it checks if any related VNF remains
 orphan (which means without any NS using it). In that case the VNF will enter in
@@
 other NS it will be deleted. This mechanism assure faster deployments in
-environments where NSs are continuously taken up and down.
+environments where NSs are continuously taken up and down. \todo{I've made an
+  Harbor implementation schema, I dunno if its usefult to put it here}
 
 \subsection{SFC}
+The tools analysis revealed little about already existent VNF or NS
+implementations, so it was neccesary to fill the gap. Since this one of our
+thesis goals, we devepled a complete solution to create a complete SFCs, trying
+to standardize our vocabulary as much as possible, and giving basic
+documentation when possible.
 
 \subsubsection{Roulette Controller}
-\label{chap:prjan:sec:Stech:sub:SFC:sub:roulette}
+\label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
+Dynamic communications between VNFs would not be possible without a thrid
+component acting as a route controller. Roulette fulfulls this role, providing a
+backend where the single VNF can ask the next packet hop. Indeed, routes must be
+submitted by another component, that we identified as Harbor (component
+description available above in
+sub-section~\ref{chap:prjan:sec:tech:sub:MANO:sub:harbor}). After a NS
+deployment, Roulette gets instructed with a new route describing a NS. Every NS
+has a proper ID (called \verb!sfid!), and thus when the classifier checks a
+packet typology, it labels that packet with the specific NS ID (for a detailed
+description about the SFC header see Section~\todo{Add section/chapter/paragraph
+  ref!}). Finally, thanks to this label, every VNF knows the packet's chain and
+it will correctly forward the it to the next VNF.
 

### 2018-10-29 res/sections/10-ProjectAnalysis.tex
Fix typo
Add Ironhide and Astaire description

@@
 Dynamic communications between VNFs would not be possible without a thrid
-component acting as a route controller. Roulette fulfulls this role, providing a
+component acting as a route controller. Roulette fulfills this role, providing a
 backend where the single VNF can ask the next packet hop. Indeed, routes must be
@@
 \subsubsection{Astaire}
+As is possible to denote, a VNF doesn't only process the incoming packets, 
+but it has to manage the communications and the related forward operations. To 
+possibly ease part of the transmission problem and to make it easier to create a 
+VNF without requiring the whole knowledge about SFC, it has been decided to 
+build a framework, Astaire, that could keep the SFC related details decoupled 
+from the core of the VNF, in other words, the entity that actually elaborates 
+the packet. Built using an event-based pattern, it presents hooks that allow 
+other software written in different languages from the Astaire one (e.g. Java), 
+to be called dynamically at runtime in order to process the packet. The result 
+of this elaboration gets returned to astaire, that performs the beforehand 
+cited operations of VNFs communications and forwarding.
 
 \subsubsection{Ironhide}
+As described in chapter~\ref{chap:archimpl}, the whole framework need to have 
+and entry and an exit point. These, defined as \textit{Ingress} and 
+\textit{Egress} points, are performed by Ironhide. Ironhide acts like a proxy, 
+hiding the whole packet processing to the sender $S$ and the receiver $R$. When 
+a client $S$ performs a request to $R$, the packet actually goes through 
+Ironhide, who has an internal classifier that it is able to recognise the type 
+of connection and to choose the right NS the packet has to go through. This 
+process in foundamental since it keeps the connection between $S$ and $R$ up, 
+for both TCP and UDP based communications.
 
@@
 
-\subsubsection{Docker Swarm} Introduced in ??\todo{Find Docker Swarm date of
-birth}, Docker Swarm allows multiple Docker nodes to cluster together and be
-seen as one logical unit. This layer completely makes the underlying
-infrastructure transparent: data management, as network one, are totaly managed
-by Docker. Recently, Docker-compose support has been introduced, making this
-one-node tool available for clustered systems. At the end of the day, on one
-hand Docker Swarm provides an easy way to set up a cluster system with all the
-tools configured out-of-the-box, without the necessity to set networking
-configurations or installing ad-hoc storage solutions. On the other hand,
-though, it doesn't have all the personalization options that Kuberentes offers,
-thus making the product less flexible. This foundamental characteristic,
-although, makes the two solutions have different use cases leading to different
-market needs. \todo{Add Docker Swarm}
-
-\subsubsection{Docker-compose} As briefly discussed above, Docker-compose is a tool
-that allows services orchestration. It automates most of the tasks that should
-be performed by hand when launching one or multiple docker containers. It's
-particulary useful when more continers have to interact together,
-because\todo{Write down docker-compose functionalities and describe the product
-  a little bit} all the services rules are described in a simple YAML file. With
-the birth of Docker Swarm, work has started to port this tool to the Swarm
-framework. At the time of this writing, the procedure to port a Docker-compose
-configuration in Docker Swarm is still not completely transparent, and it
-requires a sort of compilation, where Docker-compose bundles all the necessaries
-resources into a binary file that Docker Swarm will use to allocate the
-necessary resources.
+\subsubsection{Docker-compose} 
+Docker-compose is a tool that allows services orchestration. It automates most 
+of the tasks that should be performed by hand when launching one or multiple 
+docker containers. It's particulary useful when more continers have to interact 
+together, because\todo{Write down docker-compose functionalities and describe 
+the product a little bit} all the services rules are described in a simple 
+YAML file. With the birth of Docker Swarm, work has started to port this tool to 
+the Swarm framework. At the time of this writing, the procedure to port a 
+Docker-compose configuration in Docker Swarm (which a description of the 
+product can be found in 
+sub-section~\ref{chap:prjan:sec:tech:sub:other:sub:swarm}) is still not 
+completely transparent, and it requires a sort of compilation, where 
+Docker-compose bundles all the necessaries resources into a binary file that 
+Docker Swarm will use to allocate the necessary resources.
 
@@
 
+\subsubsection{Docker Swarm}
+\label{chap:prjan:sec:tech:sub:other:sub:swarm}
+Introduced in ??\todo{Find Docker Swarm date of birth}, Docker Swarm allows 
+multiple Docker nodes to cluster together and be seen as one logical unit. This 
+layer completely makes the underlying infrastructure transparent: data 
+management, as network one, are totaly managed by Docker. Recently, 
+Docker-compose support has been introduced, making this one-node tool available 
+for clustered systems. At the end of the day, on one hand Docker Swarm provides 
+an easy way to set up a cluster system with all the tools configured 
+out-of-the-box, without the necessity to set networking configurations or 
+installing ad-hoc storage solutions. On the other hand, though, it doesn't have 
+all the personalization options that Kuberentes offers, thus making the product 
+less flexible. This foundamental characteristic, although, makes the two 
+solutions have different use cases leading to different market needs.
+
 \subsubsection{Openvswitch} \todo{Talk about Openvswitch!!}
@@
 Patrocined by the Institute Fraunhofer, it's a open-source, customizable NFV 
-MANO-compliant framework, that offers manyconfigurations options, and since it's 
+MANO-compliant framework, that offers manyconfigurations options, and since 
+it's 
 written in Java, there is the possibility to add components (plug-ins) 
@@
 ones, making the product really flexible to different platforms. Performing this
-operation requires a deep knowlege of the framework though.
+operation requires a deep knowledge of the framework though.
 \begin{figure}[h]
@@
 \end{figure}
-% \paragraph{SDN?}

### 2018-10-31 res/sections/10-ProjectAnalysis.tex
Add partial SFC introduction

@@
 include parts of RFC 7665}
+
+Even though SFCs and VNFs are part of the ETSI architecture, the complexity 
+behind was so high to require particular care and a separated study. Indeed, 
+the  SFCs are the core of the whole project: packet transmission and 
+elaboration is made here. Challenges regarding this topic are building a 
+system that is flexible enough to allow dynamic packet routing, support 
+scalability and provide methods to recovery from failures. With this challenges 
+packet transmission revealed to be a foundamental point to choose carefully.
 

### 2018-11-04 res/sections/08-Background.tex
Add some part on SDN and on virtualization

@@
 going to briefly introduce and explain the technologies involved and their role.
+
+\section{Software-defined networking}
+The SDN is an emerging approach to cloud computing that allows to configure
+network in a programmatically way, possibly enhancing monitoring and
+performances. The main sake of SDN is to the packet forwarding process,
+\emph{data plane}, from the routing one, the so called \emph{control plane}.
+The latter component serves as the centralized network intelligence and could
+consists of one or more controllers. It has some drawbacks in teorms of
+security, scalability and elasticity, because of the centralized architecture.
+
+\subsection{The SDN stack}
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=1.3]{sdn_architecture}
+ \caption{SDN architecture}
+ \label{chap:background:img:sdn_architecture}
+\end{figure}
+As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
+architecture can be splitted into three different layers, based on their
+functionalities. \emph{SDN Applications} are programs that communicate the
+behaviour to the SDN controllers and request for needed resources and data
+through APIs, often called \emph{northbound interfaces}(NBI). NBIs add a level
+of abstraction to the higher level application on the set of functionalities of
+the controllers: ideally, taking advantage of NBIs, application could use
+controllers of different vendors in the same way. So, SDN applications can be
+seen as 2 layers programs, the logic and the communication part that uses NBIs.
+Theese applications can also build an abstracted view of the network, using
+information of controllers, in order to make easier the analysis of the traffic
+and simplifing the decision making process.
+
+\emph{SDN controllers}, instead, are logically centralized components that
+receive requests from the application and either translate them in a fashion
+that can be understood from the lower layer, or collect data needed by the
+application. Because of they have to accomplish theese aims, they can be see
+as 3 layers programs, in which a layers exposes an interface to the upper
+applications (\emph{NBI agents}), the second one represent the SDN controller
+logic and finally the third layer has the task to communicate with the network
+(Control to Data-Plane Interface or southbound interface).
+
+In the lowest part of the architecture we can find the \emph{SDN Networking Devices}.
+These ones control packet forwarding and data processing capabilities of the network.
+They comprises a CDPI agent and a set of one or more traffic forwarding engines and zero or more traffic processing functions (the logic of the devices).
+
+\subsection{The OpenFlow protocol}
+The SDN strategy is often associated with OpenFlow protocol: this communication
+protocol aim is to give access to the forwarding network component to the lower
+layer network infrastructure, made of switches and routers. OpenFlow enables
+the remote administration of networks working on layer 3: packets forwarding
+tables can be modified relying on packet matching rules. Following this
+approach it is possible to change routes periodically or ad-hoc, based on
+packet types, traffic or network load.
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.65]{openflow}
+ \caption{OpenFlow protocol communication graph}
+ \label{chap:background:img:openflow_protocol}
+\end{figure}
+
+In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
+how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
+set of tables to organizedpacket control and forwarding. These tables can be
+updated by higher-level application, usually an SDN application or controller,
+taking advatage of OpenFlow channels (OF channels).
+
+\section{Virtualization}
+Cloud computing is now leading-edge Internet technology and virtualization
+technology is one of the important part of cloud computing system.
+Virtualization, in computing, refers to the act of creating a virtual (rather
+than actual) version of something, including but not limited to a virtual
+computer hardware platform, operating system (OS), storage device, or computer
+network resources. \todo{To change I do not what can I write, stolen from ``The
+research and implementation of cloud computing platform based on docker'' paper}
+
+\subsection{Network Virtualization}
+%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAccessKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41fBNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_network_virtualization.pdf
+Virtualization, in the networking context, refers to the act of combining
+hardware, software and functionalities into a virtual network. This abstraction
+runs on the top of a physical network in a hypervisor, decupling the logical
+network from the harwdare one. Network virtualization allow multiple,
+heterogeneous networks to cohabit, increasing the flexibility and manageability.
+%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
+%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
+The idea of network virtualization came from the decoupling of the role of the
+traditional ISP into two different entities: who provides the infrastucture
+(InPs) and who provides the services(SPs). In that context it is possible to
+aggregate resources, even from different InPs, creating a virtual network,
+offering an end-to-end service. Most known instances of NV are \emph{VPN},
+dedicated networks that allows the communication among multiple sites, using
+private and secure tunnels, relying on the public network such as Internet.
+Other examples of netvork virtualization are \emph{overlay networks}. They are
+logical networks, build on the top of an existing physical network, tipically
+implemented at application layer.
+
+\subsection{Network Function Virtualization}
+The NFC concept refers to the use of virtualization tecnologies to virtualize
+network nodes functions. VNFs may run different processes and softwares, on the
+top of servers, switches, storage devices or on cloud computing infrastucture,
+that substitute custom hardware. That approach worked well on the past, because
+of the need to follow rigorous standards, providing at the same time high
+quality, but it led to long product cycles and high specialized proprietary
+hardware. New Internet senarios, that include IoT and the increasing number of
+services to spread multimedia data (Skype, Netflix, Spotify, ...), requires
+more flexibles function and faster network deployment. NFV technology is
+exploited in a framework, composed at least by three components:
+\begin{enumerate}
+ \item a software that performs the network function that must be deployed,
+ that is the VNF;
+ \item NFV Infrastructure (NFVI) that are all hardware and software components
+ that build the environment where VNFs are deployed;
+ \item NFV Management And Orchestration (NFV-MANO) architectural framework that
+ is all functional blocks, data repositories used by these blocks, and
+ reference points and interfaces through which these functional blocks exchange
+ information for the purpose of managing and orchestrating NFVI and VNFs.
+\end{enumerate}
 

### 2018-11-06 res/sections/79-Conclusions.tex
Add WIP conclusions

\chapter{Conclusions}
\label{chap:conclusions}

In this thesis we have seen an ETSI MANO implementation proposal. We have 
described the technological components needed to build it, and we have 
presented the requirements of the whole implementations. Then we described 
the challenges we have encountered and how we solved it, taking choises 
valuating the pros and cons.

Finally, we corroborated our implementation with test, showing that a 
containerized approach is faster and more reliable thant a virtualized one

### 2018-11-07 res/sections/79-Conclusions.tex
Finalized conclusions

@@
 
-In this thesis we have seen an ETSI MANO implementation proposal. We have 
-described the technological components needed to build it, and we have 
-presented the requirements of the whole implementations. Then we described 
+In this thesis we have seen an ETSI MANO and RFC 7665 implementation proposal. 
+We have described the technological components needed to build it, and we have 
+presented the requirements of the whole implementation. Then we described 
 the challenges we have encountered and how we solved it, taking choises 
-valuating the pros and cons.
+valuating their pros and cons.
 
-Finally, we corroborated our implementation with test, showing that a 
-containerized approach is faster and more reliable thant a virtualized one
+Finally, we corroborated our implementation with tests, showing that a 
+containerized approach is faster and more reliable thant a virtualized one,
+and illustrating that latency in communications through SFCs, while there 
+is room for improvement, are decent considered the testbed environment.
+
+In conclusion, we can say that the ETSI MANO specification, coupled with the 
+RFC 7665 proposal and the new virtualization technologies avaiable in the 
+market could change the shape of future mobile communications, leading to the 
+creation of a 5G able to saddisfy the challenges of the present, such as IoT 
+communications, media streaming, and real-time data, with the challenges of the 
+future, as smart dust emerges as the next possible innovation.

### 2018-11-09 res/sections/08-Background.tex
80 columns

@@
 As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be splitted into three different layers, based on their
+architecture can be splitted in three different layers, based on their
 functionalities. \emph{SDN Applications} are programs that communicate the
@@
 
-In the lowest part of the architecture we can find the \emph{SDN Networking Devices}.
-These ones control packet forwarding and data processing capabilities of the network.
-They comprises a CDPI agent and a set of one or more traffic forwarding engines and zero or more traffic processing functions (the logic of the devices).
+In the lowest part of the architecture we can find the \emph{SDN Networking 
+Devices}. These ones control packet forwarding and data processing capabilities 
+of the network. They comprises a CDPI agent and a set of one or more traffic 
+forwarding engines and zero or more traffic processing functions (the logic of 
+the devices).
 
@@
 \subsection{Network Virtualization}
-%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAccessKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41fBNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_network_virtualization.pdf
+%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
+%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
+%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
+%r k _virtualization.pdf
 Virtualization, in the networking context, refers to the act of combining

### 2018-11-09 res/sections/08-Background.tex
Polpetta's review

@@
 architecture can be splitted in three different layers, based on their
-functionalities. \emph{SDN Applications} are programs that communicate the
-behaviour to the SDN controllers and request for needed resources and data
-through APIs, often called \emph{northbound interfaces}(NBI). NBIs add a level
-of abstraction to the higher level application on the set of functionalities of
-the controllers: ideally, taking advantage of NBIs, application could use
-controllers of different vendors in the same way. So, SDN applications can be
-seen as 2 layers programs, the logic and the communication part that uses NBIs.
+functionalities. The \emph{SDN Applications} role is to set the 
+behaviour of the SDN controllers and manage its hardware resources and data 
+through APIs, often called \emph{northbound interfaces}(NBI).
+These interfaces provide an additional abstraction layer that high-level 
+applications can consume, enabling the possibility to use a variety of 
+different verdors' controllers in the same way.
+Analysing the SDN applications is possible to note two defined layers, one that 
+is responsible to manage the whole logic, and another one dedicated to the 
+communication part, that in particular uses the previsously defined NBIs.
 Theese applications can also build an abstracted view of the network, using
-information of controllers, in order to make easier the analysis of the traffic
-and simplifing the decision making process.
+controllers information, in order to make the traffic analysis easier and to 
+simplify the decision making process.
 
@@
 even if we support a plethora of protocols.''} making the whole architecture 
-indipendent from the protocol a particolar flux of data. To achieve this, 
+indipendent from the protocol a particular flux of data. To achieve this, 
 several solutions have been studied, and at the end packet incapsulation with 

### 2018-11-11 res/sections/08-Background.tex
Fix SDN stack subsection

@@
 
-\subsection{The SDN stack}
+\subsection{SDN stack}
 \begin{figure}[t]
@@
  \includegraphics[scale=1.3]{sdn_architecture}
- \caption{SDN architecture}
+ \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
+denote the three layer logic: the Application plane, the Controller plane and 
+the Data plane.}
  \label{chap:background:img:sdn_architecture}
@@
 As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be splitted in three different layers, based on their
-functionalities. The \emph{SDN Applications} role is to set the 
-behaviour of the SDN controllers and manage its hardware resources and data 
-through APIs, often called \emph{northbound interfaces}(NBI).
+architecture can be splitted in three different layers, called planes, based on 
+their functionalities.
+
+\subsubsection{Application plane}
+
+The \emph{Application plane} role is to set the behaviour of the SDN controllers 
+and manage its hardware resources and data through APIs, often called 
+\emph{northbound interfaces} (NBI).
 These interfaces provide an additional abstraction layer that high-level 
@@
 communication part, that in particular uses the previsously defined NBIs.
-Theese applications can also build an abstracted view of the network, using
+These applications can also build an abstracted view of the network, using
 controllers information, in order to make the traffic analysis easier and to 
@@
 
-\emph{SDN controllers}, instead, are logically centralized components that
-receive requests from the application and either translate them in a fashion
-that can be understood from the lower layer, or collect data needed by the
-application. Because of they have to accomplish theese aims, they can be see
-as 3 layers programs, in which a layers exposes an interface to the upper
-applications (\emph{NBI agents}), the second one represent the SDN controller
-logic and finally the third layer has the task to communicate with the network
-(Control to Data-Plane Interface or southbound interface).
+\subsubsection{Controller plane}
+
+\emph{SDN controllers}, are logically centralized components that
+receive requests from the mentioned SDN applications and either translate them 
+in a fashion that can be understood from the lower layer\todo{who?}, or collect 
+data needed by the application itself. 
+Due to the variety of tasks a controller has to accomplish, its logical 
+suddivision is achieved with a three-layer architecture, in which every layer 
+exposes it's own interface, that can be consequently used by the upper 
+layers, up to the \emph{NBI agents}\todo{explain them?}.
+In particular, the middle layer represents the controller inner-logic, which 
+process the requests coming from the upper layer (aka the third layer), that 
+has the specific task to communicate with the remaining network compoents, in 
+other words the Control to Data-Plane Interface and/or the already 
+mentioned southbound interface.
+
+\subsubsection{Data plane}
 
 In the lowest part of the architecture we can find the \emph{SDN Networking 
-Devices}. These ones control packet forwarding and data processing capabilities 
-of the network. They comprises a CDPI agent and a set of one or more traffic 
-forwarding engines and zero or more traffic processing functions (the logic of 
-the devices).
+Devices}. These ones controls the packet forwarding and data processing 
+capabilities of the network. They consist of a CDPI agent and a set of one or 
+multiple traffic forwarding engines and of a zero or more traffic processing 
+functions, that are, as a matter of fact, the devices logic.
 
-\subsection{The OpenFlow protocol}
+\subsection{OpenFlow protocol}
 The SDN strategy is often associated with OpenFlow protocol: this communication
-protocol aim is to give access to the forwarding network component to the lower
+protocol aims is to give access to the forwarding network component to the lower
 layer network infrastructure, made of switches and routers. OpenFlow enables

### 2018-11-11 res/sections/08-Background.tex
Fix english

@@
 how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
-set of tables to organizedpacket control and forwarding. These tables can be
+set of tables to organized packet control and forwarding. These tables can be
 updated by higher-level application, usually an SDN application or controller,
-taking advatage of OpenFlow channels (OF channels).
+taking advatage of OpenFlow channels (called OF channels).
 
 \section{Virtualization}
-Cloud computing is now leading-edge Internet technology and virtualization
-technology is one of the important part of cloud computing system.
+Cloud computing is now a leading-edge Internet technology and the virtualization
+technology is one of the most important part of cloud computing systems.
 Virtualization, in computing, refers to the act of creating a virtual (rather
-than actual) version of something, including but not limited to a virtual
+than an actual) version of something, including but not limited to a virtual
 computer hardware platform, operating system (OS), storage device, or computer
@@
 aggregate resources, even from different InPs, creating a virtual network,
-offering an end-to-end service. Most known instances of NV are \emph{VPN},
+offering an end-to-end service. Most known instances of NV are \emph{VPNs},
 dedicated networks that allows the communication among multiple sites, using
@@
 logical networks, build on the top of an existing physical network, tipically
-implemented at application layer.
+implemented at the application layer.
 
 \subsection{Network Function Virtualization}
-The NFC concept refers to the use of virtualization tecnologies to virtualize
-network nodes functions. VNFs may run different processes and softwares, on the
-top of servers, switches, storage devices or on cloud computing infrastucture,
-that substitute custom hardware. That approach worked well on the past, because
-of the need to follow rigorous standards, providing at the same time high
-quality, but it led to long product cycles and high specialized proprietary
-hardware. New Internet senarios, that include IoT and the increasing number of
-services to spread multimedia data (Skype, Netflix, Spotify, ...), requires
-more flexibles function and faster network deployment. NFV technology is
-exploited in a framework, composed at least by three components:
+The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
+virtualization tecnologies to virtualize network nodes functions. VNFs may run 
+different processes and softwares, on the top of servers, switches, storage 
+devices or on cloud computing infrastuctures, that substitute custom hardware. 
+That approach worked well in the past, because of the need to follow rigorous 
+standards, providing at the same time high quality, but it led to long product 
+cycles and high specialized proprietary hardware. New Internet senarios, that 
+include IoT and the increasing number of services to spread multimedia data 
+(Skype, Netflix, Spotify, ...), requires more flexible functions and faster 
+network deployments. The NFV technology is exploited in a framework, organized
+at least by three components:
 \begin{enumerate}
@@
  that is the VNF;
- \item NFV Infrastructure (NFVI) that are all hardware and software components
- that build the environment where VNFs are deployed;
+ \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
+components that builds the environment where the VNFs are deployed;
  \item NFV Management And Orchestration (NFV-MANO) architectural framework that
- is all functional blocks, data repositories used by these blocks, and
+ is all the functional blocks, data repositories used by these blocks,
  reference points and interfaces through which these functional blocks exchange

### 2018-11-12 res/sections/25-Tests.tex
Add stuff about tests

\chapter{Tests and metrics}

After the testbed implementation, tests about its performance and metrics were 
gathered in order to understand the real project effectiveness. We performed 
tests relatively at two main sections: one testing the architecture 
overall responsivness and one checking the SFC implementation and efficiency.

Regarding the overall system responsivness, a metric we considered worth to 
test was the VNF start up time in a docker environment versus one virtualized 
through a common virtual machine system (we actually used Virtual Box). To 
make this simulation as fair as possible, we used the same node (an Openstack 
VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the same ``boot 
sequence'' for both the VNFs: first we launched the Astaire framework, 
ensuring it's availability to process data and then we notified the test 
machine of the successful boot. The startup time was measured from the 
beginning of the startup command (\verb!docker run! for docker and 
\verb!VBoxManage -s! for Virtual Box) till the first TCP hit to the test 
backend received (to make things as smooth as possible, \verb!netcat! was used 
to listen to incoming TCP data).

It is worth saying that Docker metrics are generally more precise than the 
Virtual Box one: docker allows to inspect the container and to gain startup and 
shutdown times with a precision of millseconds, while to gather this 
information for VirtualBox instances we had to use the command line utility 
\verb!time!, that registered timing with a precision of centiseconds\todo{For 
the love of god check it out}. Nonetheless, this didn't influence the test too 
much % explain that the difference is so big it doesn't really matter at the end

### 2018-11-12 res/sections/08-Background.tex
add explaination of NBI agent

@@
 exposes it's own interface, that can be consequently used by the upper 
-layers, up to the \emph{NBI agents}\todo{explain them?}.
+layers, up to the \emph{NBI agents}.
+NBI agents manages the communication with SDN applications, exposing interfaces that usually provide an abstract view of the network and a set of functionalities of the controller.
 In particular, the middle layer represents the controller inner-logic, which 

### 2018-11-12 res/sections/08-Background.tex
Reword todo

@@
 layers, up to the \emph{NBI agents}.
-NBI agents manages the communication with SDN applications, exposing interfaces that usually provide an abstract view of the network and a set of functionalities of the controller.
+NBI agents manages the communication with SDN applications, exposing interfaces
+that usually provide an abstract view of the network and a set of
+functionalities of the controller.
 In particular, the middle layer represents the controller inner-logic, which 
@@
 computer hardware platform, operating system (OS), storage device, or computer
-network resources. \todo{To change I do not what can I write, stolen from ``The
-research and implementation of cloud computing platform based on docker'' paper}
+network resources. \todo{This part was taken from ``The research and
+implementation of cloud computing platform based on docker'' paper, rework it
+in a way it doesn't seem to be copied.}
 

### 2018-11-13 res/sections/25-Tests.tex
Add part regarding tests

@@
 gathered in order to understand the real project effectiveness. We performed 
-tests relatively at two main sections: one testing the architecture 
+tests relatively two main sections: one testing the architecture 
 overall responsivness and one checking the SFC implementation and efficiency.
@@
 test was the VNF start up time in a docker environment versus one virtualized 
-through a common virtual machine system (we actually used Virtual Box). To 
-make this simulation as fair as possible, we used the same node (an Openstack 
-VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the same ``boot 
-sequence'' for both the VNFs: first we launched the Astaire framework, 
+through a common virtual machine system (in this case Virtual Box was choosen). 
+To make this simulation as fair as possible, we used the same node (an 
+Openstack VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the same 
+``boot sequence'' for both the VNFs: first we launched the Astaire framework, 
 ensuring it's availability to process data and then we notified the test 
@@
 beginning of the startup command (\verb!docker run! for docker and 
-\verb!VBoxManage -s! for Virtual Box) till the first TCP hit to the test 
-backend received (to make things as smooth as possible, \verb!netcat! was used 
+\verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
+test backend (to make things as smooth as possible, \verb!netcat! was used 
 to listen to incoming TCP data).
@@
 It is worth saying that Docker metrics are generally more precise than the 
-Virtual Box one: docker allows to inspect the container and to gain startup and 
+Virtual Box one: Docker allows to inspect the container and to gain startup and 
 shutdown times with a precision of millseconds, while to gather this 
 information for VirtualBox instances we had to use the command line utility 
-\verb!time!, that registered timing with a precision of centiseconds\todo{For 
-the love of god check it out}. Nonetheless, this didn't influence the test too 
-much % explain that the difference is so big it doesn't really matter at the end
+\verb!time!, that registered timing with a precision of hundredths of a 
+second. Nonetheless, this didn't influence the test too much. To avoid possible 
+outliers we perfomed the test $100$ times. With this $N$, the total time spent 
+by the Virtual Machines to boot up was of $2000$\todo{Update with the real 
+value} seconds, while with containers the same number of start ups resulted in a 
+total of $1.xx$\todo{Update with the real value}. % Insert graphs, explain the 
+%problems of virtual machines
+
+% Talk about the next test

### 2018-11-13 res/sections/25-Tests.tex
Add content

@@
 %problems of virtual machines
+\begin{figure}[t] % TODO set subfigure and put the pie plot on the right!
+ \centering
+ \includegraphics[scale=0.5]{docker_vs_vmBarplotGraph}
+ \caption[Docker vs Virtual Box barplot results]{Docker vs Virtual Box barplot 
+results. It is possible to see how the total time to start up a virtual machine 
+through Virtual Box is very high compared to the time required to the overall 
+Docker containers to boot. A single VirtualBox instance requires on average 20 
+seconds to boot up, while a container requires, on average, only 0.02 seconds. 
+The main reason behind this difference, as already explained in other parts of 
+this thesis, it is caused by the guest's kernel that has to initialized itself, 
+before starting any program. On Docker, instead, the kernel is the same of the 
+host, so the only operations needed are to isolate the program from the rest of 
+the OS and add it to the process list.}
+ \label{chap:tests:sec:dockervsvb:img:barplot}
+\end{figure}
 

### 2018-11-14 res/sections/25-Tests.tex
Add subfigures

@@
 %problems of virtual machines
-\begin{figure}[t] % TODO set subfigure and put the pie plot on the right!
+
+\begin{figure}[t] % TODO fix image dimensions!
     \centering
- \includegraphics[scale=0.5]{docker_vs_vmBarplotGraph}
- \caption[Docker vs Virtual Box barplot results]{Docker vs Virtual Box barplot 
-results. It is possible to see how the total time to start up a virtual machine 
-through Virtual Box is very high compared to the time required to the overall 
-Docker containers to boot. A single VirtualBox instance requires on average 20 
-seconds to boot up, while a container requires, on average, only 0.02 seconds. 
-The main reason behind this difference, as already explained in other parts of 
-this thesis, it is caused by the guest's kernel that has to initialized itself, 
-before starting any program. On Docker, instead, the kernel is the same of the 
-host, so the only operations needed are to isolate the program from the rest of 
-the OS and add it to the process list.}
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{docker_vs_vmBarplotGraph}
+        \caption{Barplot between Virtual Box and Docker startup times.}
         \label{chap:tests:sec:dockervsvb:img:barplot}
+    \end{subfigure}
+    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
+      %(or a blank line to force the subfigure onto a new line)
+    \begin{subfigure}[b]{0.3\textwidth}
+        \includegraphics[width=\textwidth]{docker_vs_vmBoxplotGraph}
+        \caption{Boxplot between Virtual Box and Docker with values normalized 
+in $0-1$ range.}
+        \label{chap:tests:sec:dockervsvb:img:boxplot}
+    \end{subfigure}
+    \caption[Virtual Box vs Docker start up comparison]{In these graphs it is 
+possible to see how the total time to start up a virtual machine through Virtual 
+Box is very high compared to the time required to the overall Docker containers 
+to boot. A single VirtualBox instance requires on average 20 seconds to boot up, 
+while a container requires, on average, only 0.02 seconds. The main reason 
+behind this difference, as already explained in other parts of this thesis, it 
+is caused by the guest's kernel that has to initialized itself, before starting 
+any program. On Docker, instead, the kernel is the same of the host, so the only 
+operations needed are to isolate the program from the rest of the OS and add it 
+to the process list.}
+    \label{chap:tests:sec:dockervsvb:subimg:plots}
 \end{figure}

### 2018-11-14 res/sections/25-Tests.tex
Update image, add next test draft

@@
 gathered in order to understand the real project effectiveness. We performed
-tests relatively two main sections: one testing the architecture 
-overall responsivness and one checking the SFC implementation and efficiency.
+tests relatively two main sections: one testing the architecture overall
+responsivness and one checking the SFC implementation and efficiency.
+
+\section{System responsivness}
+
+\subsection{Docker vs Virtual Box startup measurement}
 
@@
 Virtual Box one: Docker allows to inspect the container and to gain startup and
-shutdown times with a precision of millseconds, while to gather this 
-information for VirtualBox instances we had to use the command line utility 
-\verb!time!, that registered timing with a precision of hundredths of a 
-second. Nonetheless, this didn't influence the test too much. To avoid possible 
-outliers we perfomed the test $100$ times. With this $N$, the total time spent 
-by the Virtual Machines to boot up was of $2000$\todo{Update with the real 
-value} seconds, while with containers the same number of start ups resulted in a 
-total of $1.xx$\todo{Update with the real value}. % Insert graphs, explain the 
-%problems of virtual machines
+shutdown times with a precision of millseconds, while to gather this information
+for VirtualBox instances we had to use the command line utility \verb!time!,
+that registered timing with a precision of hundredths of a second. Nonetheless,
+this didn't influence the test too much. To avoid possible outliers we perfomed
+the test $100$ times. With this $N$, the total time spent by the Virtual
+Machines to boot up was of $2021,15$ seconds, while with containers the same
+number of start ups resulted in a total of $1.638$. Summing up, Docker employs
+only the $0,08\%$ of the overall time spent for the test, underlining how
+isolating the process with a set of defined policy and avoiding a virtualized
+kernel boot-up saves a considerable amount of time.
 
-\begin{figure}[t] % TODO fix image dimensions!
+\begin{figure}[t]
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{docker_vs_vmBarplotGraph}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[scale=0.4]{docker_vs_vmBarplotGraph}
         \caption{Barplot between Virtual Box and Docker startup times.}
@@
     \end{subfigure}
-    ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
-      %(or a blank line to force the subfigure onto a new line)
-    \begin{subfigure}[b]{0.3\textwidth}
-        \includegraphics[width=\textwidth]{docker_vs_vmBoxplotGraph}
+    ~
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[scale=0.35]{docker_vs_vmBoxplotGraph}
         \caption{Boxplot between Virtual Box and Docker with values normalized 
@@
 
-% Talk about the next test
+\subsection{SFC chain deploy measurement}
+
+The previous test make us see how much Docker is faster than a traditional
+virtual machine. The test, though, was performed on a single node, interacting
+directly with Docker and not through the orchestrator we used in the thesis
+development, in other words Kubernetes. The following measurement will allow us
+to see the time required in order to deploy an SFC of different elements in a
+Kubernetes environment. The deployment consisted in a Pod inside a Deployment
+definition with a Service related, to simulate a real VNF. After the Pod
+startup, a request is made to a backend designed to count the chain elements
+that have successfully boot-up that will return, eventually, the timing when all
+the VNFs are up and running.
+
+% Add SFC definition, graphs, talk about them

### 2018-11-15 res/sections/25-Tests.tex
Add listing, change some settings

@@
 to see the time required in order to deploy an SFC of different elements in a
-Kubernetes environment. The deployment consisted in a Pod inside a Deployment
-definition with a Service related, to simulate a real VNF. After the Pod
+Kubernetes environment. The deployment consist in a Pod inside a Deployment
+definition with a related Service, to simulate a real VNF. After the Pod
 startup, a request is made to a backend designed to count the chain elements
-that have successfully boot-up that will return, eventually, the timing when all
-the VNFs are up and running.
+that have successfully boot-up returning, eventually, the overall start up 
+timing.
 
+% TODO: missing YAML definition: [language=YAML]
+%https://tex.stackexchange.com/questions/152829/how-can-i-highlight-yaml-code-in
+%- a-pretty-way-with-listings
+\lstinputlisting{res/code/sfc_test.yaml}
+
+As is possible to see from the YAML definition, the test % talk about 
+%imagepull policy
 % Add SFC definition, graphs, talk about them

### 2018-11-15 res/sections/25-Tests.tex
Add new test description and graphs

@@
 ``boot sequence'' for both the VNFs: first we launched the Astaire framework, 
-ensuring it's availability to process data and then we notified the test 
+ensuring its availability to process data and then we notified the test 
 machine of the successful boot. The startup time was measured from the 
@@
 that registered timing with a precision of hundredths of a second. Nonetheless,
-this didn't influence the test too much. To avoid possible outliers we perfomed
+this didn't influence the test too much. To reduce possible outliers we perfomed
 the test $100$ times. With this $N$, the total time spent by the Virtual
 Machines to boot up was of $2021,15$ seconds, while with containers the same
-number of start ups resulted in a total of $1.638$. Summing up, Docker employs
-only the $0,08\%$ of the overall time spent for the test, underlining how
-isolating the process with a set of defined policy and avoiding a virtualized
-kernel boot-up saves a considerable amount of time.
+number of start ups resulted in a total of $1.638$ seconds. Summing up, Docker
+employs only the $0,08\%$ of the overall time spent for the test, underlining
+how isolating the process with a set of defined policy and avoiding a
+virtualized kernel boot-up saves a considerable amount of time.
 
-\begin{figure}[t]
+\begin{figure}[H]
     \centering
@@
         \includegraphics[scale=0.4]{docker_vs_vmBarplotGraph}
-        \caption{Barplot between Virtual Box and Docker startup times.}
+        \caption{Barplot between Virtual Box and Docker startup times in seconds
+          (y-axis).}
         \label{chap:tests:sec:dockervsvb:img:barplot}
@@
         \caption{Boxplot between Virtual Box and Docker with values normalized 
-in $0-1$ range.}
+in range $0-1$.}
         \label{chap:tests:sec:dockervsvb:img:boxplot}
@@
     \caption[Virtual Box vs Docker start up comparison]{In these graphs it is
-possible to see how the total time to start up a virtual machine through Virtual 
-Box is very high compared to the time required to the overall Docker containers 
-to boot. A single VirtualBox instance requires on average 20 seconds to boot up, 
-while a container requires, on average, only 0.02 seconds. The main reason 
-behind this difference, as already explained in other parts of this thesis, it 
-is caused by the guest's kernel that has to initialized itself, before starting 
-any program. On Docker, instead, the kernel is the same of the host, so the only 
-operations needed are to isolate the program from the rest of the OS and add it 
-to the process list.}
+      possible to see how the total time to start up a virtual machine through
+      Virtual Box is very high compared to the time required to the overall
+      Docker containers to boot. A single VirtualBox instance requires on
+      average 20,211 seconds to boot up, while a container requires, on average,
+      only 0.016 seconds. The main reason behind this difference, as already
+      explained in other parts of this thesis, it is caused by the guest's
+      kernel that has to initialized itself, before starting any program. On
+      Docker, instead, the kernel is the same of the host, so the only
+      operations needed are to isolate the program from the rest of the OS and
+      add it to the process list.}
     \label{chap:tests:sec:dockervsvb:subimg:plots}
 \end{figure}
+
+\begin{table}[t]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{l|c|c}
+                     & \textbf{Total start up time (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
+\textbf{Docker}      & 1,638                                 & 0,016                                     \\
+\textbf{Virtual Box} & 2021,150                              & 20,211                                 
+\end{tabular}%
+}
+\caption[Docker vs Virtual Box start up times comparison]{Docker vs Virtual Box
+  start up times comparison. Is possible to se how the overall time required to
+  start $100$ containers is less than the time to start one traditional virtual
+  machine. The overall time required by the traitional virtualization system is
+  approximately of 33 minutes, witch gives a significant insight of how much
+  time is required to create and boot a virtualized enviroment.}
+\label{my-label}
+\end{table}
+
+% The overall time required by the traitional virtualization system is
+% approximately of 33 minutes, witch gives a significant insight of how much time
+% is required to create and boot a virtualized enviroment.
 
@@
 directly with Docker and not through the orchestrator we used in the thesis
-development, in other words Kubernetes. The following measurement will allow us
-to see the time required in order to deploy an SFC of different elements in a
-Kubernetes environment. The deployment consist in a Pod inside a Deployment
-definition with a related Service, to simulate a real VNF. After the Pod
-startup, a request is made to a backend designed to count the chain elements
-that have successfully boot-up returning, eventually, the overall start up 
-timing.
+development, Kubernetes. The following measurement will allow us to see the time
+required in order to deploy an SFC of different elements in a Kubernetes
+environment. The deployment consists in a Pod inside a Deployment definition
+with a related Service, to simulate a real VNF. After the Pod startup, a request
+is made to a backend designed to count the chain elements that have successfully
+boot-up returning, eventually, the overall start up timing.
 
 % TODO: missing YAML definition: [language=YAML]
-%https://tex.stackexchange.com/questions/152829/how-can-i-highlight-yaml-code-in
+%https://tex.stackexchan
+%ge.com/questions/152829/how-can-i-highlight-yaml-code-in
 %- a-pretty-way-with-listings
-\lstinputlisting{res/code/sfc_test.yaml}
+\lstinputlisting[caption={Kubernetes YAML configuration that has been used to
+    simulate a single VNF deployment}, captionpos=b]{res/code/sfc_test.yaml}
 
-As is possible to see from the YAML definition, the test % talk about 
-%imagepull policy
-% Add SFC definition, graphs, talk about them
+\vspace{0.5cm}
+
+\noindent As is possible to see from the YAML definition, the test involved, for
+every launch, the VNF download from the Docker hub repositories. This, while it
+forced every node to delete and download the image for each test, it ensured
+consistent measurements in cases where the selected node for the deployment did
+not have the image already pulled. In addition to that, another consideration
+needs to be taken into account: when the chain consists of two or more elements,
+the container image which is used for the container deployment is the same for
+every VNF involved, thus reducing the start up time for the subsequent pods that
+reside in nodes that have already deployed one.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{sfc_startupBarplotGraph}
+  \caption[SFC start-up time barplot]{SFC start-up time barplot. It is
+    noticeable how there seems to be a linear correlation between the SFC
+    startup time and the SFC length, even though for the chain with eight
+    elements it is possible to see that the overall deploy time is more compared
+    to chains with nine and ten elements.}
+  \label{chap:tests:sec:sfclength:img:barplot}
+\end{figure}
+
+\begin{longtable}[c]{c|c|c}
+\textbf{SFC length} & \textbf{Total deploy time (seconds)} & \textbf{Average deploy time (seconds)} \\ \hline
+\endhead
+%
+\textbf{1}          & 1329,119                   & 13,291                       \\
+\textbf{2}          & 1420,693                   & 14,207                       \\
+\textbf{3}          & 1534,487                   & 15,345                       \\
+\textbf{4}          & 1645,298                   & 16,453                       \\
+\textbf{5}          & 1663,905                   & 16,634                       \\
+\textbf{6}          & 1741,825                   & 17,418                       \\
+\textbf{7}          & 1881,525                   & 18,815                       \\
+\textbf{8}          & 2085,622                   & 20,856                       \\
+\textbf{9}          & 2018,718                   & 20,187                       \\
+\textbf{10}         & 2047,645                   & 20,476                       \\
+\caption[SFC start up time]{Table of the results. It is possible to see how the
+  time required for the SFC to go up increases with the length, reaching an
+  average of 20 seconds for chain of eight, nine and ten elements.}
+\label{chap:tests:sec:sfclength:tab:sfcdata}\\
+\end{longtable}
+
+\newpage
+
+\paragraph*{Kubernetes specifications}
+A Kubernetes installation was performed in order to carry out the test above.
+All the machines (master node and minions) involved had the same specifications
+i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage. Nonetheless the constrained
+resources, absolutely far from a real Kubernetes production environment, is
+possible to notice how, even with a chain of ten elements, the average time
+required to deploy an SFC is equal, if not smaller, of the time required to
+deploy a Virtual Box machine on an instance with 32GB of RAM, 8 vCPU and 50GB of
+SSD storage.
+% graphs, talk about them

### 2018-11-15 res/sections/25-Tests.tex
Add other fixes

@@
 \paragraph*{Kubernetes specifications}
-A Kubernetes installation was performed in order to carry out the test above.
-All the machines (master node and minions) involved had the same specifications
-i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage. Nonetheless the constrained
-resources, absolutely far from a real Kubernetes production environment, is
-possible to notice how, even with a chain of ten elements, the average time
-required to deploy an SFC is equal, if not smaller, of the time required to
-deploy a Virtual Box machine on an instance with 32GB of RAM, 8 vCPU and 50GB of
-SSD storage.
+A Kubernetes installation was performed in order to carry out the test
+descripted above. All the machines (master node and minions) involved had the
+same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
+Nonetheless the constrained resources, absolutely far from a real Kubernetes
+production environment, is possible to notice how, even with a chain of ten
+elements, the average time required to deploy an SFC is equal, if not smaller,
+of the time required to deploy a Virtual Box machine on an instance with 32GB of
+RAM, 8 vCPU and 50GB of SSD storage.
 % graphs, talk about them
+
+\section{Final considerations}
+
+In these tests is possible to see one critical aspect concerning an architecture
+that follows the containerization approach: great flexibility in terms of
+deployments. This possibility to deploy services in the form of SFCs and VNFs in
+a very small amount of time is one of the key point required in the new 5G
+architecture, that the tests above demonstrated this project is able to adhere.
+Finally, the ability to rapidly redeploy software means smooth and easier
+software development lifecycle, where developers have the possibility to
+automatically push ``fresh'' software in production, letting the MANO and thus
+the container orchestrator handling all the difficult and delicate tasks to put
+software in a production environment, eliminating the possibility of human
+errors and assuring longer uptimes.

### 2018-11-15 res/sections/25-Tests.tex
Little fix

@@
 \noindent As is possible to see from the YAML definition, the test involved, for
-every launch, the VNF download from the Docker hub repositories. This, while it
-forced every node to delete and download the image for each test, it ensured
-consistent measurements in cases where the selected node for the deployment did
-not have the image already pulled. In addition to that, another consideration
-needs to be taken into account: when the chain consists of two or more elements,
-the container image which is used for the container deployment is the same for
-every VNF involved, thus reducing the start up time for the subsequent pods that
-reside in nodes that have already deployed one.
+every launch (that included and $N = 100$, for a total for 1000 tests), the VNF
+download from the Docker hub repositories. This, while it forced every node to
+delete and download the image for each test, it ensured consistent measurements
+in cases where the selected node for the deployment did not have the image
+already pulled. In addition to that, another consideration needs to be taken
+into account: when the chain consists of two or more elements, the container
+image which is used for the container deployment is the same for every VNF
+involved, thus reducing the start up time for the subsequent pods that reside in
+nodes that have already deployed one.
 

### 2018-11-15 res/sections/25-Tests.tex
Change tables with deviation standard, add fix graphs

@@
 \begin{tabular}{l|c|c}
-                     & \textbf{Total start up time (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
-\textbf{Docker}      & 1,638                                 & 0,016                                     \\
-\textbf{Virtual Box} & 2021,150                              & 20,211                                 
+                     & \textbf{Standard deviation (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
+\textbf{Docker}      & 0.009620874                           & 0,016                                    \\
+\textbf{Virtual Box} & 0.755                                 & 20,211                                 
 \end{tabular}%
@@
 \begin{longtable}[c]{c|c|c}
-\textbf{SFC length} & \textbf{Total deploy time (seconds)} & \textbf{Average deploy time (seconds)} \\ \hline
+\textbf{SFC length} & \textbf{Stamdard deviation (seconds)} & \textbf{Average deploy time (seconds)} \\ \hline
 \endhead
 %
-\textbf{1}          & 1329,119                   & 13,291                       \\
-\textbf{2}          & 1420,693                   & 14,207                       \\
-\textbf{3}          & 1534,487                   & 15,345                       \\
-\textbf{4}          & 1645,298                   & 16,453                       \\
-\textbf{5}          & 1663,905                   & 16,634                       \\
-\textbf{6}          & 1741,825                   & 17,418                       \\
-\textbf{7}          & 1881,525                   & 18,815                       \\
-\textbf{8}          & 2085,622                   & 20,856                       \\
-\textbf{9}          & 2018,718                   & 20,187                       \\
-\textbf{10}         & 2047,645                   & 20,476                       \\
+\textbf{1}          & 2.560102                   & 13,291                       \\
+\textbf{2}          & 0.6301444                  & 14,207                       \\
+\textbf{3}          & 1.14903                    & 15,345                       \\
+\textbf{4}          & 3.592179                   & 16,453                       \\
+\textbf{5}          & 0.9728132                  & 16,634                       \\
+\textbf{6}          & 1.103802                   & 17,418                       \\
+\textbf{7}          & 1.561451                   & 18,815                       \\
+\textbf{8}          & 0.7698537                  & 20,856                       \\
+\textbf{9}          & 1.889197                   & 20,187                       \\
+\textbf{10}         & 1.709227                   & 20,476                       \\
 \caption[SFC start up time]{Table of the results. It is possible to see how the

### 2018-11-16 res/sections/25-Tests.tex
Fix graph description
Minor english fixes

@@
 machine of the successful boot. The startup time was measured from the 
-beginning of the startup command (\verb!docker run! for docker and 
+beginning of the startup command (\verb!docker run! for Docker and 
 \verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
@@
         \includegraphics[scale=0.4]{docker_vs_vmBarplotGraph}
-        \caption{Barplot between Virtual Box and Docker startup times in seconds
-          (y-axis).}
+        \caption{Barplot between Virtual Box and Docker startup times.}
         \label{chap:tests:sec:dockervsvb:img:barplot}
@@
 \noindent As is possible to see from the YAML definition, the test involved, for
-every launch (that included and $N = 100$, for a total for 1000 tests), the VNF
-download from the Docker hub repositories. This, while it forced every node to
-delete and download the image for each test, it ensured consistent measurements
-in cases where the selected node for the deployment did not have the image
-already pulled. In addition to that, another consideration needs to be taken
-into account: when the chain consists of two or more elements, the container
-image which is used for the container deployment is the same for every VNF
-involved, thus reducing the start up time for the subsequent pods that reside in
-nodes that have already deployed one.
+every launch (that included and $N = 100$, for a total for 1000 tests), the 
+download of the container image from the Docker hub repositories. This, while 
+it forced every node to delete and download the image for each test, it ensured 
+consistent measurements in cases where the selected node for the deployment did 
+not have the image already pulled. In addition to that, another consideration 
+needs to be taken into account: when the chain consists of two or more 
+elements, the container image which is used for the container deployment is the 
+same for every VNF involved, thus reducing the start up time for the subsequent 
+pods that reside in nodes that have already deployed one.
 
@@
 
+\paragraph*{Kubernetes specifications}
+A Kubernetes installation was performed in order to carry out the test
+descripted above. All the machines involved (master node and minions) had the
+same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
+Nonetheless the constrained resources, absolutely far from a real Kubernetes
+production environment, it is possible to notice how, even with a chain of ten
+elements, the average time required to deploy an SFC is equal, if not smaller,
+of the average time required to deploy a Virtual Box machine on an instance 
+with 32GB of RAM, 8 vCPU and 50GB of SSD storage.
+
+% TODO is it worth to keep it as a long table?
 \begin{longtable}[c]{c|c|c}
-\textbf{SFC length} & \textbf{Stamdard deviation (seconds)} & \textbf{Average deploy time (seconds)} \\ \hline
+\textbf{SFC length} & \textbf{Stamdard deviation (seconds)} & \textbf{Average 
+deploy time (seconds)} \\ \hline
 \endhead
@@
 
-\newpage
-
-\paragraph*{Kubernetes specifications}
-A Kubernetes installation was performed in order to carry out the test
-descripted above. All the machines (master node and minions) involved had the
-same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
-Nonetheless the constrained resources, absolutely far from a real Kubernetes
-production environment, is possible to notice how, even with a chain of ten
-elements, the average time required to deploy an SFC is equal, if not smaller,
-of the time required to deploy a Virtual Box machine on an instance with 32GB of
-RAM, 8 vCPU and 50GB of SSD storage.
-% graphs, talk about them
-
 \section{Final considerations}
@@
 deployments. This possibility to deploy services in the form of SFCs and VNFs in
-a very small amount of time is one of the key point required in the new 5G
+a very small amount of time is one of the key points required in the new 5G
 architecture, that the tests above demonstrated this project is able to adhere.
@@
 software development lifecycle, where developers have the possibility to
-automatically push ``fresh'' software in production, letting the MANO and thus
+automatically push ``fresh'' software to production, letting the MANO and thus
 the container orchestrator handling all the difficult and delicate tasks to put

### 2018-11-16 res/sections/07-Introduction.tex
Add some part on SDN and on virtualization
Fix introductions errors

--- 2018-10-17 res/sections/07-Introduction.tex
@@
 3G, LTE) and the creation of common, ad-hoc versions, transmission protocols 
-ad-hoc versions (i.e. TCP versions specialized for exploit the characteristic of 
+(i.e. TCP versions specialized for exploit the characteristic of 
 wireless networks). It is for this reason that, beside improved radio 
@@
 to the requirements of the network. With this trend, in-hardware solutions 
-will not be able to keep up. Is with this considerations that software a 
-software approach is being proposed to become the new way to deploy network 
+will not be able to keep up. Is with this considerations that a software 
+approach is being proposed to become the new way to deploy network 
 components. While less efficent, they are faster to bring them up, offering an 

### 2018-11-16 res/sections/08-Background.tex
Fix little errors

--- 2018-11-11 res/sections/08-Background.tex
@@
 
-\section{Software-defined networking}
-The SDN is an emerging approach to cloud computing that allows to configure
-network in a programmatically way, possibly enhancing monitoring and
-performances. The main sake of SDN is to the packet forwarding process,
-\emph{data plane}, from the routing one, the so called \emph{control plane}.
-The latter component serves as the centralized network intelligence and could
-consists of one or more controllers. It has some drawbacks in teorms of
-security, scalability and elasticity, because of the centralized architecture.
-
-\subsection{SDN stack}
-\begin{figure}[t]
- \centering
- \includegraphics[scale=1.3]{sdn_architecture}
- \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
-denote the three layer logic: the Application plane, the Controller plane and 
-the Data plane.}
- \label{chap:background:img:sdn_architecture}
-\end{figure}
-As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be splitted in three different layers, called planes, based on 
-their functionalities.
-
-\subsubsection{Application plane}
-
-The \emph{Application plane} role is to set the behaviour of the SDN controllers 
-and manage its hardware resources and data through APIs, often called 
-\emph{northbound interfaces} (NBI).
-These interfaces provide an additional abstraction layer that high-level 
-applications can consume, enabling the possibility to use a variety of 
-different verdors' controllers in the same way.
-Analysing the SDN applications is possible to note two defined layers, one that 
-is responsible to manage the whole logic, and another one dedicated to the 
-communication part, that in particular uses the previsously defined NBIs.
-These applications can also build an abstracted view of the network, using
-controllers information, in order to make the traffic analysis easier and to 
-simplify the decision making process.
-
-\subsubsection{Controller plane}
-
-\emph{SDN controllers}, are logically centralized components that
-receive requests from the mentioned SDN applications and either translate them 
-in a fashion that can be understood from the lower layer\todo{who?}, or collect 
-data needed by the application itself. 
-Due to the variety of tasks a controller has to accomplish, its logical 
-suddivision is achieved with a three-layer architecture, in which every layer 
-exposes it's own interface, that can be consequently used by the upper 
-layers, up to the \emph{NBI agents}\todo{explain them?}.
-In particular, the middle layer represents the controller inner-logic, which 
-process the requests coming from the upper layer (aka the third layer), that 
-has the specific task to communicate with the remaining network compoents, in 
-other words the Control to Data-Plane Interface and/or the already 
-mentioned southbound interface.
-
-\subsubsection{Data plane}
-
-In the lowest part of the architecture we can find the \emph{SDN Networking 
-Devices}. These ones controls the packet forwarding and data processing 
-capabilities of the network. They consist of a CDPI agent and a set of one or 
-multiple traffic forwarding engines and of a zero or more traffic processing 
-functions, that are, as a matter of fact, the devices logic.
-
-\subsection{OpenFlow protocol}
-The SDN strategy is often associated with OpenFlow protocol: this communication
-protocol aims is to give access to the forwarding network component to the lower
-layer network infrastructure, made of switches and routers. OpenFlow enables
-the remote administration of networks working on layer 3: packets forwarding
-tables can be modified relying on packet matching rules. Following this
-approach it is possible to change routes periodically or ad-hoc, based on
-packet types, traffic or network load.
-
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.65]{openflow}
- \caption{OpenFlow protocol communication graph}
- \label{chap:background:img:openflow_protocol}
-\end{figure}
-
-In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
-how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
-set of tables to organized packet control and forwarding. These tables can be
-updated by higher-level application, usually an SDN application or controller,
-taking advatage of OpenFlow channels (called OF channels).
-
-\section{Virtualization}
-Cloud computing is now a leading-edge Internet technology and the virtualization
-technology is one of the most important part of cloud computing systems.
-Virtualization, in computing, refers to the act of creating a virtual (rather
-than an actual) version of something, including but not limited to a virtual
-computer hardware platform, operating system (OS), storage device, or computer
-network resources. \todo{To change I do not what can I write, stolen from ``The
-research and implementation of cloud computing platform based on docker'' paper}
-
-\subsection{Network Virtualization}
-%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
-%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
-%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
-%r k _virtualization.pdf
-Virtualization, in the networking context, refers to the act of combining
-hardware, software and functionalities into a virtual network. This abstraction
-runs on the top of a physical network in a hypervisor, decupling the logical
-network from the harwdare one. Network virtualization allow multiple,
-heterogeneous networks to cohabit, increasing the flexibility and manageability.
-%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
-%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
-The idea of network virtualization came from the decoupling of the role of the
-traditional ISP into two different entities: who provides the infrastucture
-(InPs) and who provides the services(SPs). In that context it is possible to
-aggregate resources, even from different InPs, creating a virtual network,
-offering an end-to-end service. Most known instances of NV are \emph{VPNs},
-dedicated networks that allows the communication among multiple sites, using
-private and secure tunnels, relying on the public network such as Internet.
-Other examples of netvork virtualization are \emph{overlay networks}. They are
-logical networks, build on the top of an existing physical network, tipically
-implemented at the application layer.
-
-\subsection{Network Function Virtualization}
-The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
-virtualization tecnologies to virtualize network nodes functions. VNFs may run 
-different processes and softwares, on the top of servers, switches, storage 
-devices or on cloud computing infrastuctures, that substitute custom hardware. 
-That approach worked well in the past, because of the need to follow rigorous 
-standards, providing at the same time high quality, but it led to long product 
-cycles and high specialized proprietary hardware. New Internet senarios, that 
-include IoT and the increasing number of services to spread multimedia data 
-(Skype, Netflix, Spotify, ...), requires more flexible functions and faster 
-network deployments. The NFV technology is exploited in a framework, organized
-at least by three components:
-\begin{enumerate}
- \item a software that performs the network function that must be deployed,
- that is the VNF;
- \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
-components that builds the environment where the VNFs are deployed;
- \item NFV Management And Orchestration (NFV-MANO) architectural framework that
- is all the functional blocks, data repositories used by these blocks,
- reference points and interfaces through which these functional blocks exchange
- information for the purpose of managing and orchestrating NFVI and VNFs.
-\end{enumerate}
-
 \section{The upcoming connectivity standard: 5G}
@@
 connectivity, the 5G aims to offer not only at the end-user a new way to browse 
-the web,
-download and watch interactive content but also to create an ad-hoc solution 
-for Machine-to-Machine (M2M) data traffic, which is increasing more than ever 
-thanks to the spreading of IoT devices, which sensors need to continuosly send 
-data to servers/datacenters. Relatively to LTE\todo{Explain acronym}, 5G points 
-to have data rates $10$ times better, with $10$ times smaller end-to-end 
-latency and an increased connection densisty by $100$ times\todo{CITE}.
+the web, download and watch interactive content but also to create an ad-hoc 
+solution for Machine-to-Machine (M2M) data traffic, which is increasing more 
+than ever thanks to the spreading of IoT devices, which sensors need to 
+continuosly send data to servers/datacenters. Relatively to LTE\todo{Explain 
+acronym}, 5G points to have data rates $10$ times better, with $10$ times 
+smaller end-to-end latency and an increased connection densisty by $100$ 
+times\todo{CITE}.
 
@@
 
-This separation in layers allows to easly manage different identities 
-differently: an Internet Service Provider (ISP) could manage different physical
-infrastructure and have multiple business emblements dislocated 
+This separation in layers allows to easly manage mutiple identities 
+differently: an Internet Service Provider (ISP) could manage a multitude of 
+physical infrastructures and have multiple business emblements dislocated 
 along an entire continet for example, but it could decide to have only a single 
@@
 even if we support a plethora of protocols.''} making the whole architecture 
-indipendent from the protocol a particular flux of data. To achieve this, 
+indipendent from the protocol a particolar flux of data. To achieve this, 
 several solutions have been studied, and at the end packet incapsulation with 

### 2018-11-16 res/sections/08-Background.tex
Grammar
Co-Authored-By: mzanella <mzanella@users.noreply.github.com>

--- 2018-11-12 res/sections/08-Background.tex
@@
 that usually provide an abstract view of the network and a set of
-functionalities of the controller.
+the controller functionalities.
 In particular, the middle layer represents the controller inner-logic, which 

### 2018-11-16 res/sections/08-Background.tex
Grammar
Co-Authored-By: mzanella <mzanella@users.noreply.github.com>

@@
 layers, up to the \emph{NBI agents}.
-NBI agents manages the communication with SDN applications, exposing interfaces
+NBI agent manages the communications with SDN applications, exposing interfaces
 that usually provide an abstract view of the network and a set of

### 2018-11-16 res/sections/25-Tests.tex
Fix tests with new ones
Add RTT tests

@@
 errors and assuring longer uptimes.
+
+\section{System performances}
+
+\subsection{Test system}
+All the machine used for testing purposes are Openstack virtual machines, with the following specifications:
+\begin{itemize}
+\item 1 machine with 32GB of RAM, 8vCPU and solid state storage, on which the receiver of packets is running;
+\item 4 machines with 16GB of RAM, 6vCPU and solid state storage, that form a Kubernetes cluster on which SFC infrastructure is deployed;
+\item 1 machine with 4GB of RAM, 2vCPU and solid state storage, on which scripts to send packets run.
+\end{itemize}
+
+\subsection{RTT measurement}
+Regarding the delivery performance, a metric to take in account is the \emph{round trip time}(RTT), that is the time interval between the instant in whic a packet is sent and the response is received.
+To test it there was developed a script that sends a packet and waits for the corrispondant response, recording the instants (expressed as \emph{epoch}) of send and receive.
+The sake of thi test is to compare the RTT measured without any SFC (only sending the packet to the receiver waiting for the response) and making packets pass through different length SFCs, in particular with 1, 2, 3, 4, 5, and 10 hops.
+Packets traverse the SFC on both sides and the results are collected after 100 tests and the SF only prints the packet received and the total number of packets that have passed throught it.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{@{}cc@{}}
+\toprule
+\textbf{SFC length} & \textbf{Mean RTT (milliseconds)} \\ \midrule
+len. 0              & 0.0009668279      \\
+len. 1              & 0.0555884027      \\
+len. 2              & 0.0702438903      \\
+len. 3              & 0.0888472247      \\
+len. 4              & 0.1045638967      \\
+len. 5              & 0.1239343858      \\
+len. 10             & 0.2135494161      \\ \bottomrule
+\end{tabular}
+\caption{Mean RTT measured with different SFC length}
+\label{chap:tests:sec:rtt:tab:meanrtt}
+\end{table}
+
+Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
+the RTT increase with the number of elements of the chain.
+There is a big difference between the mean RTT of the configuration wiothout any chain and the one with an SFC with a single link: this outcome has different explaination.
+As explained in \todo{Add reference to implementation} the chain has always at least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3 different running software.
+Another explaination lays into how the code was developed: it was not developed as a production ready product, but for a feasibility test, so connections and trasmissions can be enhanced.
+Finally, SFs are meant to perform some modifications to the packet, so, unavoidably, some time is waisted.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.35]{meanrtt}
+  \caption{Mean RTT calculate using SFCs with different lengths}
+  \label{chap:tests:sec:rtt:img:meanstt}
+\end{figure}
+
+The results are better represented in Figure~\ref{chap:tests:sec:rtt:img:meanstt}.
+In that graph we can easier see the difference among the chains with different lengths and the trasmission that does not use any SFC.
+Looking at the bars representing the SFC composed by 1, 2, 3, 4 and 5 links we can see a pretty linear increase of the RTT.
+This trend strongly depends on the nature of the SF involved in the SFC.

### 2018-11-16 res/sections/08-Background.tex
Fix grammar

@@
 \subsection{OpenFlow protocol}
-The SDN strategy is often associated with OpenFlow protocol: this communication
-protocol aims is to give access to the forwarding network component to the lower
-layer network infrastructure, made of switches and routers. OpenFlow enables
-the remote administration of networks working on layer 3: packets forwarding
-tables can be modified relying on packet matching rules. Following this
-approach it is possible to change routes periodically or ad-hoc, based on
-packet types, traffic or network load.
+The SDN strategy is often associated with the OpenFlow protocol: this 
+communication protocol aims is to give access to the forwarding network 
+component to the lower layer network infrastructure, made of switches and 
+routers. OpenFlow enables the remote administration of networks working on 
+layer 3: packets forwarding tables can be modified relying on packet matching 
+rules. Following this approach it is possible to change routes periodically or 
+ad-hoc, based on packet types, traffic or network load.
 

### 2018-11-16 res/sections/08-Background.tex

@@
 creation of a new standard, the 5G, which is estimated to arrive to the market
-constumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
-alliace, a group composed by the major players in the field of mobile
+consumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
+alliance, a group composed by the major players in the field of mobile
 connectivity, the 5G aims to offer not only at the end-user a new way to browse 
-the web,
-download and watch interactive content but also to create an ad-hoc solution 
-for Machine-to-Machine (M2M) data traffic, which is increasing more than ever 
-thanks to the spreading of IoT devices, which sensors need to continuosly send 
-data to servers/datacenters. Relatively to LTE\todo{Explain acronym}, 5G points 
-to have data rates $10$ times better, with $10$ times smaller end-to-end 
-latency and an increased connection densisty by $100$ times\todo{CITE}.
+the web, download and watch interactive content but also to create an ad-hoc 
+solution for Machine-to-Machine (M2M) data traffic, which is increasing more 
+than ever thanks to the spreading of IoT devices, which sensors need to 
+continuously send data to servers/data-centers. Relatively to LTE\todo{Explain 
+acronym}, 5G points to have data rates $10$ times better, with $10$ times 
+smaller end-to-end latency and an increased connection density by $100$ 
+times\todo{CITE}.
 
@@
 \caption[5G improvement over LTE]{An extract from the official 5G white paper 
-illustating the improvements of 5G relatively to LTE connections.}
+illustrating the improvements of 5G relatively to LTE connections.}
 \label{chap:intro:table:ltevs5g}
@@
 higher frequencies. The utilization of higher frequencies, though, mean that the
-radio signals can be easly disrupted by physical objects, like buildings and
+radio signals can be easily disrupted by physical objects, like buildings and
 many geographical elements (such as hills and mountains), clashing with the
@@
 plays an important role. In fact, the re-design of some network components today
-existing via hardware can transform a monolitic networking approach to a modular
-one, exploting the flexibility that Virtual Network Functions (VNFs)\todo{CITE
-  VNF-P: A Model for Efficient Placement of Virtualized Network Functions} can
-offer thanks to virtualization, closing the gap to the use-case fullfilment
-defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture} which 
-require a descreased time to set up and deploy network services (specificatly, 
-from 90 hours to 90 minutes\todo{Search this requirement, check 90h 'cause 
-I'm not sure about it}).
+existing via hardware can transform a monolithic networking approach to a 
+modular one, exploiting the flexibility that Virtual Network Functions 
+(VNFs)\todo{CITE VNF-P: A Model for Efficient Placement of Virtualized Network 
+Functions} can offer thanks to virtualization, closing the gap to the use-case 
+fulfillment defined by the NGMN alliance\todo{CITE: NGMN View on 5G 
+architecture} which require a decreased time to set up and deploy network 
+services (specifically, from 90 hours to 90 minutes\todo{Search this 
+requirement, check 90h 'cause I'm not sure about it}).
 
@@
 \begin{itemize}
-\item \textbf{infrastructure resouce layer}: physical resources that are exposed
-  via a virtualized interface, and that can be monitored using specific APIs
+\item \textbf{infrastructure resource layer}: physical resources that are 
+exposed via a virtualized interface, and that can be monitored using specific 
+APIs
 \item \textbf{business enablement layer}: where a library of functions and
@@
 
-This separation in layers allows to easly manage different identities 
-differently: an Internet Service Provider (ISP) could manage different physical
-infrastructure and have multiple business emblements dislocated 
-along an entire continet for example, but it could decide to have only a single 
-centralized business application deployment that manage all the other layers 
-resources.
+This separation in layers allows to easily manage multiple identities 
+differently: an Internet Service Provider (ISP) could manage a multitude of 
+physical infrastructures and have multiple business enablements dislocated 
+along an entire continent for example, but it could decide to have only a 
+single centralized business application deployment that manage all the other 
+layers resources.
 
 \subsubsection{Network slicing}
-The role of a ``network slice'' in a 5G architecture is to specificatly handle 
-the Control-plane\todo{Explain what a control plane is?} of a praticular 
+The role of a ``network slice'' in a 5G architecture is to specifically handle 
+the Control-plane\todo{Explain what a control plane is?} of a particular 
 service (e.g. smartphones traffic, autonomous driving, massive IoT), deploying 
@@
  technology and use the NFV-MANO architecture to perform first packets
- elaboration and performance improvment and finally TCP/IP satellite chunk
+ elaboration and performance improvement and finally TCP/IP satellite chunk
  optimization with the Performance Enhancing Proxy (PEP). The VIBES project
@@
  \item VNF-PEP validation and performance tests on 5G use-cases
- \item Demonstration testbed management
+ \item Demonstration test-bed management
 \end{enumerate}
@@
 different protocols that would become infeasible to support all of them at the
-same time, packet incapsulation present itself as the only faseable solution:
+same time, packet encapsulation present itself as the only feasible solution:
 every packet incoming in the VNFs has already been encapsulated by a generic
-packet incapsulator/decapsulator,
+packet encapsulator/decapsulator,
 \todo{Also I think that the main purpose of UDP encapsulation is not to hide 
-the protocol used on the edges but 1) usa a ``quick protocol to exchange data 
+the protocol used on the edges but 1) use a ``quick protocol to exchange data 
 among VNFs and 2) we are hiding the path not the protocol itself. As we were 
@@
 even if we support a plethora of protocols.''} making the whole architecture 
-indipendent from the protocol a particular flux of data. To achieve this, 
-several solutions have been studied, and at the end packet incapsulation with 
-TCP split (for TCP sessions) have been choosen. The rationale that guided us on 
+independent from the protocol a particular flux of data. To achieve this, 
+several solutions have been studied, and at the end packet encapsulation with 
+TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
 this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
@@
 Since the requirement for the whole system (MANO+PEP) were too challenging, the 
-goal shifted into creating a MANO testbed and a working NFVI, excluding PEP. 
+goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
 The VNF architecture was shaped following the container orchestrator we decided 
@@
 packets through VNF functions, we encountered that many networking tools already
-present in the market required some tweaking and some integrations, shifting our
+present in the market required some tweaking and some integration, shifting our
 goal to create a complete European Telecommunications Standards Institutes
-(ETSI) Management and orchestrator (MANO) testbed instead, following the
+(ETSI) Management and orchestrator (MANO) test-bed instead, following the
 specifications suggested in the RFC 7665, thus implementing only the first three
@@
 particular, we discovered how, these tools, were suitable to create ETSI MANO
-and VNFs using virtual machine or exploiting cloud tecnhologies, while they were
+and VNFs using virtual machine or exploiting cloud technologies, while they were
 not designed with enough flexibility an integration with Docker. In the next

### 2018-11-16 res/sections/07-Introduction.tex

@@
  Through the history of internet, many factors changed and new innovations came,
- but at the end of the day a foundamental key remained: connection performances
+ but at the end of the day a fundamental key remained: connection performances
  are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
-   performes quite well in wired connections, while in wireless one, such as
+   performs quite well in wired connections, while in wireless one, such as
    wireless/satellite, sucks ass}
@@
 wireless networks). It is for this reason that, beside improved radio 
-communications, backend infrastructure began to be refined too. During their 
+communications, back-end infrastructure began to be refined too. During their 
 communication to the destination packets are elaborated, compressed and 
@@
 backbones having more and more data to deliver, making in-hardware backbone 
-solutions less efficent. Their time-to-deploy, in fact, is very high compared 
+solutions less efficient. Their time-to-deploy, in fact, is very high compared 
 to the requirements of the network. With this trend, in-hardware solutions 
@@
 approach is being proposed to become the new way to deploy network 
-components. While less efficent, they are faster to bring them up, offering an 
+components. While less efficient, they are faster to bring them up, offering an 
 unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network 
@@
  environments, where new frameworks, called orchestrators (e.g. Kubernetes,
- Docker Swarm), allow to easily manage the application lifecyle, i.e.
+ Docker Swarm), allow to easily manage the application lifecycle, i.e.
  deployment, scale up/down and finally, removal.

### 2018-11-16 res/sections/08-Background.tex

@@
 going to briefly introduce and explain the technologies involved and their role.
-
-\section{Software-defined networking}
-The SDN is an emerging approach to cloud computing that allows to configure
-network in a programmatically way, possibly enhancing monitoring and
-performances. The main sake of SDN is to the packet forwarding process,
-\emph{data plane}, from the routing one, the so called \emph{control plane}.
-The latter component serves as the centralized network intelligence and could
-consists of one or more controllers. It has some drawbacks in teorms of
-security, scalability and elasticity, because of the centralized architecture.
-
-\subsection{SDN stack}
-\begin{figure}[t]
- \centering
- \includegraphics[scale=1.3]{sdn_architecture}
- \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
-denote the three layer logic: the Application plane, the Controller plane and 
-the Data plane.}
- \label{chap:background:img:sdn_architecture}
-\end{figure}
-As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be splitted in three different layers, called planes, based on 
-their functionalities.
-
-\subsubsection{Application plane}
-
-The \emph{Application plane} role is to set the behaviour of the SDN controllers 
-and manage its hardware resources and data through APIs, often called 
-\emph{northbound interfaces} (NBI).
-These interfaces provide an additional abstraction layer that high-level 
-applications can consume, enabling the possibility to use a variety of 
-different verdors' controllers in the same way.
-Analysing the SDN applications is possible to note two defined layers, one that 
-is responsible to manage the whole logic, and another one dedicated to the 
-communication part, that in particular uses the previsously defined NBIs.
-These applications can also build an abstracted view of the network, using
-controllers information, in order to make the traffic analysis easier and to 
-simplify the decision making process.
-
-\subsubsection{Controller plane}
-
-\emph{SDN controllers}, are logically centralized components that
-receive requests from the mentioned SDN applications and either translate them 
-in a fashion that can be understood from the lower layer\todo{who?}, or collect 
-data needed by the application itself. 
-Due to the variety of tasks a controller has to accomplish, its logical 
-suddivision is achieved with a three-layer architecture, in which every layer 
-exposes it's own interface, that can be consequently used by the upper 
-layers, up to the \emph{NBI agents}.
-NBI agent manages the communications with SDN applications, exposing interfaces
-that usually provide an abstract view of the network and a set of
-the controller functionalities.
-In particular, the middle layer represents the controller inner-logic, which 
-process the requests coming from the upper layer (aka the third layer), that 
-has the specific task to communicate with the remaining network compoents, in 
-other words the Control to Data-Plane Interface and/or the already 
-mentioned southbound interface.
-
-\subsubsection{Data plane}
-
-In the lowest part of the architecture we can find the \emph{SDN Networking 
-Devices}. These ones controls the packet forwarding and data processing 
-capabilities of the network. They consist of a CDPI agent and a set of one or 
-multiple traffic forwarding engines and of a zero or more traffic processing 
-functions, that are, as a matter of fact, the devices logic.
-
-\subsection{OpenFlow protocol}
-The SDN strategy is often associated with the OpenFlow protocol: this 
-communication protocol aims is to give access to the forwarding network 
-component to the lower layer network infrastructure, made of switches and 
-routers. OpenFlow enables the remote administration of networks working on 
-layer 3: packets forwarding tables can be modified relying on packet matching 
-rules. Following this approach it is possible to change routes periodically or 
-ad-hoc, based on packet types, traffic or network load.
-
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.65]{openflow}
- \caption{OpenFlow protocol communication graph}
- \label{chap:background:img:openflow_protocol}
-\end{figure}
-
-In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
-how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
-set of tables to organized packet control and forwarding. These tables can be
-updated by higher-level application, usually an SDN application or controller,
-taking advatage of OpenFlow channels (called OF channels).
-
-\section{Virtualization}
-Cloud computing is now a leading-edge Internet technology and the virtualization
-technology is one of the most important part of cloud computing systems.
-Virtualization, in computing, refers to the act of creating a virtual (rather
-than an actual) version of something, including but not limited to a virtual
-computer hardware platform, operating system (OS), storage device, or computer
-network resources. \todo{This part was taken from ``The research and
-implementation of cloud computing platform based on docker'' paper, rework it
-in a way it doesn't seem to be copied.}
-
-\subsection{Network Virtualization}
-%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
-%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
-%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
-%r k _virtualization.pdf
-Virtualization, in the networking context, refers to the act of combining
-hardware, software and functionalities into a virtual network. This abstraction
-runs on the top of a physical network in a hypervisor, decupling the logical
-network from the harwdare one. Network virtualization allow multiple,
-heterogeneous networks to cohabit, increasing the flexibility and manageability.
-%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
-%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
-The idea of network virtualization came from the decoupling of the role of the
-traditional ISP into two different entities: who provides the infrastucture
-(InPs) and who provides the services(SPs). In that context it is possible to
-aggregate resources, even from different InPs, creating a virtual network,
-offering an end-to-end service. Most known instances of NV are \emph{VPNs},
-dedicated networks that allows the communication among multiple sites, using
-private and secure tunnels, relying on the public network such as Internet.
-Other examples of netvork virtualization are \emph{overlay networks}. They are
-logical networks, build on the top of an existing physical network, tipically
-implemented at the application layer.
-
-\subsection{Network Function Virtualization}
-The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
-virtualization tecnologies to virtualize network nodes functions. VNFs may run 
-different processes and softwares, on the top of servers, switches, storage 
-devices or on cloud computing infrastuctures, that substitute custom hardware. 
-That approach worked well in the past, because of the need to follow rigorous 
-standards, providing at the same time high quality, but it led to long product 
-cycles and high specialized proprietary hardware. New Internet senarios, that 
-include IoT and the increasing number of services to spread multimedia data 
-(Skype, Netflix, Spotify, ...), requires more flexible functions and faster 
-network deployments. The NFV technology is exploited in a framework, organized
-at least by three components:
-\begin{enumerate}
- \item a software that performs the network function that must be deployed,
- that is the VNF;
- \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
-components that builds the environment where the VNFs are deployed;
- \item NFV Management And Orchestration (NFV-MANO) architectural framework that
- is all the functional blocks, data repositories used by these blocks,
- reference points and interfaces through which these functional blocks exchange
- information for the purpose of managing and orchestrating NFVI and VNFs.
-\end{enumerate}
 

### 2018-11-16 res/sections/10-ProjectAnalysis.tex

@@
 integration of multiple tools (some of them have been already briefly introduced
-in~\ref{chap:intro}). The tools were choosen after their analysis, where pro and
+in~\ref{chap:intro}). The tools were chosen after their analysis, where pro and
 cons have been studied and compared to the necessity of this thesis. Since the
@@
 
-During the testbed development the software architecture diverged from the
+During the test-bed development the software architecture diverged from the
 original one, experiencing different evolutions and modifications.
@@
 technologies, is possible to identify different macro-components, each of these
-specialized in fullfilling a specific goal. For the ETSI standards, the Network
+specialized in fulfilling a specific goal. For the ETSI standards, the Network
 Function Virtualization (NFV) design should be composed as follows:
@@
 \begin{itemize} 
- \item \textbf{Hypervisor domain} that is composed of the Virtual Compute, the
+ \item \textbf{Hyper-visor domain} that is composed of the Virtual Compute, the
    Virtual Storage, and the Virtualization Layer. These components are used to
@@
 In conventional virtualization systems these resources are usually separated
-from the host operating system, providing better isolation but wrost
+from the host operating system, providing better isolation but worst
 performance. The ETSI infrastructure, instead, tries with container
@@
   term - I don't think it actually exists}, removing the additional operating
-system (called ``guest'' OS) required by the usual hypervisor-oriented approach.
+system (called ``guest'' OS) required by the usual hyper-visor-oriented 
+approach.
 
@@
 
-We performed an accured analysis of the tools choosen based on their maturity,
+We performed an accurate analysis of the tools chosen based on their maturity,
 on the community and on the technical support offered (e.g. user manuals,
-developer documentation), which, at the end, commited us to choose different
+developer documentation), which, at the end, committed us to choose different
 tools from the suggested ones. A detailed description of the architectural
@@
   manipulation. This components are design to be lightweight services able to
-  elaborate a great amount of packets per seconds. Manupulation can be of two
+  elaborate a great amount of packets per seconds. Manipulation can be of two
   typologies: stateful or stateless. In the first case, a state is maintained,
@@
   orchestration of network services (NS), also called SFC. The MANO has
-  different responsabilities, such as managing the
-  istantiation/destruction/modification of the SFCs, that achieves interacting
+  different responsibilities, such as managing the
+  creation/destruction/modification of the SFCs, that achieves interacting
   with the VIM.\todo{CITE: SDN/NFV-Based Mobile Packet Core Network
-Architectures: A Survey}. Finally, the mano can integrate with 
+Architectures: A Survey}. Finally, the MANO can integrate with 
 Operational and Business Support Systems (OSS/BSS), \todo{Expand this part with 
@@
 scalability and provide methods to recovery from failures. With this challenges 
-packet transmission revealed to be a foundamental point to choose carefully.
+packet transmission revealed to be a fundamental point to choose carefully.
 
@@
 
-When a packets are sent from a seder $S$, to a receiver $R$, they start
+When a packets are sent from a sender $S$, to a receiver $R$, they start
 travelling to the backbone receiving appropriate in-hardware packet processing
@@
 aware of the packets elaboration at all. To perform that, many solutions are
-available: TUN/TAP packet incapsulation, IP Encapsulation within IP or TCP
+available: TUN/TAP packet encapsulation, IP Encapsulation within IP or TCP
 session recreation (know also as TCP split).
@@
 aim to deliver IP packets over Mobile IP for mobile networks) maintain the
-connection transparency incapsulating the packet in another one.
+connection transparency encapsulating the packet in another one.
 \begin{figure}[t]
@@
     Encapsulation within IP. Is noticeable how the original packet becomes the
-    payload of a new one, thus being encapsulated. All the informations
+    payload of a new one, thus being encapsulated. All the information
     regarding the original packet remains untouched: when this kind of packet
@@
     payload: a IP packet. Packets encapsulation can be applied with higher
-    layers too, thus allowing for TCP and UDP packet incapsulation.}
+    layers too, thus allowing for TCP and UDP packet encapsulation.}
   \label{chap:prjan:img:ip_over_ip}
@@
 that communicate with the packet receiver creating a multi-overlay-hop path
-where for each hop there is an possible indipendent TCP connection. In order to
+where for each hop there is an possible independent TCP connection. In order to
 achieve that, a session table needs to be established and continuously updated,
@@
 packets need to follow the same (or an equivalent one) path for the same
-instatiated connection. This phenomenom is based on the intermediate proxies
+instantiated connection. This phenomenon is based on the intermediate proxies
 that create the route: it is the case, for example, when an intermediate node
 sends back a spoofed ACK before the original packet is delivered to the real
-consegnee. In this senario, the sender, unaware of the proxy, will miscalculate
+consignee. In this scenario, the sender, unaware of the proxy, will miscalculate
 the sending window based on the too low RTT. Other problems of TCP splitting are
@@
 that all the packets have reached the destination. TCP session recreation seems
-burdersome and it doesn't seem to add any significant benefit at the first
+burdensome and it doesn't seem to add any significant benefit at the first
 sight, but it offers the flexibility to perform extensive packet manipulation:
@@
 
-During the testbed development we first tried the TUN/TAP solution, that
+During the test-bed development we first tried the TUN/TAP solution, that
 revealed to be not completely transparent and with major performance drawbacks
@@
 Before starting any effective work on the project it has been decided to 
-perform an analysis of the possible technologies avvailable in the market, 
-performing a choice between them and exploiting their possibile ``weak 
-points''. Proceding with a top-down approach, the frameworks were choosen from 
-the one that had the biggest role in the testbed to conclude with the 
+perform an analysis of the possible technologies available in the market, 
+performing a choice between them and exploiting their possible ``weak 
+points''. Proceeding with a top-down approach, the frameworks were chosen from 
+the one that had the biggest role in the test-bed to conclude with the 
 less-important components.
@@
 allocating containerized applications on cloud platforms, we decided to start 
-analizing the right framework for this component first.
+analyzing the right framework for this component first.
 
@@
 Created in ??\todo{Find out Openstack date of creation}, Openstack allows
-on-permise cloud installations, using bare-metal resources to provide common
+on-premise cloud installations, using bare-metal resources to provide common
 cloud services as object storage, virtual machine deploy, virtual networking.
@@
 machines (called nodes from now on) to create an abstraction layer and to form a
-cluster where is possible to run Docker containers without brothering about
+cluster where is possible to run Docker containers without bothering about
 hardware constraints. Nodes can have specific roles and specialization, all of
@@
 duplication of loaded libraries and assets in the memory of the host. For large
-deployment containing only simple services (e.g. a backend application serving a
+deployment containing only simple services (e.g. a back-end application serving 
+a
 website or a database) this ends up in a waste of resources. \todo{Talk about
@@
 
-% The development of the ETSI MANO testbed involved the use of several 
+% The development of the ETSI MANO test-bed involved the use of several 
 % tecnhologies already in the market, and in some cases the creation of ad-hoc 
@@
 % well-estabilished technologies used is performed, describing even solutions 
-% that at the end weren't choosen because they didn't fit with the defined goals.
+% that at the end weren't chosen because they didn't fit with the defined goals.
 
@@
 
-Another foundamental component in the architecture is the MANO, that has to 
+Another fundamental component in the architecture is the MANO, that has to 
 manage the Network Services and the deployment of new VNFs. In coordination 
-with Kubernetes, it composes the core of the testbed architecture. Initially 
+with Kubernetes, it composes the core of the test-bed architecture. Initially 
 identified with Openbaton (the component is described later 
 in~\ref{chap:prjan:sec:tech:sub:other:sub:openbaton}), afterword it has been 
-discovered to be not fexible enough to support other kinds of virtualization 
-tecniques than pure or in-cloud virtual machines\todo{AAAAAA rewrite this}. 
+discovered to be not flexible enough to support other kinds of virtualization 
+techniques than pure or in-cloud virtual machines\todo{Rewrite this}. 
 Since this was a strong project requisite that could not be easily changed, it 
@@
 CRUD operations. It offers functionalities similar to Openbaton: possibility to
-save and retrieve VNF and NS definitions (basically, Kuberentes YAML
+save and retrieve VNF and NS definitions (basically, Kubernetes YAML
 deployments). When an NS is saved, a consistency check is performed to assure
@@
 orphan (which means without any NS using it). In that case the VNF will enter in
-a state of pruning: after five minutes, if the VNF has not be reutilized by any
+a state of pruning: after five minutes, if the VNF has not be re-utilized by any
 other NS it will be deleted. This mechanism assure faster deployments in
 environments where NSs are continuously taken up and down. \todo{I've made an
-  Harbor implementation schema, I dunno if its usefult to put it here}
+  Harbor implementation schema, I dunno if its useful to put it here}
 
@@
 The tools analysis revealed little about already existent VNF or NS
-implementations, so it was neccesary to fill the gap. Since this one of our
-thesis goals, we devepled a complete solution to create a complete SFCs, trying
+implementations, so it was necessary to fill the gap. Since this one of our
+thesis goals, we developed a complete solution to create a complete SFCs, trying
 to standardize our vocabulary as much as possible, and giving basic
@@
 \label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
-Dynamic communications between VNFs would not be possible without a thrid
+Dynamic communications between VNFs would not be possible without a third
 component acting as a route controller. Roulette fulfills this role, providing a
-backend where the single VNF can ask the next packet hop. Indeed, routes must be
+back-end where the single VNF can ask the next packet hop. Indeed, routes must 
+be
 submitted by another component, that we identified as Harbor (component
@@
 but it has to manage the communications and the related forward operations. To 
-possibly ease part of the transmission problem and to make it easier to create a 
+possibly ease part of the transmission problem and to make it easier to create 
+a 
 VNF without requiring the whole knowledge about SFC, it has been decided to 
@@
 to be called dynamically at runtime in order to process the packet. The result 
-of this elaboration gets returned to astaire, that performs the beforehand 
+of this elaboration gets returned to Astaire, that performs the beforehand 
 cited operations of VNFs communications and forwarding.
@@
 a client $S$ performs a request to $R$, the packet actually goes through 
-Ironhide, who has an internal classifier that it is able to recognise the type 
+Ironhide, who has an internal classifier that it is able to recognize the type 
 of connection and to choose the right NS the packet has to go through. This 
-process in foundamental since it keeps the connection between $S$ and $R$ up, 
+process in fundamental since it keeps the connection between $S$ and $R$ up, 
 for both TCP and UDP based communications.
@@
 
-Since the testbed is composed by different components and thanks to its 
+Since the test-bed is composed by different components and thanks to its 
 modularity, the implementation of the single components has been done with the 
@@
 of the tasks that should be performed by hand when launching one or multiple 
-docker containers. It's particulary useful when more continers have to interact 
+docker containers. It's particularly useful when more containers have to 
+interact 
 together, because\todo{Write down docker-compose functionalities and describe 
 the product a little bit} all the services rules are described in a simple 
-YAML file. With the birth of Docker Swarm, work has started to port this tool to 
+YAML file. With the birth of Docker Swarm, work has started to port this tool 
+to 
 the Swarm framework. At the time of this writing, the procedure to port a 
@@
 Additional components have been evaluated during the development of the 
-testbed, and here we propose a list of the most important ones. Some of them 
+test-bed, and here we propose a list of the most important ones. Some of them 
 where discarded as no more necessary, or during the evolution of the project 
@@
 layer completely makes the underlying infrastructure transparent: data 
-management, as network one, are totaly managed by Docker. Recently, 
+management, as network one, are totally managed by Docker. Recently, 
 Docker-compose support has been introduced, making this one-node tool available 
@@
 installing ad-hoc storage solutions. On the other hand, though, it doesn't have 
-all the personalization options that Kuberentes offers, thus making the product 
-less flexible. This foundamental characteristic, although, makes the two 
+all the customization options that Kubernetes offers, thus making the product 
+less flexible. This fundamental characteristic, although, makes the two 
 solutions have different use cases leading to different market needs.
 
-\subsubsection{Openvswitch} \todo{Talk about Openvswitch!!}
+\subsubsection{OpenvSwitch} \todo{Talk about OpenvSwitch!!}
 \subsubsection{Openbaton}
 \label{chap:prjan:sec:tech:sub:other:sub:openbaton}
-Patrocined by the Institute Fraunhofer, it's a open-source, customizable NFV 
-MANO-compliant framework, that offers manyconfigurations options, and since 
+Patronized by the Institute Fraunhofer, it's a open-source, customizable NFV 
+MANO-compliant framework, that offers many configurations options, and since 
 it's 
@@
 containers too. It's designed to work with cloud providers like Amazon or 
-Google Compute engine, but it offers compability with on-permise cloud 
+Google Compute engine, but it offers compatibility with on-premise cloud 
 solutions like Openstack, where it deploys virtual machines using the API
@@
 YAML\todo{What does TOSCA mean? Find out} and it can handle multiple
-PoPs\todo{Write down PoP meaning}. It offers a VNF lifecyle management
+PoPs\todo{Write down PoP meaning}. It offers a VNF lifecycle management
 out-of-the-box, with the possibility to customize it based on the user needs.

### 2018-11-16 res/sections/08-Background.tex
Fix typos

--- 2018-11-16 res/sections/08-Background.tex
@@
 The latter component serves as the centralized network intelligence and could
-consists of one or more controllers. It has some drawbacks in teorms of
+consists of one or more controllers. It has some drawbacks in terms of
 security, scalability and elasticity, because of the centralized architecture.
@@
 As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be splitted in three different layers, called planes, based on 
+architecture can be split in three different layers, called planes, based on 
 their functionalities.
@@
 applications can consume, enabling the possibility to use a variety of 
-different verdors' controllers in the same way.
-Analysing the SDN applications is possible to note two defined layers, one that 
+different vendors' controllers in the same way.
+Analyzing the SDN applications is possible to note two defined layers, one that 
 is responsible to manage the whole logic, and another one dedicated to the 
-communication part, that in particular uses the previsously defined NBIs.
+communication part, that in particular uses the previously defined NBIs.
 These applications can also build an abstracted view of the network, using
@@
 Due to the variety of tasks a controller has to accomplish, its logical 
-suddivision is achieved with a three-layer architecture, in which every layer 
+subdivision is achieved with a three-layer architecture, in which every layer 
 exposes it's own interface, that can be consequently used by the upper 
@@
 process the requests coming from the upper layer (aka the third layer), that 
-has the specific task to communicate with the remaining network compoents, in 
+has the specific task to communicate with the remaining network components, in 
 other words the Control to Data-Plane Interface and/or the already 
@@
 In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
-how switches works in the OpenFlow protocol. Basically, OpenFlow swiches own a
+how switches works in the OpenFlow protocol. Basically, OpenFlow switches own a
 set of tables to organized packet control and forwarding. These tables can be
 updated by higher-level application, usually an SDN application or controller,
-taking advatage of OpenFlow channels (called OF channels).
+taking advantage of OpenFlow channels (called OF channels).
 
@@
 hardware, software and functionalities into a virtual network. This abstraction
-runs on the top of a physical network in a hypervisor, decupling the logical
-network from the harwdare one. Network virtualization allow multiple,
+runs on the top of a physical network in a hyper-visor, decoupling the logical
+network from the hardware one. Network virtualization allow multiple,
 heterogeneous networks to cohabit, increasing the flexibility and manageability.
@@
 The idea of network virtualization came from the decoupling of the role of the
-traditional ISP into two different entities: who provides the infrastucture
+traditional ISP into two different entities: who provides the infrastructure
 (InPs) and who provides the services(SPs). In that context it is possible to
@@
 private and secure tunnels, relying on the public network such as Internet.
-Other examples of netvork virtualization are \emph{overlay networks}. They are
-logical networks, build on the top of an existing physical network, tipically
+Other examples of network virtualization are \emph{overlay networks}. They are
+logical networks, build on the top of an existing physical network, typically
 implemented at the application layer.
@@
 The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
-virtualization tecnologies to virtualize network nodes functions. VNFs may run 
-different processes and softwares, on the top of servers, switches, storage 
-devices or on cloud computing infrastuctures, that substitute custom hardware. 
+virtualization technologies to virtualize network nodes functions. VNFs may run 
+different processes and software, on the top of servers, switches, storage 
+devices or on cloud computing infrastructures, that substitute custom hardware. 
 That approach worked well in the past, because of the need to follow rigorous 
 standards, providing at the same time high quality, but it led to long product 
-cycles and high specialized proprietary hardware. New Internet senarios, that 
+cycles and high specialized proprietary hardware. New Internet scenarios, that 
 include IoT and the increasing number of services to spread multimedia data 
-(Skype, Netflix, Spotify, ...), requires more flexible functions and faster 
+(Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
 network deployments. The NFV technology is exploited in a framework, organized

### 2018-11-16 res/sections/79-Conclusions.tex
Minor fix

--- 2018-11-06 res/sections/79-Conclusions.tex
@@
 
-In this thesis we have seen an ETSI MANO implementation proposal. We have 
-described the technological components needed to build it, and we have 
-presented the requirements of the whole implementations. Then we described 
-the challenges we have encountered and how we solved it, taking choises 
-valuating the pros and cons.
+In this thesis we have seen an ETSI MANO and RFC 7665 implementation proposal. 
+We described the technological components needed to build it, and we presented 
+the requirements of the whole implementation. Then we described the challenges 
+we encountered and how we solved it, taking choises valuating their pros and 
+cons.
 
-Finally, we corroborated our implementation with test, showing that a 
-containerized approach is faster and more reliable thant a virtualized one
+Finally, we corroborated our implementation with tests and metrics, showing 
+that a containerized approach is faster and more reliable and faster than a 
+virtualized one, and illustrating that latency in communications through SFCs, 
+while there is room for improvement, are decent considered that this is just 
+a test-bed implementation.
+
+In conclusion, we can say that the ETSI MANO specification, coupled with the 
+RFC 7665 proposal and the new virtualization technologies avaiable in the 
+market could change the shape of future mobile communications, leading to the 
+creation of a 5G able to either saddisfy the challenges of the present, such as 
+IoT communications, media streaming, and real-time data, with the challenges of 
+the future, as smart dust and flying autonomous vehicles emerge as the next 
+possible innovations where connectivity will play an essential role.

### 2018-11-16 res/sections/25-Tests.tex
Add graphs and jitter table

@@
 Regarding the delivery performance, a metric to take in account is the \emph{round trip time}(RTT), that is the time interval between the instant in whic a packet is sent and the response is received.
-To test it there was developed a script that sends a packet and waits for the corrispondant response, recording the instants (expressed as \emph{epoch}) of send and receive.
+To test it there was developed a script that sends a packet with 8B of payload and waits for the corrispondant response, recording the instants (expressed as \emph{epoch}) of send and receive.
 The sake of thi test is to compare the RTT measured without any SFC (only sending the packet to the receiver waiting for the response) and making packets pass through different length SFCs, in particular with 1, 2, 3, 4, 5, and 10 hops.
@@
 \begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.35]{meanrtt}
-  \caption{Mean RTT calculate using SFCs with different lengths}
+  %\centering
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[scale=0.4]{meanrtt}
+        \caption{Mean RTT calculate using SFCs of different lengths}
         \label{chap:tests:sec:rtt:img:meanstt}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[b]{0.45\textwidth}
+        \includegraphics[scale=0.4]{differencenosfc}
+        \caption{Time increase using SFCs of different lengths}
+        \label{chap:tests:sec:rtt:img:differencertt}
+    \end{subfigure}
 \end{figure}
@@
 Looking at the bars representing the SFC composed by 1, 2, 3, 4 and 5 links we can see a pretty linear increase of the RTT.
-This trend strongly depends on the nature of the SF involved in the SFC.
+This trend strongly depends on the nature of the SF involved in the SFC, in that case the linear increase is highligted due to the use of the same funcion.
+In Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference between the RTT measured with a normal packet transmission and using the chains with different size, that represent the additional time needed for each transmission.
+Even in that case we can see that some time is waisted against a normal transmission and that this delta of time depends on the number of SFC involved in the chain.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.4]{differenceonehop}
+  \caption{Time difference between SFCs that have a length difference of one SF}
+  \label{chap:tests:sec:rtt:img:differenceonehop}
+\end{figure}
+
+In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the difference in terms of RTT among chains that have a length difference of one link.
+The first bar represent the comparison between an SFC composed by only one element and the direct trasmission to the receiver: that explain why that value is higher than the others.
+Other differences among SFC, instead, are quite similar.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\centering
+\begin{tabular}{@{}cc@{}}
+\toprule
+\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
+len. 0              & 0.0001074767    \\
+len. 1              & 0.0028695822    \\
+len. 2              & 0.0032037902    \\
+len. 3              & 0.0031961584    \\
+len. 4              & 0.0027278513    \\
+len. 5              & 0.0036802184    \\
+len. 10             & 0.0063434506    \\ \bottomrule
+\end{tabular}
+\caption{Jitterf for different SFC lengths}
+\label{chap:tests:sec:rtt:tab:jitter}
+\end{table}
+
+Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for different SFC length, based on values of RTT measured in the previous test.
+Jitter is the variation in the delay of received packet, and gives a measure of the congestion of the network.
+As we can see values are really low, even if the lowest one is always the one calculated without the chain.

### 2018-11-16 res/sections/25-Tests.tex
Spellcheck and 80 columns

@@
 tests relatively two main sections: one testing the architecture overall
-responsivness and one checking the SFC implementation and efficiency.
+responsiveness and one checking the SFC implementation and efficiency.
 
-\section{System responsivness}
+\section{System responsiveness}
 
@@
 
-Regarding the overall system responsivness, a metric we considered worth to 
+Regarding the overall system responsiveness, a metric we considered worth to 
 test was the VNF start up time in a docker environment versus one virtualized 
-through a common virtual machine system (in this case Virtual Box was choosen). 
+through a common virtual machine system (in this case Virtual Box was chosen). 
 To make this simulation as fair as possible, we used the same node (an 
-Openstack VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the same 
+Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the same 
 ``boot sequence'' for both the VNFs: first we launched the Astaire framework, 
@@
 that registered timing with a precision of hundredths of a second. Nonetheless,
-this didn't influence the test too much. To reduce possible outliers we perfomed
+this didn't influence the test too much. To reduce possible outliers we performed
 the test $100$ times. With this $N$, the total time spent by the Virtual
@@
 \caption[Docker vs Virtual Box start up times comparison]{Docker vs Virtual Box
-  start up times comparison. Is possible to se how the overall time required to
+  start up times comparison. Is possible to see how the overall time required to
   start $100$ containers is less than the time to start one traditional virtual
-  machine. The overall time required by the traitional virtualization system is
+  machine. The overall time required by the traditional virtualization system is
   approximately of 33 minutes, witch gives a significant insight of how much
-  time is required to create and boot a virtualized enviroment.}
+  time is required to create and boot a virtualized environment.}
 \label{my-label}
@@
 A Kubernetes installation was performed in order to carry out the test
-descripted above. All the machines involved (master node and minions) had the
+described above. All the machines involved (master node and minions) had the
 same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
@@
 \subsection{Test system}
-All the machine used for testing purposes are Openstack virtual machines, with the following specifications:
+All the machine used for testing purposes are Openstack virtual machines, with
+the following specifications:
 \begin{itemize}
-\item 1 machine with 32GB of RAM, 8vCPU and solid state storage, on which the receiver of packets is running;
-\item 4 machines with 16GB of RAM, 6vCPU and solid state storage, that form a Kubernetes cluster on which SFC infrastructure is deployed;
-\item 1 machine with 4GB of RAM, 2vCPU and solid state storage, on which scripts to send packets run.
+\item 1 machine with 32GB of RAM, 8vCPU and solid state storage, on which the
+receiver of packets is running;
+\item 4 machines with 16GB of RAM, 6vCPU and solid state storage, that form a
+Kubernetes cluster on which SFC infrastructure is deployed;
+\item 1 machine with 4GB of RAM, 2vCPU and solid state storage, on which
+scripts to send packets run.
 \end{itemize}
@@
 \subsection{RTT measurement}
-Regarding the delivery performance, a metric to take in account is the \emph{round trip time}(RTT), that is the time interval between the instant in whic a packet is sent and the response is received.
-To test it there was developed a script that sends a packet with 8B of payload and waits for the corrispondant response, recording the instants (expressed as \emph{epoch}) of send and receive.
-The sake of thi test is to compare the RTT measured without any SFC (only sending the packet to the receiver waiting for the response) and making packets pass through different length SFCs, in particular with 1, 2, 3, 4, 5, and 10 hops.
-Packets traverse the SFC on both sides and the results are collected after 100 tests and the SF only prints the packet received and the total number of packets that have passed throught it.
+Regarding the delivery performance, a metric to take in account is the
+\emph{round trip time}(RTT), that is the time interval between the instant in
+which a packet is sent and the response is received. To test it there was
+developed a script that sends a packet with 8B of payload and waits for the
+correspondent response, recording the instants (expressed as \emph{epoch}) of
+send and receive. The sake of this test is to compare the RTT measured without
+any SFC (only sending the packet to the receiver waiting for the response) and
+making packets pass through different length SFCs, in particular with 1, 2, 3,
+4, 5, and 10 hops. Packets traverse the SFC on both sides and the results are
+collected after 100 tests and the SF only prints the packet received and the
+total number of packets that have passed through it.
 
@@
 Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
-the RTT increase with the number of elements of the chain.
-There is a big difference between the mean RTT of the configuration wiothout any chain and the one with an SFC with a single link: this outcome has different explaination.
-As explained in \todo{Add reference to implementation} the chain has always at least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3 different running software.
-Another explaination lays into how the code was developed: it was not developed as a production ready product, but for a feasibility test, so connections and trasmissions can be enhanced.
-Finally, SFs are meant to perform some modifications to the packet, so, unavoidably, some time is waisted.
+the RTT increase with the number of elements of the chain. There is a big
+difference between the mean RTT of the configuration without any chain and the
+one with an SFC with a single link: this outcome has different explanation. As
+explained in \todo{Add reference to implementation} the chain has always at
+least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3
+different running software. Another explanation lays into how the code was
+developed: it was not developed as a production ready product, but for a
+feasibility test, so connections and transmissions can be enhanced. Finally, SFs
+are meant to perform some modifications to the packet, so, unavoidably, some
+time is wasted.
 
@@
 
-The results are better represented in Figure~\ref{chap:tests:sec:rtt:img:meanstt}.
-In that graph we can easier see the difference among the chains with different lengths and the trasmission that does not use any SFC.
-Looking at the bars representing the SFC composed by 1, 2, 3, 4 and 5 links we can see a pretty linear increase of the RTT.
-This trend strongly depends on the nature of the SF involved in the SFC, in that case the linear increase is highligted due to the use of the same funcion.
-In Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference between the RTT measured with a normal packet transmission and using the chains with different size, that represent the additional time needed for each transmission.
-Even in that case we can see that some time is waisted against a normal transmission and that this delta of time depends on the number of SFC involved in the chain.
+The results are better represented in
+Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easier see
+the difference among the chains with different lengths and the transmission that
+does not use any SFC. Looking at the bars representing the SFC composed by 1,
+2, 3, 4 and 5 links we can see a pretty linear increase of the RTT. This trend
+strongly depends on the nature of the SF involved in the SFC, in that case the
+linear increase is highlighted due to the use of the same function. In
+Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference
+between the RTT measured with a normal packet transmission and using the chains
+with different size, that represent the additional time needed for each
+transmission. Even in that case we can see that some time is wasted against a
+normal transmission and that this delta of time depends on the number of SFC
+involved in the chain.
 
@@
 
-In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the difference in terms of RTT among chains that have a length difference of one link.
-The first bar represent the comparison between an SFC composed by only one element and the direct trasmission to the receiver: that explain why that value is higher than the others.
-Other differences among SFC, instead, are quite similar.
+In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the
+difference in terms of RTT among chains that have a length difference of one
+link. The first bar represent the comparison between an SFC composed by only
+one element and the direct transmission to the receiver: that explain why that
+value is higher than the others. Other differences among SFC, instead, are
+quite similar.
 
@@
 \end{tabular}
-\caption{Jitterf for different SFC lengths}
+\caption{Jitter for different SFC lengths}
 \label{chap:tests:sec:rtt:tab:jitter}
@@
 
-Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for different SFC length, based on values of RTT measured in the previous test.
-Jitter is the variation in the delay of received packet, and gives a measure of the congestion of the network.
-As we can see values are really low, even if the lowest one is always the one calculated without the chain.
+Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for
+different SFC length, based on values of RTT measured in the previous test.
+Jitter is the variation in the delay of received packet, and gives a measure of
+the congestion of the network. As we can see values are really low, even if the
+lowest result is always the one calculated without the chain.

### 2018-11-16 res/sections/25-Tests.tex
Spellcheck and 80 columns

@@
 To make this simulation as fair as possible, we used the same node (an 
-Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the same 
-``boot sequence'' for both the VNFs: first we launched the Astaire framework, 
-ensuring its availability to process data and then we notified the test 
-machine of the successful boot. The startup time was measured from the 
+Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the
+same ``boot sequence'' for both the VNFs: first we launched the Astaire
+framework, ensuring its availability to process data and then we notified the
+test machine of the successful boot. The startup time was measured from the 
 beginning of the startup command (\verb!docker run! for Docker and 
@@
 Virtual Box one: Docker allows to inspect the container and to gain startup and
-shutdown times with a precision of millseconds, while to gather this information
-for VirtualBox instances we had to use the command line utility \verb!time!,
-that registered timing with a precision of hundredths of a second. Nonetheless,
-this didn't influence the test too much. To reduce possible outliers we performed
+shutdown times with a precision of milliseconds, while to gather this
+information for VirtualBox instances we had to use the command line utility
+\verb!time!, that registered timing with a precision of hundredths of a second.
+Nonetheless,this didn't influence the test too much. To reduce possible
+outliers we performed
 the test $100$ times. With this $N$, the total time spent by the Virtual
@@
 \begin{longtable}[c]{c|c|c}
-\textbf{SFC length} & \textbf{Stamdard deviation (seconds)} & \textbf{Average 
+\textbf{SFC length} & \textbf{Standard deviation (seconds)} & \textbf{Average 
 deploy time (seconds)} \\ \hline

### 2018-11-16 res/sections/25-Tests.tex
Fix jitter
Add part on test concerning resilience

@@
 \toprule
-\textbf{SFC length} & \textbf{Mean RTT (milliseconds)} \\ \midrule
+\textbf{SFC length} & \textbf{Mean RTT (Seconds)} \\ \midrule
 len. 0              & 0.0009668279      \\
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
 \begin{table}[]
@@
 \toprule
-\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
-len. 0              & 0.0001074767    \\
-len. 1              & 0.0028695822    \\
-len. 2              & 0.0032037902    \\
-len. 3              & 0.0031961584    \\
-len. 4              & 0.0027278513    \\
-len. 5              & 0.0036802184    \\
-len. 10             & 0.0063434506    \\ \bottomrule
+\textbf{SFC Length} & \textbf{Jitter(Seconds)} \\ \midrule
+len. 0    & 0.0001484817    \\
+len. 1    & 0.0034793676    \\
+len. 2    & 0.0040213103    \\
+len. 3    & 0.0036568939    \\
+len. 4    & 0.0031962546    \\
+len. 5    & 0.0046836929    \\
+len. 10   & 0.0073287308    \\ \bottomrule
 \end{tabular}
@@
 lowest result is always the one calculated without the chain.
+
+\subsection{Resilience}
+We also developed also a test to analyze the resilience of the system.
+One of the most important requirements of our implementation is to make the system to be able to recover after an SF crash.
+Thanks to the exploit of Docker container technology and Kubernetes platform it is quite straightforward to restart a container after a fault.
+In fact, \texttt{yaml} files that allows to deploy containers on Kubernetes, make possible to set pods restart policy to always(the default one), on failure or never, but the overall system must be able to recognize that the SF is brought up again.
+
+During this test test we send out 500 UDP packets (1 packet each 50ms) from a sender to a receiver through an SFC of length 3.
+The configuration of the test is represented in Figure~\ref{chap:tests:sec:fault:img:testfaultconf}.
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.4]{testfaultconf}
+  \caption{Host logical configuration for testing}
+  \label{chap:tests:sec:fault:img:testfaultconf}
+\end{figure}
+The receiver record the instant on which it receives a packet and, to emulate a SF crash, we bring down the central link of the chain after 5 seconds using the command \verb!kubectl delete pods <pod-id> --grace-period=0 --force!, to force the stop of a pod.
+
+The behavior of the system in the case of the crash of a SF is depicted in Figure~\ref{chap:tests:sec:fault:img:faultgraph}.
+We repeated the test 20 times and we measured an average interval of time on which the receiver does not receive anything of 13.5793391228s.
+In the graph are only shown the best and the worst cases.
+In both cases we can see that the interval of time in which packets are not delivered is similar and it is not significantly different from the average (around 12.75 the best case, 14.41 the worst one).
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.3]{faultgraph}
+  \caption{Packet time arrival in case of SF fault}
+  \label{chap:tests:sec:fault:img:faultgraph}
+\end{figure}
+
+After those results, we repeated the test with the configuration shown in Figure~\ref{chap:tests:sec:fault:img:testfaultconfwithreplica}, in order to test the resilience of the system in case of replicas of SF.
+In that case after repeating the test 10 times, bringing down only one of the 2 equal SFs.
+In that case we were not able to record any significant interval of time in which the receiver does not received packet (receiving times are consistent with sending times).
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.4]{testfaultconfwithreplica}
+  \caption{Host logical configuration in case of a replicated SF for testing}
+  \label{chap:tests:sec:fault:img:testfaultconfwithreplica}
+\end{figure}

### 2018-11-17 res/sections/25-TestsResponsiveness.tex
Correct part on performance tests and splitted tests in 3 files

--- 2018-11-16 res/sections/25-Tests.tex
@@
 \textbf{6}          & 1.103802                   & 17,418                     \\
-\textbf{7}          & 1.561451                   & 18,815                     \\
-\textbf{8}          & 0.7698537                  & 20,856                     \\
-\textbf{9}          & 1.889197                   & 20,187                     \\
-\textbf{10}         & 1.709227                   & 20,476                     \\
+\textbf{7}          & 2.944699                   & 17.984                     \\
+\textbf{8}          & 1.733722                   & 18.480                     \\
+\textbf{9}          & 1.746296                   & 18.607                     \\
+\textbf{10}         & 1.291508                   & 20.511                     \\
 \caption[SFC start up time]{Table of the results. It is possible to see how the
@@
 \end{longtable}
-
-\section{Final considerations}
-
-In these tests is possible to see one critical aspect concerning an architecture
-that follows the containerization approach: great flexibility in terms of
-deployments. This possibility to deploy services in the form of SFCs and VNFs in
-a very small amount of time is one of the key points required in the new 5G
-architecture, that the tests above demonstrated this project is able to adhere.
-Finally, the ability to rapidly redeploy software means smooth and easier
-software development lifecycle, where developers have the possibility to
-automatically push ``fresh'' software to production, letting the MANO and thus
-the container orchestrator handling all the difficult and delicate tasks to put
-software in a production environment, eliminating the possibility of human
-errors and assuring longer uptimes.
-
-\section{System performances}
-
-\subsection{Test system}
-All the machine used for testing purposes are Openstack virtual machines, with
-the following specifications:
-\begin{itemize}
-\item 1 machine with 32GB of RAM, 8vCPU and solid state storage, on which the
-receiver of packets is running;
-\item 4 machines with 16GB of RAM, 6vCPU and solid state storage, that form a
-Kubernetes cluster on which SFC infrastructure is deployed;
-\item 1 machine with 4GB of RAM, 2vCPU and solid state storage, on which
-scripts to send packets run.
-\end{itemize}
-
-\subsection{RTT measurement}
-Regarding the delivery performance, a metric to take in account is the
-\emph{round trip time}(RTT), that is the time interval between the instant in
-which a packet is sent and the response is received. To test it there was
-developed a script that sends a packet with 8B of payload and waits for the
-correspondent response, recording the instants (expressed as \emph{epoch}) of
-send and receive. The sake of this test is to compare the RTT measured without
-any SFC (only sending the packet to the receiver waiting for the response) and
-making packets pass through different length SFCs, in particular with 1, 2, 3,
-4, 5, and 10 hops. Packets traverse the SFC on both sides and the results are
-collected after 100 tests and the SF only prints the packet received and the
-total number of packets that have passed through it.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{@{}cc@{}}
-\toprule
-\textbf{SFC length} & \textbf{Mean RTT (milliseconds)} \\ \midrule
-len. 0              & 0.0009668279      \\
-len. 1              & 0.0555884027      \\
-len. 2              & 0.0702438903      \\
-len. 3              & 0.0888472247      \\
-len. 4              & 0.1045638967      \\
-len. 5              & 0.1239343858      \\
-len. 10             & 0.2135494161      \\ \bottomrule
-\end{tabular}
-\caption{Mean RTT measured with different SFC length}
-\label{chap:tests:sec:rtt:tab:meanrtt}
-\end{table}
-
-Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
-the RTT increase with the number of elements of the chain. There is a big
-difference between the mean RTT of the configuration without any chain and the
-one with an SFC with a single link: this outcome has different explanation. As
-explained in \todo{Add reference to implementation} the chain has always at
-least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3
-different running software. Another explanation lays into how the code was
-developed: it was not developed as a production ready product, but for a
-feasibility test, so connections and transmissions can be enhanced. Finally, SFs
-are meant to perform some modifications to the packet, so, unavoidably, some
-time is wasted.
-
-\begin{figure}[H]
-  %\centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[scale=0.4]{meanrtt}
-        \caption{Mean RTT calculate using SFCs of different lengths}
-        \label{chap:tests:sec:rtt:img:meanstt}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[scale=0.4]{differencenosfc}
-        \caption{Time increase using SFCs of different lengths}
-        \label{chap:tests:sec:rtt:img:differencertt}
-    \end{subfigure}
-\end{figure}
-
-The results are better represented in
-Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easier see
-the difference among the chains with different lengths and the transmission that
-does not use any SFC. Looking at the bars representing the SFC composed by 1,
-2, 3, 4 and 5 links we can see a pretty linear increase of the RTT. This trend
-strongly depends on the nature of the SF involved in the SFC, in that case the
-linear increase is highlighted due to the use of the same function. In
-Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference
-between the RTT measured with a normal packet transmission and using the chains
-with different size, that represent the additional time needed for each
-transmission. Even in that case we can see that some time is wasted against a
-normal transmission and that this delta of time depends on the number of SFC
-involved in the chain.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.4]{differenceonehop}
-  \caption{Time difference between SFCs that have a length difference of one SF}
-  \label{chap:tests:sec:rtt:img:differenceonehop}
-\end{figure}
-
-In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the
-difference in terms of RTT among chains that have a length difference of one
-link. The first bar represent the comparison between an SFC composed by only
-one element and the direct transmission to the receiver: that explain why that
-value is higher than the others. Other differences among SFC, instead, are
-quite similar.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\centering
-\begin{tabular}{@{}cc@{}}
-\toprule
-\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
-len. 0              & 0.0001074767    \\
-len. 1              & 0.0028695822    \\
-len. 2              & 0.0032037902    \\
-len. 3              & 0.0031961584    \\
-len. 4              & 0.0027278513    \\
-len. 5              & 0.0036802184    \\
-len. 10             & 0.0063434506    \\ \bottomrule
-\end{tabular}
-\caption{Jitter for different SFC lengths}
-\label{chap:tests:sec:rtt:tab:jitter}
-\end{table}
-
-Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for
-different SFC length, based on values of RTT measured in the previous test.
-Jitter is the variation in the delay of received packet, and gives a measure of
-the congestion of the network. As we can see values are really low, even if the
-lowest result is always the one calculated without the chain.

### 2018-11-17 res/sections/26-TestsPerformance.tex
Correct part on performance tests and splitted tests in 3 files

\section{System performances}

\subsection{Test system}
All the machine used for testing purposes are Openstack virtual machines, with
the following specifications:
\begin{itemize}
\item 1 machine with 32GB of RAM, 8vCPU and solid state storage. This host was
used as final destination of packets;
\item 4 machines with 16GB of RAM, 6vCPU and solid state storage. These VMs
were utilized to form the Kubernetes cluster, on the top of which the SFC
infrastructure developed was running;
\item 1 machine with 4GB of RAM, 2vCPU and solid state storage. This host on
was used as client and on which run script to send packets.
\end{itemize}

\subsection{RTT measurement}
Regarding the delivery performance, a metric to take in account is the
\emph{round trip time}(RTT), that is the time interval between the instant in
which a packet is sent and when the response is received. To analyze it we
developed a Python script that sends a packet (with 8B of payload) and waits for
the correspondent response, recording the instants (expressed as \emph{epoch})
of transmission and receiving. The goal of this test is to compare the average
RTT measured without involving any SFC (therefore simply delivering the packet
to the receiver and get back the reply) with the RTT in the case in which
packets pass through different length SFCs. In particular, this experiments was
repeated with chains that have 1, 2, 3, 4, 5, and 10 hops respectively. It is
useful to remember that packets traverse the SFC on both sides: SF that are
used only prints the packet received and the total number of packets that have
passed through it. Results are collected after 100 repetitions of the test.

\begin{table}[H]
\centering
\begin{tabular}{@{}cc@{}}
\toprule
\textbf{SFC length} & \textbf{Mean RTT (Seconds)} \\ \midrule
len. 0              & 0.0009668279      \\
len. 1              & 0.0555884027      \\
len. 2              & 0.0702438903      \\
len. 3              & 0.0888472247      \\
len. 4              & 0.1045638967      \\
len. 5              & 0.1239343858      \\
len. 10             & 0.2135494161      \\ \bottomrule
\end{tabular}
\caption{Mean RTT measured with different SFC length}
\label{chap:tests:sec:rtt:tab:meanrtt}
\end{table}

Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
the RTT is bigger when packets travel through the chain and it increases with
the total number of elements of the SFC. There is a large
difference between the mean RTT of the configuration without any chain and the
one with an SFC with a single link: this outcome has different explanation. As
described in \todo{Add reference to implementation} the chain has always at
least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3
different running software. Another interpretation lays into how the code was
written: components of these framework are not developed as a production ready
software, but for a proof of concept, so connections and transmissions can be
enhanced. Finally, SFs are meant to perform some modifications to the packet,
so, unavoidably, some time is required for their computation.

\begin{figure}[H]
  %\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[scale=0.4]{meanrtt}
        \caption{Mean RTT calculate using SFCs of different lengths}
        \label{chap:tests:sec:rtt:img:meanstt}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[scale=0.4]{differencenosfc}
        \caption{Time increase using SFCs of different lengths}
        \label{chap:tests:sec:rtt:img:differencertt}
    \end{subfigure}
\end{figure}

The results on average RTT are better represented in
Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easily see
the difference among the different configurations. Looking at the bars in the
center, representing the SFCs composed by 1, 2, 3, 4 and 5 links, we can see a
fairly linear increase of the RTT. This trend strongly depends on the nature of
the SF involved in the chain, and in this experiment the linear increase is
highlighted due to the usage of the same function for each hop. In
Figure~\ref{chap:tests:sec:rtt:img:differencertt} we can see the additional
time required for each transmission with the different SFCs, calculated as the
difference between the average RTT of a specific chain and the RTT measured
without involving any SFC. Event in that plot we can see a quite linear
increase depending on the length of the chain.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.4]{differenceonehop}
  \caption{Time difference between SFCs that have a length difference of one SF}
  \label{chap:tests:sec:rtt:img:differenceonehop}
\end{figure}

In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the
difference in terms of average RTT between chains whose lengths differs only by
one link. The highest value was calculated bringing into comparison the direct
deliver of packets and the chain of only one element: even in that case is
remarked that SFC cost time in the context of transmission of single packets.
Other differences among SFC, instead, are quite similar, highlighting the
linear trend.

\begin{table}[H]
\centering
\begin{tabular}{@{}cc@{}}
\toprule
\textbf{SFC Length} & \textbf{Jitter(Seconds)} \\ \midrule
len. 0    & 0.0001484817    \\
len. 1    & 0.0034793676    \\
len. 2    & 0.0040213103    \\
len. 3    & 0.0036568939    \\
len. 4    & 0.0031962546    \\
len. 5    & 0.0046836929    \\
len. 10   & 0.0073287308    \\ \bottomrule
\end{tabular}
\caption{Jitter for different SFC lengths}
\label{chap:tests:sec:rtt:tab:jitter}
\end{table}

Another metric to take in account is jitter. Jitter is the variation in the
delay of received packet, and gives a measure of the congestion of the network
or underline other communication problems.
Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for
different SFC length, derived from values of RTT measured in the previous test.
As we can see values are really low, indicating no particular problems on the
network, even if the lowest result is always the one calculated without the
chain.

\subsection{Resilience}
Regarding the overall system resilience, we performed some test in which some
nodes of our solution are stopped during a transmission. As a matter of fact
that one of the most important requirements of our implementation is to make
the system  able to recover after an SF crash. Exploiting Docker container
technology and Kubernetes platform it is quite straightforward to restart a
container after a fault. In fact, \texttt{yaml} files that allows to deploy
containers on Kubernetes, make possible to set pods restart policy to
``always'' (the default one), ``on failure'' or ``never'', but the overall
system must be able to recognize that the SF is brought up again.

During this test test we forward 500 packets (1 packet each 50ms) from a sender
host to a receiver through an SFC of length 3. The \texttt{yaml} file
describing the chain is the following:

\lstinputlisting[caption={Kubernetes YAML configuration that has been used for
resilience test}, captionpos=b]{res/code/fault_test.yaml}

The previous configuration can be represented as in
Figure~\ref{chap:tests:sec:fault:img:testfaultconf}.
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.4]{testfaultconf}
  \caption{Host logical configuration for testing}
  \label{chap:tests:sec:fault:img:testfaultconf}
\end{figure}

The receiver record the instant on which it receives a packet and, to emulate a
SF crash, we bring down the central link of the chain after 5 seconds using the
command \verb!kubectl delete pods <pod-id> --grace-period=0 --force!, to force
the stop of a pod.

We performed the test using both UDP and TCP protocols, repeating for each
protocol the measurement 20 times.
Figure~\ref{chap:tests:sec:fault:img:faultgraphudp} and
Figure~\ref{chap:tests:sec:fault:img:faultgraphtcp} represent the behavior of
the system in the worst case and in the best case measured, respectively for
UDP and TCP. The average interval of time in which the receiver does not
receive anything was 13.5793391228s in case of UDP, 13.7828050137s in case of
TCP. This is a good result, because in about 13s a machine can crash and can be
recreated and be operative again. We can also notice that there are no
significant difference between worst cases and best cases measured 
respectively 12.75 and 14.41 for UDP and 13.14 and 14.63 for TCP)

\begin{figure}[]
  %\centering
    \begin{subfigure}[b]{0.9\textwidth}
        \includegraphics[scale=0.3]{nolog_udp}
        \caption{Packet time arrival in case of SF fault using UDP}
        \label{chap:tests:sec:fault:img:faultgraphudp}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.9\textwidth}
        \includegraphics[scale=0.3]{nolog_tcp}
        \caption{Packet time arrival in case of SF fault using TCP}
        \label{chap:tests:sec:fault:img:faultgraphtcp}
    \end{subfigure}
    \caption{Comparison among TCP and UDP in terms of packet time arrival in
    case of SF fault}
    \label{chap:tests:sec:fault:img:faultgrapht}
\end{figure}

After those results, we repeated the test (this time only using UDP) with the
configuration and shown in
Figure~\ref{chap:tests:sec:fault:img:testfaultconfwithreplica}, in order to
test the resilience of the system in case of a replica of a certain SF. We
repeated this experiment 10 times, bringing down only one of the 2 replica of
the same SF and we were not able to record any significant interval of time in
which the receiver does not received packet (receiving times are consistent
with sending times). So we can consider that in case of light traffic only 2
replicas of a specific function can be enough to offer to final user a reliable
service.

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.4]{testfaultconfwithreplica}
  \caption{Host logical configuration in case of a replicated SF for testing}
  \label{chap:tests:sec:fault:img:testfaultconfwithreplica}
\end{figure}

### 2018-11-17 res/sections/27-TestsFinalConsiderations.tex
Correct part on performance tests and splitted tests in 3 files

--- 2018-11-16 res/sections/25-Tests.tex
@@
-\chapter{Tests and metrics}
-
-After the testbed implementation, tests about its performance and metrics were
-gathered in order to understand the real project effectiveness. We performed
-tests relatively two main sections: one testing the architecture overall
-responsiveness and one checking the SFC implementation and efficiency.
-
-\section{System responsiveness}
-
-\subsection{Docker vs Virtual Box startup measurement}
-
-Regarding the overall system responsiveness, a metric we considered worth to
-test was the VNF start up time in a docker environment versus one virtualized
-through a common virtual machine system (in this case Virtual Box was chosen).
-To make this simulation as fair as possible, we used the same node (an 
-Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the
-same ``boot sequence'' for both the VNFs: first we launched the Astaire
-framework, ensuring its availability to process data and then we notified the
-test machine of the successful boot. The startup time was measured from the 
-beginning of the startup command (\verb!docker run! for Docker and 
-\verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
-test backend (to make things as smooth as possible, \verb!netcat! was used 
-to listen to incoming TCP data).
-
-It is worth saying that Docker metrics are generally more precise than the
-Virtual Box one: Docker allows to inspect the container and to gain startup and
-shutdown times with a precision of milliseconds, while to gather this
-information for VirtualBox instances we had to use the command line utility
-\verb!time!, that registered timing with a precision of hundredths of a second.
-Nonetheless,this didn't influence the test too much. To reduce possible
-outliers we performed
-the test $100$ times. With this $N$, the total time spent by the Virtual
-Machines to boot up was of $2021,15$ seconds, while with containers the same
-number of start ups resulted in a total of $1.638$ seconds. Summing up, Docker
-employs only the $0,08\%$ of the overall time spent for the test, underlining
-how isolating the process with a set of defined policy and avoiding a
-virtualized kernel boot-up saves a considerable amount of time.
-
-\begin{figure}[H]
-    \centering
-    \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[scale=0.4]{docker_vs_vmBarplotGraph}
-        \caption{Barplot between Virtual Box and Docker startup times.}
-        \label{chap:tests:sec:dockervsvb:img:barplot}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.4\textwidth}
-        \includegraphics[scale=0.35]{docker_vs_vmBoxplotGraph}
-        \caption{Boxplot between Virtual Box and Docker with values normalized 
-in range $0-1$.}
-        \label{chap:tests:sec:dockervsvb:img:boxplot}
-    \end{subfigure}
-    \caption[Virtual Box vs Docker start up comparison]{In these graphs it is
-      possible to see how the total time to start up a virtual machine through
-      Virtual Box is very high compared to the time required to the overall
-      Docker containers to boot. A single VirtualBox instance requires on
-      average 20,211 seconds to boot up, while a container requires, on average,
-      only 0.016 seconds. The main reason behind this difference, as already
-      explained in other parts of this thesis, it is caused by the guest's
-      kernel that has to initialized itself, before starting any program. On
-      Docker, instead, the kernel is the same of the host, so the only
-      operations needed are to isolate the program from the rest of the OS and
-      add it to the process list.}
-    \label{chap:tests:sec:dockervsvb:subimg:plots}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{l|c|c}
-                     & \textbf{Standard deviation (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
-\textbf{Docker}      & 0.009620874                           & 0,016                                    \\
-\textbf{Virtual Box} & 0.755                                 & 20,211                                 
-\end{tabular}%
-}
-\caption[Docker vs Virtual Box start up times comparison]{Docker vs Virtual Box
-  start up times comparison. Is possible to see how the overall time required to
-  start $100$ containers is less than the time to start one traditional virtual
-  machine. The overall time required by the traditional virtualization system is
-  approximately of 33 minutes, witch gives a significant insight of how much
-  time is required to create and boot a virtualized environment.}
-\label{my-label}
-\end{table}
-
-% The overall time required by the traitional virtualization system is
-% approximately of 33 minutes, witch gives a significant insight of how much time
-% is required to create and boot a virtualized enviroment.
-
-\subsection{SFC chain deploy measurement}
-
-The previous test make us see how much Docker is faster than a traditional
-virtual machine. The test, though, was performed on a single node, interacting
-directly with Docker and not through the orchestrator we used in the thesis
-development, Kubernetes. The following measurement will allow us to see the time
-required in order to deploy an SFC of different elements in a Kubernetes
-environment. The deployment consists in a Pod inside a Deployment definition
-with a related Service, to simulate a real VNF. After the Pod startup, a request
-is made to a backend designed to count the chain elements that have successfully
-boot-up returning, eventually, the overall start up timing.
-
-% TODO: missing YAML definition: [language=YAML]
-%https://tex.stackexchan
-%ge.com/questions/152829/how-can-i-highlight-yaml-code-in
-%- a-pretty-way-with-listings
-\lstinputlisting[caption={Kubernetes YAML configuration that has been used to
-    simulate a single VNF deployment}, captionpos=b]{res/code/sfc_test.yaml}
-
-\vspace{0.5cm}
-
-\noindent As is possible to see from the YAML definition, the test involved, for
-every launch (that included and $N = 100$, for a total for 1000 tests), the 
-download of the container image from the Docker hub repositories. This, while 
-it forced every node to delete and download the image for each test, it ensured 
-consistent measurements in cases where the selected node for the deployment did 
-not have the image already pulled. In addition to that, another consideration 
-needs to be taken into account: when the chain consists of two or more 
-elements, the container image which is used for the container deployment is the 
-same for every VNF involved, thus reducing the start up time for the subsequent 
-pods that reside in nodes that have already deployed one.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.5]{sfc_startupBarplotGraph}
-  \caption[SFC start-up time barplot]{SFC start-up time barplot. It is
-    noticeable how there seems to be a linear correlation between the SFC
-    startup time and the SFC length, even though for the chain with eight
-    elements it is possible to see that the overall deploy time is more compared
-    to chains with nine and ten elements.}
-  \label{chap:tests:sec:sfclength:img:barplot}
-\end{figure}
-
-\paragraph*{Kubernetes specifications}
-A Kubernetes installation was performed in order to carry out the test
-described above. All the machines involved (master node and minions) had the
-same specifications i.e., 4GB of RAM, 2 vCPU and 10GB of SSD storage.
-Nonetheless the constrained resources, absolutely far from a real Kubernetes
-production environment, it is possible to notice how, even with a chain of ten
-elements, the average time required to deploy an SFC is equal, if not smaller,
-of the average time required to deploy a Virtual Box machine on an instance 
-with 32GB of RAM, 8 vCPU and 50GB of SSD storage.
-
-% TODO is it worth to keep it as a long table?
-\begin{longtable}[c]{c|c|c}
-\textbf{SFC length} & \textbf{Standard deviation (seconds)} & \textbf{Average 
-deploy time (seconds)} \\ \hline
-\endhead
-%
-\textbf{1}          & 2.560102                   & 13,291                     \\
-\textbf{2}          & 0.6301444                  & 14,207                     \\
-\textbf{3}          & 1.14903                    & 15,345                     \\
-\textbf{4}          & 3.592179                   & 16,453                     \\
-\textbf{5}          & 0.9728132                  & 16,634                     \\
-\textbf{6}          & 1.103802                   & 17,418                     \\
-\textbf{7}          & 1.561451                   & 18,815                     \\
-\textbf{8}          & 0.7698537                  & 20,856                     \\
-\textbf{9}          & 1.889197                   & 20,187                     \\
-\textbf{10}         & 1.709227                   & 20,476                     \\
-\caption[SFC start up time]{Table of the results. It is possible to see how the
-  time required for the SFC to go up increases with the length, reaching an
-  average of 20 seconds for chain of eight, nine and ten elements.}
-\label{chap:tests:sec:sfclength:tab:sfcdata}\\
-\end{longtable}
-
 \section{Final considerations}
@@
 errors and assuring longer uptimes.
-
-\section{System performances}
-
-\subsection{Test system}
-All the machine used for testing purposes are Openstack virtual machines, with
-the following specifications:
-\begin{itemize}
-\item 1 machine with 32GB of RAM, 8vCPU and solid state storage, on which the
-receiver of packets is running;
-\item 4 machines with 16GB of RAM, 6vCPU and solid state storage, that form a
-Kubernetes cluster on which SFC infrastructure is deployed;
-\item 1 machine with 4GB of RAM, 2vCPU and solid state storage, on which
-scripts to send packets run.
-\end{itemize}
-
-\subsection{RTT measurement}
-Regarding the delivery performance, a metric to take in account is the
-\emph{round trip time}(RTT), that is the time interval between the instant in
-which a packet is sent and the response is received. To test it there was
-developed a script that sends a packet with 8B of payload and waits for the
-correspondent response, recording the instants (expressed as \emph{epoch}) of
-send and receive. The sake of this test is to compare the RTT measured without
-any SFC (only sending the packet to the receiver waiting for the response) and
-making packets pass through different length SFCs, in particular with 1, 2, 3,
-4, 5, and 10 hops. Packets traverse the SFC on both sides and the results are
-collected after 100 tests and the SF only prints the packet received and the
-total number of packets that have passed through it.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{@{}cc@{}}
-\toprule
-\textbf{SFC length} & \textbf{Mean RTT (milliseconds)} \\ \midrule
-len. 0              & 0.0009668279      \\
-len. 1              & 0.0555884027      \\
-len. 2              & 0.0702438903      \\
-len. 3              & 0.0888472247      \\
-len. 4              & 0.1045638967      \\
-len. 5              & 0.1239343858      \\
-len. 10             & 0.2135494161      \\ \bottomrule
-\end{tabular}
-\caption{Mean RTT measured with different SFC length}
-\label{chap:tests:sec:rtt:tab:meanrtt}
-\end{table}
-
-Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
-the RTT increase with the number of elements of the chain. There is a big
-difference between the mean RTT of the configuration without any chain and the
-one with an SFC with a single link: this outcome has different explanation. As
-explained in \todo{Add reference to implementation} the chain has always at
-least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3
-different running software. Another explanation lays into how the code was
-developed: it was not developed as a production ready product, but for a
-feasibility test, so connections and transmissions can be enhanced. Finally, SFs
-are meant to perform some modifications to the packet, so, unavoidably, some
-time is wasted.
-
-\begin{figure}[H]
-  %\centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[scale=0.4]{meanrtt}
-        \caption{Mean RTT calculate using SFCs of different lengths}
-        \label{chap:tests:sec:rtt:img:meanstt}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[scale=0.4]{differencenosfc}
-        \caption{Time increase using SFCs of different lengths}
-        \label{chap:tests:sec:rtt:img:differencertt}
-    \end{subfigure}
-\end{figure}
-
-The results are better represented in
-Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easier see
-the difference among the chains with different lengths and the transmission that
-does not use any SFC. Looking at the bars representing the SFC composed by 1,
-2, 3, 4 and 5 links we can see a pretty linear increase of the RTT. This trend
-strongly depends on the nature of the SF involved in the SFC, in that case the
-linear increase is highlighted due to the use of the same function. In
-Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference
-between the RTT measured with a normal packet transmission and using the chains
-with different size, that represent the additional time needed for each
-transmission. Even in that case we can see that some time is wasted against a
-normal transmission and that this delta of time depends on the number of SFC
-involved in the chain.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.4]{differenceonehop}
-  \caption{Time difference between SFCs that have a length difference of one SF}
-  \label{chap:tests:sec:rtt:img:differenceonehop}
-\end{figure}
-
-In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the
-difference in terms of RTT among chains that have a length difference of one
-link. The first bar represent the comparison between an SFC composed by only
-one element and the direct transmission to the receiver: that explain why that
-value is higher than the others. Other differences among SFC, instead, are
-quite similar.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\centering
-\begin{tabular}{@{}cc@{}}
-\toprule
-\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
-len. 0              & 0.0001074767    \\
-len. 1              & 0.0028695822    \\
-len. 2              & 0.0032037902    \\
-len. 3              & 0.0031961584    \\
-len. 4              & 0.0027278513    \\
-len. 5              & 0.0036802184    \\
-len. 10             & 0.0063434506    \\ \bottomrule
-\end{tabular}
-\caption{Jitter for different SFC lengths}
-\label{chap:tests:sec:rtt:tab:jitter}
-\end{table}
-
-Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for
-different SFC length, based on values of RTT measured in the previous test.
-Jitter is the variation in the delay of received packet, and gives a measure of
-the congestion of the network. As we can see values are really low, even if the
-lowest result is always the one calculated without the chain.

### 2018-11-18 res/sections/26-TestsPerformance.tex
Change images of resilience tests and fix description

@@
 
-The receiver record the instant on which it receives a packet and, to emulate a
-SF crash, we bring down the central link of the chain after 5 seconds using the
-command \verb!kubectl delete pods <pod-id> --grace-period=0 --force!, to force
-the stop of a pod.
+In that test we bring down the central link of the chain after 5 seconds using
+the command \verb!kubectl delete pods <pod-id> --grace-period=0 --force!, to
+force the stop of a pod and we recorded the time on which packets arrived to
+the receiver and passed through the VNF just before the one that will be make
+stopped.
 
@@
 Figure~\ref{chap:tests:sec:fault:img:faultgraphtcp} represent the behavior of
-the system in the worst case and in the best case measured, respectively for
-UDP and TCP. The average interval of time in which the receiver does not
-receive anything was 13.5793391228s in case of UDP, 13.7828050137s in case of
-TCP. This is a good result, because in about 13s a machine can crash and can be
-recreated and be operative again. We can also notice that there are no
-significant difference between worst cases and best cases measured 
-respectively 12.75 and 14.41 for UDP and 13.14 and 14.63 for TCP)
+the system during the experiment: blue points describe packet inter-arrival
+time on the receiver, green once the same metric on the SF that precede the
+central one. The average interval of time in which the
+receiver does not receive anything was 13.5793391228s in case of UDP,
+13.7828050137s in case of TCP. This is a good result, because in about 13s a
+machine can crash and can be recreated and be operative again.
 
@@
     \begin{subfigure}[b]{0.9\textwidth}
-        \includegraphics[scale=0.3]{nolog_udp}
+        \includegraphics[scale=0.3]{udp_fault}
         \caption{Packet time arrival in case of SF fault using UDP}
@@
     \begin{subfigure}[b]{0.9\textwidth}
-        \includegraphics[scale=0.3]{nolog_tcp}
+        \includegraphics[scale=0.3]{tcp_fault}
         \caption{Packet time arrival in case of SF fault using TCP}

### 2018-11-18 res/sections/08-Background.tex
Add part on network slicing

@@
 going to briefly introduce and explain the technologies involved and their role.
+
+\section{Network slicing}
+A network slice can be defined as a logical virtual network, tailored for a
+given use case, on the top a common network using technologies such as
+Software-Defined Networking and Network Function Chaining that allows network
+softwarization~\cite{ordonez2017network}. The notion of separated virtual
+networks deployed on the same infrastructure is not new, e.g. VPN, but it has
+some specifics that make it a novel idea. The concept of network slicing was
+first introduced in 2002 with the paper~\cite{peterson2003blueprint}, in which
+is described an overlay network that allow services to access different slices
+of overlay's resources and in which overlay management services run in their
+own slice. This idea was be refined during time and
+in~\cite{alliance2016description} the design of network slicing is based on a
+three-layers model, as depicted in
+Figure~\ref{chap:background:img:network_slicing}.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.2]{networkslicing}
+  \caption{The three layers model of network slicing}
+  \label{chap:background:img:network_slicing}
+\end{figure}
+
+The first layer is the \emph{Service Instance Layer} used to represent the
+business or user services supported, the second one is the \emph{Network Slice
+Instance Layer}. This is a set of network functions that allows to
+satisfy certain requirements~\cite{kotulski2017end}. Network functions, in
+general, can be both physical, thus a combination of vendor-specific hardware
+and/or virtualized. In that case the function is implemented as software and it
+is decoupled from the hardware on which it is running on. Each instance must be
+isolated, in order to ensure that function running at the top of it will always
+provide the same Quality of Service, without degrading because of a change on
+another slice~\cite{richart2016resource}.
+It also allows to provide better security.
+In fact, in case of attack or fault of a certain slice, others must not be affected.
+Moreover, each slice must implement it own set of security function to preventing unauthorized access.
+Furthermore, isolation let to manage individually each slice~\cite{ordonez2017network}.
+Network slice instances are defined by a \emph{Network Slice Blueprint}: a complete description of the structure and data flows that allow the creation and the control of the Network Slice Instance during its whole lifecycle.
+The blueprint also refers to the resources needed to assemble the network slice and defines its characteristics (e.g.
+ultra-low latency, ultra-reliability, value-added services for enterprises, etc.).
+The last layer is the \emph{Resource Layer}, containing both physical and logical resources, that involve all physical assets and resources for computation, storage or transport, including radio access.
 
@@
 \subsection{SDN stack}
-\begin{figure}[t]
+\begin{figure}[ht]
  \centering
@@
 
-The \emph{Application plane} role is to set the behaviour of the SDN controllers 
+The \emph{Application plane} role is to set the behavior of the SDN controllers 
 and manage its hardware resources and data through APIs, often called 

### 2018-11-18 res/sections/79-Conclusions.tex
Fix spellcheck on conclusions

@@
 the requirements of the whole implementation. Then we described the challenges 
-we encountered and how we solved it, taking choises valuating their pros and 
+we encountered and how we solved it, taking choices valuating their pros and 
 cons.
@@
 In conclusion, we can say that the ETSI MANO specification, coupled with the 
-RFC 7665 proposal and the new virtualization technologies avaiable in the 
+RFC 7665 proposal and the new virtualization technologies available in the 
 market could change the shape of future mobile communications, leading to the 
-creation of a 5G able to either saddisfy the challenges of the present, such as 
+creation of a 5G able to either satisfy the challenges of the present, such as 
 IoT communications, media streaming, and real-time data, with the challenges of 

### 2018-11-18 res/sections/25-Tests.tex
Fix some typos

--- 2018-11-16 res/sections/25-Tests.tex
@@
 
-After the testbed implementation, tests about its performance and metrics were
+After the test-bed implementation, tests about its performance and metrics were
 gathered in order to understand the real project effectiveness. We performed
@@
 
-\subsection{Docker vs Virtual Box startup measurement}
+\subsection{Docker vs Virtual Box start-up measurement}
 
@@
 To make this simulation as fair as possible, we used the same node (an 
-Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the
+Openstack VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the 
 same ``boot sequence'' for both the VNFs: first we launched the Astaire 
 framework, ensuring its availability to process data and then we notified the 
-test machine of the successful boot. The startup time was measured from the 
-beginning of the startup command (\verb!docker run! for Docker and 
+test machine of the successful boot. The start-up time was measured from the 
+beginning of the start-up command (\verb!docker run! for Docker and 
 \verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
-test backend (to make things as smooth as possible, \verb!netcat! was used 
+test back-end (to make things as smooth as possible, \verb!netcat! was used 
 to listen to incoming TCP data).
@@
 It is worth saying that Docker metrics are generally more precise than the
-Virtual Box one: Docker allows to inspect the container and to gain startup and
+Virtual Box one: Docker allows to inspect the container and to gain start-up and
 shutdown times with a precision of milliseconds, while to gather this 
@@
 \verb!time!, that registered timing with a precision of hundredths of a second. 
-Nonetheless,this didn't influence the test too much. To reduce possible
-outliers we performed
-the test $100$ times. With this $N$, the total time spent by the Virtual
-Machines to boot up was of $2021,15$ seconds, while with containers the same
-number of start ups resulted in a total of $1.638$ seconds. Summing up, Docker
-employs only the $0,08\%$ of the overall time spent for the test, underlining
-how isolating the process with a set of defined policy and avoiding a
-virtualized kernel boot-up saves a considerable amount of time.
+Nonetheless, this didn't influence the test too much. To reduce possible 
+outliers we performed the test $100$ times. With this $N$, the total time spent 
+by the Virtual Machines to boot up was of $2021,15$ seconds, while with 
+containers the same number of start ups resulted in a total of $1.638$ seconds. 
+Summing up, Docker employs only the $0,08\%$ of the overall time spent for the 
+test, underlining how isolating the process with a set of defined policy and 
+avoiding a virtualized kernel boot-up saves a considerable amount of time.
 
@@
         \includegraphics[scale=0.4]{docker_vs_vmBarplotGraph}
-        \caption{Barplot between Virtual Box and Docker startup times.}
+        \caption{Barplot between Virtual Box and Docker start-up times.}
         \label{chap:tests:sec:dockervsvb:img:barplot}
@@
 \begin{tabular}{l|c|c}
-                     & \textbf{Standard deviation (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
-\textbf{Docker}      & 0.009620874                           & 0,016                                    \\
+                     & \textbf{Standard deviation (seconds)} & \textbf{Average 
+start up time (seconds)} \\ \hline
+\textbf{Docker}      & 0.009620874                           & 0,016            
+                        \\
 \textbf{Virtual Box} & 0.755                                 & 20,211
+
 \end{tabular}%
@@
 \caption[Docker vs Virtual Box start up times comparison]{Docker vs Virtual Box
-  start up times comparison. Is possible to see how the overall time required to
+  start up times comparison. Is possible to se how the overall time required to
   start $100$ containers is less than the time to start one traditional virtual
@@
 
-% The overall time required by the traitional virtualization system is
-% approximately of 33 minutes, witch gives a significant insight of how much time
-% is required to create and boot a virtualized enviroment.
+% The overall time required by the traditional virtualization system is
+% approximately of 33 minutes, witch gives a significant insight of how much 
+time
+% is required to create and boot a virtualized environment.
 
@@
 environment. The deployment consists in a Pod inside a Deployment definition
-with a related Service, to simulate a real VNF. After the Pod startup, a request
-is made to a backend designed to count the chain elements that have successfully
+with a related Service, to simulate a real VNF. After the Pod start-up, a 
+request
+is made to a back-end designed to count the chain elements that have 
+successfully
 boot-up returning, eventually, the overall start up timing.
@@
     noticeable how there seems to be a linear correlation between the SFC
-    startup time and the SFC length, even though for the chain with eight
+    start-up time and the SFC length, even though for the chain with eight
     elements it is possible to see that the overall deploy time is more compared
@@
 software in a production environment, eliminating the possibility of human
-errors and assuring longer uptimes.
-
-\section{System performances}
-
-\subsection{Test system}
-All the machine used for testing purposes are Openstack virtual machines, with
-the following specifications:
-\begin{itemize}
-\item 1 machine with 32GB of RAM, 8vCPU and solid state storage, on which the
-receiver of packets is running;
-\item 4 machines with 16GB of RAM, 6vCPU and solid state storage, that form a
-Kubernetes cluster on which SFC infrastructure is deployed;
-\item 1 machine with 4GB of RAM, 2vCPU and solid state storage, on which
-scripts to send packets run.
-\end{itemize}
-
-\subsection{RTT measurement}
-Regarding the delivery performance, a metric to take in account is the
-\emph{round trip time}(RTT), that is the time interval between the instant in
-which a packet is sent and the response is received. To test it there was
-developed a script that sends a packet with 8B of payload and waits for the
-correspondent response, recording the instants (expressed as \emph{epoch}) of
-send and receive. The sake of this test is to compare the RTT measured without
-any SFC (only sending the packet to the receiver waiting for the response) and
-making packets pass through different length SFCs, in particular with 1, 2, 3,
-4, 5, and 10 hops. Packets traverse the SFC on both sides and the results are
-collected after 100 tests and the SF only prints the packet received and the
-total number of packets that have passed through it.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{@{}cc@{}}
-\toprule
-\textbf{SFC length} & \textbf{Mean RTT (milliseconds)} \\ \midrule
-len. 0              & 0.0009668279      \\
-len. 1              & 0.0555884027      \\
-len. 2              & 0.0702438903      \\
-len. 3              & 0.0888472247      \\
-len. 4              & 0.1045638967      \\
-len. 5              & 0.1239343858      \\
-len. 10             & 0.2135494161      \\ \bottomrule
-\end{tabular}
-\caption{Mean RTT measured with different SFC length}
-\label{chap:tests:sec:rtt:tab:meanrtt}
-\end{table}
-
-Not surprisingly, as depicted in the Table~\ref{chap:tests:sec:rtt:tab:meanrtt},
-the RTT increase with the number of elements of the chain. There is a big
-difference between the mean RTT of the configuration without any chain and the
-one with an SFC with a single link: this outcome has different explanation. As
-explained in \todo{Add reference to implementation} the chain has always at
-least 3 hops (ingress, SF, egress), so the packet must be elaborated by 3
-different running software. Another explanation lays into how the code was
-developed: it was not developed as a production ready product, but for a
-feasibility test, so connections and transmissions can be enhanced. Finally, SFs
-are meant to perform some modifications to the packet, so, unavoidably, some
-time is wasted.
-
-\begin{figure}[H]
-  %\centering
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[scale=0.4]{meanrtt}
-        \caption{Mean RTT calculate using SFCs of different lengths}
-        \label{chap:tests:sec:rtt:img:meanstt}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[b]{0.45\textwidth}
-        \includegraphics[scale=0.4]{differencenosfc}
-        \caption{Time increase using SFCs of different lengths}
-        \label{chap:tests:sec:rtt:img:differencertt}
-    \end{subfigure}
-\end{figure}
-
-The results are better represented in
-Figure~\ref{chap:tests:sec:rtt:img:meanstt}. In that graph we can easier see
-the difference among the chains with different lengths and the transmission that
-does not use any SFC. Looking at the bars representing the SFC composed by 1,
-2, 3, 4 and 5 links we can see a pretty linear increase of the RTT. This trend
-strongly depends on the nature of the SF involved in the SFC, in that case the
-linear increase is highlighted due to the use of the same function. In
-Figure~\ref{chap:tests:sec:rtt:img:differencertt} is depicted the difference
-between the RTT measured with a normal packet transmission and using the chains
-with different size, that represent the additional time needed for each
-transmission. Even in that case we can see that some time is wasted against a
-normal transmission and that this delta of time depends on the number of SFC
-involved in the chain.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.4]{differenceonehop}
-  \caption{Time difference between SFCs that have a length difference of one SF}
-  \label{chap:tests:sec:rtt:img:differenceonehop}
-\end{figure}
-
-In Figure~\ref{chap:tests:sec:rtt:img:differenceonehop} is represented the
-difference in terms of RTT among chains that have a length difference of one
-link. The first bar represent the comparison between an SFC composed by only
-one element and the direct transmission to the receiver: that explain why that
-value is higher than the others. Other differences among SFC, instead, are
-quite similar.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-\centering
-\begin{tabular}{@{}cc@{}}
-\toprule
-\textbf{SFC length} & \textbf{Jitter (Milliseconds)} \\ \midrule
-len. 0              & 0.0001074767    \\
-len. 1              & 0.0028695822    \\
-len. 2              & 0.0032037902    \\
-len. 3              & 0.0031961584    \\
-len. 4              & 0.0027278513    \\
-len. 5              & 0.0036802184    \\
-len. 10             & 0.0063434506    \\ \bottomrule
-\end{tabular}
-\caption{Jitter for different SFC lengths}
-\label{chap:tests:sec:rtt:tab:jitter}
-\end{table}
-
-Table~\ref{chap:tests:sec:rtt:tab:jitter} represent the jitter calculated for
-different SFC length, based on values of RTT measured in the previous test.
-Jitter is the variation in the delay of received packet, and gives a measure of
-the congestion of the network. As we can see values are really low, even if the
-lowest result is always the one calculated without the chain.
+errors and assuring longer up-times.

### 2018-11-19 res/sections/08-Background.tex
80 columns, preliminar check

@@
 some specifics that make it a novel idea. The concept of network slicing was
-first introduced in 2002 with the paper~\cite{peterson2003blueprint}, in which
-is described an overlay network that allow services to access different slices
+first introduced in 2002 with the in\cite{peterson2003blueprint}, where
+is described an overlay network that allows services to access different slices
 of overlay's resources and in which overlay management services run in their
-own slice. This idea was be refined during time and
-in~\cite{alliance2016description} the design of network slicing is based on a
+own slice. Lately this idea has be refined  and
+in\cite{alliance2016description} the design of network slicing was based on a
 three-layers model, as depicted in
@@
 provide the same Quality of Service, without degrading because of a change on
-another slice~\cite{richart2016resource}.
-It also allows to provide better security.
-In fact, in case of attack or fault of a certain slice, others must not be affected.
-Moreover, each slice must implement it own set of security function to preventing unauthorized access.
-Furthermore, isolation let to manage individually each slice~\cite{ordonez2017network}.
-Network slice instances are defined by a \emph{Network Slice Blueprint}: a complete description of the structure and data flows that allow the creation and the control of the Network Slice Instance during its whole lifecycle.
-The blueprint also refers to the resources needed to assemble the network slice and defines its characteristics (e.g.
-ultra-low latency, ultra-reliability, value-added services for enterprises, etc.).
-The last layer is the \emph{Resource Layer}, containing both physical and logical resources, that involve all physical assets and resources for computation, storage or transport, including radio access.
+another slice~\cite{richart2016resource}. It also allows to provide better 
+security. In fact, in case of attack or fault of a certain slice, others must 
+not be affected. Moreover, each slice must implement it own set of security 
+function to preventing unauthorized access. Furthermore, isolation let to manage 
+individually each slice~\cite{ordonez2017network}. Network slice instances are 
+defined by a \emph{Network Slice Blueprint}: a complete description of the 
+structure and data flows that allow the creation and the control of the Network 
+Slice Instance during its whole lifecycle. The blueprint also refers to the 
+resources needed to assemble the network slice and defines its characteristics 
+(e.g. ultra-low latency, ultra-reliability, value-added services for 
+enterprises, etc.). The last layer is the \emph{Resource Layer}, containing both 
+physical and logical resources, that involve all physical assets and resources 
+for computation, storage or transport, including radio access.
 
@@
 \textbf{LTE capability}
+
                                                                            & 
@@
 100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).
+
                                                                      & 10X 
@@
 Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
-civil aviation.                                                                & 
+civil aviation.                                                                
+& 
 1.5X                                                                  \\
@@
 Typically $\sim$2,000 active users/km2.
+
                                                                            & 

### 2018-11-19 res/sections/08-Background.tex
Fix spaces
Fix sentences

@@
 some specifics that make it a novel idea. The concept of network slicing was
-first introduced in 2002 with the in\cite{peterson2003blueprint}, where
-is described an overlay network that allows services to access different slices
-of overlay's resources and in which overlay management services run in their
-own slice. Lately this idea has be refined  and
-in\cite{alliance2016description} the design of network slicing was based on a
-three-layers model, as depicted in
+first introduced in 2002 in\cite{peterson2003blueprint}, where a network 
+overlay is illustrated, allowing services to run continuously and to access the 
+overlay resources across different slices. Services operating over the overlay 
+management have their own slice instead. This idea was lately refined in 2016, 
+where the design of network slicing was based on a three-layers 
+model~\cite{alliance2016description}, as depicted in 
 Figure~\ref{chap:background:img:network_slicing}.

### 2018-11-19 res/sections/08-Background.tex
Change part on SDN

--- 2018-11-18 res/sections/08-Background.tex
@@
   \includegraphics[scale=0.2]{networkslicing}
-  \caption{The three layers model of network slicing}
+  \caption{The three layers model of network
+  slicing~\cite{alliance2016description}}
   \label{chap:background:img:network_slicing}
@@
 provide the same Quality of Service, without degrading because of a change on
-another slice~\cite{richart2016resource}.
-It also allows to provide better security.
-In fact, in case of attack or fault of a certain slice, others must not be affected.
-Moreover, each slice must implement it own set of security function to preventing unauthorized access.
-Furthermore, isolation let to manage individually each slice~\cite{ordonez2017network}.
-Network slice instances are defined by a \emph{Network Slice Blueprint}: a complete description of the structure and data flows that allow the creation and the control of the Network Slice Instance during its whole lifecycle.
-The blueprint also refers to the resources needed to assemble the network slice and defines its characteristics (e.g.
-ultra-low latency, ultra-reliability, value-added services for enterprises, etc.).
-The last layer is the \emph{Resource Layer}, containing both physical and logical resources, that involve all physical assets and resources for computation, storage or transport, including radio access.
+another slice~\cite{richart2016resource}. It also allows to provide better
+security. In fact, in case of attack or fault of a certain slice, others must
+not be affected. Moreover, each slice must implement it own set of security
+function to preventing unauthorized access. Furthermore, isolation let to
+manage individually each slice~\cite{ordonez2017network}. Network slice
+instances are defined by a \emph{Network Slice Blueprint}: a complete
+description of the structure and data flows that allow the creation and the
+control of the Network Slice Instance during its whole lifecycle. The blueprint
+also refers to the resources needed to assemble the network slice and defines
+its characteristics (e.g. ultra-low latency, ultra-reliability, value-added
+services for enterprises, etc.). The last layer is the \emph{Resource Layer},
+containing both physical and logical resources, that involve all physical
+assets and resources for computation, storage or transport, including radio
+access.
 
 \section{Software-defined networking}
-The SDN is an emerging approach to cloud computing that allows to configure
+The SDN is an emerging approach to cloud computing and networking technology
+capable of supporting the dynamic nature of future network functions and
+intelligent applications while lowering operating costs through simplified
+hardware, software, and management~\cite{sezer2013we}, that allows to configure
 network in a programmatically way, possibly enhancing monitoring and
-performances. The main sake of SDN is to the packet forwarding process,
-\emph{data plane}, from the routing one, the so called \emph{control plane}.
-The latter component serves as the centralized network intelligence and could
-consists of one or more controllers. It has some drawbacks in terms of
+performances. One of the main principles of SDN architecture is to decouple the
+packet forwarding process, \emph{data plane}, from the routing one, the so
+called \emph{control plane}. The architecture also contemplate that network
+intelligence and state are logically centralized\cite{fundation2012software}.
+
+It has some drawbacks in terms of
 security, scalability and elasticity, because of the centralized architecture.
@@
 denote the three layer logic: the Application plane, the Controller plane and 
-the Data plane.}
+the Data plane~\cite{fundation2013software}.}
  \label{chap:background:img:sdn_architecture}
@@
 As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be split in three different layers, called planes, based on 
-their functionalities.
+architecture can be split in three different layers, also known as planes,
+based on their functionalities~\cite{fundation2012software} as explained
+in~\cite{fundation2013software}.
 
@@
 
-The \emph{Application plane} role is to set the behavior of the SDN controllers 
+In the \emph{Application plane} lays \emph{SDN applications} that are programs
+whose role is to set the behavior of the SDN controllers 
 and manage its hardware resources and data through APIs, often called 
-\emph{northbound interfaces} (NBI).
+\emph{northbound interfaces} (NBI). In order to do so they can also build an
+abstracted view of the network, using controllers information, in order to make
+the traffic analysis easier and to simplify the decision making process..
 These interfaces provide an additional abstraction layer that high-level 
 applications can consume, enabling the possibility to use a variety of 
-different vendors' controllers in the same way.
-Analyzing the SDN applications is possible to note two defined layers, one that 
-is responsible to manage the whole logic, and another one dedicated to the 
-communication part, that in particular uses the previously defined NBIs.
-These applications can also build an abstracted view of the network, using
-controllers information, in order to make the traffic analysis easier and to 
-simplify the decision making process.
+different vendors' controllers in the same way. Applications consists both of
+the two defined layers, one that is responsible to manage the whole logic, and
+another one dedicated to the communication part, that in particular uses the
+previously defined NBIs.
 
@@
 \emph{SDN controllers}, are logically centralized components that
-receive requests from the mentioned SDN applications and either translate them 
-in a fashion that can be understood from the lower layer\todo{who?}, or collect 
-data needed by the application itself. 
+receive requests from SDN applications and either translate them 
+in a fashion that can be understood from the SDN Datapath, or collect 
+data as statics and events needed by the application to build an abstract view
+of the network. 
 Due to the variety of tasks a controller has to accomplish, its logical 
@@
 NBI agent manages the communications with SDN applications, exposing interfaces
-that usually provide an abstract view of the network and a set of
-the controller functionalities.
-In particular, the middle layer represents the controller inner-logic, which 
-process the requests coming from the upper layer (aka the third layer), that 
-has the specific task to communicate with the remaining network components, in 
-other words the Control to Data-Plane Interface and/or the already 
-mentioned southbound interface.
+to access network data and to the set of the controller functionalities.
+In particular, the middle layer represents the \emph{SDN Controller Logic},
+which process the requests coming from the upper layer applications. Finally
+requests are communicated to the remaining network components, exploiting the
+Control to Data-Plane Interface driver.
+
+Despite the centralized design of controllers there are no limitations on the
+implementation, so it possible to create federations of controllers or organize
+them into a hierarchy.
 
 \subsubsection{Data plane}
-
-In the lowest part of the architecture we can find the \emph{SDN Networking 
-Devices}. These ones controls the packet forwarding and data processing 
-capabilities of the network. They consist of a CDPI agent and a set of one or 
-multiple traffic forwarding engines and of a zero or more traffic processing 
-functions, that are, as a matter of fact, the devices logic.
+In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
+network device, which exposes the possibility to access and control to its
+forwarding and processing capabilities.It is composed by a CDPI agent and a set
+of one or  multiple traffic forwarding engines and of a zero or more traffic
+processing functions, that are, as a matter of fact, the devices logic.
+Datapath can be mapped either to a single physical device or multiple ones and
+the implementation of the its logic does not preclude neither the
+virtualization or the sharing of physical resources nor the interoperability
+with non-SDN networking.
 
 \subsection{OpenFlow protocol}
-The SDN strategy is often associated with the OpenFlow protocol: this 
-communication protocol aims is to give access to the forwarding network 
-component to the lower layer network infrastructure, made of switches and 
+The SDN strategy is often associated with the OpenFlow protocol because it is
+the first standard communications interface defined between the control and the
+datapath in SDN architecture~\cite{fundation2013software}, making possible to
+access from controller the components of infrastructure, such as switches and
 routers. OpenFlow enables the remote administration of networks working on 

### 2018-11-19 res/sections/08-Background.tex

@@
 some specifics that make it a novel idea. The concept of network slicing was
-first introduced in 2002 with the paper~\cite{peterson2003blueprint}, in which
-is described an overlay network that allow services to access different slices
-of overlay's resources and in which overlay management services run in their
-own slice. This idea was be refined during time and
-in~\cite{alliance2016description} the design of network slicing is based on a
-three-layers model, as depicted in
+first introduced in 2002 in\cite{peterson2003blueprint}, where a network 
+overlay that allows services to run continuously and to access the overlay
+resources across different slices was illustrated.
+Services operating over the overlay management have their own slice instead.
+This idea was lately refined in 2016, where the design of network slicing was
+based on a three-layers model~\cite{alliance2016description}, as depicted in 
 Figure~\ref{chap:background:img:network_slicing}.

### 2018-11-19 res/sections/08-Background.tex
Reward virtualization and network virtualization adn add references

@@
 \section{Virtualization}
-Cloud computing is now a leading-edge Internet technology and the virtualization
-technology is one of the most important part of cloud computing systems.
-Virtualization, in computing, refers to the act of creating a virtual (rather
-than an actual) version of something, including but not limited to a virtual
-computer hardware platform, operating system (OS), storage device, or computer
-network resources. \todo{This part was taken from ``The research and
-implementation of cloud computing platform based on docker'' paper, rework it
-in a way it doesn't seem to be copied.}
+Cloud computing is nowadays one of the cutting edge Internet technologies, and
+it is possible thanks to virtualization, that allows to isolate applications,
+optimizing the utilization of hardware resources and providing more flexibility.
+Virtualization, in computing, usually refers to the act of creating a virtual
+version of a certain asset, such as virtual computer hardware platform,
+operating system (OS), storage device, or computer network
+resources~\cite{liu2014research}. Conventional virtualization
+(hypervisor-based) uses virtual machines: it provides strong isolation and a
+complete operating system~\cite{eder2016hypervisor}. In hypervisor
+virtualization every pieces of hardware required to running software is
+emulated, and it is common that the hypervisor provides interfaces to access
+the real hardware, for instance to access physical devices or to communicate
+with the network. However all this functionalities comes to a resource
+overhead~\cite{scheepers2014virtualization}Container-based virtualization,
+instead, could be a more efficient in certain context, because it does not
+emulate the entire computer. A container is seen as an isolate process running
+on the host, to which the Linux kernel gives a set of capabilities.
 
 \subsection{Network Virtualization}
-%https://s3.amazonaws.com/academia.edu.documents/29016225/cs-2008-25.pdf?AWSAcce
-%ssKeyId=AKIAIWOWYYGZ2Y53UL3A&Expires=1541352965&Signature=On6XJ5daSAmnfJMnNf41f
-%BNHFgg%3D&response-content-disposition=inline%3B%20filename%3DA_survey_of_netwo
-%r k _virtualization.pdf
+%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
+%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
 Virtualization, in the networking context, refers to the act of combining
@@
 heterogeneous networks to cohabit, increasing the flexibility and manageability.
-%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
-%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
 The idea of network virtualization came from the decoupling of the role of the
@@
 aggregate resources, even from different InPs, creating a virtual network,
-offering an end-to-end service. Most known instances of NV are \emph{VPNs},
-dedicated networks that allows the communication among multiple sites, using
+offering an end-to-end service~\cite{chowdhury2009network}. Most known
+instances of NV are \emph{VPNs}, dedicated networks that allows the
+communication among multiple sites, using
 private and secure tunnels, relying on the public network such as Internet.

### 2018-11-19 res/sections/09-RelatedWork.tex
Add some related works

\chapter{Related Work}
\label{chap:rel_wk}

The possibility to solve the constant network capacity requests and to make
network infrastracture lighter and faster to deploy has rouse the interests of
many internet service providers (ISP). In literacy, several NFV implementations
and Network Slicing solutions have been proposed and some of them will be
discussed in the following chapter. An opportunity identified
in~\cite{nogales2018nfv} shows that ETSI MANO implementations can be deployed
inside Unmanned Aerial Vehicles (UAVs), giving the possibility to rapidly deploy
network connectivity in areas where the common communication systems are
temporarily unavailable. This can be very useful in places where natural
disasters occurred and there is the necessity to restore the communications
within a very short span of time. On top of that, this system is flexible enough
to be able to provide different kind of services, in addition to simple
broadband connectivity. The only downside of this approach are the constrained
system rources: the authors had to install the VIM (in particular, Openstack) in
a separated machine, located on the ground. A solution based only on
containerization, instead, could lead to interesting results, since Kubernetes
does not require too much resources \textit{per se} and in literature is
possible to find solutions of Kubernetes installations performed on machines
with the same harware installed on the UAVs~\cite{pahl2016container}.

A virtualized EPC implementation has been describe in~\cite{hawilo2014nfv}
% TODO talk about vEPC

It is interesting to note that, between all the possibile and alternative
related works we have been albe to find, none of these used Docker as an
alternative virtualization approach.

### 2018-11-19 res/sections/08-Background.tex
Add part on NVFI and NFV challenges

@@
 \subsection{Network Virtualization}
-%http://www.csc.villanova.edu/~nadi/csc8580/S11/DeeptiNune.pdf
-%https://www.sciencedirect.com/science/article/abs/pii/S1389128609003387
 Virtualization, in the networking context, refers to the act of combining
@@
 The \emph{Network Function Virtualization} (NFV) concept refers to the use of 
-virtualization technologies to virtualize network nodes functions. VNFs may run 
-different processes and software, on the top of servers, switches, storage 
-devices or on cloud computing infrastructures, that substitute custom hardware. 
+virtualization technologies to virtualize network nodes functions. This
+emerging technology focuses on the decoupling of function logic and hardware,
+allowing VNFs to run on the top of servers, switches, storage 
+devices or on cloud computing infrastructures. In other word the software logic
+is separated from the custom platform on which it runs, enabling the use of
+commodity hardware~\cite{gray2016network}.
 That approach worked well in the past, because of the need to follow rigorous 
 standards, providing at the same time high quality, but it led to long product 
-cycles and high specialized proprietary hardware. New Internet scenarios, that 
+cycles and high specialized services that use a combination of software and
+hardware resources, typically vendor-specific, and only largely
+non-interoperable within other vendor solutions. New Internet scenarios, that 
 include IoT and the increasing number of services to spread multimedia data 
 (Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
-network deployments. The NFV technology is exploited in a framework, organized
-at least by three components:
+network deployments. 
+This detachment provides greater flexibility to scale VNF based on traffic and
+performances and also improve the automation of function deploy and update:
+removing dedicated hardware also will reduce costs machines and for
+maintenance.
+
+The NFV technology is exploited in a framework, organized
+at least by three components~\cite{etsi2013gs}, as represented in
+Figure~\ref{chap:background:img:etsi_arch}:
 \begin{enumerate}
- \item a software that performs the network function that must be deployed,
- that is the VNF;
- \item NFV Infrastructure (NFVI) that is composed by all hardware and software 
-components that builds the environment where the VNFs are deployed;
- \item NFV Management And Orchestration (NFV-MANO) architectural framework that
- is all the functional blocks, data repositories used by these blocks,
- reference points and interfaces through which these functional blocks exchange
+  \item a software, able to run on over the NFVI, that performs the network
+  function that must be deployed, that is the VNF;
+  \item NFV Infrastructure (NFVI) that is composed by all physical resources
+  and how these can be virtualized, that forms the environment where the VNFs
+  are deployed;
+  \item NFV Management And Orchestration (NFV-MANO) that covers the
+  orchestration and management of both physical and logical resources, the
+  functional blocks, data repositories used by these blocks, reference points
+  and interfaces through which these functional blocks exchange
   information for the purpose of managing and orchestrating NFVI and VNFs.
 \end{enumerate}
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=2]{etsi_arch}
+  \caption{Network Function Virtualization reference
+  architecture~\cite{etsi2013gs}}
+  \label{chap:background:img:etsi_arch}
+\end{figure}
+
+\subsection{Network Function Virtualized Infrastructure}
+As mentioned before, NFVI is the combination of physical and logical resources,
+that combined together create the environment on which the VNFs are deployed
+and run. Physical supplies can be proprietary systems, commercial-off-the-shelf
+hardware, software and network resources or cloud computing capabilities. The
+abstraction of resources is made taking advantage of a virtualization layer,
+that hides the underlying infrastructure. Storage and computing resources may
+be represented in terms of virtual machines or
+containers~\cite{mijumbi2016network}.
+
+\subsection{VNF challenges}
+This new concept open a plethora of challenges~\cite{han2015network}. Regarding
+performances, functions running on general purposes servers cloud affect
+throughput and latency, moreover the capacity of the virtualized function can
+be less than the hardware and software version. To reach comparable levels in
+terms of performances efficient algorithms are needed, as a excellent resource
+management. Another problems is the VNF management. In case of hardware
+functions resources are almost equivalent, at the contrary, virtualized ones
+could have costs and values that differs significantly depending on the point
+of presence on which are deployed. On the other hand, since VNFs are software
+components, can elastically scale up and down, depending on the traffic loads
+and user demand. Another management challenge refers to keeping track where the
+VNF is running, that is of paramount importance in case of issues. Related to
+possible dysfunctions, software function must reach a certain level of
+reliability. Generally, purpose-build network equipment can ensure the
+five-nines reliability requirement, that is more challenging to follow in that
+scenario due to the higher number of possible points of failure, such as during
+the scaling operation or VNF migrating from a point of presence to another.
+Finally even security must be taken into account. In fact, infrastructure on
+which VNFs run can be outsourced to third parties. Furthermore, orchestration
+and management components must implement some sort of intrusion prevention and
+detection system and the sharing of resources can introduce security threads.
 

### 2018-11-19 res/sections/09-RelatedWork.tex
Add preliminar part about related works

@@
 
-A virtualized EPC implementation has been describe in~\cite{hawilo2014nfv}
-% TODO talk about vEPC
+A virtualized EPC implementation, designed to allow mobile broadband services in
+the UMTS/LTE world, has been described in~\cite{hawilo2014nfv}, showing how this
+implementation is able, thanks to entities grouping, to perform better than a
+traditional EPC installation. In~\cite{afolabi2017end} is possible to read
+another vEPC implementation, where the Wireless Netowrk Virtualization (WNV),
+SDN and NFV (identified from the authors as key elements in order to implement
+the 5G architecture) are virtualized, and ported as a service (defined as
+EPCaaS) across multi-domains clouds. The realization of such framework is
+pursued thanks to end-to-end network slicing.
+% talk about end-to-end network slicing
 
 It is interesting to note that, between all the possibile and alternative
-related works we have been albe to find, none of these used Docker as an
-alternative virtualization approach.
+related works we have been albe to find, none of these used an only
+container-based approach.

### 2018-11-20 res/sections/08-Background.tex
Start SFC explaination

@@
   are deployed;
-  \item NFV Management And Orchestration (NFV-MANO) that covers the
+  \item NFV MANagement and Orchestration (NFV-MANO) that covers the
   orchestration and management of both physical and logical resources, the
@@
 
+\subsection{NFV MANagement and Orchestration}
+The MANO, represented in Figure~\ref{chap:background:img:etsimano}, can be splitted into three logical functional components:
+\begin{itemize}
+  \item the virtual infrastructure manager (VIM);
+  \item the VNF Manager (VNFM);
+  \item the NFV Orchestrator (NFVO).
+\end{itemize}
+And into four data repositories:
+\begin{itemize}
+  \item Network Service (NS) catalog;
+  \item VNF catalog;
+  \item NFV instances;
+  \item NFVI resources.
+\end{itemize}
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.3]{etsimano}
+  \caption{NFV MANO components}
+  \label{chap:background:img:etsimano}
+\end{figure}
+
+\paragraph*{VIM}
+The VIM aim is to manage and control NFVI resources.
+Every MANO can contain more VIM instances, each of them control NFVI resources from a single infrastructure.
+VIM can be developed either to administer one type of NFVI or more.
+VIM have to monitor the allocation of virtual resource in order to orchestrate allocation, upgrade, release, and reclamation of resources and optimize their usage, including virtual links, networks, subnets, and ports.
+It is also the component in charge to manage security group policies to ensure access control.
+In addition, VIM also has to record data on performance and faults, that can be queried.
+Thank to this component it is possible to dynamically allocate resource, based on demand, that is particularly using IaaS.
+
+\paragraph*{VNFM} 
+In this framework all VNFs, once created, have to register to a VNFM. This
+component is accountable to manage their lifecycle, that consists of
+instantiation, scaling, termination and, possibly, update/upgrade, and
+communicate the state of a certain VNF to others MANO blocks: this enable the
+monitoring of key performance indicators (KPI), that can help to make decision
+on VNF scaling. VNFM can manage a single type of VNF instances or multiple.
+
+\paragraph*{NFVO}
+The NFVO is responsible of combining functions to create end-to-end services.
+Its functionalities are to ensure there are adequate compute, storage, and
+network resources available to provide a network service independently of the
+VIM and the creation of end-to-end services by composing different VNFs, that
+can be deployed on different point-of-presence and registered to different VNF.
+It also managing the topology of these services.
+
+\paragraph*{Data repositories} Data repositories are use to keep information on
+the NFV MANO. The NS catalog is a set of predefined templates, which define
+services, functions involved and their connectivity. The VNF catalog is,
+instead, a set of templates describing the deployment and characteristics of
+VNFs. The NFVI resources repository holds information about available/allocated
+NFVI resources. The NFV instances repository keep information about functions
+and services instances during their lifetimes.
+
 \subsection{VNF challenges}
@@
 detection system and the sharing of resources can introduce security threads.
+
+\section{Service Function Chaining}
+The Service Function Chaining (SFC) concept was defined  in~\cite{halpern2015service} and refers to an ordered set of Service Functions (SF) and the constraint, defined after a classification process, to packets and flows.
+SF are the particular functions whose aim is the treatment of packets received.
+They can work on different levels of the network layer.
+These components can be implemented as virtual element or a combination of a physical network element and the software running on the top of it.
+There are no limitations on the number of SFs running on the same network element, so in general it is possible to have one or more functions running on the same element.
+SFs can be aware or not that are part of a certain chain, in the matter case they are called \emph{SFC aware} functions, in the latter \emph{SFC unaware}.
+
+An SFC can be seen as a graph in which each node is an SF and edges are the path through which a packet have to pass through.
+Because of this definitions there are no limitation on possible cycles that can occur in the chain, in case that a certain SF have to be reapplied on the same packet, frame or flow.
+These scenarios must be adequately managed to avoid endless loops.
+This enable even the possibility of branching on a certain node in order to define the treatment that has to be imposed from that point of the chain.
+The graph representation also permit to think to \emph{unidirectional} and \emph{bidirectional} chains.
+An SFC is defined unidirectional if traffic can traverse it only in one direction, so if it is composed by 3 SF, named \texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go from \texttt{SF0} to \texttt{SF2}.
+On the contrary, a bidirectional SFC allows traffic in both directions.
+So, having the same chain as in the previous example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from \texttt{SF2} to \texttt{SF0}.
 

### 2018-11-20 res/sections/09-RelatedWork.tex
Add new related works

@@
 
-The possibility to solve the constant network capacity requests and to make
-network infrastracture lighter and faster to deploy has rouse the interests of
-many internet service providers (ISP). In literacy, several NFV implementations
-and Network Slicing solutions have been proposed and some of them will be
-discussed in the following chapter. An opportunity identified
-in~\cite{nogales2018nfv} shows that ETSI MANO implementations can be deployed
-inside Unmanned Aerial Vehicles (UAVs), giving the possibility to rapidly deploy
-network connectivity in areas where the common communication systems are
-temporarily unavailable. This can be very useful in places where natural
-disasters occurred and there is the necessity to restore the communications
-within a very short span of time. On top of that, this system is flexible enough
-to be able to provide different kind of services, in addition to simple
-broadband connectivity. The only downside of this approach are the constrained
-system rources: the authors had to install the VIM (in particular, Openstack) in
-a separated machine, located on the ground. A solution based only on
-containerization, instead, could lead to interesting results, since Kubernetes
-does not require too much resources \textit{per se} and in literature is
-possible to find solutions of Kubernetes installations performed on machines
+The possibility to solve the increasing network capacity necessities and to make
+the network infrastracture lighter and faster to deploy has rouse the interests
+of many internet service providers (ISP).
+
+\section{NFV alternative implementations}
+
+In literacy, several NFV implementations and Network Slicing solutions have been
+proposed and some of them will be discussed in the following chapter. For
+example, an opportunity identified in~\cite{nogales2018nfv} shows that ETSI MANO
+implementations can be deployed inside Unmanned Aerial Vehicles (UAVs), giving
+the possibility to rapidly deploy network connectivity in areas where the common
+communication systems are temporarily unavailable. This can be very useful in
+places where natural disasters occurred and there is the necessity to restore
+the communications within a very short span of time. On top of that, this system
+is flexible enough to be able to provide different kind of services, in addition
+to simple broadband connectivity. The only downside of this approach are the
+constrained system rources: the authors had to install the VIM (in particular,
+Openstack) in a separated machine, located on the ground. A solution based only
+on containerization, instead, could lead to interesting results, since
+Kubernetes does not require too much resources \textit{per se} and in literature
+is possible to find solutions of Kubernetes installations performed on machines
 with the same harware installed on the UAVs~\cite{pahl2016container}.
+
+\section{EPC virtual implementations}
 
@@
 pursued thanks to end-to-end network slicing.
-% talk about end-to-end network slicing
 
-It is interesting to note that, between all the possibile and alternative
+EPC and ETSI MANO defitinions are not the only solutions to have been explored
+at this time. An SDN/NFV-based end to end network slicing for 5G has been
+created in~\cite{chartsias2017sdn}, where the authors provided examples to
+demonstrate how end-to-end services can be provisioned using SDN networking, in
+particular with controllers like such as OpenDaylight, Pox, Ryu ONOD and
+Floodlight. The last one was studied by us too during the technologies analysis,
+and a description is available in
+Section~\ref{chap:prjan:sec:tech:sub:other:sub:floodlight}.
+%TODO section missing...write it down!!
+
+\section{Network slicing security concerns}
+
+Another important aspect related to 5G networking and network slice is the one
+regarding security. Security in this context plays a fundamental role, since
+secure communications depends on the security of the overall system.
+In~\cite{kotulski2017end} the authors carry out a deep analysis about the
+network slice security, putting an emphasis on how this new element (the network
+slice) brings additional security questions and problems, especially about
+isolation. Moreover, slicing in 5G introduces the concept of shared services,
+that is contrary with the definition of isolation itself. In particular, resouce
+isolation gives the possibility to avoid attacks such us slice resources
+exaustion. Additionally, the authors give a critic regarding the ETSI NFV-MANO
+standard making too general assumptions regarding the whole system. Finally, a
+Service-Level Agreement (SLA) based orchestration criteria between network
+slices is proposed as a viable solution to solve the secutiry problems about
+isolation.
+
+\vspace{1cm}
+
+\noindent It is interesting to note that, between all the possibile and alternative
 related works we have been albe to find, none of these used an only

### 2018-11-20 res/sections/08-Background.tex
Add part on encapsulation and continue part on SFC

@@
 \subsection{NFV MANagement and Orchestration}
-The MANO, represented in Figure~\ref{chap:background:img:etsimano}, can be splitted into three logical functional components:
+The MANO, represented in Figure~\ref{chap:background:img:etsimano}, can be
+splitted into three logical functional components:
 \begin{itemize}
@@
 \paragraph*{VIM}
-The VIM aim is to manage and control NFVI resources.
-Every MANO can contain more VIM instances, each of them control NFVI resources from a single infrastructure.
-VIM can be developed either to administer one type of NFVI or more.
-VIM have to monitor the allocation of virtual resource in order to orchestrate allocation, upgrade, release, and reclamation of resources and optimize their usage, including virtual links, networks, subnets, and ports.
-It is also the component in charge to manage security group policies to ensure access control.
-In addition, VIM also has to record data on performance and faults, that can be queried.
-Thank to this component it is possible to dynamically allocate resource, based on demand, that is particularly using IaaS.
+The VIM aim is to manage and control NFVI resources. Every MANO can contain more
+VIM instances, each of them control NFVI resources from a single infrastructure.
+VIM can be developed either to administer one type of NFVI or more. VIM have to
+monitor the allocation of virtual resource in order to orchestrate allocation,
+upgrade, release, and reclamation of resources and optimize their usage,
+including virtual links, networks, subnets, and ports. It is also the component
+in charge to manage security group policies to ensure access control. In
+addition, VIM also has to record data on performance and faults, that can be
+queried. Thank to this component it is possible to dynamically allocate
+resource, based on demand, that is particularly using IaaS.
 
@@
 \section{Service Function Chaining}
-The Service Function Chaining (SFC) concept was defined  in~\cite{halpern2015service} and refers to an ordered set of Service Functions (SF) and the constraint, defined after a classification process, to packets and flows.
-SF are the particular functions whose aim is the treatment of packets received.
-They can work on different levels of the network layer.
-These components can be implemented as virtual element or a combination of a physical network element and the software running on the top of it.
-There are no limitations on the number of SFs running on the same network element, so in general it is possible to have one or more functions running on the same element.
-SFs can be aware or not that are part of a certain chain, in the matter case they are called \emph{SFC aware} functions, in the latter \emph{SFC unaware}.
+The Service Function Chaining (SFC) concept was defined 
+in~\cite{halpern2015service} and refers to an ordered set of Service Functions
+(SF) and the constraint, defined after a classification process, to packets and
+flows. SF are the particular functions whose aim is the treatment of packets
+received. They can work on different levels of the network layer. These
+components can be implemented as virtual element or a combination of a physical
+network element and the software running on the top of it. There are no
+limitations on the number of SFs running on the same network element, so in
+general it is possible to have one or more functions running on the same
+element. SFs can be aware or not that are part of a certain chain, in the
+matter case they are called \emph{SFC aware} functions, in the latter one \emph
+SFC unaware}.
 
-An SFC can be seen as a graph in which each node is an SF and edges are the path through which a packet have to pass through.
-Because of this definitions there are no limitation on possible cycles that can occur in the chain, in case that a certain SF have to be reapplied on the same packet, frame or flow.
-These scenarios must be adequately managed to avoid endless loops.
-This enable even the possibility of branching on a certain node in order to define the treatment that has to be imposed from that point of the chain.
-The graph representation also permit to think to \emph{unidirectional} and \emph{bidirectional} chains.
-An SFC is defined unidirectional if traffic can traverse it only in one direction, so if it is composed by 3 SF, named \texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go from \texttt{SF0} to \texttt{SF2}.
-On the contrary, a bidirectional SFC allows traffic in both directions.
-So, having the same chain as in the previous example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from \texttt{SF2} to \texttt{SF0}.
+\subsection{Architecture principles}
+The SFC technology was though to pursue some concept. One of the most important
+principles is the \emph{decoupling}. Decoupling in terms on independence on the
+infrastructure and the network, so no change must be performed to the
+forwarding underlying topology to deploy SFCs, and in terms of separation
+between packet handling (by SFs) and forwarding. This decoupling refers even to
+architecture and functions: one must be independent from the other. Another
+main point is the possibility to dynamically set routes on the incoming traffic
+to decide the best treatment based on its typology. The classification can
+occur on different point of the chain and can be updated during the packet
+delivery. Finally creation, modification and deletion of a certain SFC must has
+To impact on the others.
+
+\subsection{Graph view of an SFC}
+An SFC can be seen as a graph in which each node is an SF and edges are the
+path through which a packet have to pass through. Because of this definitions
+there are no limitation on possible cycles that can occur in the chain, in case
+that a certain SF have to be reapplied on the same packet, frame or flow. These
+scenarios must be adequately managed to avoid endless loops. This enable even
+the possibility of branching on a certain node in order to define the treatment
+that has to be imposed from that point of the chain. Traffic path in the SFC
+graph are managed and described by the \emph{Service Function Forwarder} (SFF)
+and \emph{Service Function Path} (SFP) components. The matter is the element of
+the architecture responsible of traffic forwarding. The main goals are to
+redirect traffic according to the SFC specified for a specific packet flow,
+query the classifier if needed and manage traffic coming back from a SF. The
+SFP, instead, is an indirection layer between the fully abstract notion of SFC
+and the list of SF that the traffic has to traverse. It is used for dictate the
+constraint specification on the path of where packets assigned to a certain
+service function path must go. There are no limitation on how specific it must
+be: it can specify the exact locations, identifying both the SFF and the SF, or
+only give some recommendation or, for example, defining only the SFF and
+demanding to it the decision on the SF instance.
+
+The graph representation also permit to think to \emph{unidirectional} and
+\emph{bidirectional} chains. An SFC is defined unidirectional if traffic can
+traverse it only in one direction, so if it is composed by 3 SF, named
+\texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go
+from \texttt{SF0} to \texttt{SF2}. On the contrary, a bidirectional SFC allows
+traffic in both directions. So, having the same chain as in the previous
+example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from
+\texttt{SF2} to \texttt{SF0}.
+
+\subsection{SFC Encapsulation}
+In order to exchange packet and information in the SFC ecosystem, traffic is
+encapsulated. The so called SFC encapsulation carries data used by the framework
+to identify SFPs, but it differs from a transport encapsulation and it is
+independent from the transport one: as such any transport protocol can be
+exploited to forward traffic.
+
+\todo{add sfc proxy}
+
+The \emph{Network Service Header} (NSH), used by SFC environment for
+encapsulation is described in~\cite{rfc8300} and consider three main fields as
+represent in Figure~\ref{chap:background:img:nshformat}: Base Header, Service
+Path Header and Context Header.
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.6]{nshformat}
+  \caption{Network Service Header~\cite{rfc8300}}
+  \label{chap:background:img:nshformat}
+\end{figure}
+
+\subsubsection{Base header}
+The Base Header utility is provide information about the service header and the
+payload protocol.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.6]{nshbase}
+  \caption{Network Service Header~\cite{rfc8300}}
+  \label{chap:background:img:nshformat}
+\end{figure}
+
+Fields are:
+\begin{description}
+  \item[Version]: This field is used to set the version of the header, used for
+  ensuring backward compatibility through different NSH definitions;
+  \item[0 bit]: Bit set only in case of orchestration or management packets, and
+  must not be changed by the SFC;
+  \item[TTL]: Time To Live, indicate the maximum number of hops among SFF and
+  its initial value should be configurable by the control plane. At each hop it
+  must be decremented to 1 until 0: in the latter case the packet must be
+  discarded. The default value is 63. This field is useful in loop prevention
+  mechanism. The default value is 63;
+  \item[Length]: the total length of the header (in 4-byte words), including
+  the \texttt{Fixed-Length Context Header} in NSH type 0x1 or the
+  \texttt{Variable-Length Context Header} in NSH type 0x2. In case of variable
+  length header data must be padded to a multiple of 4;
+  \item[Unassigned bits]: unassigned bits for future usages;
+  \item[MD]: Metadata Type, indicate the type of the NSH. Values that now
+  accepted are \texttt{0x0} that is a reserved value and if set the packet must
+  be discarded, \texttt{0x1} indicate that context header length is
+  fixed~\ref{subsubsec:background:flch}, \texttt{0x2} indicate that context
+  header length is variable~\ref{subsubsec:background:vlch} and \texttt{0xF} for
+  testing purposes (if not part of the experiment the packet must be discarded).
+  Base header length and Service Path Header have always the same size;
+  \item[Next Protocol]: indicates the protocol type of encapsulated data
+  (protocols codes can be found at~\cite{rfc8300}).
+\end{description}
+
+\subsubsection{Service Path Header}
+Provides path identification and location within a service path.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.6]{nshservicepath}
+  \caption{NSH Service Path Header~\cite{rfc8300}}
+  \label{chap:background:img:nshpath}
+\end{figure}
+
+\begin{description}
+  \item[Service Path Identifier (SPI)]: 24 bits, identify unambiguously an SFP.
+  The first classifier have to set it;
+  \item[Service Index (SI)]: 8 bits, provide location within the SFP. Initially
+  must be set to 255 and it must be decremented on each hop.
+\end{description}
+
+SPI combined to SI determine the next SF on each link in the chain. SI change
+provide a minimal loop prevention mechanism.
+
+\subsubsection{NSH with Fixed-Length Context Header}
+\label{subsubsec:background:flch}
+The NSH of type \texttt{0x1} have a 16-Byte Fixed-Length Context Header that
+can be used to share data along the service path.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.6]{nshtype1}
+  \caption{NSH MD Type \texttt{0x1}~\cite{rfc8300}}
+  \label{chap:background:img:nshtype1}
+\end{figure}
+
+The specification does not specify the content of this field and if SF or SFC
+Proxy does not known its semantics they have to discard the entire packet.
+
+\subsubsection{NSH with Variable-Length Context Header}
+\label{subsubsec:background:vlch}
+In NSH of type \texttt{0x2}, zero or more Variable-Length Context Headers may
+be added, but lengths must always be a multiple of 4-bytes (otherwise they must
+be padded).
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.6]{nshtype2}
+  \caption{NSH MD Type \texttt{0x2}~\cite{rfc8300}}
+  \label{chap:background:img:nshtype2}
+\end{figure}
+
+The structure of the variable length part is depicted in
+Figure~\ref{chap:background:img:nshtype2variable}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.6]{nshtype2variable}
+  \caption{Variable-Length Context Header~\cite{rfc8300}}
+  \label{chap:background:img:nshtype2variable}
+\end{figure}
+
+\begin{description}
+  \item[Metadata Class (MD Class)]: field used in order to provide hierarchical
+  namespaces for standards, vendors or other scopes;
+  \item[Type]: indicates the type of metadata, depending on the MD Class;
+\end{description}
+
+\subsection{Classification of traffic}
+In order to define the path of which traffic have to be forwarded, it must be
+classified. The classification granularity is not specified and can occur even
+during the SFC traverse. At the ingress of SFC environment packets have to be
+classified to determine to choose an SFC and eventually, if the classification
+granularity is too coarse, can be repeated after the treatment of some SFs. The
+reclassification, also called branching, may result in a different SFP choice or
+only to a metadata change.
+
+\subsection{Considerations}
 

### 2018-11-20 res/sections/09-RelatedWork.tex
80 columns
Fix spell check errors

@@
 The possibility to solve the increasing network capacity necessities and to make
-the network infrastracture lighter and faster to deploy has rouse the interests
+the network infrastructure lighter and faster to deploy has rouse the interests
 of many internet service providers (ISP).
@@
 to simple broadband connectivity. The only downside of this approach are the
-constrained system rources: the authors had to install the VIM (in particular,
+constrained system sources: the authors had to install the VIM (in particular,
 Openstack) in a separated machine, located on the ground. A solution based only
@@
 is possible to find solutions of Kubernetes installations performed on machines
-with the same harware installed on the UAVs~\cite{pahl2016container}.
+with the same hardware installed on the UAVs~\cite{pahl2016container}.
 
@@
 traditional EPC installation. In~\cite{afolabi2017end} is possible to read
-another vEPC implementation, where the Wireless Netowrk Virtualization (WNV),
+another vEPC implementation, where the Wireless Network Virtualization (WNV),
 SDN and NFV (identified from the authors as key elements in order to implement
@@
 
-EPC and ETSI MANO defitinions are not the only solutions to have been explored
+EPC and ETSI MANO definitions are not the only solutions to have been explored
 at this time. An SDN/NFV-based end to end network slicing for 5G has been
@@
 isolation. Moreover, slicing in 5G introduces the concept of shared services,
-that is contrary with the definition of isolation itself. In particular, resouce
-isolation gives the possibility to avoid attacks such us slice resources
-exaustion. Additionally, the authors give a critic regarding the ETSI NFV-MANO
-standard making too general assumptions regarding the whole system. Finally, a
-Service-Level Agreement (SLA) based orchestration criteria between network
-slices is proposed as a viable solution to solve the secutiry problems about
-isolation.
+that is contrary with the definition of isolation itself. In particular, 
+resource isolation gives the possibility to avoid attacks such us slice 
+resources exhaustion. Additionally, the authors give a critic regarding the 
+ETSI NFV-MANO standard making too general assumptions regarding the whole 
+system. Finally, a Service-Level Agreement (SLA) based orchestration criteria 
+between network slices is proposed as a viable solution to solve the security 
+problems about isolation.
 
@@
 
-\noindent It is interesting to note that, between all the possibile and alternative
-related works we have been albe to find, none of these used an only
+\noindent It is interesting to note that, between all the possible and 
+alternative related works we have been able to find, none of these used an only
 container-based approach.

### 2018-11-20 res/sections/09-RelatedWork.tex
Add new related works

@@
 EPC and ETSI MANO definitions are not the only solutions to have been explored
-at this time. An SDN/NFV-based end to end network slicing for 5G has been
+at this time. An SDN/NFV-based end-to-end network slicing for 5G has been
 created in~\cite{chartsias2017sdn}, where the authors provided examples to
@@
 %TODO section missing...write it down!!
+
+Compared to our solution, these implementations does not present an unified
+framework that gives the possbility orchestrate different network slices
+consistently. Adopting Kubernetes gives the advantage to use the Kubernetes
+Federation feature~\cite{kubeFederation}, enabling the possibility to simplify
+the management of multiple network slice in end-to-end services.
 
@@
 resource isolation gives the possibility to avoid attacks such us slice
-resources exhaustion. Additionally, the authors give a critic regarding the 
-ETSI NFV-MANO standard making too general assumptions regarding the whole 
-system. Finally, a Service-Level Agreement (SLA) based orchestration criteria 
-between network slices is proposed as a viable solution to solve the security 
-problems about isolation.
+resources exhaustion. Additionally, the authors give a critic regarding the ETSI
+NFV-MANO standard making too general assumptions regarding the whole system.
+Finally, a Service-Level Agreement (SLA) based orchestration criteria between
+network slices is proposed as a viable solution to solve the security problems
+about isolation. We tried to mitigate the security consern presented above using
+the integrated Role Base Access Control system (RBAC) and other security policy
+bundled in by default in Kubernetes.
+
+\section{Provision capabilities}
+We noted how, in some other NFV/ETSI-MANO proposals like~\cite{soares2015toward}
+and~\cite{abujoda2015midas}, the implementations relays on a static provision
+system, where the SFC paths are somehow defined at deploy time, and can't change
+without decomissioning, editing and deploying the chain again. Our proposal,
+instead, is able to support chain editing on-the-fly, allowing NFV to dynamically
+change SFP.
 

### 2018-11-20 res/sections/08-Background.tex
Add part on SFC proxy and other considerations

@@
 components can be implemented as virtual element or a combination of a physical
-network element and the software running on the top of it. There are no
-limitations on the number of SFs running on the same network element, so in
-general it is possible to have one or more functions running on the same
-element. SFs can be aware or not that are part of a certain chain, in the
-matter case they are called \emph{SFC aware} functions, in the latter one \emph
-SFC unaware}.
+network element and the software running on the top of it. 
+There are no limitations on the number of SFs running on the same network
+element, so in general it is possible to have one or more functions running on
+the same element. SFs can be aware or not that are part of a certain chain, in
+the matter case they are called \emph{SFC aware} functions, in the latter one
+\emph{SFC unaware}.
+
+The lifecycle management of the chain is enabled by two emerged technologies,
+software defined networking and network defined
+virtualization~\cite{medhat2017service}.
 
@@
 In order to exchange packet and information in the SFC ecosystem, traffic is
-encapsulated. The so called SFC encapsulation carries data used by the framework
-to identify SFPs, but it differs from a transport encapsulation and it is
-independent from the transport one: as such any transport protocol can be
+encapsulated. The so called SFC encapsulation carries data used by the
+framework to identify SFPs, but it differs from a transport encapsulation and it
+is independent from the transport one: as such any transport protocol can be
 exploited to forward traffic.
 
-\todo{add sfc proxy}
+This encapsulation is neither supported from SFC unaware SF nor from legacy 
+SF\footnote{Legacy SF refers to network function created before SFC framework}
+so \emph{SFC Proxies} has to be included in this ecosystem. These components
+sits the SFF and the SFs. They accept packets from the SFF and remove the
+encapsulation before delivering the packet to the SFs. In a mirrored way, it
+receive the result of the SF and reapplies the SFC header, in order to be
+managed by the SFF and the overall system. The communication among SFFs and
+proxies is the same as the SFFs and SFC aware SF.
 
@@
 
-\subsection{Considerations}
+\subsection{Other considerations}
+In this architecture the SFs are viewed are resource and managed by the control
+plane. Capabilities, availability and location are all accountable to the
+control part as their creation. It is also in charge of collecting information
+needed by the classifier to make proper decisions. The control plane could be
+either distributed or centralized and even a combination is possible.
+
+SFC system add to the original packets information as a NSH, that both contains
+metadata and information on the path that it must traverse. It is possible that
+the size of the packet, in this manner, exceed the MTU on some link in the
+chain. Solution must take in account this problem and the cost to overcome it.
 

### 2018-11-20 res/sections/08-Background.tex
Add citations and solve TODOs into 5g background

@@
 creation of a new standard, the 5G, which is estimated to arrive to the market
-consumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
-alliance, a group composed by the major players in the field of mobile
-connectivity, the 5G aims to offer not only at the end-user a new way to browse 
-the web, download and watch interactive content but also to create an ad-hoc 
-solution for Machine-to-Machine (M2M) data traffic, which is increasing more 
-than ever thanks to the spreading of IoT devices, which sensors need to 
-continuously send data to servers/data-centers. Relatively to LTE\todo{Explain 
-acronym}, 5G points to have data rates $10$ times better, with $10$ times 
-smaller end-to-end latency and an increased connection density by $100$ 
-times\todo{CITE}.
+consumer in 2020~\cite{iwamura2015ngmn}. Lead by the Next Generation Mobile
+Network (NGMN) alliance, a group composed by the major players in the field of
+mobile connectivity, the 5G aims to offer not only at the end-user a new way to
+browse the web, download and watch interactive content but also to create an
+ad-hoc solution for Machine-to-Machine (M2M) data traffic, which is increasing
+more than ever thanks to the spreading of IoT devices, which sensors need to 
+continuously send data to servers/data-centers. Relatively to \emph{Long Term
+Evolution} (LTE), 5G points to have data rates $10$ times better, with $10$
+times smaller end-to-end latency and an increased connection density by $100$ 
+times~\cite{alliance20155g}.
 
@@
 modular one, exploiting the flexibility that Virtual Network Functions 
-(VNFs)\todo{CITE VNF-P: A Model for Efficient Placement of Virtualized Network 
-Functions} can offer thanks to virtualization, closing the gap to the use-case 
-fulfillment defined by the NGMN alliance\todo{CITE: NGMN View on 5G 
-architecture} which require a decreased time to set up and deploy network 
-services (specifically, from 90 hours to 90 minutes\todo{Search this 
-requirement, check 90h 'cause I'm not sure about it}).
+(VNFs)\cite{moens2014vnf} can offer thanks to virtualization, closing the gap to
+the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
+which require a decreased time to set up and deploy network services
+(specifically, from 90 hours to 90 minutes~\cite{networld20202014role}).
 
@@
 With the constraints placed on the requirements formulated by the NGMN alliance,
-5G envisage a multi-layered architecture, based on three main layers:\todo{CITE:
-  NGMN View on 5G architecture}
+5G envisage a multi-layered architecture, based on three main
+layers~\cite{alliance20155g}:
 \begin{itemize}
@@
 
-\todo{Add three-layer image here?}
+\begin{figure}[H]
+  \centering
+  \includegraphics[scale=0.55]{5garchitecture}
+  \caption{5G Architecture~\cite{alliance20155g}}
+  \label{chap:background:img:5garchitecture}
+\end{figure}
 
@@
 \subsubsection{Network slicing}
-The role of a ``network slice'' in a 5G architecture is to specifically handle 
-the Control-plane\todo{Explain what a control plane is?} of a particular 
-service (e.g. smartphones traffic, autonomous driving, massive IoT), deploying 
-resources in a manner that assure the required latency, security and 
-reliability. While some very peculiar legacy services could require specific 
-hardware, the common resources between services could be shared in a 
-virtualized way, providing auto-scaling capabilities in services that are under 
-heavy network pressure.
+The role of a \emph{network slice} in a 5G architecture is to specifically
+handle the Control-plane of a particular service (e.g. smartphones traffic,
+autonomous driving, massive IoT), deploying resources in a manner that assure
+the required latency, security and reliability. While some very peculiar legacy
+services could require specific hardware, the common resources between services
+could be shared in a virtualized way, providing auto-scaling capabilities in
+services that are under heavy network pressure.
 
@@
 
-% \section{Technologies used}
-
 \vspace{0.5cm}

### 2018-11-20 res/sections/08-Background.tex
Rewrite some part on vibes and add reference to the website

@@
 \begin{description}
-  \item[Version]: This field is used to set the version of the header, used for
+  \item[Version:] This field is used to set the version of the header, used for
   ensuring backward compatibility through different NSH definitions;
-  \item[0 bit]: Bit set only in case of orchestration or management packets, and
+  \item[0 bit:] Bit set only in case of orchestration or management packets, and
   must not be changed by the SFC;
-  \item[TTL]: Time To Live, indicate the maximum number of hops among SFF and
+  \item[TTL:] Time To Live, indicate the maximum number of hops among SFF and
   its initial value should be configurable by the control plane. At each hop it
@@
   mechanism. The default value is 63;
-  \item[Length]: the total length of the header (in 4-byte words), including
+  \item[Length:] the total length of the header (in 4-byte words), including
   the \texttt{Fixed-Length Context Header} in NSH type 0x1 or the
@@
   length header data must be padded to a multiple of 4;
-  \item[Unassigned bits]: unassigned bits for future usages;
-  \item[MD]: Metadata Type, indicate the type of the NSH. Values that now
+  \item[Unassigned bits:] unassigned bits for future usages;
+  \item[MD:] Metadata Type, indicate the type of the NSH. Values that now
   accepted are \texttt{0x0} that is a reserved value and if set the packet must
@@
   Base header length and Service Path Header have always the same size;
-  \item[Next Protocol]: indicates the protocol type of encapsulated data
+  \item[Next Protocol:] indicates the protocol type of encapsulated data
   (protocols codes can be found at~\cite{rfc8300}).
@@
 \begin{description}
-  \item[Service Path Identifier (SPI)]: 24 bits, identify unambiguously an SFP.
+  \item[Service Path Identifier (SPI):] 24 bits, identify unambiguously an SFP.
   The first classifier have to set it;
-  \item[Service Index (SI)]: 8 bits, provide location within the SFP. Initially
+  \item[Service Index (SI):] 8 bits, provide location within the SFP. Initially
   must be set to 255 and it must be decremented on each hop.
@@
 \begin{description}
-  \item[Metadata Class (MD Class)]: field used in order to provide hierarchical
+  \item[Metadata Class (MD Class):] field used in order to provide hierarchical
   namespaces for standards, vendors or other scopes;
-  \item[Type]: indicates the type of metadata, depending on the MD Class;
+  \item[Type:] indicates the type of metadata, depending on the MD Class;
 \end{description}
@@
 \section{The VIBES project}
-
- This thesis is part of the VIBES project\todo{Talk about VIBES project, add
-   some reference, explain what it is.}, where the necessity for better TCP/IP
- transmission through satellite connections is the main requirement. To reach
- this goal, the project specifications suggest to exploit the 5G incoming
- technology and use the NFV-MANO architecture to perform first packets
- elaboration and performance improvement and finally TCP/IP satellite chunk
- optimization with the Performance Enhancing Proxy (PEP). The VIBES project
+This thesis is part of the VIBES project~\cite{vibesesa}, whose main requirement
+is to enhance the performance of end-to-end IP based that involves satellite
+connections. To reach this goal, the project specifications suggest to exploit
+the 5G incoming technology and use the NFV-MANO architecture to perform first
+packets elaboration and performance improvement and finally TCP/IP satellite
+chunk optimization with the Performance Enhancing Proxy (PEP). The VIBES project
 proposed five technical requirements:
@@
 \subsection{VNF-PEP architecture and internet protocols}
+VNF-PEP are the softwarization of usual PEP technology. One of the sake of the
+project is to research and include this technology in the usual VNF ecosystem to
+include without change satellite communications to the SFC. The architecture of
+the systems will be based on clusters of containers whose management and
+orchestration will follows the ETSI MANO architecture and SDN principles. Thanks
+to VNF-PEP will be possible to dynamically take decisions on how to optimize a
+certain communication.
 
-The analysis of this topic revealed to be trivial: since internet has many
-different protocols that would become infeasible to support all of them at the
-same time, packet encapsulation present itself as the only feasible solution:
-every packet incoming in the VNFs has already been encapsulated by a generic
-packet encapsulator/decapsulator,
-\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
-the protocol used on the edges but 1) use a ``quick protocol to exchange data 
-among VNFs and 2) we are hiding the path not the protocol itself. As we were 
-discussing, we are creating some sort of proxy, so the aim will be the same 
-even if we support a plethora of protocols.''} making the whole architecture 
-independent from the protocol a particular flux of data. To achieve this, 
-several solutions have been studied, and at the end packet encapsulation with 
-TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
-this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
-reference with the section of the explanation}
-
-\subsection{VNF-PEP prototype}
-
-Since the requirement for the whole system (MANO+PEP) were too challenging, the 
-goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
-The VNF architecture was shaped following the container orchestrator we decided 
-to use. An more detailed architectural implementation can be found at 
-Chapter~\ref{chap:archimpl}.
-
-\vspace{0.5cm}
-
-Starting with the first step of creating a MANO able to process incoming data
-packets through VNF functions, we encountered that many networking tools already
-present in the market required some tweaking and some integration, shifting our
-goal to create a complete European Telecommunications Standards Institutes
-(ETSI) Management and orchestrator (MANO) test-bed instead, following the
-specifications suggested in the RFC 7665, thus implementing only the first three
-requisites, without digging in the satellite data flow optimization. In
-particular, we discovered how, these tools, were suitable to create ETSI MANO
-and VNFs using virtual machine or exploiting cloud technologies, while they were
-not designed with enough flexibility an integration with Docker. In the next
-chapter we are going to take a deep analysis of the cited technologies 
-(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
-context to be able to understand our design choices we are going to describe 
-later.
-
-\noindent With this in mind, we performed a requirements analysis described in
-the next chapter.\todo{This should be put at the end of the introduction.}
+%The analysis of this topic revealed to be trivial: since internet has many
+%different protocols that would become infeasible to support all of them at the
+%same time, packet encapsulation present itself as the only feasible solution:
+%every packet incoming in the VNFs has already been encapsulated by a generic
+%packet encapsulator/decapsulator,
+%\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
+%the protocol used on the edges but 1) use a ``quick protocol to exchange data 
+%among VNFs and 2) we are hiding the path not the protocol itself. As we were 
+%discussing, we are creating some sort of proxy, so the aim will be the same 
+%even if we support a plethora of protocols.''} making the whole architecture 
+%independent from the protocol a particular flux of data. To achieve this, 
+%several solutions have been studied, and at the end packet encapsulation with 
+%TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
+%this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
+%reference with the section of the explanation}
+%
+%\subsection{VNF-PEP prototype}
+%
+%Since the requirement for the whole system (MANO+PEP) were too challenging, the 
+%goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
+%The VNF architecture was shaped following the container orchestrator we decided 
+%to use. An more detailed architectural implementation can be found at 
+%Chapter~\ref{chap:archimpl}.
+%
+%\vspace{0.5cm}
+%
+%Starting with the first step of creating a MANO able to process incoming data
+%packets through VNF functions, we encountered that many networking tools already
+%present in the market required some tweaking and some integration, shifting our
+%goal to create a complete European Telecommunications Standards Institutes
+%(ETSI) Management and orchestrator (MANO) test-bed instead, following the
+%specifications suggested in the RFC 7665, thus implementing only the first three
+%requisites, without digging in the satellite data flow optimization. In
+%particular, we discovered how, these tools, were suitable to create ETSI MANO
+%and VNFs using virtual machine or exploiting cloud technologies, while they were
+%not designed with enough flexibility an integration with Docker. In the next
+%chapter we are going to take a deep analysis of the cited technologies 
+%(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
+%context to be able to understand our design choices we are going to describe 
+%later.
+%
+%\noindent With this in mind, we performed a requirements analysis described in
+%the next chapter.\todo{This should be put at the end of the introduction.}
+%

### 2018-11-20 res/sections/25-Tests.tex
Add explanation about K8s vs Docker

@@
 start up time (seconds)} \\ \hline
-\textbf{Docker}      & 0.009620874                           & 0,016            
-                        \\
-\textbf{Virtual Box} & 0.755                                 & 20,211
-
+\textbf{Docker}      & 0.009620874                           & 0.016        \\
+\textbf{Virtual Box} & 0.755                                 & 20.211
 \end{tabular}%
@@
 % approximately of 33 minutes, witch gives a significant insight of how much 
-time
-% is required to create and boot a virtualized environment.
+% time is required to create and boot a virtualized environment.
 
@@
 
+\newpage
+
 % TODO is it worth to keep it as a long table?
@@
 %
-\textbf{1}          & 2.560102                   & 13,291                     \\
-\textbf{2}          & 0.6301444                  & 14,207                     \\
-\textbf{3}          & 1.14903                    & 15,345                     \\
-\textbf{4}          & 3.592179                   & 16,453                     \\
-\textbf{5}          & 0.9728132                  & 16,634                     \\
-\textbf{6}          & 1.103802                   & 17,418                     \\
+\textbf{1}          & 2.560102                   & 13.291                     \\
+\textbf{2}          & 0.6301444                  & 14.207                     \\
+\textbf{3}          & 1.14903                    & 15.345                     \\
+\textbf{4}          & 3.592179                   & 16.453                     \\
+\textbf{5}          & 0.9728132                  & 16.634                     \\
+\textbf{6}          & 1.103802                   & 17.418                     \\
 \textbf{7}          & 2.944699                   & 17.984                     \\
@@
 \end{longtable}
+
+\paragraph*{Considerations on container start-up in Kubernetes}
+Regarding the start-up times in Kubernetes is possible to denote a considerable
+difference between the deployment of a SFC with only one element
+(\textit{de-facto} only one container) with a single container directly launched
+on Docker. The difference, of many seconds, can be easily explained: the
+creation of a Docker container, \textit{per se}, is quite fast compared to the
+scheduling job the Kubernetes cluster has to do before launching the pod. A
+Kubernetes deployment is indeed a rather complex operation than launching a
+simple container. When a new deployment is launched on Kubernetes, the operation
+initiate with the scheduling of a Pod. A Pod can have several lifecycle
+states~\cite{kubePodLifecycle}:
+\begin{description}
+\item[pending:] The Pod has been accepted by the Kubernetes system, but one or
+  more of the Container images has not been created. This includes time before
+  being scheduled as well as time spent downloading images over the network,
+  which could take a while;
+\item[running:] The Pod has been bound to a node, and all of the Containers have
+  been created. At least one Container is still running, or is in the process of
+  starting or restarting;
+\item[succeeded:] All Containers in the Pod have terminated in success, and will
+  not be restarted;
+\item[failed:] All Containers in the Pod have terminated, and at least one
+  Container has terminated in failure. That is, the Container either exited with
+  non-zero status or was terminated by the system;
+\item[unknown:] For some reason the state of the Pod could not be obtained,
+  typically due to an error in communicating with the host of the Pod.
+\end{description}
+
+On top of that, different conditions can affect a Pod:
+\begin{itemize}
+\item \verb!PodScheduled!: the Pod has been scheduled to a node;
+\item \verb!Ready!: the Pod is able to serve requests and should be added to the
+  load balancing pools of all matching Services;
+\item \verb!Initialized!: all ``Init Containers'' have started successfully;
+\item \verb!Unschedulable!: the scheduler cannot schedule the Pod right now, for
+  example due to lacking of resources or other constraints.
+\end{itemize}
+
+These additional operations implicate some time overhead, that cause the
+deployment of a single Pod with a single container to be way longer that the
+deployment of a simple container directly from Docker.
 

### 2018-11-21 res/sections/08-Background.tex
Move section of background
Minor english fixes

@@
 
+\section{Virtualization}
+Cloud computing is nowadays one of the cutting edge Internet technologies, and
+it is possible thanks to virtualization, that allows to isolate applications,
+optimizing the utilization of hardware resources and providing more flexibility.
+Virtualization, in computing, usually refers to the act of creating a virtual
+version of a certain asset, such as virtual computer hardware platform,
+operating system (OS), storage device, or computer network
+resources~\cite{liu2014research}. Conventional virtualization
+(hypervisor-based) uses virtual machines: it provides strong isolation and a
+complete operating system~\cite{eder2016hypervisor}. In hypervisor
+virtualization every pieces of hardware required to running software is
+emulated, and it is common that the hypervisor provides interfaces to access
+the real hardware, for instance to access physical devices or to communicate
+with the network. However all this functionalities comes to a resource
+overhead~\cite{scheepers2014virtualization}. Container-based virtualization,
+instead, could be a more efficient in certain context, because it does not
+emulate the entire computer. A container is seen as an isolate process running
+on the host, to which the Linux kernel gives a set of capabilities.
+
+\subsection{Network Virtualization}
+Virtualization, in the networking context, refers to the act of combining
+hardware, software and functionalities into a virtual network. This abstraction
+runs on the top of a physical network in a hyper-visor, decoupling the logical
+network from the hardware one. Network virtualization allow multiple,
+heterogeneous networks to cohabit, increasing the flexibility and manageability.
+The idea of network virtualization came from the decoupling of the role of the
+traditional ISP into two different entities: who provides the infrastructure
+(InPs) and who provides the services (SPs). In that context it is possible to
+aggregate resources, even from different InPs, creating a virtual network,
+offering an end-to-end service~\cite{chowdhury2009network}. Most known instances
+of NV are \emph{VPNs}, dedicated networks that allows the communication among
+multiple sites, using private and secure tunnels, relying on the public network
+such as Internet. Other examples of network virtualization are \emph{overlay
+  networks}. They are logical networks, build on the top of an existing physical
+network, typically implemented at the application layer.
+
+\section{Software-defined networking}
+The SDN is an emerging approach to cloud computing and networking technology
+capable of supporting the dynamic nature of future network functions and
+intelligent applications while lowering operating costs through simplified
+hardware, software, and management~\cite{sezer2013we}, that allows to configure
+network in a programmatically way, possibly enhancing monitoring and
+performances. One of the main principles of SDN architecture is to decouple the
+packet forwarding process, \emph{data plane}, from the routing one, the so
+called \emph{control plane}. The architecture also contemplate that network
+intelligence and state are logically centralized~\cite{fundation2012software}.
+
+It has some drawbacks in terms of security, scalability and elasticity, because
+of the centralized architecture.
+
+\subsection{SDN stack}
+\begin{figure}[ht]
+ \centering
+ \includegraphics[scale=1.3]{sdn_architecture}
+ \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
+denote the three layer logic: the Application plane, the Controller plane and 
+the Data plane~\cite{fundation2013software}.}
+ \label{chap:background:img:sdn_architecture}
+\end{figure}
+As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
+architecture can be split in three different layers, also known as planes,
+based on their functionalities~\cite{fundation2012software} as explained
+in~\cite{fundation2013software}.
+
+\subsubsection{Application plane}
+
+In the \emph{Application plane} lays \emph{SDN applications} that are programs
+whose role is to set the behavior of the SDN controllers 
+and manage its hardware resources and data through APIs, often called 
+\emph{northbound interfaces} (NBI). In order to do so they can also build an
+abstracted view of the network, using controllers information, in order to make
+the traffic analysis easier and to simplify the decision making process.
+These interfaces provide an additional abstraction layer that high-level 
+applications can consume, enabling the possibility to use a variety of 
+different vendors' controllers in the same way. Applications consists both of
+the two defined layers, one that is responsible to manage the whole logic, and
+another one dedicated to the communication part, that in particular uses the
+previously defined NBIs.
+
+\subsubsection{Controller plane}
+
+\emph{SDN controllers} are logically centralized components that
+receive requests from SDN applications and either translate them 
+in a fashion that can be understood from the SDN Datapath, or collect 
+data as statics and events needed by the application to build an abstract view
+of the network. 
+Due to the variety of tasks a controller has to accomplish, its logical 
+subdivision is achieved with a three-layer architecture, in which every layer 
+exposes its own interface, that can be consequently used by the upper 
+layers, up to the \emph{NBI agents}.
+NBI agent manages the communications with SDN applications, exposing interfaces
+to access network data and to the set of the controller functionalities.
+In particular, the middle layer represents the \emph{SDN Controller Logic},
+which process the requests coming from the upper layer applications. Finally
+requests are communicated to the remaining network components, exploiting the
+Control to Data-Plane Interface driver.
+
+Despite the centralized design of controllers there are no limitations on the
+implementation, so it possible to create federations of controllers or organize
+them into a hierarchy.
+
+\subsubsection{Data plane}
+In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
+network device, which exposes the possibility to access and control to its
+forwarding and processing capabilities. It is composed by a CDPI agent and a set
+of one or multiple traffic forwarding engines and of a zero or more traffic
+processing functions that are, as a matter of fact, the devices logic. Datapath
+can be mapped either to a single physical device or multiple ones and the
+implementation of the its logic does not preclude neither the virtualization or
+the sharing of physical resources nor the interoperability with non-SDN
+networking.
+
+\subsection{OpenFlow protocol}
+The SDN strategy is often associated with the OpenFlow protocol because it is
+the first standard communications interface defined between the control and the
+datapath in SDN architecture~\cite{fundation2013software}, making possible to
+access from controller the components of infrastructure, such as switches and
+routers. OpenFlow enables the remote administration of networks working on 
+layer 3: packets forwarding tables can be modified relying on packet matching 
+rules. Following this approach it is possible to change routes periodically or 
+ad-hoc, based on packet types, traffic or network load.
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.65]{openflow}
+ \caption{OpenFlow protocol communication graph}
+ \label{chap:background:img:openflow_protocol}
+\end{figure}
+
+In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
+how switches works in the OpenFlow protocol. Basically, OpenFlow switches own a
+set of tables to organized packet control and forwarding. These tables can be
+updated by higher-level application, usually an SDN application or controller,
+taking advantage of OpenFlow channels (called OF channels).
+
 \section{Network slicing}
@@
 some specifics that make it a novel idea. The concept of network slicing was
-first introduced in 2002 in\cite{peterson2003blueprint}, where a network 
+first introduced in 2002 in~\cite{peterson2003blueprint}, where a network 
 overlay that allows services to run continuously and to access the overlay
@@
 security. In fact, in case of attack or fault of a certain slice, others must
-not be affected. Moreover, each slice must implement it own set of security
+not be affected. Moreover, each slice must implement its own set of security
 function to preventing unauthorized access. Furthermore, isolation let to
@@
 
-\section{Software-defined networking}
-The SDN is an emerging approach to cloud computing and networking technology
-capable of supporting the dynamic nature of future network functions and
-intelligent applications while lowering operating costs through simplified
-hardware, software, and management~\cite{sezer2013we}, that allows to configure
-network in a programmatically way, possibly enhancing monitoring and
-performances. One of the main principles of SDN architecture is to decouple the
-packet forwarding process, \emph{data plane}, from the routing one, the so
-called \emph{control plane}. The architecture also contemplate that network
-intelligence and state are logically centralized\cite{fundation2012software}.
-
-It has some drawbacks in terms of
-security, scalability and elasticity, because of the centralized architecture.
-
-\subsection{SDN stack}
-\begin{figure}[ht]
- \centering
- \includegraphics[scale=1.3]{sdn_architecture}
- \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
-denote the three layer logic: the Application plane, the Controller plane and 
-the Data plane~\cite{fundation2013software}.}
- \label{chap:background:img:sdn_architecture}
-\end{figure}
-As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be split in three different layers, also known as planes,
-based on their functionalities~\cite{fundation2012software} as explained
-in~\cite{fundation2013software}.
-
-\subsubsection{Application plane}
-
-In the \emph{Application plane} lays \emph{SDN applications} that are programs
-whose role is to set the behavior of the SDN controllers 
-and manage its hardware resources and data through APIs, often called 
-\emph{northbound interfaces} (NBI). In order to do so they can also build an
-abstracted view of the network, using controllers information, in order to make
-the traffic analysis easier and to simplify the decision making process..
-These interfaces provide an additional abstraction layer that high-level 
-applications can consume, enabling the possibility to use a variety of 
-different vendors' controllers in the same way. Applications consists both of
-the two defined layers, one that is responsible to manage the whole logic, and
-another one dedicated to the communication part, that in particular uses the
-previously defined NBIs.
-
-\subsubsection{Controller plane}
-
-\emph{SDN controllers}, are logically centralized components that
-receive requests from SDN applications and either translate them 
-in a fashion that can be understood from the SDN Datapath, or collect 
-data as statics and events needed by the application to build an abstract view
-of the network. 
-Due to the variety of tasks a controller has to accomplish, its logical 
-subdivision is achieved with a three-layer architecture, in which every layer 
-exposes it's own interface, that can be consequently used by the upper 
-layers, up to the \emph{NBI agents}.
-NBI agent manages the communications with SDN applications, exposing interfaces
-to access network data and to the set of the controller functionalities.
-In particular, the middle layer represents the \emph{SDN Controller Logic},
-which process the requests coming from the upper layer applications. Finally
-requests are communicated to the remaining network components, exploiting the
-Control to Data-Plane Interface driver.
-
-Despite the centralized design of controllers there are no limitations on the
-implementation, so it possible to create federations of controllers or organize
-them into a hierarchy.
-
-\subsubsection{Data plane}
-In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
-network device, which exposes the possibility to access and control to its
-forwarding and processing capabilities.It is composed by a CDPI agent and a set
-of one or  multiple traffic forwarding engines and of a zero or more traffic
-processing functions, that are, as a matter of fact, the devices logic.
-Datapath can be mapped either to a single physical device or multiple ones and
-the implementation of the its logic does not preclude neither the
-virtualization or the sharing of physical resources nor the interoperability
-with non-SDN networking.
-
-\subsection{OpenFlow protocol}
-The SDN strategy is often associated with the OpenFlow protocol because it is
-the first standard communications interface defined between the control and the
-datapath in SDN architecture~\cite{fundation2013software}, making possible to
-access from controller the components of infrastructure, such as switches and
-routers. OpenFlow enables the remote administration of networks working on 
-layer 3: packets forwarding tables can be modified relying on packet matching 
-rules. Following this approach it is possible to change routes periodically or 
-ad-hoc, based on packet types, traffic or network load.
-
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.65]{openflow}
- \caption{OpenFlow protocol communication graph}
- \label{chap:background:img:openflow_protocol}
-\end{figure}
-
-In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
-how switches works in the OpenFlow protocol. Basically, OpenFlow switches own a
-set of tables to organized packet control and forwarding. These tables can be
-updated by higher-level application, usually an SDN application or controller,
-taking advantage of OpenFlow channels (called OF channels).
-
-\section{Virtualization}
-Cloud computing is nowadays one of the cutting edge Internet technologies, and
-it is possible thanks to virtualization, that allows to isolate applications,
-optimizing the utilization of hardware resources and providing more flexibility.
-Virtualization, in computing, usually refers to the act of creating a virtual
-version of a certain asset, such as virtual computer hardware platform,
-operating system (OS), storage device, or computer network
-resources~\cite{liu2014research}. Conventional virtualization
-(hypervisor-based) uses virtual machines: it provides strong isolation and a
-complete operating system~\cite{eder2016hypervisor}. In hypervisor
-virtualization every pieces of hardware required to running software is
-emulated, and it is common that the hypervisor provides interfaces to access
-the real hardware, for instance to access physical devices or to communicate
-with the network. However all this functionalities comes to a resource
-overhead~\cite{scheepers2014virtualization}Container-based virtualization,
-instead, could be a more efficient in certain context, because it does not
-emulate the entire computer. A container is seen as an isolate process running
-on the host, to which the Linux kernel gives a set of capabilities.
-
-\subsection{Network Virtualization}
-Virtualization, in the networking context, refers to the act of combining
-hardware, software and functionalities into a virtual network. This abstraction
-runs on the top of a physical network in a hyper-visor, decoupling the logical
-network from the hardware one. Network virtualization allow multiple,
-heterogeneous networks to cohabit, increasing the flexibility and manageability.
-The idea of network virtualization came from the decoupling of the role of the
-traditional ISP into two different entities: who provides the infrastructure
-(InPs) and who provides the services(SPs). In that context it is possible to
-aggregate resources, even from different InPs, creating a virtual network,
-offering an end-to-end service~\cite{chowdhury2009network}. Most known
-instances of NV are \emph{VPNs}, dedicated networks that allows the
-communication among multiple sites, using
-private and secure tunnels, relying on the public network such as Internet.
-Other examples of network virtualization are \emph{overlay networks}. They are
-logical networks, build on the top of an existing physical network, typically
-implemented at the application layer.
-
 \subsection{Network Function Virtualization}
 The \emph{Network Function Virtualization} (NFV) concept refers to the use of
-virtualization technologies to virtualize network nodes functions. This
-emerging technology focuses on the decoupling of function logic and hardware,
-allowing VNFs to run on the top of servers, switches, storage 
-devices or on cloud computing infrastructures. In other word the software logic
-is separated from the custom platform on which it runs, enabling the use of
-commodity hardware~\cite{gray2016network}.
-That approach worked well in the past, because of the need to follow rigorous 
-standards, providing at the same time high quality, but it led to long product 
-cycles and high specialized services that use a combination of software and
-hardware resources, typically vendor-specific, and only largely
-non-interoperable within other vendor solutions. New Internet scenarios, that 
-include IoT and the increasing number of services to spread multimedia data 
-(Skype, Netflix, Spotify, \dots), requires more flexible functions and faster 
-network deployments. 
-This detachment provides greater flexibility to scale VNF based on traffic and
-performances and also improve the automation of function deploy and update:
-removing dedicated hardware also will reduce costs machines and for
-maintenance.
+virtualization technologies to virtualize network nodes functions. This emerging
+technology focuses on the decoupling of function logic and hardware, allowing
+VNFs to run on the top of servers, switches, storage devices or on cloud
+computing infrastructures. In other words the software logic is separated from
+the custom platform on which it runs, enabling the use of commodity
+hardware~\cite{gray2016network}. That approach worked well in the past, because
+of the need to follow rigorous standards, providing at the same time high
+quality, but it led to long product cycles and high specialized services that
+use a combination of software and hardware resources, typically vendor-specific,
+and only largely non-interoperable within other vendor solutions. New Internet
+scenarios, that include IoT and the increasing number of services to spread
+multimedia data (Skype, Netflix, Spotify, \dots), requires more flexible
+functions and faster network deployments. This detachment provides greater
+flexibility to scale VNF based on traffic and performances and also improve the
+automation of function deploy and update: removing dedicated hardware also will
+reduce costs machines and for maintenance.
 
@@
 performances, functions running on general purposes servers cloud affect
-throughput and latency, moreover the capacity of the virtualized function can
-be less than the hardware and software version. To reach comparable levels in
-terms of performances efficient algorithms are needed, as a excellent resource
+throughput and latency, moreover the capacity of the virtualized function can be
+less than the hardware and software version. To reach comparable levels in terms
+of performances efficient algorithms are needed, as a excellent resource
 management. Another problems is the VNF management. In case of hardware
 functions resources are almost equivalent, at the contrary, virtualized ones
-could have costs and values that differs significantly depending on the point
-of presence on which are deployed. On the other hand, since VNFs are software
-components, can elastically scale up and down, depending on the traffic loads
-and user demand. Another management challenge refers to keeping track where the
-VNF is running, that is of paramount importance in case of issues. Related to
-possible dysfunctions, software function must reach a certain level of
-reliability. Generally, purpose-build network equipment can ensure the
+could have costs and values that differs significantly depending on the point of
+presence on which are deployed. On the other hand, since VNFs are software
+components, they can elastically scale up and down, depending on the traffic
+loads and user demand. Another management challenge refers to keeping track
+where the VNF is running, that is of paramount importance in case of issues.
+Related to possible dysfunctions, software function must reach a certain level
+of reliability. Generally, purpose-build network equipment can ensure the
 five-nines reliability requirement, that is more challenging to follow in that
@@
 received. They can work on different levels of the network layer. These
-components can be implemented as virtual element or a combination of a physical
-network element and the software running on the top of it. 
-There are no limitations on the number of SFs running on the same network
-element, so in general it is possible to have one or more functions running on
-the same element. SFs can be aware or not that are part of a certain chain, in
-the matter case they are called \emph{SFC aware} functions, in the latter one
-\emph{SFC unaware}.
+components can be implemented as virtual elements only or in a combination of a
+physical and software network elements. There are no limitations on the number
+of SFs running on the same network element, so in general it is possible to have
+one or more functions running on the same element. SFs can be aware or not that
+are part of a certain chain, in the matter case they are called \emph{SFC aware}
+functions, in the latter one \emph{SFC unaware}.
 
@@
 principles is the \emph{decoupling}. Decoupling in terms on independence on the
-infrastructure and the network, so no change must be performed to the
-forwarding underlying topology to deploy SFCs, and in terms of separation
-between packet handling (by SFs) and forwarding. This decoupling refers even to
-architecture and functions: one must be independent from the other. Another
-main point is the possibility to dynamically set routes on the incoming traffic
-to decide the best treatment based on its typology. The classification can
-occur on different point of the chain and can be updated during the packet
-delivery. Finally creation, modification and deletion of a certain SFC must has
-To impact on the others.
+infrastructure and the network, so no change must be performed to the forwarding
+underlying topology to deploy SFCs, and in terms of separation between packet
+handling (by SFs) and forwarding. This decoupling refers even to architecture
+and functions: one must be independent from the other. Another main point is the
+possibility to dynamically set routes on the incoming traffic to decide the best
+treatment based on its typology. The classification can occur on different point
+of the chain and can be updated during the packet delivery. Finally creation,
+modification and deletion of a certain SFC must not impact on the others.
 
 \subsection{Graph view of an SFC}
-An SFC can be seen as a graph in which each node is an SF and edges are the
-path through which a packet have to pass through. Because of this definitions
-there are no limitation on possible cycles that can occur in the chain, in case
-that a certain SF have to be reapplied on the same packet, frame or flow. These
+An SFC can be seen as a graph in which each node is an SF and edges are the path
+through which a packet have to pass through. Because of this definitions there
+are no limitations on possible cycles that can occur in the chain, in case that
+a certain SF has to be reapplied on the same packet, frame or flow. These
 scenarios must be adequately managed to avoid endless loops. This enable even
@@
 that has to be imposed from that point of the chain. Traffic path in the SFC
-graph are managed and described by the \emph{Service Function Forwarder} (SFF)
-and \emph{Service Function Path} (SFP) components. The matter is the element of
+graph is managed and described by the \emph{Service Function Forwarder} (SFF)
+and \emph{Service Function Path} (SFP) components. The latter is the element of
 the architecture responsible of traffic forwarding. The main goals are to
@@
 constraint specification on the path of where packets assigned to a certain
-service function path must go. There are no limitation on how specific it must
+service function path must go. There are no limitations on how specific it must
 be: it can specify the exact locations, identifying both the SFF and the SF, or
@@
 The \emph{Network Service Header} (NSH), used by SFC environment for
-encapsulation is described in~\cite{rfc8300} and consider three main fields as
-represent in Figure~\ref{chap:background:img:nshformat}: Base Header, Service
-Path Header and Context Header.
+encapsulation is described in~\cite{rfc8300} and it consider three main fields
+as represented in Figure~\ref{chap:background:img:nshformat}: Base Header,
+Service Path Header and Context Header.
 \begin{figure}[H]
@@
 \subsubsection{Base header}
-The Base Header utility is provide information about the service header and the
-payload protocol.
+The Base Header utility is to provide information about the service header and
+the payload protocol.
 
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
   \centering
@@
 autonomous driving, massive IoT), deploying resources in a manner that assure
-the required latency, security and reliability. While some very peculiar legacy
-services could require specific hardware, the common resources between services
-could be shared in a virtualized way, providing auto-scaling capabilities in
-services that are under heavy network pressure.
+the required latency, security and reliability. While some very peculiar and
+legacy service could require specific hardware, the common resources between
+services could be shared in a virtualized way, providing auto-scaling
+capabilities in services that are under heavy network pressure.
 
 \section{The VIBES project}
-This thesis is part of the VIBES project~\cite{vibesesa}, whose main requirement
-is to enhance the performance of end-to-end IP based that involves satellite
-connections. To reach this goal, the project specifications suggest to exploit
-the 5G incoming technology and use the NFV-MANO architecture to perform first
-packets elaboration and performance improvement and finally TCP/IP satellite
-chunk optimization with the Performance Enhancing Proxy (PEP). The VIBES project
-proposed five technical requirements:
+This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
+requirement is to enhance the performance of end-to-end IP based that involves
+satellite connections. To reach this goal, the project specifications suggest to
+exploit the 5G incoming technology and use the NFV-MANO architecture to perform
+first packets elaboration and performance improvement and finally TCP/IP
+satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
+VIBES project proposed five technical requirements:
 \begin{enumerate}
@@
 \subsection{VNF-PEP architecture and internet protocols}
-VNF-PEP are the softwarization of usual PEP technology. One of the sake of the
-project is to research and include this technology in the usual VNF ecosystem to
-include without change satellite communications to the SFC. The architecture of
-the systems will be based on clusters of containers whose management and
-orchestration will follows the ETSI MANO architecture and SDN principles. Thanks
-to VNF-PEP will be possible to dynamically take decisions on how to optimize a
-certain communication.
+VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
+the project is to research and include this technology in the usual VNF
+ecosystem to include without change satellite communications to the SFC. The
+architecture of the systems will be based on clusters of containers whose
+management and orchestration will follows the ETSI MANO architecture and SDN
+principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
+how to optimize a certain communication.
 

### 2018-11-21 res/sections/09-RelatedWork.tex
Add some related works

@@
 container-based approach.
+
+\subsection{Other SFC implementation}
+Since the SFC technology was proposed many work came out proposing different implementation solutions. In this part there is an overview of some of them.
+
+In~\cite{GhaznaviSAB16} proposed optimization model to calculate the optimal
+location of VNFs based on different metrics such as bandwidth, load balancing
+and routes. This decomposes SFC on performance requirements eliminating the
+throughput bound of a chain to a single VNF or a single physical machine.
+Mathematical models developed was tested in a simulation enviroment. A critical
+enhancement to this work could be to test algorithm in a real environment, under
+different conditions of traffic load: if it efficiant even in this senario it
+can be useful to the orchestration and management part of the ecosystem because
+it can reduce resource waste.
+
+Other implementation tested in a simulated environment
+are~\cite{csoma2014escape} and~\cite{kim2016evaluations}. In both of them the
+network functions are deployed on Mininet. The matter work provide a full
+implementation of the SFC environment, including SFC configuration, allocating
+VNFs into the physical resources, flow routing across the VNFs based on
+policies, and provisioning live management information on operating VNF
+instances, but no performance evaluation of the work is presented. In the
+latter case chains are identified by VLAN tag and the VNF chaining is created
+exploiting only using Floodlight controllers.
+
+Another noteworthy work is~\cite{xia2015optical} that discuss the packet
+delivery through function chaining in high capacity network exploiting aggregate
+flow steering. The architecture proposed make use of SDN and cloud platform to
+deploy VMs on which run VNFs assuming that the SDN and network element are
+connected using optical circuit switching, taking advantage of Wave Selective
+Switchig (WSS), to route (switch) signals between optical fibres on a
+per-wavelength basis. Thank to the implementation of the SDN controller and the
+communication between the previous element and the VNFM, this architecture is
+able to provide good capabilities in terms of flexibilities. Another element
+that helps the flexibility is the high performance link: in this way they get
+rid of problems in terms of connection time and bandwidth. However, in the paper
+authors assumes really specific hardware in that underlying infrastructure that
+connects the overall system: this is not suitable to be deployed in a general
+purpose infrastructure or using IaaS. It will be interesting to evaluated the
+same work on different instrumentation. A clashing aspect is the time to make a
+VM up and running compared to times of servvice chaining configuration/packet
+delivery. Using containers the instantiation of virtual functions should employ
+less time.
+
+A rilevant work on function chaining in network is~\cite{trajkovska2017sdn}. In
+this work author efforts are concentrated on joining SDN and VNF technologies to
+create SFCs. They developed an SDN-driven SDK, a virtual traffic classifier to
+analyze traffic in real-time and decide the optimal chain for it and a VNF
+specialized on video transdecoding, used event to evaluate the performances of
+the whole system. Chaining is performed exploiting layer 2 connectivity,
+using OpenvSwitch as virtual network svitches an OpenDayLight in order to
+route the traffic on a set of policies. However the solution does not provide
+any mechanism to automatic deployment of chains and services: this can be an
+issue in particular in case of VNF fault. Orchestration and some mechanisms of
+fault tollerance are critical in particular in a real environment, enabling
+autoscaling on user demand or traffic load. Automated deployment also make
+possible the recreation of a virtual machine in case of fault. As a matter of
+fact that a virtual function can fail depending on an erroneus load balancing
+policy or too heavy computational load and it is critical in order to do not
+interrupt end-to-end communication to bring up the VNF as soon as possible. This
+recreation process can be slow due to the usage of VMs to deploy the
+functions, that could take minutes to startup.
+
+The paper~\cite{kriti2017dnfc} introduce the possibility using container in SFC
+deployment. The chain of container is created exploiting Openflow to create flow
+to route traffic to the containers, that have to be installed on the host
+machine. This work however does present any form of management or orchestration
+of the chain, neither an automatic deployment. Moreover it requires installation
+of part of the system on the host machine: in this way there is not a fully
+virtualized environment. The matter issue has some implication: the VNF
+has a strong relation on the host on which is running and the system could not
+be simply be moved from one machine to another as it could be done only using
+containers. Finally this paper does not provide any form of system evaluation
+neither in terms of performances nor in terms of resilience. Fault recovering
+and autoscaling could be achieved only improving the system introducing other
+technologies, such as Docker Compose, Kubernetes or Docker Swarm.

### 2018-11-21 res/sections/09-RelatedWork.tex
Add some other works

@@
 
+One of the first proposal that take advantage of container in to take on network
+edges NFV is~\cite{cziva2017container}. The authors present the work, based on
+previous definition of the platform defined in~\cite{cziva2015container} to
+orchestrate the virtual function and to deploy it as close as possible to the
+data sender to classify traffic and choose the appropriate functions
+immediately.
+
 \section{EPC virtual implementations}
@@
 Compared to our solution, these implementations does not present an unified
-framework that gives the possbility orchestrate different network slices
+framework that gives the possibility orchestrate different network slices
 consistently. Adopting Kubernetes gives the advantage to use the Kubernetes
@@
 network slices is proposed as a viable solution to solve the security problems
-about isolation. We tried to mitigate the security consern presented above using
+about isolation. We tried to mitigate the security concern presented above using
 the integrated Role Base Access Control system (RBAC) and other security policy
@@
 system, where the SFC paths are somehow defined at deploy time, and can't change
-without decomissioning, editing and deploying the chain again. Our proposal,
-instead, is able to support chain editing on-the-fly, allowing NFV to dynamically
-change SFP.
+without decommissioning, editing and deploying the chain again. Our proposal,
+instead, is able to support chain editing on-the-fly, allowing NFV to
+dynamically change SFP.
 
@@
 
-\subsection{Other SFC implementation}
-Since the SFC technology was proposed many work came out proposing different implementation solutions. In this part there is an overview of some of them.
+\subsection{SFC implementation}
+Since the SFC technology was proposed many work came out proposing different
+implementation solutions. In this part there is an overview of some of them.
 
@@
 throughput bound of a chain to a single VNF or a single physical machine.
-Mathematical models developed was tested in a simulation enviroment. A critical
+Mathematical models developed was tested in a simulation environment. A critical
 enhancement to this work could be to test algorithm in a real environment, under
-different conditions of traffic load: if it efficiant even in this senario it
+different conditions of traffic load: if it efficient even in this scenario it
 can be useful to the orchestration and management part of the ecosystem because
@@
 connected using optical circuit switching, taking advantage of Wave Selective
-Switchig (WSS), to route (switch) signals between optical fibres on a
+Switching (WSS), to route (switch) signals between optical fibers on a
 per-wavelength basis. Thank to the implementation of the SDN controller and the
 communication between the previous element and the VNFM, this architecture is
-able to provide good capabilities in terms of flexibilities. Another element
+able to provide good capabilities in terms of flexibility. Another element
 that helps the flexibility is the high performance link: in this way they get
@@
 same work on different instrumentation. A clashing aspect is the time to make a
-VM up and running compared to times of servvice chaining configuration/packet
+VM up and running compared to times of service chaining configuration/packet
 delivery. Using containers the instantiation of virtual functions should employ
@@
 
-A rilevant work on function chaining in network is~\cite{trajkovska2017sdn}. In
+A relevant work on function chaining in network is~\cite{trajkovska2017sdn}. In
 this work author efforts are concentrated on joining SDN and VNF technologies to
@@
 the whole system. Chaining is performed exploiting layer 2 connectivity,
-using OpenvSwitch as virtual network svitches an OpenDayLight in order to
+using OpenvSwitch as virtual network switches an OpenDayLight in order to
 route the traffic on a set of policies. However the solution does not provide
@@
 issue in particular in case of VNF fault. Orchestration and some mechanisms of
-fault tollerance are critical in particular in a real environment, enabling
+fault tolerance are critical in particular in a real environment, enabling
 autoscaling on user demand or traffic load. Automated deployment also make
 possible the recreation of a virtual machine in case of fault. As a matter of
-fact that a virtual function can fail depending on an erroneus load balancing
+fact that a virtual function can fail depending on an erroneous load balancing
 policy or too heavy computational load and it is critical in order to do not

### 2018-11-21 res/sections/08-Background.tex
Fix English

@@
 one or more functions running on the same element. SFs can be aware or not that
-are part of a certain chain, in the matter case they are called \emph{SFC aware}
+are part of a certain chain, in the former case they are called \emph{SFC aware}
 functions, in the latter one \emph{SFC unaware}.
@@
 graph is managed and described by the \emph{Service Function Forwarder} (SFF)
-and \emph{Service Function Path} (SFP) components. The latter is the element of
+and \emph{Service Function Path} (SFP) components. The former is the element of
 the architecture responsible of traffic forwarding. The main goals are to

### 2018-11-21 res/sections/09-RelatedWork.tex
Fix English

@@
 instead, is able to support chain editing on-the-fly, allowing NFV to
-dynamically change SFP.
+dynamically change the SFP.
 
@@
 
-In~\cite{GhaznaviSAB16} proposed optimization model to calculate the optimal
+\cite{GhaznaviSAB16} proposed an optimization model to calculate the optimal
 location of VNFs based on different metrics such as bandwidth, load balancing
-and routes. This decomposes SFC on performance requirements eliminating the
+and routes. This decomposes the SFC on performance requirements eliminating the
 throughput bound of a chain to a single VNF or a single physical machine.
@@
 are~\cite{csoma2014escape} and~\cite{kim2016evaluations}. In both of them the
-network functions are deployed on Mininet. The matter work provide a full
+network functions are deployed on Mininet. The former work provides a full
 implementation of the SFC environment, including SFC configuration, allocating
@@
 policies, and provisioning live management information on operating VNF
-instances, but no performance evaluation of the work is presented. In the
-latter case chains are identified by VLAN tag and the VNF chaining is created
+instances, but no performance evaluation of the work is presented. In the latter
+case chains are identified by VLAN tag and the VNF chaining is created
 exploiting only using Floodlight controllers.
@@
 Another noteworthy work is~\cite{xia2015optical} that discuss the packet
-delivery through function chaining in high capacity network exploiting aggregate
-flow steering. The architecture proposed make use of SDN and cloud platform to
-deploy VMs on which run VNFs assuming that the SDN and network element are
-connected using optical circuit switching, taking advantage of Wave Selective
-Switching (WSS), to route (switch) signals between optical fibers on a
-per-wavelength basis. Thank to the implementation of the SDN controller and the
+delivery through function chaining in high capacity networks exploiting
+aggregate flow steering. The architecture proposed makes use of SDN and cloud
+platforms to deploy VMs that will host VNFs. Assumption are made about the SDN
+network elements: connection through optical circuit switching and use of Wave
+Selective Switching (WSS) to route (switch) signals between optical fibers on a
+per-wavelength basis. Thanks to the implementation of the SDN controller and the
 communication between the previous element and the VNFM, this architecture is
-able to provide good capabilities in terms of flexibility. Another element
-that helps the flexibility is the high performance link: in this way they get
-rid of problems in terms of connection time and bandwidth. However, in the paper
-authors assumes really specific hardware in that underlying infrastructure that
+able to provide good capabilities in terms of flexibility. Another element that
+helps the flexibility is the high performance link: in this way they get rid of
+problems in terms of connection time and bandwidth. However, in the paper
+authors assumes really specific hardware in the underlying infrastructure that
 connects the overall system: this is not suitable to be deployed in a general
-purpose infrastructure or using IaaS. It will be interesting to evaluated the
+purpose infrastructure or using IaaS. It will be interesting to evaluate the
 same work on different instrumentation. A clashing aspect is the time to make a
@@
 specialized on video transdecoding, used event to evaluate the performances of
-the whole system. Chaining is performed exploiting layer 2 connectivity,
-using OpenvSwitch as virtual network switches an OpenDayLight in order to
-route the traffic on a set of policies. However the solution does not provide
-any mechanism to automatic deployment of chains and services: this can be an
-issue in particular in case of VNF fault. Orchestration and some mechanisms of
-fault tolerance are critical in particular in a real environment, enabling
+the whole system. Chaining is performed exploiting layer 2 connectivity, using
+OpenvSwitch as virtual network switches an OpenDayLight in order to route the
+traffic on a set of policies. However the solution does not provide any
+mechanism to perform an automatic deployment of chains and services: this can be
+an issue in particular in case of VNF fault. Orchestration and some mechanisms
+of fault tolerance are critical in particular in a real environment, enabling
 autoscaling on user demand or traffic load. Automated deployment also make
@@
 interrupt end-to-end communication to bring up the VNF as soon as possible. This
-recreation process can be slow due to the usage of VMs to deploy the
-functions, that could take minutes to startup.
+recreation process can be slow due to the usage of VMs to deploy the functions,
+that could take minutes to startup.
 
@@
 of part of the system on the host machine: in this way there is not a fully
-virtualized environment. The matter issue has some implication: the VNF
-has a strong relation on the host on which is running and the system could not
-be simply be moved from one machine to another as it could be done only using
+virtualized environment. The former issue has some implication: the VNF has a
+strong relation on the host on which is running and the system could not be
+simply be moved from one machine to another as it could be done only using
 containers. Finally this paper does not provide any form of system evaluation

### 2018-11-21 res/sections/25-TestsResponsiveness.tex

@@
 \begin{tabular}{l|c|c}
-                     & \textbf{Standard deviation (seconds)} & \textbf{Average start up time (seconds)} \\ \hline
-\textbf{Docker}      & 0.009620874                           & 0,016                                    \\
-\textbf{Virtual Box} & 0.755                                 & 20,211                                 
+                     & \textbf{Standard deviation (seconds)} & \textbf{Average 
+start up time (seconds)} \\ \hline
+\textbf{Docker}      & 0.009620874                           & 0.016        \\
+\textbf{Virtual Box} & 0.755                                 & 20.211
 \end{tabular}%
@@
 
-% The overall time required by the traitional virtualization system is
-% approximately of 33 minutes, witch gives a significant insight of how much time
-% is required to create and boot a virtualized enviroment.
+% The overall time required by the traditional virtualization system is
+% approximately of 33 minutes, witch gives a significant insight of how much 
+% time is required to create and boot a virtualized environment.
 
@@
 \label{chap:tests:sec:sfclength:tab:sfcdata}\\
+
 \end{longtable}
+
+\paragraph*{Considerations on container start-up in Kubernetes}
+Regarding the start-up times in Kubernetes is possible to denote a considerable
+difference between the deployment of a SFC with only one element
+(\textit{de-facto} only one container) with a single container directly launched
+on Docker. The difference, of many seconds, can be easily explained: the
+creation of a Docker container, \textit{per se}, is quite fast compared to the
+scheduling job the Kubernetes cluster has to do before launching the pod. A
+Kubernetes deployment is indeed a rather complex operation than launching a
+simple container. When a new deployment is launched on Kubernetes, the operation
+initiate with the scheduling of a Pod. A Pod can have several lifecycle
+states~\cite{kubePodLifecycle}:
+\begin{description}
+\item[pending:] The Pod has been accepted by the Kubernetes system, but one or
+  more of the Container images has not been created. This includes time before
+  being scheduled as well as time spent downloading images over the network,
+  which could take a while;
+\item[running:] The Pod has been bound to a node, and all of the Containers have
+  been created. At least one Container is still running, or is in the process of
+  starting or restarting;
+\item[succeeded:] All Containers in the Pod have terminated in success, and will
+  not be restarted;
+\item[failed:] All Containers in the Pod have terminated, and at least one
+  Container has terminated in failure. That is, the Container either exited with
+  non-zero status or was terminated by the system;
+\item[unknown:] For some reason the state of the Pod could not be obtained,
+  typically due to an error in communicating with the host of the Pod.
+\end{description}
+
+On top of that, different conditions can affect a Pod:
+\begin{itemize}
+\item \verb!PodScheduled!: the Pod has been scheduled to a node;
+\item \verb!Ready!: the Pod is able to serve requests and should be added to the
+  load balancing pools of all matching Services;
+\item \verb!Initialized!: all ``Init Containers'' have started successfully;
+\item \verb!Unschedulable!: the scheduler cannot schedule the Pod right now, for
+  example due to lacking of resources or other constraints.
+\end{itemize}
+
+These additional operations implicate some time overhead, that cause the
+deployment of a single Pod with a single container to be way longer that the
+deployment of a simple container directly from Docker.

### 2018-11-21 res/sections/27-TestsFinalConsiderations.tex

@@
 software in a production environment, eliminating the possibility of human
-errors and assuring longer uptimes.
+errors and assuring longer up-times.

### 2018-11-21 res/sections/25-TestsResponsiveness.tex
Fix dictionary

@@
 To make this simulation as fair as possible, we used the same node (an 
-Openstack VM vith 32GB of RAM, 8 vCPU and solid state storage) and used the 
+Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the 
 same ``boot sequence'' for both the VNFs: first we launched the Astaire 
@@
 difference between the deployment of a SFC with only one element
-(\textit{de-facto} only one container) with a single container directly launched
+(\textit{de facto} only one container) with a single container directly launched
 on Docker. The difference, of many seconds, can be easily explained: the

### 2018-11-21 res/sections/10-ProjectAnalysis.tex

@@
   Function (SF)\footnote{From now on the two terminologies will be use in a
-    interchangeable way.}. \todo{Add ref to chap about SFC?}
+    interchangeable way~\cite{medhat2017service}.}.
+  \todo{Add ref to chap about SFC?}
 

### 2018-11-21 res/sections/10-ProjectAnalysis.tex
Add bib and initial description

--- 2018-11-16 res/sections/10-ProjectAnalysis.tex
@@
 \subsubsection{OpenvSwitch} \todo{Talk about OpenvSwitch!!}
+
+\subsubsection{Floodlight}
+\label{chap:prjan:sec:tech:sub:other:sub:floodlight}
+Community-developed and open source, Floodlight is an OpenFlow Controller
+written in Java born in 2011~\cite{floodlightGit}. It allows various
+applications to run on top of it. Floodlight has the role of controlling and
+inquiring an OpenFlow network, querying the applications integrated with the
+purpose of solving different user needs over the network. When instantiated,
+OpenSwitch routers have to register to the controller in order to gain routing
+rules from the controller.
+
 \subsubsection{Openbaton}

### 2018-11-21 res/sections/10-ProjectAnalysis.tex
Add new software descriptions and fix TODO

--- 2018-11-16 res/sections/10-ProjectAnalysis.tex
@@
 performance. The ETSI infrastructure, instead, tries with container
-virtualization technology to push for a better NFV responsiveness\todo{Check
-  term - I don't think it actually exists}, removing the additional operating
-system (called ``guest'' OS) required by the usual hyper-visor-oriented 
-approach.
+virtualization technology to push for a better NFV responsiveness, removing the
+additional operating system (called ``guest'' OS) required by the usual
+hyper-visor-oriented approach.
 
@@
   case, no state is maintained, and every packet is treated without any
-  particular assumption. Following the RFC 7665, the VNF are also called Service
-  Function (SF)\footnote{From now on the two terminologies will be use in a
-    interchangeable way.}. \todo{Add ref to chap about SFC?}
+  particular assumption. Following the RFC 7665, the VNFs are also called
+  Service Function (SF)\footnote{From now on the two terminologies will be use
+    in a interchangeable way.}. Additional infomation can be found in
+  Section~\ref{chap:prjan:sfc}.
 
@@
   different responsibilities, such as managing the
-  creation/destruction/modification of the SFCs, that achieves interacting
-  with the VIM.\todo{CITE: SDN/NFV-Based Mobile Packet Core Network
-Architectures: A Survey}. Finally, the MANO can integrate with 
-Operational and Business Support Systems (OSS/BSS), \todo{Expand this part with 
-the following papers: https://www.tandfonline.com/doi/abs/10.1057/jors.1995.176 
-https://ieeexplore.ieee.org/abstract/document/6963800 
-https://www.computer.org/csdl/proceedings/ifcsta/2009/3930/01/3930a466-abs.html}
-
+  creation/destruction/modification of the SFCs, that achieves interacting with
+  the VIM~\cite{nguyen2017sdn}. Finally, the MANO can be integrated with
+  Operational and Business Support Systems (OSS/BSS).
 \end{itemize}
@@
 \section{SFC}
+\label{chap:prjan:sfc}
 
@@
 \label{chap:prjan:sec:openstack}
-Created in ??\todo{Find out Openstack date of creation}, Openstack allows
-on-premise cloud installations, using bare-metal resources to provide common
-cloud services as object storage, virtual machine deploy, virtual networking.
-Furthermore its modularity offers the possibility to add additional components,
-even proprietary, to achieve a complete cloud solution.
+Created in 2010 by a joint operation between Rackspace Hosting and
+NASA~\cite{openstackWebsite}, Openstack allows on-premise cloud installations,
+using bare-metal resources to provide common cloud services as object storage,
+virtual machine deploy, virtual networking. Furthermore its modularity offers
+the possibility to add additional components, even proprietary, to achieve a
+complete cloud solution.
 \begin{figure}[t]
@@
 specialized in container management and orchestration. Developed by Google in
-??\todo{Find out the actual date} and written in Go, it allows different
-machines (called nodes from now on) to create an abstraction layer and to form a
-cluster where is possible to run Docker containers without bothering about
-hardware constraints. Nodes can have specific roles and specialization, all of
-them managed by one or many master nodes, that actually manage the
-orchestration. On top of that, Kubernetes is virtualization-agnostic, meaning
-that it offers the possibility to change the virtualization software in use
-(i.e. from Container-based technology to Virtual machine software).
+2014 and written in Go~\cite{k8sGit}, it allows different machines (called nodes
+from now on) to create an abstraction layer and to form a cluster where is
+possible to run Docker containers without bothering about hardware constraints.
+Nodes can have specific roles and specialization, all of them managed by one or
+many master nodes, that actually manage the orchestration. On top of that,
+Kubernetes is virtualization-agnostic, meaning that it offers the possibility to
+change the virtualization software in use (i.e. from Container-based technology
+to Virtual machine software).
 \begin{figure}[h]
@@
 deployment containing only simple services (e.g. a back-end application serving
-a
-website or a database) this ends up in a waste of resources. \todo{Talk about
-  lxc containers and how Docker solved magically the problem.} It's here that,
-in ??\todo{Find Docker date of birth}, Docker was created, basing its solution
-on an already existing product: Linux Container (LXC). From this framework,
-Docker built an entire ecosystem, consisting in a client/server model giving the
+a website or a database) this ends up in a waste of resources. \todo{Talk about
+  lxc containers and how Docker solved magically the problem.} It is here that,
+in 2013~\cite{dockerWebsite}, Docker was created, basing its solution on an
+already existing product: Linux Container (LXC). From this framework, Docker
+built an entire ecosystem, consisting in a client/server model giving the
 possibility for users to simply launch, scale and delete containers (locally or
@@
 \label{chap:prjan:sec:tech:sub:other:sub:swarm}
-Introduced in ??\todo{Find Docker Swarm date of birth}, Docker Swarm allows 
-multiple Docker nodes to cluster together and be seen as one logical unit. This 
-layer completely makes the underlying infrastructure transparent: data 
-management, as network one, are totally managed by Docker. Recently, 
-Docker-compose support has been introduced, making this one-node tool available 
-for clustered systems. At the end of the day, on one hand Docker Swarm provides 
-an easy way to set up a cluster system with all the tools configured 
-out-of-the-box, without the necessity to set networking configurations or 
-installing ad-hoc storage solutions. On the other hand, though, it doesn't have 
-all the customization options that Kubernetes offers, thus making the product 
-less flexible. This fundamental characteristic, although, makes the two 
-solutions have different use cases leading to different market needs.
+Introduced in 2014~\cite{swarmGit}, Docker Swarm allows multiple Docker nodes to
+cluster together and be seen as one logical unit. This layer completely makes
+the underlying infrastructure transparent: data management, as network one, are
+totally managed by Docker. Recently, Docker-compose support has been introduced,
+making this one-node tool available for clustered systems. At the end of the
+day, on one hand Docker Swarm provides an easy way to set up a cluster system
+with all the tools configured out-of-the-box, without the necessity to set
+networking configurations or installing ad-hoc storage solutions. On the other
+hand, though, it doesn't have all the customization options that Kubernetes
+offers, thus making the product less flexible. This fundamental characteristic,
+although, makes the two solutions have different use cases leading to different
+market needs.
 
-\subsubsection{OpenvSwitch} \todo{Talk about OpenvSwitch!!}
+\subsubsection{OpenVSwitch}
+Started in 2009~\cite{ovsGit}, the OpenVSwitch project aims to create an
+open-source muti-layer virtual switch. It provides a switching stack for
+hardware virtualization environments, implementing several interfaces and
+protocols. It has also the possibility to run through Hypervisor virtualization
+but it can run under Docker too~\cite{ovsDocker}. OpenVSwitch can route packets
+from incoming connections following rules, that can be ``pushed'' in its routing
+table from software defined as \textit{controllers}. Additionally, it supports the
+OpenFlow communication protocol, in order to set previously cited routing
+definitions.
+
+\subsubsection{Floodlight}
+\label{chap:prjan:sec:tech:sub:other:sub:floodlight}
+Community-developed and open source, Floodlight is an OpenFlow Controller
+written in Java at the end of 2011~\cite{floodlightGit}. It allows various
+applications to run on top of it. Floodlight has the role of controlling and
+inquiring an OpenFlow network, querying the applications integrated with the
+purpose of solving different user needs over it. When instantiated, OpenSwitch
+routers have to register to the controller in order to gain routing rules.
+
 \subsubsection{Openbaton}
 \label{chap:prjan:sec:tech:sub:other:sub:openbaton}
-Patronized by the Institute Fraunhofer, it's a open-source, customizable NFV 
+Patronized by the Institute Fraunhofer, it is an open-source, customizable NFV
 MANO-compliant framework, that offers many configurations options, and since
-it's 
-written in Java, there is the possibility to add components (plug-ins) 
-dynamically. Openbaton can be installed as a normal computer program (they 
-offer packages for the most common distros) but it can be deployed on Docker 
-containers too. It's designed to work with cloud providers like Amazon or 
-Google Compute engine, but it offers compatibility with on-premise cloud 
-solutions like Openstack, where it deploys virtual machines using the API
-the platform offers. It can store VNF configurations saved as TOSCA
-YAML\todo{What does TOSCA mean? Find out} and it can handle multiple
-PoPs\todo{Write down PoP meaning}. It offers a VNF lifecycle management
-out-of-the-box, with the possibility to customize it based on the user needs.
-Its modular design allows developers to change parts of the codebase with custom
-ones, making the product really flexible to different platforms. Performing this
-operation requires a deep knowledge of the framework though.
+it's written in Java, there is the possibility to add components (plug-ins)
+dynamically. Openbaton can be installed as a normal computer program (they offer
+packages for the most common distros) but it can be deployed on Docker
+containers too. It's designed to work with cloud providers like Amazon or Google
+Compute engine, but it offers compatibility with on-premise cloud solutions like
+Openstack, where it deploys virtual machines using the API the platform offers.
+It can store VNF configurations saved as Topology and Orchestration
+Specification for Cloud Applications (TOSCA) Yet Another Markup Language (YAML)
+and it can handle multiple Point of Presence (PoP). It offers a VNF lifecycle
+management out-of-the-box, with the possibility to customize it based on the
+user needs. Its modular design allows developers to change parts of the codebase
+with custom ones, making the product really flexible to different platforms.
+Performing this operation requires a deep knowledge of the framework though.
 \begin{figure}[h]

### 2018-11-21 res/sections/10-ProjectAnalysis.tex

@@
   case, no state is maintained, and every packet is treated without any
-  particular assumption. Following the RFC 7665, the VNFs are also called
-  Service Function (SF)\footnote{From now on the two terminologies will be use
-    in a interchangeable way.}. Additional infomation can be found in
-  Section~\ref{chap:prjan:sfc}.
+  particular assumption. Following the RFC 7665, the VNF are also called Service
+  Function (SF)\footnote{From now on the two terminologies will be use in a
+    interchangeable way~\cite{medhat2017service}.}. Additional infomation can 
+  be found in Section~\ref{chap:prjan:sfc}.
 

### 2018-11-21 res/sections/10-ProjectAnalysis.tex
Fix typo

--- 2018-11-21 res/sections/10-ProjectAnalysis.tex
@@
   Service Function (SF)\footnote{From now on the two terminologies will be use
-    in a interchangeable way.}. Additional infomation can be found in
+    in a interchangeable way.}. Additional information can be found in
   Section~\ref{chap:prjan:sfc}.
@@
 Started in 2009~\cite{ovsGit}, the OpenVSwitch project aims to create an
-open-source muti-layer virtual switch. It provides a switching stack for
+open-source multi-layer virtual switch. It provides a switching stack for
 hardware virtualization environments, implementing several interfaces and
@@
 from incoming connections following rules, that can be ``pushed'' in its routing
-table from software defined as \textit{controllers}. Additionally, it supports the
-OpenFlow communication protocol, in order to set previously cited routing
+table from software defined as \textit{controllers}. Additionally, it supports
+the OpenFlow communication protocol, in order to set previously cited routing
 definitions.

### 2018-11-21 res/sections/10-ProjectAnalysis.tex
Fix additional typos

@@
   case, no state is maintained, and every packet is treated without any
-  particular assumption. Following the RFC 7665, the VNFs are also called
-  Service Function (SF)\footnote{From now on the two terminologies will be use
-    in a interchangeable way.}. Additional information can be found in
-  Section~\ref{chap:prjan:sfc}.
+  particular assumption. Following the RFC 7665, the VNF are also called Service
+  Function (SF)\footnote{From now on the two terminologies will be use in a
+    interchangeable way~\cite{medhat2017service}.}. Additional information can
+  be found in Section~\ref{chap:prjan:sfc}.
 
@@
 inquiring an OpenFlow network, querying the applications integrated with the
-purpose of solving different user needs over it. When instantiated, OpenSwitch
+purpose of solving different user needs over it. When instantiated, OpenVSwitch
 routers have to register to the controller in order to gain routing rules.

### 2018-11-21 res/sections/07-Introduction.tex
Add some related works
Fix TODOs, add citation

@@
  Through the history of internet, many factors changed and new innovations came,
- but at the end of the day a fundamental key remained: connection performances
- are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
-   performs quite well in wired connections, while in wireless one, such as
-   wireless/satellite, sucks ass}
-With the arrival of mobile connectivity, wireless networks started to have an 
-important role in everyone's life. In this context network efficiency plays 
-an important role with the development of new connectivity standards (e.g 2G, 
-3G, LTE) and the creation of common, ad-hoc versions, transmission protocols 
-(i.e. TCP versions specialized for exploit the characteristic of 
-wireless networks). It is for this reason that, beside improved radio 
-communications, back-end infrastructure began to be refined too. During their 
-communication to the destination packets are elaborated, compressed and 
-processed in order to make their transmission as fast as possible.
-On top of that, mobile traffic is continuously surging every year, with the 
-backbones having more and more data to deliver, making in-hardware backbone 
-solutions less efficient. Their time-to-deploy, in fact, is very high compared 
-to the requirements of the network. With this trend, in-hardware solutions 
-will not be able to keep up. Is with this considerations that a software 
-approach is being proposed to become the new way to deploy network 
-components. While less efficient, they are faster to bring them up, offering an 
-unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile Packet Core Network 
-Architectures: A Survey}.
- Furthermore, new virtualization technologies such as Cloud computing and Docker
- are becoming more and more mature for being employed in large-scale production
- environments, where new frameworks, called orchestrators (e.g. Kubernetes,
- Docker Swarm), allow to easily manage the application lifecycle, i.e.
- deployment, scale up/down and finally, removal.
+ but at the end of the day a fundamental key remains: connection performances
+ are, indeed, what the users really want. With the arrival of mobile
+ connectivity, wireless networks started to have an important role in everyone's
+ life. In this context network efficiency is a decisive factor in the
+ development of new connectivity standards (e.g 2G, 3G, LTE) and in the creation
+ of ad-hoc transmission protocols (i.e. TCP versions specialized to exploit the
+ characteristic of wireless networks, such as Hybla and Vegas). It is for this
+ reason that, beside improved radio communications, back-end infrastructure
+ began to be refined too. During their communication to the destination packets
+ are elaborated, compressed and processed in order to make their transmission as
+ fast as possible. On top of that, mobile traffic is continuously surging every
+ year, with the backbones having more and more data to deliver, making
+ in-hardware backbone solutions less efficient. Their time-to-deploy, in fact,
+ is very high compared to the requirements of the network. With this trend,
+ in-hardware solutions will not be able to keep up. Is with this considerations
+ that a software approach is being proposed to become the new way to deploy
+ network components. While less efficient, they are faster to bring them up,
+ offering an unprecedented flexibility~\cite{nguyen2017sdn}. Furthermore, new
+ virtualization technologies such as Cloud computing and Docker are becoming
+ more and more mature for being employed in large-scale production environments,
+ where new frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm),
+ allow to easily manage the application lifecycle, i.e. deployment, scale
+ up/down and finally, removal.
+
+ \newpage
 
@@
 and the technologies used to build an ETSI compliant architecture
+  \item Chapter~\ref{chap:rel_wk} gives an insight of correlated works on this 
+field
   \item Chapter~\ref{chap:prjan} explains the analysis performed on the 

### 2018-11-21 res/sections/08-Background.tex

@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
   \includegraphics[scale=0.2]{networkslicing}
-  \caption{The three layers model of network
-  slicing~\cite{alliance2016description}}
+  \caption[The three layers model of network slicing]{The three layers model of
+    network slicing~\cite{alliance2016description}}
   \label{chap:background:img:network_slicing}
@@
   \includegraphics[scale=2]{etsi_arch}
-  \caption{Network Function Virtualization reference
-  architecture~\cite{etsi2013gs}}
+  \caption[Network Function Virtualization reference architecture]{Network
+    Function Virtualization reference architecture~\cite{etsi2013gs}}
   \label{chap:background:img:etsi_arch}
@@
   \includegraphics[scale=0.6]{nshformat}
-  \caption{Network Service Header~\cite{rfc8300}}
+  \caption[Network Service Header]{Network Service Header~\cite{rfc8300}}
   \label{chap:background:img:nshformat}
@@
   \includegraphics[scale=0.6]{nshbase}
-  \caption{Network Service Header~\cite{rfc8300}}
+  \caption[Network Service Header]{Network Service Header~\cite{rfc8300}}
   \label{chap:background:img:nshformat}
@@
   \includegraphics[scale=0.6]{nshservicepath}
-  \caption{NSH Service Path Header~\cite{rfc8300}}
+  \caption[NSH Service Path Header]{NSH Service Path Header~\cite{rfc8300}}
   \label{chap:background:img:nshpath}
@@
   \includegraphics[scale=0.6]{nshtype1}
-  \caption{NSH MD Type \texttt{0x1}~\cite{rfc8300}}
+  \caption[NSH MD Type \texttt{0x1}]{NSH MD Type \texttt{0x1}~\cite{rfc8300}}
   \label{chap:background:img:nshtype1}
@@
   \includegraphics[scale=0.6]{nshtype2}
-  \caption{NSH MD Type \texttt{0x2}~\cite{rfc8300}}
+  \caption[NSH MD Type \texttt{0x2}]{NSH MD Type \texttt{0x2}~\cite{rfc8300}}
   \label{chap:background:img:nshtype2}
@@
   \includegraphics[scale=0.6]{nshtype2variable}
-  \caption{Variable-Length Context Header~\cite{rfc8300}}
+  \caption[Variable-Length Context Header]{Variable-Length Context
+    Header~\cite{rfc8300}}
   \label{chap:background:img:nshtype2variable}
@@
   \includegraphics[scale=0.55]{5garchitecture}
-  \caption{5G Architecture~\cite{alliance20155g}}
+  \caption[5G Architecture]{5G Architecture~\cite{alliance20155g}}
   \label{chap:background:img:5garchitecture}

### 2018-11-21 res/sections/81-k8s.tex
Add k8s appendix

\chapter{Kubernetes}\label{kubernetes}

In this appendix we provide a basic guide to Kubernetes concept and
installation. This is not thought as an complete overview of functionalities but
there are only reported a description of concept an feature used in the
implementation of our solution.

\section{Kubernetes basic concepts}\label{kubernetes-basic-concepts}

\subsection{Pod}\label{pod}

A pod is the most basic building block for Kubernetes. A pod usually
contains one or more container (if the containers are highly coupled),
even if Kubernetes it's able to manage different virtualization technologies
(e.g.
Virtual Machines\footnote{\url{https://github.com/kubevirt/kubevirt}}).
Pods have a life cycle, where it can assume these
states\footnote{
\url{https://kubernetes.io/docs/concepts/workloads/pods/pod-lifecycle/}}:
% TODO davide must delete from test
\begin{itemize}
\item
  \textbf{Pending}: The Pod has been accepted by the Kubernetes system,
  but one or more of the Container images has not been created. This
  includes time before being scheduled as well as time spent downloading
  images over the network, which could take a while.
\item
  \textbf{Running}: The Pod has been bound to a node, and all of the
  Containers have been created. At least one Container is still running,
  or is in the process of starting or restarting.
\item
  \textbf{Succeeded}: All Containers in the Pod have terminated in
  success, and will not be restarted.
\item
  \textbf{Failed}: All Containers in the Pod have terminated, and at
  least one Container has terminated in failure. That is, the Container
  either exited with non-zero status or was terminated by the system.
\item
  \textbf{Unknown}: For some reason the state of the Pod could not be
  obtained, typically due to an error in communicating with the host of
  the Pod.
\end{itemize}

A pod can have different conditions too:
% TODO davide must delete from test
\begin{itemize}
\item \textbf{PodScheduled}: the Pod has been scheduled to a node;
\item \textbf{Ready}: the Pod is able to serve requests and should be added
  to the load balancing pools of all matching Services;
\item \textbf{Initialized}: all init containers\footnote{
  \url{https://kubernetes.io/docs/concepts/workloads/pods/init-containers}}
  have started successfully;
\item \textbf{Unschedulable}: the scheduler cannot schedule the Pod right
  now, for example due to lacking of resources or other constraints;
\end{itemize}

Each pod can have different QoS classes:
\begin{description}
  \item [Guaranteed:] every container in the Pod must have a memory and CPU
  limits and memory and CPU requests, and they must be the same;
  \item [Burstable:] the Pod does not meet the criteria for QoS class
  Guaranteed and at least one container in the Pod has a memory or CPU request;
  \item [BestEffor:] the containers in the Pod must not have any memory or CPU
  limits or requests.
\end{description}

\subsection{Service}\label{service}

Since pods for their nature are highly volatile, applications that use
them can not rely on their IP address. To solve this problem, the
Service implementation in Kubernetes assure a stable IP, even if
different pods in this service crash and gets restarted by a ReplicaSet
(witch will not be explain in this guide, but it basically scale up and
down pods\footnote{
\url{https://kubernetes.io/docs/concepts/workloads/controllers/replicaset/}})

Citing the official documentation\footnote{
\url{https://kubernetes.io/docs/concepts/services-networking/service/}}, the
Service is:\\
\textquote{A Kubernetes Service is an abstraction which defines a logical set
of Pods and a policy by which to access them - sometimes called a
micro-service.}

\begin{figure}[htbp]
\centering
\includegraphics{k8s_services}
\caption{Example of a service}
\end{figure}

\subsection{Volume}\label{volume}

When Pods saves data on a node, this information is designated to perish.
Kubernetes offers a way to save all data and to possibly make it
available to all the pods replicas, offering a transparent layer where
pods are not aware of working with information shared other pods, even located
in other nodes. This functionality is called Volume, and there are many
different drivers, that range from a simple local storage to glusterfs.

\subsection{Namespace}\label{namespace}

Namespaces are intend to be used when there are multiple users using the
cluster, or to separate multiple projects.

\section{Kubernetes architectural concepts}
\label{kubernetes-architectural-concepts}

\subsection{Node}\label{node}

A node is the building block for a Kubernetes cluster. Everyone must have
at least the kubelet daemon (to manage Kubernetes functionalities) and a
container manager (e.g. Docker Engine, VirtualBox) running. A node runs pods
and can communicate with other nodes in the same cluster

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.8]{k8s_nodes}
\caption{Schema illustrating the elements in a node}
\end{figure}

\subsection{Master}\label{master}

The \emph{master node} is the one that starts a cluster, and that manage the
slaves. Only some pods can run n the master node (e.g. the dashboard).

The master generates the token for the slaves to join, and eventually
coordinates the federation with other clusters. In a cluster, there is the
possibility to have multiple master to ensure high availability.

\subsection{Slave}\label{slave}

A \emph{slave} is a node that run the pods that the master schedule. It can
also be part of a \emph{PersistantVolume} storage (like GlusterFS) and if it
goes down, another slave will take care of the pods that where running on
it. Slaves can join existing cluster via a given token.

\subsection{Networking}\label{networking}

Kubernetes has its own networking structure, even if it uses Docker for running
containers. In particular, it delegates the network operations (like packet
forwarding) to different managers, such as Flannel or Calico.

Every nodes in the cluster runs a pod called \texttt{kubeproxy}, that
communicates with the \texttt{kube-api} pod (running on master) and
the \texttt{kube-dns} (for Kubernetes \textless{}=1.9.x) or
\texttt{CoreDNS} (for Kubernetes \textgreater{}=1.10.x). \texttt{kubeproxy}
assure transparent communication between pods in the same cluster, and
\texttt{kube-api} allows to call Kubernetes services. kube-dns/CoreDNS manage
DNS request coming from the pods, differentiating the one that goes outside
the cluster from the other that point internally\footnote{Additional
information about Flannel, Kubernetes and Docker networking
can at:\\
\sloppy
\url{https://www.sdxcentral.com/cloud/containers/definitions/what-is-coreos-flannel-definition/}}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.3]{flannelredirect}
\caption{And example of how flannel redirect packages between pods}
\end{figure}

\section{Focusing on Storage}
\label{focusing-on-storage}

This section is based on official Kubernetes documentation\footnote{
\url{https://kubernetes.io/docs/concepts/storage/persistent-volumes/}} and
from \texttt{newstack site}\footnote{
\sloppy
\url{https://thenewstack.io/strategies-running-stateful-applications-kubernetes-persistent-volumes-claims/}}
.

\subsection{Volumes, Claims and Classes}
\label{volumes-claims-and-classes}

Managing compute and storage are totally different problems and in order
to abstract the last one Kubernetes relies on \textbf{PersistantVolume}
and \textbf{PersistantVolumeClaim} resources.

A PersistantVolume (\textbf{PV}) is something similar to a node, in fact
is a piece of storage provided by the administrator: it captures the
details of the implementation of the storage and has a lifecycle that is
independent of any individual pod. To not expose developers to storage
boring details admins can expose different PV types through
\textbf{StorageClasses}.

A PersistantVolumeClaim (\textbf{PVC}) instead is a request for storage
and it is similar a pod: instead of requesting memory and CPU, a PVC
requests certain size and access mode to storage.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.18]{kubernetes_pvc}
\caption{PersistantVolumeClaims management}
\end{figure}

It is important to understand the difference on \textbf{Volumes} and
\textbf{PersistantVolumes}/\textbf{PersistantVolumeClaims}. The former is
something like Docker volumes, used for persistence, host-based or on external
block storage devices but a user does not expect resources be reserved before
using volumes. In persistent volumes and claims, there is a strict enforcement
of resource utilization dictated by the policy defined during the
creation of resources. Kubernetes pod can use claims as volumes.

\textbf{StorageClasses}\footnote{more information on different provisioners,
policies and parameters can be found here: \\\url{
https://kubernetes.io/docs/concepts/storage/storage-classes/
}} are a way to describe different type of storage
that can be used. Different classes can be used to describe different
levels of service or policies. PVCs that don't specify a particular class
can use the default one: to define it you can use annotation
\verb!"storageclass.kubernetes.io/is-default-class":"true"!, as for
instance:

\begin{lstlisting}[language=bash]
kubectl patch storageclass glusterfs-storage -p '{"metadata": {"annotations":{"storageclass.kubernetes.io/is-default-class":"true"}}}'
\end{lstlisting}

\subsection{Lifecycle}\label{lifecycle}

\begin{enumerate}
  \item \textbf{Provisioning}: there are 2 ways: statically or dynamically.
  \emph{Static} means that an administrator creates some PVs that are
  available for consumption. \emph{Dynamic} means that the cluster
  dynamically try to provide a volume that matches a PVC. This is done
  with \emph{StorageClasses}: a claim requests storage for a certain
  storage class that can be provisioned. The class \texttt{""} disable
  dynamic provisioning;
\item
  \textbf{Binding}: a user requests a certain amount of storage with a
  claim and a control loop in the master look for a PV that can bind the
  claim. In case of dynamic provisioning the loop always bind a PVC to a
  PV, otherwise user will always get at least what they asked for, but
  the volume may be larger. Claims can remain unbound if a matching
  volume does not exists.
\item
  \textbf{Using}: after the binding phase the pod start using the volume
  (PVC is described in the pod yaml
  \footnote{\url{https://kubernetes.io/docs/concepts/storage/persistent-volumes/\#claims-as-volumes}}
\item
  \textbf{Releasing}: when an application is done with using the volume,
  developers can delete the PVC objects through the API, releasing the
  claim. This step will initiate the reclamation process.
\item
  \textbf{Reclaiming}: this phase is defined as a policy by the
  administrators. The reclaim policy for a PersistantVolume tells the
  cluster what to do with the volume after it has been released of its
  claim. Persistent volumes can either be \emph{retained},
  \emph{recycled} or \emph{deleted}.
\end{enumerate}

\section{Installing Kubernetes}\label{installing-kubernetes}

We performed our Kubernetes installation in a Openstack environment.
Since we used the Openstack of our school department, the administrator
have given us a virtual machine, Openstack-CLI, to perform
operations. From now, our scripts assume you're in a machine with access
to Openstack, and that you have sourced your openrc.sh script before
continuing.

\subsection{Installing Kubernetes on CentOS}
\label{installing-kubernetes-on-centos}

To install Kubernetes on CentOS 7 we develop a script\footnote{available
at \url{https://github.com/Augugrumi/init-script/tree/centos/kubernetes}}.
With CentOS is possible to have a complete installation with
a GlusterFS-PersistentStorage correctly set.

After there is an explanation for the most important part of the
scripts. The script \verb!on_openstack_cli.sh!
\footnote{\url{https://github.com/Augugrumi/init-script/raw/centos/kubernetes/on_openstack_cli.sh}}
main aim is to run other file retrieving information using even
\texttt{nova} commands.

First all the software needed must to be installed an for this the
script \\\verb!install_kubectl_repo.sh!
\footnote{\url{https://raw.githubusercontent.com/Augugrumi/init-script/centos/kubernetes/install_kubectl_repo.sh}}
was created. This only call another script and wait until all machine
are ready. The other script is \verb!bootstrap.sh!
\footnote{\url{https://raw.githubusercontent.com/Augugrumi/vagrantfiles/oldversion/kubernetes/centos/bootstrap.sh}}
whose aim is to configure the machine and the install the necessary
software.\\
Boostrap script first \textbf{disable SELinux} on all machines (needed
for Kubernetes networking)

\begin{lstlisting}
sudo setenforce 0
sudo sed -i --follow-symlinks 's/SELINUX=enforcing/SELINUX=disabled/g' /etc/sysconfig/selinux
\end{lstlisting}

then \textbf{enable br\_netfilter} module (needed for Kubernetes
networking)

\begin{lstlisting}
sudo modprobe br_netfilter
sudo sh -c 'echo "1" > /proc/sys/net/bridge/bridge-nf-call-iptables'
\end{lstlisting}

\textbf{Switch off the swap} by the commands

\begin{lstlisting}
sudo swapoff -a
sudo sed -i '/swap/d' /etc/fstab
\end{lstlisting}

And finally add \texttt{rbd}, \texttt{ip\_vs}, \texttt{ip\_vs\_rr},
\texttt{ip\_vs\_wrr}, \texttt{ip\_vs\_sh}, \texttt{dm\_snapshot},
\texttt{dm\_mirror} and \texttt{dm\_thin\_pool} to modules that will be
loaded to machine startup.

\begin{lstlisting}
sudo sh -c 'echo -e "rbd\nip_vs\nip_vs_rr\nip_vs_wrr\nip_vs_sh" >  /etc/modules-load.d/ip_vs.conf'
sudo sh -c 'echo -e "dm_snapshot\ndm_mirror\ndm_thin_pool" >  /etc/modules-load.d/gluster.conf'
\end{lstlisting}

Note: Last 3 modules will be useful in the case that you want to install
GlusterFS on your machines.

In the end the script install \texttt{docker},
\texttt{device-mapper-persistent-data}, \texttt{lvm2}, \texttt{kubelet},
\texttt{kubectl} and \texttt{kubeadm} and start (and enable)
\texttt{docker} and \texttt{kubelet} services.

After rebooting all the machine, the installation go on adding in the
\texttt{/etc/hosts} file of all node that will be part of the Kubernetes
cluster the IP and the name of all the nodes.

After the installation continues in the \textbf{master} node, with the script
\verb!on_master.sh!
\footnote{\url{https://raw.githubusercontent.com/Augugrumi/init-script/centos/kubernetes/on_master.sh}}.
We run \texttt{kubeadm} by the command

\begin{lstlisting}
sudo kubeadm init --apiserver-advertise-address=<ip-of-the-master> --pod-network-cidr=10.244.0.0/16 --ignore-preflight-errors cri | grep "kubeadm join" > /home/centos/joincommand
\end{lstlisting}

Saving part of the output that will be necessary for other nodes to join
the cluster.\\
Note 1:\texttt{-\/-pod-network-cidr} depends on which agent you want to
use to configure Kubernetes network. We choose Flannel so we put
\texttt{10.244.0.0/16}.\\
Note 2: we don't know at the moment why some preflight errors was
launched by the command and in order to do so (since everything seems
working) we added \texttt{-\/-ignore-preflight-errors\ cri} to go on.\\
This save in the \texttt{joincommand} file the command to be run on
slave node to join the cluster.

And after give the following commands to create the folder needed by
Kubernetes

\begin{lstlisting}
mkdir -p $HOME/.kube
sudo cp -i /etc/kubernetes/admin.conf $HOME/.kube/config
sudo chown $(id -u):$(id -g) $HOME/.kube/config
\end{lstlisting}

Then wait for all pods that are ready
(\texttt{kubectl\ get\ pods\ -\/-all-namespaces}).

Following, install Flannel\footnote{\url{https://github.com/coreos/flannel}}
for the networking and wait that it is up:

\begin{lstlisting}
kubectl apply -f https://raw.githubusercontent.com/coreos/flannel/master/Documentation/kube-flannel.yml
kubectl apply -f https://raw.githubusercontent.com/coreos/flannel/master/Documentation/k8s-manifests/kube-flannel-rbac.yml
\end{lstlisting}

And then it is possible to setup the dashboard to access Kubernetes resources:

\begin{lstlisting}
kubectl create serviceaccount dashboard -n default
  kubectl create clusterrolebinding dashboard-admin -n default \
   --clusterrole=cluster-admin \
   --serviceaccount=default:dashboard
  kubectl get secret $(kubectl get serviceaccount dashboard -o jsonpath="{.secrets[0].name}") -o jsonpath="{.data.token}" | base64 --decode >  secret.txt
  screen -dmS kubectl_proxy_screen bash
  screen -S kubectl_proxy_screen -X stuff "kubectl proxy
  "
\end{lstlisting}

With this command we make the dashboard accessible on port 8001 of the
master node at this URL
\begin{verbatim}
https://<master-ip>:<apiserver-port>/api/v1/namespaces/kube-system/services/https:kubernetes-dashboard:/proxy/}
\end{verbatim}

The previous command also the token to access the dashboard on
\texttt{secret.txt} file on the home folder of the master node.

The last thing to do is to make other node join the cluster, using the
\texttt{joincommand} file that we created before.

\section{Persistent Storage}\label{persistent-storage}

\subsection{Prerequisites}\label{prerequisites}

\begin{itemize}
\item A working installation of Kubernetes
\item Python (2.7) installed and available on all Kubernetes nodes
\end{itemize}

\subsection{GlusterFS}\label{glusterfs}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.3]{glusterkube}
\caption{Installation of GlusterFS on Kubernetes cluster hosts}
\end{figure}

To enable persistent storage in Kubernetes with Glusterfs we need at
least three nodes. As we've already written above, we deployed our
Kubernetes cluster using CentOS 7.\\
Note: this procedure was tested on \texttt{Kubernetes\ v1.11.1.}

Since we're running our cluster on a Openstack environment, we attached
one volume per VM. If you're in our shoes, just create some empty
volumes and attach them to nodes you want to configure with glusterfs.

First of all, you need to install \texttt{gluster-server} on all the
nodes that will join the storage and \texttt{gluster-fuse} to provide
\texttt{mount.glusterfs} as an option to mount Gluster file systems.
Open a terminal and type:

\begin{lstlisting}
sudo yum install -y centos-release-gluster && \
sudo yum install -y glusterfs-server glusterfs-fuse
\end{lstlisting}

Our next step is to load different kernel modules, using
\texttt{modprobe}:

\begin{lstlisting}
sudo sh -c 'modprobe dm_snapshot && modprobe dm_mirror && modprobe dm_thin_pool'
\end{lstlisting}

To make this change permanent, create a new file in
\texttt{/etc/modules-load.d/gluster.conf} with the following content:

\begin{lstlisting}
dm_snapshot
dm_mirror
dm_thin_pool
\end{lstlisting}

At this point, enable the \texttt{glusterd} daemon to start at every
system start:

\begin{lstlisting}
sudo systemctl start glusterd && sudo systemctl enable glusterd
\end{lstlisting}

For all the nodes you installed glusterfs you need to apply a particular
label in you Kubernetes cluster. In a terminal, type:

\begin{lstlisting}
kubectl label node <node name> storagenode=glusterfs
\end{lstlisting}

According to the \emph{gluster-kubernetes setup guide}
\footnote{\url{https://github.com/gluster/gluster-kubernetes/blob/master/docs/setup-guide.md}},
you should also check that the following ports are reachable:

\begin{itemize}
\item 2222
\item 24007
\item 24008
\item 49152-49251
\end{itemize}

On top of that, you'll need to grant access to your root account while
installing all the stuff we need. In CentOS, if like us you're using the
CloudImg image, you need to access every machine and type:

\begin{lstlisting}
sudo su
cd /root/
nano .ssh/authorized_keys
# edit the first line in the follwing way: from the sub-string starting
# as "ssh-rsa" until the end put the content in a new line, while comment
# the whole line
nano /etc/ssh/sshd_config # un-comment "PermitRootLogin: yes"
systemctl restart sshd
\end{lstlisting}

At this point, you need to clone \texttt{kubernetes-glusterfs}, the repository
that contains a script that will figure out all the necessary stuff
(e.g. heketi), \texttt{gk-deploy}.

\begin{lstlisting}
git clone https://github.com/gluster/gluster-kubernetes.git
cd gluster-kubernetes/deploy
\end{lstlisting}

Now we need to define the topology. To do so, a \texttt{topology.json}
file must be created, from the sample \texttt{topology.json.sample}, and edit to
fit your topology. As in the following example:

\begin{lstlisting}
{
 "clusters": [
   {
     "nodes": [
       {
         "node": {
           "hostnames": {
             "manage": [
               "192.168.29.26"
             ],
             "storage": [
               "192.168.29.26"
             ]
           },
           "zone": 1
         },
         "devices": [
           "/dev/vdb"
         ]
       },
       {
         "node": {
           "hostnames": {
             "manage": [
               "192.168.29.19"
             ],
             "storage": [
               "192.168.29.19"
             ]
           },
           "zone": 1
         },
         "devices": [
           "/dev/vdb"
         ]
       },
       {
         "node": {
           "hostnames": {
             "manage": [
               "192.168.29.22"
             ],
             "storage": [
               "192.168.29.22"
             ]
           },
           "zone": 1
         },
         "devices": [
           "/dev/vdb"
         ]
       }
     ]
   }
 ]
}
\end{lstlisting}

Now it is possible to deploy a Gluster cluster. From the master the following
command must be run:

\begin{lstlisting}[language=bash]
# assuming you still are in ~/blabla/gluster-kubernetes/deploy/
./gk-deploy -s <path/to/your/private/key> --ssh-user root --ssh-port 22 topology.json -y
\end{lstlisting}

At the end, if everything goes correctly, a \texttt{StorageClass} template have
to be created. To your Kubernetes through \texttt{kubectl}:

\begin{lstlisting}[language=bash]
kubectl create -f storageclass.yaml
\end{lstlisting}

Optionally, it can be set as default \texttt{StorageClass} with the following
command:

\begin{lstlisting}[language=bash]
kubectl patch -f storageclass.yaml \
-p'{"metadata": {"annotations":{"storageclass.kubernetes.io/is-default-class":"true"}}}'
\end{lstlisting}

\section{Ingress}\label{ingress}

Ingress makes service available outside the cluster. There are different ways
Kubernetes handle service reachability, and these are:

\begin{description}
  \item [ClusterIP:] Exposes the service on a cluster-internal IP. Choosing this
  value makes the service only reachable from within the cluster. This
  is the default ServiceType.
  \item [NodePort:]
  \footnote{\url{https://kubernetes.io/docs/concepts/services-networking/service/\#nodeport}}
  Exposes the service on each Node's IP at a static port (the NodePort).
  A ClusterIP service, to which the NodePort service will route, is
  automatically created. You'll be able to contact the NodePort service,
  from outside the cluster, by requesting
  \texttt{\textless{}NodeIP\textgreater{}:\textless{}NodePort\textgreater{}}.
  \item [LoadBalancer:]
  \footnote{\url{https://kubernetes.io/docs/concepts/services-networking/service/\#loadbalancer}}
  Exposes the service externally using a cloud provider's load balancer.
  NodePort and ClusterIP services, to which the external load balancer
  will route, are automatically created.
  \item [ExternalName:]
  \footnote{\url{https://kubernetes.io/docs/concepts/services-networking/service/\#externalname}}
  Maps the service to the contents of the ExternalName field
  (e.g.~foo.bar.example.com), by returning a CNAME record with its
  value. No proxying of any kind is set up. This requires version 1.7 or
  higher of kube-dns.
\end{description}

Since for LoadBalancer we would need a Cloud service like AWS or GCE, we
opted for NodePort. This is not a good solution to use in
a production environment since it opens a couple of port for every pods
that runs in a node, but since we are building a test environment, and this
was the most suitable solution for our cluster.

In the market are available different \texttt{ingresses-controller} an we
analyzed some of them.
\subsection{Nginx}\label{nginx}

Kubernetes supports it's own version of Nginx\footnote{
\url{https://github.com/kubernetes/ingress-nginx}}, but there is also an
ingress-controller based supported by the official Nginx team\footnote{
\url{https://github.com/nginxinc/kubernetes-ingress}}.
We've chosen the second one, since it seems more stable.

\subsection{Installation}\label{installation}

The installation is pretty simple, and can be done with Helm:

\begin{lstlisting}[language=bash]
git clone https://github.com/nginxinc/kubernetes-ingress.git --depth 1 # Clone the project repository
kubernetes-ingress/helm-chart
helm install --name nginxcontroller . --set controller.replicaCount=2,controller.service.type=NodePort,controller.service.externalTrafficPolicy=Cluster # Install the helm package
\end{lstlisting}

At this point, if you want to expose a service you need to create a file
similar to this one:

\begin{lstlisting}[language=yaml]
apiVersion: extensions/v1beta1
kind: Ingress
metadata:
 name: wordpress-ingress
spec:
 rules:
 - host: wordpress.example.com
   http:
     paths:
     - path: /<your-url-path>
       backend:
         serviceName: <your-service-name>
         servicePort: 80
\end{lstlisting}

So, to apply your forwarding to Kubernetes:

\begin{lstlisting}[language=bash]
kubectl create -f ingress-forwarding.yaml
\end{lstlisting}

In the dashboard, you should see now something similar to this:

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.3]{ingress}
\caption{Example of ingress in Kubernetes dashboard}
\end{figure}

\subsection{Træfik}\label{truxe6fik}

\subsection{Installation}\label{installation-1}

Since Kubernetes 1.6 Role based access control was introduces so in
order to make Træfik properly it must be configured. For simplicity,
ClusterRoleBinding can be used as in the Træfik
installation guide\footnote{
\url{https://docs.traefik.io/user-guide/kubernetes/}}:

\begin{lstlisting}[language=yaml]
---
kind: ClusterRole
apiVersion: rbac.authorization.k8s.io/v1beta1
metadata:
  name: traefik-ingress-controller
rules:
  - apiGroups:
      - ""
    resources:
      - services
      - endpoints
      - secrets
    verbs:
      - get
      - list
      - watch
  - apiGroups:
      - extensions
    resources:
      - ingresses
    verbs:
      - get
      - list
      - watch
---
kind: ClusterRoleBinding
apiVersion: rbac.authorization.k8s.io/v1beta1
metadata:
  name: traefik-ingress-controller
roleRef:
  apiGroup: rbac.authorization.k8s.io
  kind: ClusterRole
  name: traefik-ingress-controller
subjects:
- kind: ServiceAccount
  name: traefik-ingress-controller
  namespace: kube-system
\end{lstlisting}

Save that in a file \texttt{traefik-rbac.yaml} and run

\begin{lstlisting}[language=bash]
kubectl apply -f traefik-rbac.yaml
\end{lstlisting}

After the ingress controller can be set up in to 2 way: using a
deployment or a daemon set. We chose the first option for easier
management deploying it with this YAML:

\begin{lstlisting}[language=yaml]
---
apiVersion: v1
kind: ServiceAccount
metadata:
  name: traefik-ingress-controller
  namespace: kube-system
---
kind: Deployment
apiVersion: extensions/v1beta1
metadata:
  name: traefik-ingress-controller
  namespace: kube-system
  labels:
    k8s-app: traefik-ingress-lb
spec:
  replicas: 1
  selector:
    matchLabels:
      k8s-app: traefik-ingress-lb
  template:
    metadata:
      labels:
        k8s-app: traefik-ingress-lb
        name: traefik-ingress-lb
    spec:
      serviceAccountName: traefik-ingress-controller
      terminationGracePeriodSeconds: 60
      containers:
      - image: traefik
        name: traefik-ingress-lb
        ports:
        - name: http
          containerPort: 80
        - name: admin
          containerPort: 8080
        args:
        - --api
        - --kubernetes
        - --logLevel=INFO
---
kind: Service
apiVersion: v1
metadata:
  name: traefik-ingress-service
  namespace: kube-system
spec:
  selector:
    k8s-app: traefik-ingress-lb
  ports:
    - protocol: TCP
      port: 80
      name: web
    - protocol: TCP
      port: 8080
      name: admin
  type: NodePort
\end{lstlisting}

As before save it in a file, say \texttt{traefik-deployment.yaml} and run:

\begin{lstlisting}[language=bash]
kubectl apply -f traefik-deployment.yaml
\end{lstlisting}

This deployment will expose two NodePorts which allow access both to the
ingress and to the web management interface.

\section{Additional Tools}\label{additional-tools}

\subsection{Helm}\label{helm}

To install Helm you need to be in the master node\footnote{
\url{https://docs.helm.sh/using_helm/\#installing-helm}} and run the
command:
\begin{lstlisting}[language=bash]
curl https://raw.githubusercontent.com/kubernetes/helm/master/scripts/get | bash
\end{lstlisting}

Now that we have Helm installed in our master node, let's define first
some additional permission to get it work correctly. creating a file like this:

\begin{lstlisting}[language=yaml]
apiVersion: v1
kind: ServiceAccount
metadata:
  name: helm
  namespace: kube-system
---
apiVersion: rbac.authorization.k8s.io/v1beta1
kind: ClusterRoleBinding
metadata:
  name: helm
roleRef:
  apiGroup: rbac.authorization.k8s.io
  kind: ClusterRole
  name: cluster-admin
subjects:
  - kind: ServiceAccount
    name: helm
    namespace: kube-system
\end{lstlisting}

And after deploy the service:

\begin{lstlisting}[language=bash]
kubectl create -f helm.yaml
\end{lstlisting}

At this point, we have defined a new ClusterRole, \emph{helm}, that has
the possibility to interact with \texttt{kube-system} namespace. Now
we are able to deploy Tiller in our cluster (Helm backend), by running:

\begin{lstlisting}[language=bash]
 helm init --service-account helm --wait
 \end{lstlisting}

### 2018-11-21 res/sections/08-Background.tex
Splitting

--- 2018-11-21 res/sections/08-Background.tex
@@
 going to briefly introduce and explain the technologies involved and their role.
-
-\section{Virtualization}
-Cloud computing is nowadays one of the cutting edge Internet technologies, and
-it is possible thanks to virtualization, that allows to isolate applications,
-optimizing the utilization of hardware resources and providing more flexibility.
-Virtualization, in computing, usually refers to the act of creating a virtual
-version of a certain asset, such as virtual computer hardware platform,
-operating system (OS), storage device, or computer network
-resources~\cite{liu2014research}. Conventional virtualization
-(hypervisor-based) uses virtual machines: it provides strong isolation and a
-complete operating system~\cite{eder2016hypervisor}. In hypervisor
-virtualization every pieces of hardware required to running software is
-emulated, and it is common that the hypervisor provides interfaces to access
-the real hardware, for instance to access physical devices or to communicate
-with the network. However all this functionalities comes to a resource
-overhead~\cite{scheepers2014virtualization}. Container-based virtualization,
-instead, could be a more efficient in certain context, because it does not
-emulate the entire computer. A container is seen as an isolate process running
-on the host, to which the Linux kernel gives a set of capabilities.
-
-\subsection{Network Virtualization}
-Virtualization, in the networking context, refers to the act of combining
-hardware, software and functionalities into a virtual network. This abstraction
-runs on the top of a physical network in a hyper-visor, decoupling the logical
-network from the hardware one. Network virtualization allow multiple,
-heterogeneous networks to cohabit, increasing the flexibility and manageability.
-The idea of network virtualization came from the decoupling of the role of the
-traditional ISP into two different entities: who provides the infrastructure
-(InPs) and who provides the services (SPs). In that context it is possible to
-aggregate resources, even from different InPs, creating a virtual network,
-offering an end-to-end service~\cite{chowdhury2009network}. Most known instances
-of NV are \emph{VPNs}, dedicated networks that allows the communication among
-multiple sites, using private and secure tunnels, relying on the public network
-such as Internet. Other examples of network virtualization are \emph{overlay
-  networks}. They are logical networks, build on the top of an existing physical
-network, typically implemented at the application layer.
-
-\section{Software-defined networking}
-The SDN is an emerging approach to cloud computing and networking technology
-capable of supporting the dynamic nature of future network functions and
-intelligent applications while lowering operating costs through simplified
-hardware, software, and management~\cite{sezer2013we}, that allows to configure
-network in a programmatically way, possibly enhancing monitoring and
-performances. One of the main principles of SDN architecture is to decouple the
-packet forwarding process, \emph{data plane}, from the routing one, the so
-called \emph{control plane}. The architecture also contemplate that network
-intelligence and state are logically centralized~\cite{fundation2012software}.
-
-It has some drawbacks in terms of security, scalability and elasticity, because
-of the centralized architecture.
-
-\subsection{SDN stack}
-\begin{figure}[ht]
- \centering
- \includegraphics[scale=1.3]{sdn_architecture}
- \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
-denote the three layer logic: the Application plane, the Controller plane and 
-the Data plane~\cite{fundation2013software}.}
- \label{chap:background:img:sdn_architecture}
-\end{figure}
-As depicted in Figure~\ref{chap:background:img:sdn_architecture} the SDN
-architecture can be split in three different layers, also known as planes,
-based on their functionalities~\cite{fundation2012software} as explained
-in~\cite{fundation2013software}.
-
-\subsubsection{Application plane}
-
-In the \emph{Application plane} lays \emph{SDN applications} that are programs
-whose role is to set the behavior of the SDN controllers 
-and manage its hardware resources and data through APIs, often called 
-\emph{northbound interfaces} (NBI). In order to do so they can also build an
-abstracted view of the network, using controllers information, in order to make
-the traffic analysis easier and to simplify the decision making process.
-These interfaces provide an additional abstraction layer that high-level 
-applications can consume, enabling the possibility to use a variety of 
-different vendors' controllers in the same way. Applications consists both of
-the two defined layers, one that is responsible to manage the whole logic, and
-another one dedicated to the communication part, that in particular uses the
-previously defined NBIs.
-
-\subsubsection{Controller plane}
-
-\emph{SDN controllers} are logically centralized components that
-receive requests from SDN applications and either translate them 
-in a fashion that can be understood from the SDN Datapath, or collect 
-data as statics and events needed by the application to build an abstract view
-of the network. 
-Due to the variety of tasks a controller has to accomplish, its logical 
-subdivision is achieved with a three-layer architecture, in which every layer 
-exposes its own interface, that can be consequently used by the upper 
-layers, up to the \emph{NBI agents}.
-NBI agent manages the communications with SDN applications, exposing interfaces
-to access network data and to the set of the controller functionalities.
-In particular, the middle layer represents the \emph{SDN Controller Logic},
-which process the requests coming from the upper layer applications. Finally
-requests are communicated to the remaining network components, exploiting the
-Control to Data-Plane Interface driver.
-
-Despite the centralized design of controllers there are no limitations on the
-implementation, so it possible to create federations of controllers or organize
-them into a hierarchy.
-
-\subsubsection{Data plane}
-In the lowest plane we can find the \emph{SDN Datapath}, defined as a logical
-network device, which exposes the possibility to access and control to its
-forwarding and processing capabilities. It is composed by a CDPI agent and a set
-of one or multiple traffic forwarding engines and of a zero or more traffic
-processing functions that are, as a matter of fact, the devices logic. Datapath
-can be mapped either to a single physical device or multiple ones and the
-implementation of the its logic does not preclude neither the virtualization or
-the sharing of physical resources nor the interoperability with non-SDN
-networking.
-
-\subsection{OpenFlow protocol}
-The SDN strategy is often associated with the OpenFlow protocol because it is
-the first standard communications interface defined between the control and the
-datapath in SDN architecture~\cite{fundation2013software}, making possible to
-access from controller the components of infrastructure, such as switches and
-routers. OpenFlow enables the remote administration of networks working on 
-layer 3: packets forwarding tables can be modified relying on packet matching 
-rules. Following this approach it is possible to change routes periodically or 
-ad-hoc, based on packet types, traffic or network load.
-
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.65]{openflow}
- \caption{OpenFlow protocol communication graph}
- \label{chap:background:img:openflow_protocol}
-\end{figure}
-
-In Figure~\ref{chap:background:img:openflow_protocol} it is possible to see the
-how switches works in the OpenFlow protocol. Basically, OpenFlow switches own a
-set of tables to organized packet control and forwarding. These tables can be
-updated by higher-level application, usually an SDN application or controller,
-taking advantage of OpenFlow channels (called OF channels).
-
-\section{Network slicing}
-A network slice can be defined as a logical virtual network, tailored for a
-given use case, on the top a common network using technologies such as
-Software-Defined Networking and Network Function Chaining that allows network
-softwarization~\cite{ordonez2017network}. The notion of separated virtual
-networks deployed on the same infrastructure is not new, e.g. VPN, but it has
-some specifics that make it a novel idea. The concept of network slicing was
-first introduced in 2002 in~\cite{peterson2003blueprint}, where a network 
-overlay that allows services to run continuously and to access the overlay
-resources across different slices was illustrated.
-Services operating over the overlay management have their own slice instead.
-This idea was lately refined in 2016, where the design of network slicing was
-based on a three-layers model~\cite{alliance2016description}, as depicted in 
-Figure~\ref{chap:background:img:network_slicing}.
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.2]{networkslicing}
-  \caption{The three layers model of network
-  slicing~\cite{alliance2016description}}
-  \label{chap:background:img:network_slicing}
-\end{figure}
-
-The first layer is the \emph{Service Instance Layer} used to represent the
-business or user services supported, the second one is the \emph{Network Slice
-Instance Layer}. This is a set of network functions that allows to
-satisfy certain requirements~\cite{kotulski2017end}. Network functions, in
-general, can be both physical, thus a combination of vendor-specific hardware
-and/or virtualized. In that case the function is implemented as software and it
-is decoupled from the hardware on which it is running on. Each instance must be
-isolated, in order to ensure that function running at the top of it will always
-provide the same Quality of Service, without degrading because of a change on
-another slice~\cite{richart2016resource}. It also allows to provide better
-security. In fact, in case of attack or fault of a certain slice, others must
-not be affected. Moreover, each slice must implement its own set of security
-function to preventing unauthorized access. Furthermore, isolation let to
-manage individually each slice~\cite{ordonez2017network}. Network slice
-instances are defined by a \emph{Network Slice Blueprint}: a complete
-description of the structure and data flows that allow the creation and the
-control of the Network Slice Instance during its whole lifecycle. The blueprint
-also refers to the resources needed to assemble the network slice and defines
-its characteristics (e.g. ultra-low latency, ultra-reliability, value-added
-services for enterprises, etc.). The last layer is the \emph{Resource Layer},
-containing both physical and logical resources, that involve all physical
-assets and resources for computation, storage or transport, including radio
-access.
-
-\subsection{Network Function Virtualization}
-The \emph{Network Function Virtualization} (NFV) concept refers to the use of
-virtualization technologies to virtualize network nodes functions. This emerging
-technology focuses on the decoupling of function logic and hardware, allowing
-VNFs to run on the top of servers, switches, storage devices or on cloud
-computing infrastructures. In other words the software logic is separated from
-the custom platform on which it runs, enabling the use of commodity
-hardware~\cite{gray2016network}. That approach worked well in the past, because
-of the need to follow rigorous standards, providing at the same time high
-quality, but it led to long product cycles and high specialized services that
-use a combination of software and hardware resources, typically vendor-specific,
-and only largely non-interoperable within other vendor solutions. New Internet
-scenarios, that include IoT and the increasing number of services to spread
-multimedia data (Skype, Netflix, Spotify, \dots), requires more flexible
-functions and faster network deployments. This detachment provides greater
-flexibility to scale VNF based on traffic and performances and also improve the
-automation of function deploy and update: removing dedicated hardware also will
-reduce costs machines and for maintenance.
-
-The NFV technology is exploited in a framework, organized
-at least by three components~\cite{etsi2013gs}, as represented in
-Figure~\ref{chap:background:img:etsi_arch}:
-\begin{enumerate}
-  \item a software, able to run on over the NFVI, that performs the network
-  function that must be deployed, that is the VNF;
-  \item NFV Infrastructure (NFVI) that is composed by all physical resources
-  and how these can be virtualized, that forms the environment where the VNFs
-  are deployed;
-  \item NFV MANagement and Orchestration (NFV-MANO) that covers the
-  orchestration and management of both physical and logical resources, the
-  functional blocks, data repositories used by these blocks, reference points
-  and interfaces through which these functional blocks exchange
-  information for the purpose of managing and orchestrating NFVI and VNFs.
-\end{enumerate}
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=2]{etsi_arch}
-  \caption{Network Function Virtualization reference
-  architecture~\cite{etsi2013gs}}
-  \label{chap:background:img:etsi_arch}
-\end{figure}
-
-\subsection{Network Function Virtualized Infrastructure}
-As mentioned before, NFVI is the combination of physical and logical resources,
-that combined together create the environment on which the VNFs are deployed
-and run. Physical supplies can be proprietary systems, commercial-off-the-shelf
-hardware, software and network resources or cloud computing capabilities. The
-abstraction of resources is made taking advantage of a virtualization layer,
-that hides the underlying infrastructure. Storage and computing resources may
-be represented in terms of virtual machines or
-containers~\cite{mijumbi2016network}.
-
-\subsection{NFV MANagement and Orchestration}
-The MANO, represented in Figure~\ref{chap:background:img:etsimano}, can be
-splitted into three logical functional components:
-\begin{itemize}
-  \item the virtual infrastructure manager (VIM);
-  \item the VNF Manager (VNFM);
-  \item the NFV Orchestrator (NFVO).
-\end{itemize}
-And into four data repositories:
-\begin{itemize}
-  \item Network Service (NS) catalog;
-  \item VNF catalog;
-  \item NFV instances;
-  \item NFVI resources.
-\end{itemize}
-
-\begin{figure}
-  \centering
-  \includegraphics[scale=0.3]{etsimano}
-  \caption{NFV MANO components}
-  \label{chap:background:img:etsimano}
-\end{figure}
-
-\paragraph*{VIM}
-The VIM aim is to manage and control NFVI resources. Every MANO can contain more
-VIM instances, each of them control NFVI resources from a single infrastructure.
-VIM can be developed either to administer one type of NFVI or more. VIM have to
-monitor the allocation of virtual resource in order to orchestrate allocation,
-upgrade, release, and reclamation of resources and optimize their usage,
-including virtual links, networks, subnets, and ports. It is also the component
-in charge to manage security group policies to ensure access control. In
-addition, VIM also has to record data on performance and faults, that can be
-queried. Thank to this component it is possible to dynamically allocate
-resource, based on demand, that is particularly using IaaS.
-
-\paragraph*{VNFM} 
-In this framework all VNFs, once created, have to register to a VNFM. This
-component is accountable to manage their lifecycle, that consists of
-instantiation, scaling, termination and, possibly, update/upgrade, and
-communicate the state of a certain VNF to others MANO blocks: this enable the
-monitoring of key performance indicators (KPI), that can help to make decision
-on VNF scaling. VNFM can manage a single type of VNF instances or multiple.
-
-\paragraph*{NFVO}
-The NFVO is responsible of combining functions to create end-to-end services.
-Its functionalities are to ensure there are adequate compute, storage, and
-network resources available to provide a network service independently of the
-VIM and the creation of end-to-end services by composing different VNFs, that
-can be deployed on different point-of-presence and registered to different VNF.
-It also managing the topology of these services.
-
-\paragraph*{Data repositories} Data repositories are use to keep information on
-the NFV MANO. The NS catalog is a set of predefined templates, which define
-services, functions involved and their connectivity. The VNF catalog is,
-instead, a set of templates describing the deployment and characteristics of
-VNFs. The NFVI resources repository holds information about available/allocated
-NFVI resources. The NFV instances repository keep information about functions
-and services instances during their lifetimes.
-
-\subsection{VNF challenges}
-This new concept open a plethora of challenges~\cite{han2015network}. Regarding
-performances, functions running on general purposes servers cloud affect
-throughput and latency, moreover the capacity of the virtualized function can be
-less than the hardware and software version. To reach comparable levels in terms
-of performances efficient algorithms are needed, as a excellent resource
-management. Another problems is the VNF management. In case of hardware
-functions resources are almost equivalent, at the contrary, virtualized ones
-could have costs and values that differs significantly depending on the point of
-presence on which are deployed. On the other hand, since VNFs are software
-components, they can elastically scale up and down, depending on the traffic
-loads and user demand. Another management challenge refers to keeping track
-where the VNF is running, that is of paramount importance in case of issues.
-Related to possible dysfunctions, software function must reach a certain level
-of reliability. Generally, purpose-build network equipment can ensure the
-five-nines reliability requirement, that is more challenging to follow in that
-scenario due to the higher number of possible points of failure, such as during
-the scaling operation or VNF migrating from a point of presence to another.
-Finally even security must be taken into account. In fact, infrastructure on
-which VNFs run can be outsourced to third parties. Furthermore, orchestration
-and management components must implement some sort of intrusion prevention and
-detection system and the sharing of resources can introduce security threads.
-
-\section{Service Function Chaining}
-The Service Function Chaining (SFC) concept was defined
-in~\cite{halpern2015service} and refers to an ordered set of Service Functions
-(SF) and the constraint, defined after a classification process, to packets and
-flows. SF are the particular functions whose aim is the treatment of packets
-received. They can work on different levels of the network layer. These
-components can be implemented as virtual elements only or in a combination of a
-physical and software network elements. There are no limitations on the number
-of SFs running on the same network element, so in general it is possible to have
-one or more functions running on the same element. SFs can be aware or not that
-are part of a certain chain, in the former case they are called \emph{SFC aware}
-functions, in the latter one \emph{SFC unaware}.
-
-The lifecycle management of the chain is enabled by two emerged technologies,
-software defined networking and network defined
-virtualization~\cite{medhat2017service}.
-
-\subsection{Architecture principles}
-The SFC technology was though to pursue some concept. One of the most important
-principles is the \emph{decoupling}. Decoupling in terms on independence on the
-infrastructure and the network, so no change must be performed to the forwarding
-underlying topology to deploy SFCs, and in terms of separation between packet
-handling (by SFs) and forwarding. This decoupling refers even to architecture
-and functions: one must be independent from the other. Another main point is the
-possibility to dynamically set routes on the incoming traffic to decide the best
-treatment based on its typology. The classification can occur on different point
-of the chain and can be updated during the packet delivery. Finally creation,
-modification and deletion of a certain SFC must not impact on the others.
-
-\subsection{Graph view of an SFC}
-An SFC can be seen as a graph in which each node is an SF and edges are the path
-through which a packet have to pass through. Because of this definitions there
-are no limitations on possible cycles that can occur in the chain, in case that
-a certain SF has to be reapplied on the same packet, frame or flow. These
-scenarios must be adequately managed to avoid endless loops. This enable even
-the possibility of branching on a certain node in order to define the treatment
-that has to be imposed from that point of the chain. Traffic path in the SFC
-graph is managed and described by the \emph{Service Function Forwarder} (SFF)
-and \emph{Service Function Path} (SFP) components. The former is the element of
-the architecture responsible of traffic forwarding. The main goals are to
-redirect traffic according to the SFC specified for a specific packet flow,
-query the classifier if needed and manage traffic coming back from a SF. The
-SFP, instead, is an indirection layer between the fully abstract notion of SFC
-and the list of SF that the traffic has to traverse. It is used for dictate the
-constraint specification on the path of where packets assigned to a certain
-service function path must go. There are no limitations on how specific it must
-be: it can specify the exact locations, identifying both the SFF and the SF, or
-only give some recommendation or, for example, defining only the SFF and
-demanding to it the decision on the SF instance.
-
-The graph representation also permit to think to \emph{unidirectional} and
-\emph{bidirectional} chains. An SFC is defined unidirectional if traffic can
-traverse it only in one direction, so if it is composed by 3 SF, named
-\texttt{SF0}, \texttt{SF1} and \texttt{SF2} in this order, traffic can only go
-from \texttt{SF0} to \texttt{SF2}. On the contrary, a bidirectional SFC allows
-traffic in both directions. So, having the same chain as in the previous
-example, traffic can pass both from \texttt{SF0} to \texttt{SF2} and from
-\texttt{SF2} to \texttt{SF0}.
-
-\subsection{SFC Encapsulation}
-In order to exchange packet and information in the SFC ecosystem, traffic is
-encapsulated. The so called SFC encapsulation carries data used by the
-framework to identify SFPs, but it differs from a transport encapsulation and it
-is independent from the transport one: as such any transport protocol can be
-exploited to forward traffic.
-
-This encapsulation is neither supported from SFC unaware SF nor from legacy 
-SF\footnote{Legacy SF refers to network function created before SFC framework}
-so \emph{SFC Proxies} has to be included in this ecosystem. These components
-sits the SFF and the SFs. They accept packets from the SFF and remove the
-encapsulation before delivering the packet to the SFs. In a mirrored way, it
-receive the result of the SF and reapplies the SFC header, in order to be
-managed by the SFF and the overall system. The communication among SFFs and
-proxies is the same as the SFFs and SFC aware SF.
-
-The \emph{Network Service Header} (NSH), used by SFC environment for
-encapsulation is described in~\cite{rfc8300} and it consider three main fields
-as represented in Figure~\ref{chap:background:img:nshformat}: Base Header,
-Service Path Header and Context Header.
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.6]{nshformat}
-  \caption{Network Service Header~\cite{rfc8300}}
-  \label{chap:background:img:nshformat}
-\end{figure}
-
-\subsubsection{Base header}
-The Base Header utility is to provide information about the service header and
-the payload protocol.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.6]{nshbase}
-  \caption{Network Service Header~\cite{rfc8300}}
-  \label{chap:background:img:nshformat}
-\end{figure}
-
-Fields are:
-\begin{description}
-  \item[Version:] This field is used to set the version of the header, used for
-  ensuring backward compatibility through different NSH definitions;
-  \item[0 bit:] Bit set only in case of orchestration or management packets, and
-  must not be changed by the SFC;
-  \item[TTL:] Time To Live, indicate the maximum number of hops among SFF and
-  its initial value should be configurable by the control plane. At each hop it
-  must be decremented to 1 until 0: in the latter case the packet must be
-  discarded. The default value is 63. This field is useful in loop prevention
-  mechanism. The default value is 63;
-  \item[Length:] the total length of the header (in 4-byte words), including
-  the \texttt{Fixed-Length Context Header} in NSH type 0x1 or the
-  \texttt{Variable-Length Context Header} in NSH type 0x2. In case of variable
-  length header data must be padded to a multiple of 4;
-  \item[Unassigned bits:] unassigned bits for future usages;
-  \item[MD:] Metadata Type, indicate the type of the NSH. Values that now
-  accepted are \texttt{0x0} that is a reserved value and if set the packet must
-  be discarded, \texttt{0x1} indicate that context header length is
-  fixed~\ref{subsubsec:background:flch}, \texttt{0x2} indicate that context
-  header length is variable~\ref{subsubsec:background:vlch} and \texttt{0xF} for
-  testing purposes (if not part of the experiment the packet must be discarded).
-  Base header length and Service Path Header have always the same size;
-  \item[Next Protocol:] indicates the protocol type of encapsulated data
-  (protocols codes can be found at~\cite{rfc8300}).
-\end{description}
-
-\subsubsection{Service Path Header}
-Provides path identification and location within a service path.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.6]{nshservicepath}
-  \caption{NSH Service Path Header~\cite{rfc8300}}
-  \label{chap:background:img:nshpath}
-\end{figure}
-
-\begin{description}
-  \item[Service Path Identifier (SPI):] 24 bits, identify unambiguously an SFP.
-  The first classifier have to set it;
-  \item[Service Index (SI):] 8 bits, provide location within the SFP. Initially
-  must be set to 255 and it must be decremented on each hop.
-\end{description}
-
-SPI combined to SI determine the next SF on each link in the chain. SI change
-provide a minimal loop prevention mechanism.
-
-\subsubsection{NSH with Fixed-Length Context Header}
-\label{subsubsec:background:flch}
-The NSH of type \texttt{0x1} have a 16-Byte Fixed-Length Context Header that
-can be used to share data along the service path.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.6]{nshtype1}
-  \caption{NSH MD Type \texttt{0x1}~\cite{rfc8300}}
-  \label{chap:background:img:nshtype1}
-\end{figure}
-
-The specification does not specify the content of this field and if SF or SFC
-Proxy does not known its semantics they have to discard the entire packet.
-
-\subsubsection{NSH with Variable-Length Context Header}
-\label{subsubsec:background:vlch}
-In NSH of type \texttt{0x2}, zero or more Variable-Length Context Headers may
-be added, but lengths must always be a multiple of 4-bytes (otherwise they must
-be padded).
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.6]{nshtype2}
-  \caption{NSH MD Type \texttt{0x2}~\cite{rfc8300}}
-  \label{chap:background:img:nshtype2}
-\end{figure}
-
-The structure of the variable length part is depicted in
-Figure~\ref{chap:background:img:nshtype2variable}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[scale=0.6]{nshtype2variable}
-  \caption{Variable-Length Context Header~\cite{rfc8300}}
-  \label{chap:background:img:nshtype2variable}
-\end{figure}
-
-\begin{description}
-  \item[Metadata Class (MD Class):] field used in order to provide hierarchical
-  namespaces for standards, vendors or other scopes;
-  \item[Type:] indicates the type of metadata, depending on the MD Class;
-\end{description}
-
-\subsection{Classification of traffic}
-In order to define the path of which traffic have to be forwarded, it must be
-classified. The classification granularity is not specified and can occur even
-during the SFC traverse. At the ingress of SFC environment packets have to be
-classified to determine to choose an SFC and eventually, if the classification
-granularity is too coarse, can be repeated after the treatment of some SFs. The
-reclassification, also called branching, may result in a different SFP choice or
-only to a metadata change.
-
-\subsection{Other considerations}
-In this architecture the SFs are viewed are resource and managed by the control
-plane. Capabilities, availability and location are all accountable to the
-control part as their creation. It is also in charge of collecting information
-needed by the classifier to make proper decisions. The control plane could be
-either distributed or centralized and even a combination is possible.
-
-SFC system add to the original packets information as a NSH, that both contains
-metadata and information on the path that it must traverse. It is possible that
-the size of the packet, in this manner, exceed the MTU on some link in the
-chain. Solution must take in account this problem and the cost to overcome it.
 

### 2018-11-21 res/sections/81-k8s.tex
English fix

@@
-\chapter{Kubernetes}\label{kubernetes}
+\chapter{Kubernetes}
+\label{kubernetes}
 
-In this appendix we provide a basic guide to Kubernetes concept and
-installation. This is not thought as an complete overview of functionalities but
-there are only reported a description of concept an feature used in the
-implementation of our solution.
+In this appendix we provide a basic summary to the Kubernetes concepts and
+we will show a simple way to perform a basic installation. This is not thought 
+as an complete overview of functionalities but there are only reported a 
+description of concepts an features used in the implementation of our solution.
 
-\section{Kubernetes basic concepts}\label{kubernetes-basic-concepts}
+\section{Kubernetes basic concepts}
+\label{kubernetes-basic-concepts}
 
-\subsection{Pod}\label{pod}
+\subsection{Pod}
+\label{pod}
 
-A pod is the most basic building block for Kubernetes. A pod usually
+A Pod is the most basic building block for Kubernetes. A Pod usually
 contains one or more container (if the containers are highly coupled),
-even if Kubernetes it's able to manage different virtualization technologies
-(e.g.
-Virtual Machines\footnote{\url{https://github.com/kubevirt/kubevirt}}).
+even if Kubernetes is able to manage different virtualization technologies
+(e.g. Virtual Machines\footnote{\url{https://github.com/kubevirt/kubevirt}}).
 Pods have a life cycle, where it can assume these
@@
 \begin{itemize}
-\item
-  \textbf{Pending}: The Pod has been accepted by the Kubernetes system,
-  but one or more of the Container images has not been created. This
+\item \textbf{Pending}: The Pod has been accepted by the Kubernetes system,
+  but one or more of the Container images have not been created yet. This
   includes time before being scheduled as well as time spent downloading
   images over the network, which could take a while.
-\item
-  \textbf{Running}: The Pod has been bound to a node, and all of the
+\item \textbf{Running}: The Pod has been bound to a node, and all of the
   Containers have been created. At least one Container is still running,
   or is in the process of starting or restarting.
-\item
-  \textbf{Succeeded}: All Containers in the Pod have terminated in
-  success, and will not be restarted.
-\item
-  \textbf{Failed}: All Containers in the Pod have terminated, and at
+\item \textbf{Succeeded}: All Containers in the Pod have terminated with
+  success, and they will not be restarted.
+\item \textbf{Failed}: All Containers in the Pod have terminated, and at
   least one Container has terminated in failure. That is, the Container
   either exited with non-zero status or was terminated by the system.
-\item
-  \textbf{Unknown}: For some reason the state of the Pod could not be
+\item \textbf{Unknown}: For some reason the state of the Pod could not be
   obtained, typically due to an error in communicating with the host of
@@
 
-A pod can have different conditions too:
+A Pod can have different conditions too:
 % TODO davide must delete from test
@@
 
-Each pod can have different QoS classes:
+Each Pod can have different QoS classes:
 \begin{description}
@@
 
-\subsection{Service}\label{service}
+\subsection{Service}
+\label{service}
 
-Since pods for their nature are highly volatile, applications that use
+Since Pods for their nature are highly volatile, applications that use
 them can not rely on their IP address. To solve this problem, the
 Service implementation in Kubernetes assure a stable IP, even if
-different pods in this service crash and gets restarted by a ReplicaSet
+different Pods in this service crash and gets restarted by a ReplicaSet
 (witch will not be explain in this guide, but it basically scale up and
-down pods\footnote{
+down Pods\footnote{
 \url{https://kubernetes.io/docs/concepts/workloads/controllers/replicaset/}})
@@
 
-\subsection{Volume}\label{volume}
+\subsection{Volume}
+\label{volume}
 
@@
 Kubernetes offers a way to save all data and to possibly make it
-available to all the pods replicas, offering a transparent layer where
-pods are not aware of working with information shared other pods, even located
+available to all the Pods replicas, offering a transparent layer where
+Pods are not aware of working with information shared other Pods, even located
 in other nodes. This functionality is called Volume, and there are many
@@
 
-\subsection{Namespace}\label{namespace}
+\subsection{Namespace}
+\label{namespace}
 
@@
 at least the kubelet daemon (to manage Kubernetes functionalities) and a
-container manager (e.g. Docker Engine, VirtualBox) running. A node runs pods
+container manager (e.g. Docker Engine, VirtualBox) running. A node runs Pods
 and can communicate with other nodes in the same cluster
@@
 
-\subsection{Master}\label{master}
+\subsection{Master}
+\label{master}
 
 The \emph{master node} is the one that starts a cluster, and that manage the
-slaves. Only some pods can run n the master node (e.g. the dashboard).
+slaves. Only some Pods can run in the master node (e.g. the dashboard).
 
@@
 
-\subsection{Slave}\label{slave}
+\subsection{Slave}
+\label{slave}
 
-A \emph{slave} is a node that run the pods that the master schedule. It can
+A \emph{slave} is a node that run the Pods that the master schedule. It can
 also be part of a \emph{PersistantVolume} storage (like GlusterFS) and if it
-goes down, another slave will take care of the pods that where running on
+goes down, another slave will take care of the Pods that where running on
 it. Slaves can join existing cluster via a given token.
 
-\subsection{Networking}\label{networking}
+\subsection{Networking}
+\label{networking}
 
@@
 
-Every nodes in the cluster runs a pod called \texttt{kubeproxy}, that
-communicates with the \texttt{kube-api} pod (running on master) and
+Every node in the cluster runs a Pod called \texttt{kubeproxy}, that
+communicates with the \texttt{kube-api} Pod (running on master) and
 the \texttt{kube-dns} (for Kubernetes \textless{}=1.9.x) or
 \texttt{CoreDNS} (for Kubernetes \textgreater{}=1.10.x). \texttt{kubeproxy}
-assure transparent communication between pods in the same cluster, and
-\texttt{kube-api} allows to call Kubernetes services. kube-dns/CoreDNS manage
-DNS request coming from the pods, differentiating the one that goes outside
-the cluster from the other that point internally\footnote{Additional
-information about Flannel, Kubernetes and Docker networking
-can at:\\
+assure transparent communications between Pods in the same cluster, and
+\texttt{kube-api} allows to call Kubernetes services. \texttt{kube-dns/CoreDNS} 
+manage DNS requests coming from the Pods, differentiating the one that goes 
+outside the cluster from the other that points internally\footnote{Additional
+information about Flannel, Kubernetes and Docker networking can at:\\
 \sloppy
@@
 \includegraphics[scale=0.3]{flannelredirect}
-\caption{And example of how flannel redirect packages between pods}
+\caption{And example of how flannel redirect packages between Pods}
 \end{figure}
@@
 Managing compute and storage are totally different problems and in order
-to abstract the last one Kubernetes relies on \textbf{PersistantVolume}
-and \textbf{PersistantVolumeClaim} resources.
+to abstract the last one Kubernetes relies on the \textbf{PersistantVolume}
+and the \textbf{PersistantVolumeClaim} resources.
 
-A PersistantVolume (\textbf{PV}) is something similar to a node, in fact
+A PersistantVolume (PV) is something similar to a node, in fact
 is a piece of storage provided by the administrator: it captures the
 details of the implementation of the storage and has a lifecycle that is
-independent of any individual pod. To not expose developers to storage
+independent of any individual Pod. To not expose developers to storage
 boring details admins can expose different PV types through
@@
 
-A PersistantVolumeClaim (\textbf{PVC}) instead is a request for storage
-and it is similar a pod: instead of requesting memory and CPU, a PVC
-requests certain size and access mode to storage.
+A PersistantVolumeClaim (PVC) instead is a request for storage and it is 
+similar a Pod: instead of requesting memory and CPU, a PVC requests certain 
+size and access mode to storage.
 
@@
 of resource utilization dictated by the policy defined during the
-creation of resources. Kubernetes pod can use claims as volumes.
+creation of resources. Kubernetes Pod can use claims as volumes.
 
@@
 that can be used. Different classes can be used to describe different
-levels of service or policies. PVCs that don't specify a particular class
+levels of service or policies. PVCs that do not specify a particular class
 can use the default one: to define it you can use annotation
@@
   dynamic provisioning;
-\item
-  \textbf{Binding}: a user requests a certain amount of storage with a
+\item \textbf{Binding}: a user requests a certain amount of storage with a
   claim and a control loop in the master look for a PV that can bind the
@@
   volume does not exists.
-\item
-  \textbf{Using}: after the binding phase the pod start using the volume
-  (PVC is described in the pod yaml
+\item \textbf{Using}: after the binding phase the Pod starts using the volume
+  (PVC is described in the Pod yaml
   \footnote{\url{https://kubernetes.io/docs/concepts/storage/persistent-volumes/\#claims-as-volumes}}
-\item
-  \textbf{Releasing}: when an application is done with using the volume,
+\item \textbf{Releasing}: when an application is done with using the volume,
   developers can delete the PVC objects through the API, releasing the
   claim. This step will initiate the reclamation process.
-\item
-  \textbf{Reclaiming}: this phase is defined as a policy by the
+\item \textbf{Reclaiming}: this phase is defined as a policy by the
   administrators. The reclaim policy for a PersistantVolume tells the
@@
 We performed our Kubernetes installation in a Openstack environment.
-Since we used the Openstack of our school department, the administrator
-have given us a virtual machine, Openstack-CLI, to perform
-operations. From now, our scripts assume you're in a machine with access
-to Openstack, and that you have sourced your openrc.sh script before
-continuing.
+Since we used the Openstack of our school department, the administrators have 
+given us a virtual machine, Openstack-CLI, to perform operations. From now, our 
+scripts assume you're in a machine with access to Openstack, and that you have 
+sourced your openrc.sh script before continuing.
 
@@
 
-After there is an explanation for the most important part of the
-scripts. The script \verb!on_openstack_cli.sh!
+After there is an explanation for the most important part of the scripts. The 
+script \verb!on_openstack_cli.sh!
 \footnote{\url{https://github.com/Augugrumi/init-script/raw/centos/kubernetes/on_openstack_cli.sh}}
-main aim is to run other file retrieving information using even
-\texttt{nova} commands.
+main aim is to run other file retrieving information using even \texttt{nova} 
+commands.
 
-First all the software needed must to be installed an for this the
-script \\\verb!install_kubectl_repo.sh!
+First of all the software needed must be installed and for this the script 
+\\\verb!install_kubectl_repo.sh!
 \footnote{\url{https://raw.githubusercontent.com/Augugrumi/init-script/centos/kubernetes/install_kubectl_repo.sh}}
-was created. This only call another script and wait until all machine
+was created. This only calls another script and waits until all the machines
 are ready. The other script is \verb!bootstrap.sh!
 \footnote{\url{https://raw.githubusercontent.com/Augugrumi/vagrantfiles/oldversion/kubernetes/centos/bootstrap.sh}}
-whose aim is to configure the machine and the install the necessary
-software.\\
-Boostrap script first \textbf{disable SELinux} on all machines (needed
-for Kubernetes networking)
-
+whose aim is to configure the machine and to install all the necessary
+software. The boostrap script first \textbf{disables SELinux} on all machines 
+(needed for Kubernetes networking):
 \begin{lstlisting}
@@
 
-then \textbf{enable br\_netfilter} module (needed for Kubernetes
-networking)
-
+and then it enables the \verb!br_netfilter! module (needed for Kubernetes
+networking):
 \begin{lstlisting}
@@
 
-\textbf{Switch off the swap} by the commands
-
+The swap is then switched off with the following commands:
 \begin{lstlisting}
@@
 
-And finally add \texttt{rbd}, \texttt{ip\_vs}, \texttt{ip\_vs\_rr},
+And it finally adds \texttt{rbd}, \texttt{ip\_vs}, \texttt{ip\_vs\_rr},
 \texttt{ip\_vs\_wrr}, \texttt{ip\_vs\_sh}, \texttt{dm\_snapshot},
-\texttt{dm\_mirror} and \texttt{dm\_thin\_pool} to modules that will be
-loaded to machine startup.
-
+\texttt{dm\_mirror} and \texttt{dm\_thin\_pool} to the modules that will be
+loaded to the machine startup.
 \begin{lstlisting}
-sudo sh -c 'echo -e "rbd\nip_vs\nip_vs_rr\nip_vs_wrr\nip_vs_sh" >  /etc/modules-load.d/ip_vs.conf'
-sudo sh -c 'echo -e "dm_snapshot\ndm_mirror\ndm_thin_pool" >  /etc/modules-load.d/gluster.conf'
+sudo sh -c 'echo -e "rbd ip_vs ip_vs_rr ip_vs_wrr ip_vs_sh" >  
+/etc/modules-load.d/ip_vs.conf'
+sudo sh -c 'echo -e "dm_snapshot dm_mirror dm_thin_pool" >  
+/etc/modules-load.d/gluster.conf'
 \end{lstlisting}
 
-Note: Last 3 modules will be useful in the case that you want to install
-GlusterFS on your machines.
+Note: the last 3 modules will be useful in case GlusterFs has to be installed 
+in the system.
 
-In the end the script install \texttt{docker},
+In the end the script installs \texttt{docker}, 
 \texttt{device-mapper-persistent-data}, \texttt{lvm2}, \texttt{kubelet},
-\texttt{kubectl} and \texttt{kubeadm} and start (and enable)
+\texttt{kubectl}, \texttt{kubeadm} and it starts (and enables)
 \texttt{docker} and \texttt{kubelet} services.
 
-After rebooting all the machine, the installation go on adding in the
-\texttt{/etc/hosts} file of all node that will be part of the Kubernetes
-cluster the IP and the name of all the nodes.
+After rebooting all the machines, the installation goes on adding in the
+\texttt{/etc/hosts} file of all nodes IP that will be part of the Kubernetes
+cluster and the names of all the nodes.
 
-After the installation continues in the \textbf{master} node, with the script
-\verb!on_master.sh!
+After that the installation continues in the \textbf{master} node, with the 
+script \verb!on_master.sh!
 \footnote{\url{https://raw.githubusercontent.com/Augugrumi/init-script/centos/kubernetes/on_master.sh}}.
-We run \texttt{kubeadm} by the command
-
+The following command initiates the Kubernetes cluster:
 \begin{lstlisting}
-sudo kubeadm init --apiserver-advertise-address=<ip-of-the-master> --pod-network-cidr=10.244.0.0/16 --ignore-preflight-errors cri | grep "kubeadm join" > /home/centos/joincommand
+sudo kubeadm init --apiserver-advertise-address=<ip-of-the-master> 
+--Pod-network-cidr=10.244.0.0/16 --ignore-preflight-errors cri | grep "kubeadm 
+join" > /home/centos/joincommand
 \end{lstlisting}
 
-Saving part of the output that will be necessary for other nodes to join
-the cluster.\\
-Note 1:\texttt{-\/-pod-network-cidr} depends on which agent you want to
+\noindent Saving part of the output that will be necessary for other nodes to 
+join the cluster. There are some considerations to keep in mind though:
+\begin{enumerate}
+ \item \texttt{-\/-pod-network-cidr} depends on which agent you want to
 use to configure Kubernetes network. We choose Flannel so we put
-\texttt{10.244.0.0/16}.\\
-Note 2: we don't know at the moment why some preflight errors was
+\texttt{10.244.0.0/16};
+ \item we don't know at the moment why some preflight errors was
 launched by the command and in order to do so (since everything seems
@@
 slave node to join the cluster.
+\end{enumerate}
 
-And after give the following commands to create the folder needed by
-Kubernetes
-
+And after its necessary to give the following commands to create the folder 
+needed by Kubernetes to properly work:
 \begin{lstlisting}
@@
 
-Then wait for all pods that are ready
+Then is necessary to wait for all Pods to be ready
 (\texttt{kubectl\ get\ pods\ -\/-all-namespaces}).
 
-Following, install Flannel\footnote{\url{https://github.com/coreos/flannel}}
-for the networking and wait that it is up:
-
+The next step is to install 
+Flannel\footnote{\url{https://github.com/coreos/flannel}}
+in order to get the networking working. Some time is necessay until the 
+networking is up:
 \begin{lstlisting}
@@
 And then it is possible to setup the dashboard to access Kubernetes resources:
-
 \begin{lstlisting}
@@
 
-With this command we make the dashboard accessible on port 8001 of the
+Finally, with this command we make the dashboard accessible on port 8001 of the 
 master node at this URL
@@
 
-The previous command also the token to access the dashboard on
-\texttt{secret.txt} file on the home folder of the master node.
+The previous command also necessitate of the token to access the dashboard, 
+that is available in the file \texttt{secret.txt}, present on the home of the 
+master node.
 
-The last thing to do is to make other node join the cluster, using the
+The last thing to do is to make other nodes to join the cluster, using the
 \texttt{joincommand} file that we created before.
 
-\section{Persistent Storage}\label{persistent-storage}
+\section{Persistent Storage}
+\label{persistent-storage}
 
-\subsection{Prerequisites}\label{prerequisites}
+\subsection{Prerequisites}
+\label{prerequisites}
 
@@
 
-\subsection{GlusterFS}\label{glusterfs}
+\subsection{GlusterFS}
+\label{glusterfs}
 
@@
 To enable persistent storage in Kubernetes with Glusterfs we need at
-least three nodes. As we've already written above, we deployed our
-Kubernetes cluster using CentOS 7.\\
-Note: this procedure was tested on \texttt{Kubernetes\ v1.11.1.}
+least three nodes. As we have already written above, we deployed our
+Kubernetes cluster using CentOS 7\footnote{This procedure was tested on 
+\texttt{Kubernetes\ v1.11.1.}}.
 
-Since we're running our cluster on a Openstack environment, we attached
+Since we are running our cluster on a Openstack environment, we attached
 one volume per VM. If you're in our shoes, just create some empty
-volumes and attach them to nodes you want to configure with glusterfs.
+volumes and attach them to the nodes you want to configure with Glusterfs.
 
@@
 
-For all the nodes you installed glusterfs you need to apply a particular
+For all the nodes where you installed Glusterfs you need to apply a particular
 label in you Kubernetes cluster. In a terminal, type:
-
 \begin{lstlisting}
@@
 you should also check that the following ports are reachable:
-
 \begin{itemize}
@@
 
-On top of that, you'll need to grant access to your root account while
-installing all the stuff we need. In CentOS, if like us you're using the
-CloudImg image, you need to access every machine and type:
-
+On top of that, you will need to grant access to your root account while
+installing all the stuff we need. In CentOS, if like us you are using the
+\texttt{CloudImg} image, you need to access every machine and type:
 \begin{lstlisting}
@@
 (e.g. heketi), \texttt{gk-deploy}.
-
 \begin{lstlisting}
@@
 file must be created, from the sample \texttt{topology.json.sample}, and edit to
-fit your topology. As in the following example:
-
+fit your topology, like in the following example:
 \begin{lstlisting}
@@
 Now it is possible to deploy a Gluster cluster. From the master the following
-command must be run:
+commands must be run:
 
@@
 
-At the end, if everything goes correctly, a \texttt{StorageClass} template have
-to be created. To your Kubernetes through \texttt{kubectl}:
-
+At the end, if everything goes correctly, a \texttt{StorageClass} template needs
+to be created. To your Kubernetes through \texttt{kubectl} type:
 \begin{lstlisting}[language=bash]
@@
 command:
-
 \begin{lstlisting}[language=bash]
@@
 
-\section{Ingress}\label{ingress}
+\section{Ingress}
+\label{ingress}
 
-Ingress makes service available outside the cluster. There are different ways
-Kubernetes handle service reachability, and these are:
-
+Ingress allows services to be available from outside the cluster. There are 
+different ways Kubernetes can handle service reachability, and these are:
 \begin{description}
@@
 
-Since for LoadBalancer we would need a Cloud service like AWS or GCE, we
-opted for NodePort. This is not a good solution to use in
-a production environment since it opens a couple of port for every pods
-that runs in a node, but since we are building a test environment, and this
+Since for the \texttt{LoadBalancer} typology we would need a Cloud service like 
+AWS or GCE, we opted for the \texttt{NodePort} one. This is not a good solution 
+to use in a production environment since it opens a couple of ports for every 
+Pods that runs in a node, but since we are building a test environment, and this
 was the most suitable solution for our cluster.
 
-In the market are available different \texttt{ingresses-controller} an we
-analyzed some of them.
-\subsection{Nginx}\label{nginx}
+The market is available with plenty of \texttt{ingresses-controller} and 
+therefore we decided to analyze some of them.
 
-Kubernetes supports it's own version of Nginx\footnote{
+\subsection{Nginx}
+\label{nginx}
+Kubernetes supports its own version of Nginx\footnote{
 \url{https://github.com/kubernetes/ingress-nginx}}, but there is also an
@@
 \url{https://github.com/nginxinc/kubernetes-ingress}}.
-We've chosen the second one, since it seems more stable.
+We have chosen the second one, since it seems more stable.
 
-\subsection{Installation}\label{installation}
+\subsection{Installation}
+\label{installation}
 
-The installation is pretty simple, and can be done with Helm:
-
+The installation is pretty simple, and it can be done with Helm:
 \begin{lstlisting}[language=bash]
@@
 similar to this one:
-
 \begin{lstlisting}[language=yaml]
@@
 So, to apply your forwarding to Kubernetes:
-
 \begin{lstlisting}[language=bash]
@@
 In the dashboard, you should see now something similar to this:
-
 \begin{figure}[htbp]
@@
 
-\subsection{Træfik}\label{truxe6fik}
+\subsection{Træfik}
+\label{truxe6fik}
 
-\subsection{Installation}\label{installation-1}
-
-Since Kubernetes 1.6 Role based access control was introduces so in
-order to make Træfik properly it must be configured. For simplicity,
-ClusterRoleBinding can be used as in the Træfik
-installation guide\footnote{
+\subsection{Installation}
+\label{installation-1}
+Since in Kubernetes 1.6 the Role based access control was introduced in order to 
+make Træfik working properly it must be configured. For simplicity, 
+ClusterRoleBinding can be used as in the Træfik installation guide\footnote{
 \url{https://docs.traefik.io/user-guide/kubernetes/}}:
@@
 
-Save that in a file \texttt{traefik-rbac.yaml} and run
-
+Save that in a file \texttt{traefik-rbac.yaml} and run:
 \begin{lstlisting}[language=bash]
@@
 
-After the ingress controller can be set up in to 2 way: using a
-deployment or a daemon set. We chose the first option for easier
-management deploying it with this YAML:
-
+After this the ingress controller can be set up in to 2 way: using a deployment 
+or a daemon set. We chose the first option for easier management. We deployed 
+it with this YAML:
 \begin{lstlisting}[language=yaml]
@@
 
-As before save it in a file, say \texttt{traefik-deployment.yaml} and run:
-
+As before, save it in a file, say \texttt{traefik-deployment.yaml} and run:
 \begin{lstlisting}[language=bash]
@@
 
-This deployment will expose two NodePorts which allow access both to the
-ingress and to the web management interface.
+This deployment will expose two \texttt{NodePorts} which allows access both to 
+the ingress and to the web management interface.
 
-\section{Additional Tools}\label{additional-tools}
+\section{Additional Tools}
+\label{additional-tools}
 
-\subsection{Helm}\label{helm}
+\subsection{Helm}
+\label{helm}
 
@@
 And after deploy the service:
-
 \begin{lstlisting}[language=bash]
@@
 we are able to deploy Tiller in our cluster (Helm backend), by running:
-
 \begin{lstlisting}[language=bash]

### 2018-11-22 res/sections/15-ArchitecturalImplementation.tex
Start the architectural implementation

@@
 \label{chap:archimpl}
+
+In this chapter we will dig into the actual architectural implementation, 
+seeing all the problems related to it and how they were solved.
+
+\section{Adopting Openstack for container orchestration}
+
+Our first step was to try to use Openstack to create a container orchestration. 
+The reason behind this approach was simple: Openbaton, on paper, provides an 
+out-of-the-box support for Openstack. In fact, it natively supports Openstack 
+PoPs, giving to us a great possibility to reuse the two frameworks and to ease 
+the development phase of our software.
+\subsection{Openstack for developers: Devstack}
+% TODO talk about Devstack installation

### 2018-11-22 res/sections/81-k8s.tex
Fix typos

@@
 whose aim is to configure the machine and to install all the necessary
-software. The boostrap script first \textbf{disables SELinux} on all machines 
+software. The bootstrap script first \textbf{disables SELinux} on all machines 
 (needed for Kubernetes networking):
@@
 Flannel\footnote{\url{https://github.com/coreos/flannel}}
-in order to get the networking working. Some time is necessay until the 
+in order to get the networking working. Some time is necessary until the 
 networking is up:

### 2018-11-22 res/sections/12-Requirements.tex
Add requirements table

\section{Requirements}\label{chap:prjan:sec:req}
In the following section are described the main requirements of this projects.

% Please add the following required packages to your document preamble:
% \usepackage{longtable}
% Note: It may be necessary to compile the document several times to get a multi-page table to line up properly
\begin{longtable}[c]{p{0.3\textwidth}p{0.7\textwidth}}
\hline
\multicolumn{1}{c}{\textbf{Requirement}}                                & \multicolumn{1}{c}{\textbf{Description}}                                                                                                                                                       \\ \hline
\endhead
%
\hline
\endfoot
%
\endlastfoot
%
1. Automatic creation and configuration of NFVI                         & It must be possible to create the environment on which SFC will be deployed in an automatic manner. The system must support Docker container and exploit Kubernetes as container orchestrator. \\
2. Implementation of chain orchestration and management tools           & The final work must be comprehensive of a minimal software implementation that allows the deployment and the management and the orchestration of VNF and their lifecycle.                      \\
3. Automatic configuration of chain orchestration and managements tools & Software instrumentation for VNF and SFC management must be configured automatically, at startup.                                                                                              \\
4. Developing software for manual SFC deployment                        & Instrumentation to deploy a chain of containers in which VNF can run must be develop. These containers must be able to communicate each others.                                                \\
5. VNF fault management                                                 & The system must be able to recover after a container fault automatically.                                                                                                                      \\
6. VNF scalability                                                      & The system must be able to scale the number of container implementing a function, both augmenting and decrementing the replicas.                                                               \\
7. SFC definition saving                                                & The system must be able to save a definition of a chain                                                                                                                                        \\
8. Deployment/stop/update of a chain from definition                    & The system must be able to allow the deployment of a chain from a saved definition, stop a running SFC and update it.                                                                          \\ \hline
\caption{Table of requirements}
\label{chap:prjan:tab:req}\\
\end{longtable}

### 2018-11-22 res/sections/11-ProjectScheduling.tex
Add schedule analysis and related problems

\section{Project Scheduling}
\label{chap:prjan:sec:prjsche}

Before starting this project a time requirements has been carried out, to see
its feasibility. We calculated at least six month necessary to be able to create
a working proof of concept (PoC) that allowed us to perform tests and, in
general, to have a feedback about the solution.

\subsection{Scheduled timeline}
Since our knowledge about the technologies was very poor at the beginning, we
preferred to invest a good chunk of time on them, to understand the already
present products available in the market and to choose what we really needed.
Additionally, at the end of this phase we though it was necessary to try these
products to see it they really fitted our necessities. After this phase, we
planned to actually starting implementing the missing components to create and
to fulfill the requirements defined in Section~\ref{chap:prjan:sec:req}. Since
the required technologies, Java and C/C++ are known to the authors, we estimated
an eight-week period related to coding and software implementation. After that,
a two weeks were planned in order to perform tests and gather metrics about the
system built, and too see how much the job done satisfied our expectations.
Finally, we estimated four weeks to group and write down all the information and
considerations we got during the thesis.

\subsection{Actual timeline}
Unfortunately, many difficulties slowed our project development, starting from
the beginning. The total time required to finish all the requirements decided
was of 37 weeks, namely nine months circa. Many problems emerged during the
study of the technologies, that revealed to be more complicated than expected,
especially to deploy and play with. Time was lost to obtain resources from the
University that was not initially planned. On top of that, we encountered little
bugs we had to manually fix, like for example in the Floodlight
Controller\footnote{See
  \url{https://github.com/floodlight/floodlight-webui/pull/10}}, that
additionally slowed our technologies exploration. This not-planned digging in
other projects' code cost us time and resources, delaying this phase of eight
weeks, for a total of 17 weeks. The software implementation revealed to be
irksome too. We end up underestimating the time required for this phase, because
we did not consider all the issues related to work with system calls. Moreover,
we had to implement more components than planned, in order to be able to
actually transmit data between two senders. Python, another program language,
had to be used to create proxy in order to carry out proper packet
encapsulation. This language, however simple, required some time to be picked up
by us, resulting in additional delay. In a nutshell, the additional time
required was of eight weeks, for a total of 16 weeks. Finally, the time to write
down the following thesis was limited due, and we had to write it down in a
total of two weeks, instead of the four initially planned.

### 2018-11-22 res/sections/11-ProjectScheduling.tex
Fix sentence

@@
 study of the technologies, that revealed to be more complicated than expected,
-especially to deploy and play with. Time was lost to obtain resources from the
-University that was not initially planned. On top of that, we encountered little
-bugs we had to manually fix, like for example in the Floodlight
-Controller\footnote{See
+especially since some framworks were more complicated to deploy and try them
+out. Time was lost to obtain resources from the University that was not
+initially planned. On top of that, we encountered little bugs we had to manually
+fix, like for example in the Floodlight Controller\footnote{See
   \url{https://github.com/floodlight/floodlight-webui/pull/10}}, that

### 2018-11-22 res/sections/11-ProjectScheduling.tex
Update res/sections/11-ProjectScheduling.tex
Co-Authored-By: Polpetta <poloniodavide@gmail.com>

@@
 Since our knowledge about the technologies was very poor at the beginning, we
-preferred to invest a good chunk of time on them, to understand the already
+preferred to invest a good chunk of time on studying them, to understand the already
 present products available in the market and to choose what we really needed.
@@
 we had to implement more components than planned, in order to be able to
-actually transmit data between two senders. Python, another program language,
-had to be used to create proxy in order to carry out proper packet
-encapsulation. This language, however simple, required some time to be picked up
-by us, resulting in additional delay. In a nutshell, the additional time
+actually transmit data between two senders. In a nutshell, the additional time
 required was of eight weeks, for a total of 16 weeks. Finally, the time to write

### 2018-11-22 res/sections/12-Requirements.tex
Code clean up
Update res/sections/12-Requirements.tex
Co-Authored-By: mzanella <mzanella@users.noreply.github.com>

@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{longtable}
-% Note: It may be necessary to compile the document several times to get a multi-page table to line up properly
 \begin{longtable}[c]{p{0.3\textwidth}p{0.7\textwidth}}
@@
 1.
-Automatic creation and configuration of NFVI                         & It must be possible to create the environment on which SFC will be deployed in an automatic manner.
+Automatic creation and configuration of NFVI                         & It must be possible to create the environment on which the SFC will be deployed in an automatic manner.
 The system must support Docker container and exploit Kubernetes as container orchestrator.
@@
 3.
-Automatic configuration of chain orchestration and managements tools & Software instrumentation for VNF and SFC management must be configured automatically, at startup.
+Automatic configuration of chain orchestration and management tools & Software instrumentation for VNFs and SFCs management must be configured automatically, at startup.
 \\

### 2018-11-22 res/sections/12-Requirements.tex
Update res/sections/12-Requirements.tex
Co-Authored-By: mzanella <mzanella@users.noreply.github.com>

@@
 2.
-Implementation of chain orchestration and management tools           & The final work must be comprehensive of a minimal software implementation that allows the deployment and the management and the orchestration of VNF and their lifecycle.
+Implementation of chain orchestration and management tools           & The final work must be comprehensive of a minimal software implementation that allows the deployment, the management and the orchestration of VNFs and their lifecycle.
 \\
@@
 5.
-VNF fault management                                                 & The system must be able to recover after a container fault automatically.
+VNF fault management                                                 & The system must be able to recover a container that has fault automatically.
 \\

### 2018-11-22 res/sections/12-Requirements.tex
Update res/sections/12-Requirements.tex
Co-Authored-By: mzanella <mzanella@users.noreply.github.com>

@@
 4.
-Developing software for manual SFC deployment                        & Instrumentation to deploy a chain of containers in which VNF can run must be develop.
-These containers must be able to communicate each others.
+Developing software for manual SFC deployment                        & An instrumentation to deploy a chain of containers in which VNFs can run must be developed.
+These containers must be able to communicate with each other.
 \\

### 2018-11-22 res/sections/11-ProjectScheduling.tex
Remove Python part
Update res/sections/11-ProjectScheduling.tex
Co-Authored-By: Polpetta <poloniodavide@gmail.com>
Fix typo

@@
 Additionally, at the end of this phase we though it was necessary to try these
-products to see it they really fitted our necessities. After this phase, we
+products to see if they really fitted our necessities. After this phase, we
 planned to actually starting implementing the missing components to create and
@@
 study of the technologies, that revealed to be more complicated than expected,
-especially since some framworks were more complicated to deploy and try them
+especially since some frameworks were more complicated to deploy and try them
 out. Time was lost to obtain resources from the University that was not

### 2018-11-22 res/sections/25-TestsResponsiveness.tex

@@
 \lstinputlisting[caption={Kubernetes YAML configuration that has been used to
-    simulate a single VNF deployment}, captionpos=b]{res/code/sfc_test.yaml}
+    simulate a single VNF deployment}, captionpos=b,
+  language=yaml]{res/code/sfc_test.yaml}
 

### 2018-11-22 res/sections/08-Background.tex
Add part about virtualization

@@
 going to briefly introduce and explain the technologies involved and their role.
+
+\section{Virtualization}
+It is remarkable how cloud computing has surged in the recent years. This is due
+the numerous features it carries, such as a great way to isolate processes and
+mitigate possible attacks from malicious users. On top of that, it allows to run
+``virtual'' machines with different configurations, without apporting any
+changes to the host machine. Indeed, the virtualization concepts refer to the
+property of a machine to be virtualized. To achieve this, the components that
+forms a common computer need to be virtualized too. Thus, these assets are
+recreated in this virtualized manner, e.g. Storage devices and computer network
+resources~\cite{liu2014research}. At the time writing, there are two common
+virtualization options: the virtualization through Hypervisor (defined also as
+hypervisor-based) and the container virtualization.
+
+\paragraph*{Hypervisor}
+The first one is known for providing a very good isolation from the host
+machine~\cite{eder2016hypervisor}, even if studies found possibilties for an
+attacker to escape from the a virtual machine and operate on the host
+machine\footnote{An example can be found in
+  \url{https://github.com/MorteNoir1/virtualbox_e1000_0day}}. Even though every
+component in a hypervisor environment is virtualized, there are possibilties for
+the ``guest'' OS (the virtualized one), to access to physical devices (like USB
+keys attacched on the ``host'' OS).
+
+\paragraph*{Container-based}
+Container-based virtualization uses a differnt approach to reach isolation.
+Instead of creating a whole new stack, it applies very strict policies to the
+programs running on the ``host'' machine. This concept has its roots back to the
+jail environments, where group of process where isolated by the main OS, that
+provided a virtualized filesystem~\cite{canonico2007virtualization}. In this
+context, a container-based applications can be seen as a jailed application with
+policies aimed to protect the host machine.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{docker_vs_hypervisor}
+  \caption[Container-based emulation vs hypervisor emulation]{Container-based
+    emulation, such as Docker vs hypervisor-based emulation~\cite{hung2016guidock}}
+\end{figure}
+
+\subsection{Network Virtulization}
+
+Virtualization assumes another context when referred to networking: virtualized
+networking allow multiple, heterogeneous networks to cohabit, increasing the
+flexibility and manageability. This network separation creates two different
+entities, that becomes decouples: who provides the infrastructure (commonly
+called InPs) and who provides the services (SPs). A virtual network can be
+create aggregating one or more InPs, thus creating a virtual network end-to-end
+service~\cite{chowdhury2009network}. An example of this are the Virtual Private
+Networks (VPNs), that are able to cross public networks, giving access to
+content that otherwise would not be available locally.
 

### 2018-11-22 res/sections/08-Background.tex
Add part about SDN and OpenFlow

@@
 content that otherwise would not be available locally.
+
+\section{Software-defined networking}
+The concept of Software-defined networking (SDN) is not new: starting from
+1996~\cite{sezer2013we} it aims to provide user-controlled management of
+forwarding in network nodes. This means to have the possibility to control the
+forwarding procedures. A close examples of SDN implementations are Ethane and
+OpenFlow, respectively created in 2007 and 2008. In an SDN architecture, there
+is the decoupling of data and control plane\todo{ref}, but there is a
+centralization of the networking logic and state~\cite{fundation2012software}.
+The use of an SDN network is estimated to reduce the Capital Expenditure (CapEX)
+and Operational Expenditure (OpEX) costs~\cite{benzekki2016software}.
+
+\subsection{SDN stack}
+\begin{figure}[ht]
+ \centering
+ \includegraphics[scale=1.3]{sdn_architecture}
+ \caption[SDN architecture schema]{SDN architecture schema. Is possible to 
+denote the three layer logic: the Application plane, the Controller plane and 
+the Data plane~\cite{fundation2013software}.}
+ \label{chap:background:img:sdn_architecture}
+\end{figure}
+
+As the Figure~\ref{chap:background:img:sdn_architecture} illustrats, the SDN
+architecture is formed by three layers, defined as planes, that accomplish
+determinated tasks~\cite{fundation2012software} as
+in~\cite{fundation2013software}. Below we will perform a brief exaplanation of
+each plane.
+
+\subsubsection{Application plane}
+In general, this layer is responsible of abstracting the SDN network control
+management~\cite{benzekki2016software}. The application plane holds SDN
+applications, programs that have the possibility to change the way the SDN
+network forwards the data. These applications have the possibility to access
+hardware capabilities through specific API, that are called \emph{northbound}
+interfaces (NBI). These interfaces usually, acts like a bridge between the SDN
+applications and the SDN controllers and are usually implemented in an open and
+vedor-neutral way~\cite{fundation2013software}.
+
+\subsubsection{Control plane}
+The SDN controllers, as just said in the previously, receives requests from the
+SND applications. Since the controller is centralized, it can therefore exploit
+the complete knowledge of the network. This gives the possibility to apply
+networking optimizations like flow management~\cite{sezer2013we}. The controller
+abstracts from the network complexity, and it can collect informations about the
+network performing requests to the \emph{southbound} interfaces. The control
+plane can have multiple controllers, that are in turn federated thanks to the
+\emph{eastbound} and \emph{westbound} (e.g.
+HyperFlow)~\cite{benzekki2016software}, that allow several existing controller
+written with different languages to coexists.
+
+\subsubsection{Data Plane}
+The data plane is the lowest plane, and it provides the aforementioned
+\emph{southbound} API. This kind of API can be deployed in two different kind of
+scenarios: for in-band communications and for out-of-band
+communications~\cite{benzekki2016software}. The data plane gives access to the
+physical and virtual switches, routers and access-points, and it gives the
+possibility to perform programmatic forwarding operations, but it is also able
+to provide statistic reportings and it supports event notifications.
+
+\subsection{OpenFlow protocol}
+OpenFlow is an open communication protocols that enables the possibility to
+actually program networks, offering the possibility to change the flow-tables
+for layer 3 switches and for routers~\cite{mckeown2008openflow}. In other words,
+the OpenFlow protocol gives the possibility for controller to access at these
+components. Routes can be changed on the typology of the packet, or on the state
+of the network (i.e. if a network is under heavy load, certain packages can be
+discarded).
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.65]{openflow}
+ \caption{OpenFlow protocol communication graph}
+ \label{chap:background:img:openflow_protocol}
+\end{figure}
+
+In the Figure~\ref{chap:background:img:openflow_protocol} we have a
+schematization of what we said before: every switch has a flow table that gets
+updated thanks the OpenFlow protocol by the SDN controller, that is the one
+managing the whole network.
 

### 2018-11-22 res/sections/08-Background.tex
Add stuff about Network Slice

@@
 
+\section{Network slicing}
+Internet Service Providers (ISP) provides networking services to a wide range of
+users, that have different necessities between them, and that can require
+different kind of characteristics, e.g. a network connectivity in a train has
+different networking requirements from the one regarding a company. It is with
+these presuppositions that emerges the necessity of service optimization based
+on the necessities of the users. To solve this problem in the recent years the
+concept of Networks Slice (NS) has been introducted~\cite{alliance20155g}. The
+NS aims to logically separate this set of services that present different
+use-cases, and to achieve this it expoits the SDN and Network Function Chaining
+(NFC) that allows network softwarization~\cite{ordonez2017network}.
+
+Even if the first idea of NS goes back to 2002~\cite{peterson2003blueprint}, NSs
+ancestor can be found in VLANs, that allows different hosts with no physical
+connectivity to be in the same network. The NS concept has attracted a lot of
+interest escpecially from the Industries, that is trying to introduce them in
+the next mobile network standard, the 5G. More details are provided in
+Section~\ref{chap:background:sec:5g:sub:ns}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.2]{networkslicing}
+  \caption[The three layers model of network slicing]{The three layers model of
+    network slicing~\cite{alliance2016description}}
+  \label{chap:background:img:network_slicing}
+\end{figure}
+
+As is possible to see from the Figure~\ref{chap:background:img:network_slicing},
+a network slice is divided in multiple layers, namely the Service Istance Layer,
+the Network Slice Instance Layer and the Resource Layer.
+
+\subsection{Service Instance Layer}
+The first layer we presented, the \emph{Service Instance Layer}, represents the
+business or user services supported. These services are achieved thanks to the
+Network Functions (NF)~\cite{kotulski2017end}.
+
+\subsection{Network Slice Instance Layer}
+Instead, the \emph{Network Slice Instance Layer} as is possible to see it
+contains multiple sub-network instances, tahtn can be virtualized or
+non-virtualized.
+
+\paragraph*{Network Functions}
+The concept behind NF is to usually elaborate incoming data and forward it to
+destination. NF can be physical of virtualized: in the first case is commonly a
+set of harware, in the second case instead it can be a software implementation,
+decoupled from the hardware where its running.
+
+\vspace{0.5cm}
+
+Every NF should be isolated, so that it only has the possibility to consumed a
+defined set of resources.
+
 \section{The upcoming connectivity standard: 5G}
@@
 \subsubsection{Network slicing}
+\label{chap:background:sec:5g:sub:ns}
 The role of a \emph{network slice} in a 5G architecture is to specifically

### 2018-11-23 res/sections/15-ArchitecturalImplementation.tex
Add something

@@
 Our first step was to try to use Openstack to create a container orchestration. 
-The reason behind this approach was simple: Openbaton, on paper, provides an 
+The reason behind this approach was simple: Openbaton, on paper, comes with an 
 out-of-the-box support for Openstack. In fact, it natively supports Openstack 
@@
 \subsection{Openstack for developers: Devstack}
+Openstack is composed by modules, giving the possibility to the user who wants 
+to install it to choose only the components it really needs. So the first thing 
+we decided was to perform a selection of these modules between the 33 available 
+in the installation page.
 % TODO talk about Devstack installation

### 2018-11-23 res/sections/15-ArchitecturalImplementation.tex
Add additional stuff

@@
 we decided was to perform a selection of these modules between the 33 available 
-in the installation page.
-% TODO talk about Devstack installation
+in the installation page. The installation of this components span between 
+multiple machines (nodes), therefore creating a whole cluster of resources. 
+Since we considered too time-consuming to install Openstack in a distribuited 
+manner, we considered to test first the developer version, called 
+\emph{Devstack}. Devstack offers the possibility to install a subset of modules 
+in the local machine, and to leverage a Openstack environment able to perform 
+the same operations of the production-grade installation. Thus, we starting 
+configurating the Devstack installation, installing first of all the Nova 
+module, that is required to perform computer operations such as the launch of 
+virtual machines, and lately we installed Taker too.\todo{exaplain Taker 
+functionalities}
+
+\paragraph*{Devstack configuration}
+The Openstack installation is not an easy one, particularly for newcomers, that 
+have to deal with a great set of tools and not-always-clear instructions. Here 
+we will describe how we managed to get a single-node deployment up and running.
+
+First of all, Devstack needs a machine with at least 16GB of ram and 50GB of HD 
+space.
+We decided to install these components:
+\begin{itemize}
+ \item Keystone (for identity management)
+ \item Object storage
+ \item Compute
+ \item Tacker (we tested this module to see is Openstack was a suitable VIM too)
+\end{itemize}
+
+Using Devstack is a simple operation, but the time required to reach a correct 
+configuration is very high. In particular, it required from 20 minutes to 1 
+hour to install Devstack we a particular configuration, without knowing if the 
+final installation was working or not. This trial-and-error approach consumed a 
+considerable amount of time, but it was the only possibility we had.

### 2018-11-23 res/sections/08-Background.tex
Finish part about network slicing

@@
 
+\subsection{Network slice layers}
+
 Even if the first idea of NS goes back to 2002~\cite{peterson2003blueprint}, NSs
@@
 
-\subsection{Service Instance Layer}
+\subsubsection{Service Instance Layer}
 The first layer we presented, the \emph{Service Instance Layer}, represents the
@@
 
-\subsection{Network Slice Instance Layer}
+\subsubsection{Network Slice Instance Layer}
 Instead, the \emph{Network Slice Instance Layer} as is possible to see it
@@
 set of harware, in the second case instead it can be a software implementation,
-decoupled from the hardware where its running.
+decoupled from the hardware where its running. The Virtualized Network Functions
+(VNFs), that are recently emerging propose a multitude of challenges, especially
+regarding their latency, reliability and efficiency due their virtualized nature
+of network function.
 
@@
 
-Every NF should be isolated, so that it only has the possibility to consumed a
-defined set of resources.
+\noindent Every NF should be isolated, so that it only has the possibility to
+consume a defined set of resources, otherwise the functions running on top of it
+could suffer performance degradation~\cite{richart2016resource}. Isolation has
+the benefits of offering more security and the possibility to individually
+manage the slice~\cite{ordonez2017network}. Slices have to respond to a defined
+set o requisites, first of all the possibilty to being created by a template,
+defined as Network Slice Template (NST). After the NS and the related resources
+are created, a NS should be activable. Obviously, for a NS there should be the
+possibility to apport changes on-the-fly, and to decommission or deactivate it,
+freeing all the occupied resources~\cite{slice20183gpp}.
+
+\subsubsection{Resource Layer}
+This layer contains all the physical and virtualized resources, that are needed
+to the network slice to properly function. The resources are divided in
+computational and storage resources.
+
+\subsection{Network Function Virtualization}
+Concept recently developed in 2012 by the ETSI organization, the Network
+Function Virtualization (NFV), is a network architecture concept that aims, with
+the virtualization technologies available in the market, to virtualize network
+nodes. It is belived tht NFV will trasform the wat the global communications
+will work, thanks to its disruptive trend in the networking
+industry~\cite{gray2016network}. Essentially, a NFV, working in harmony with the
+NS aims to provide sufficient packet processing efficency, especially in
+environment such as the cloud one.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=2]{etsi_arch}
+  \caption[Network Function Virtualization reference architecture]{Network
+    Function Virtualization reference architecture~\cite{etsi2013gs}}
+  \label{chap:background:img:etsi_arch}
+\end{figure}
+
+\noindent The NFV has its own architecture, divided in three main components:
+\begin{description}
+\item[Virtualised Network Function] that is a software implementation able to fun
+  a network function over a NFVI;
+\item[NFVI Infrastructure (NFVI)] that includes different physical resuources
+  and thei respective virtualization, that have to support the execution of the
+  aforementioned VNFs;
+\item[NFV Management and Orchestrator] that has the main role to orchestrate the
+  lifecycle of the available resources and of the network functions running on
+  top of them.
+\end{description}
+
+In the next section we are going to dig a little bit about the component that
+have not been explained yet.
+
+\subsubsection{Network Function Virtualized Infrastructure}
+With the term NFVI we mean the totality of all the hardware and software
+components where NFVs run~\cite{etsi2013gs}. The NFVI does not have to be
+centralized, but it can span across multiple locations, even though the
+resources should be virtualized in a manner from the VNF point of view that they
+look like coming form a single entity. The computing hardware that composes the
+NFVI can be either Commercial-of-the-shelf (COTS) hardware or a specialized one,
+while for the storage there is the possibility to have a Network Attached
+Storage (NAS) o a ``local'' storage directly on the server~\cite{etsi2013gs}.
+Finally, network resources, composed by switches, routers and links (both wired
+and wireless) can span between different domain, that can be of two typologies:
+\begin{itemize}
+\item NFVI-PoP
+\item Transport network
+\end{itemize}
+
+\subsubsection{NFV MANagement and Orchestration}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.7]{etsimano}
+  \caption[ETSI NFV MANO framework]{ETSI NFV MANO
+    framework~\cite{mijumbi2016management}}
+  \label{chap:background:img:etsimano}
+\end{figure}
+
+The MANagement and Orchestration (MANO) architecture is divided in turn in three
+main components, such as the Virtual Infrastructure Manager (VIM), the VNF
+Manager (VNFM) and the NFV Orchestrator (NFVO). The MANO organize and holds
+important data, regarding the Network Service (NS) catalog, the list of the VNFs
+definition (VNF catalog) and the current NFV instances and the NFVI resources.
+
+It is important to note how the mano is a critical aspect in ensuring
+cooperation between the NFVI and the VNFs. On top of that, is able to provide
+and manage the VNFs lifecycle~\cite{mijumbi2016management}.
+
+\paragraph{VIM}
+As we can see from the Figure~\ref{chap:background:img:etsimano}, it is possible
+to denote the position of this component in the schema. The VIM has the role to
+communicate with the NFV infrstructure, and to control its physical and virtual
+resources. A VIM should has the possibility to control one or more of NFVI. In
+general a VIM can be in charge of control a defined set of
+NFVI~\cite{mijumbi2016management}
+
+\paragraph{VNFM}
+The VNFM, as we can see from the schema, interacts with the Data repositories,
+the NFVO and with the VNFs. Its role is to manage the lifecycle of VNFs. The
+VNFM can handle one or multiple VNF istances.
+
+\paragraph{NFVO}
+We can see how the NFVO is the component with the most connections in the whiole
+architecture, and its able to interact with the OSS/BSS, the data repositories,
+the VNFM and the VIM. The aim of the NFVO is thus to orchestrate both the
+resources and the services.
+
+\paragraph{Data Repositories}
+Data repositories can be seen as a sort of databases, holding informations about
+the VNF definitions, the VNF instances with the rispective lifecycle status, and
+list of the Network Services defined and the resources available and occupied of
+the NFVI.
 

### 2018-11-23 res/sections/12-Requirements.tex
Update res/sections/12-Requirements.tex
Co-Authored-By: mzanella <mzanella@users.noreply.github.com>

@@
 Automatic creation and configuration of NFVI                         & It must be possible to create the environment on which the SFC will be deployed in an automatic manner.
-The system must support Docker container and exploit Kubernetes as container orchestrator.
+The system must support Docker container and exploit must be deployed on a clustered environment.
 \\
@@
 6.
-VNF scalability                                                      & The system must be able to scale the number of container implementing a function, both augmenting and decrementing the replicas.
+VNF scalability                                                      & The system must be able to scale the number of containers implementing a function, both augmenting and decrementing their replicas.
 \\

### 2018-11-23 res/sections/08-Background.tex
Finish SFC part

@@
 
+\section{Service Function Caining (SFC)}
+Another important component we have to talk about is the SFC, defined as a
+concept that it collocates logically above the NFVs. As said in~\cite{rfc7665}:
+\begin{displayquote}
+The definition and instantiation of an ordered set of service functions and
+subsequent ``steering'' of traffic through them is termed Service Function
+Chaining (SFC).
+\end{displayquote}
+
+\paragraph*{Service Function}
+A single element of an SFC is called Service Function (SF), and its duty is to
+elaborate the packets incoming and to forward them to the next receiver. SFs, as
+for NFs, can be virtualized or be a set of hardware functions. SF does not
+necessary know its part of a chain, and this defines it as SFC unaware or, on
+the contrary, SFC aware.
+
+\paragraph*{Architectural view}
+About the architectural view of SFCs, proposal defines several assumptions on
+them, starting from considering every element of the chain an opaque processing
+unit. Another important assumpion is that there is no global or standard SF
+chaining logic~\cite{rfc7665}. SFC can be easily seen as a graph, because some
+SF can have multiple address to forward the elaborated data, thus creating what
+is defined as \emph{fork} that ends up forming differen branchs. This coulds
+leeds to loops, that must be treated consequently.
+
+\paragraph*{Forwarding strategies}
+At high level, SFC proposes an abstracted view of a service, with a defined set
+of required SFs that must be executed in a particular order. These services can
+be unidirectional if they support data coming only from one direction, thus
+following the chain from the first SF to the last one, or they can be
+bidirectional, when they supports data coming from both directions, and thus the
+first SF can be the last one~\cite{rfc7665}. The forwarding logic has to be
+managed by the Service Function Forwarder (SFF), that forwards the data
+following the Service Function Path (SFP) defined.
+
+\paragraph*{Data encapsulation}
+The data passing inside an SFC must be encapsulated to ensure its integrity.
+Data encapsulation enables the possibility of sharing metadata/content, and to
+share information about the SFP to follow. Data gets continuously encapsulated
+and decapsulated between SFs, to give them the possibility to elaborate the
+payload contained.
+
+\subparagraph*{Network Serivce Header (NSH)}
+The encapsulation has an header that is well defined and descripted in detail
+in~\cite{rfc8300}. This header is composed by a 4-byte base header, and a 4-byte
+Service Path Header (SPH). The base header provides information about the services
+header and the kind of protocol contained as the payload, while the service path
+header provides path identification and location within a define Service Path
+(SP). Finally, the Context Header carries metadata.
+
+To further digging in the rabbit hole, the base header has several fields,
+identifying the NSH version, the total Time To Live (TTL) of a packet, the
+lenght of it, an additional metadata field and at the end it has a field to
+identify the protocol of the encapsulated data.
+
+The SPH contains, more than the 24-bit Service Path Identifier (SPI), a Service
+Index (SI), that provides the location within the SFP.
+
+\subparagraph*{Context header}
+There are two kind of context header definition: one with a fixed context header
+length (16-Byte) and one with a dynamic context header length.
+
+\noindent Regarding this field, the standard does not make any assuption about
+the content, and it is left free to be filled with custom metadata. The only
+requirement is to set this filed with zero when is not used.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.6]{nshtype1}
+  \caption[NSH MD Type \texttt{0x1}]{NSH MD Type \texttt{0x1}~\cite{rfc8300}.
+    The context header in this NSH has a fixed length. There is another possible
+    context header that has a variable length.}
+  \label{chap:background:img:nshtype1}
+\end{figure}
+
 \section{The upcoming connectivity standard: 5G}

### 2018-11-23 res/sections/08-Background.tex
Fix typos

@@
 mitigate possible attacks from malicious users. On top of that, it allows to run
-``virtual'' machines with different configurations, without apporting any
+``virtual'' machines with different configurations, without bringing any
 changes to the host machine. Indeed, the virtualization concepts refer to the
@@
 The first one is known for providing a very good isolation from the host
-machine~\cite{eder2016hypervisor}, even if studies found possibilties for an
+machine~\cite{eder2016hypervisor}, even if studies found possibilities for an
 attacker to escape from the a virtual machine and operate on the host
@@
   \url{https://github.com/MorteNoir1/virtualbox_e1000_0day}}. Even though every
-component in a hypervisor environment is virtualized, there are possibilties for
+component in a hypervisor environment is virtualized, there are possibilities for
 the ``guest'' OS (the virtualized one), to access to physical devices (like USB
-keys attacched on the ``host'' OS).
+keys attached on the ``host'' OS).
 
 \paragraph*{Container-based}
-Container-based virtualization uses a differnt approach to reach isolation.
+Container-based virtualization uses a different approach to reach isolation.
 Instead of creating a whole new stack, it applies very strict policies to the
@@
 jail environments, where group of process where isolated by the main OS, that
-provided a virtualized filesystem~\cite{canonico2007virtualization}. In this
+provided a virtualized file-system~\cite{canonico2007virtualization}. In this
 context, a container-based applications can be seen as a jailed application with
@@
 
-\subsection{Network Virtulization}
+\subsection{Network Virtualization}
 
@@
 
-As the Figure~\ref{chap:background:img:sdn_architecture} illustrats, the SDN
+As the Figure~\ref{chap:background:img:sdn_architecture} illustrates, the SDN
 architecture is formed by three layers, defined as planes, that accomplish
-determinated tasks~\cite{fundation2012software} as
-in~\cite{fundation2013software}. Below we will perform a brief exaplanation of
+specific tasks~\cite{fundation2012software} as
+in~\cite{fundation2013software}. Below we will perform a brief explanation of
 each plane.
@@
 applications and the SDN controllers and are usually implemented in an open and
-vedor-neutral way~\cite{fundation2013software}.
+vendor-neutral way~\cite{fundation2013software}.
 
@@
 The SDN controllers, as just said in the previously, receives requests from the
-SND applications. Since the controller is centralized, it can therefore exploit
+SDN applications. Since the controller is centralized, it can therefore exploit
 the complete knowledge of the network. This gives the possibility to apply
-networking optimizations like flow management~\cite{sezer2013we}. The controller
-abstracts from the network complexity, and it can collect informations about the
+networking optimization like flow management~\cite{sezer2013we}. The controller
+abstracts from the network complexity, and it can collect information about the
 network performing requests to the \emph{southbound} interfaces. The control
@@
 possibility to perform programmatic forwarding operations, but it is also able
-to provide statistic reportings and it supports event notifications.
+to provide statistic reporting and it supports event notifications.
 
@@
 
-In the Figure~\ref{chap:background:img:openflow_protocol} we have a
-schematization of what we said before: every switch has a flow table that gets
-updated thanks the OpenFlow protocol by the SDN controller, that is the one
-managing the whole network.
+In the Figure~\ref{chap:background:img:openflow_protocol} we have a schema of
+what we said before: every switch has a flow table that gets updated thanks the
+OpenFlow protocol by the SDN controller, that is the one managing the whole
+network.
 
@@
 on the necessities of the users. To solve this problem in the recent years the
-concept of Networks Slice (NS) has been introducted~\cite{alliance20155g}. The
+concept of Networks Slice (NS) has been introduced~\cite{alliance20155g}. The
 NS aims to logically separate this set of services that present different
-use-cases, and to achieve this it expoits the SDN and Network Function Chaining
+use-cases, and to achieve this it exploits the SDN and Network Function Chaining
 (NFC) that allows network softwarization~\cite{ordonez2017network}.
@@
 connectivity to be in the same network. The NS concept has attracted a lot of
-interest escpecially from the Industries, that is trying to introduce them in
+interest especially from the Industries, that is trying to introduce them in
 the next mobile network standard, the 5G. More details are provided in
@@
 As is possible to see from the Figure~\ref{chap:background:img:network_slicing},
-a network slice is divided in multiple layers, namely the Service Istance Layer,
-the Network Slice Instance Layer and the Resource Layer.
+a network slice is divided in multiple layers, namely the Service Instance
+Layer, the Network Slice Instance Layer and the Resource Layer.
 
@@
 Instead, the \emph{Network Slice Instance Layer} as is possible to see it
-contains multiple sub-network instances, tahtn can be virtualized or
+contains multiple sub-network instances, that can be virtualized or
 non-virtualized.
@@
 destination. NF can be physical of virtualized: in the first case is commonly a
-set of harware, in the second case instead it can be a software implementation,
+set of hardware, in the second case instead it can be a software implementation,
 decoupled from the hardware where its running. The Virtualized Network Functions
@@
 manage the slice~\cite{ordonez2017network}. Slices have to respond to a defined
-set o requisites, first of all the possibilty to being created by a template,
+set o requisites, first of all the possibility to being created by a template,
 defined as Network Slice Template (NST). After the NS and the related resources
-are created, a NS should be activable. Obviously, for a NS there should be the
-possibility to apport changes on-the-fly, and to decommission or deactivate it,
+are created, a NS should be activated. Obviously, for a NS there should be the
+possibility to bring changes on-the-fly, and to decommission or deactivate it,
 freeing all the occupied resources~\cite{slice20183gpp}.
@@
 the virtualization technologies available in the market, to virtualize network
-nodes. It is belived tht NFV will trasform the wat the global communications
+nodes. It is believed that NFV will transform the way the global communications
 will work, thanks to its disruptive trend in the networking
 industry~\cite{gray2016network}. Essentially, a NFV, working in harmony with the
-NS aims to provide sufficient packet processing efficency, especially in
+NS aims to provide sufficient packet processing efficiency, especially in
 environment such as the cloud one.
@@
 \begin{description}
-\item[Virtualised Network Function] that is a software implementation able to fun
+\item[Virtualized Network Function] that is a software implementation able to fun
   a network function over a NFVI;
-\item[NFVI Infrastructure (NFVI)] that includes different physical resuources
-  and thei respective virtualization, that have to support the execution of the
+\item[NFVI Infrastructure (NFVI)] that includes different physical resources
+  and their respective virtualization, that have to support the execution of the
   aforementioned VNFs;
@@
 to denote the position of this component in the schema. The VIM has the role to
-communicate with the NFV infrstructure, and to control its physical and virtual
+communicate with the NFV infrastructure, and to control its physical and virtual
 resources. A VIM should has the possibility to control one or more of NFVI. In
@@
 the NFVO and with the VNFs. Its role is to manage the lifecycle of VNFs. The
-VNFM can handle one or multiple VNF istances.
+VNFM can handle one or multiple VNF instances.
 
 \paragraph{NFVO}
-We can see how the NFVO is the component with the most connections in the whiole
+We can see how the NFVO is the component with the most connections in the whole
 architecture, and its able to interact with the OSS/BSS, the data repositories,
@@
 \paragraph{Data Repositories}
-Data repositories can be seen as a sort of databases, holding informations about
-the VNF definitions, the VNF instances with the rispective lifecycle status, and
+Data repositories can be seen as a sort of databases, holding information about
+the VNF definitions, the VNF instances with the respective lifecycle status, and
 list of the Network Services defined and the resources available and occupied of
@@
 
-\section{Service Function Caining (SFC)}
+\section{Service Function Chaining (SFC)}
 Another important component we have to talk about is the SFC, defined as a
@@
 them, starting from considering every element of the chain an opaque processing
-unit. Another important assumpion is that there is no global or standard SF
+unit. Another important assumption is that there is no global or standard SF
 chaining logic~\cite{rfc7665}. SFC can be easily seen as a graph, because some
 SF can have multiple address to forward the elaborated data, thus creating what
-is defined as \emph{fork} that ends up forming differen branchs. This coulds
-leeds to loops, that must be treated consequently.
+is defined as \emph{fork} that ends up forming different branches. This could
+lead to loops, that must be treated consequently.
 
@@
 
-\subparagraph*{Network Serivce Header (NSH)}
-The encapsulation has an header that is well defined and descripted in detail
+\subparagraph*{Network Service Header (NSH)}
+The encapsulation has an header that is well defined and described in detail
 in~\cite{rfc8300}. This header is composed by a 4-byte base header, and a 4-byte
@@
 identifying the NSH version, the total Time To Live (TTL) of a packet, the
-lenght of it, an additional metadata field and at the end it has a field to
+length of it, an additional metadata field and at the end it has a field to
 identify the protocol of the encapsulated data.
@@
 
-\noindent Regarding this field, the standard does not make any assuption about
+\noindent Regarding this field, the standard does not make any assumption about
 the content, and it is left free to be filled with custom metadata. The only

### 2018-11-24 res/sections/07-Introduction.tex
Fix thesis name and supervisor
Add some intro
New intro

@@
- % This is just a draft, nothing suitable for the final product: sources are in
- % form of a todo
-
- % --- INTRODUCTION ---
 \chapter{Introduction}
@@
 
- Through the history of internet, many factors changed and new innovations came,
+Through the history of Internet, many factors changed and new innovations came,
 but at the end of the day a fundamental key remains: connection performances
- are, indeed, what the users really want. With the arrival of mobile
- connectivity, wireless networks started to have an important role in everyone's
- life. In this context network efficiency is a decisive factor in the
- development of new connectivity standards (e.g 2G, 3G, LTE) and in the creation
- of ad-hoc transmission protocols (i.e. TCP versions specialized to exploit the
- characteristic of wireless networks, such as Hybla and Vegas). It is for this
- reason that, beside improved radio communications, back-end infrastructure
- began to be refined too. During their communication to the destination packets
- are elaborated, compressed and processed in order to make their transmission as
- fast as possible. On top of that, mobile traffic is continuously surging every
- year, with the backbones having more and more data to deliver, making
- in-hardware backbone solutions less efficient. Their time-to-deploy, in fact,
- is very high compared to the requirements of the network. With this trend,
- in-hardware solutions will not be able to keep up. Is with this considerations
- that a software approach is being proposed to become the new way to deploy
- network components. While less efficient, they are faster to bring them up,
- offering an unprecedented flexibility~\cite{nguyen2017sdn}. Furthermore, new
- virtualization technologies such as Cloud computing and Docker are becoming
- more and more mature for being employed in large-scale production environments,
- where new frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm),
- allow to easily manage the application lifecycle, i.e. deployment, scale
- up/down and finally, removal.
+are, indeed, what the users really want.
+Today, connectivity plays an important role in everyone's life. Several decades 
+brought solutions with enhanced wired and wireless connectivity, like 
+Ethernet, Fast Ethernet, 2G, 3G, LTE, and better communication protocols such 
+as a whole plethora of TCP standards in order to keep up with the continuous 
+surging of network traffic. Internet Service Providers to be able to 
+efficiently process the whole quantity of packets passing through the network, 
+applies several optimizations along the packet path. Until this day, these 
+optimizations have mostly been based on costly, in hardware solutions. In 
+particular, these Network Functions are able to perform only generic packet 
+elaboration, since they do not know the Quality of Service required by every 
+connection. It is in this context that 5G proposes a new way to approach 
+networking, suggesting the virtualization of this Network Functions, making 
+their deployments faster and cheaper. On top of that, it proposes specific 
+services for determined kind of data. If in the past a train running a 
+300km/h had to use the same network services of an IoT device sending data 
+updates every hour, now 5G proposes two different and efficient solutions for 
+this completely different services. This separation is performed through the 
+creation of determined network components called Network Slices. This new 
+possibility of offering the network with specialized class of services, 
+combined with emerging virtualization technologies like cloud computing, 
+hypervisor-virtualized machine or container-based solutions enables the 
+deployment of Virtualized Network Functions within seconds running on generic 
+purpose or custom hardware. While there may be some performances degreadations, 
+it has been demonstrated that their time-to-deploy and the possibility of 
+automation they offers are unprecedented~\cite{nguyen2017sdn}. On top of 
+that, specialized software acts as a MANager and Orchestrator (MANO) for this 
+VNFs, regulating their lifecycle. While most of the current proposals are based 
+on hypervisor on cloud solutions, in this thesis we are going to see a 
+proof-of-concept of a part of a Network Slice, equipped with a custom MANO 
+solution, based on the new container-based technology, called Docker, on a 
+recent orchestrator framework: Kubernetes.
 
- \newpage
-
- \section{Document organization}
+\section*{Document organization}
 

### 2018-11-24 res/sections/08-Background.tex

@@
 In this chapter we are going to explain the most important concepts of the
-upcoming 5G standard, talking about the Virtual Network Functions (VNFs) and 
-how first experiments with this new technology already begun. Finally we are 
-going to briefly introduce and explain the technologies involved and their role.
+upcoming 5G standard, talking about the Virtual Network Functions (VNFs), the
+Network Slice (NS) concept and how first experiments with these new innovations
+already begun. Finally we are going to briefly introduce and explain the
+technologies involved and their role.
 

### 2018-11-24 res/sections/08-Background.tex

@@
 mitigate possible attacks from malicious users. On top of that, it allows to run
-``virtual'' machines with different configurations, without bringing any
-changes to the host machine. Indeed, the virtualization concepts refer to the
-property of a machine to be virtualized. To achieve this, the components that
-forms a common computer need to be virtualized too. Thus, these assets are
-recreated in this virtualized manner, e.g. Storage devices and computer network
+virtual machines with different configurations, without bringing any changes to
+the host machine. Indeed, the virtualizations concepts refer to the property of
+a machine to be virtualized. To achieve this, the components that forms a common
+computer need to be virtualized too. Thus, these assets are recreated in this
+virtualized manner, e.g. storage devices and computer network
 resources~\cite{liu2014research}. At the time writing, there are two common
@@
 
-\paragraph*{Hypervisor}
+\paragraph*{Hypervisor-based}
 The first one is known for providing a very good isolation from the host
@@
 attacker to escape from the a virtual machine and operate on the host
-machine\footnote{An example can be found in
+one\footnote{An example can be found in
   \url{https://github.com/MorteNoir1/virtualbox_e1000_0day}}. Even though every
-component in a hypervisor environment is virtualized, there are possibilities for
-the ``guest'' OS (the virtualized one), to access to physical devices (like USB
-keys attached on the ``host'' OS).
+component in a hypervisor environment is virtualized, there are possibilities
+for the ``guest'' OS (the virtualized one), to access to physical devices (like
+USB keys attached on the ``host'' OS).
 
@@
 jail environments, where group of process where isolated by the main OS, that
-provided a virtualized file-system~\cite{canonico2007virtualization}. In this
-context, a container-based applications can be seen as a jailed application with
-policies aimed to protect the host machine.
+provided a virtualized file-system without any possibility to access to
+different paths~\cite{canonico2007virtualization}. In this context, a
+container-based application can be seen as a jailed application with policies
+aimed to protect the host machine.
 
@@
 Virtualization assumes another context when referred to networking: virtualized
-networking allow multiple, heterogeneous networks to cohabit, increasing the
+networking allows multiple, heterogeneous networks to cohabit, increasing the
 flexibility and manageability. This network separation creates two different
-entities, that becomes decouples: who provides the infrastructure (commonly
+entities, that becomes decoupled: who provides the infrastructure (commonly
 called InPs) and who provides the services (SPs). A virtual network can be
-create aggregating one or more InPs, thus creating a virtual network end-to-end
+created aggregating one or more InPs, thus creating a virtual network end-to-end
 service~\cite{chowdhury2009network}. An example of this are the Virtual Private
@@
 The concept of Software-defined networking (SDN) is not new: starting from
-1996~\cite{sezer2013we} it aims to provide user-controlled management of
+1996~\cite{sezer2013we} it aims to provide user-controlled management of data
 forwarding in network nodes. This means to have the possibility to control the
@@
 OpenFlow, respectively created in 2007 and 2008. In an SDN architecture, there
-is the decoupling of data and control plane\todo{ref}, but there is a
-centralization of the networking logic and state~\cite{fundation2012software}.
-The use of an SDN network is estimated to reduce the Capital Expenditure (CapEX)
-and Operational Expenditure (OpEX) costs~\cite{benzekki2016software}.
+is the decoupling of data and control plane (which will be explained in
+Section~\ref{chap:background:sec:sdn:sub:cp}), but there is a centralization of
+the networking logic and state~\cite{fundation2012software}. The use of an SDN
+network is estimated to reduce the Capital Expenditure (CapEX) and Operational
+Expenditure (OpEX) costs~\cite{benzekki2016software}.
 
@@
 architecture is formed by three layers, defined as planes, that accomplish
-specific tasks~\cite{fundation2012software} as
+specific tasks~\cite{fundation2012software} as explained
 in~\cite{fundation2013software}. Below we will perform a brief explanation of
@@
 \subsubsection{Control plane}
+\label{chap:background:sec:sdn:sub:cp}
 The SDN controllers, as just said in the previously, receives requests from the
@@
 \emph{eastbound} and \emph{westbound} (e.g.
-HyperFlow)~\cite{benzekki2016software}, that allow several existing controller
-written with different languages to coexists.
+HyperFlow)~\cite{benzekki2016software} interfaces, that allow several existing
+controller written even with different languages to coexists.
 
 \subsubsection{Data Plane}
-The data plane is the lowest plane, and it provides the aforementioned
-\emph{southbound} API. This kind of API can be deployed in two different kind of
+The data plane is the lowest plane and it provides the aforementioned
+\emph{southbound} API. This API can be deployed in two different kind of
 scenarios: for in-band communications and for out-of-band
@@
 the OpenFlow protocol gives the possibility for controller to access at these
-components. Routes can be changed on the typology of the packet, or on the state
-of the network (i.e. if a network is under heavy load, certain packages can be
-discarded).
+components. Routes can be changed based on the typology of the packet, or based
+on the state of the network (i.e. if a network is under heavy load, certain
+packages can be discarded).
 
@@
 users, that have different necessities between them, and that can require
-different kind of characteristics, e.g. a network connectivity in a train has
-different networking requirements from the one regarding a company. It is with
-these presuppositions that emerges the necessity of service optimization based
-on the necessities of the users. To solve this problem in the recent years the
-concept of Networks Slice (NS) has been introduced~\cite{alliance20155g}. The
-NS aims to logically separate this set of services that present different
-use-cases, and to achieve this it exploits the SDN and Network Function Chaining
-(NFC) that allows network softwarization~\cite{ordonez2017network}.
+different kind of characteristics, e.g. a network connectivity in a travelling
+train has different networking requirements from the one deployed in a company.
+It is with these presuppositions that emerges the necessity of service
+optimization based on the users necessities. To solve this problem in the recent
+years the concept of Networks Slice (NS) has been
+introduced~\cite{alliance20155g}. The NS aims to logically separate this set of
+services that present different use-cases, and in order to achieve this it
+exploits the SDN and Network Function Chaining (NFC) that allows network
+softwarization~\cite{ordonez2017network}.
 
@@
 connectivity to be in the same network. The NS concept has attracted a lot of
-interest especially from the Industries, that is trying to introduce them in
-the next mobile network standard, the 5G. More details are provided in
+interest especially from the Industries, that is trying to introduce them in the
+5G mobile network standard. About this, more details are provided in
 Section~\ref{chap:background:sec:5g:sub:ns}.
@@
 The concept behind NF is to usually elaborate incoming data and forward it to
-destination. NF can be physical of virtualized: in the first case is commonly a
-set of hardware, in the second case instead it can be a software implementation,
-decoupled from the hardware where its running. The Virtualized Network Functions
-(VNFs), that are recently emerging propose a multitude of challenges, especially
-regarding their latency, reliability and efficiency due their virtualized nature
-of network function.
+destination. NFs can be physical of virtualized: in the first case is commonly a
+set of hardware devices, in the second case instead it can be a software
+implementation, decoupled from the hardware where its running. The Virtualized
+Network Functions (VNFs), that are recently emerging propose a multitude of
+challenges, especially regarding their latency, reliability and efficiency due
+their virtualized nature of network function.
 
@@
 NS aims to provide sufficient packet processing efficiency, especially in
-environment such as the cloud one.
+environments such as the cloud one.
 
@@
 important data, regarding the Network Service (NS) catalog, the list of the VNFs
-definition (VNF catalog) and the current NFV instances and the NFVI resources.
+definitions (VNF catalog) and the current NFV instances and the NFVI resources.
 
@@
 
-The SPH contains, more than the 24-bit Service Path Identifier (SPI), a Service
-Index (SI), that provides the location within the SFP.
+The SPH contains the 24-bit Service Path Identifier (SPI) and a Service Index
+(SI), that provides the location within the SFP.
 
 \subparagraph*{Context header}
-There are two kind of context header definition: one with a fixed context header
-length (16-Byte) and one with a dynamic context header length.
+There are two kind of context header definitions: one with a fixed context
+header length (16-Byte) and one with a dynamic context header length.
 
@@
 the content, and it is left free to be filled with custom metadata. The only
-requirement is to set this filed with zero when is not used.
+requirement is to set this filed with zeroes when is not used.
 
@@
 
-\section{The upcoming connectivity standard: 5G}
-The continuous innovation in the mobile network connectivity is leading to the
+\section{5G}
+The continuous innovations in the mobile network connectivity is leading to the
 creation of a new standard, the 5G, which is estimated to arrive to the market

### 2018-11-24 res/sections/08-Background.tex
Chap reorder

@@
 technologies involved and their role.
+
+\section{The VIBeS project}
+This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
+requirement is to enhance the performance of end-to-end IP based that involves
+satellite connections. To reach this goal, the project specifications suggest to
+exploit the 5G incoming technology and use the NFV-MANO architecture to perform
+first packets elaboration and performance improvement and finally TCP/IP
+satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
+VIBES project proposed five technical requirements:
+\begin{enumerate}
+  \item Analysis of the applicability of current and new Internet protocols in
+  the proposed VNF-PEP architecture
+  \item Implementation of a VNF-PEP prototype
+  \item Building of a PoC test platform
+  \item VNF-PEP validation and performance tests on 5G use-cases
+  \item Demonstration test-bed management
+\end{enumerate}
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=1]{vibes_logo}
+ \caption{ViBeS project logo}
+ \label{chap:background:img:vibes_logo}
+\end{figure}
+
+\subsection{VNF-PEP architecture and internet protocols}
+VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
+the project is to research and include this technology in the usual VNF
+ecosystem to include without change satellite communications to the SFC. The
+architecture of the systems will be based on clusters of containers whose
+management and orchestration will follows the ETSI MANO architecture and SDN
+principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
+how to optimize a certain communication.
+
+%The analysis of this topic revealed to be trivial: since internet has many
+%different protocols that would become infeasible to support all of them at the
+%same time, packet encapsulation present itself as the only feasible solution:
+%every packet incoming in the VNFs has already been encapsulated by a generic
+%packet encapsulator/decapsulator,
+%\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
+%the protocol used on the edges but 1) use a ``quick protocol to exchange data 
+%among VNFs and 2) we are hiding the path not the protocol itself. As we were 
+%discussing, we are creating some sort of proxy, so the aim will be the same 
+%even if we support a plethora of protocols.''} making the whole architecture 
+%independent from the protocol a particular flux of data. To achieve this, 
+%several solutions have been studied, and at the end packet encapsulation with 
+%TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
+%this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
+%reference with the section of the explanation}
+%
+%\subsection{VNF-PEP prototype}
+%
+%Since the requirement for the whole system (MANO+PEP) were too challenging, the 
+%goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
+%The VNF architecture was shaped following the container orchestrator we decided 
+%to use. An more detailed architectural implementation can be found at 
+%Chapter~\ref{chap:archimpl}.
+%
+%\vspace{0.5cm}
+%
+%Starting with the first step of creating a MANO able to process incoming data
+%packets through VNF functions, we encountered that many networking tools already
+%present in the market required some tweaking and some integration, shifting our
+%goal to create a complete European Telecommunications Standards Institutes
+%(ETSI) Management and orchestrator (MANO) test-bed instead, following the
+%specifications suggested in the RFC 7665, thus implementing only the first three
+%requisites, without digging in the satellite data flow optimization. In
+%particular, we discovered how, these tools, were suitable to create ETSI MANO
+%and VNFs using virtual machine or exploiting cloud technologies, while they were
+%not designed with enough flexibility an integration with Docker. In the next
+%chapter we are going to take a deep analysis of the cited technologies 
+%(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
+%context to be able to understand our design choices we are going to describe 
+%later.
+%
+%\noindent With this in mind, we performed a requirements analysis described in
+%the next chapter.\todo{This should be put at the end of the introduction.}
+%
+
+\section{5G}
+The continuous innovations in the mobile network connectivity is leading to the
+creation of a new standard, the 5G, which is estimated to arrive to the market
+consumer in 2020~\cite{iwamura2015ngmn}. Lead by the Next Generation Mobile
+Network (NGMN) alliance, a group composed by the major players in the field of
+mobile connectivity, the 5G aims to offer not only at the end-user a new way to
+browse the web, download and watch interactive content but also to create an
+ad-hoc solution for Machine-to-Machine (M2M) data traffic, which is increasing
+more than ever thanks to the spreading of IoT devices, which sensors need to
+continuously send data to servers/data-centers. Relatively to \emph{Long Term
+  Evolution} (LTE), 5G points to have data rates $10$ times better, with $10$
+times smaller end-to-end latency and an increased connection density by $100$
+times~\cite{alliance20155g}.
+
+\begin{table}[t]
+\centering
+\resizebox{\textwidth}{!}{%
+  \begin{tabular}{p{4,5cm}|p{5,5cm}|p{5cm}}
+\textbf{Attribute}                                                    & 
+\textbf{LTE capability}
+
+                                                                           & 
+\textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
+\textbf{Data rate (per user)}                                         & Up to 
+100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).
+
+                                                                     & 10X 
+expected on average and peak rates and 100X expected on cell edge \\
+\textbf{End-to-end latency}                                           & 10 ms 
+for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other 
+factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X 
+(smaller)                                                         \\
+\textbf{Mobility}                                                     & 
+Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
+civil aviation.                                                                
+& 
+1.5X                                                                  \\
+\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & 
+Typically $\sim$2,000 active users/km2.
+
+                                                                           & 
+100X                                                                 
+\end{tabular}%
+}
+\caption[5G improvement over LTE]{An extract from the official 5G white paper 
+illustrating the improvements of 5G relatively to LTE connections.}
+\label{chap:intro:table:ltevs5g}
+\end{table}
+
+Part of the new requirements can be satisfied using a large radio spectrum with
+higher frequencies. The utilization of higher frequencies, though, mean that the
+radio signals can be easily disrupted by physical objects, like buildings and
+many geographical elements (such as hills and mountains), clashing with the
+expectation of an ever-reachable connectivity. It is here where virtualization
+plays an important role. In fact, the re-design of some network components today
+existing via hardware can transform a monolithic networking approach to a 
+modular one, exploiting the flexibility that Virtual Network Functions 
+(VNFs)\cite{moens2014vnf} can offer thanks to virtualization, closing the gap to
+the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
+which require a decreased time to set up and deploy network services
+(specifically, from 90 hours to 90 minutes~\cite{networld20202014role}).
+
+\subsection{5G architecture}
+
+With the constraints placed on the requirements formulated by the NGMN alliance,
+5G envisage a multi-layered architecture, based on three main
+layers~\cite{alliance20155g}:
+\begin{itemize}
+\item \textbf{infrastructure resource layer}: physical resources that are 
+exposed via a virtualized interface, and that can be monitored using specific 
+APIs
+\item \textbf{business enablement layer}: where a library of functions and
+  deployment is contained, and its configuration is accessible via APIs
+\item \textbf{business application layer}: layer that contains specific
+  applications and services of the operator
+\end{itemize}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.55]{5garchitecture}
+  \caption[5G Architecture]{5G Architecture~\cite{alliance20155g}}
+  \label{chap:background:img:5garchitecture}
+\end{figure}
+
+This separation in layers allows to easily manage multiple identities 
+differently: an Internet Service Provider (ISP) could manage a multitude of 
+physical infrastructures and have multiple business enablements dislocated 
+along an entire continent for example, but it could decide to have only a 
+single centralized business application deployment that manage all the other 
+layers resources.
+
+\subsubsection{Network slicing}
+\label{chap:background:sec:5g:sub:ns}
+The role of a \emph{network slice} in a 5G architecture is to specifically
+handle the Control-plane of a particular service (e.g. smartphones traffic,
+autonomous driving, massive IoT), deploying resources in a manner that assure
+the required latency, security and reliability. While some very peculiar and
+legacy service could require specific hardware, the common resources between
+services could be shared in a virtualized way, providing auto-scaling
+capabilities in services that are under heavy network pressure.
 
@@
 \end{figure}
-
-\section{5G}
-The continuous innovations in the mobile network connectivity is leading to the
-creation of a new standard, the 5G, which is estimated to arrive to the market
-consumer in 2020~\cite{iwamura2015ngmn}. Lead by the Next Generation Mobile
-Network (NGMN) alliance, a group composed by the major players in the field of
-mobile connectivity, the 5G aims to offer not only at the end-user a new way to
-browse the web, download and watch interactive content but also to create an
-ad-hoc solution for Machine-to-Machine (M2M) data traffic, which is increasing
-more than ever thanks to the spreading of IoT devices, which sensors need to
-continuously send data to servers/data-centers. Relatively to \emph{Long Term
-  Evolution} (LTE), 5G points to have data rates $10$ times better, with $10$
-times smaller end-to-end latency and an increased connection density by $100$
-times~\cite{alliance20155g}.
-
-\begin{table}[t]
-\centering
-\resizebox{\textwidth}{!}{%
-  \begin{tabular}{p{4,5cm}|p{5,5cm}|p{5cm}}
-\textbf{Attribute}                                                    & 
-\textbf{LTE capability}
-
-                                                                           & 
-\textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
-\textbf{Data rate (per user)}                                         & Up to 
-100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).
-
-                                                                     & 10X 
-expected on average and peak rates and 100X expected on cell edge \\
-\textbf{End-to-end latency}                                           & 10 ms 
-for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other 
-factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X 
-(smaller)                                                         \\
-\textbf{Mobility}                                                     & 
-Functional up to 350 km/h (for certain bands up to 500 km/h). No support for 
-civil aviation.                                                                
-& 
-1.5X                                                                  \\
-\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & 
-Typically $\sim$2,000 active users/km2.
-
-                                                                           & 
-100X                                                                 
-\end{tabular}%
-}
-\caption[5G improvement over LTE]{An extract from the official 5G white paper 
-illustrating the improvements of 5G relatively to LTE connections.}
-\label{chap:intro:table:ltevs5g}
-\end{table}
-
-Part of the new requirements can be satisfied using a large radio spectrum with
-higher frequencies. The utilization of higher frequencies, though, mean that the
-radio signals can be easily disrupted by physical objects, like buildings and
-many geographical elements (such as hills and mountains), clashing with the
-expectation of an ever-reachable connectivity. It is here where virtualization
-plays an important role. In fact, the re-design of some network components today
-existing via hardware can transform a monolithic networking approach to a 
-modular one, exploiting the flexibility that Virtual Network Functions 
-(VNFs)\cite{moens2014vnf} can offer thanks to virtualization, closing the gap to
-the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
-which require a decreased time to set up and deploy network services
-(specifically, from 90 hours to 90 minutes~\cite{networld20202014role}).
-
-\subsection{5G architecture}
-
-With the constraints placed on the requirements formulated by the NGMN alliance,
-5G envisage a multi-layered architecture, based on three main
-layers~\cite{alliance20155g}:
-\begin{itemize}
-\item \textbf{infrastructure resource layer}: physical resources that are 
-exposed via a virtualized interface, and that can be monitored using specific 
-APIs
-\item \textbf{business enablement layer}: where a library of functions and
-  deployment is contained, and its configuration is accessible via APIs
-\item \textbf{business application layer}: layer that contains specific
-  applications and services of the operator
-\end{itemize}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.55]{5garchitecture}
-  \caption[5G Architecture]{5G Architecture~\cite{alliance20155g}}
-  \label{chap:background:img:5garchitecture}
-\end{figure}
-
-This separation in layers allows to easily manage multiple identities 
-differently: an Internet Service Provider (ISP) could manage a multitude of 
-physical infrastructures and have multiple business enablements dislocated 
-along an entire continent for example, but it could decide to have only a 
-single centralized business application deployment that manage all the other 
-layers resources.
-
-\subsubsection{Network slicing}
-\label{chap:background:sec:5g:sub:ns}
-The role of a \emph{network slice} in a 5G architecture is to specifically
-handle the Control-plane of a particular service (e.g. smartphones traffic,
-autonomous driving, massive IoT), deploying resources in a manner that assure
-the required latency, security and reliability. While some very peculiar and
-legacy service could require specific hardware, the common resources between
-services could be shared in a virtualized way, providing auto-scaling
-capabilities in services that are under heavy network pressure.
-
-\section{The VIBES project}
-This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
-requirement is to enhance the performance of end-to-end IP based that involves
-satellite connections. To reach this goal, the project specifications suggest to
-exploit the 5G incoming technology and use the NFV-MANO architecture to perform
-first packets elaboration and performance improvement and finally TCP/IP
-satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
-VIBES project proposed five technical requirements:
-\begin{enumerate}
-  \item Analysis of the applicability of current and new Internet protocols in
-  the proposed VNF-PEP architecture
-  \item Implementation of a VNF-PEP prototype
-  \item Building of a PoC test platform
-  \item VNF-PEP validation and performance tests on 5G use-cases
-  \item Demonstration test-bed management
-\end{enumerate}
-
-\begin{figure}[t]
- \centering
- \includegraphics[scale=1]{vibes_logo}
- \caption{ViBeS project logo}
- \label{chap:background:img:vibes_logo}
-\end{figure}
-
-\subsection{VNF-PEP architecture and internet protocols}
-VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
-the project is to research and include this technology in the usual VNF
-ecosystem to include without change satellite communications to the SFC. The
-architecture of the systems will be based on clusters of containers whose
-management and orchestration will follows the ETSI MANO architecture and SDN
-principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
-how to optimize a certain communication.
-
-%The analysis of this topic revealed to be trivial: since internet has many
-%different protocols that would become infeasible to support all of them at the
-%same time, packet encapsulation present itself as the only feasible solution:
-%every packet incoming in the VNFs has already been encapsulated by a generic
-%packet encapsulator/decapsulator,
-%\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
-%the protocol used on the edges but 1) use a ``quick protocol to exchange data 
-%among VNFs and 2) we are hiding the path not the protocol itself. As we were 
-%discussing, we are creating some sort of proxy, so the aim will be the same 
-%even if we support a plethora of protocols.''} making the whole architecture 
-%independent from the protocol a particular flux of data. To achieve this, 
-%several solutions have been studied, and at the end packet encapsulation with 
-%TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
-%this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
-%reference with the section of the explanation}
-%
-%\subsection{VNF-PEP prototype}
-%
-%Since the requirement for the whole system (MANO+PEP) were too challenging, the 
-%goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
-%The VNF architecture was shaped following the container orchestrator we decided 
-%to use. An more detailed architectural implementation can be found at 
-%Chapter~\ref{chap:archimpl}.
-%
-%\vspace{0.5cm}
-%
-%Starting with the first step of creating a MANO able to process incoming data
-%packets through VNF functions, we encountered that many networking tools already
-%present in the market required some tweaking and some integration, shifting our
-%goal to create a complete European Telecommunications Standards Institutes
-%(ETSI) Management and orchestrator (MANO) test-bed instead, following the
-%specifications suggested in the RFC 7665, thus implementing only the first three
-%requisites, without digging in the satellite data flow optimization. In
-%particular, we discovered how, these tools, were suitable to create ETSI MANO
-%and VNFs using virtual machine or exploiting cloud technologies, while they were
-%not designed with enough flexibility an integration with Docker. In the next
-%chapter we are going to take a deep analysis of the cited technologies 
-%(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
-%context to be able to understand our design choices we are going to describe 
-%later.
-%
-%\noindent With this in mind, we performed a requirements analysis described in
-%the next chapter.\todo{This should be put at the end of the introduction.}
-%

### 2018-11-24 res/sections/07-Introduction.tex
Fix typos

@@
 but at the end of the day a fundamental key remains: connection performances
-are, indeed, what the users really want.
-Today, connectivity plays an important role in everyone's life. Several decades 
-brought solutions with enhanced wired and wireless connectivity, like 
-Ethernet, Fast Ethernet, 2G, 3G, LTE, and better communication protocols such 
-as a whole plethora of TCP standards in order to keep up with the continuous 
-surging of network traffic. Internet Service Providers to be able to 
-efficiently process the whole quantity of packets passing through the network, 
-applies several optimizations along the packet path. Until this day, these 
-optimizations have mostly been based on costly, in hardware solutions. In 
-particular, these Network Functions are able to perform only generic packet 
-elaboration, since they do not know the Quality of Service required by every 
-connection. It is in this context that 5G proposes a new way to approach 
-networking, suggesting the virtualization of this Network Functions, making 
-their deployments faster and cheaper. On top of that, it proposes specific 
-services for determined kind of data. If in the past a train running a 
+are, indeed, what the users really want. Today, connectivity plays an important
+role in everyone's life. Several decades brought solutions with enhanced wired
+and wireless connectivity, like Ethernet, Fast Ethernet, 2G, 3G, LTE, and better
+communication protocols such as a whole plethora of TCP standards in order to
+keep up with the continuous surging of network traffic. Internet Service
+Providers to be able to efficiently process the whole quantity of packets
+passing through the network, applies several optimization along the packet path.
+Until this day, these optimization have mostly been based on costly, in hardware
+solutions. In particular, these Network Functions are able to perform only
+generic packet elaboration, since they do not know the Quality of Service
+required by every connection. It is in this context that 5G proposes a new way
+to approach networking, suggesting the virtualization of this Network Functions,
+making their deployments faster and cheaper. On top of that, it proposes
+specific services for determined kind of data. If in the past a train running a
 300km/h had to use the same network services of an IoT device sending data
@@
 creation of determined network components called Network Slices. This new
-possibility of offering the network with specialized class of services, 
-combined with emerging virtualization technologies like cloud computing, 
+possibility of offering the network with specialized class of services, combined
+with emerging virtualization technologies like cloud computing,
 hypervisor-virtualized machine or container-based solutions enables the
 deployment of Virtualized Network Functions within seconds running on generic
-purpose or custom hardware. While there may be some performances degreadations, 
+purpose or custom hardware. While there may be some performances degreadation,
 it has been demonstrated that their time-to-deploy and the possibility of
-automation they offers are unprecedented~\cite{nguyen2017sdn}. On top of 
-that, specialized software acts as a MANager and Orchestrator (MANO) for this 
-VNFs, regulating their lifecycle. While most of the current proposals are based 
-on hypervisor on cloud solutions, in this thesis we are going to see a 
+automation they offers are unprecedented~\cite{nguyen2017sdn}. On top of that,
+specialized software acts as a MANager and Orchestrator (MANO) for this VNFs,
+regulating their lifecycle. While most of the current proposals are based on
+hypervisor on cloud solutions, in this thesis we are going to see a
 proof-of-concept of a part of a Network Slice, equipped with a custom MANO

### 2018-11-24 res/sections/07-Introduction.tex
Another fix

@@
 deployment of Virtualized Network Functions within seconds running on generic
-purpose or custom hardware. While there may be some performances degreadation,
+purpose or custom hardware. While there may be some performances degradation,
 it has been demonstrated that their time-to-deploy and the possibility of

### 2018-11-24 res/sections/07-Introduction.tex
Fix document organization

@@
  \item Chapter~\ref{chap:background} describes background information about 5G
-and the technologies used to build an ETSI compliant architecture
+   and the technologies used to build an ETSI compliant architecture;
  \item Chapter~\ref{chap:rel_wk} gives an insight of correlated works on this
-field
+   field;
  \item Chapter~\ref{chap:prjan} explains the analysis performed on the
-requirements provided and the design choices performed.
- \item Chapter~\ref{chap:archimpl} provides an in-depth view of the 
-high-level architectural implementation.
+   requirements provided and the design choices performed;
+ \item Chapter~\ref{chap:archimpl} provides an in-depth view of the high-level
+   architectural implementation;
+ \item Chapter~\ref{chap:vnf_ns_impl} gives a bief explanation of the VNF and NS
+   implementation;
+ \item Chapter~\ref{chap:test} describes the tests done on the plaftorm;
  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.

### 2018-11-24 res/sections/07-Introduction.tex
Fix typos

@@
    architectural implementation;
- \item Chapter~\ref{chap:vnf_ns_impl} gives a bief explanation of the VNF and NS
+ \item Chapter~\ref{chap:vnf_ns_impl} gives a brief explanation of the VNF and NS
    implementation;
- \item Chapter~\ref{chap:test} describes the tests done on the plaftorm;
+ \item Chapter~\ref{chap:test} describes the tests done on the platform;
  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.

### 2018-11-24 res/sections/15-ArchitecturalImplementation.tex
Add content about first implementation attempt

@@
 
-In this chapter we will dig into the actual architectural implementation, 
-seeing all the problems related to it and how they were solved.
+In this chapter we will dig into the actual architectural implementation, seeing
+all the problems related to it and how they were solved. We will show all the
+various attemps made, and we will discussion about the technological issue and
+changes we had to perform.
 
-\section{Adopting Openstack for container orchestration}
+\section{First attempt}
+
+In our first attempt we started building our infrastructure on top of Ubuntu
+16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton).
+
+\subsection{Adopting Openstack for container orchestration}
 
@@
 the development phase of our software.
-\subsection{Openstack for developers: Devstack}
+\subsubsection{Openstack for developers: Devstack}
 Openstack is composed by modules, giving the possibility to the user who wants 
@@
 module, that is required to perform computer operations such as the launch of 
-virtual machines, and lately we installed Taker too.\todo{exaplain Taker 
-functionalities}
+virtual machines, and lately we installed Tacker too.
+
+\paragraph*{Taker}
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.3]{tacker_architecture}
+  \caption[Taker architecture schema]{Taker architecture
+    schema~\cite{tackerOpenstackwiki}}
+\end{figure}
+Tacker is a VNFM and a NFVO. It is an official Openstack module, and it has the
+ability to deploy, edit, stop and do periodic health checks VNF directly on it.
+The develpers claim it has the possbility to make deployments to platforms
+compatible with Openstack too. It supports VNF definition through the TOSCA
+standard, and a proposal for SFC integration has been proposed in
+2015~\cite{tackerOpenstackwiki}.
+
+We sperimented Tacker to see if its functionalities were good enough to be
+adopted in our testbed instead of Openbaton. Its complete integration with
+Openstack, indeed, was the main reason for its testing. We felt like the product
+was not fully ready to be used in a real deployment though, since it lacked of
+options like SFC management and the fact that it only officially supported
+Openstack and not other cloud platforms was another reason to prefer Openbaton
+eventually.
 
@@
 considerable amount of time, but it was the only possibility we had.
+
+For testing purposes, and in order to better understanding the Openstack
+functionalities, we configured a first version with Virtual Machine support
+only.
+
+\paragraph*{Openbaton configuration}
+With Devstack ready to be used as NFVI, we started working on Openbaton.
+Openbaton currently offers three ways to be deployed:
+\begin{itemize}
+  \item Normal installation. Openbaton offers official Ubuntu repositories that,
+    added into the system, offer a very smooth installation via the
+    \verb!aptitude! command line utility, with the only cons of having to wait
+    for the automatic Openbaton download, installation and configuration;
+  \item Using Vagrant. Vagrant is a software that helps the deployment of
+    software, and it offers drivers to deploy virtual virtual machines too;
+  \item Docker Compose. At the time of our study, the Docker Compose had three
+    different configurations, based on the necessity of the user. A minimal,
+    standard and full deployment were offered.
+\end{itemize}
+
+\subparagraph*{Installation via repository}
+
+\subparagraph*{Deployment via Vagrant}
+
+\subparagraph*{Deployment via Docker Compose}

### 2018-11-25 res/sections/15-ArchitecturalImplementation.tex
Add schema and content about the first attempt

@@
 In our first attempt we started building our infrastructure on top of Ubuntu
-16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton).
+16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton). Our main goal was to set
+up Openstack as our VIM while setting Openbaton as our MANO. Then, Openbaton
+would have Openstack registered as PoP, where TOSCA definition would be launched
+on Openstack that, instead of deploying hypervisor based virtual machines, it
+would have deployed container instances. The VNF lifecycles would have been
+controller through an Element Management System (EMS), transmitting the various
+VNF states via the message broker RabbitMQ. Is possible to see the interactions
+between elements in a minimum schema available in
+Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{Attempt1}
+  \caption[Fist attempt components organization schema]{Fist attempt components
+    organization schema. We can see how the main Openstack role is to deploy and
+    manage the resources, while Openbaton has the role of maangement and
+    orchestration. Finally, RabbitMQ sends and receives status update from the
+    EMS for every VNF deployed.}
+  \label{chap:archimpl:sec:fistattempt:img:schema1}
+\end{figure}
 
@@
 \paragraph*{Taker}
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
@@
 
+\subsubsection{Openbaton installation}
 \paragraph*{Openbaton configuration}
@@
 
+Now we are going to do a short description about every installation/deploy method.
+
 \subparagraph*{Installation via repository}
+The installation via repositories granted us the possibility to select only the
+necessary parts we needed. The back of the medal of this approach was the
+impossibility to automate it: in fact, during the installation proces user
+interaction was required, that we were not able to write down in a scripted
+fashion.
 
 \subparagraph*{Deployment via Vagrant}
+The deployment via Vagrant confirmed our expectations because it revealed to be
+was always smooth. The downside of this approach was the lack of customization
+and of component management. The possibility to remove or add Openbaton Modules
+dynamically later become fundamental to us. On top of that, the Vagrant
+deployment uses a hyperviso virtualization system (based on Virtual Box), in
+contrast with our goal to build a fully container-based system.
 
 \subparagraph*{Deployment via Docker Compose}
+The deployment with Docker Compose tourned out to be best of the three
+possibilities. Docker Compose configuration offered an unprecedented
+configuration flexibility, with the possibility to deploy only selected modules.
+On top of that, this method let us manage the Openbaton version freely, having
+more control on update especially during the development of our components.
+Finally, a Docker based solution seemed to be the best approach since all of our
+components were meant to be put inside a Docker container.

### 2018-11-25 res/sections/15-ArchitecturalImplementation.tex
Add first attempt draft

@@
 \subparagraph*{Deployment via Docker Compose}
-The deployment with Docker Compose tourned out to be best of the three
-possibilities. Docker Compose configuration offered an unprecedented
+The deployment with Docker Compose turned out to be best of the three
+possibilities. The Docker Compose configuration offered an unprecedented
 configuration flexibility, with the possibility to deploy only selected modules.
@@
 more control on update especially during the development of our components.
-Finally, a Docker based solution seemed to be the best approach since all of our
-components were meant to be put inside a Docker container.
+Additionally, using this Docker Compose configuration allowed us to seamlessly
+integrate the monitoring tools Zabbix, in order to have the possibility to
+graphically see the data fetch from Openbaton about the system state. Finally, a
+Docker based solution seemed to be the best approach since all of our components
+were meant to be put inside a Docker container.
+
+\subsection{Docker integration in Openstack}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{dockeropenstackmodule}
+  \caption[Docker module for Openstack working schema]{Docker module for Openstack
+    working schema~\cite{openstackDockerModule}}
+  \label{chap:archimpl:sec:fistattempt:img:dockeropnestackmodule}
+\end{figure}
+
+After the initial Openstack and Openbaton configuration we started making the
+components talking to each other. First, we tried a simple \verb!iperf! data
+transmission with two VNFs deployed by Openbaton as a common hypervisor-based
+Openstack instances, as explained in~\cite{openbatonIperf}. With this test
+completed, we started tinkering Openstack for using Docker containers instead of
+traditional hypervisor instances. When we started developing this solution, we
+adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
+Unfortunately, we were not able to make it properly work, in particular, it was
+not possible to open selected desired ports on the Docker containers. This lead
+to the impossibility to perform packet forwarding between VNFs. On top of that,
+we found difficulties regarding the integration of Docker images in TOSCA
+definitions and Openstack. These two issues are related: the TOSCA definition
+does not have any keyword to specificate in any way that the image used for the
+VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
+time of the implementation did not performed the image pull from any Docker
+registry. The impossibility to easily set the containers port and the lack of
+some support in between the TOSCA definition and the Openstack Zun plugin pushed
+us to drop Openstack as a possible solution and to start looking for another
+valid VIM component.

### 2018-11-25 res/sections/15-ArchitecturalImplementation.tex
Start second attempt part

@@
 Unfortunately, we were not able to make it properly work, in particular, it was
-not possible to open selected desired ports on the Docker containers. This lead
-to the impossibility to perform packet forwarding between VNFs. On top of that,
-we found difficulties regarding the integration of Docker images in TOSCA
+not possible to open desired ports on the Docker containers. This lead to the
+impossibility to perform packet forwarding between VNFs. On top of that, we
+found difficulties regarding the integration of Docker images in TOSCA
 definitions and Openstack. These two issues are related: the TOSCA definition
@@
 time of the implementation did not performed the image pull from any Docker
-registry. The impossibility to easily set the containers port and the lack of
+registry. The impossibility to easily set the containers ports and the lack of
 some support in between the TOSCA definition and the Openstack Zun plugin pushed
@@
 valid VIM component.
+
+\section{Second attempt}
+
+After the failure to find a viable setting for Openstack, we decided to look
+around for other solutions. We started studing proper Docker container
+orchestrators, that could offer all the features we needed: container lifecycle
+management, scaling, healing and monitoring. We decided to pick two of the most
+mature and famous ones: Docker Swarm and Kubenetes. The decision to study only
+these two relies on their big community supporting them and the possibility to
+easily found projects that used the forementioned frameworks.
+
+\subsection{Docker Swarm}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.35]{swarm_diagram}
+  \caption[Docker Swarm schema]{Docker Swarm schema~\cite{dockerSwarmWiki}.}
+  \label{chap:archimpl:sec:secondattempt:img:dockerswarm}
+\end{figure}
+
+First we started playing with Docker Swarm. Docker Swarm is a product directly
+made by the Docker Inc., so it seamlessly integrates with other Docker tools
+such as Docker Compose. The installation and set-up is pretty straightforward,
+and the configuration is almost automatically handed by Docker itself. The
+minimum number of nodes in order to create a proper Docker Swarm cluster is of
+free nodes, where one must be a manager node (designed to orchestrate the other
+nodes in the cluster), while the other can be simple workers. Multiple managers
+can coexist in the same cluster, as long as only one effectively operates. The
+others act like backup manager in case the one orchestrating goes down for some
+reason. The networking is given out-of-the box, and the same holds for
+distributed persistence. While this for most of the uses cases can be considered
+a pro, in our specific situation the impossibility to tweak the network
+configuration was a cons, since it was important for us to have the full control
+over the network, especially regarding the SFC implementation, explained better
+in Chapter~\ref{chap:vnf_ns_impl}.
+
+\subsection{Kubernetes}

### 2018-11-26 res/sections/15-ArchitecturalImplementation.tex
Add something

@@
 around for other solutions. We started studing proper Docker container
-orchestrators, that could offer all the features we needed: container lifecycle
+orchestrators that could offer all the features we needed: container lifecycle
 management, scaling, healing and monitoring. We decided to pick two of the most
@@
 \subsection{Kubernetes}
+Kubenetes, as already described in the technology section, is a orchestrator 
+created by Google able to create a software abstraction from the underlying 
+virtualization technology. This enables Kubernetes to create an integration 
+with the Docker technology. Kubernetes has a architecture based on plug-ins, 
+giving  the possibility to the users to change part of Kubernetes itself, such 
+as the networking routing strategy or the persistence layer. With this 
+flexibility Kubernetes offered to us what Docker Swarm was not able to give. 
+Additionally, we had the possibility to set custom Domain Name System (DNS) 
+lookups and distribute other internal components in order to avoid requests 
+bottleneck.
+
+\subsection{Kubernetes configuration}
+Kubernetes does not have a minimum number of nodes in order to properly 
+function. Nonetheless, we decided to use a four node cluster configured in the 
+following way: three slaves node (also known as \emph{minions}) and one master 
+node. The Kubernetes version installed was the 1.11, enabling us to exploit new 
+features like a reworked DNS lookup system, a better intergrated\todo{find our 
+1.11 new features}.

### 2018-11-26 res/sections/15-ArchitecturalImplementation.tex
Add k8s explanation

@@
 \subsection{Kubernetes configuration}
-Kubernetes does not have a minimum number of nodes in order to properly 
-function. Nonetheless, we decided to use a four node cluster configured in the 
-following way: three slaves node (also known as \emph{minions}) and one master 
-node. The Kubernetes version installed was the 1.11, enabling us to exploit new 
-features like a reworked DNS lookup system, a better intergrated\todo{find our 
-1.11 new features}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.25]{kubernetes-control-plane}
+  \caption[Kubernetes control plane schema]{Kubernetes control plane
+    schema~\cite{k8scp}. It is possible to denothe important components in the
+    Kubernetes Control plane. In this high-level component architecture the
+    master keeps an internal database, \emph{etcd}, the entry point for the API
+    Server and the Pod scheduler. The nodes instead contain the kubelet daemon,
+    the container runtime and the networking configuration.}
+  \label{chap:archimpl:sec:secondattempt:img:k8scp}
+\end{figure}
+
+Kubernetes does not have a minimum number of nodes in order to properly work.
+Nonetheless, we decided to use a four node cluster configured in the following
+way: three slaves nodes (also known as \emph{minions}) and one master node. The
+Kubernetes version installed was the 1.11, enabling us to exploit new features
+like a reworked DNS lookup system and new storage features\footnote{A full
+  kubernetes log release can be found at:
+  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
+We initially deployed from the internal Openstack of the University four nodes
+with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines. We
+performed the installation using the official Kubernetes Repositories. After the
+initial installation, we obtained a configuration similar to the one illustrated
+in Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
+possible to denote different components on the master node and on the minions,
+which will be described.
+
+\subsubsection{Master architecture}
+\begin{description}
+\item[API Server] Fist of all, the User interacts with an API Server located in
+  the master node. This APIs are developed using the Swagger tools, ensuring
+  consistency and uniformity. From the image, is possible to see that the API
+  Server communicates with all the main components of the architecture.
+\item[etcd] The \verb!etcd! component communicates via the gRCP protocol, a
+  efficient Remote Procedure Call (RPC) protocol to serve data to the API
+  Server. This component is a distributed key-value storage, and allows data to
+  be shared in a reliable way across different machines~\cite{etcddatamodel}.
+\item[Controller Manager] The Controller Manager is the main Kubernetes control
+  loop, and its role is to periodically check the system health and integrity.
+  It communicates with the API Server via the Protobuf protocol, a
+  language-neutral, platform-neutral and extensible data serialization
+  mechanism.
+\item[Scheduler] The Scheduler regulates the Pods distribution in the minions,
+  having a knowlege of the current cluster typologies and of the current
+  resources available. On top of that, is able to handle hardware constraints,
+  defined policy, and defined class of QoS.
+\end{description}
+
+All of the previously descripted components are hide beghind the API server,
+which is the only components the final user can really interact with. The
+protocol between the API server and the final user used is a common JSON data
+object interchange. To make the Kubernetes more human-friendly, Kubernetes comes
+with a command line interface (CLI) tool, called \verb!kubectl!, that allows to
+easily perform deployments and admin operations on the cluster without having a
+full knowlege of the official API.
+
+\subsubsection{Minions architecture}
+
+As referred in Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}, the API
+server interacts with the components located in the minion nodes. In the image
+we can denote different modules.
+\begin{description}
+\item[Kubelet] The \verb!kubelet! daemon is the core component that runs in
+  every minion, and \emph{de facto} acts like a local manager. It receives YAML
+  or JSON definition from the API Server of Pod specifications (PodSpec) that
+  have to be deployed in the node where is running. Then, its role is to ensure
+  that the Pod deployed are up and healthy. This excludes possible
+  intereferences with other virtualized deployments.
+\item[Container Runtime] This runtime does not have to necessarily be based on
+  Docker or a container-based technology, but it can be based on an hypervisor
+  one, e.g. the KuberVirt project\footnote{Additional information about this
+    project can be found at the following link: \url{https://kubevirt.io/}}. The
+  container runtime performs the real Pods deployments.
+
+  Communications between the container runtime and the Kubelet component are
+  performed by the Container Runtime Interface (CRI), that consists of a series
+  of specifications/requirements built on top of the Protobuf protocol API.
+  Additionally, the Container Runtime has to communicate with the underlying OS,
+  and this operation is performed using the Open Container Initiative (OCI)
+  specification, which seeks to provide a standard on how to interact and run
+  the container images on the OS.
+\item[Networking] This part aims to provide to che container runtime network
+  connectivity. The role of the networking module is to provide, over than the
+  standard Internet connectivity, the possibility for Pods to see each other as
+  a part of the same network, even if Pods could be dislocated in other minions
+  working between different subnets or completely different networks. On top of
+  that, dynamic port allocation has to be kept in consideration too. The
+  Kubernetes networking model focus on three fundamental
+  requirements~\cite{k8snetworkingwiki}:
+  \begin{itemize}
+  \item all containers have to communicate with all the other containers without
+    NAT;
+  \item all nodes have the possibility to communicate with all containers (and
+    vice-versa) without NAT;
+  \item the IP that a container sees itself as is the same IP that others see it
+    as.
+  \end{itemize}
+  It is important to note that the IP addresses are at the Pod scope and not at
+  the container one. Every Pod has in fact its own networking space. Containers
+  inside a Pod can reach each other using the \verb!localhost! address. In
+  Kubernetes, several networking implementation exists, like Flannel, Google
+  Copmute Engine (GCE), Kube-route, Project Calico~\cite{k8snetworkingwiki}. For
+  our deployment, we decided to use Flannel, that demonstrated to work without
+  any particular configuration and to be mature enough to handle big workloads.
+
+  The communications between this component and the Kubelet one is carried
+  through the Cloud Native Interface (CNI) protocol, that consists of a set
+  of specification and libraries for writing plugins to configure network
+  interfaces in linux containers~\cite{cnigithub}.
+\end{description}
+
+\subsection{Openbaton integration}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{Attempt2v1}
+  \caption[Second attempt components organization schema]{Second attempt
+    components organization schema. It is possible to see, differently from the
+    first attempt, how now Kubernetes covers the old Openstack role and act as
+    VIM, receiving YAML definitions in to the master node that will schedule the
+    VNF in one of its minions.}
+  \label{chap:archimpl:sec:secondattempt:img:schema1}
+\end{figure}
+
+After the Kubernetes configuration and installation, we started working for a
+viable intergation between Kubernetes and Openbaton. Our first idea is the one
+illustrated in Figure~\ref{chap:archimpl:sec:secondattempt:img:schema1}:
+Openbaton, that had the TOSCA definitions, had to somehow convert these
+definition in a suitable YAML for Kubernetes, that had in turn to elaborate it
+and schedule to one minion in the cluster. The monitoring and management would
+remain untouched, as we planned to create an image with the integrated EMS tools
+that provided a seamless integration with Docker container. With this setting,
+the only thing remained was to solve the issue of converting TOSCA definitions in
+YAML one.
+The first thing we started thinking about was how to perform the conversion.
+% TODO talk about the problems of mapping the TOSCA definition from Openbaton
+% with the Kubernetes elements (which we actually solved), and talk about how we
+% weren't able to develop a VNFM plugin for Openbaton, and describe how Harbor
+% born from it.

### 2018-11-26 res/sections/15-ArchitecturalImplementation.tex
Start second attempt - revision one

@@
 that provided a seamless integration with Docker container. With this setting,
-the only thing remained was to solve the issue of converting TOSCA definitions in
-YAML one.
-The first thing we started thinking about was how to perform the conversion.
+the only thing remained was to solve the issue of converting TOSCA definitions
+in YAML one. The first thing we started thinking about was how to perform the
+conversion. This topic was quite important because the TOSCA definition does not
+really consider the possibility to have an alternative virtualization system
+apart from the hypervisor one. This causes the TOSCA definition to be, in our
+opinion, too much focused on the resources usage and regarding granularity
+aspects such the deployment flavour and the kind of virtual link (concept that
+does not really exsits in container based systems) to instatiate. Nonetheless,
+we performed a mapping between most of the TOSCA keywords with some of the YAML
+definition, in a way that a TOSCA configuration file was able to generate a
+valid YAML deployment. After that, we started thinking instead who, in the
+current architecture, had the task to perform this operation. Initially we
+though that Kubernetes itself could have been a perfect candidate. From a
+received TOSCA, a Kubernetes component should have validated and translated the
+TOSCA definition to generate a YAML suitable to be finally submitted to the
+Kubernetes API Server. Another feasible approach could have been to create a
+Openbaton plugin in order to take the TOSCA to deploy and convert it before
+submitting it to Kubernetes. This solution had great advantages in terms of
+responsability separation: in this way only the MANO had to know, for a VIM
+based on Kubernetes, that a conversion needed to be performed. From a Kuberentes
+point-of-view, nothing would have changed, since all the configuration
+conversion would have been performed by Openbaton.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.7]{OpenbatonMANO}
+  \caption[Openbaton architecture components]{Openbaton architecture components,
+    taken from~\cite{openbatondocumentation}.}
+  \label{chap:archimpl:sec:secondattempt:img:openbatonMANO}
+\end{figure}
+
+We started inspecting the Openbaton framework, to identify possible modules
+where the conversion should have taken place. We identified first the VIM
+driver, that is the component that directly interacts with the VIM, and then the
+VNFM, because we discovered there were programmatic dipendences between the two
+Openbaton standard components. After we identified the VIM driver and the VNFM
+as components to be rewritten and to be adapted to the Kubernetes VIM, we
+studied the original components code. We discovered that regarding these
+components the documentation was very poor, especially regarding the code. On
+top of that, the program interfaces given by the develpers did not match our
+expectations: we thought to find interfaces indipendent from the type of the
+virtualization technology, instead, we found that they were strictly coupled
+with frameworks like Openstack or cloud vendors like Google Compute engine or
+Amazon AWS. These reasons, with the fact that after two weeks we were not able
+to get any working prototype of VIM driver or VNFM we gave up, searching for
+another solution.
+
+\section{Second attempt - revision one}
+
+The impossibility to develop a custom VIM driver and VNFM manager lead us to
+search for an alternative solution. Then, we started work directly on the VIM.
+The base approach was to create an API Server which in future an Openbaton
+plugin could contact and simply deploy TOSCA definitions, interacting with a set
+of RESTful API.
+
 % TODO talk about the problems of mapping the TOSCA definition from Openbaton

### 2018-11-26 res/sections/07-Introduction.tex
Update intro

@@
 
-Through the history of Internet, many factors changed and new innovations came,
-but at the end of the day a fundamental key remains: connection performances
-are, indeed, what the users really want. Today, connectivity plays an important
-role in everyone's life. Several decades brought solutions with enhanced wired
-and wireless connectivity, like Ethernet, Fast Ethernet, 2G, 3G, LTE, and better
-communication protocols such as a whole plethora of TCP standards in order to
-keep up with the continuous surging of network traffic. Internet Service
-Providers to be able to efficiently process the whole quantity of packets
-passing through the network, applies several optimization along the packet path.
-Until this day, these optimization have mostly been based on costly, in hardware
-solutions. In particular, these Network Functions are able to perform only
-generic packet elaboration, since they do not know the Quality of Service
-required by every connection. It is in this context that 5G proposes a new way
-to approach networking, suggesting the virtualization of this Network Functions,
-making their deployments faster and cheaper. On top of that, it proposes
-specific services for determined kind of data. If in the past a train running a
-300km/h had to use the same network services of an IoT device sending data
-updates every hour, now 5G proposes two different and efficient solutions for
-this completely different services. This separation is performed through the
-creation of determined network components called Network Slices. This new
-possibility of offering the network with specialized class of services, combined
-with emerging virtualization technologies like cloud computing,
-hypervisor-virtualized machine or container-based solutions enables the
-deployment of Virtualized Network Functions within seconds running on generic
-purpose or custom hardware. While there may be some performances degradation,
-it has been demonstrated that their time-to-deploy and the possibility of
-automation they offers are unprecedented~\cite{nguyen2017sdn}. On top of that,
-specialized software acts as a MANager and Orchestrator (MANO) for this VNFs,
-regulating their lifecycle. While most of the current proposals are based on
-hypervisor on cloud solutions, in this thesis we are going to see a
-proof-of-concept of a part of a Network Slice, equipped with a custom MANO
-solution, based on the new container-based technology, called Docker, on a
-recent orchestrator framework: Kubernetes.
+Through its history, the Internet has undergone many, rather slow, innovation
+cycles, each bringing into the light its unique and important role in human
+societal evolution. The Internet has effectively become a utility, embedded into
+our everyday fabric which is more than often taken for granted. Connectivity
+plays an important role and service continuity and performance is what the users
+really strive for.
+
+Several decades of research and engineering effort have broadened the
+possibilities with which users access the network i.e., enhanced wired and
+wireless connectivity, like Ethernet, Fast Ethernet, 2G, 3G, LTE, while enjoying
+their preferred services. This pace of innovation is matched by a similar
+evolution of the TCP/IP protocol standards, trying to keep up with the
+continuous surge of network traffic. On their side, Internet Service Providers
+(ISPs) rely on ad-hoc optimizations of the end-to-end service delivery path. To
+this date, optimization is made in terms of physically placing middlebox
+devices, carrying out specialized network functions, on the delivery path. Deep
+packet inspection, firewalling, ad-hoc transport optimization are some examples
+of such functionalities deployed either at aggregation or core-network points.
+
+This modus operandi, however, is neither scalable nor cost-effective.
+Maintenance operational costs, careful planning and dedicated costly hardware
+are some of the side-effects. Yet, another more subtle symptom is that of
+coupling software lifecycle to the hardware (middlebox) hindering innovation.
+Thanks to recent developments, new possibilities have arisen. In particular, the
+trend in virtualization and softwarization techniques have open the door to the
+so-called network softwarization, allowing for a potentially faster pace of
+innovation in the network-side by exploiting the software as a driving force.
+
+The fifth-generation public mobile network architecture, commonly referred to as
+5G, has been designed to support highly reliable high-performance communications
+to satisfy the different requirements of all kinds of applications and services.
+To achieve such flexibility, 5G assigns a central role to “network
+softwarization”, i.e. the ability of the network to operate according to rules
+and policies that may be controlled by software entities. By leveraging on
+Software Defined Networking (SDN) and Network Function Virtualization (NFV), the
+5G architecture heavily relies on software entities whose behaviour may be
+easily customized to adapt to the requirements of different applications and
+tenants. The 5G vision is centered around the idea of building tenant- or
+service-specific networks on top of a collection of heterogeneous shared
+physical infrastructures. This concept is referred to as network slicing.
+Slicing is a key enabler for making the network able to support applications
+with a wide range of different requirements, including:
+\begin{enumerate*}[label=(\roman*)]
+\item enhanced mobile broadband (eMBB);
+\item ultra-reliable and low-latency communications (URLLC) and
+\item massive machine-type communications (mMTC).
+\end{enumerate*}
+
+This new possibility of provisioning a network with specialized, heterogeneous
+classes of services, combined with emerging virtualization technologies like
+cloud computing, hypervisor-virtualized machine or container-based solutions,
+enables the deployment of Virtualized Network Functions (VNFs) within seconds
+running on generic purpose or custom hardware. While there may be some
+performances degradation, it has been demonstrated that their time-to-deploy and
+the possibility of automation they offer are unprecedented~\cite{nguyen2017sdn}.
+On top of that, a specialized software entity acts as a MANager and Orchestrator
+(MANO) for these VNFs, managing their lifecycle. While most of the current
+proposals are based on a hypervisor solution, in this thesis we set the
+objective of providing a proof-of-concept implementation of an Infrastructure
+Management and Orchestration plane based on the new container-based technology,
+called Docker, on a recent orchestrator framework: Kubernetes. Moreover,
+building on recent published results of the 3GPP Release 15, we outline a
+software architecture capable of end-to-end service orchestration exploiting
+container-technology.
 

### 2018-11-26 res/sections/09-RelatedWork.tex

@@
 Section~\ref{chap:prjan:sec:tech:sub:other:sub:floodlight}.
-%TODO section missing...write it down!!
 
@@
 container-based approach.
-
-\subsection{SFC implementation}
-Since the SFC technology was proposed many work came out proposing different
-implementation solutions. In this part there is an overview of some of them.
-
-\cite{GhaznaviSAB16} proposed an optimization model to calculate the optimal
-location of VNFs based on different metrics such as bandwidth, load balancing
-and routes. This decomposes the SFC on performance requirements eliminating the
-throughput bound of a chain to a single VNF or a single physical machine.
-Mathematical models developed was tested in a simulation environment. A critical
-enhancement to this work could be to test algorithm in a real environment, under
-different conditions of traffic load: if it efficient even in this scenario it
-can be useful to the orchestration and management part of the ecosystem because
-it can reduce resource waste.
-
-Other implementation tested in a simulated environment
-are~\cite{csoma2014escape} and~\cite{kim2016evaluations}. In both of them the
-network functions are deployed on Mininet. The former work provides a full
-implementation of the SFC environment, including SFC configuration, allocating
-VNFs into the physical resources, flow routing across the VNFs based on
-policies, and provisioning live management information on operating VNF
-instances, but no performance evaluation of the work is presented. In the latter
-case chains are identified by VLAN tag and the VNF chaining is created
-exploiting only using Floodlight controllers.
-
-Another noteworthy work is~\cite{xia2015optical} that discuss the packet
-delivery through function chaining in high capacity networks exploiting
-aggregate flow steering. The architecture proposed makes use of SDN and cloud
-platforms to deploy VMs that will host VNFs. Assumption are made about the SDN
-network elements: connection through optical circuit switching and use of Wave
-Selective Switching (WSS) to route (switch) signals between optical fibers on a
-per-wavelength basis. Thanks to the implementation of the SDN controller and the
-communication between the previous element and the VNFM, this architecture is
-able to provide good capabilities in terms of flexibility. Another element that
-helps the flexibility is the high performance link: in this way they get rid of
-problems in terms of connection time and bandwidth. However, in the paper
-authors assumes really specific hardware in the underlying infrastructure that
-connects the overall system: this is not suitable to be deployed in a general
-purpose infrastructure or using IaaS. It will be interesting to evaluate the
-same work on different instrumentation. A clashing aspect is the time to make a
-VM up and running compared to times of service chaining configuration/packet
-delivery. Using containers the instantiation of virtual functions should employ
-less time.
-
-A relevant work on function chaining in network is~\cite{trajkovska2017sdn}. In
-this work author efforts are concentrated on joining SDN and VNF technologies to
-create SFCs. They developed an SDN-driven SDK, a virtual traffic classifier to
-analyze traffic in real-time and decide the optimal chain for it and a VNF
-specialized on video transdecoding, used event to evaluate the performances of
-the whole system. Chaining is performed exploiting layer 2 connectivity, using
-OpenvSwitch as virtual network switches an OpenDayLight in order to route the
-traffic on a set of policies. However the solution does not provide any
-mechanism to perform an automatic deployment of chains and services: this can be
-an issue in particular in case of VNF fault. Orchestration and some mechanisms
-of fault tolerance are critical in particular in a real environment, enabling
-autoscaling on user demand or traffic load. Automated deployment also make
-possible the recreation of a virtual machine in case of fault. As a matter of
-fact that a virtual function can fail depending on an erroneous load balancing
-policy or too heavy computational load and it is critical in order to do not
-interrupt end-to-end communication to bring up the VNF as soon as possible. This
-recreation process can be slow due to the usage of VMs to deploy the functions,
-that could take minutes to startup.
-
-The paper~\cite{kriti2017dnfc} introduce the possibility using container in SFC
-deployment. The chain of container is created exploiting Openflow to create flow
-to route traffic to the containers, that have to be installed on the host
-machine. This work however does present any form of management or orchestration
-of the chain, neither an automatic deployment. Moreover it requires installation
-of part of the system on the host machine: in this way there is not a fully
-virtualized environment. The former issue has some implication: the VNF has a
-strong relation on the host on which is running and the system could not be
-simply be moved from one machine to another as it could be done only using
-containers. Finally this paper does not provide any form of system evaluation
-neither in terms of performances nor in terms of resilience. Fault recovering
-and autoscaling could be achieved only improving the system introducing other
-technologies, such as Docker Compose, Kubernetes or Docker Swarm.

### 2018-11-26 res/sections/10-ProjectAnalysis.tex

@@
 
-\subsection{Packet transmission strategies}
-
-When a packets are sent from a sender $S$, to a receiver $R$, they start
-travelling to the backbone receiving appropriate in-hardware packet processing
-before being received by $R$. During this elaboration, it is important to
-maintain the transparency of the connection, so that the network users are not
-aware of the packets elaboration at all. To perform that, many solutions are
-available: TUN/TAP packet encapsulation, IP Encapsulation within IP or TCP
-session recreation (know also as TCP split).
-
-TUN/TAP and IP Encapsulation within IP (initially defined in RFC 2003 with the
-aim to deliver IP packets over Mobile IP for mobile networks) maintain the
-connection transparency encapsulating the packet in another one.
-\begin{figure}[t]
-  \centering \includegraphics[scale=0.5]{IPoverIP}
-  \caption[IP Encapsulation within IP packet schema]{A schema of IP
-    Encapsulation within IP. Is noticeable how the original packet becomes the
-    payload of a new one, thus being encapsulated. All the information
-    regarding the original packet remains untouched: when this kind of packet
-    gets send from a machine to another, the OS will remove the new packet
-    header and will present to the user-space program receiving the data its
-    payload: a IP packet. Packets encapsulation can be applied with higher
-    layers too, thus allowing for TCP and UDP packet encapsulation.}
-  \label{chap:prjan:img:ip_over_ip}
-\end{figure}
-
-In this way, the original packet remains preserved and it becomes the new packet
-payload, that can be elaborated and modified accordingly, until it gets
-decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is not
-be aware of the packet elaboration, since all the original headers were
-preserved (or slightly modified). In particular, IP Encapsulation within IP
-allow the packets to make intermediate destinations that otherwise would not be
-selected. TCP session recreation, instead, ``split'' the TCP session in two
-endpoints: an Ingress, that manage the connection with the client and an Egress,
-that communicate with the packet receiver creating a multi-overlay-hop path
-where for each hop there is an possible independent TCP connection. In order to
-achieve that, a session table needs to be established and continuously updated,
-while TCP headers need to be accordingly modified to allow data transmission
-through the Ingress and Egress points. Nonetheless, care should be given to
-packets path: in order to avoid ill-calculated congestion windows, the incoming
-packets need to follow the same (or an equivalent one) path for the same
-instantiated connection. This phenomenon is based on the intermediate proxies
-that create the route: it is the case, for example, when an intermediate node
-sends back a spoofed ACK before the original packet is delivered to the real
-consignee. In this scenario, the sender, unaware of the proxy, will miscalculate
-the sending window based on the too low RTT. Other problems of TCP splitting are
-reliability and security, especially because the end to end TCP logic is no
-longer maintained: a server failure may cause to an unaware client to believe
-that all the packets have reached the destination. TCP session recreation seems
-burdensome and it doesn't seem to add any significant benefit at the first
-sight, but it offers the flexibility to perform extensive packet manipulation:
-since the packet its recreated every time at the Ingress and Egress points,
-while in the Network Function Virtualization Infrastructure (NFVI), the original
-packet can be stripped of its headers (that could be saved in the session table)
-to slightly improve internal packet transmission. \todo{Insert TCP session
-  recreation schema} On top of that, it offers the possibility to use different
-TCP flavours, hence increasing the communication speed in heterogeneous networks
-(e.g. in case of satellite connections TCP Hybla can be employed, while it is
-known that HighSpeed TCP gives best results when applied in optical fiber-based
-connections).
-
-During the test-bed development we first tried the TUN/TAP solution, that
-revealed to be not completely transparent and with major performance drawbacks
-(i.e. packet transmission in a LAN connection reached latencies of 50ms)
-\todo{Explain the constraint of point-to-point connections, while we need
-  dynamic SFC)}, so we choose to opt for a TCP session recreation solution,
-implementing a little back-end with the aim of saving in a persistent storage
-the SFC the packet has to perform and additional metadata.
-
-\section{Available technologies in the market}
+\section{Available technologies in the market} \todo{Add images}
 \label{chap:prjan:sec:tech}

### 2018-11-26 res/sections/25-TestsResponsiveness.tex

@@
 \chapter{Tests and metrics}
-
+\label{chap:test}
 After the test-bed implementation, tests about its performance and metrics were

### 2018-11-26 res/sections/15-ArchitecturalImplementation.tex
Fix typos

@@
 The impossibility to develop a custom VIM driver and VNFM manager lead us to
-search for an alternative solution. Then, we started work directly on the VIM.
-The base approach was to create an API Server which in future an Openbaton
+search for an alternative solution. Then, we started to work directly on the
+VIM. The base approach was to create an API Server which in future an Openbaton
 plugin could contact and simply deploy TOSCA definitions, interacting with a set
-of RESTful API.
+of RESTful APIs.
 

### 2018-11-27 res/sections/10-ProjectAnalysis.tex
Move Harbor section from 10 to 15. Add Harbor schema

@@
 
-\subsubsection{Harbor}
-\label{chap:prjan:sec:tech:sub:MANO:sub:harbor}
-Started as a VNFM (Virtual Network Function Virtualization Manager) to offer a
-bridge for Openbaton to interact with Kubernetes, during the project it has been
-expanded to become a complete replacement of the original product, adding the
-possibility to manage Network Services and to update the routing tables
-contained in the Roulette Controller (the tool description can be found below
-in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web
-interface, but it has a RESTful API that allows users to interact with it using
-CRUD operations. It offers functionalities similar to Openbaton: possibility to
-save and retrieve VNF and NS definitions (basically, Kubernetes YAML
-deployments). When an NS is saved, a consistency check is performed to assure
-that all the VNFs have already been defined in Harbor. On an NS deployment, with
-a mechanism similar to the Openbaton one, all the associated VNFs are set up
-too, making the whole operation of launching a NS atomic from the user
-point-of-view. What Openbaton doesn't do is a smart resource utilization:
-Harbor, in fact, during a NS deletion, it checks if any related VNF remains
-orphan (which means without any NS using it). In that case the VNF will enter in
-a state of pruning: after five minutes, if the VNF has not be re-utilized by any
-other NS it will be deleted. This mechanism assure faster deployments in
-environments where NSs are continuously taken up and down. \todo{I've made an
-  Harbor implementation schema, I dunno if its useful to put it here}
-
 \subsection{SFC}
@@
 submitted by another component, that we identified as Harbor (component
-description available above in
-sub-section~\ref{chap:prjan:sec:tech:sub:MANO:sub:harbor}). After a NS
+description available in
+section~\ref{chap:archimpl:sec:secondattempt:sub:harbor}). After a NS
 deployment, Roulette gets instructed with a new route describing a NS. Every NS

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Move Harbor section from 10 to 15. Add Harbor schema

@@
 plugin could contact and simply deploy TOSCA definitions, interacting with a set
-of RESTful APIs.
+of RESTful APIs. We called this simple backend \emph{Harbor}.
+
+\subsubsection{Harbor}
+\label{chap:archimpl:sec:secondattempt:sub:harbor}
+Started to offer a bridge between Openbaton and Kubernetes, during the project 
+it has been expanded to become a complete replacement of the original product, 
+adding the possibility to manage Network Services and to update the routing 
+tables contained in the Roulette Controller (the tool description can be found 
+in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web 
+interface, but as already written it has a RESTful API that allows users to 
+interact with it using CRUD operations. It offers functionalities similar to 
+Openbaton: possibility to save and retrieve VNF and SFC definitions (basically, 
+Kubernetes YAML deployments). When an SFC is saved, a consistency check is 
+performed to assure that all the VNFs have already been defined in Harbor. On 
+an SFC deployment, with a mechanism similar to the Openbaton one, all the 
+associated VNFs are set up too, making the whole operation of launching a SFC 
+atomic from the user point-of-view. What Openbaton doesn't do is a smart 
+resource utilization: Harbor, in fact, during an SFC deletion, it checks if any 
+related VNF remains orphan (which means without any SFC using it). In that case 
+the VNF will enter in a state of pruning: after five minutes, if the VNF has 
+not be re-utilized by any other SFC it will be deleted. This mechanism assure 
+faster deployments in environments where NSs are continuously taken up and 
+down.
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.5]{Harbor_architecture}
+ \caption[Harbor architectural schema]{Harbor architectural schema}
+ \label{chap:archimpl:sec:secondattempt:img:harborarchitecture}
+\end{figure}
 

### 2018-11-27 res/sections/08-Background.tex
Add final part about arch impl

@@
 virtual machines with different configurations, without bringing any changes to
-the host machine. Indeed, the virtualizations concepts refer to the property of
+the host machine. Indeed, the virtualization concepts refer to the property of
 a machine to be virtualized. To achieve this, the components that forms a common
@@
 \section{Service Function Chaining (SFC)}
+\label{chap:background:sec:sfc}
 Another important component we have to talk about is the SFC, defined as a

### 2018-11-27 res/sections/10-ProjectAnalysis.tex
Add final part about arch impl

@@
 
-\subsubsection{Roulette Controller}
-\label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
-Dynamic communications between VNFs would not be possible without a third
-component acting as a route controller. Roulette fulfills this role, providing a
-back-end where the single VNF can ask the next packet hop. Indeed, routes must 
-be
-submitted by another component, that we identified as Harbor (component
-description available in
-section~\ref{chap:archimpl:sec:secondattempt:sub:harbor}). After a NS
-deployment, Roulette gets instructed with a new route describing a NS. Every NS
-has a proper ID (called \verb!sfid!), and thus when the classifier checks a
-packet typology, it labels that packet with the specific NS ID (for a detailed
-description about the SFC header see Section~\todo{Add section/chapter/paragraph
-  ref!}). Finally, thanks to this label, every VNF knows the packet's chain and
-it will correctly forward the it to the next VNF.
-
-\subsubsection{Astaire}
-As is possible to denote, a VNF doesn't only process the incoming packets, 
-but it has to manage the communications and the related forward operations. To 
-possibly ease part of the transmission problem and to make it easier to create 
-a 
-VNF without requiring the whole knowledge about SFC, it has been decided to 
-build a framework, Astaire, that could keep the SFC related details decoupled 
-from the core of the VNF, in other words, the entity that actually elaborates 
-the packet. Built using an event-based pattern, it presents hooks that allow 
-other software written in different languages from the Astaire one (e.g. Java), 
-to be called dynamically at runtime in order to process the packet. The result 
-of this elaboration gets returned to Astaire, that performs the beforehand 
-cited operations of VNFs communications and forwarding.
-
-\subsubsection{Ironhide}
-As described in chapter~\ref{chap:archimpl}, the whole framework need to have 
-and entry and an exit point. These, defined as \textit{Ingress} and 
-\textit{Egress} points, are performed by Ironhide. Ironhide acts like a proxy, 
-hiding the whole packet processing to the sender $S$ and the receiver $R$. When 
-a client $S$ performs a request to $R$, the packet actually goes through 
-Ironhide, who has an internal classifier that it is able to recognize the type 
-of connection and to choose the right NS the packet has to go through. This 
-process in fundamental since it keeps the connection between $S$ and $R$ up, 
-for both TCP and UDP based communications.
-
 \subsection{Technologies used during development}

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Add final part about arch impl

--- 2018-11-26 res/sections/15-ArchitecturalImplementation.tex
@@
 plugin could contact and simply deploy TOSCA definitions, interacting with a set
-of RESTful APIs.
+of RESTful APIs. We called this simple backend \emph{Harbor}.
 
-% TODO talk about the problems of mapping the TOSCA definition from Openbaton
-% with the Kubernetes elements (which we actually solved), and talk about how we
-% weren't able to develop a VNFM plugin for Openbaton, and describe how Harbor
-% born from it.
+\subsection{Harbor}
+\label{chap:archimpl:sec:secondattempt:sub:harbor}
+Harbor grew considerably during the project development, as it revealed to be a
+key component of our architecture. Harbor does not have a web interface, but as
+already written it has a RESTful API that allows users to interact with it using
+CRUD operations. It offers functionalities similar to Openbaton: possibility to
+save and retrieve VNFs and SFCs definitions. When an SFC is saved, a consistency
+check is performed to assure that all the VNFs have already been defined in
+Harbor. On an SFC deployment, with a mechanism similar to the Openbaton one, all
+the associated VNFs are set up too, making the whole operation of launching a
+SFC atomic from the user point-of-view. Since this architecture component grew
+until it became a Kubernetes MANO, we added a feature that, in our opinion,
+Openbaton lacks: smart resource utilization. Harbor, in fact, during an SFC
+deletion, it checks if any related VNF remains orphan (which means without any
+SFC using it). In that case the VNF will enter in a state of pruning: after five
+minutes, if the VNF has not be re-utilized by any other SFC it will be deleted.
+This mechanism assure faster deployments in environments where NSs are
+continuously taken up and down.
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.5]{Harbor_architecture}
+ \caption[Harbor architectural schema]{Harbor architectural schema}
+ \label{chap:archimpl:sec:secondattempt:img:harborarchitecture}
+\end{figure}
+
+\subsection{The need for additional components}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{Attempt2v1Harbor}
+  \caption[Second attempt revision one schema]{Second attempt revision one
+    schema. It is possible to see how Harbor completely substituted Openbaton.
+    Additionally, Harbor now operates directly with the Kubernetes API,
+    retrieving informations about the VNFs status from it. This causes the VNFs
+    to be lighter because they don't have to include any reporting tools, like
+    the EMS, in it.}
+  \label{chap:archimple:sec:secondattempt:img:attempt2v1harbor}
+\end{figure}
+
+Terminated the Harbor development, we ended up in a situation similar the one
+depicted in Figure~\ref{chap:archimple:sec:secondattempt:img:attempt2v1harbor}.
+It is possible to see how the Openbaton framework is not necessary anymore:
+Harbor, in fact, ended up substituting it. We then started to think about the
+challenges of integrating the SFC structure inside a Kubernetes cluster. We
+identified the following challenges:
+\begin{itemize}
+\item whole system scalability: we wanted that every component had the
+  possibility to scale during intense workloads, without the necessity of human
+  intervention.
+\item ingress and egress points: there should be at least two points in the
+  cluster were the packect enters and exits;
+\item dynamic package routing: as formulated by one of our requirements, packet
+  routing had to be dynamic. We immediatly identified a possibility to create
+  this functionality with the already-built features of Kubernetes;
+\item a framework to automatically forward the data elaborated by the VNF: we
+  identified the possibility to build a framework that automatically handled the
+  packet receiving from another VNF, it properly called the function that had to
+  elaborate the packet payload and then it forwarded the result to the next VNF;
+\item connection status: having ingress and egress points implicates the
+  necessity to save some persistent data in a distributed manner.
+\end{itemize}
+
+\subsubsection{Scalability considerations}
+When adopting Kubernetes, we realized the possibility to make every component
+easily scalable. Indeed, all the following components can be described in
+Kubernetes as a \verb!Service!. For a \verb!Service! a \verb!ReplicaSet! and a
+consequent \verb!Deployment! can be associated, giving the possibility for it to
+scale under heavy workloads. The introduction of scalability meant that was not
+possible to refer to the components based on their IP, but we had to refer to
+their service name. As it will be explained later, this introduced challenges
+especially for the ingress and egress points of our network, that have to keep a
+connection with the sender and the receiver. On top of that, the scalability and
+the required DNS lookup meant a possible bottleneck. With a great number of
+components and VNFs constantly quering the centralized Kubernetes DNS lookup
+system a performance degradation could have easily happened. We decided to
+distribute the DNS service too, integrating the CoreDNS Kubernetes default
+service with \emph{DNSMasq}, a DNS server able to perform query caching for
+faster name resolutions. We created a DNSMasq \verb!DaemonSets!, which in turn
+deployed one Pod with DNSMasq running inside for every minion in the cluster. At
+the time of the VNF launch, we set that the DNS resolution should pass first for
+the local \verb!DNSMasq! DNS, and then for the Kubernetes CoreDNS service. In
+this way, we were able to distribute the DNS query workload from the master,
+using a local caching system to speed up the name resolution between services.
+
+\subsubsection{Ingress and Egress: Ironhide}
+As just said, the whole framework needs to have and entry and an exit point.
+These, defined as \emph{Ingress} and \emph{Egress} points, are performed by
+\emph{Ironhide}. Ironhide acts like a hides the whole packet processing to the
+sender $S$ and the receiver $R$. When a client $S$ performs a request to $R$,
+the packet actually goes through Ironhide, who has an internal classifier that
+it is able to recognize the type of connection and to choose the right NS the
+packet has to go through. This process is fundamental since it keeps the
+connection between $S$ and $R$ up, for both TCP and UDP based communications.
+Since the connection need to be end-to-end, the state of the connections are
+saved in the \emph{Roulette Controller} component.
+
+\subsubsection{Dyamic package routing: the Roulette component}
+\label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
+Dynamic end-to-end communications between VNFs would not be possible without a
+third component acting as a route controller. Roulette fulfills this role,
+providing a back-end where the single VNF can ask the next packet hop. Indeed,
+routes must be submitted by another component, that we identified as Harbor.
+After a NS deployment, Roulette gets instructed with a new route describing a
+NS. Every NS has a proper ID (called \verb!sfid!), and thus when the classifier
+checks a packet typology, it labels that packet with the specific NS ID (for a
+detailed description about the SFC header see
+Section~\ref{chap:background:sec:sfc}). Finally, thanks to this label, every VNF
+knows the packet's chain and it will correctly forward the it to the next VNF.
+
+\subsubsection{Keeping the connection status}
+The scalability challenge introduces new issues in the system. The most
+important is regarding keeping up stateful connections, since the ingress and
+egress points could be more than one container. With this considerations in
+mind, we started thinking at how and where keeping information about the ingress
+container handling the connection with the sender $S$ and about the egress
+container talking with the receiver $R$. Since the Roulette component has a
+complete end-to-end topology view of the system, we decided to store these
+additional informations (we called it \emph{endpoints}) in it.
+
+\subsubsection{A VNF framework: Astaire}
+As is possible to denote, a VNF doesn't only process the incoming packets, but
+it has to manage the communications and the related forward operations. To
+possibly ease part of the transmission problems and to make it easier to create
+a VNF without requiring the whole knowledge about SFC, we decided to build a
+framework, Astaire, that could kept the SFC related details decoupled from the
+core of the VNF, in other words, the entity that actually elaborates the packet.
+Built using an event-based pattern, it presents hooks that allow other software
+written in different languages from the Astaire one (e.g. Java), to be called
+dynamically at runtime in order to process the packet. The result of this
+elaboration gets returned to Astaire, that performs the beforehand cited
+operations of VNFs communications and forwarding. In the SFC specification,
+Astaire is defined as a Service Function Forwarder (SFF).
+
+\subsection{Final architecture structure}
+At the end, we were able to create a scalable architecture that followed the
+ETSI MANO standards, with Kubernetes as VIM and Harbor has MANO. We faced
+additional challenges regarding the system scalability, that we solved using the
+Roulette controller. We eased the creation of VNFs with Astaire, and we created
+proper ingress and egress points with Ironhide, that supported scalability
+issues and performed a protocol-based packet classification. Is possible to see
+a final architectural schema in
+Figure~\ref{chap:archimpl:sec:secondattempt:img:attempt2v1total}.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.48]{Attempt2v1Total}
+  \caption[Final components organization]{Final components organization. In the
+    image is possible to denote how Egress, Ingress, the Roulette Controller and
+    the VNFs instances are optional or could not be present in the minion, while
+    DNSMasq is always deployed.}
+  \label{chap:archimpl:sec:secondattempt:img:attempt2v1total}
+\end{figure}

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Add images as required

@@
 changes we had to perform.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{Architecture}
+  \caption[Architecture schema depicting the expected
+    implementation]{Architecture schema depicting the expected implementation.
+    It is possible to see how Openbaton interacts with the various components in
+    the architecture, and how RabbitMQ is used as a message broker between the
+    different parts to enable message exchanging.}
+\end{figure}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
@@
 \paragraph*{Taker}
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 easily scalable. Indeed, all the following components can be described in
-Kubernetes as a \verb!Service!. For a \verb!Service! a \verb!ReplicaSet! and a
-consequent \verb!Deployment! can be associated, giving the possibility for it to
-scale under heavy workloads. The introduction of scalability meant that was not
+Kubernetes as a \verb!Service!. For a \verb!Service! a consequent
+\verb!Deployment! can be associated, giving the possibility for it to scale
+under heavy workloads. The introduction of scalability meant that was not
 possible to refer to the components based on their IP, but we had to refer to

### 2018-11-27 res/sections/08-Background.tex
Add image as required

--- 2018-11-24 res/sections/08-Background.tex
@@
 
-\section{The VIBeS project}
-This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
-requirement is to enhance the performance of end-to-end IP based that involves
-satellite connections. To reach this goal, the project specifications suggest to
-exploit the 5G incoming technology and use the NFV-MANO architecture to perform
-first packets elaboration and performance improvement and finally TCP/IP
-satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
-VIBES project proposed five technical requirements:
-\begin{enumerate}
-  \item Analysis of the applicability of current and new Internet protocols in
-  the proposed VNF-PEP architecture
-  \item Implementation of a VNF-PEP prototype
-  \item Building of a PoC test platform
-  \item VNF-PEP validation and performance tests on 5G use-cases
-  \item Demonstration test-bed management
-\end{enumerate}
+\section{The 5G connectivity standard}
 
@@
   \centering
- \includegraphics[scale=1]{vibes_logo}
- \caption{ViBeS project logo}
- \label{chap:background:img:vibes_logo}
+  \includegraphics[scale=0.8]{5GEnablers}
+  \caption[5G enablers]{5G enablers. It is possible to see that the NFV/SDN
+    approach is considered an enabler for the challenge regarding a massive
+    number of connections~\cite{agyapong2014design}.}
+  \label{chap:background:sec:5G:img:enablers}
 \end{figure}
 
-\subsection{VNF-PEP architecture and internet protocols}
-VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
-the project is to research and include this technology in the usual VNF
-ecosystem to include without change satellite communications to the SFC. The
-architecture of the systems will be based on clusters of containers whose
-management and orchestration will follows the ETSI MANO architecture and SDN
-principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
-how to optimize a certain communication.
-
-%The analysis of this topic revealed to be trivial: since internet has many
-%different protocols that would become infeasible to support all of them at the
-%same time, packet encapsulation present itself as the only feasible solution:
-%every packet incoming in the VNFs has already been encapsulated by a generic
-%packet encapsulator/decapsulator,
-%\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
-%the protocol used on the edges but 1) use a ``quick protocol to exchange data 
-%among VNFs and 2) we are hiding the path not the protocol itself. As we were 
-%discussing, we are creating some sort of proxy, so the aim will be the same 
-%even if we support a plethora of protocols.''} making the whole architecture 
-%independent from the protocol a particular flux of data. To achieve this, 
-%several solutions have been studied, and at the end packet encapsulation with 
-%TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
-%this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
-%reference with the section of the explanation}
-%
-%\subsection{VNF-PEP prototype}
-%
-%Since the requirement for the whole system (MANO+PEP) were too challenging, the 
-%goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
-%The VNF architecture was shaped following the container orchestrator we decided 
-%to use. An more detailed architectural implementation can be found at 
-%Chapter~\ref{chap:archimpl}.
-%
-%\vspace{0.5cm}
-%
-%Starting with the first step of creating a MANO able to process incoming data
-%packets through VNF functions, we encountered that many networking tools already
-%present in the market required some tweaking and some integration, shifting our
-%goal to create a complete European Telecommunications Standards Institutes
-%(ETSI) Management and orchestrator (MANO) test-bed instead, following the
-%specifications suggested in the RFC 7665, thus implementing only the first three
-%requisites, without digging in the satellite data flow optimization. In
-%particular, we discovered how, these tools, were suitable to create ETSI MANO
-%and VNFs using virtual machine or exploiting cloud technologies, while they were
-%not designed with enough flexibility an integration with Docker. In the next
-%chapter we are going to take a deep analysis of the cited technologies 
-%(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
-%context to be able to understand our design choices we are going to describe 
-%later.
-%
-%\noindent With this in mind, we performed a requirements analysis described in
-%the next chapter.\todo{This should be put at the end of the introduction.}
-%
-
-\section{5G}
 The continuous innovations in the mobile network connectivity is leading to the
@@
 
+\section{The VIBeS project}
+This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
+requirement is to enhance the performance of end-to-end IP based that involves
+satellite connections. To reach this goal, the project specifications suggest to
+exploit the 5G incoming technology and use the NFV-MANO architecture to perform
+first packets elaboration and performance improvement and finally TCP/IP
+satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
+VIBES project proposed five technical requirements:
+\begin{enumerate}
+  \item Analysis of the applicability of current and new Internet protocols in
+  the proposed VNF-PEP architecture
+  \item Implementation of a VNF-PEP prototype
+  \item Building of a PoC test platform
+  \item VNF-PEP validation and performance tests on 5G use-cases
+  \item Demonstration test-bed management
+\end{enumerate}
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=1]{vibes_logo}
+ \caption{ViBeS project logo}
+ \label{chap:background:img:vibes_logo}
+\end{figure}
+
+\subsection{VNF-PEP architecture and internet protocols}
+VNF-PEP is the softwarization of the usual PEP technology. One of the goals of
+the project is to research and include this technology in the usual VNF
+ecosystem to include without change satellite communications to the SFC. The
+architecture of the systems will be based on clusters of containers whose
+management and orchestration will follows the ETSI MANO architecture and SDN
+principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
+how to optimize a certain communication.
+
+%The analysis of this topic revealed to be trivial: since internet has many
+%different protocols that would become infeasible to support all of them at the
+%same time, packet encapsulation present itself as the only feasible solution:
+%every packet incoming in the VNFs has already been encapsulated by a generic
+%packet encapsulator/decapsulator,
+%\todo{Also I think that the main purpose of UDP encapsulation is not to hide 
+%the protocol used on the edges but 1) use a ``quick protocol to exchange data 
+%among VNFs and 2) we are hiding the path not the protocol itself. As we were 
+%discussing, we are creating some sort of proxy, so the aim will be the same 
+%even if we support a plethora of protocols.''} making the whole architecture 
+%independent from the protocol a particular flux of data. To achieve this, 
+%several solutions have been studied, and at the end packet encapsulation with 
+%TCP split (for TCP sessions) have been chosen. The rationale that guided us on 
+%this choice is described in Chapter~\ref{chap:vnf_ns_impl}. \todo{Update 
+%reference with the section of the explanation}
+%
+%\subsection{VNF-PEP prototype}
+%
+%Since the requirement for the whole system (MANO+PEP) were too challenging, the 
+%goal shifted into creating a MANO test-bed and a working NFVI, excluding PEP. 
+%The VNF architecture was shaped following the container orchestrator we decided 
+%to use. An more detailed architectural implementation can be found at 
+%Chapter~\ref{chap:archimpl}.
+%
+%\vspace{0.5cm}
+%
+%Starting with the first step of creating a MANO able to process incoming data
+%packets through VNF functions, we encountered that many networking tools already
+%present in the market required some tweaking and some integration, shifting our
+%goal to create a complete European Telecommunications Standards Institutes
+%(ETSI) Management and orchestrator (MANO) test-bed instead, following the
+%specifications suggested in the RFC 7665, thus implementing only the first three
+%requisites, without digging in the satellite data flow optimization. In
+%particular, we discovered how, these tools, were suitable to create ETSI MANO
+%and VNFs using virtual machine or exploiting cloud technologies, while they were
+%not designed with enough flexibility an integration with Docker. In the next
+%chapter we are going to take a deep analysis of the cited technologies 
+%(Section~\ref{chap:prjan:sec:tech}, in order to provide to the reader enough 
+%context to be able to understand our design choices we are going to describe 
+%later.
+%
+%\noindent With this in mind, we performed a requirements analysis described in
+%the next chapter.\todo{This should be put at the end of the introduction.}
+%
+
 \section{Virtualization}

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Change chap name
Fix typos

@@
-\chapter{Architectural implementation}
+\chapter{Architectural proposal}
 \label{chap:archimpl}
@@
 all the problems related to it and how they were solved. We will show all the
-various attemps made, and we will discussion about the technological issue and
+various attempts made, and we will discussion about the technological issue and
 changes we had to perform.
@@
 on Openstack that, instead of deploying hypervisor based virtual machines, it
-would have deployed container instances. The VNF lifecycles would have been
+would have deployed container instances. The VNFs lifecycles would have been
 controller through an Element Management System (EMS), transmitting the various
@@
     organization schema. We can see how the main Openstack role is to deploy and
-    manage the resources, while Openbaton has the role of maangement and
+    manage the resources, while Openbaton has the role of management and
     orchestration. Finally, RabbitMQ sends and receives status update from the
@@
 multiple machines (nodes), therefore creating a whole cluster of resources. 
-Since we considered too time-consuming to install Openstack in a distribuited 
+Since we considered too time-consuming to install Openstack in a distributed 
 manner, we considered to test first the developer version, called 
@@
 the same operations of the production-grade installation. Thus, we starting 
-configurating the Devstack installation, installing first of all the Nova 
+configuring the Devstack installation, installing first of all the Nova 
 module, that is required to perform computer operations such as the launch of 
@@
 ability to deploy, edit, stop and do periodic health checks VNF directly on it.
-The develpers claim it has the possbility to make deployments to platforms
+The developers claim it has the possibility to make deployments to platforms
 compatible with Openstack too. It supports VNF definition through the TOSCA
@@
 
-We sperimented Tacker to see if its functionalities were good enough to be
+We tried Tacker to see if its functionalities were good enough to be
 adopted in our testbed instead of Openbaton. Its complete integration with
@@
 necessary parts we needed. The back of the medal of this approach was the
-impossibility to automate it: in fact, during the installation proces user
+impossibility to automate it: in fact, during the installation process user
 interaction was required, that we were not able to write down in a scripted
@@
 dynamically later become fundamental to us. On top of that, the Vagrant
-deployment uses a hyperviso virtualization system (based on Virtual Box), in
+deployment uses a hypervisor virtualization system (based on Virtual Box), in
 contrast with our goal to build a fully container-based system.
@@
 After the failure to find a viable setting for Openstack, we decided to look
-around for other solutions. We started studing proper Docker container
+around for other solutions. We started studying proper Docker container
 orchestrators that could offer all the features we needed: container lifecycle
 management, scaling, healing and monitoring. We decided to pick two of the most
-mature and famous ones: Docker Swarm and Kubenetes. The decision to study only
+mature and famous ones: Docker Swarm and Kubernetes. The decision to study only
 these two relies on their big community supporting them and the possibility to
-easily found projects that used the forementioned frameworks.
+easily found projects that used the aforementioned frameworks.
 
@@
 \subsection{Kubernetes}
-Kubenetes, as already described in the technology section, is a orchestrator 
+Kubernetes, as already described in the technology section, is a orchestrator
 created by Google able to create a software abstraction from the underlying
-virtualization technology. This enables Kubernetes to create an integration 
-with the Docker technology. Kubernetes has a architecture based on plug-ins, 
-giving  the possibility to the users to change part of Kubernetes itself, such 
-as the networking routing strategy or the persistence layer. With this 
-flexibility Kubernetes offered to us what Docker Swarm was not able to give. 
-Additionally, we had the possibility to set custom Domain Name System (DNS) 
-lookups and distribute other internal components in order to avoid requests 
-bottleneck.
+virtualization technology. This enables Kubernetes to create an integration with
+the Docker technology. Kubernetes has a architecture based on plug-ins, giving
+the possibility to the users to change part of Kubernetes itself, such as the
+networking routing strategy or the persistence layer. With this flexibility
+Kubernetes offered to us what Docker Swarm was not able to give. Additionally,
+we had the possibility to set custom Domain Name System (DNS) lookups and
+distribute other internal components in order to avoid requests bottleneck.
 
@@
   \caption[Kubernetes control plane schema]{Kubernetes control plane
-    schema~\cite{k8scp}. It is possible to denothe important components in the
+    schema~\cite{k8scp}. It is possible to denote important components in the
     Kubernetes Control plane. In this high-level component architecture the
@@
 \item[Scheduler] The Scheduler regulates the Pods distribution in the minions,
-  having a knowlege of the current cluster typologies and of the current
+  having a knowledge of the current cluster typologies and of the current
   resources available. On top of that, is able to handle hardware constraints,
@@
 
-All of the previously descripted components are hide beghind the API server,
+All of the previously descripted components are hide behind the API server,
 which is the only components the final user can really interact with. The
@@
 easily perform deployments and admin operations on the cluster without having a
-full knowlege of the official API.
+full knowledge of the official API.
 
@@
   that the Pod deployed are up and healthy. This excludes possible
-  intereferences with other virtualized deployments.
+  interference with other virtualized deployments.
 \item[Container Runtime] This runtime does not have to necessarily be based on
@@
   the container images on the OS.
-\item[Networking] This part aims to provide to che container runtime network
+\item[Networking] This part aims to provide to the container runtime network
   connectivity. The role of the networking module is to provide, over than the
@@
   Kubernetes, several networking implementation exists, like Flannel, Google
-  Copmute Engine (GCE), Kube-route, Project Calico~\cite{k8snetworkingwiki}. For
+  Compute Engine (GCE), Kube-route, Project Calico~\cite{k8snetworkingwiki}. For
   our deployment, we decided to use Flannel, that demonstrated to work without
@@
   of specification and libraries for writing plugins to configure network
-  interfaces in linux containers~\cite{cnigithub}.
+  interfaces in Linux containers~\cite{cnigithub}.
 \end{description}
@@
 After the Kubernetes configuration and installation, we started working for a
-viable intergation between Kubernetes and Openbaton. Our first idea is the one
+viable integration between Kubernetes and Openbaton. Our first idea is the one
 illustrated in Figure~\ref{chap:archimpl:sec:secondattempt:img:schema1}:
@@
 aspects such the deployment flavour and the kind of virtual link (concept that
-does not really exsits in container based systems) to instatiate. Nonetheless,
+does not really exists in container based systems) to instantiate. Nonetheless,
 we performed a mapping between most of the TOSCA keywords with some of the YAML
@@
 submitting it to Kubernetes. This solution had great advantages in terms of
-responsability separation: in this way only the MANO had to know, for a VIM
-based on Kubernetes, that a conversion needed to be performed. From a Kuberentes
+responsibility separation: in this way only the MANO had to know, for a VIM
+based on Kubernetes, that a conversion needed to be performed. From a Kubernetes
 point-of-view, nothing would have changed, since all the configuration
@@
 driver, that is the component that directly interacts with the VIM, and then the
-VNFM, because we discovered there were programmatic dipendences between the two
+VNFM, because we discovered there were programmatic dependencies between the two
 Openbaton standard components. After we identified the VIM driver and the VNFM
@@
 components the documentation was very poor, especially regarding the code. On
-top of that, the program interfaces given by the develpers did not match our
-expectations: we thought to find interfaces indipendent from the type of the
+top of that, the program interfaces given by the developers did not match our
+expectations: we thought to find interfaces independent from the type of the
 virtualization technology, instead, we found that they were strictly coupled
@@
     Additionally, Harbor now operates directly with the Kubernetes API,
-    retrieving informations about the VNFs status from it. This causes the VNFs
+    retrieving information about the VNFs status from it. This causes the VNFs
     to be lighter because they don't have to include any reporting tools, like
@@
 \item ingress and egress points: there should be at least two points in the
-  cluster were the packect enters and exits;
+  cluster were the packet enters and exits;
 \item dynamic package routing: as formulated by one of our requirements, packet
-  routing had to be dynamic. We immediatly identified a possibility to create
+  routing had to be dynamic. We immediately identified a possibility to create
   this functionality with the already-built features of Kubernetes;
@@
 the required DNS lookup meant a possible bottleneck. With a great number of
-components and VNFs constantly quering the centralized Kubernetes DNS lookup
+components and VNFs constantly querying the centralized Kubernetes DNS lookup
 system a performance degradation could have easily happened. We decided to
@@
 
-\subsubsection{Dyamic package routing: the Roulette component}
+\subsubsection{Dynamic package routing: the Roulette component}
 \label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
@@
 complete end-to-end topology view of the system, we decided to store these
-additional informations (we called it \emph{endpoints}) in it.
+additional information (we called it \emph{endpoints}) in it.
 

### 2018-11-27 res/sections/10-ProjectAnalysis.tex

@@
 in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
-
-\section{Architecture overview}
-
-During the test-bed development the software architecture diverged from the
-original one, experiencing different evolutions and modifications.
-
-\subsection{VIBES architectural description}
-
-In the original VIBES architectural design, based on leveraging virtualization
-technologies, is possible to identify different macro-components, each of these
-specialized in fulfilling a specific goal. For the ETSI standards, the Network
-Function Virtualization (NFV) design should be composed as follows:
-\begin{figure}[t]
- \centering \includegraphics[scale=2]{etsi_arch}
- \caption{A schematic image representing the high level NFV architecture
-   proposed by ESA.}
- \label{chap:prjan:img:etsi_arch}
-\end{figure}
-
-\begin{itemize}
- \item \textbf{Network Function Virtualization Infrastructure (NFVI)}: it's the
-   heart of the whole computation component, and here three sub-domains can be
-   identified: \todo{Expand this section}
-\begin{itemize} 
- \item \textbf{Hyper-visor domain} that is composed of the Virtual Compute, the
-   Virtual Storage, and the Virtualization Layer. These components are used to
-   virtualize the computational and storage resources, providing a layer for
-   accessing it.
- \item \textbf{Compute domain} composed by the Physical Compute and the Physical
-   Storage. These are the real resources which are virtualized via software.
- \item \textbf{Infrastructure domain} which has the Physical Network, and its
-   virtualized counterpart as made available by the Virtualization Layer.
-\end{itemize}
-
-In conventional virtualization systems these resources are usually separated
-from the host operating system, providing better isolation but worst
-performance. The ETSI infrastructure, instead, tries with container
-virtualization technology to push for a better NFV responsiveness, removing the
-additional operating system (called ``guest'' OS) required by the usual
-hyper-visor-oriented approach.
-
-To realize the required components, specific tools were suggested in the VNF
-technical proposal: regarding the NFV (network infrastructure) domain
-Docker-Compose was proposed as a viable solution, while to store PEP instances
-docker swarm was described as a possible candidate.
-
-We performed an accurate analysis of the tools chosen based on their maturity,
-on the community and on the technical support offered (e.g. user manuals,
-developer documentation), which, at the end, committed us to choose different
-tools from the suggested ones. A detailed description of the architectural
-implementation can be found in chapter~\ref{chap:archimpl}.\todo{Include prof.
-  survey about community components here}
-
-\item \textbf{VNFs} this are the core components that perform packets
-  manipulation. This components are design to be lightweight services able to
-  elaborate a great amount of packets per seconds. Manipulation can be of two
-  typologies: stateful or stateless. In the first case, a state is maintained,
-  and the VNF can perform heuristics about data coming through it (especially
-  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
-  case, no state is maintained, and every packet is treated without any
-  particular assumption. Following the RFC 7665, the VNF are also called Service
-  Function (SF)\footnote{From now on the two terminologies will be use in a
-    interchangeable way~\cite{medhat2017service}.}. Additional information can
-  be found in Section~\ref{chap:prjan:sfc}.
-
-\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
-  Orchestrator (NFVO) is responsible for the end-to-end management and
-  orchestration of network services (NS), also called SFC. The MANO has
-  different responsibilities, such as managing the
-  creation/destruction/modification of the SFCs, that achieves interacting with
-  the VIM~\cite{nguyen2017sdn}. Finally, the MANO can be integrated with
-  Operational and Business Support Systems (OSS/BSS).
-\end{itemize}
-
-\section{SFC}
-\label{chap:prjan:sfc}
-
-\todo{Here it's missing an introduction about SFC and its terminology, need to 
-include parts of RFC 7665}
-
-Even though SFCs and VNFs are part of the ETSI architecture, the complexity 
-behind was so high to require particular care and a separated study. Indeed, 
-the  SFCs are the core of the whole project: packet transmission and 
-elaboration is made here. Challenges regarding this topic are building a 
-system that is flexible enough to allow dynamic packet routing, support 
-scalability and provide methods to recovery from failures. With this challenges 
-packet transmission revealed to be a fundamental point to choose carefully.
-
-\section{Available technologies in the market} \todo{Add images}
-\label{chap:prjan:sec:tech}
-
-Before starting any effective work on the project it has been decided to 
-perform an analysis of the possible technologies available in the market, 
-performing a choice between them and exploiting their possible ``weak 
-points''. Proceeding with a top-down approach, the frameworks were chosen from 
-the one that had the biggest role in the test-bed to conclude with the 
-less-important components.
-
-\subsection{NFVI}
-
-A critical component, as the MANO one, is the NFVI. Without it, network 
-resources can't be allocated, and since our project was focused on dynamically 
-allocating containerized applications on cloud platforms, we decided to start 
-analyzing the right framework for this component first.
 

### 2018-11-27 res/sections/10-ProjectAnalysis.tex
Minor fix

@@
 in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
+
+\section{Available technologies in the market} \todo{Add images}
+\label{chap:prjan:sec:tech}
+Before starting any effective work on the project it has been decided to 
+perform an analysis of the possible technologies available in the market, 
+performing a choice between them and exploiting their possible ``weak 
+points''. Proceeding with a top-down approach, the frameworks were chosen from 
+the one that had the biggest role in the test-bed to conclude with the 
+less-important components.
+\subsection{VIM}
+A critical component, as the MANO one, is the VIM. Without it, network 
+resources can't be managed, and since our project was focused on dynamically 
+allocating containerized applications on cloud platforms, we decided to start 
+analyzing the right framework for this component first.
 
@@
 
-\vspace{0.5cm}
+% \vspace{0.5cm}
 
@@
 Another fundamental component in the architecture is the MANO, that has to
-manage the Network Services and the deployment of new VNFs. In coordination 
-with Kubernetes, it composes the core of the test-bed architecture. Initially 
-identified with Openbaton (the component is described later 
+manage the Network Services and the deployment of new VNFs. Initially identified
+with Openbaton (the component is described later
 in~\ref{chap:prjan:sec:tech:sub:other:sub:openbaton}), afterword it has been
 discovered to be not flexible enough to support other kinds of virtualization
-techniques than pure or in-cloud virtual machines\todo{Rewrite this}. 
-Since this was a strong project requisite that could not be easily changed, it 
-has been decided to opt for the creation of a smaller and simpler clone that 
-offered a full integration with Docker and Kubernetes.
+techniques than pure or in-cloud virtual machines. Since this was a strong
+project requisite that could not be easily changed, it has been decided to opt
+for the creation of a smaller and simpler clone that offered a full integration
+with Docker and Kubernetes.
 
@@
 docker containers. It's particularly useful when more containers have to
-interact 
-together, because\todo{Write down docker-compose functionalities and describe 
-the product a little bit} all the services rules are described in a simple 
-YAML file. With the birth of Docker Swarm, work has started to port this tool 
-to 
+interact together, because all the services rules are described in a simple YAML
+file. On top of that, it is possible to restart/stop/destroy selected services,
+while keeping the other up. Additionally, it supports containers scaling and it
+offers the possibility to simply get logs, interact of inspect with the defined
+services. With the birth of Docker Swarm, work has started to port this tool to
 the Swarm framework. At the time of this writing, the procedure to port a
-Docker-compose configuration in Docker Swarm (which a description of the 
-product can be found in 
-sub-section~\ref{chap:prjan:sec:tech:sub:other:sub:swarm}) is still not 
-completely transparent, and it requires a sort of compilation, where 
-Docker-compose bundles all the necessaries resources into a binary file that 
-Docker Swarm will use to allocate the necessary resources.
+Docker-compose configuration in Docker Swarm (which a description of the product
+can be found in sub-section~\ref{chap:prjan:sec:tech:sub:other:sub:swarm}) it
+has become completely transparent, bundling all the necessaries resources.
 
@@
 networking configurations or installing ad-hoc storage solutions. On the other
-hand, though, it doesn't have all the customization options that Kubernetes
+hand, though, it does not have all the customization options that Kubernetes
 offers, thus making the product less flexible. This fundamental characteristic,
@@
 \subsubsection{OpenVSwitch}
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.5]{ovs}
+  \caption[OpenVSwitch logo]{OpenVSwitch logo}
+\end{figure}
 Started in 2009~\cite{ovsGit}, the OpenVSwitch project aims to create an

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Add introduction on chapter 5

@@
 
+\section{Architecture introduction}
+Before describing the real implementation, we are going to briefly introduce
+some of its functional components and the way they work.
+
+\subsection{A two-level orchestration framework}
 \begin{figure}[t]
@@
     different parts to enable message exchanging.}
+  \label{chap:archimpl:sec:fistattempt:img:architecture}
 \end{figure}
+
+It is important to keep in consideration that the ETSI NFV architectural
+framework is to be considered service and/or application agnostic. This excludes
+this framework from the burden of being application aware for NS configuration
+and management and for the VNF application layer configuration and management.
+
+The lack of these two management requires a service aware software, that has the
+task to perform service lifecycle management and monitoring via the API offered
+by the ETSI MANO. In this context, is possible to identify two different type of
+orchestrator, as represented in
+Figure~\ref{chap:archimpl:sec:fistattempt:img:architecture}:
+\begin{itemize}
+\item a high layer, service-aware;
+\item a low layer, service-unaware.
+\end{itemize}
+Both the layer would use Openbaton. The messages would be exchanged through
+RabbitMQ, using the Advance Message Queuing Protocol (AMQP).
+
+\paragraph*{Low-layer orchestrator}
+The low-layer orchestrator jobs corresponds to the one defined in the ETSI MANO
+definition, with the role of managing and orchestrating VNFs and NSs. The
+low-layer MANO holds different data repositories where VNF and NS definitions
+are stored, various monitoring tools and logs and the network services
+repository. The low layer MANO exposes a set of APIs to the upper layer (related
+to handling services) MANO, that is able to configure and manage the services.
+In other words, the service MANO has the possibility to configure and launch
+services to the low-layer orchestrator, providing the necessary information.
+
+\subsection{Low-layer functional components}
+In the following section we are going to briefly talk about the role of the most
+peculiar components in the defined architecture. A description of the tasks that
+the components have to abide in the ETSI MANO architecture is available in
+Section~\ref{chap:archimpl:sec:secondattempt:sec:k8s}.
+\begin{description}
+\item[VNFM] As is possible to see from the
+  Figure~\ref{chap:archimpl:sec:fistattempt:img:architecture}, the following
+  architecture envisages the possibility of multiple VNFM components, required
+  for the control and management of different virtualized functions. In
+  particular, the use of different and specialized VNFMs allows to control and
+  manage different functional aspects of the control/data plan.
+\item[Network service management] in this architectural proposal, it corresponds
+  the the brain of the infrastructure MANO, where service instantiation and
+  lifecycle management happens. This part interacts with the VNFM/VIM and with
+  the RabbitMQ middleware. In this context, the NFVO is involved in chaining the
+  required VNFs in order to create the service delivery path. Inside this
+  component is possible to identify the Service Lifecycle Management, that
+  listens to status report coming from the VNFM and thus generating logs and
+  monitoring information.
+\item[Monitoring and Information Management] this block gathers data from the
+  data layers (identified in the image as the repository), which can be object
+  of consultation from other applications.
+\end{description}
+
+From now on we are going to describe out attempt in implementing the
+aforementioned architectural proposal, and we will describe the challenges and
+the issue we encountered.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 \subsection{Kubernetes configuration}
-
+\label{chap:archimpl:sec:secondattempt:sec:k8s}
 \begin{figure}[t]

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Other minor fixes

@@
 
-\subsection{Adopting Openstack for container orchestration}
+\subsection{Openstack for container orchestration}
 
@@
 the development phase of our software.
-\subsubsection{Openstack for developers: Devstack}
+\subsubsection{Devstack: Openstack for developers}
 Openstack is composed by modules, giving the possibility to the user who wants 

### 2018-11-27 res/sections/40-FutureWork.tex
Add partial future works

\chapter{Future Work}
\label{chap:fw}

The realization of this architecture proposal is not finished yet. Indeed, a lot
of functionalities could be implemented and new components coul be added. We are
now going to see possible future additions to improve the system.

\section{Security concerns}

One of the aspects that is usually not really given the required attention is
the one regarding security. In this context, the security we talk is the one
about the whole system, and not the single VNFs implementations. Fortunatly,
Kubernetes is equipped with out-of-the-box access management options. A good way
to manage user authentication in the system should be integrate the Role Based
Access model provided by Kubernetes directly in Harbor. In fact, Harbor does not
come with security features, and the exposed APIs are usable by everyone. This
could lead to possible Denial of Service (DoS) attack of the disruption of NS
services by a malicious attacker.

Another security aspect which is important to provide is the isolation between
the network slices. This isolation, in our implementation, should be achieved
clearly identifing sets of nodes (the different slices) and restricting their
possibility to talk and interact with the other sets. This kind of restrinction
should be achieved using different group Policy through Kubernetes.

\section{Kubernetes federation}

A single Kubernetes cluster presents limitations in terms of number of nodes
available on a single cluster and regarding the geographical position of this
nodes. Indeed, nodes in a cluster should be located or in the same network or
geographically closer to each other, in order to ensure show latences between
data exchange and Pods communications. An available solution, introduced in
version 1.9 of Kubernetes, is the possibility to create federated Kubernetes
cluster. This solution ensures a proper division between groups of nodes, and
the possibility to ensure better performances creating cluster with nodes closer
to each other. Unfortunatly, we werent able to create a proper Kubernetes
federation. On of the reason was related the current immaturity of the project,
that could have created and instable cluster federation.

\section{Dynamic chain update}

The importance of having NS services updated as smooth as possible.

### 2018-11-27 res/sections/03-Inscription.tex
Add acknowledgement and first citation

@@
 \thispagestyle{empty}
-\pdfbookmark{Dedica}{Dedica}
+\pdfbookmark{Inscription}{Inscription}
 
@@
 	\slshape
-	dedica} \\
+	Do not shorten the morning by getting up late, or waste it in unworthy occupations or in talk; look upon it as the quintessence of life, as to a certain extent sacred.
+Evening is like old age: we are languid, talkative, silly.
+Each day is a little life: every waking and rising a little birth, every fresh morning a little youth, every going to rest and sleep a little death.} \\
 	\medskip
+  Arthur Schopenhauer
 \end{flushright}

### 2018-11-27 res/sections/04-Credits.tex
Add acknowledgement and first citation

@@
 
-\section*{Credits}
+%\section*{Credits}
 
-\noindent \textit{Ringraziamenti}\\
+\noindent \textit{Acknowledgements}\\
 \bigskip
+
+\noindent When I began my journey at University, five years ago, I was only
+a boy. A boy curious about IT and geeky stuff, with the humble dream of becoming
+proficient with computing programming. During the years, the dream gradually
+became a goal to reach, a destination. The goal changed, becoming more
+ambitious, and now I am finally going to reach it\dots But I would not be able
+to make it without the multitude of people that I luckly have around me. It is
+for the power of this people if today I am here writing these acknowledgements.
+
+\vspace{0.3cm}
+
+\noindent A big thank for my friends, whose power of cheering me up and make
+me smile saved myself different times from the risk of failing this journey.
+Thanks because you where always here when I needed you, and you always found the
+right words to say, or the right thing to do. Big part of my success is thanks
+to them, who have pushed me in places I would not be able to reach alone. In
+particular I would like to thank Giovanni, Giorgio, Michele, Gianluca and
+Umberto for the good times we end up spending every weekend. On top of that, I
+would like to say also thanks to Marco, Federico, Matteo, Mirko, Emanuele,
+Stefano, Sara, Lisa and Eduard that closely witnessed with me the good and the
+bad moments of the last two years of this journey. The time spent together with
+you guys will not be easily forgotten.
+
+\vspace{0.3cm}
+
+\noindent Last but not least, I would like to praise my best friend, Franco,
+which has the power to keep me rolling in every aspect of my life, even when I
+am without energy to keep up or when I am completely lost, and which is alway
+with me, regardless of the physical distance that separate us. Thanks.
+
+\vspace{0.3cm}
+
+\noindent Not only friends helped me to complete my journey. Additional
+acknowledgements go to my favourite sport, Judo, that allowed me to mentally e
+spiritually grow and to become what I am now. It is for people that never gave
+up like you, Pierluigi, if I am now able to keep a cool head even in the worst
+of the storms.
+
+\vspace{0.3cm}
+
+\noindent Other thanks go to my parents, which supports my every day and always
+believed in me, no matter what.
+
+\vspace{0.3cm}
+
+\noindent Finally, thanks to my supervisor, Prof. Armir Bujari, to have
+mentored me during the design and development of this thesis.
+
+\vspace{0.5cm}
 

### 2018-11-28 res/sections/40-FutureWork.tex
Finish future work draft

@@
 
-The importance of having NS services updated as smooth as possible.
+The importance of having NS services updated as smooth as possible is without 
+doubts a feature that would ensure a continue service availability avoiding its 
+disruption. This kind of approach, though, requires particular attention about 
+keeping the current connections alive. This means at least keeping the old 
+chain up while deploying the new one, and gradefully redirect all the 
+new traffic from the chain to replace, until none is left. Only after this 
+operation, the old chain definition can be completely decomissioned and the 
+related VNFs removed. In our current implementation, the NS update 
+functionality relies on a simpler approach: when an update gets ordered, 
+Kubernetes gets instructed to remove the old chain and deploy the new one. 
+Meanwhile, all the Roulette controllers get updated with the new NS route. The 
+back of the medal of this soulution is, while it keeps the sistem implementation 
+simpler, that packets current travelling along the old NS could be lost during 
+the chain updating.

### 2018-11-28 res/sections/40-FutureWork.tex
Fix future work

@@
 The realization of this architecture proposal is not finished yet. Indeed, a lot
-of functionalities could be implemented and new components coul be added. We are
-now going to see possible future additions to improve the system.
+of functionalities could be implemented and new components could be added. We
+are now going to see possible future additions to improve the system.
 
@@
 Another security aspect which is important to provide is the isolation between
-the network slices. This isolation, in our implementation, should be achieved
+network slices. This isolation, in our implementation, should be achieved
 clearly identifing sets of nodes (the different slices) and restricting their
 possibility to talk and interact with the other sets. This kind of restrinction
-should be achieved using different group Policy through Kubernetes.
+should be achieved using different policy groups through Kubernetes.
 
@@
 nodes. Indeed, nodes in a cluster should be located or in the same network or
-geographically closer to each other, in order to ensure show latences between
+geographically closer to each other, in order to ensure low latences between
 data exchange and Pods communications. An available solution, introduced in
 version 1.9 of Kubernetes, is the possibility to create federated Kubernetes
-cluster. This solution ensures a proper division between groups of nodes, and
-the possibility to ensure better performances creating cluster with nodes closer
-to each other. Unfortunatly, we werent able to create a proper Kubernetes
-federation. On of the reason was related the current immaturity of the project,
-that could have created and instable cluster federation.
+clusters. This solution ensures a proper division between groups of nodes, and
+the possibility to ensure better performances creating clusters only with nodes
+closer to each other. Unfortunatly, we were not able to create a proper
+Kubernetes federation. One of the main reasons was related the current
+immaturity of the project, that could have created and instable cluster
+federation.
 
@@
 disruption. This kind of approach, though, requires particular attention about
-keeping the current connections alive. This means at least keeping the old 
-chain up while deploying the new one, and gradefully redirect all the 
-new traffic from the chain to replace, until none is left. Only after this 
-operation, the old chain definition can be completely decomissioned and the 
-related VNFs removed. In our current implementation, the NS update 
-functionality relies on a simpler approach: when an update gets ordered, 
-Kubernetes gets instructed to remove the old chain and deploy the new one. 
-Meanwhile, all the Roulette controllers get updated with the new NS route. The 
-back of the medal of this soulution is, while it keeps the sistem implementation 
-simpler, that packets current travelling along the old NS could be lost during 
-the chain updating.
+keeping the current connections alive. This means at least keeping the old chain
+up while deploying the new one, and gracefully redirect all the new traffic from
+the chain to replace, until none is left. Only after this operation, the old
+chain definition can be completely decomissioned and the related VNFs removed.
+In our current implementation, the NS update functionality relies on a simpler
+approach: when an update gets ordered, Kubernetes gets instructed to remove the
+old chain and deploy the new one. Meanwhile, all the Roulette controllers get
+updated with the new NS route. While this keeps the system implementation
+simpler the back of the medal of this soulution is that packets currently
+travelling along the old NS could be lost during the chain update.

### 2018-11-28 res/sections/40-FutureWork.tex
Fix typos

@@
 the one regarding security. In this context, the security we talk is the one
-about the whole system, and not the single VNFs implementations. Fortunatly,
+about the whole system, and not the single VNFs implementations. Fortunately,
 Kubernetes is equipped with out-of-the-box access management options. A good way
@@
 network slices. This isolation, in our implementation, should be achieved
-clearly identifing sets of nodes (the different slices) and restricting their
-possibility to talk and interact with the other sets. This kind of restrinction
+clearly identifying sets of nodes (the different slices) and restricting their
+possibility to talk and interact with the other sets. This kind of restriction
 should be achieved using different policy groups through Kubernetes.
@@
 nodes. Indeed, nodes in a cluster should be located or in the same network or
-geographically closer to each other, in order to ensure low latences between
+geographically closer to each other, in order to ensure low latencies between
 data exchange and Pods communications. An available solution, introduced in
@@
 the possibility to ensure better performances creating clusters only with nodes
-closer to each other. Unfortunatly, we were not able to create a proper
+closer to each other. Unfortunately, we were not able to create a proper
 Kubernetes federation. One of the main reasons was related the current
-immaturity of the project, that could have created and instable cluster
+immaturity of the project, that could have created and unstable cluster
 federation.
@@
 the chain to replace, until none is left. Only after this operation, the old
-chain definition can be completely decomissioned and the related VNFs removed.
+chain definition can be completely decommissioned and the related VNFs removed.
 In our current implementation, the NS update functionality relies on a simpler
@@
 updated with the new NS route. While this keeps the system implementation
-simpler the back of the medal of this soulution is that packets currently
+simpler the back of the medal of this solution is that packets currently
 travelling along the old NS could be lost during the chain update.

### 2018-11-28 res/sections/08-Background.tex
Other minor fixes
Fix image size
Fix image position

@@
 
-\section{The 5G connectivity standard}
+\section{Fifth generation mobile networks}
 
@@
 
-\subsection{5G architecture}
+\subsection{5G 3GPP MANO}
 
@@
 \section{The VIBeS project}
+
 This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
@@
 
-\begin{figure}[t]
- \centering
- \includegraphics[scale=1]{vibes_logo}
- \caption{ViBeS project logo}
- \label{chap:background:img:vibes_logo}
-\end{figure}
-
 \subsection{VNF-PEP architecture and internet protocols}
@@
 how to optimize a certain communication.
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[scale=1]{vibes_logo}
+ \caption{ViBeS project logo}
+ \label{chap:background:img:vibes_logo}
+\end{figure}
 

### 2018-11-28 res/sections/15-ArchitecturalImplementation.tex
Fix sentences

@@
 
-In this chapter we will dig into the actual architectural implementation, seeing
+In this chapter we will dig into the actual architectural proposal, seeing
 all the problems related to it and how they were solved. We will show all the
@@
 
-We tried Tacker to see if its functionalities were good enough to be
-adopted in our testbed instead of Openbaton. Its complete integration with
-Openstack, indeed, was the main reason for its testing. We felt like the product
-was not fully ready to be used in a real deployment though, since it lacked of
-options like SFC management and the fact that it only officially supported
-Openstack and not other cloud platforms was another reason to prefer Openbaton
-eventually.
+We tried Tacker to see if its functionalities were good enough to be adopted in
+our testbed instead of Openbaton. Its complete integration with Openstack,
+indeed, was the main reason for its testing. At the end though, we felt like the
+product was not ready for a real production environment. SFC management and the
+lack of support for alternative platforms other than Openstack was another
+reason to prefer Openbaton eventually.
 
@@
 The deployment via Vagrant confirmed our expectations because it revealed to be
-was always smooth. The downside of this approach was the lack of customization
-and of component management. The possibility to remove or add Openbaton Modules
-dynamically later become fundamental to us. On top of that, the Vagrant
-deployment uses a hypervisor virtualization system (based on Virtual Box), in
-contrast with our goal to build a fully container-based system.
+always a smooth operation. The downside of this approach was the lack of
+customization and component management. The possibility to remove or add
+Openbaton Modules dynamically become later fundamental to us. On top of that,
+the Vagrant deployment uses a hypervisor virtualization system (based on Virtual
+Box), in contrast with our goal to build a fully container-based system.
 
 \subparagraph*{Deployment via Docker Compose}
-The deployment with Docker Compose turned out to be best of the three
+The deployment with Docker Compose turned out to be the best of the three
 possibilities. The Docker Compose configuration offered an unprecedented
 configuration flexibility, with the possibility to deploy only selected modules.
-On top of that, this method let us manage the Openbaton version freely, having
-more control on update especially during the development of our components.
-Additionally, using this Docker Compose configuration allowed us to seamlessly
-integrate the monitoring tools Zabbix, in order to have the possibility to
-graphically see the data fetch from Openbaton about the system state. Finally, a
-Docker based solution seemed to be the best approach since all of our components
-were meant to be put inside a Docker container.
+On top of that, this method let us to manage the Openbaton version freely,
+having more control on update especially during the development of our
+components. Additionally, using this Docker Compose configuration allowed us to
+seamlessly integrate the monitoring tools Zabbix, in order to have the
+possibility to graphically see the data fetch from Openbaton about the system
+state. Finally, a Docker based solution seemed to be the best approach since all
+of our components were meant to be put inside a Docker container.
 
@@
   \includegraphics[scale=0.5]{dockeropenstackmodule}
-  \caption[Docker module for Openstack working schema]{Docker module for Openstack
-    working schema~\cite{openstackDockerModule}}
+  \caption[Docker module for Openstack working schema]{Docker module for
+    Openstack working schema~\cite{openstackDockerModule}}
   \label{chap:archimpl:sec:fistattempt:img:dockeropnestackmodule}
@@
 
-The impossibility to develop a custom VIM driver and VNFM manager lead us to
+The impossibility to develop a custom VIM driver and VNFM manager led us to
 search for an alternative solution. Then, we started to work directly on the
@@
 These, defined as \emph{Ingress} and \emph{Egress} points, are performed by
-\emph{Ironhide}. Ironhide acts like a hides the whole packet processing to the
-sender $S$ and the receiver $R$. When a client $S$ performs a request to $R$,
-the packet actually goes through Ironhide, who has an internal classifier that
-it is able to recognize the type of connection and to choose the right NS the
-packet has to go through. This process is fundamental since it keeps the
-connection between $S$ and $R$ up, for both TCP and UDP based communications.
-Since the connection need to be end-to-end, the state of the connections are
-saved in the \emph{Roulette Controller} component.
+\emph{Ironhide}. Ironhide hides the whole packet processing to the sender $S$
+and the receiver $R$. When a client $S$ performs a request to $R$, the packet
+actually goes through Ironhide, who has an internal classifier that it is able
+to recognize the type of connection and to choose the right NS the packet has to
+go through. This process is fundamental since it keeps the connection between
+$S$ and $R$ up, for both TCP and UDP based communications. Since the connection
+need to be end-to-end, the state of the connections are saved in the
+\emph{Roulette Controller} component.
 

### 2018-11-28 res/sections/15-ArchitecturalImplementation.tex
Fix errors in 15

@@
 third component acting as a route controller. Roulette fulfills this role,
-providing a back-end where the single VNF can ask the next packet hop. Indeed,
-routes must be submitted by another component, that we identified as Harbor.
-After a NS deployment, Roulette gets instructed with a new route describing a
-NS. Every NS has a proper ID (called \verb!sfid!), and thus when the classifier
-checks a packet typology, it labels that packet with the specific NS ID (for a
-detailed description about the SFC header see
+providing a back-end where the single Service Function Forwarder (SFF) can ask
+the next packet hop. Indeed, routes must be submitted by another component, that
+we identified as Harbor. After a NS deployment, Roulette gets instructed with a
+new route describing a NS. Every NS has a proper ID (called \verb!sfid!), and
+thus when the classifier checks a packet typology, it labels that packet with
+the specific NS ID (for a detailed description about the SFC header see
 Section~\ref{chap:background:sec:sfc}). Finally, thanks to this label, every VNF
@@
 
-\subsubsection{A VNF framework: Astaire}
-As is possible to denote, a VNF doesn't only process the incoming packets, but
-it has to manage the communications and the related forward operations. To
-possibly ease part of the transmission problems and to make it easier to create
-a VNF without requiring the whole knowledge about SFC, we decided to build a
-framework, Astaire, that could kept the SFC related details decoupled from the
-core of the VNF, in other words, the entity that actually elaborates the packet.
-Built using an event-based pattern, it presents hooks that allow other software
-written in different languages from the Astaire one (e.g. Java), to be called
-dynamically at runtime in order to process the packet. The result of this
-elaboration gets returned to Astaire, that performs the beforehand cited
-operations of VNFs communications and forwarding. In the SFC specification,
-Astaire is defined as a Service Function Forwarder (SFF).
+\subsubsection{A Service Function Forwarder: Astaire}
+As is possible to denote, an SFC-unaware VNF can not elaborate a packet directly
+coming from an SFC. The packet coming from an SFC contains additional
+information and metadata that have to be removed first. In this context, we
+decided to build a framework, Astaire, that could kept the SFC related details
+decoupled from the core of the VNF, in other words, the entity that actually
+elaborates the packet. Built using an event-based pattern, it presents hooks
+that allow other software written in different languages from the Astaire one
+(e.g. Java), to be called dynamically at runtime in order to process the packet.
+The result of this elaboration gets returned to Astaire, that performs the
+beforehand cited operations of VNFs communications and forwarding. In the SFC
+specification, Astaire is defined as a SFF.
 

### 2018-11-28 res/sections/10-ProjectAnalysis.tex
Fix todo and change sec names

@@
 
-\section{Available technologies in the market} \todo{Add images}
+\section{Available technologies in the market}
 \label{chap:prjan:sec:tech}
@@
 \subsection{VIM}
-A critical component, as the MANO one, is the VIM. Without it, network 
-resources can't be managed, and since our project was focused on dynamically 
-allocating containerized applications on cloud platforms, we decided to start 
-analyzing the right framework for this component first.
+A critical component, as the MANO one, is the VIM. Without it, network resources
+can not be managed, and since our project was focused on dynamically allocating
+containerized applications on cloud platforms, we decided to start analyzing the
+right framework for this component first.
 
@@
 deployment containing only simple services (e.g. a back-end application serving
-a website or a database) this ends up in a waste of resources. \todo{Talk about
-  lxc containers and how Docker solved magically the problem.} It is here that,
+a website or a database) this ends up in a waste of resources. It is here that,
 in 2013~\cite{dockerWebsite}, Docker was created, basing its solution on an
@@
 
-\subsection{MANO}
+\subsection{Available MANO solutions}
 
@@
 
-\subsection{SFC}
+\subsection{Available SFC solutions}
 The tools analysis revealed little about already existent VNF or NS

### 2018-11-28 res/sections/08-Background.tex
Fix image position

@@
 \subsection{OpenFlow protocol}
+
+\begin{figure}[h]
+ \centering
+ \includegraphics[scale=0.65]{openflow}
+ \caption{OpenFlow protocol communication graph}
+ \label{chap:background:img:openflow_protocol}
+\end{figure}
+
 OpenFlow is an open communication protocols that enables the possibility to
@@
 packages can be discarded).
-
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.65]{openflow}
- \caption{OpenFlow protocol communication graph}
- \label{chap:background:img:openflow_protocol}
-\end{figure}
 

### 2018-11-28 res/sections/20-VNFImplementaton.tex
Started sfc implementation

@@
-\chapter{VNF and Network Service implementation} 
+\chapter{Service Function Chain Implementation} 
 \label{chap:vnf_ns_impl}
+
+The complexity behind the creation of a working Service Function Chain system
+was so high to require particular care and a separated study. Indeed, the SFCs
+are the at the base of the whole architecture: packet transmission and
+elaboration is made here. Challenges regarding this topic are building a system
+that is flexible enough to allow dynamic packet routing, support scalability and
+provide methods to recovery from failures. In this chapter, a brief explanation
+about the challenges and the issues encountered building this system will be
+given.
+
+\section{Harbor}
+
+As already cited, Harbor acts as a MANO, orchestrating and managing VNFs and the
+related services. From the SFC point of view, it carries consistency checks when
+a NS definition is created. The checks are carried out based on the VNF ID.
+While the VNF definitions corresponds to a YAML Kubernetes configuration, the NS
+one is a JSON array, for example:
+
+\begin{lstlisting}[caption={Example of SFC definition}, captionpos=b,
+                   language=json, label=chap:impl:lst:sfcexample]
+{
+    "ns" : [
+                {
+                    "id": "astaire-service"
+                },
+                ...
+            ]
+}
+\end{lstlisting}
+
+\section{Kubernetes Pods}
+
+All the SFC-related components have been developed with the idea of being
+containerized. This is a foundamental concept, because in this way all the
+components have the possibility to run inside Kubernetes, granting them the
+scalability and persistence benefits. On top of that, the use of
+\verb!Deployment! granted the possibility to perform easy an update of already
+existing components.
+
+\section{Approved proposal}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{finalreview}
+  \caption{Final review schema}
+  \label{chap:impl:img:finalreview}
+\end{figure}
+
+The approved proposal for the creation of a proper chain has been based on four components:
+\begin{itemize}
+\item Astaire;
+\item Harbor;
+\item Ironhide;
+\item Roulette.
+\end{itemize}

### 2018-11-28 res/sections/20-VNFImplementaton.tex
Finish SFC implementation

@@
 \end{itemize}
+
+In this context our chain has the ability to perform packet classification,
+making the whole system able to separate the traffic based on a packet-type
+policy.
+
+Another important aspect to keep in consideration is how communications inside
+the chain are performed. The packet arriving in the ingress needs to be
+perserved, so that VNFs can perform packet elaboration. In order to otain this,
+packet encapsulation has been used. IP in IP encapsulation allow to preserve the
+original packet, making the incoming packet the payload of a new one.
+
+Ironhide, which was introduced in the previous chapter, plays a crucial role in
+the establishment of a end-to-end connection between the client and the
+receiver. Ironhide is able to fulfill the ingress and the egress role, and it
+provides the necessary tools to add and remove the SFC headers that gets appended
+and removed.
+
+Finally, we were able to create bidirectional chains, resistent to faults and
+degradations that saddisfty the defined requirements.

### 2018-11-28 res/sections/20-VNFImplementaton.tex
Fix typos

@@
 All the SFC-related components have been developed with the idea of being
-containerized. This is a foundamental concept, because in this way all the
+containerized. This is a fundamental concept, because in this way all the
 components have the possibility to run inside Kubernetes, granting them the
@@
 the chain are performed. The packet arriving in the ingress needs to be
-perserved, so that VNFs can perform packet elaboration. In order to otain this,
+preserved, so that VNFs can perform packet elaboration. In order to obtain this,
 packet encapsulation has been used. IP in IP encapsulation allow to preserve the
@@
 
-Finally, we were able to create bidirectional chains, resistent to faults and
-degradations that saddisfty the defined requirements.
+Finally, we were able to create bidirectional chains, resistant to faults and
+degradation that satisfy the defined requirements.

### 2018-11-28 res/sections/05-Abstract.tex
Applied thesis template
Add abstract draft

--- 2018-10-14 res/sections/04-Credits.tex
@@
 %**************************************************************
-% Ringraziamenti
+% Abstract
 %**************************************************************
@@
 \phantomsection
+
 \thispagestyle{empty}
-\pdfbookmark{Credits}{credits}
 
+\newenvironment{changemargin}[2]{%
+\begin{list}{}{%
+\setlength{\topsep}{0pt}%
+\setlength{\leftmargin}{#1}%
+\setlength{\rightmargin}{#2}%
+\setlength{\listparindent}{\parindent}%
+\setlength{\itemindent}{\parindent}%
+\setlength{\parsep}{\parskip}%
+}%
+\item[]}{\end{list}}
+
+\vspace*{\fill}
+\begin{changemargin}{1.5cm}{1.5cm}
+\section*{Abstract}
+The importance of Internet is growing every year. Mobile and wired traffic are
+continuously surging, asking for better performances. On this day, ISPs rely on
+optimized, ad-hoc and often in-hardware nework solutions. These network
+functions are costly and require time to be deployed and mantained. In this
+context, upcoming virtualization technolgies like hypervisor-based and
+container-based are providing the possibility to create virtualized network
+functions, that have the ability to be up and running in a matter of seconds.
+VNFs are the managed and orchestrated by a specific application, the MANO that
+handles their lifecycle.
+
+In this thesis we are going to propose an Infrastructure Management and
+Orchestrator plane Proof-of-Concept, based on the container virtualization
+technology and on the 3GPP Release 15 specifications. In the current PoC, Docker
+and the container orchestrator Kubernetes play a fundamntal role.
+\vspace*{\fill}
+\end{changemargin}
 %\bigskip
-
-\begingroup
-\let\clearpage\relax
-\let\cleardoublepage\relax
-\let\cleardoublepage\relax
-
-\section*{Credits}
-
-\noindent \textit{Ringraziamenti}\\
-\bigskip
-
-\noindent\textit{\myLocation, \myTime}
-\hfill \myName
-
-\endgroup

### 2018-11-28 res/sections/25-TestsResponsiveness.tex
Remove duplicated part

@@
 initiate with the scheduling of a Pod. A Pod can have several lifecycle
-states~\cite{kubePodLifecycle}:
-\begin{description}
-\item[pending:] The Pod has been accepted by the Kubernetes system, but one or
-  more of the Container images has not been created. This includes time before
-  being scheduled as well as time spent downloading images over the network,
-  which could take a while;
-\item[running:] The Pod has been bound to a node, and all of the Containers have
-  been created. At least one Container is still running, or is in the process of
-  starting or restarting;
-\item[succeeded:] All Containers in the Pod have terminated in success, and will
-  not be restarted;
-\item[failed:] All Containers in the Pod have terminated, and at least one
-  Container has terminated in failure. That is, the Container either exited with
-  non-zero status or was terminated by the system;
-\item[unknown:] For some reason the state of the Pod could not be obtained,
-  typically due to an error in communicating with the host of the Pod.
-\end{description}
-
-On top of that, different conditions can affect a Pod:
-\begin{itemize}
-\item \verb!PodScheduled!: the Pod has been scheduled to a node;
-\item \verb!Ready!: the Pod is able to serve requests and should be added to the
-  load balancing pools of all matching Services;
-\item \verb!Initialized!: all ``Init Containers'' have started successfully;
-\item \verb!Unschedulable!: the scheduler cannot schedule the Pod right now, for
-  example due to lacking of resources or other constraints.
-\end{itemize}
+states, that are explained in~\ref{pod}.
 

### 2018-11-29 res/sections/05-Abstract.tex
Change commands order
Fix margins and section title
Fix typo

@@
 
-\newenvironment{changemargin}[2]{%
-\begin{list}{}{%
-\setlength{\topsep}{0pt}%
-\setlength{\leftmargin}{#1}%
-\setlength{\rightmargin}{#2}%
-\setlength{\listparindent}{\parindent}%
-\setlength{\itemindent}{\parindent}%
-\setlength{\parsep}{\parskip}%
-}%
-\item[]}{\end{list}}
-
 \vspace*{\fill}
-\begin{changemargin}{1.5cm}{1.5cm}
-\section*{Abstract}
+\par
+\begingroup
+\leftskip4em
+\rightskip\leftskip
+\section*{\centering Abstract}
 The importance of Internet is growing every year. Mobile and wired traffic are
@@
 functions, that have the ability to be up and running in a matter of seconds.
-VNFs are the managed and orchestrated by a specific application, the MANO that
+VNFs are then managed and orchestrated by a specific application, the MANO that
 handles their lifecycle.
@@
 and the container orchestrator Kubernetes play a fundamntal role.
+\par
+\endgroup
 \vspace*{\fill}
-\end{changemargin}
 %\bigskip

### 2018-11-29 res/sections/05-Abstract.tex
Fix typo

@@
 functions, that have the ability to be up and running in a matter of seconds.
-VNFs are then managed and orchestrated by a specific application, the MANO that
+VNFs are then managed and orchestrated by a specific application, the MANO, that
 handles their lifecycle.
@@
 technology and on the 3GPP Release 15 specifications. In the current PoC, Docker
-and the container orchestrator Kubernetes play a fundamntal role.
+and the container orchestrator Kubernetes play a fundamental role.
 \par

### 2018-11-29 res/sections/08-Background.tex
Fix citation in background

@@
 modular one, exploiting the flexibility that Virtual Network Functions
-(VNFs)\cite{moens2014vnf} can offer thanks to virtualization, closing the gap to
-the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
+(VNFs)~\cite{moens2014vnf} can offer thanks to virtualization, closing the gap
+to the use-case fulfillment defined by the NGMN alliance~\cite{iwamura2015ngmn}
 which require a decreased time to set up and deploy network services
@@
   \centering
-  \includegraphics[scale=0.2]{networkslicing}
+  \includegraphics[scale=0.3]{networkslicing}
   \caption[The three layers model of network slicing]{The three layers model of

### 2018-11-29 res/sections/10-ProjectAnalysis.tex
Add DNSMasq software description (#51)

--- 2018-11-27 res/sections/10-ProjectAnalysis.tex
@@
 in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
+
+\section{Architecture overview}
+
+During the test-bed development the software architecture diverged from the
+original one, experiencing different evolutions and modifications.
+
+\subsection{VIBES architectural description}
+
+In the original VIBES architectural design, based on leveraging virtualization
+technologies, is possible to identify different macro-components, each of these
+specialized in fulfilling a specific goal. For the ETSI standards, the Network
+Function Virtualization (NFV) design should be composed as follows:
+\begin{figure}[t]
+ \centering \includegraphics[scale=2]{etsi_arch}
+ \caption{A schematic image representing the high level NFV architecture
+   proposed by ESA.}
+ \label{chap:prjan:img:etsi_arch}
+\end{figure}
+
+\begin{itemize}
+ \item \textbf{Network Function Virtualization Infrastructure (NFVI)}: it's the
+   heart of the whole computation component, and here three sub-domains can be
+   identified: \todo{Expand this section}
+\begin{itemize} 
+ \item \textbf{Hyper-visor domain} that is composed of the Virtual Compute, the
+   Virtual Storage, and the Virtualization Layer. These components are used to
+   virtualize the computational and storage resources, providing a layer for
+   accessing it.
+ \item \textbf{Compute domain} composed by the Physical Compute and the Physical
+   Storage. These are the real resources which are virtualized via software.
+ \item \textbf{Infrastructure domain} which has the Physical Network, and its
+   virtualized counterpart as made available by the Virtualization Layer.
+\end{itemize}
+
+In conventional virtualization systems these resources are usually separated
+from the host operating system, providing better isolation but worst
+performance. The ETSI infrastructure, instead, tries with container
+virtualization technology to push for a better NFV responsiveness, removing the
+additional operating system (called ``guest'' OS) required by the usual
+hyper-visor-oriented approach.
+
+To realize the required components, specific tools were suggested in the VNF
+technical proposal: regarding the NFV (network infrastructure) domain
+Docker-Compose was proposed as a viable solution, while to store PEP instances
+docker swarm was described as a possible candidate.
+
+We performed an accurate analysis of the tools chosen based on their maturity,
+on the community and on the technical support offered (e.g. user manuals,
+developer documentation), which, at the end, committed us to choose different
+tools from the suggested ones. A detailed description of the architectural
+implementation can be found in chapter~\ref{chap:archimpl}.\todo{Include prof.
+  survey about community components here}
+
+\item \textbf{VNFs} this are the core components that perform packets
+  manipulation. This components are design to be lightweight services able to
+  elaborate a great amount of packets per seconds. Manipulation can be of two
+  typologies: stateful or stateless. In the first case, a state is maintained,
+  and the VNF can perform heuristics about data coming through it (especially
+  useful to detect patterns in the data flow, e.g. DDos attacks). In the second
+  case, no state is maintained, and every packet is treated without any
+  particular assumption. Following the RFC 7665, the VNF are also called Service
+  Function (SF)\footnote{From now on the two terminologies will be use in a
+    interchangeable way~\cite{medhat2017service}.}. Additional information can
+  be found in Section~\ref{chap:prjan:sfc}.
+
+\item \textbf{NFV Management and Orchestration (MANO)} also known as NFV
+  Orchestrator (NFVO) is responsible for the end-to-end management and
+  orchestration of network services (NS), also called SFC. The MANO has
+  different responsibilities, such as managing the
+  creation/destruction/modification of the SFCs, that achieves interacting with
+  the VIM~\cite{nguyen2017sdn}. Finally, the MANO can be integrated with
+  Operational and Business Support Systems (OSS/BSS).
+\end{itemize}
+
+\section{SFC}
+\label{chap:prjan:sfc}
+
+\todo{Here it's missing an introduction about SFC and its terminology, need to 
+include parts of RFC 7665}
+
+Even though SFCs and VNFs are part of the ETSI architecture, the complexity 
+behind was so high to require particular care and a separated study. Indeed, 
+the  SFCs are the core of the whole project: packet transmission and 
+elaboration is made here. Challenges regarding this topic are building a 
+system that is flexible enough to allow dynamic packet routing, support 
+scalability and provide methods to recovery from failures. With this challenges 
+packet transmission revealed to be a fundamental point to choose carefully.
+
+\subsection{Packet transmission strategies}
+
+When a packets are sent from a sender $S$, to a receiver $R$, they start
+travelling to the backbone receiving appropriate in-hardware packet processing
+before being received by $R$. During this elaboration, it is important to
+maintain the transparency of the connection, so that the network users are not
+aware of the packets elaboration at all. To perform that, many solutions are
+available: TUN/TAP packet encapsulation, IP Encapsulation within IP or TCP
+session recreation (know also as TCP split).
+
+TUN/TAP and IP Encapsulation within IP (initially defined in RFC 2003 with the
+aim to deliver IP packets over Mobile IP for mobile networks) maintain the
+connection transparency encapsulating the packet in another one.
+\begin{figure}[t]
+  \centering \includegraphics[scale=0.5]{IPoverIP}
+  \caption[IP Encapsulation within IP packet schema]{A schema of IP
+    Encapsulation within IP. Is noticeable how the original packet becomes the
+    payload of a new one, thus being encapsulated. All the information
+    regarding the original packet remains untouched: when this kind of packet
+    gets send from a machine to another, the OS will remove the new packet
+    header and will present to the user-space program receiving the data its
+    payload: a IP packet. Packets encapsulation can be applied with higher
+    layers too, thus allowing for TCP and UDP packet encapsulation.}
+  \label{chap:prjan:img:ip_over_ip}
+\end{figure}
+
+In this way, the original packet remains preserved and it becomes the new packet
+payload, that can be elaborated and modified accordingly, until it gets
+decapsulated and sent to $R$ eventually. When $R$ receives the packet, it is not
+be aware of the packet elaboration, since all the original headers were
+preserved (or slightly modified). In particular, IP Encapsulation within IP
+allow the packets to make intermediate destinations that otherwise would not be
+selected. TCP session recreation, instead, ``split'' the TCP session in two
+endpoints: an Ingress, that manage the connection with the client and an Egress,
+that communicate with the packet receiver creating a multi-overlay-hop path
+where for each hop there is an possible independent TCP connection. In order to
+achieve that, a session table needs to be established and continuously updated,
+while TCP headers need to be accordingly modified to allow data transmission
+through the Ingress and Egress points. Nonetheless, care should be given to
+packets path: in order to avoid ill-calculated congestion windows, the incoming
+packets need to follow the same (or an equivalent one) path for the same
+instantiated connection. This phenomenon is based on the intermediate proxies
+that create the route: it is the case, for example, when an intermediate node
+sends back a spoofed ACK before the original packet is delivered to the real
+consignee. In this scenario, the sender, unaware of the proxy, will miscalculate
+the sending window based on the too low RTT. Other problems of TCP splitting are
+reliability and security, especially because the end to end TCP logic is no
+longer maintained: a server failure may cause to an unaware client to believe
+that all the packets have reached the destination. TCP session recreation seems
+burdensome and it doesn't seem to add any significant benefit at the first
+sight, but it offers the flexibility to perform extensive packet manipulation:
+since the packet its recreated every time at the Ingress and Egress points,
+while in the Network Function Virtualization Infrastructure (NFVI), the original
+packet can be stripped of its headers (that could be saved in the session table)
+to slightly improve internal packet transmission. \todo{Insert TCP session
+  recreation schema} On top of that, it offers the possibility to use different
+TCP flavours, hence increasing the communication speed in heterogeneous networks
+(e.g. in case of satellite connections TCP Hybla can be employed, while it is
+known that HighSpeed TCP gives best results when applied in optical fiber-based
+connections).
+
+During the test-bed development we first tried the TUN/TAP solution, that
+revealed to be not completely transparent and with major performance drawbacks
+(i.e. packet transmission in a LAN connection reached latencies of 50ms)
+\todo{Explain the constraint of point-to-point connections, while we need
+  dynamic SFC)}, so we choose to opt for a TCP session recreation solution,
+implementing a little back-end with the aim of saving in a persistent storage
+the SFC the packet has to perform and additional metadata.
+
+\section{Available technologies in the market}
+\label{chap:prjan:sec:tech}
+
+Before starting any effective work on the project it has been decided to 
+perform an analysis of the possible technologies available in the market, 
+performing a choice between them and exploiting their possible ``weak 
+points''. Proceeding with a top-down approach, the frameworks were chosen from 
+the one that had the biggest role in the test-bed to conclude with the 
+less-important components.
+
+\subsection{NFVI}
+
+A critical component, as the MANO one, is the NFVI. Without it, network 
+resources can't be allocated, and since our project was focused on dynamically 
+allocating containerized applications on cloud platforms, we decided to start 
+analyzing the right framework for this component first.
 
@@
 
+\subsubsection{Harbor}
+\label{chap:prjan:sec:tech:sub:MANO:sub:harbor}
+Started as a VNFM (Virtual Network Function Virtualization Manager) to offer a
+bridge for Openbaton to interact with Kubernetes, during the project it has been
+expanded to become a complete replacement of the original product, adding the
+possibility to manage Network Services and to update the routing tables
+contained in the Roulette Controller (the tool description can be found below
+in~\ref{chap:prjan:sec:tech:sub:SFC:sub:roulette}). Harbor doesn't have a web
+interface, but it has a RESTful API that allows users to interact with it using
+CRUD operations. It offers functionalities similar to Openbaton: possibility to
+save and retrieve VNF and NS definitions (basically, Kubernetes YAML
+deployments). When an NS is saved, a consistency check is performed to assure
+that all the VNFs have already been defined in Harbor. On an NS deployment, with
+a mechanism similar to the Openbaton one, all the associated VNFs are set up
+too, making the whole operation of launching a NS atomic from the user
+point-of-view. What Openbaton doesn't do is a smart resource utilization:
+Harbor, in fact, during a NS deletion, it checks if any related VNF remains
+orphan (which means without any NS using it). In that case the VNF will enter in
+a state of pruning: after five minutes, if the VNF has not be re-utilized by any
+other NS it will be deleted. This mechanism assure faster deployments in
+environments where NSs are continuously taken up and down. \todo{I've made an
+  Harbor implementation schema, I dunno if its useful to put it here}
+
 \subsection{SFC}
@@
 documentation when possible.
+
+\subsubsection{Roulette Controller}
+\label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
+Dynamic communications between VNFs would not be possible without a third
+component acting as a route controller. Roulette fulfills this role, providing a
+back-end where the single VNF can ask the next packet hop. Indeed, routes must 
+be
+submitted by another component, that we identified as Harbor (component
+description available above in
+sub-section~\ref{chap:prjan:sec:tech:sub:MANO:sub:harbor}). After a NS
+deployment, Roulette gets instructed with a new route describing a NS. Every NS
+has a proper ID (called \verb!sfid!), and thus when the classifier checks a
+packet typology, it labels that packet with the specific NS ID (for a detailed
+description about the SFC header see Section~\todo{Add section/chapter/paragraph
+  ref!}). Finally, thanks to this label, every VNF knows the packet's chain and
+it will correctly forward the it to the next VNF.
+
+\subsubsection{Astaire}
+As is possible to denote, a VNF doesn't only process the incoming packets, 
+but it has to manage the communications and the related forward operations. To 
+possibly ease part of the transmission problem and to make it easier to create 
+a 
+VNF without requiring the whole knowledge about SFC, it has been decided to 
+build a framework, Astaire, that could keep the SFC related details decoupled 
+from the core of the VNF, in other words, the entity that actually elaborates 
+the packet. Built using an event-based pattern, it presents hooks that allow 
+other software written in different languages from the Astaire one (e.g. Java), 
+to be called dynamically at runtime in order to process the packet. The result 
+of this elaboration gets returned to Astaire, that performs the beforehand 
+cited operations of VNFs communications and forwarding.
+
+\subsubsection{Ironhide}
+As described in chapter~\ref{chap:archimpl}, the whole framework need to have 
+and entry and an exit point. These, defined as \textit{Ingress} and 
+\textit{Egress} points, are performed by Ironhide. Ironhide acts like a proxy, 
+hiding the whole packet processing to the sender $S$ and the receiver $R$. When 
+a client $S$ performs a request to $R$, the packet actually goes through 
+Ironhide, who has an internal classifier that it is able to recognize the type 
+of connection and to choose the right NS the packet has to go through. This 
+process in fundamental since it keeps the connection between $S$ and $R$ up, 
+for both TCP and UDP based communications.
 
@@
 \end{figure}
+
+\subsubsection{Dnsmasq}
+\label{chap:prjan:sec:tech:sub:other:sub:dnsmasq}
+Dnsmasq is an open source project, started in 2001~\cite{dnsmasqweb}. Its
+features includes the possibility to serve as DNS, DHCP or router advertisement
+and network boot. Dnsmasq is designed to have a small footprint memory, and it
+is suited for medium to small network infrastructures. Along with the DNS
+server, it has the possibility to perform request caching, to speed up
+additional requests of the same type. On top of that, it provides different
+compilation options in order to eliminate unused features at compile time.

### 2018-11-29 res/sections/03-Inscription.tex
80 columns

@@
 \begin{flushright}{
-	\slshape
-	Do not shorten the morning by getting up late, or waste it in unworthy occupations or in talk; look upon it as the quintessence of life, as to a certain extent sacred.
-Evening is like old age: we are languid, talkative, silly.
-Each day is a little life: every waking and rising a little birth, every fresh morning a little youth, every going to rest and sleep a little death.} \\
-	\medskip
-  Arthur Schopenhauer
+	\slshape Do not shorten the morning by getting up late, or waste it in
+	unworthy occupations or in talk; look upon it as the quintessence of life, as
+	to a certain extent sacred. Evening is like old age: we are languid,
+	talkative, silly. Each day is a little life: every waking and rising a little
+	birth, every fresh morning a little youth, every going to rest and sleep a
+	little death.} \\ \medskip Arthur Schopenhauer
 \end{flushright}

### 2018-11-29 res/sections/04-Credits.tex

@@
 
-\noindent A big thank for my friends, whose power of cheering me up and make
-me smile saved myself different times from the risk of failing this journey.
-Thanks because you where always here when I needed you, and you always found the
-right words to say, or the right thing to do. Big part of my success is thanks
-to them, who have pushed me in places I would not be able to reach alone. In
-particular I would like to thank Giovanni, Giorgio, Michele, Gianluca and
-Umberto for the good times we end up spending every weekend. On top of that, I
-would like to say also thanks to Marco, Federico, Matteo, Mirko, Emanuele,
-Stefano, Sara, Lisa and Eduard that closely witnessed with me the good and the
-bad moments of the last two years of this journey. The time spent together with
-you guys will not be easily forgotten.
+\noindent A big thank for my friends, whose power of cheer me up and make me
+smile saved myself different times from the risk of failing this journey. Thanks
+because you where always here when I needed you, and you always found the right
+words to say, or the right thing to do. Big part of my success is thanks to
+them, who have pushed me in places I would not be able to reach alone. In
+particular I would like to thank Giovanni, Stefano, Giorgio, Michele, Gianluca
+and Umberto for the good times we end up spending every weekend. On top of that,
+I would like to say also thanks to Marco, Federico, Matteo, Mirko, Emanuele,
+Stefano (yes, again, you deserve it!), Sara, Lisa and Eduard that closely
+witnessed with me the good and the bad moments of the last two years of this
+journey. The time spent together with you guys will not be easily forgotten.
 

### 2018-11-29 res/sections/05-Abstract.tex

--- 2018-11-29 res/sections/05-Abstract.tex
@@
 \section*{\centering Abstract}
-The importance of Internet is growing every year. Mobile and wired traffic are
-continuously surging, asking for better performances. On this day, ISPs rely on
-optimized, ad-hoc and often in-hardware nework solutions. These network
-functions are costly and require time to be deployed and mantained. In this
-context, upcoming virtualization technolgies like hypervisor-based and
-container-based are providing the possibility to create virtualized network
-functions, that have the ability to be up and running in a matter of seconds.
-VNFs are then managed and orchestrated by a specific application, the MANO that
-handles their lifecycle.
-
-In this thesis we are going to propose an Infrastructure Management and
-Orchestrator plane Proof-of-Concept, based on the container virtualization
-technology and on the 3GPP Release 15 specifications. In the current PoC, Docker
-and the container orchestrator Kubernetes play a fundamntal role.
+Network resource management and service optimization are two basic
+functionalities widely adopted by network operators. However, provisioning these
+two key capability is becoming increasingly difficult considering the high
+number of services delivered through today networks. The usual approach of
+optimization through middleboxes deployed in the end-to-end path between
+user/service is proving incapable in keeping up with the current pace of change.
+On the other side, advances in virtualization and softwarization techniques
+could provide the much needed flexibility, serving as enablers for an
+end-to-end, softwarized network management plane. In this context, taking
+inspiration from the recent proposal of the 3GPP Release 15, we propose a
+two-layer service orchestration architecture exploiting container technology. In
+specifics, we provide a proof of concept implementation of an infrastructure
+Management and Orchestration (MANO) plane, replying on the Kubernetes container
+orchestration framework.
 \par

### 2018-11-29 res/sections/07-Introduction.tex

@@
  \item Chapter~\ref{chap:prjan} explains the analysis performed on the
-   requirements provided and the design choices performed;
+   requirements provided and the overall project roadmap;
  \item Chapter~\ref{chap:archimpl} provides an in-depth view of the high-level
    architectural implementation;
- \item Chapter~\ref{chap:vnf_ns_impl} gives a brief explanation of the VNF and NS
+ \item Chapter~\ref{chap:vnf_ns_impl} gives a brief explanation of the SFC
    implementation;
  \item Chapter~\ref{chap:test} describes the tests done on the platform;
+ \item Chapter~\ref{chap:fw} illustrates possible future works;
  \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.

### 2018-11-29 res/sections/08-Background.tex

@@
 exposed via a virtualized interface, and that can be monitored using specific 
-APIs
+APIs;
 \item \textbf{business enablement layer}: where a library of functions and
-  deployment is contained, and its configuration is accessible via APIs
+  deployment is contained, and its configuration is accessible via APIs;
 \item \textbf{business application layer}: layer that contains specific
-  applications and services of the operator
+  applications and services of the operator.
 \end{itemize}
@@
   \item Analysis of the applicability of current and new Internet protocols in
-  the proposed VNF-PEP architecture
-  \item Implementation of a VNF-PEP prototype
-  \item Building of a PoC test platform
-  \item VNF-PEP validation and performance tests on 5G use-cases
-  \item Demonstration test-bed management
+  the proposed VNF-PEP architecture;
+  \item Implementation of a VNF-PEP prototype;
+  \item Building of a PoC test platform;
+  \item VNF-PEP validation and performance tests on 5G use-cases;
+  \item Demonstration test-bed management.
 \end{enumerate}

### 2018-11-29 res/sections/79-Conclusions.tex
Add personal conclusions

@@
 possible innovations where connectivity will play an essential role.
+
+\section*{Personal conclusions}
+
+Overally speaking, I have never designed and though a software architecture of
+these dimensions. The multiple attempts and issues I faced during the thesis
+taught me more that I would have tought, showing that, in life, you never end to
+learn something new. Its for this reason that I think this project is something
+more than a simple thesis, because I was put in front of challenges that seemed
+over my possibilities at first, but that revealed to be no more than any other
+common problem after having spent time analyzing, studing and understanding
+them. At the end, I consider every second of my time spent on this thesis worth
+it, and it gave me the confidence that every limit you face in life can be
+broken down.

### 2018-11-29 res/sections/79-Conclusions.tex
Fix typos

--- 2018-11-18 res/sections/79-Conclusions.tex
@@
 possible innovations where connectivity will play an essential role.
+
+\section*{Personal conclusions}
+
+Generally speaking, I have never designed and though a software architecture of
+these dimensions. The multiple attempts and issues I faced during the thesis
+taught me more that I would have though, showing that, in life, you never end to
+learn something new. Its for this reason that I think this project is something
+more than a simple thesis, because I was put in front of challenges that seemed
+over my possibilities at first, but that revealed to be no more than any other
+common problem after having spent time analyzing, studyng and understanding
+them. At the end, I consider every second of my time spent on this thesis worth
+it, and it gave me the confidence that every limit you face in life can be
+broken down.

### 2018-11-29 res/sections/79-Conclusions.tex
Fix more typos

@@
 over my possibilities at first, but that revealed to be no more than any other
-common problem after having spent time analyzing, studyng and understanding
+common problem after having spent time analyzing, studying and understanding
 them. At the end, I consider every second of my time spent on this thesis worth

### 2018-11-29 res/sections/08-Background.tex
Fix a bunch of typos

@@
 along an entire continent for example, but it could decide to have only a 
-single centralized business application deployment that manage all the other 
+single centralized business application deployment that manages all the other 
 layers resources.
@@
 the project is to research and include this technology in the usual VNF
-ecosystem to include without change satellite communications to the SFC. The
-architecture of the systems will be based on clusters of containers whose
-management and orchestration will follows the ETSI MANO architecture and SDN
-principles. Thanks to VNF-PEP will be possible to dynamically take decisions on
-how to optimize a certain communication.
+ecosystem, with the final aim to include satellite communications in the SFC.
+The architecture of the system will be based on a container cluster which
+management and orchestration will follow the ETSI MANO architecture and SDN
+principles. Thanks to VNF-PEP, it will be possible to dynamically take decisions
+on how to optimize certain communications.
 
@@
 virtual machines with different configurations, without bringing any changes to
-the host machine. Indeed, the virtualization concepts refer to the property of
-a machine to be virtualized. To achieve this, the components that forms a common
+the host machine. Indeed, the virtualization concepts refer to the property of a
+machine to be virtualized. To achieve this, the components that forms a common
 computer need to be virtualized too. Thus, these assets are recreated in this
 virtualized manner, e.g. storage devices and computer network
-resources~\cite{liu2014research}. At the time writing, there are two common
+resources~\cite{liu2014research}. At the time of writing, there are two common
 virtualization options: the virtualization through Hypervisor (defined also as
-hypervisor-based) and the container virtualization.
+hypervisor-based) and the container through virtualization.
 
@@
 flexibility and manageability. This network separation creates two different
-entities, that becomes decoupled: who provides the infrastructure (commonly
+entities, that become decoupled: who provides the infrastructure (commonly
 called InPs) and who provides the services (SPs). A virtual network can be
@@
 \label{chap:background:sec:sdn:sub:cp}
-The SDN controllers, as just said in the previously, receives requests from the
+The SDN controllers, as just said in the previously, receive requests from the
 SDN applications. Since the controller is centralized, it can therefore exploit
@@
 
-OpenFlow is an open communication protocols that enables the possibility to
+OpenFlow is an open communication protocol that enables the possibility to
 actually program networks, offering the possibility to change the flow-tables

### 2018-11-29 res/sections/08-Background.tex
Minor fixes

@@
 manage the slice~\cite{ordonez2017network}. Slices have to respond to a defined
-set o requisites, first of all the possibility to being created by a template,
+set of requisites, first of all the possibility to being created by a template,
 defined as Network Slice Template (NST). After the NS and the related resources
@@
 \begin{description}
-\item[Virtualized Network Function] that is a software implementation able to fun
-  a network function over a NFVI;
+\item[Virtualized Network Function] that is a software implementation able to
+  run a network function over a NFVI;
 \item[NFVI Infrastructure (NFVI)] that includes different physical resources
@@
   \centering
-  \includegraphics[scale=0.7]{etsimano}
+  \includegraphics[scale=0.8]{etsimano}
   \caption[ETSI NFV MANO framework]{ETSI NFV MANO
@@
 main components, such as the Virtual Infrastructure Manager (VIM), the VNF
-Manager (VNFM) and the NFV Orchestrator (NFVO). The MANO organize and holds
+Manager (VNFM) and the NFV Orchestrator (NFVO). The MANO organizes and holds
 important data, regarding the Network Service (NS) catalog, the list of the VNFs
-definitions (VNF catalog) and the current NFV instances and the NFVI resources.
+definitions (VNF catalog), the current NFV instances and the NFVI resources.
 
@@
 communicate with the NFV infrastructure, and to control its physical and virtual
-resources. A VIM should has the possibility to control one or more of NFVI. In
-general a VIM can be in charge of control a defined set of
+resources. A VIM has the possibility to control one or more of NFVI. In general,
+a VIM can be in charge of control a defined set of
 NFVI~\cite{mijumbi2016management}
@@
 Data repositories can be seen as a sort of databases, holding information about
-the VNF definitions, the VNF instances with the respective lifecycle status, and
-list of the Network Services defined and the resources available and occupied of
-the NFVI.
+the VNF definitions, the VNF instances with the respective lifecycle status, a
+list of the Network Services defined and the resources rispectively available
+and occupied of the NFVI.
 
@@
 \paragraph*{Service Function}
-A single element of an SFC is called Service Function (SF), and its duty is to
+A single SFC element is called Service Function (SF), and its duty is to
 elaborate the packets incoming and to forward them to the next receiver. SFs, as

### 2018-11-30 res/sections/08-Background.tex
Fix typo

@@
 the VNF definitions, the VNF instances with the respective lifecycle status, a
-list of the Network Services defined and the resources rispectively available
+list of the Network Services defined and the resources respectively available
 and occupied of the NFVI.

### 2018-11-30 res/sections/10-ProjectAnalysis.tex
Add new logos in technology section

@@
 complete cloud solution.
-\begin{figure}[t]
+\begin{figure}[h]
  \centering \includegraphics[scale=0.58]{openstack_logo}
@@
 respectively, orchestrate containers in a single or a clustered system.
-\begin{figure}[t]
+\begin{figure}[h]
  \centering \includegraphics[scale=0.7]{docker_logo}
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=1]{dockercompose_logo}
+  \caption[Docker-Compose logo]{Docker-Compose logo}
+\end{figure}
+
 \subsection{Other technologies evaluated}
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=1]{dockerswarm_logo}
+  \caption[Docker Swarm logo]{Docker Swarm logo}
+\end{figure}
+
 \subsubsection{OpenVSwitch}
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.5]{ovs}
-  \caption[OpenVSwitch logo]{OpenVSwitch logo}
-\end{figure}
 Started in 2009~\cite{ovsGit}, the OpenVSwitch project aims to create an
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{ovs}
+  \caption[OpenVSwitch logo]{OpenVSwitch logo}
+\end{figure}
+
 \subsubsection{Floodlight}
@@
 routers have to register to the controller in order to gain routing rules.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{floodlight_logo}
+  \caption[Floodlight logo]{Floodlight logo}
+\end{figure}
 
@@
 Performing this operation requires a deep knowledge of the framework though.
+
 \begin{figure}[h]
@@
 compilation options in order to eliminate unused features at compile time.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.25]{dnsmasq_logo}
+  \caption{Dnsmasq logo}
+\end{figure}

### 2018-11-30 res/sections/08-Background.tex
Fix typos and english

@@
 \paragraph*{Architectural view}
-About the architectural view of SFCs, proposal defines several assumptions on
+About the SFCs architectural view, the proposal defines several assumptions on
 them, starting from considering every element of the chain an opaque processing
 unit. Another important assumption is that there is no global or standard SF
-chaining logic~\cite{rfc7665}. SFC can be easily seen as a graph, because some
-SF can have multiple address to forward the elaborated data, thus creating what
-is defined as \emph{fork} that ends up forming different branches. This could
-lead to loops, that must be treated consequently.
+chaining logic~\cite{rfc7665}. An SFC can be easily seen as a graph, because
+some SF can have multiple address to forward the elaborated data, thus creating
+what is defined as \emph{fork} that ends up forming different branches. This
+could lead to loops, that must be treated consequently.
 
@@
 the content, and it is left free to be filled with custom metadata. The only
-requirement is to set this filed with zeroes when is not used.
+requirement is to set this field with zeroes when is not used.
 

### 2018-11-30 res/sections/09-RelatedWork.tex
Fix typos and english

@@
 
-One of the first proposal that take advantage of container in to take on network
-edges NFV is~\cite{cziva2017container}. The authors present the work, based on
-previous definition of the platform defined in~\cite{cziva2015container} to
-orchestrate the virtual function and to deploy it as close as possible to the
-data sender to classify traffic and choose the appropriate functions
-immediately.
+One of the first proposals that takes advantage of the container technology and
+to take to the network edges the NFV is~\cite{cziva2017container}. The authors
+present the work, based on previous definition of the platform defined
+in~\cite{cziva2015container} to orchestrate the virtual function and to deploy
+it as close as possible to the data sender to classify traffic and choose the
+appropriate functions immediately.
 
@@
 demonstrate how end-to-end services can be provisioned using SDN networking, in
-particular with controllers like such as OpenDaylight, Pox, Ryu ONOD and
-Floodlight. The last one was studied by us too during the technologies analysis,
-and a description is available in
+particular with controllers like OpenDaylight, Pox, Ryu ONOD and Floodlight. The
+last one was studied by us too during the technologies analysis, and a
+description is available in
 Section~\ref{chap:prjan:sec:tech:sub:other:sub:floodlight}.
 
-Compared to our solution, these implementations does not present an unified
+Compared to our solution, these implementations do not present an unified
 framework that gives the possibility orchestrate different network slices
@@
 the integrated Role Base Access Control system (RBAC) and other security policy
-bundled in by default in Kubernetes.
+bundled by default in Kubernetes.
 
 \section{Provision capabilities}
-We noted how, in some other NFV/ETSI-MANO proposals like~\cite{soares2015toward}
-and~\cite{abujoda2015midas}, the implementations relays on a static provision
-system, where the SFC paths are somehow defined at deploy time, and can't change
-without decommissioning, editing and deploying the chain again. Our proposal,
-instead, is able to support chain editing on-the-fly, allowing NFV to
-dynamically change the SFP.
+We noticed how, in some other NFV/ETSI-MANO proposals
+like~\cite{soares2015toward} and~\cite{abujoda2015midas}, the implementations
+relay on a static provision system, where the SFC paths are somehow defined at
+deploy time, and can not change without decommissioning, editing and deploying
+the chain again. Our proposal, instead, is able to support chain editing
+on-the-fly, allowing NFV to dynamically change the SFP.
 

### 2018-11-30 res/sections/10-ProjectAnalysis.tex
Fix typos and english

@@
 
-As in chapter~\ref{chap:intro}, the creation of a full ETSI MANO compliant
-architecture, following the suggestions described in RFC 7665, require the
-integration of multiple tools (some of them have been already briefly introduced
-in~\ref{chap:intro}). The tools were chosen after their analysis, where pro and
-cons have been studied and compared to the necessity of this thesis. Since the
-project has been developed with the cloud provided by the University of Padua,
-Openstack (which a product analysis can be found later
+As described in chapter~\ref{chap:intro}, the creation of a full ETSI MANO
+compliant architecture, following the suggestions described in RFC 7665,
+requires the integration of multiple tools (some of them have been already
+briefly introduced in~\ref{chap:intro}). The tools were chosen after their
+analysis, where pro and cons have been studied and compared to the necessity of
+this thesis. Since the project has been developed with the cloud provided by the
+University of Padua, Openstack (which a product analysis can be found later
 in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
@@
 using bare-metal resources to provide common cloud services as object storage,
-virtual machine deploy, virtual networking. Furthermore its modularity offers
+virtual machine deploy and virtual networking. Furthermore its modularity offers
 the possibility to add additional components, even proprietary, to achieve a
@@
 possible to run Docker containers without bothering about hardware constraints.
-Nodes can have specific roles and specialization, all of them managed by one or
+Nodes can have specific roles and specializations, all of them managed by one or
 many master nodes, that actually manage the orchestration. On top of that,
@@
 change the virtualization software in use (i.e. from Container-based technology
-to Virtual machine software).
+to Virtual machine one).
 \begin{figure}[h]
@@
 project requisite that could not be easily changed, it has been decided to opt
-for the creation of a smaller and simpler clone that offered a full integration
-with Docker and Kubernetes.
+for the creation of a smaller and simpler alternative that offered a full
+integration with Docker and Kubernetes.
 
 \subsection{Available SFC solutions}
-The tools analysis revealed little about already existent VNF or NS
-implementations, so it was necessary to fill the gap. Since this one of our
-thesis goals, we developed a complete solution to create a complete SFCs, trying
-to standardize our vocabulary as much as possible, and giving basic
-documentation when possible.
+The tools analysis revealed little about already existent VNF or NS reusable
+framework implementations, so it was necessary to fill the gap. We then
+developed a complete solution to create complete SFCs, trying to standardize
+our vocabulary as much as possible, and giving basic documentation.
 
@@
 of the tasks that should be performed by hand when launching one or multiple
-docker containers. It's particularly useful when more containers have to
+docker containers. It is particularly useful when more containers have to
 interact together, because all the services rules are described in a simple YAML
@@
 packages for the most common distros) but it can be deployed on Docker
-containers too. It's designed to work with cloud providers like Amazon or Google
-Compute engine, but it offers compatibility with on-premise cloud solutions like
-Openstack, where it deploys virtual machines using the API the platform offers.
-It can store VNF configurations saved as Topology and Orchestration
-Specification for Cloud Applications (TOSCA) Yet Another Markup Language (YAML)
-and it can handle multiple Point of Presence (PoP). It offers a VNF lifecycle
-management out-of-the-box, with the possibility to customize it based on the
-user needs. Its modular design allows developers to change parts of the codebase
-with custom ones, making the product really flexible to different platforms.
-Performing this operation requires a deep knowledge of the framework though.
+containers too. It is designed to work with cloud providers like Amazon or
+Google Compute engine, but it offers compatibility with on-premise cloud
+solutions like Openstack, where it deploys virtual machines using the API the
+platform offers. It can store VNF configurations saved as Topology and
+Orchestration Specification for Cloud Applications (TOSCA) Yet Another Markup
+Language (YAML) and it can handle multiple Point of Presence (PoP). It offers a
+VNF lifecycle management out-of-the-box, with the possibility to customize it
+based on the user needs. Its modular design allows developers to change parts of
+the codebase with custom ones, making the product really flexible to different
+platforms. Performing this operation requires a deep knowledge of the framework
+though.
 

### 2018-11-30 res/sections/15-ArchitecturalImplementation.tex
Fix English typos

@@
 
-In this chapter we will dig into the actual architectural proposal, seeing
-all the problems related to it and how they were solved. We will show all the
-various attempts made, and we will discussion about the technological issue and
+In this chapter we will dig into the actual architectural proposal, seeing all
+the problems related to it and how they were solved. We will show all the
+various attempts made, and we will discussion about the technological issues and
 changes we had to perform.
@@
   \centering
-  \includegraphics[scale=0.5]{Architecture}
+  \includegraphics[scale=0.6]{Architecture}
   \caption[Architecture schema depicting the expected
@@
 framework is to be considered service and/or application agnostic. This excludes
-this framework from the burden of being application aware for NS configuration
-and management and for the VNF application layer configuration and management.
+this framework from the burden of being application aware for any possible NS
+configuration and management, and for the VNF application layer configuration
+and management.
 
@@
 task to perform service lifecycle management and monitoring via the API offered
-by the ETSI MANO. In this context, is possible to identify two different type of
-orchestrator, as represented in
+by the ETSI MANO. In this context, is possible to identify two different types
+of orchestrators, as represented in
 Figure~\ref{chap:archimpl:sec:fistattempt:img:architecture}:
@@
 \end{itemize}
-Both the layer would use Openbaton. The messages would be exchanged through
+Both the layers would use Openbaton. The messages would be exchanged through
 RabbitMQ, using the Advance Message Queuing Protocol (AMQP).
@@
 \paragraph*{Low-layer orchestrator}
-The low-layer orchestrator jobs corresponds to the one defined in the ETSI MANO
+The low-layer orchestrator job corresponds to the one defined in the ETSI MANO
 definition, with the role of managing and orchestrating VNFs and NSs. The
@@
 In the following section we are going to briefly talk about the role of the most
-peculiar components in the defined architecture. A description of the tasks that
-the components have to abide in the ETSI MANO architecture is available in
-Section~\ref{chap:archimpl:sec:secondattempt:sec:k8s}.
+peculiar components in the defined architecture. A detailed description of the
+tasks that the components have to abide in the ETSI MANO architecture is
+available in Section~\ref{chap:archimpl:sec:secondattempt:sec:k8s}.
 \begin{description}
@@
 \item[Network service management] in this architectural proposal, it corresponds
-  the the brain of the infrastructure MANO, where service instantiation and
+  to the brain of the infrastructure MANO, where service instantiation and
   lifecycle management happens. This part interacts with the VNFM/VIM and with

### 2018-11-30 res/sections/12-Requirements.tex

@@
 \section{Requirements}\label{chap:prjan:sec:req}
-In the following section are described the main requirements of this projects.
+In the following section are described the main requirements of this project.
 
@@
 1.
-Automatic creation and configuration of NFVI                         & It must be possible to create the environment on which the SFC will be deployed in an automatic manner.
-The system must support Docker container and exploit must be deployed on a clustered environment.
+Automatic NFVI creation and configuration                            & It must be possible to create the environment in which the SFC will be deployed in an automatic manner.
+The system must support Docker container and must be deployed on a clustered environment.
 \\
@@
 5.
-VNF fault management                                                 & The system must be able to recover a container that has fault automatically.
+VNF fault management                                                 & The system must be able to recover a faulty container automatically.
 \\
@@
 7.
-SFC definition saving                                                & The system must be able to save a definition of a chain                                                                                                                                        \\
+SFC definition saving                                                & The system must be able to save a definition of a chain.
+\\
 8.
-Deployment/stop/update of a chain from definition                    & The system must be able to allow the deployment of a chain from a saved definition, stop a running SFC and update it.
+Chain definition deploy/stop/update                    & The system must be able to allow the deployment of a chain from a saved definition, stop a running SFC and update it.
 \\ \hline

### 2018-11-30 res/sections/15-ArchitecturalImplementation.tex
Fix English typos

@@
 low-layer MANO holds different data repositories where VNF and NS definitions
-are stored, various monitoring tools and logs and the network services
-repository. The low layer MANO exposes a set of APIs to the upper layer (related
-to handling services) MANO, that is able to configure and manage the services.
-In other words, the service MANO has the possibility to configure and launch
-services to the low-layer orchestrator, providing the necessary information.
+are stored, alongside with various monitoring tools, logs and network services
+repositories. The low layer MANO exposes a set of APIs to the upper layer
+(related to handling services) MANO, that is able to configure and manage the
+services. In other words, the service MANO has the possibility to configure and
+launch services to the low-layer orchestrator, providing the necessary
+information.
 
@@
   manage different functional aspects of the control/data plan.
-\item[Network service management] in this architectural proposal, it corresponds
-  to the brain of the infrastructure MANO, where service instantiation and
-  lifecycle management happens. This part interacts with the VNFM/VIM and with
-  the RabbitMQ middleware. In this context, the NFVO is involved in chaining the
-  required VNFs in order to create the service delivery path. Inside this
-  component is possible to identify the Service Lifecycle Management, that
-  listens to status report coming from the VNFM and thus generating logs and
-  monitoring information.
+\item[Network service management] in this architectural proposal, the Network
+  service management corresponds to the brain of the infrastructure MANO, where
+  service instantiation and lifecycle management happens. This part interacts
+  with the VNFM/VIM and with the RabbitMQ middleware. In this context, the NFVO
+  is involved in chaining the required VNFs in order to create the service
+  delivery path. Inside this component is possible to identify the Service
+  Lifecycle Management, that listens to status report coming from the VNFM and
+  it generates logs and monitoring information.
 \item[Monitoring and Information Management] this block gathers data from the
@@
 up Openstack as our VIM while setting Openbaton as our MANO. Then, Openbaton
-would have Openstack registered as PoP, where TOSCA definition would be launched
-on Openstack that, instead of deploying hypervisor based virtual machines, it
-would have deployed container instances. The VNFs lifecycles would have been
-controller through an Element Management System (EMS), transmitting the various
-VNF states via the message broker RabbitMQ. Is possible to see the interactions
-between elements in a minimum schema available in
+would have registered Openstack as PoP, where TOSCA definitions would have been
+launched on Openstack that, instead of deploying hypervisor based virtual
+machines, it would have deployed container instances. The VNFs lifecycles would
+have been controlled through an Element Management System (EMS), transmitting
+the various VNF states via the message broker RabbitMQ. It is possible to see the
+interactions between elements in a minimum schema available in
 Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.
@@
 
-Our first step was to try to use Openstack to create a container orchestration. 
-The reason behind this approach was simple: Openbaton, on paper, comes with an 
-out-of-the-box support for Openstack. In fact, it natively supports Openstack 
+Our first step was to try and use Openstack to create a container orchestration.
+The reason behind this approach was simple: Openbaton, on the paper, comes with
+an out-of-the-box support for Openstack. In fact, it natively supports Openstack
 PoPs, giving to us a great possibility to reuse the two frameworks and to ease
@@
 in the local machine, and to leverage a Openstack environment able to perform
-the same operations of the production-grade installation. Thus, we starting 
-configuring the Devstack installation, installing first of all the Nova 
-module, that is required to perform computer operations such as the launch of 
-virtual machines, and lately we installed Tacker too.
+the same operations of the production-grade installation. Thus, we started
+configuring the Devstack installation, installing first of all the Nova module,
+that is required to perform computing operations such as the launch of virtual
+machines, and lately we installed Tacker too.
 
-\paragraph*{Taker}
+\paragraph*{Tacker}
 \begin{figure}[t]
@@
 Tacker is a VNFM and a NFVO. It is an official Openstack module, and it has the
-ability to deploy, edit, stop and do periodic health checks VNF directly on it.
+ability to deploy, edit, stop and do periodic VNF health checks directly on it.
 The developers claim it has the possibility to make deployments to platforms
 compatible with Openstack too. It supports VNF definition through the TOSCA
-standard, and a proposal for SFC integration has been proposed in
+standard, and a proposal for SFC integration has been made in
 2015~\cite{tackerOpenstackwiki}.
@@
 configuration is very high. In particular, it required from 20 minutes to 1 
-hour to install Devstack we a particular configuration, without knowing if the 
+hour to install Devstack with a particular configuration, without knowing if the 
 final installation was working or not. This trial-and-error approach consumed a 
@@
 \begin{itemize}
-  \item Normal installation. Openbaton offers official Ubuntu repositories that,
-    added into the system, offer a very smooth installation via the
+  \item Normal installation. Openbaton delivers official Ubuntu repositories
+    that, added into the system, offer a very smooth installation via the
     \verb!aptitude! command line utility, with the only cons of having to wait
     for the automatic Openbaton download, installation and configuration;
-  \item Using Vagrant. Vagrant is a software that helps the deployment of
-    software, and it offers drivers to deploy virtual virtual machines too;
+  \item Using Vagrant. Vagrant is a software that helps the software deployment,
+    and it offers drivers to deploy virtual virtual machines too;
   \item Docker Compose. At the time of our study, the Docker Compose had three
@@
 
-Now we are going to do a short description about every installation/deploy method.
+A short description about every installation/deploy method will be briefly
+illustrated.
 
 \subparagraph*{Installation via repository}
-The installation via repositories granted us the possibility to select only the
+The installation via repository granted us the possibility to select only the
 necessary parts we needed. The back of the medal of this approach was the
@@
 customization and component management. The possibility to remove or add
-Openbaton Modules dynamically become later fundamental to us. On top of that,
+Openbaton modules dynamically became later fundamental to us. On top of that,
 the Vagrant deployment uses a hypervisor virtualization system (based on Virtual
@@
 On top of that, this method let us to manage the Openbaton version freely,
-having more control on update especially during the development of our
+having more control on updates especially during the development of our
 components. Additionally, using this Docker Compose configuration allowed us to
-seamlessly integrate the monitoring tools Zabbix, in order to have the
+seamlessly integrate the monitoring tool Zabbix, in order to have the
 possibility to graphically see the data fetch from Openbaton about the system
 state. Finally, a Docker based solution seemed to be the best approach since all
-of our components were meant to be put inside a Docker container.
+of our components were meant to be put inside Docker containers.
 
@@
 registry. The impossibility to easily set the containers ports and the lack of
-some support in between the TOSCA definition and the Openstack Zun plugin pushed
-us to drop Openstack as a possible solution and to start looking for another
-valid VIM component.
+some support between the TOSCA definition and the Openstack Zun plugin pushed us
+to drop Openstack as a possible solution and to start looking for another valid
+VIM component.
 
@@
 such as Docker Compose. The installation and set-up is pretty straightforward,
-and the configuration is almost automatically handed by Docker itself. The
+and the configuration is almost automatically handled by Docker itself. The
 minimum number of nodes in order to create a proper Docker Swarm cluster is of
-free nodes, where one must be a manager node (designed to orchestrate the other
-nodes in the cluster), while the other can be simple workers. Multiple managers
+three, where one must be a manager node (designed to orchestrate the other
+nodes in the cluster), while the others can be simple workers. Multiple managers
 can coexist in the same cluster, as long as only one effectively operates. The
@@
 \subsection{Kubernetes}
-Kubernetes, as already described in the technology section, is a orchestrator
+Kubernetes, as already described in the technology section, is an orchestrator
 created by Google able to create a software abstraction from the underlying
 virtualization technology. This enables Kubernetes to create an integration with
-the Docker technology. Kubernetes has a architecture based on plug-ins, giving
-the possibility to the users to change part of Kubernetes itself, such as the
-networking routing strategy or the persistence layer. With this flexibility
-Kubernetes offered to us what Docker Swarm was not able to give. Additionally,
-we had the possibility to set custom Domain Name System (DNS) lookups and
-distribute other internal components in order to avoid requests bottleneck.
+Docker. Kubernetes has an architecture based on plug-ins, giving the possibility
+to the users to change part of Kubernetes itself, such as the networking routing
+strategy or the persistence layer. With this flexibility Kubernetes offered to
+us what Docker Swarm was not able to give. Additionally, we had the possibility
+to set custom Domain Name System (DNS) lookups and distribute other internal
+components in order to avoid requests bottleneck.
 
@@
 We initially deployed from the internal Openstack of the University four nodes
-with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines. We
-performed the installation using the official Kubernetes Repositories. After the
-initial installation, we obtained a configuration similar to the one illustrated
-in Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
+with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines (that
+we later swapped with CentOS 7). We performed the installation using the
+official Kubernetes Repositories. After the initial installation, we obtained a
+configuration similar to the one illustrated in
+Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
 possible to denote different components on the master node and on the minions,

### 2018-11-30 res/sections/15-ArchitecturalImplementation.tex
Fix English typos

@@
 we later swapped with CentOS 7). We performed the installation using the
-official Kubernetes Repositories. After the initial installation, we obtained a
+official Kubernetes repositories. After the initial installation, we obtained a
 configuration similar to the one illustrated in
@@
 \item[API Server] Fist of all, the User interacts with an API Server located in
-  the master node. This APIs are developed using the Swagger tools, ensuring
+  the master node. These APIs are developed using the Swagger tool, ensuring
   consistency and uniformity. From the image, is possible to see that the API
   Server communicates with all the main components of the architecture.
-\item[etcd] The \verb!etcd! component communicates via the gRCP protocol, a
+\item[etcd] The \verb!etcd! component communicates via the gRCP protocol, an
   efficient Remote Procedure Call (RPC) protocol to serve data to the API
@@
 \item[Scheduler] The Scheduler regulates the Pods distribution in the minions,
-  having a knowledge of the current cluster typologies and of the current
-  resources available. On top of that, is able to handle hardware constraints,
-  defined policy, and defined class of QoS.
+  having a knowledge of the current cluster typology and of the current
+  resources available. On top of that, it is able to handle hardware
+  constraints, defined policy, and defined class of QoS.
 \end{description}
 
-All of the previously descripted components are hide behind the API server,
-which is the only components the final user can really interact with. The
-protocol between the API server and the final user used is a common JSON data
-object interchange. To make the Kubernetes more human-friendly, Kubernetes comes
-with a command line interface (CLI) tool, called \verb!kubectl!, that allows to
-easily perform deployments and admin operations on the cluster without having a
-full knowledge of the official API.
+All of the previously descripted components are hidden behind the API server,
+which is the only component the final user can really interact with. The
+communication protocol between the API server and the final user used is a
+common JSON data object interchange. To make the Kubernetes more human-friendly,
+Kubernetes comes with a command line interface (CLI) tool, called
+\verb!kubectl!, that allows to easily perform deployments and admin operations
+on the cluster without having a full knowledge of the official API.
 
@@
   every minion, and \emph{de facto} acts like a local manager. It receives YAML
-  or JSON definition from the API Server of Pod specifications (PodSpec) that
+  or JSON definitions from the API Server of Pod specifications (PodSpec) that
   have to be deployed in the node where is running. Then, its role is to ensure
-  that the Pod deployed are up and healthy. This excludes possible
+  that the Pods deployed are up and healthy. This excludes possible
   interference with other virtualized deployments.
@@
   specification, which seeks to provide a standard on how to interact and run
-  the container images on the OS.
+  the container images on the host OS.
 \item[Networking] This part aims to provide to the container runtime network
@@
   a part of the same network, even if Pods could be dislocated in other minions
-  working between different subnets or completely different networks. On top of
-  that, dynamic port allocation has to be kept in consideration too. The
+  working between different subnets or completely different network domains. On
+  top of that, dynamic port allocation has to be kept in consideration too. The
   Kubernetes networking model focus on three fundamental
@@
   inside a Pod can reach each other using the \verb!localhost! address. In
-  Kubernetes, several networking implementation exists, like Flannel, Google
+  Kubernetes, several networking implementations exist, like Flannel, Google
   Compute Engine (GCE), Kube-route, Project Calico~\cite{k8snetworkingwiki}. For
@@
 
-  The communications between this component and the Kubelet one is carried
+  The communications between this component and the Kubelet one are carried
   through the Cloud Native Interface (CNI) protocol, that consists of a set
@@
 illustrated in Figure~\ref{chap:archimpl:sec:secondattempt:img:schema1}:
-Openbaton, that had the TOSCA definitions, had to somehow convert these
-definition in a suitable YAML for Kubernetes, that had in turn to elaborate it
-and schedule to one minion in the cluster. The monitoring and management would
-remain untouched, as we planned to create an image with the integrated EMS tools
-that provided a seamless integration with Docker container. With this setting,
-the only thing remained was to solve the issue of converting TOSCA definitions
-in YAML one. The first thing we started thinking about was how to perform the
-conversion. This topic was quite important because the TOSCA definition does not
-really consider the possibility to have an alternative virtualization system
-apart from the hypervisor one. This causes the TOSCA definition to be, in our
-opinion, too much focused on the resources usage and regarding granularity
-aspects such the deployment flavour and the kind of virtual link (concept that
-does not really exists in container based systems) to instantiate. Nonetheless,
-we performed a mapping between most of the TOSCA keywords with some of the YAML
-definition, in a way that a TOSCA configuration file was able to generate a
-valid YAML deployment. After that, we started thinking instead who, in the
-current architecture, had the task to perform this operation. Initially we
-though that Kubernetes itself could have been a perfect candidate. From a
+Openbaton, that usually handles the TOSCA definitions storing them in its
+database, had to somehow convert these definition in a suitable YAML for
+Kubernetes, that had in turn to elaborate it and schedule to one minion in the
+cluster. The monitoring and management would remain untouched, as we planned to
+create an image with the integrated EMS tools that provided a seamless
+integration with Docker container. With this setting, the only thing to solve
+was how to convert TOSCA definitions in YAML one. The first thing we started
+thinking about was how to perform the conversion. This topic was quite important
+because the TOSCA definition does not really consider the possibility to have an
+alternative virtualization system apart from the hypervisor one. This causes the
+TOSCA definition to be, in our opinion, too much focused on the resources usage
+and granularity aspects such the deployment flavour and the kind of virtual link
+(concept that does not really exists in container based systems) to instantiate.
+Nonetheless, we performed a mapping between most of the TOSCA keywords with some
+of the YAML definitions, in a way that a TOSCA configuration file was able to
+generate a valid YAML deployment. After that, we started thinking instead who,
+in the current architecture, had the task to perform this operation. Initially
+we thought that Kubernetes itself could have been a perfect candidate. From a
 received TOSCA, a Kubernetes component should have validated and translated the
@@
 Kubernetes API Server. Another feasible approach could have been to create a
-Openbaton plugin in order to take the TOSCA to deploy and convert it before
-submitting it to Kubernetes. This solution had great advantages in terms of
+Openbaton plugin in order to convert the TOSCA definition before submitting it
+to Kubernetes for the deployment. This solution had great advantages in terms of
 responsibility separation: in this way only the MANO had to know, for a VIM
-based on Kubernetes, that a conversion needed to be performed. From a Kubernetes
+based on Kubernetes, that a conversion was needed. From a Kubernetes
 point-of-view, nothing would have changed, since all the configuration
@@
 VNFM, because we discovered there were programmatic dependencies between the two
-Openbaton standard components. After we identified the VIM driver and the VNFM
-as components to be rewritten and to be adapted to the Kubernetes VIM, we
-studied the original components code. We discovered that regarding these
-components the documentation was very poor, especially regarding the code. On
-top of that, the program interfaces given by the developers did not match our
-expectations: we thought to find interfaces independent from the type of the
-virtualization technology, instead, we found that they were strictly coupled
-with frameworks like Openstack or cloud vendors like Google Compute engine or
-Amazon AWS. These reasons, with the fact that after two weeks we were not able
-to get any working prototype of VIM driver or VNFM we gave up, searching for
-another solution.
+Openbaton default components. After we identified the VIM driver and the VNFM as
+components to be rewritten and to be adapted to the Kubernetes VIM, we studied
+the original components code. We discovered that regarding these components the
+documentation was very poor, especially looking the code. On top of that, the
+program interfaces given by the developers did not match our expectations: we
+thought to find interfaces independent from the type of the virtualization
+technology, instead, we found that they were strictly coupled with frameworks
+like Openstack or cloud vendors like Google Compute engine or Amazon AWS. These
+reasons, with the fact that after two weeks we were not able to get any working
+prototype of VIM driver or VNFM we gave up, searching for another solution.
 
@@
 Openbaton lacks: smart resource utilization. Harbor, in fact, during an SFC
-deletion, it checks if any related VNF remains orphan (which means without any
+deletion, checks if any related VNF remains orphan (which means without any
 SFC using it). In that case the VNF will enter in a state of pruning: after five
@@
 \verb!Deployment! can be associated, giving the possibility for it to scale
-under heavy workloads. The introduction of scalability meant that was not
+under heavy workloads. The introduction of scalability meant that it was not
 possible to refer to the components based on their IP, but we had to refer to
@@
 connection with the sender and the receiver. On top of that, the scalability and
-the required DNS lookup meant a possible bottleneck. With a great number of
+the required DNS lookups meant a possible bottleneck. With a great number of
 components and VNFs constantly querying the centralized Kubernetes DNS lookup
@@
 distribute the DNS service too, integrating the CoreDNS Kubernetes default
-service with \emph{DNSMasq}, a DNS server able to perform query caching for
-faster name resolutions. We created a DNSMasq \verb!DaemonSets!, which in turn
-deployed one Pod with DNSMasq running inside for every minion in the cluster. At
+service with \emph{Dnsmasq}, a DNS server able to perform query caching for
+faster name resolution. We created a Dnsmasq \verb!DaemonSets!, which in turn
+deployed one Pod with Dnsmasq running inside for every minion in the cluster. At
 the time of the VNF launch, we set that the DNS resolution should pass first for
-the local \verb!DNSMasq! DNS, and then for the Kubernetes CoreDNS service. In
-this way, we were able to distribute the DNS query workload from the master,
-using a local caching system to speed up the name resolution between services.
+the local Dnsmasq DNS, and then for the Kubernetes CoreDNS service. In this way,
+we were able to distribute the DNS query workload from the master, using a local
+caching system to speed up the name resolution between services.
 
@@
 $S$ and $R$ up, for both TCP and UDP based communications. Since the connection
-need to be end-to-end, the state of the connections are saved in the
+needs to be end-to-end, the state of the connections are saved in the
 \emph{Roulette Controller} component.
@@
 The scalability challenge introduces new issues in the system. The most
-important is regarding keeping up stateful connections, since the ingress and
-egress points could be more than one container. With this considerations in
-mind, we started thinking at how and where keeping information about the ingress
+important is keeping up stateful connections, since the ingress and egress
+points could be more than one container. With this consideration in mind, we
+started thinking at how and where keeping information about the ingress
 container handling the connection with the sender $S$ and about the egress
@@
 The result of this elaboration gets returned to Astaire, that performs the
-beforehand cited operations of VNFs communications and forwarding. In the SFC
+beforehand cited operations of VNFs communication and forwarding. In the SFC
 specification, Astaire is defined as a SFF.
@@
 At the end, we were able to create a scalable architecture that followed the
-ETSI MANO standards, with Kubernetes as VIM and Harbor has MANO. We faced
+ETSI MANO standard, with Kubernetes as VIM and Harbor has MANO. We faced
 additional challenges regarding the system scalability, that we solved using the

### 2018-11-30 res/sections/20-VNFImplementaton.tex
Fix English typos

@@
 a NS definition is created. The checks are carried out based on the VNF ID.
-While the VNF definitions corresponds to a YAML Kubernetes configuration, the NS
-one is a JSON array, for example:
+While the VNF definition corresponds to a YAML Kubernetes configuration, the NS
+one is a JSON array, as illustrated in this example:
 
@@
 components have the possibility to run inside Kubernetes, granting them the
-scalability and persistence benefits. On top of that, the use of
-\verb!Deployment! granted the possibility to perform easy an update of already
-existing components.
+scalability and persistence benefits. On top of that, the use of the Kubernetes
+\verb!Deployment! feature granted the possibility to easily perform updates of
+already existing components.
 
@@
 preserved, so that VNFs can perform packet elaboration. In order to obtain this,
-packet encapsulation has been used. IP in IP encapsulation allow to preserve the
-original packet, making the incoming packet the payload of a new one.
+packet encapsulation has been used. IP in IP encapsulation allows to preserve
+the original packet, making the incoming packet the payload of a new one.
 
 Ironhide, which was introduced in the previous chapter, plays a crucial role in
-the establishment of a end-to-end connection between the client and the
+the establishment of an end-to-end connection between the client and the
 receiver. Ironhide is able to fulfill the ingress and the egress role, and it
-provides the necessary tools to add and remove the SFC headers that gets appended
-and removed.
+provides the necessary tools to add and remove the SFC headers.
 
 Finally, we were able to create bidirectional chains, resistant to faults and
-degradation that satisfy the defined requirements.
+degradations that satisfy the defined requirements.

### 2018-11-30 res/sections/25-TestsResponsiveness.tex
Fix English typos

@@
 gathered in order to understand the real project effectiveness. We performed
-tests relatively two main sections: one testing the architecture overall
-responsiveness and one checking the SFC implementation and efficiency.
+tests relatively the architecture overall responsiveness.
 
@@
 through a common virtual machine system (in this case Virtual Box was chosen).
-To make this simulation as fair as possible, we used the same node (an 
-Openstack VM with 32GB of RAM, 8 vCPU and solid state storage) and used the 
-same ``boot sequence'' for both the VNFs: first we launched the Astaire 
-framework, ensuring its availability to process data and then we notified the 
-test machine of the successful boot. The start-up time was measured from the 
-beginning of the start-up command (\verb!docker run! for Docker and 
-\verb!VBoxManage -s! for Virtual Box) till the first TCP hit received by the 
-test back-end (to make things as smooth as possible, \verb!netcat! was used 
-to listen to incoming TCP data).
+To make this simulation as fair as possible, we used the same node (an Openstack
+VM with 32GB of RAM, 8 vCPU and solid state storage) and used the same ``boot
+sequence'' for both the VNFs: first we launched the Astaire framework, ensuring
+its availability to process data and then we notified the test machine of the
+successful boot. The start-up time was measured from the beginning of the
+start-up command (\verb!docker run! for Docker and \verb!VBoxManage -s! for
+Virtual Box) till the first packet was received by the test back-end (to make
+things as smooth as possible, \verb!netcat! was used to listen to incoming
+data).
 
@@
 \noindent As is possible to see from the YAML definition, the test involved, for
-every launch (that included and $N = 100$, for a total for 1000 tests), the 
-download of the container image from the Docker hub repositories. This, while 
-it forced every node to delete and download the image for each test, it ensured 
-consistent measurements in cases where the selected node for the deployment did 
+every launch (that included an $N = 100$, for a total for 1000 tests), the
+download of the container image from the Docker hub repositories. This, while it
+forced every node to delete and download the image for each test, it ensured
+consistent measurements in cases where the selected nodes for the deployment did
 not have the image already pulled. In addition to that, another consideration
-needs to be taken into account: when the chain consists of two or more 
-elements, the container image which is used for the container deployment is the 
-same for every VNF involved, thus reducing the start up time for the subsequent 
-pods that reside in nodes that have already deployed one.
+needs to be taken into account: when the chain consists of two or more elements,
+the container image which is used for the container deployment is the same for
+every VNF involved, thus reducing the start up time for the subsequent pods that
+reside in nodes that have already deployed one.
 
@@
 
-These additional operations implicate some time overhead, that cause the
+These additional operations implicate some time overhead, that causes the
 deployment of a single Pod with a single container to be way longer that the

### 2018-11-30 res/sections/79-Conclusions.tex
Fix English typos

@@
 that a containerized approach is faster and more reliable and faster than a 
-virtualized one, and illustrating that latency in communications through SFCs, 
-while there is room for improvement, are decent considered that this is just 
-a test-bed implementation.
+virtualized one.
 
@@
 
-Generally speaking, I have never designed and though a software architecture of
+Generally speaking, I have never designed and thought a software architecture of
 these dimensions. The multiple attempts and issues I faced during the thesis
@@
 it, and it gave me the confidence that every limit you face in life can be
-broken down.
+broken.

### 2018-12-01 res/sections/07-Introduction.tex
Add prof corrections

@@
 enables the deployment of Virtualized Network Functions (VNFs) within seconds
-running on generic purpose or custom hardware. While there may be some
-performances degradation, it has been demonstrated that their time-to-deploy and
-the possibility of automation they offer are unprecedented~\cite{nguyen2017sdn}.
-On top of that, a specialized software entity acts as a MANager and Orchestrator
-(MANO) for these VNFs, managing their lifecycle. While most of the current
-proposals are based on a hypervisor solution, in this thesis we set the
-objective of providing a proof-of-concept implementation of an Infrastructure
-Management and Orchestration plane based on the new container-based technology,
-called Docker, on a recent orchestrator framework: Kubernetes. Moreover,
-building on recent published results of the 3GPP Release 15, we outline a
-software architecture capable of end-to-end service orchestration exploiting
-container-technology.
+running on generic purpose or custom hardware. Even though the use
+virtualization is sometimes paid in terms of computational overhead, the
+technology has demostrated its maturity and the possibilities of automation it
+offers are unprecedented~\cite{nguyen2017sdn}. On top of that, a specialized
+software entity acts as a MANager and Orchestrator (MANO) for these VNFs,
+managing their lifecycle. While most of the current proposals are based on a
+hypervisor solution, in this thesis we set the objective of providing a
+proof-of-concept implementation of an Infrastructure Management and
+Orchestration plane based on the new container-based technology, called Docker,
+on a recent orchestrator framework: Kubernetes. Moreover, building on recent
+published results of the 3GPP Release 15, we outline a software architecture
+capable of end-to-end service orchestration exploiting container-technology.
 
@@
 \begin{itemize}
- \item Chapter~\ref{chap:background} describes background information about 5G
-   and the technologies used to build an ETSI compliant architecture;
- \item Chapter~\ref{chap:rel_wk} gives an insight of correlated works on this
-   field;
- \item Chapter~\ref{chap:prjan} explains the analysis performed on the
-   requirements provided and the overall project roadmap;
+ \item Chapter~\ref{chap:background} provides some background information about
+   5G and the technologies used to build an ETSI compliant architecture;
+ \item Chapter~\ref{chap:rel_wk} discusses some prior work of relevance in the
+   context;
+ \item Chapter~\ref{chap:prjan} provides a detailed analysis of the domain,
+   fixing some design requirments, stating the overall project roadmap.
  \item Chapter~\ref{chap:archimpl} provides an in-depth view of the high-level
@@
    implementation;
- \item Chapter~\ref{chap:test} describes the tests done on the platform;
+ \item Chapter~\ref{chap:test} discusses the experimental setting along with
+   some obtained preliminary results
  \item Chapter~\ref{chap:fw} illustrates possible future works;

### 2018-12-01 res/sections/08-Background.tex
Change paper images
Add prof corrections

@@
 
-In this chapter we are going to explain the most important concepts of the
-upcoming 5G standard, talking about the Virtual Network Functions (VNFs), the
-Network Slice (NS) concept and how first experiments with these new innovations
-already begun. Finally we are going to briefly introduce and explain the
-technologies involved and their role.
+This chapter provides some background information related to 5G technology
+confined to the domain under scrutiny. We start by introducing the concept the
+ETSI NFV- Management and Orchestration (MANO) proposal, outlining outlining key
+functional components and their modus operandi. Next, we provide some background
+information on the Software Defined Networking (SDN) paradigm and its
+data/control plane softwarization effort. Concluding, we introduce the Service
+Function Chaining paradigm, arguing how all the above mentioned technology could
+be employed in synergy in order to implement an end-to-end service delivery
+scheme.
 
@@
 \label{chap:background:sec:5g:sub:ns}
-The role of a \emph{network slice} in a 5G architecture is to specifically
-handle the Control-plane of a particular service (e.g. smartphones traffic,
-autonomous driving, massive IoT), deploying resources in a manner that assure
-the required latency, security and reliability. While some very peculiar and
-legacy service could require specific hardware, the common resources between
-services could be shared in a virtualized way, providing auto-scaling
-capabilities in services that are under heavy network pressure.
+
+Network Slicing is a 5G related concept referring to the possibility of
+provisioning an end-to-end optimized service within the network, that responds
+to particular Quality of Service requirements (e.g. low latency, high
+throughput). The service under consideration might be constituted by one or more
+functional, either virtualized of physical components, all controlled and
+management by a logically centralized network brain. The service provider
+exposes a set of well-defined APIs to the consumer, through which it is possible
+to define and instantiate a fully fledged service whose deployment is enacted on
+a shared infrastructure layer. It is noteworthy to point out that no exact
+definition of a network slice exists, and different approaches on how to achieve
+it have been advanced.
+
+While some very peculiar and legacy service could require specific hardware, the
+common resources between services could be shared in a virtualized way,
+providing auto-scaling capabilities in services that are under heavy network
+pressure.
 
@@
 
-This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
-requirement is to enhance the performance of end-to-end IP based that involves
-satellite connections. To reach this goal, the project specifications suggest to
-exploit the 5G incoming technology and use the NFV-MANO architecture to perform
-first packets elaboration and performance improvement and finally TCP/IP
-satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
-VIBES project proposed five technical requirements:
+VIBeS is a concrete example of a 5G project combining the use of SDN and NFV
+technology, working in concert to provide specialized network slices for several
+envisioned use-cases~\cite{vibesesa}. VIbES goals are ambitious, aiming to
+demonstrate the practical convergence of satellite and 5G in a unified frame. To
+reach this goal, the project specifications suggest to exploit the 5G incoming
+technology and use the NFV-MANO architecture to perform first packets
+elaboration and performance improvement and finally TCP/IP satellite chunk
+optimization with the Performance Enhancing Proxy (PEP). The VIBES project
+proposed five technical requirements:
 \begin{enumerate}
@@
 
-\section{Virtualization}
+\section{Virtualization Technology}
 It is remarkable how cloud computing has surged in the recent years. This is due
@@
   \centering
-  \includegraphics[scale=2]{etsi_arch}
+  \includegraphics[scale=0.5]{etsi_arch}
   \caption[Network Function Virtualization reference architecture]{Network

### 2018-12-01 res/sections/09-RelatedWork.tex
Add prof corrections

@@
 
-The possibility to solve the increasing network capacity necessities and to make
-the network infrastructure lighter and faster to deploy has rouse the interests
-of many internet service providers (ISP).
+As argued, service differentiation and network management are the swiss army
+kife of network operators. Through the rest of this chapter we discuss some
+prior work in the context of ETSI-NFV, providing end-to-end management
+capabilities into the network.
 
@@
 
-In literacy, several NFV implementations and Network Slicing solutions have been
+In literature, several NFV implementations and Network Slicing solutions have been
 proposed and some of them will be discussed in the following chapter. For

### 2018-12-01 res/sections/10-ProjectAnalysis.tex
Add prof corrections

@@
 
-As described in chapter~\ref{chap:intro}, the creation of a full ETSI MANO
-compliant architecture, following the suggestions described in RFC 7665,
-requires the integration of multiple tools (some of them have been already
-briefly introduced in~\ref{chap:intro}). The tools were chosen after their
-analysis, where pro and cons have been studied and compared to the necessity of
-this thesis. Since the project has been developed with the cloud provided by the
-University of Padua, Openstack (which a product analysis can be found later
-in~\ref{chap:prjan:sec:openstack}) has been used as base for our deployments.
+As discussed through Chapter ~\ref{chap:intro}, the implementation of an
+ETSI-compliant MANO architecture, contemplating service chain proovisioning
+capabilities as mandated by RFC 7665, requires the integration of multiple
+tools. The tools were chosen after their analysis, where pro and cons have been
+studied and compared to the necessity of this thesis. Since the project has been
+developed within the cloud environment provided by the Department of
+Mathematics, University of Padua, we initially relied on Openstack (which a
+product analysis can be found later in~\ref{chap:prjan:sec:openstack}) as the
+default infrastructure manager choice.
 
@@
 less-important components.
-\subsection{VIM}
+
+\subsection{Virtualized Infrastructure Manager}
 A critical component, as the MANO one, is the VIM. Without it, network resources

### 2018-12-01 res/sections/15-ArchitecturalImplementation.tex
Add prof corrections

@@
 
-In this chapter we will dig into the actual architectural proposal, seeing all
-the problems related to it and how they were solved. We will show all the
-various attempts made, and we will discussion about the technological issues and
-changes we had to perform.
+In this chapter we dig into the actual architectural proposal, seeing all the
+problems related to it and how they were solved. We will show all the various
+attempts made, and we will discussion about the technological issues and changes
+we had to perform.
 
 \section{Architecture introduction}
-Before describing the real implementation, we are going to briefly introduce
-some of its functional components and the way they work.
+
+Before devling into the details of the actual implementation, we briefly
+introduce some of its functional components and the way they work.
 
@@
   \centering
-  \includegraphics[scale=0.6]{Architecture}
+  \includegraphics[width=\textwidth]{Architecture}
   \caption[Architecture schema depicting the expected
@@
 \end{itemize}
-Both the layers would use Openbaton. The messages would be exchanged through
+Both layers rely on the Openbaton NFVO. The messages would be exchanged through
 RabbitMQ, using the Advance Message Queuing Protocol (AMQP).
@@
 
-\section{Second attempt - revision one}
+\section{Final Design: End-to-end MANO}
 
@@
 
-\subsection{Harbor}
+\subsection{Harbor: extending the capabilities}
 \label{chap:archimpl:sec:secondattempt:sub:harbor}
@@
 Openbaton lacks: smart resource utilization. Harbor, in fact, during an SFC
-deletion, checks if any related VNF remains orphan (which means without any
-SFC using it). In that case the VNF will enter in a state of pruning: after five
+deletion, checks if any related VNF remains orphan (which means without any SFC
+using it). In that case the VNF will enter in a state of pruning: after five
 minutes, if the VNF has not be re-utilized by any other SFC it will be deleted.
@@
 
-\subsection{Final architecture structure}
+\subsection{Discussion}
 At the end, we were able to create a scalable architecture that followed the

### 2018-12-02 res/sections/25-TestsResponsiveness.tex
Swap table columns, add saved configurations

@@
 \begin{tabular}{l|c|c}
-                     & \textbf{Standard deviation (seconds)} & \textbf{Average 
-start up time (seconds)} \\ \hline
-\textbf{Docker}      & 0.009620874                           & 0.016        \\
-\textbf{Virtual Box} & 0.755                                 & 20.211
+                     & \textbf{Average start up time (seconds)} & \textbf{Standard deviation (seconds)} \\ \hline
+\textbf{Docker}      & 0.016                                    & 0.009620874                           \\
+\textbf{Virtual Box} & 20.211                                   & 0.755                                
 \end{tabular}%
@@
   time is required to create and boot a virtualized environment.}
-\label{my-label}
+\label{chap:tests:sec:dockervsvb:tab:startup}
 \end{table}
@@
 
-% TODO is it worth to keep it as a long table?
 \begin{longtable}[c]{c|c|c}
-\textbf{SFC length} & \textbf{Standard deviation (seconds)} & \textbf{Average 
-deploy time (seconds)} \\ \hline
+\textbf{SFC Length} & \textbf{Average deploy time (seconds)} & \textbf{Standard deviation (seconds)} \\ \hline
+\endfirsthead
+%
 \endhead
 %
-\textbf{1}          & 2.560102                   & 13.291                     \\
-\textbf{2}          & 0.6301444                  & 14.207                     \\
-\textbf{3}          & 1.14903                    & 15.345                     \\
-\textbf{4}          & 3.592179                   & 16.453                     \\
-\textbf{5}          & 0.9728132                  & 16.634                     \\
-\textbf{6}          & 1.103802                   & 17.418                     \\
-\textbf{7}          & 2.944699                   & 17.984                     \\
-\textbf{8}          & 1.733722                   & 18.480                     \\
-\textbf{9}          & 1.746296                   & 18.607                     \\
-\textbf{10}         & 1.291508                   & 20.511                     \\
+\textbf{1} & 13.291 & 2.560102 \\
+\textbf{2} & 14.207 & 0.6301444 \\
+\textbf{3} & 15.345 & 1.14903 \\
+\textbf{4} & 16.453 & 3.592179 \\
+\textbf{5} & 16.634 & 0.9728132 \\
+\textbf{6} & 17.418 & 1.103802 \\
+\textbf{7} & 17.984 & 2.944699 \\
+\textbf{8} & 18.480 & 1.733722 \\
+\textbf{9} & 18.607 & 1.746296 \\
+\textbf{10} & 20.511 & 1.291508 \\
 \caption[SFC start up time]{Table of the results. It is possible to see how the
@@
 \label{chap:tests:sec:sfclength:tab:sfcdata}\\
-
 \end{longtable}

### 2018-12-02 res/sections/07-Introduction.tex
Add prof corrections (#15)
- Fix #14 - Add prof corrections
- Fix #13 - Swap table columns, add saved configurations

@@
 virtualization is sometimes paid in terms of computational overhead, the
-technology has demostrated its maturity and the possibilities of automation it
+technology has demonstrated its maturity and the possibilities of automation it
 offers are unprecedented~\cite{nguyen2017sdn}. On top of that, a specialized
@@
  \item Chapter~\ref{chap:prjan} provides a detailed analysis of the domain,
-   fixing some design requirments, stating the overall project roadmap.
+   fixing some design requirements, stating the overall project roadmap.
  \item Chapter~\ref{chap:archimpl} provides an in-depth view of the high-level

### 2018-12-02 res/sections/10-ProjectAnalysis.tex
Add prof corrections (#15)
- Fix #14 - Add prof corrections
- Fix #13 - Swap table columns, add saved configurations

@@
 As discussed through Chapter ~\ref{chap:intro}, the implementation of an
-ETSI-compliant MANO architecture, contemplating service chain proovisioning
+ETSI-compliant MANO architecture, contemplating service chain provisioning
 capabilities as mandated by RFC 7665, requires the integration of multiple

### 2018-12-02 res/sections/08-Background.tex
Fix typos
Fix spacing

@@
 technology, working in concert to provide specialized network slices for several
-envisioned use-cases~\cite{vibesesa}. VIbES goals are ambitious, aiming to
+envisioned use-cases~\cite{vibesesa}. VIBeS goals are ambitious, aiming to
 demonstrate the practical convergence of satellite and 5G in a unified frame. To
@@
 \subparagraph*{Network Service Header (NSH)}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.6]{nshtype1}
+  \caption[NSH MD Type \texttt{0x1}]{NSH MD Type \texttt{0x1}~\cite{rfc8300}.
+    The context header in this NSH has a fixed length. There is another possible
+    context header that has a variable length.}
+  \label{chap:background:img:nshtype1}
+\end{figure}
+
 The encapsulation has an header that is well defined and described in detail
@@
 requirement is to set this field with zeroes when is not used.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.6]{nshtype1}
-  \caption[NSH MD Type \texttt{0x1}]{NSH MD Type \texttt{0x1}~\cite{rfc8300}.
-    The context header in this NSH has a fixed length. There is another possible
-    context header that has a variable length.}
-  \label{chap:background:img:nshtype1}
-\end{figure}

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Add prof corrections (#15)
- Fix #14 - Add prof corrections
- Fix #13 - Swap table columns, add saved configurations
Add two missing corrections

@@
 
-\section{Architecture introduction}
+\section{A top-down overview}
 
-Before devling into the details of the actual implementation, we briefly
+Before digging into the details of the actual implementation, we briefly
 introduce some of its functional components and the way they work.
@@
 \paragraph*{Low-layer orchestrator}
-The low-layer orchestrator job corresponds to the one defined in the ETSI MANO
-definition, with the role of managing and orchestrating VNFs and NSs. The
-low-layer MANO holds different data repositories where VNF and NS definitions
-are stored, alongside with various monitoring tools, logs and network services
-repositories. The low layer MANO exposes a set of APIs to the upper layer
-(related to handling services) MANO, that is able to configure and manage the
-services. In other words, the service MANO has the possibility to configure and
-launch services to the low-layer orchestrator, providing the necessary
-information.
+The low-layer orchestrator job corresponds to the one defined in the ETSI MANO,
+with the role of managing and orchestrating VNFs and NSs. The low-layer MANO
+holds different data repositories where VNF and NS definitions are stored,
+alongside with various monitoring tools, logs and network services repositories.
+The low layer MANO exposes a set of APIs to the upper layer (related to handling
+services) MANO, that is able to configure and manage the services. In other
+words, the service MANO has the possibility to configure and launch services to
+the low-layer orchestrator, providing the necessary information.
 

### 2018-12-03 res/sections/25-TestsResponsiveness.tex

@@
 
-\begin{figure}[H]
+\begin{figure}[t]
     \centering
@@
       average 20,211 seconds to boot up, while a container requires, on average,
-      only 0.016 seconds. The main reason behind this difference, as already
-      explained in other parts of this thesis, it is caused by the guest's
-      kernel that has to initialized itself, before starting any program. On
-      Docker, instead, the kernel is the same of the host, so the only
-      operations needed are to isolate the program from the rest of the OS and
-      add it to the process list.}
+      only 0.016 seconds.}
     \label{chap:tests:sec:dockervsvb:subimg:plots}
@@
 
-\begin{table}[t]
+\begin{table}[h]
 \centering
@@
 with a related Service, to simulate a real VNF. After the Pod start-up, a
-request
-is made to a back-end designed to count the chain elements that have 
-successfully
-boot-up returning, eventually, the overall start up timing.
-
-% TODO: missing YAML definition: [language=YAML]
-%https://tex.stackexchan
-%ge.com/questions/152829/how-can-i-highlight-yaml-code-in
-%- a-pretty-way-with-listings
+request is made to a back-end designed to count the chain elements that have
+successfully boot-up returning, eventually, the overall start up timing.
 \lstinputlisting[caption={Kubernetes YAML configuration that has been used to
@@
   language=yaml]{res/code/sfc_test.yaml}
-
-\vspace{0.5cm}
-
 \noindent As is possible to see from the YAML definition, the test involved, for

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Fix Arch impl

@@
 16.04 (for Openstack) and Ubuntu 18.04 (for Openbaton). Our main goal was to set
-up Openstack as our VIM while setting Openbaton as our MANO. Then, Openbaton
-would have registered Openstack as PoP, where TOSCA definitions would have been
-launched on Openstack that, instead of deploying hypervisor based virtual
-machines, it would have deployed container instances. The VNFs lifecycles would
-have been controlled through an Element Management System (EMS), transmitting
-the various VNF states via the message broker RabbitMQ. It is possible to see the
-interactions between elements in a minimum schema available in
+up Openstack as our VIM while setting Openbaton as our MANO. Then, Openstack
+would register with Openbaton and TOSCA definitions would be deployed on
+Openstack following the Infrastructure as Code (IaC) paradigm. In this context
+we are interested in deploying container-based instances instead of hypervisor
+based virtual machines. The VNFs lifecycles would have been controlled through
+an Element Management System (EMS), transmitting the various VNF state changes
+via the message broker RabbitMQ. It is possible to see the interactions between
+the elements in a minimal schema available in
 Figure~\ref{chap:archimpl:sec:fistattempt:img:schema1}.
@@
 
-Our first step was to try and use Openstack to create a container orchestration.
-The reason behind this approach was simple: Openbaton, on the paper, comes with
-an out-of-the-box support for Openstack. In fact, it natively supports Openstack
-PoPs, giving to us a great possibility to reuse the two frameworks and to ease
-the development phase of our software.
+Our first step was to try and use Openstack to create a container orchestration
+environment. The reason behind this approach was simple: Openbaton, on the
+paper, comes with an out-of-the-box support for Openstack. In fact, it natively
+supports Openstack registered as a PoP, giving to us a great possibility to
+reuse the two frameworks and to ease the development phase of our software.
 \subsubsection{Devstack: Openstack for developers}
@@
 multiple machines (nodes), therefore creating a whole cluster of resources.
-Since we considered too time-consuming to install Openstack in a distributed
-manner, we considered to test first the developer version, called
-\emph{Devstack}. Devstack offers the possibility to install a subset of modules
-in the local machine, and to leverage a Openstack environment able to perform
-the same operations of the production-grade installation. Thus, we started
-configuring the Devstack installation, installing first of all the Nova module,
-that is required to perform computing operations such as the launch of virtual
-machines, and lately we installed Tacker too.
+
+Since a distributed Openstack installation was too time consuming, we considered
+to test first the developer version, called \emph{Devstack}. Devstack offers the
+possibility to install a subset of modules in the local machine, and to leverage
+a Openstack environment able to perform the same operations of the
+production-grade installation. Thus, we started configuring the Devstack
+installation, installing first of all the Nova module, that is required to
+perform computing operations such as the launch of virtual machines, and lately
+we installed Tacker too.
 
@@
 First of all, Devstack needs a machine with at least 16GB of ram and 50GB of HD
-space.
-We decided to install these components:
+space. We decided to install these components:
 \begin{itemize}
- \item Keystone (for identity management)
+ \item Keystone: identity management module
  \item Object storage
  \item Compute
- \item Tacker (we tested this module to see is Openstack was a suitable VIM too)
+ \item Tacker: a VNF and NFVO (as explained), tested to see if Openstack was a
+   suitable VIM too
 \end{itemize}
@@
 
-A short description about every installation/deploy method will be briefly
+A short description about every installation/deploy method is briefly
 illustrated.
@@
 
+\vfill
+
 \subsection{Docker integration in Openstack}
@@
 
-After the initial Openstack and Openbaton configuration we started making the
-components talking to each other. First, we tried a simple \verb!iperf! data
-transmission with two VNFs deployed by Openbaton as a common hypervisor-based
-Openstack instances, as explained in~\cite{openbatonIperf}. With this test
-completed, we started tinkering Openstack for using Docker containers instead of
-traditional hypervisor instances. When we started developing this solution, we
-adopted the Zun module, that aims to smoothly integrate Docker in Openstack.
-Unfortunately, we were not able to make it properly work, in particular, it was
-not possible to open desired ports on the Docker containers. This lead to the
-impossibility to perform packet forwarding between VNFs. On top of that, we
-found difficulties regarding the integration of Docker images in TOSCA
-definitions and Openstack. These two issues are related: the TOSCA definition
-does not have any keyword to specificate in any way that the image used for the
-VNF definition is a Docker image, and Openstack, with the Zun plugin, at the
-time of the implementation did not performed the image pull from any Docker
-registry. The impossibility to easily set the containers ports and the lack of
-some support between the TOSCA definition and the Openstack Zun plugin pushed us
-to drop Openstack as a possible solution and to start looking for another valid
-VIM component.
+After the deployment of Openstack and Openbaton, the next logical step is to try
+and make the component talk to each other. First, we tried a simple \verb!iperf!
+data transmission with two VNFs deployed by Openbaton as a common
+hypervisor-based Openstack instances, as explained in~\cite{openbatonIperf}.
+With this test completed, we started experimenting Openstack so as to use Docker
+containers instead of traditional hypervisor instances. When we started
+developing this solution, we adopted the Zun module~\cite{zunmoduleos}, that
+aims to smoothly integrate Docker in Openstack. Unfortunately, we were not able
+to make it work properly. In particular, it was not possible to open
+communications ports on the Docker containers, hence making it impossible for
+the VNFs to communicate. On top of that, we found difficulties regarding the
+integration of Docker images in TOSCA definitions and Openstack. These two
+issues are related: the TOSCA definition does not have any keyword to
+specificate in any way that the image used for the VNF definition is a Docker
+image, and Openstack, with the Zun plugin, at the time of the implementation did
+not performed the image pull from any Docker registry. The impossibility to
+easily set the containers ports and the lack of some support between the TOSCA
+definition and the Openstack Zun plugin pushed us to drop Openstack as a
+possible solution and to start looking for another valid VIM component.
 
@@
 mature and famous ones: Docker Swarm and Kubernetes. The decision to study only
-these two relies on their big community supporting them and the possibility to
-easily found projects that used the aforementioned frameworks.
+these two, relies on the big communities supporting them and the possibility to
+easily find Open Source projects built on them that used the aforementioned
+frameworks.
 
@@
 
-First we started playing with Docker Swarm. Docker Swarm is a product directly
-made by the Docker Inc., so it seamlessly integrates with other Docker tools
-such as Docker Compose. The installation and set-up is pretty straightforward,
-and the configuration is almost automatically handled by Docker itself. The
-minimum number of nodes in order to create a proper Docker Swarm cluster is of
-three, where one must be a manager node (designed to orchestrate the other
-nodes in the cluster), while the others can be simple workers. Multiple managers
-can coexist in the same cluster, as long as only one effectively operates. The
-others act like backup manager in case the one orchestrating goes down for some
-reason. The networking is given out-of-the box, and the same holds for
-distributed persistence. While this for most of the uses cases can be considered
-a pro, in our specific situation the impossibility to tweak the network
-configuration was a cons, since it was important for us to have the full control
-over the network, especially regarding the SFC implementation, explained better
-in Chapter~\ref{chap:vnf_ns_impl}.
+First we started assesing the Docker Swarm orchestrator. Docker Swarm is a
+product directly made by the Docker Inc., so it seamlessly integrates with other
+Docker tools such as Docker Compose. The installation and set-up is pretty
+straightforward, and the configuration is almost automatically handled by Docker
+itself. The minimum number of nodes needed to create a proper Docker Swarm
+cluster is three, where one must be a manager node (designed to orchestrate the
+other nodes in the cluster), while the others can be simple workers. Multiple
+managers can coexist in the same cluster, as long as only one effectively
+operates. The others act like backup managers in case the one orchestrating goes
+down for some reason. The networking is given out-of-the box, and the same holds
+for distributed persistence. While this for most of the uses cases can be
+considered a pro, in our specific situation the impossibility to tweak the
+network configuration was a cons, since it was important for us to have the full
+control over the network, especially regarding the SFC implementation, explained
+better in Chapter~\ref{chap:vnf_ns_impl}.
 
@@
 strategy or the persistence layer. With this flexibility Kubernetes offered to
-us what Docker Swarm was not able to give. Additionally, we had the possibility
-to set custom Domain Name System (DNS) lookups and distribute other internal
-components in order to avoid requests bottleneck.
+us what Docker Swarm was not able to give. Additionally, we developed a custom
+Domain Name System (DNS) lookup process and distribute other internal components
+in order to avoid requests bottleneck (as explained better in
+Section~\ref{chap:archimple:sec:secondattempt:sub:scalability}).
 
@@
 
-Kubernetes does not have a minimum number of nodes in order to properly work.
+Kubernetes does not require a minimum number of nodes in order to properly work.
 Nonetheless, we decided to use a four node cluster configured in the following
@@
   \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
-We initially deployed from the internal Openstack of the University four nodes
-with 4GB of RAM, 10 GB of SSD storage and 2 vCPUs on Ubuntu 16.04 machines (that
-we later swapped with CentOS 7). We performed the installation using the
-official Kubernetes repositories. After the initial installation, we obtained a
+For the initial deployment we relied on an internal Openstack instance installed
+at the University of Padova made of four nodes with 4GB of RAM, 10GB of SSD
+storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
+7). We performed the installation using the official Kubernetes repositories,
+taking care of performing a accurated cluster configuration, as better explained
+in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
 configuration similar to the one illustrated in
@@
 create an image with the integrated EMS tools that provided a seamless
-integration with Docker container. With this setting, the only thing to solve
-was how to convert TOSCA definitions in YAML one. The first thing we started
-thinking about was how to perform the conversion. This topic was quite important
+integration with Docker container. Under these settings, a key issue is the
+conversion of TOSCA definitions in YAML ones. This topic was quite important
 because the TOSCA definition does not really consider the possibility to have an
-alternative virtualization system apart from the hypervisor one. This causes the
-TOSCA definition to be, in our opinion, too much focused on the resources usage
-and granularity aspects such the deployment flavour and the kind of virtual link
-(concept that does not really exists in container based systems) to instantiate.
-Nonetheless, we performed a mapping between most of the TOSCA keywords with some
-of the YAML definitions, in a way that a TOSCA configuration file was able to
-generate a valid YAML deployment. After that, we started thinking instead who,
-in the current architecture, had the task to perform this operation. Initially
-we thought that Kubernetes itself could have been a perfect candidate. From a
-received TOSCA, a Kubernetes component should have validated and translated the
-TOSCA definition to generate a YAML suitable to be finally submitted to the
-Kubernetes API Server. Another feasible approach could have been to create a
-Openbaton plugin in order to convert the TOSCA definition before submitting it
-to Kubernetes for the deployment. This solution had great advantages in terms of
-responsibility separation: in this way only the MANO had to know, for a VIM
-based on Kubernetes, that a conversion was needed. From a Kubernetes
-point-of-view, nothing would have changed, since all the configuration
-conversion would have been performed by Openbaton.
+alternative virtualization system apart from the hypervisor one. In particular,
+we found the Openbaton TOSCA definition to be virtual machine centric, without
+providing the possibility of formulating a agnostic approach to virtualization
+enabling container-based deployments. Nonetheless, we performed a mapping
+between most of the TOSCA keywords with some of the YAML definitions, in a way
+that a TOSCA configuration file was able to generate a valid YAML deployment.
+After that, we started thinking who, in the current architecture, had the task
+to perform this operation. Initially we thought that Kubernetes itself could
+have been a perfect candidate. From a received TOSCA, a Kubernetes component
+should have validated and translated the TOSCA definition to generate a YAML
+suitable to be finally submitted to the Kubernetes API Server. Another feasible
+approach could have been to create a Openbaton plugin in order to convert the
+TOSCA definition before submitting it to Kubernetes for the deployment. This
+solution had great advantages in terms of responsibility separation: in this way
+only the MANO had to know, for a VIM based on Kubernetes, that a conversion was
+needed. From a Kubernetes point-of-view, nothing would have changed, since all
+the configuration conversion would have been performed by Openbaton.
 
@@
 
-The impossibility to develop a custom VIM driver and VNFM manager led us to
-search for an alternative solution. Then, we started to work directly on the
-VIM. The base approach was to create an API Server which in future an Openbaton
-plugin could contact and simply deploy TOSCA definitions, interacting with a set
-of RESTful APIs. We called this simple backend \emph{Harbor}.
+The impossibility to develop a custom VIM driver and VNFM manager, given the
+resources of hand, led us to search for an alternative solution. Then, we
+started to work directly on the VIM. The base approach was to create an API
+Server which in future an Openbaton plugin could contact and simply deploy TOSCA
+definitions, interacting with a set of RESTful APIs. We called this simple
+backend \emph{Harbor}.
 
@@
 \begin{itemize}
-\item whole system scalability: we wanted that every component had the
-  possibility to scale during intense workloads, without the necessity of human
-  intervention.
-\item ingress and egress points: there should be at least two points in the
-  cluster were the packet enters and exits;
-\item dynamic package routing: as formulated by one of our requirements, packet
+\item whole system scalability: we want for every component to have the
+  possibility of scaling during intense workloads, without the necessity of
+  human intervention;
+\item ingress and egress points: specularized entities of our framework handling
+  SFC encapsulation/decapsulation session data;
+\item dynamic packet routing: as formulated by one of our requirements, packet
   routing had to be dynamic. We immediately identified a possibility to create
@@
 \end{itemize}
+In the following, we provide some details on how we achieved the aforementioned
+properties.
 
 \subsubsection{Scalability considerations}
+\label{chap:archimple:sec:secondattempt:sub:scalability}
 When adopting Kubernetes, we realized the possibility to make every component
 easily scalable. Indeed, all the following components can be described in
-Kubernetes as a \verb!Service!. For a \verb!Service! a consequent
-\verb!Deployment! can be associated, giving the possibility for it to scale
-under heavy workloads. The introduction of scalability meant that it was not
-possible to refer to the components based on their IP, but we had to refer to
-their service name. As it will be explained later, this introduced challenges
-especially for the ingress and egress points of our network, that have to keep a
-connection with the sender and the receiver. On top of that, the scalability and
-the required DNS lookups meant a possible bottleneck. With a great number of
-components and VNFs constantly querying the centralized Kubernetes DNS lookup
-system a performance degradation could have easily happened. We decided to
-distribute the DNS service too, integrating the CoreDNS Kubernetes default
-service with \emph{Dnsmasq}, a DNS server able to perform query caching for
-faster name resolution. We created a Dnsmasq \verb!DaemonSets!, which in turn
-deployed one Pod with Dnsmasq running inside for every minion in the cluster. At
-the time of the VNF launch, we set that the DNS resolution should pass first for
-the local Dnsmasq DNS, and then for the Kubernetes CoreDNS service. In this way,
-we were able to distribute the DNS query workload from the master, using a local
-caching system to speed up the name resolution between services.
+Kubernetes as a \verb!Service!. A \verb!Service! can have zero or many
+\verb!Deployment!s, giving the possibility for it to scale under heavy
+workloads. Contemplating for scalability meant that it was not possible to refer
+to the components based on their IP, but we had to refer to their service name.
+As it will be explained later, this introduced challenges especially for the
+ingress and egress points of our network, that have to keep a connection with
+the sender and the receiver. On top of that, the scalability and the required
+DNS lookups meant a possible bottleneck. With a great number of components and
+VNFs constantly querying the centralized Kubernetes DNS lookup system a
+performance degradation could have easily happened. We decided to distribute the
+DNS service too, integrating the CoreDNS Kubernetes default service with
+\emph{Dnsmasq}, a DNS server able to perform query caching for faster name
+resolution. We created a Dnsmasq \verb!DaemonSets!, which in turn deployed one
+Pod with Dnsmasq running inside for every minion in the cluster. At the time of
+the VNF launch, we set that the DNS resolution should pass first for the local
+Dnsmasq DNS, and then for the Kubernetes CoreDNS service. In this way, we were
+able to distribute the DNS query workload from the master, using a local caching
+system to speed up the name resolution between services.
 
@@
 As just said, the whole framework needs to have and entry and an exit point.
-These, defined as \emph{Ingress} and \emph{Egress} points, are performed by
-\emph{Ironhide}. Ironhide hides the whole packet processing to the sender $S$
-and the receiver $R$. When a client $S$ performs a request to $R$, the packet
-actually goes through Ironhide, who has an internal classifier that it is able
-to recognize the type of connection and to choose the right NS the packet has to
-go through. This process is fundamental since it keeps the connection between
-$S$ and $R$ up, for both TCP and UDP based communications. Since the connection
-needs to be end-to-end, the state of the connections are saved in the
-\emph{Roulette Controller} component.
+These entities, defined as \emph{Ingress} and \emph{Egress} points, are
+provisioned by \emph{Ironhide}. Ironhide hides the whole packet processing to
+the sender $S$ and the receiver $R$. When a client $S$ performs a request to
+$R$, the packet actually goes through Ironhide, who has an internal classifier
+that it is able to recognize the type of connection and to choose the right NS
+the packet has to go through. This process is fundamental since it keeps the
+connection between $S$ and $R$ up, for both TCP and UDP based communications.
+Since the connection needs to be end-to-end, the state of the connections are
+saved in the \emph{Roulette Controller} component.
 
@@
 \subsubsection{Keeping the connection status}
-The scalability challenge introduces new issues in the system. The most
-important is keeping up stateful connections, since the ingress and egress
-points could be more than one container. With this consideration in mind, we
-started thinking at how and where keeping information about the ingress
-container handling the connection with the sender $S$ and about the egress
-container talking with the receiver $R$. Since the Roulette component has a
-complete end-to-end topology view of the system, we decided to store these
-additional information (we called it \emph{endpoints}) in it.
+Desining for scalability gave rise to new issues. The most important is keeping
+up stateful connections, since the ingress and egress points could be more than
+one container. With this consideration in mind, we started thinking at how and
+where keeping information about the ingress container handling the connection
+with the sender $S$ and about the egress container talking with the receiver
+$R$. Since the Roulette component has a complete end-to-end topology view of the
+system, we decided to store these additional information (we called it
+\emph{endpoints}) in it.
 
@@
 At the end, we were able to create a scalable architecture that followed the
-ETSI MANO standard, with Kubernetes as VIM and Harbor has MANO. We faced
+ETSI MANO standard, with Kubernetes as VIM and Harbor as MANO. We faced
 additional challenges regarding the system scalability, that we solved using the

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Fix typos

@@
 
-First we started assesing the Docker Swarm orchestrator. Docker Swarm is a
+First we started assessing the Docker Swarm orchestrator. Docker Swarm is a
 product directly made by the Docker Inc., so it seamlessly integrates with other
@@
 For the initial deployment we relied on an internal Openstack instance installed
-at the University of Padova made of four nodes with 4GB of RAM, 10GB of SSD
+at the University of Padua made of four nodes with 4GB of RAM, 10GB of SSD
 storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
 7). We performed the installation using the official Kubernetes repositories,
-taking care of performing a accurated cluster configuration, as better explained
+taking care of performing an accurate cluster configuration, as better explained
 in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
@@
   human intervention;
-\item ingress and egress points: specularized entities of our framework handling
+\item ingress and egress points: specialized entities of our framework handling
   SFC encapsulation/decapsulation session data;
@@
 \subsubsection{Keeping the connection status}
-Desining for scalability gave rise to new issues. The most important is keeping
+Designing for scalability gave rise to new issues. The most important is keeping
 up stateful connections, since the ingress and egress points could be more than

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Add missing
Fix minor typos

@@
 After the deployment of Openstack and Openbaton, the next logical step is to try
-and make the component talk to each other. First, we tried a simple \verb!iperf!
-data transmission with two VNFs deployed by Openbaton as a common
+and make the components talk to each other. First, we tried a simple
+\verb!iperf! data transmission with two VNFs deployed by Openbaton as a common
 hypervisor-based Openstack instances, as explained in~\cite{openbatonIperf}.
@@
 us what Docker Swarm was not able to give. Additionally, we developed a custom
-Domain Name System (DNS) lookup process and distribute other internal components
-in order to avoid requests bottleneck (as explained better in
+Domain Name System (DNS) lookup process and distributed other internal
+components in order to avoid requests bottleneck (as explained better in
 Section~\ref{chap:archimple:sec:secondattempt:sub:scalability}).
@@
 For the initial deployment we relied on an internal Openstack instance installed
-at the University of Padua made of four nodes with 4GB of RAM, 10GB of SSD
-storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
+at the University of Padua, creating four nodes with 4GB of RAM, 10GB of SSD
+storage and 2 vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
 7). We performed the installation using the official Kubernetes repositories,
-taking care of performing an accurate cluster configuration, as better explained
-in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
-configuration similar to the one illustrated in
+taking care of performing an accurate cluster configuration complete with
+additional modules (e.g. networking and distributed persistent storage), as
+better explained in Appendix~\ref{kubernetes}. After the initial installation,
+we obtained a configuration similar to the one illustrated in
 Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
@@
 The impossibility to develop a custom VIM driver and VNFM manager, given the
-resources of hand, led us to search for an alternative solution. Then, we
+resources at hand, led us to search for an alternative solution. Then, we
 started to work directly on the VIM. The base approach was to create an API

### 2018-12-03 res/sections/05-Abstract.tex
Little fix

@@
 specifics, we provide a proof of concept implementation of an infrastructure
-Management and Orchestration (MANO) plane, replying on the Kubernetes container
+Management and Orchestration (MANO) plane, relying on the Kubernetes container
 orchestration framework.

### 2018-12-04 thesis/res/sections/10-ProjectAnalysis.tex
Add presentation and fix travis

--- 2018-12-02 res/sections/10-ProjectAnalysis.tex
@@
 \label{chap:prjan:sec:tech:sub:other:sub:swarm}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=1]{dockerswarm_logo}
+  \caption[Docker Swarm logo]{Docker Swarm logo}
+\end{figure}
+
 Introduced in 2014~\cite{swarmGit}, Docker Swarm allows multiple Docker nodes to
@@
 market needs.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=1]{dockerswarm_logo}
-  \caption[Docker Swarm logo]{Docker Swarm logo}
-\end{figure}
 

### 2018-12-04 thesis/res/sections/81-k8s.tex
Add presentation and fix travis

\chapter{Kubernetes}
\label{kubernetes}

\newcommand{\gluster}{GlusterFS}

In this appendix we provide a basic summary to the Kubernetes concepts and
we will show a simple way to perform a basic installation. This is not thought 
as an complete overview of functionalities but there are only reported a 
description of concepts an features used in the implementation of our solution.

\section{Kubernetes basic concepts}
\label{kubernetes-basic-concepts}

\subsection{Pod}
\label{pod}

A Pod is the most basic building block for Kubernetes. Pods wrap up the
underlying virtualization technology used: in case its a container based
solution, it can contain one or more container (although this configuration is
worth only if the containers are highly coupled). Pods have a lifecycle, and
they can assume these states\footnote{
  \url{https://kubernetes.io/docs/concepts/workloads/pods/pod-lifecycle/}}:

\begin{description}
\item[Pending] The Pod has been accepted by the Kubernetes system, but
  one or more container images have not been created yet. This includes the Pod
  scheduling and image downloading which could take a while.
\item[Running] The Pod has been bound to a node, and all of the
  Containers have been created. At least one container is still running, or is
  in the process of starting or restarting.
\item[Succeeded] All containers in the Pod have terminated with
  success, and they will not be restarted.
\item[Failed] All containers in the Pod have terminated, and at least
  one container has terminated with a failure. That is, the container either
  exited with non-zero status or was terminated by the system.
\item[Unknown] For some reason the state of the Pod could not be
  obtained, typically due to an error communicating with the host of the Pod.
\end{description}

A Pod can have different conditions too:

\begin{description}
\item[PodScheduled] The Pod has been scheduled to a node;
\item[Ready]: the Pod is able to serve requests and should be added
  to the load balancing pools of all matching Services;
\item[Initialized] All init containers\footnote{
  \url{https://kubernetes.io/docs/concepts/workloads/pods/init-containers}}
  have started successfully;
\item[Unschedulable] The scheduler cannot schedule the Pod right
  now, for example due to the lacking of resources or other constraints;
\end{description}

Each Pod can have different QoS classes:
\begin{description}
  \item [Guaranteed] Every container in the Pod must have a memory and CPU
  limits and memory and CPU requests, and they must be the same;
  \item [Burstable] The Pod does not meet the criteria for QoS class
  Guaranteed and at least one container in the Pod has a memory or CPU request;
  \item [BestEffor] The containers in the Pod must not have any memory or CPU
  limits or requests.
\end{description}

\subsection{Service}
\label{service}

Since Pods for their nature are highly volatile, applications that use them can
not rely on their IP address. To solve this problem, the Service implementation
in Kubernetes assure a stable IP, even if different Pods in this service crash
and gets restarted by a ReplicaSet (which will not be explain in this appendix,
but it basically scale up and down Pods\footnote{
  \url{https://kubernetes.io/docs/concepts/workloads/controllers/replicaset/}}).
Citing the official documentation\footnote{
\url{https://kubernetes.io/docs/concepts/services-networking/service/}}, a
Service is:\\
\textquote{A Kubernetes Service is an abstraction which defines a logical set
of Pods and a policy by which to access them - sometimes called a
micro-service.}

\begin{figure}[htbp]
\centering
\includegraphics{k8s_services}
\caption{Example of a service}
\end{figure}

\subsection{Volume}
\label{volume}

When Pods saves data on a node, this information is designated to perish.
Kubernetes offers a way to save all the data and to possibly make it available
to all the Pods replicas, offering a transparent layer where Pods are not aware
of working with information shared other Pods, even located in other nodes. This
functionality is called Volume, and there are many different drivers, that range
from a simple local storage to \emph{\gluster{}}.

\subsection{Namespace}
\label{namespace}

Namespaces are intended to be used when there are multiple users accessing the
cluster, or to separate multiple projects.

\section{Kubernetes architectural concepts}
\label{kubernetes-architectural-concepts}

\subsection{Node}\label{node}

A node is the building block for a Kubernetes cluster. Everyone must have
at least the kubelet daemon (to manage Kubernetes functionalities) and a
container manager (e.g. Docker Engine, VirtualBox) running. A node runs Pods
and can communicate with other nodes in the same cluster.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.8]{k8s_nodes}
\caption{Schema illustrating the elements in a node}
\end{figure}

\subsection{Master}
\label{master}

The \emph{master node} is the one that starts a cluster, and that manage the
slaves. Only some Pods can run in the master node (e.g. the dashboard).

The master generates the token for the slaves to join, and eventually
coordinates the federation with other clusters. In a cluster, there is the
possibility to have multiple master to ensure high availability.

\subsection{Slave}
\label{slave}

A \emph{slave}, or \emph{minion} is a node that runs the Pods that the master
schedule. It can also be part of a \emph{PersistantVolume} storage (like
\gluster{}) and if it goes down, another slave will take care of the Pods that
where running on it. Minions can join an existing cluster via a given token.

\subsection{Networking}
\label{networking}

Kubernetes has its own networking structure, even if it uses Docker for running
containers. In particular, it delegates the network operations (like packet
forwarding) to different managers, such as Flannel or Calico.

Every node in the cluster runs a Pod called \texttt{kubeproxy}, that
communicates with the \texttt{kube-api} Pod (running on master) and
\texttt{kube-dns} (for Kubernetes \textless{}=1.9.x) or \texttt{CoreDNS} (for
Kubernetes \textgreater{}=1.10.x) services. \texttt{kubeproxy} assure
transparent communications between Pods in the same cluster, and
\texttt{kube-api} allows to call Kubernetes services. \texttt{kube-dns/CoreDNS}
manage DNS requests coming from the Pods, differentiating the one that goes
outside the cluster from the other that points internally\footnote{Additional
  information about Flannel, Kubernetes and Docker networking can at:\\\sloppy
  \url{https://www.sdxcentral.com/cloud/containers/definitions/what-is-coreos-flannel-definition/}}.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.3]{flannelredirect}
\caption{And example of how Flannel redirects packages between Pods}
\end{figure}

\section{Focusing on Storage}
\label{focusing-on-storage}

This section is based on official Kubernetes documentation\footnote{%
  \url{https://kubernetes.io/docs/concepts/storage/persistent-volumes/}} and
from the \texttt{newstack site}\footnote{\sloppy
  \url{https://thenewstack.io/strategies-running-stateful-applications-kubernetes-persistent-volumes-claims/}}.

\subsection{Volumes, Claims and Classes}
\label{volumes-claims-and-classes}

Managing compute and storage resources are totally different problems and in
order to create a proper storage abstraction Kubernetes relies on the
\textbf{PersistantVolume} and the \textbf{PersistantVolumeClaim} features.

A PersistantVolume (PV) is something similar to a node, in fact is a piece of
storage provided by the administrator: it captures the details of the
implementation of the storage and has a lifecycle that is independent of any
individual Pod. To not expose developers to storage boring details admins can
expose different PV types through \textbf{StorageClasses}.

A PersistantVolumeClaim (PVC) instead is a request for storage and it is similar
to a Pod: instead of requesting memory and CPU, a PVC requests certain size and
access mode to storage.

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.18]{kubernetes_pvc}
\caption{PersistantVolumeClaims management}
\end{figure}

It is important to understand the difference between \textbf{Volumes} and
\textbf{PersistantVolumes}/\textbf{PersistantVolumeClaims}. The former is
something like Docker volumes, used for persistence, host-based or on external
block storage devices but a user does not expect the resources to be reserved
before using it. In persistent volumes and claims, there is a strict enforcement
of resource utilization dictated by the policy defined during the creation of
resources. Kubernetes Pods can use claims as volumes.

\textbf{StorageClasses}\footnote{more information on different provisioners,
  policies and parameters can be found here: \\\url{
 https://kubernetes.io/docs/concepts/storage/storage-classes/ }} are a way to
describe different type of storage that can be used. Different classes can be
used to describe different levels of service or policies. PVCs that do not
specify a particular class can use the default one: to define it is possible to
use the annotation \verb!"storageclass.kubernetes.io/is-default-class":"true"!,
as for instance:

\begin{lstlisting}[language=bash]
kubectl patch storageclass glusterfs-storage -p '{"metadata": {"annotations":{"storageclass.kubernetes.io/is-default-class":"true"}}}'
\end{lstlisting}

\subsection{Volume and claim lifecycles}
\label{lifecycle}

We are going to briefly explain all the possible interactions between PVs and
PVCs.

\begin{enumerate}
\item \textbf{Provisioning}: there are 2 ways: statically or dynamically.
  \emph{Static} means that an administrator creates some PVs that are available
  for consumption. \emph{Dynamic} means that the cluster dynamically tries to
  provide a volume that matches a PVC. This is done with \emph{StorageClasses}:
  a claim requests storage for a certain storage class that can be provisioned.
  The class \texttt{""} disable dynamic provisioning;
\item \textbf{Binding}: a user requests a certain amount of storage with a claim
  and a control loop in the master looks for a PV that can bind the claim. In
  case of dynamic provisioning the loop always binds a PVC to a PV, otherwise
  the user will always get at least what they asked for, but the volume may be
  larger. Claims can remain unbound if a matching volume does not exists.
\item \textbf{Using}: after the binding phase the Pod starts using the volume
  (PVC is described in the Pod yaml
  \footnote{\url{https://kubernetes.io/docs/concepts/storage/persistent-volumes/\#claims-as-volumes}})
\item \textbf{Releasing}: when an application is done using the volume,
  developers can delete the PVC objects through the API, releasing the claim.
  This step will initiate the reclamation process.
\item \textbf{Reclaiming}: this phase is defined as a policy by the
  administrators. The reclaim policy for a PersistantVolume tells the cluster
  what to do with the volume after it has been released of its claim. Persistent
  volumes can either be \emph{retained}, \emph{recycled} or \emph{deleted}.
\end{enumerate}

\section{Installing Kubernetes}
\label{installing-kubernetes}

We performed our Kubernetes installation in a Openstack environment. Since we
used the Openstack of our school department, the administrators have given us a
virtual machine, \verb!Openstack-CLI!, to perform operations. From now, our
scripts assume to be in a machine with access to Openstack, and that the file
\verb!openrc.sh! has been sourced in the command line interface before
continuing. To install Kubernetes on CentOS 7 we develop a
script\footnote{available at
  \url{https://github.com/Augugrumi/init-script/tree/centos/kubernetes}}. With
CentOS is possible to have a complete installation with a
GlusterFS-PersistentStorage correctly set. After there is an explanation for the
most important part of the scripts. The script \verb!on_openstack_cli.sh!
\footnote{\url{https://github.com/Augugrumi/init-script/raw/centos/kubernetes/on_openstack_cli.sh}}
main aim is to run other files retrieving information even using \texttt{nova}
commands. First of all the software dependencies have to be installed and for
this purpose the script \verb!install_kubectl_repo.sh!
\footnote{\url{https://raw.githubusercontent.com/Augugrumi/init-script/centos/kubernetes/install_kubectl_repo.sh}}
was created. After the software installation it calls another script and waits
until all the machines are ready to proceed to the next step. The other script
is \verb!bootstrap.sh!
\footnote{\url{https://raw.githubusercontent.com/Augugrumi/vagrantfiles/oldversion/kubernetes/centos/bootstrap.sh}}
whose aim is to configure the machine and to install all the necessary
software. The bootstrap script first \textbf{disables SELinux} on all machines 
(needed for Kubernetes networking):
\begin{lstlisting}
sudo setenforce 0
sudo sed -i --follow-symlinks 's/SELINUX=enforcing/SELINUX=disabled/g' /etc/sysconfig/selinux
\end{lstlisting}
and then it enables the \verb!br_netfilter! module (needed for Kubernetes
networking):
\begin{lstlisting}
sudo modprobe br_netfilter
sudo sh -c 'echo "1" > /proc/sys/net/bridge/bridge-nf-call-iptables'
\end{lstlisting}
The swap is then switched off with the following commands:
\begin{lstlisting}
sudo swapoff -a
sudo sed -i '/swap/d' /etc/fstab
\end{lstlisting}
and it finally adds \texttt{rbd}, \texttt{ip\_vs}, \texttt{ip\_vs\_rr},
\texttt{ip\_vs\_wrr}, \texttt{ip\_vs\_sh}, \texttt{dm\_snapshot},
\texttt{dm\_mirror} and \texttt{dm\_thin\_pool} to the modules that will be
loaded at the machine startup.
\begin{lstlisting}
sudo sh -c 'echo -e "rbd ip_vs ip_vs_rr ip_vs_wrr ip_vs_sh" >  
/etc/modules-load.d/ip_vs.conf'
sudo sh -c 'echo -e "dm_snapshot dm_mirror dm_thin_pool" >  
/etc/modules-load.d/gluster.conf'
\end{lstlisting}
It is important to note that the last 3 modules will be useful in case
\gluster{} will be installed in the system. In the end the script installs
\texttt{docker}, \texttt{device-mapper-persistent-data}, \texttt{lvm2},
\texttt{kubelet}, \texttt{kubectl}, \texttt{kubeadm} and it starts (and enables)
\texttt{docker} and \texttt{kubelet} services. After this the machines are
rebooted. Considering $M$ to be the set of machines that are going to be added
to the Kubernetes cluster, for each host that is part of $M$ a set of entries
corresponding of the hostname and the IP address of every other node in $M$ will
be added in the \texttt{/etc/hosts} file. Then the installation continues in the
\textbf{master} node, with the script \verb!on_master.sh!
\footnote{\url{https://raw.githubusercontent.com/Augugrumi/init-script/centos/kubernetes/on_master.sh}}.
The following command initiates the Kubernetes cluster:
\begin{lstlisting}
sudo kubeadm init --apiserver-advertise-address=<ip-of-the-master> 
--Pod-network-cidr=10.244.0.0/16 --ignore-preflight-errors cri | grep "kubeadm 
join" > /home/centos/joincommand
\end{lstlisting}
Saving in the \texttt{joincommand} file the command that it will be run on the
slave node to join the cluster. There are some considerations to keep in mind
though:
\begin{enumerate}
  \item \verb!--pod-network-cidr! value depends on which agent you want to use
    to configure Kubernetes network. We choose Flannel so we put
    \texttt{10.244.0.0/16};
  \item at the moment we do not know why some preflight errors came out by
    running the aforementioned command and in order to ignore them we added the
    \verb!--ignore-preflight-errors cri! flag to in order to complete the
    set-up.
\end{enumerate}
Next it is necessary to run the following commands to create a folder
needed by Kubernetes to properly work:
\begin{lstlisting}
mkdir -p $HOME/.kube
sudo cp -i /etc/kubernetes/admin.conf $HOME/.kube/config
sudo chown $(id -u):$(id -g) $HOME/.kube/config
\end{lstlisting}
It is now necessary to wait until all the Pods get initialized, and this
operation is performed using the \verb!kubectl get pods --all-namespaces! in an
automated fashion. The next step is to install
Flannel\footnote{\url{https://github.com/coreos/flannel}} in order to make
networking between pods working:
\begin{lstlisting}
kubectl apply -f https://raw.githubusercontent.com/coreos/flannel/master/Documentation/kube-flannel.yml
kubectl apply -f https://raw.githubusercontent.com/coreos/flannel/master/Documentation/k8s-manifests/kube-flannel-rbac.yml
\end{lstlisting}
Later it is possible to setup the dashboard to access Kubernetes resources
running:
\begin{lstlisting}
kubectl create serviceaccount dashboard -n default
  kubectl create clusterrolebinding dashboard-admin -n default \
   --clusterrole=cluster-admin \
   --serviceaccount=default:dashboard
  kubectl get secret $(kubectl get serviceaccount dashboard -o jsonpath="{.secrets[0].name}") -o jsonpath="{.data.token}" | base64 --decode >  secret.txt
  screen -dmS kubectl_proxy_screen bash
  screen -S kubectl_proxy_screen -X stuff "kubectl proxy
  "
\end{lstlisting}
The dashboard will be accessible on the port 8001 only on the master node and it
will typically be reachable at this URL:
\url{https://<master-ip>:<apiserver-port>/api/v1/namespaces/kube-system/services/https:kubernetes-dashboard:/proxy/}
To access the information available on the dashboard the token stored into the
file \texttt{secret.txt} present on the home of the master node is needed.
Finally, in order to make a node join the cluster you have to run the command
saved into the \texttt{joincommand} file that we created before.

\section{Persistent Storage}
\label{persistent-storage}

\subsection{Prerequisites}
\label{prerequisites}

\begin{itemize}
\item A working installation of Kubernetes
\item Python (2.7) installed and available on all Kubernetes nodes
\end{itemize}

\subsection{\gluster{}}
\label{glusterfs}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.3]{glusterkube}
\caption{Installation of \gluster{} on Kubernetes cluster hosts}
\end{figure}

To enable persistent storage in Kubernetes with \gluster{} we need at least
three nodes. As described above, we deployed our Kubernetes cluster using CentOS
7\footnote{This procedure was tested on \texttt{Kubernetes\ v1.11.1.}}. Since we
are running our cluster on an Openstack environment, we attached a volume to
each VMs. First of all, you need to install \texttt{gluster-server} on all the
nodes that will join the storage and \texttt{gluster-fuse} to provide
\texttt{mount.glusterfs} as an option to mount \gluster{} file system. On a
command line interface we typed:
\begin{lstlisting}
sudo yum install -y centos-release-gluster && \
sudo yum install -y glusterfs-server glusterfs-fuse
\end{lstlisting}
The next step is to load different kernel modules, using \texttt{modprobe}:
\begin{lstlisting}
sudo sh -c 'modprobe dm_snapshot && modprobe dm_mirror && modprobe dm_thin_pool'
\end{lstlisting}
To make this change permanent, a new file
\texttt{/etc/modules-load.d/gluster.conf} needs to be created, with the
following content:
\begin{lstlisting}
dm_snapshot
dm_mirror
dm_thin_pool
\end{lstlisting}
At this point, you can make \texttt{glusterd} daemon run at system start up by
the command:
\begin{lstlisting}
sudo systemctl start glusterd && sudo systemctl enable glusterd
\end{lstlisting}
All the nodes on which you want to install \gluster{} must be labelled by
Kubernetes running the command:
\begin{lstlisting}
kubectl label node <node name> storagenode=glusterfs
\end{lstlisting}
According to the \emph{gluster-kubernetes setup guide}
\footnote{\url{https://github.com/gluster/gluster-kubernetes/blob/master/docs/setup-guide.md}},
the ports $2222$, $24007$, $24008$ and ports between $49152$ and $49251$ have to
be reachable. On top of that, you have to to grant access to your root account
during the installation of all the \gluster{} software. In CentOS you need to
access every machine and type:
\begin{lstlisting}
sudo su
cd /root/
nano .ssh/authorized_keys
# edit the first line in the follwing way: from the sub-string starting
# as "ssh-rsa" until the end put the content in a new line, while comment
# the whole line
nano /etc/ssh/sshd_config # un-comment "PermitRootLogin: yes"
systemctl restart sshd
\end{lstlisting}
Later you need to clone the \texttt{gluster-kubernetes} repository that contains
the script \texttt{gk-deploy}. It figures out all the necessary stuff (e.g.
heketi):
\begin{lstlisting}
git clone https://github.com/gluster/gluster-kubernetes.git
cd gluster-kubernetes/deploy
\end{lstlisting}
Then, we need to define the topology. To do so, a \texttt{topology.json}
file must be created. It can be done modifying the sample file
\texttt{topology.json.sample} with your topology, like in the following example:
\begin{lstlisting}
{
 "clusters": [
   {
     "nodes": [
       {
         "node": {
           "hostnames": {
             "manage": [
               "192.168.29.26"
             ],
             "storage": [
               "192.168.29.26"
             ]
           },
           "zone": 1
         },
         "devices": [
           "/dev/vdb"
         ]
       },
       {
         "node": {
           "hostnames": {
             "manage": [
               "192.168.29.19"
             ],
             "storage": [
               "192.168.29.19"
             ]
           },
           "zone": 1
         },
         "devices": [
           "/dev/vdb"
         ]
       },
       {
         "node": {
           "hostnames": {
             "manage": [
               "192.168.29.22"
             ],
             "storage": [
               "192.168.29.22"
             ]
           },
           "zone": 1
         },
         "devices": [
           "/dev/vdb"
         ]
       }
     ]
   }
 ]
}
\end{lstlisting}
After it is possible to deploy a \gluster{} cluster. On master the following
command must be run:
\begin{lstlisting}[language=bash]
# assuming you still are in ~/blabla/gluster-kubernetes/deploy/
./gk-deploy -s <path/to/your/private/key> --ssh-user root --ssh-port 22 topology.json -y
\end{lstlisting}
At the end of the script, if no error occurs, you can find a 
\texttt{StorageClass} definition file called \texttt{storageclass.yaml} and to
add the \texttt{StorageClass} to Kubernetes you have to type:
\begin{lstlisting}[language=bash]
kubectl create -f storageclass.yaml
\end{lstlisting}
Optionally, it can be set as default \texttt{StorageClass} with the following
command:
\begin{lstlisting}[language=bash]
kubectl patch -f storageclass.yaml \
-p'{"metadata": {"annotations":{"storageclass.kubernetes.io/is-default-class":"true"}}}'
\end{lstlisting}

\section{Ingress}
\label{ingress}

Ingresses allow services to be available from outside the cluster. There are 
different ways with which Kubernetes can handle service reachability, and these
are:
\begin{description}
  \item [ClusterIP:] Exposes the service on a cluster-internal IP. This value
    makes the service only reachable from within the cluster and is the default
    ServiceType value.
  \item [NodePort:]
  \footnote{\url{https://kubernetes.io/docs/concepts/services-networking/service/\#nodeport}}
  Exposes the service on each Node IP at a static port (the NodePort). A
  ClusterIP service, to which the NodePort service will route, is automatically
  created. It will be possible to contact the NodePort service, from outside the
  cluster, by requesting
  \texttt{\textless{}NodeIP\textgreater{}:\textless{}NodePort\textgreater{}}.
  \item [LoadBalancer:]
  \footnote{\url{https://kubernetes.io/docs/concepts/services-networking/service/\#loadbalancer}}
  Exposes the service externally using a load balancer of a cloud provider.
  NodePort and ClusterIP services, to which the external load balancer will
  route, are automatically created.
  \item [ExternalName:]
  \footnote{\url{https://kubernetes.io/docs/concepts/services-networking/service/\#externalname}}
  Maps the service to the contents of the ExternalName field
  (e.g.~foo.bar.example.com), by returning a CNAME record with its value. No
  proxying of any kind is set up. This requires version 1.7 or higher of
  \texttt{kube-dns}.
\end{description}
Since for the \texttt{LoadBalancer} typology we would need a Cloud service like 
AWS or GCE, we opted for the \texttt{NodePort} one. This is not a suitable
solution for production environments since it opens a couple of ports for every
Pods that runs in a node, but since we are building a test environment this
was considered to be fine.

Plenty of \texttt{ingresses-controllers} are available in the market and we
decided to analyze some of them.

\subsection{Nginx}
\label{nginx}
Kubernetes supports its own version of Nginx\footnote{
\url{https://github.com/kubernetes/ingress-nginx}}, but there is also an
ingress-controller supported by the official Nginx team\footnote{
\url{https://github.com/nginxinc/kubernetes-ingress}}.
We have chosen the second one, since it seems more stable.

\subsubsection{Installation}
\label{installation}

The installation is pretty simple, and it can be done with Helm:
\begin{lstlisting}[language=bash]
git clone https://github.com/nginxinc/kubernetes-ingress.git --depth 1 # Clone the project repository
kubernetes-ingress/helm-chart
helm install --name nginxcontroller . --set controller.replicaCount=2,controller.service.type=NodePort,controller.service.externalTrafficPolicy=Cluster # Install the helm package
\end{lstlisting}
At this point, if you want to expose a service you need to create a file
similar to this one:
\begin{lstlisting}[language=yaml]
apiVersion: extensions/v1beta1
kind: Ingress
metadata:
 name: wordpress-ingress
spec:
 rules:
 - host: wordpress.example.com
   http:
     paths:
     - path: /<your-url-path>
       backend:
         serviceName: <your-service-name>
         servicePort: 80
\end{lstlisting}
And apply your forwarding to Kubernetes running the command:
\begin{lstlisting}[language=bash]
kubectl create -f ingress-forwarding.yaml
\end{lstlisting}
In the dashboard, you should see now something similar to this:
\begin{figure}[htbp]
\centering
\includegraphics[scale=0.3]{ingress}
\caption{Example of ingress in Kubernetes dashboard}
\end{figure}

\subsection{Træfik}
\label{truxe6fik}

\subsubsection{Installation}
\label{installation-1}
Role based access control was introduced in Kubernetes 1.6 and to make Træfik
working properly it must be configured. For simplicity, ClusterRoleBinding can
be used as in the Træfik installation guide\footnote{
\url{https://docs.traefik.io/user-guide/kubernetes/}}:
\begin{lstlisting}[language=yaml]
---
kind: ClusterRole
apiVersion: rbac.authorization.k8s.io/v1beta1
metadata:
  name: traefik-ingress-controller
rules:
  - apiGroups:
      - ""
    resources:
      - services
      - endpoints
      - secrets
    verbs:
      - get
      - list
      - watch
  - apiGroups:
      - extensions
    resources:
      - ingresses
    verbs:
      - get
      - list
      - watch
---
kind: ClusterRoleBinding
apiVersion: rbac.authorization.k8s.io/v1beta1
metadata:
  name: traefik-ingress-controller
roleRef:
  apiGroup: rbac.authorization.k8s.io
  kind: ClusterRole
  name: traefik-ingress-controller
subjects:
- kind: ServiceAccount
  name: traefik-ingress-controller
  namespace: kube-system
\end{lstlisting}
Save that in a file \texttt{traefik-rbac.yaml} and run:
\begin{lstlisting}[language=bash]
kubectl apply -f traefik-rbac.yaml
\end{lstlisting}
After the ingress controller can be set up in to two different way: using a
Deployment or a DaemonSet. We chose the first option for easier management. We
deployed it with this YAML:
\begin{lstlisting}[language=yaml]
---
apiVersion: v1
kind: ServiceAccount
metadata:
  name: traefik-ingress-controller
  namespace: kube-system
---
kind: Deployment
apiVersion: extensions/v1beta1
metadata:
  name: traefik-ingress-controller
  namespace: kube-system
  labels:
    k8s-app: traefik-ingress-lb
spec:
  replicas: 1
  selector:
    matchLabels:
      k8s-app: traefik-ingress-lb
  template:
    metadata:
      labels:
        k8s-app: traefik-ingress-lb
        name: traefik-ingress-lb
    spec:
      serviceAccountName: traefik-ingress-controller
      terminationGracePeriodSeconds: 60
      containers:
      - image: traefik
        name: traefik-ingress-lb
        ports:
        - name: http
          containerPort: 80
        - name: admin
          containerPort: 8080
        args:
        - --api
        - --kubernetes
        - --logLevel=INFO
---
kind: Service
apiVersion: v1
metadata:
  name: traefik-ingress-service
  namespace: kube-system
spec:
  selector:
    k8s-app: traefik-ingress-lb
  ports:
    - protocol: TCP
      port: 80
      name: web
    - protocol: TCP
      port: 8080
      name: admin
  type: NodePort
\end{lstlisting}
As before, save it in a file, say \texttt{traefik-deployment.yaml} and run:
\begin{lstlisting}[language=bash]
kubectl apply -f traefik-deployment.yaml
\end{lstlisting}
This deployment will expose two \texttt{NodePorts} services that allow to access
to the ingress and to the web management interface.

\section{Additional Tools}
\label{additional-tools}

\subsection{Helm}
\label{helm}

To install Helm you need to be in the master node\footnote{
\url{https://docs.helm.sh/using_helm/\#installing-helm}} and run the
command:
\begin{lstlisting}[language=bash]
curl https://raw.githubusercontent.com/kubernetes/helm/master/scripts/get | bash
\end{lstlisting}
To make it working correctly you should give it some additional permission
creating a file like this:
\begin{lstlisting}[language=yaml]
apiVersion: v1
kind: ServiceAccount
metadata:
  name: helm
  namespace: kube-system
---
apiVersion: rbac.authorization.k8s.io/v1beta1
kind: ClusterRoleBinding
metadata:
  name: helm
roleRef:
  apiGroup: rbac.authorization.k8s.io
  kind: ClusterRole
  name: cluster-admin
subjects:
  - kind: ServiceAccount
    name: helm
    namespace: kube-system
\end{lstlisting}
And after deploying the service running:
\begin{lstlisting}[language=bash]
kubectl create -f helm.yaml
\end{lstlisting}
At this point, we have defined a new ClusterRole, \emph{helm}, that has the
possibility to interact with \texttt{kube-system} namespace. Now we were able to
deploy Tiller in our cluster (Helm backend), by running:
\begin{lstlisting}[language=bash]
 helm init --service-account helm --wait
 \end{lstlisting}

### 2018-12-05 thesis/res/sections/11-ProjectScheduling.tex
Presentation integration (#18)
Integrate presentation in thesis repository

--- 2018-11-22 res/sections/11-ProjectScheduling.tex
@@
 
-Before starting this project a time requirements has been carried out, to see
-its feasibility. We calculated at least six month necessary to be able to create
+Before starting this project a time scheduling has been carried out, to see
+its feasibility. We calculated at least six months necessary to be able to create
 a working proof of concept (PoC) that allowed us to perform tests and, in
@@
 Since our knowledge about the technologies was very poor at the beginning, we
-preferred to invest a good chunk of time on studying them, to understand the already
-present products available in the market and to choose what we really needed.
-Additionally, at the end of this phase we though it was necessary to try these
-products to see if they really fitted our necessities. After this phase, we
-planned to actually starting implementing the missing components to create and
-to fulfill the requirements defined in Section~\ref{chap:prjan:sec:req}. Since
-the required technologies, Java and C/C++ are known to the authors, we estimated
-an eight-week period related to coding and software implementation. After that,
-a two weeks were planned in order to perform tests and gather metrics about the
-system built, and too see how much the job done satisfied our expectations.
-Finally, we estimated four weeks to group and write down all the information and
-considerations we got during the thesis.
+preferred to invest a good chunk of time on studying them, to understand the
+already present products available in the market and to choose what we really
+needed. Additionally, at the end of this phase we though it was necessary to try
+these products to see if they really fitted our necessities. After this phase,
+we planned to actually starting implementing the missing components to create
+and to fulfill the requirements defined in Section~\ref{chap:prjan:sec:req}.
+Since the required technologies, Java and C/C++ are known to the authors, we
+estimated an eight-week period related to coding and software implementation.
+After that, two weeks were planned in order to perform tests and gather metrics
+about the system built, and too see how much the job done satisfied our
+expectations. Finally, we estimated four weeks to group and write down all the
+information and considerations we got during the thesis.
 
@@
 Unfortunately, many difficulties slowed our project development, starting from
-the beginning. The total time required to finish all the requirements decided
+the beginning. The total time required to finish all the planned requirements
 was of 37 weeks, namely nine months circa. Many problems emerged during the
 study of the technologies, that revealed to be more complicated than expected,
-especially since some frameworks were more complicated to deploy and try them
+especially since some frameworks were more intricate to deploy and to try them
 out. Time was lost to obtain resources from the University that was not
@@
 required was of eight weeks, for a total of 16 weeks. Finally, the time to write
-down the following thesis was limited due, and we had to write it down in a
+down the following thesis was limited too, and we had to write it down in a
 total of two weeks, instead of the four initially planned.

### 2018-12-05 thesis/res/sections/20-VNFImplementaton.tex
Presentation integration (#18)
Integrate presentation in thesis repository

--- 2018-11-30 res/sections/20-VNFImplementaton.tex
@@
 The complexity behind the creation of a working Service Function Chain system
-was so high to require particular care and a separated study. Indeed, the SFCs
-are the at the base of the whole architecture: packet transmission and
+was high demanding for particular care and a separated study. Indeed, the SFCs
+are the at the core of the whole architecture: packet transmission and
 elaboration is made here. Challenges regarding this topic are building a system
@@
 provide methods to recovery from failures. In this chapter, a brief explanation
-about the challenges and the issues encountered building this system will be
-given.
+about the challenges and the issues encountered building this system are given.
 
@@
 \end{figure}
-
-The approved proposal for the creation of a proper chain has been based on four components:
+In Figure~\ref{chap:impl:img:finalreview} is shown a high level schema
+evidencing the components and their relationships. The approved proposal for the
+creation of a proper chain has been based on four components:
 \begin{itemize}
-\item Astaire;
-\item Harbor;
-\item Ironhide;
-\item Roulette.
+\item Astaire, which aim is to performing packet encapsulation and decapsulation
+  between VNFs;
+\item Harbor, which acts as our MANO component;
+\item Ironhide, that provides ingress and egress point from our architecture;
+\item Roulette, that stores endpoint information and holds the SFC topology
+  view.
 \end{itemize}
@@
 preserved, so that VNFs can perform packet elaboration. In order to obtain this,
-packet encapsulation has been used. IP in IP encapsulation allows to preserve
+packet encapsulation has been used. An IP/IP encapsulation allows to preserve
 the original packet, making the incoming packet the payload of a new one.

### 2018-12-05 thesis/res/sections/40-FutureWork.tex
Presentation integration (#18)
Integrate presentation in thesis repository

--- 2018-11-28 res/sections/40-FutureWork.tex
@@
 the one regarding security. In this context, the security we talk is the one
-about the whole system, and not the single VNFs implementations. Fortunately,
+about the whole system, and not the single VNFs implementation. Fortunately,
 Kubernetes is equipped with out-of-the-box access management options. A good way
@@
 come with security features, and the exposed APIs are usable by everyone. This
-could lead to possible Denial of Service (DoS) attack of the disruption of NS
-services by a malicious attacker.
+could lead to possible Denial of Service (DoS) attacks leading to the disruption
+of NS services by a malicious attacker.
 
